### 2023-10-07 paper/main.tex
first commit

\documentclass[a4paper, man, floatsintext]{apa7}
%dvipsnames

% Language setting
\usepackage[english]{babel}

% Useful packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[natbib,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
\usepackage{csquotes}
\DeclareLanguageMapping{american}{american-apa}
\bibliography{interferencebib}

\usepackage{gb4e}
\usepackage{pifont}
\noautomath
\usepackage{authblk}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{subcaption}

\title{\hspace{1cm}Semantic interference during encoding and retrieval \newline
%Semantic (but no syntactic) similarity-based interference
Evidence from self-paced reading and event-related potentials}
\authorsnames{Pia Schoknecht, Shravan Vasishth}

\shorttitle{Semantic interference}
\affiliation{
Department of Linguistics, University of Potsdam}

\authornote{This research was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), Project ID 317633480, SFB 1287.
We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition.
Correspondence concerning this article should be addressed to Pia Schoknecht, e-mail: pia.schoknecht@uni-potsdam.de
}

\abstract{Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed to interpret the sentence. If these retrieval cues match with not only the features on the noun that is the retrieval target but also with those on other nouns in the sentence, then processing difficulty arises at the verb. This difficulty in identifying the correct dependent is called similarity-based interference. 
% We are explaining what we did, not summarizing every detail in the paper:
%Previously, behavioral studies have consistently found syntactic interference at the verb, but semantic interference was only found in later regions or not at all. 
% What we did:
We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07}, investigating interference due to syntactic and semantic cues in German. 
% Not needed for an abstract, plus sounds boring:
%The aim of our experiments was to gain precise estimates of the effects of syntactic and semantic interference to inform theory building.
% What we found:
A Bayes factor analysis showed that 
both the self-paced reading times as well as event-related potentials provided clear evidence for a semantic interference effect. In the self-paced reading experiment, interference due to the semantic cues caused a reading time slow-down starting at the distractor. 
% The reader does not even know what that is:
%The best explanation for this effect is encoding interference. 
The event-related potentials showed a more negative N400 amplitude at the critical verb, i.e., the retrieval site, for high vs.\ low semantic interference.
% Irrelevant in an abstract:
 %and no such effects at preceding words. 
 % No new information in this sentence, a waste:
 %Therefore, the event-related potentials were modulated by retrieval interference.
% Said this more compactly already:
%Bayes factors for both reading times and event-related potentials provided evidence for the semantic interference effect. 
Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to syntactic cues. This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play a role in retrieval. We show through computational modeling that cue-based retrieval also shows no syntactic interference if the parser is assumed to keep track of which clause a noun occurs in. Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model are consistent with the observed patterns in our data.

In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that although both syntactic and semantic cues drive retrieval at the verb, syntactic cues include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.

% This cannot be the main point of this paper! 
%The present findings, which are in line with various sentence processing accounts, highlight the importance of semantic cues during sentence processing.

% Highlights must be in a separate file (not in the paper itself)
\clearpage

\section{Highlights} 
% 3-5 highlights
% only 85 characters including spaces per highlight
\begin{itemize}
\item Self-paced reading (SPR) and event-related potentials (ERP) experiments on interference
\item Bayes factors show evidence for semantic, but against syntactic interference
\item Modeling shows that the parser tracks hierarchical information, attenuating syntactic interference.
% Wasting highlights on low information content
%\item SPR: Encoding interference lead to longer reading times starting at a distractor
%\item ERP: Time course of effect clearly indicated an effect of retrieval interference
\end{itemize}

}

\keywords{Cue-based Retrieval, Interference, Syntax, Semantics, Replication, Self-paced Reading, Event-related Potentials, Computational Modeling}

\begin{document}

\maketitle

\section{Introduction}

It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval. As an example, consider 
\ref{ex:1_vandyke07}.

\begin{exe}[ht]
\ex \label{ex:1_vandyke07} 
The worker was surprised that the resident who was living near the dangerous warehouse was complaining about the investigation. \parencite{vandyke07}
\end{exe}

There is clear evidence in the literature  \parencite[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom. A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues.

Here, we consider two kinds of cues used in retrieval: semantic and  syntactic cues.
In a sentence like \ref{ex:1_vandyke07}, the retrieval cues generated at the verb are standardly assumed to include at least  the semantic cue \{+animate\} and the syntactic cue \{+grammatical subject\}. These cues pick out the target noun \textit{the resident} as the subject. The distractor \textit{warehouse}, intervening between subject and verb, does not match either of these retrieval cues. However, if the distractor were \textit{neighbor} instead of \textit{warehouse}, then it would match the \{+animate\} cue at the verb. This overlap in cues between the target noun and the distractor is called cue-overload in cue-based retrieval accounts, and is predicted to trigger semantic interference, which expresses itself as greater processing time at the verb compared to the baseline condition \ref{ex:1_vandyke07}.

By contrast, syntactic interference is expected to be triggered if the distractor matches the \{+grammatical subject\} cue. This situation can occur when the relative clause  in \ref{ex:1_vandyke07} that modifies \textit{the resident} were to be changed to \textit{who said that the warehouse was dangerous}. Now, the distractor matches the \{+grammatical subject\}.

Thus, a distractor which matches a semantic cue, here, \{+animate\}, is predicted to induce semantic interference and a distractor which matches a syntactic cue, here, \{+grammatical subject\}, is predicted to induce syntactic interference. More generally, interference that is induced by such an overlap between retrieval cues and features of items in memory is called retrieval interference. 
%In grammatical sentences, where the retrieval cues match with the features of the target of the memory retrieval (e.g., the subject \textit{the resident} in Example \ref{ex:1_vandyke07}), interference from distractors is assumed to cause processing difficulty.

Previous work has shown clear evidence for this kind of retrieval interference \parencite{vandyke07, vandyke_mcelree06, vandyke_mcelree2011, nicenboim,  vandyke_lewis03,jaeger_etal_2017}.
However, two important questions remain unanswered. First, as discussed in \cite{jaeger_etal_2017}, almost all the published studies on interference effects are severely underpowered. Low power has the consequence that published estimates will tend to be overestimates or misestimates of the effect size \parencite{vasishth2018_signficancefilter}. Over-/misestimates have important implications for theory evaluation. One example is discussed in \cite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \parencite{levy&keller_2013} could not be validated when a larger-sample study was conducted; by contrast, the original small-sample study  \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory. Another example is the recent heated debate on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
Second, although it is clear and uncontroversial  \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event. For example, \cite{dillon2013} argued, following \cite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built; instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.  \cite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference \parencite[cf. ][]{jaeger_etal_2020,yadav2021individual}. The broader implication here is that the cues used during retrieval can modulate processing difficulty in possibly surprising ways.

\subsection{The present study}

We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with a series of experiments using self-paced reading and EEG. All experiments used the 2 $\times$ 2 design by \citet{vandyke07} crossing syntactic and semantic interference. This is -- to our knowledge -- the first ERP study on this classic design investigating interference during subject-verb dependency resolution. The SPR experiments were conducted web-based via Prolific and the EEG experiments required participants to come into the lab. All experiments were run with native speakers of German. Because the high signal-to-noise ratio of EEG data requires a large number of trials per condition and participant, we constructed 120 items which were used in the EEG and SPR experiments. It would have been very taxing for the participants to read all 120 items in one experimental session, therefore we decided to split the experiment into two sessions. The ERP data that we present here, comes only from participants who completed both sessions. In contrast, most of the SPR participants (570 out of 774) finished only one session. This was intentional because we found that participants' reading speed in the second session was increased, probably due to increased familiarity with the task. Figure \ref{fig:project_str} shows the structure of the present study, i.e., how many participants saw which subset of the items and whether these participants completed one or two experimental sessions.

\begin{figure}[H]
    \caption{Structure of the present study. In total, 120 items were used for the SPR (red branches) and the ERP experiments (blue branches). The color brightness reflects how many sessions were completed by each participant (lighter color: two sessions, darker color: one session).}
    \label{fig:project_str}
    \centering
    \includegraphics[width=0.85\textwidth]{images/pandora_project_structure_figure.pdf}
\end{figure}

\subsection{Materials}
As the same materials were used for all our experiments, we describe them here once before we describe the specifics of the experiments, separately. The materials for the study consisted of 120 items with 4 conditions in a Latin square design shown in \ref{ex:materials}. The items were constructed following \citeauthor{vandyke07}'s (\citeyear{vandyke07}) items and partially re-used  \citeauthor{mertzen}'s (\citeyear{mertzen}) items.

\begin{exe}  
\ex \label{ex:materials} Example item (critical word in bold, distractor in italics) of the present study:
    \begin{xlist}   
    \ex {High syntactic interference with and high / low semantic interference:}\label{ex:hisyn} 
    \gll Die  Nachbarin glaubte,	dass	der Witwer,  der  erzählt hatte, dass der \textit{Einbrecher} / \textit{Verlust}  schrecklich war, abends regelmäßig \textbf{trank}, um zu vergessen.\\ 
    The\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} believed that the widower who told had that the burglar / loss awful was in.the.evening regularly drank in.order to forget \\
    \trans `The neighbor believed that the widower, who had told her that the loss was awful, drank regularly in the evenings to forget.' \\

    \ex {Low syntactic interference with high / low semantic interference:}\label{ex:losyn} 
    \gll Die  Nachbarin glaubte, dass der Witwer,  der  von    dem schrecklichen \textit{Einbrecher} / \textit{Verlust} erzählt hatte, abends regelmäßig \textbf{trank}, um zu vergessen. \\ 
    The\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} believed that the widower who about the awful burglar / loss told had in.the.evening regularly drank in.order to forget\\
    \trans `The neighbor believed that the widower, who had told her about the awful loss, drank regularly in the evenings to forget.'  \\
    \end{xlist}
\end{exe}

Comprehension of the sentences in \ref{ex:materials} requires a memory retrieval operation at the verb \textit{trank} (`drank') to retrieve its non-adjacent subject \textit{Witwer} (`widower'). During this retrieval, the distractor \textit{Einbrecher} (`burglar') / \textit{Verlust} (`loss') might cause interference and impede comprehension. Syntactic interference is manipulated via the syntactic status of the distractor, which is either subject (\ref{ex:hisyn}) or part of a prepositional phrase (\ref{ex:losyn}). During the retrieval of the subject, a distractor in subject position is assumed to cause more syntactic interference than a distractor inside a prepositional phrase. Semantic interference is manipulated via the animacy of the distractor. During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference. In contrast to the previous studies (see Introduction), we inserted a second pre-critical adverb, i.e., a pre-pre-critical region, to absorb potential spill-over effects from the embedded clause(s). So, potential wrap-up effects from the embedded clause(s) could spill-over into the pre-pre-critical region and the pre-critical region would be further away from clause boundaries. This way potential pre-critical effects are less likely to be caused by structural confounds.

\subsubsection{Plausibility norming}
\citeauthor{vandyke07} conducted a pre-test to ensure that the distractor in high semantic interference conditions was a sufficiently probable subject of the critical verb, i.e., that it could potentially induce semantic interference. For the same purpose, we conducted a plausibility norming experiment.

\paragraph{Participants}
Forty participants (18 female, 22 male, mean age: 27.3 years old, age range: 21 - 39 years old) from the SPR experiments were re-invited for the plausibility norming one month after they completed the SPR experiment.

\paragraph{Materials}
For each item quadruplet, four simple SVO sentences were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb. The norming materials corresponding to the item despicted in \ref{ex:materials} are shown in \ref{ex:materials_norming}.

\begin{exe}  
\ex \label{ex:materials_norming}
    \begin{xlist}   
    \ex {Subject}:\label{ex:subject} 
    \gll Der Witwer trank. \\ 
    the widower drank\\
    \trans  `The widower drank.'\\  
    \ex {Animate distractor:}\label{ex:animate_dist} 
    \gll Der Einbrecher trank. \\ 
    the burglar drank \\
    \trans `The burglar drank.' \\
   \ex {Inanimate distractor:}\label{ex:inanimate_dist} 
    \gll Der Verlust trank.  \\ 
    the loss drank\\
    \trans  `The loss drank.'\\  
    \ex {Introduction noun:}\label{ex:intro_noun} 
    \gll Die Nachbarin trank.\\ 
    the\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} drank \\
    \trans `The neighbor drank.' \\
    \end{xlist}
\end{exe}

\paragraph{Procedure}
The plausibility norming experiment was implemented in PCIbex \citep{pcibex} and was conducted online. Participants were asked to judge the plausibility of the sentences on a 7-point scale ranging from absolute implausible (1) to absolute plausible (7). All sentences belonging to one item were shown on one screen, so that they could be judged relative to each other. The plausibility norming experiment included 120 sentences from a different experiment which functioned as fillers. The participants received \pounds 3.40 as compensation for the norming experiment. A demonstration of the plausibility norming experiment can be accessed via \hyperlink{https://farm.pcibex.net/r/WIeKYM/}{https://farm.pcibex.net/r/WIeKYM/}.

\paragraph{Results}
The plausibility ratings showed that the semantic manipulation worked as intended. The subject received the highest rating (mean: 6.5, sd: 0.9) and the animate distractor (high semantic interference conditions) was almost as plausible (mean: 6.1, sd: 1.3). In contrast, the inanimate distractor (low semantic interference conditions) was very implausible (mean: 1.7, sd: 1.4). The introduction noun which was not manipulated was tested as well and it received a high plausibility rating (mean: 6.0, sd: 1.4). These results are illustrated in Figure \ref{fig:plausibility_rating}. The plausibility ratings are similar to those reported for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3.

\begin{figure}[H]
\caption{Plausibility ratings for the animate (high semantic interference conditions) and inanimate distractor (low semantic interference conditions), the introduction noun (\textit{X believed that \dots}) and the subject. The diamonds represent the mean rating.}\label{fig:plausibility_rating}
\centering
\includegraphics[width=0.6\textwidth]{images/pandora_plausibility_ratings.jpg}
\end{figure}

\section{Experiment 1: SPR}
\subsection{Methods}
Initially, we planned to conduct SPR experiment in two sessions to gain data from all participants for all 120 items. Collecting data for 120 items in one session was not considered feasible in a web-based SPR experiment, as the experiment duration would have been close to one hour. So, the first batch of participants (160 out of 774) completed two experimental sessions. This approach was dropped after we realized, that participants adapted to the task quickly which led to a considerable reading time speed-up between sessions (see Figure \ref{fig:whole_sentence} D). After this realization, all further participants took part in a single session during which they read half of the 120 items (for a visualization of the project structure, see Figure \ref{fig:project_str}).

\subsubsection{Participants}
%two sessions, session 1
% 224 participants completed
% 17 low acc
% 2 inconsistent demographic info
% 1 technical error
% final 204

% two sessions, session 2
% 178 participants completed
% 18 low acc
% final 160

% rep s1
% 387 participants completed
% 31 low acc
% 6 inconsistent demographic info
% final 350

% rep s2
% 279 participants completed
% 51 low acc
% 8 inconsistent demographic info
% final 220

% 224 + 18  + 387 + 279 = 908

Participants for the experiment were recruited via Prolific. In total, 908 participants took part in the self-paced reading experiment. 117 participants were excluded due to low comprehension question accuracy (< 70\,\%). Additionally, 16 participants were excluded because the demographic information in their Prolific profile and the information which they provided during the experiment did not match (regarding, e.g., their native language, language impairment, or age). One participant was excluded because of a technical error during their session. The data of 774 participants (mean age: 27 years old, age range: 18 - 40 years, 395 female, 369 male, 10 preferred to not provide sex information) were used for analyses. When asked about their highest level of education, 365 participants reported a bachelor's degree or higher university education. 334 participants reported to have a high-school diploma and 70 another secondary school certificate. Five participants replied with ``other.'' We conclude that the participants who were recruited via Prolific were slightly older and have a comparable or higher level of education than the typical undergraduate and graduate students who participate in lab-based studies.

A subset of the SPR experiment was carried out in two experimental sessions per participant. Drop-out between sessions led to a lower sample size in the second session (204 vs.\ 160). In this multiple-sessions part of the experiment, compensation for the first session was \pounds 3.5 and for the second session, it was \pounds 5. In the single-session part of the experiment, compensation was \pounds 4.

\subsubsection{Procedure}
Participants completed a moving window self-paced reading task \citep{just_etal_1982} on the PCIbex Farm \citep{pcibex}. They were instructed to read for comprehension at a comfortable pace. All sentences were displayed word by word. Masked words were presented as a line indicating its length. Unmasked words were presented in 18 pt Courier font. The space bar on the keyboard was used to unmask words. The length of the sentences required line breaks. These were hard-coded so that in each sentence a line break appeared i) between the subject and the relative clause and ii) two words after the critical verb (see \ref{ex:linebreak}, critical word in bold, | indicate line breaks in the moving window display). In a third of the trials, the sentence was followed by a yes/no comprehension question. The participants received no feedback on their performance.

In addition to the reading-for-comprehension task, the participants completed two types of attention and compliance checks. The attention checks required them to press one of two designated keyboard keys ten times during each session. The compliance checks required them to type in two types of fruit and two hobbies (session 1) and two German cities and two breakfast food items (session 2). No participant failed any of these checks. The reader can try the experiment via this link \hyperlink{https://farm.pcibex.net/r/CBkSKl/}{https://farm.pcibex.net/r/CBkSKl/}.

\begin{exe}  
\ex \label{ex:linebreak}
    \gll Die Nachbarin glaubte, dass der Witwer, | der erzählt hatte, dass der Verlust schrecklich war, regelmäßig abends \textbf{trank}, um zu | vergessen. \\ 
    The neighbor believed that the widower | who told had that the loss awful was regularly in.the.evening drank in.order to | forget\\
    \trans  `The neighbor believed that the widower, who had told her that the loss was awful, regularly drank in the evenings to forget.'\\  
\end{exe}

In the multiple-sessions part of the experiment, participants were re-invited for session 2 after they completed session 1. The experimental sessions were separated by 1 - 20 days. The procedure of the sessions was identical. Each session lasted approximately 25 minutes.

\subsubsection{Statistical analyses}
Comprehension question accuracy was analyzed with Bayesian generalized linear mixed models, i.e., logistic regression, in R \citep{r}, using the brms package \citep{brms}. The fixed effects were syntactic interference, semantic interference and their interaction. These were sum-contrast coded (high +0.5, low -0.5). Random intercepts for participants and items were included. For the intercept, we used a Normal(0, 1.5) prior and  for all slope parameters, a Normal(0, 0.1) prior was used. No other priors were specified. Models were run with 4 chains and 8,000 iterations in each chain. 2,000 iterations in each chain consisted of a warm-up phase.

Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in \ref{ex:materials}, constituting the retrieval site. Additionally, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. Factors were sum-contrast coded \citep{schad_et_al_contrasts}. Random intercepts for participants and items were included. Inclusion of any random slopes led to unstable Bayes factors due to increased model complexity. We used informative priors for all parameters of the models (see Table \ref{tab:spr_priors}). We used a wide range of priors on the parameters of interest, assuming effect sizes ranging from 8 ms up to 81 ms, to carry out a Bayes factors sensitivity analysis. If BF$_{10}$ > 1, it provides evidence in favor of the effect. If BF$_{10}$ < 1, it provides evidence in favor of the null model. The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence for the null. Models were run with 4 chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being the warm-up phase \citep{schad_etal_2022_BF}.

\begin{table}[h]
    \caption{Informative priors for the analysis of the self-paced reading time data. Priors on the slope parameters were truncated to a lower bound of zero, i.e., only positive values. This corresponds to the assumption that high interference conditions lead to longer reading times than low interference conditions. The standard deviation of the slope priors was varied in order to conduct a Bayes factor sensitivity analysis following \citet{schad_etal_2022_BF}. See the corresponding assumed a priori range of the difference between reading times under high vs.\ low interference on the millisecond scale.}
    \label{tab:spr_priors}
    \centering
     \begin{tabular}{lllr}
    \toprule
    Parameter&Prior & &Assumed Range\\
    \midrule
  Intercept & Normal(6, 0.6)& &\\
  \cmidrule{2-4}
  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \multirow{3}{1cm}{lb=0} & 0 - 8 ms\\
  &  Normal(0, 0.05)& &0 - 40 ms\\
  & Normal(0, 0.1) & &0 - 81 ms\\
  \cmidrule{2-4}
  sigma & Normal(0, 0.5)&&\\
  SD & Normal(0, 0.1)& &\\
  \bottomrule
  \end{tabular}
  \end{table}

%\bigskip

%\begin{table}[ht]
%    \centering
%        \caption{Contrast coding.}
%    \label{tab:contrasts}
%    \begin{tabular}{cccc}
%\toprule
%condition & syntactic interference & semantic interference & interaction\\
%\midrule
%a)  LoSyn LoSem &  $-1$& $-1$ & $+1$ \\
%b)  LoSyn HiSem & $-1$ & $+1$ & $-1$  \\
%c)  HiSyn LoSem &  $+1$ &$-1$ & $-1$    \\ 
%d)  HiSyn HiSem &  $+1$ &$+1$  & $+1$   \\
%\bottomrule
%    \end{tabular}
%\end{table}

\subsection{Results}
Data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.

\subsubsection{Comprehension question accuracy}
After exclusion of participants with accuracy below 70\,\%, overall accuracy (including fillers) was good; on average 85.9\,\% (range: 71.9 - 100\,\%). Accuracy in critical trials was 81.1\,\% (range: 20 - 100\,\%). Condition-wise accuracy is shown in Table \ref{tab:spr_acc}.

\begin{table}[]
    \caption{By-condition accuracy in critical trials in the SPR experiment.}
    \label{tab:spr_acc}
    \centering
    \begin{tabular}{llr}
    \toprule
    syntactic & semantic & accuracy \%\\
    \midrule
        low &  low & 86.0\\
        low &  high & 77.7\\
        high &  low & 85.8\\
        high &  high & 74.9\\
    \bottomrule
    \end{tabular}
\end{table}

Table \ref{tab:spr_acc_mod} shows the results of the generalized mixed model analyzing the comprehension accuracy in critical trials. The estimates and 95\,\% credible intervals in log-odds show primarily a reduction in comprehension accuracy for high compared to low semantic interference conditions.

\begin{table}[H]
    \caption{Results in log-odds from the Bayesian generalized model analyzing the comprehension accuracy in critical trials of the SPR experiment.}
    \label{tab:spr_acc_mod}
    \centering
    \begin{tabular}{lrrr}
    \toprule
    & Estimate & Est.Error & 95\% CrI  \\
    \midrule
Intercept& 1.66 & 0.14&  1.38 - 1.93\\
syntactic& -0.08 & 0.04& -0.16 - -0.01\\
semantic&  -0.61 &0.04& -0.69 - -0.54\\
interaction& -0.10& 0.06& -0.22 - 0.02\\
    \bottomrule
    \end{tabular}
\end{table}

\subsubsection{Self-paced reading times}
Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structure. 
%\clearpage

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. Panel C shows the pooled reading times of all conditions in the critical and surrounding regions. Panel D shows the reading times of all conditions in the critical and surrounding regions in the different parts of the self-paced reading experiment.}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=0.85\textwidth]{images/Pandora_all_wholesentence_pooled_zoom_exp.jpg}
\end{sidewaysfigure}
\clearpage

It is apparent from Figure \ref{fig:whole_sentence} A and B that, at the distractor, regardless of the syntactic manipulation, reading times between high and low semantic interference started to differ. Distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions. This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions. Nonetheless, we refrained from analyzing the distractor and the immediately following regions, because they differed between conditions. Instead we focused our analyses on regions which were identical across conditions: The pre-critical, critical and spill-over region. Figure \ref{fig:whole_sentence} C shows the pooled reading times of these regions (and additionally the pre-pre-critical region which was identical across conditions as well, but might be confounded by spill-over from the syntactic manipulation). The pre-pre-critical and pre-critical region show longer reading times for high vs.\ low semantic interference conditions. The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low. The other two conditions showed intermediate reading times. The spill-over region again primarily showed a difference between high and low semantic interference conditions.

Panel D shows the reading times of the critical and surrounding regions for each of the sub-parts of the SPR experiment. The sub-experiment with multiple sessions and items 1-60 (N = 204, left most sub-figure) showed results similar to the pooled results presented in Panel C. The sub-experiment with multiple sessions and items 61-120 (N = 160) showed a semantic interference effect at the critical verb and less clear results in the surrounding regions. Additionally, this sub-experiment showed shorter reading times compared to the other sub-experiments in general. This was probably caused by adaptation to the task, because the same participants had completed another session earlier. The sub-experiment with a single session and items 1-60 (N = 350) showed a semantic interference effect in all regions. The sub-experiment with a single session and items 61-120 (N = 220, right most sub-figure) showed a semantic interference effect in the regions surrounding the critical verb, but not at the verb itself. At the verb, reading times were longest when both syntactic and semantic interference were high. Reading times were a bit faster when both syntactic and semantic interference were low. The other two conditions (high syntactic/low semantic and low syntactic/high semantic) lead to the fastest verb reading times in this sub-experiment. These somewhat inconsistent results, especially in the critical region, across the sub-experiments and the random assignment of participants to the sub-experiments motivated us to gain estimates of the effects under investigation for the individual participants (see Figure \ref{fig:spr_individual}).

\begin{figure}[H]
    \caption{Individual estimates of syntactic and semantic interference and their interaction at the critical verb for all 774 participants in the self-paced reading experiment. Shown are shrunken estimates from a Bayesian hierarchical linear mixed model. Circles show individual’s mean effects, errors bars are 95\,\% credible intervals. The red dashed line shows an effect of zero (difference of zero milliseconds). Two solid red lines mark an effect size of 10 and 20 ms, respectively.}
    \label{fig:spr_individual}
    \centering
    \includegraphics[width=\textwidth]{images/individual_diffs.jpg}
\end{figure}

The individual-level effects were derived from a Bayesian linear mixed model with the priors shown in Table \ref{tab:spr_priors} with the intermediate slope prior of Normal(0, 0.05) and random slopes for syntactic and semantic interference and their interaction but without correlations of random effects.

Figure \ref{fig:spr_individual} shows that at the critical verb, the 95\,\% credible intervals for the syntactic interference effect of all participants were centered around zero. In contrast, participants consistently displayed a small effect of semantic interference with narrow 95\,\% credible intervals. The individual-level estimates of the interaction were a bit larger but they were associated with more uncertainty (wider 95\,\% credible intervals). The individual-level results help to understand the population-level posteriors presented in Figure \ref{fig:spr_posteriors} and the population-level Bayes factor analysis (see Figure \ref{fig:spr_bfs}).

The posteriors of the pre-critical, critical and spill-over regions were very similar (see Figure \ref{fig:spr_posteriors}). For the syntactic interference effect, they were close to zero in all regions. In contrast, for the semantic interference effect, the estimates ranged from 4 ms (lower bound in the critical region) to 14 ms (upper bound in the pre-critical region). The posteriors for the interaction were consistent with an interaction effect between 1 ms and 13 ms in the pre-critical and spill-over region. They were largest in the critical region (95\,\% CrI [7, 22 ms]).

\begin{figure}[H]
    \caption{Posteriors for the syntactic and semantic interference effects and their interaction at the critical verb and surrounding regions in the self-paced reading experiment. Numbers are the means and 95\,\% credible intervals. Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
    \label{fig:spr_posteriors}
    \centering
    \includegraphics[width=\textwidth]{images/posteriors_spr_pooled_774.png}
\end{figure}

Bayes factors, the Bayesian version of hypothesis testing, were used to evaluate whether there was evidence for the effects estimated above while taking into account the uncertainty \citep{schad_etal_2022_BF}. The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. Bayes factors provided extremely strong evidence for the semantic interference effect in the pre-critical, critical and spill-over region (BF$_{10}$ > 200). Bayes factors showed no effect of syntactic interference in any of the investigated regions. In fact, the Bayes factor analysis actually provided evidence against the syntactic interference effect in all regions and under all priors (BF$_{10}$ < 1).  Evidence for the interaction was only found in the critical region, where it ranged from strong (BF$_{10}$ = 40) to weak (BF$_{10}$ = 2) depending on the a priori assumed effect size.

\begin{figure}[H]
    \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times at the critical verb and surrounding regions.}
    \label{fig:spr_bfs}
    \centering
    \includegraphics[width=\textwidth]{images/BF_plot_spr_774_allregions.png}
\end{figure}

\section{Discussion}
We presented a large-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. Comprehension accuracy and reading times of the critical verb showed semantic interference, but no syntactic interference. Importantly, in the reading times, the effect did not emerge at the critical verb, i.e., the retrieval site during subject-verb dependency formation, but actually started when the distractor was read.

\subsection{Pre-critical effects}
Animate distractors (high semantic interference) were read slower than inanimate distractors (low semantic interference, see Figure \ref{fig:whole_sentence} A and B). One might argue that this could be due to differences in word length. The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9). It seems unlikely that this small difference in word length caused such a large and long-lasting effect.

We assume that the semantic interference effect starting at the distractor and persisting into later regions was caused by encoding interference \citep{Oberauer_Kliegl_2006}. The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) caused the comprehender to slow down \citep[for similar findings, see e.g., ][]{lago_etal_2021, kush_etal_2015, gordon02}.

The pre-critical effects in the previous studies \citep{vandyke07, mertzen} might have also started at the distractor. \citeauthor{mertzen}'s (\citeyear{mertzen}) Figures 5 and 6 indicated reading time differences in earlier regions, especially for their German data (see their Figure 6). \citet{vandyke07} did not report reading times for the whole sentences / distractors. While \citet{vandyke07} assumed that the pre-critical effects were due to plausibility differences between conditions, \citet{mertzen} also discussed encoding interference as the cause of pre-critical effects in their study. Our results strongly support the encoding interference explanation because the effects started at the distractor.

Alternative causes of the pre-critical effects discussed by \citet{mertzen} are parafoveal-on-foveal effects, sentence structure confounds and prediction. Our study rules out the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading. The observed difference between the high and low semantic interference conditions is independent of the syntactic manipulation. Therefore, the sentence structure confound is not a good explanation for the effect in our study. \citet{mertzen} argued that the pre-critical adverb is the first word of the verbal phrase and therefore, might trigger the retrieval of the subject. In our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence} A and B). Consequently, the prediction explanation is also ruled out by our findings, rendering the encoding interference explanation the most likely explanation of the effects found prior to the critical verb in our and previous studies. Crucially this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in the critical region as well. The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence. It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region or still due to encoding and maintaining the distractor in memory.

\subsection{Comparison with previous findings}
The results of the present study stand in contrast to previous findings \citep{vandyke07, mertzen}. The present study is the only one to find semantic and no syntactic interference effect. The difference in observed effects could be explained by differences in statistical power or by methodological differences. The present study has higher statistical power than the previous studies. \citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively. \cite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each. In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items). Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design. Consequently, the estimates from the present study are the most precise estimates for this design. The small effect size in the present study is comparable to that of other large-scale studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.

An obvious difference from previous studies is the experimental methodology. The majority of the previous experiments employed eye-tracking while reading methodology, while we used a self-paced reading task. The exact relation of reading times obtained by these two methods is unclear. Self-paced reading could facilitate interference effects compared to natural reading (as in eye-tracking studies) due to the restricted reading format. Comprehenders have no opportunity to revisit previous material during conventional self-paced reading  \citep[but see][]{BSPR}. This is more demanding on the comprehender's memory which might enhance interference effects. This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous}). The estimates of both main effects (syntactic and semantic interference) and the interaction in the spill-over region are very similar to the results of the present study in the critical region. This similarity between the two self-paced reading experiments further validates the findings of the present study. But it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.

The contrast between our results and the results from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items. So, why did we find a semantic interference effect and \citet{mertzen} did not? This could be due to lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study. While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40). This is also supported by the reduced uncertainty of the individual-level estimates in our data compared to \citeauthor{mertzen}'s (\citeyear{mertzen}). We will discuss the lack of syntactic interference in our data in the general discussion.

%We derived quantitative predictions from four versions of an adapted \citet{Lewis2005} model for the investigated 2 $\times$ 2 interference design \citep{mertzen}. The versions of the model differed in their cue-combinatorics (linear, non-linear) and cue-weighting (equal weighting, higher weighting of syntactic/structural cues). We described in our Predictions section that previous findings \citep{vandyke07, mertzen} were most compatible with linear cue-combinatorics with higher weighing of syntactic/structural cues. This was mainly driven by the lack of an interaction in the previous studies. In contrast, the results of the present study are only compatible with non-linear cue-combinatorics. As predicted by the non-linear cue-combinatorics model, only distractors which matched all retrieval cues induced interference; see the interaction / nested semantic interference effect. The predictions of non-linear cue-combinatorics models changed only slightly when the cue-weighting was varied (equal weighting or higher weighting of syntactic/structural cues). Therefore, we do not claim that our results provide evidence for one type of cue-weighting over the other \citep[for a detailed discussion of cue-weighting, see e.g.,][]{yadav2021individual}.

%The present findings can inform the debate on cue-combinatorics in cue-based retrieval models as they provide evidence in favor of non-linear cue-weighting \citep{parker2019cue,vandyke07, raaijmakers_shiffrin_1981, hintzman1984minerva}. Non-linear cue-weighting is an alternative to linear cue-weighting; the latter assumes a simple additive process in cue combination. In contrast, non-linear cue-weighting denotes that cues are combined via multiplication. One mismatching cue multiplies all other cues, even if they match, with zero. Therefore, a distractor which mismatches at least one of the retrieval cues will be dismissed. The present study provides evidence that cues are non-linearly combined. Distractors which only partially matched retrieval cues did not induce interference.

%The observed effect sizes are smaller than the predicted effect sizes, but that might just be due to the latency parameter in the models (the scaling parameter used to link activation levels to reading times). We used the latency parameter distribution which was previously used \citep[][]{jaeger_etal_2020, engelmann_etal_2019, vasishth2020_abc}. In future research, it might be adjusted to better match the smaller effect sizes found in large-scale self-paced reading studies of interference \citep[see][]{nicenboim}.

%The observed interaction at the critical verb, i.e., a semantic interference effect only within the high syntactic interference conditions, was predicted by syntax-first models of sentence processing. But as described above no main effect of syntactic interference was observed. Such an effect would be a prerequisite to claim primacy or dominance of syntactic features. The observed interaction and the corresponding nested comparisons rather suggest that syntactic and semantic cues are used simultaneously to guide retrieval (see constraint-based sentence processing models, \cite[e.g.,][]{macdonald1994}).

\clearpage
\section{Experiment 2: EEG}
\subsection{Methods}
This experiment was conducted with the approval of the ethics committee of the University of Potsdam, Germany.

\subsubsection{Participants}
146 participants from the University of Potsdam participant pool took part in the experiment. Three participants were excluded because they did not fulfill the demographic requirements (bilinguals or medical history). Four participants were excluded because they finished only one out of two experimental sessions. Seven participants were excluded due to EEG artifacts (below 20 artifact-free trials in at least one condition). Additionally, 29 participants were excluded because they showed poor comprehension question accuracy (below 70\,\%) in one of the experimental sessions. The data of 103 participants (mean age: 23.5 years old, age range: 18 -- 38 years old, 81 female, 22 male) were used for the analyses presented here. These final participants were all right-handed, mono-lingual native speakers of German with normal or corrected-to-normal vision and no reported history of psychiatric or neurological disease. All participants gave written informed consent and were compensated with 40 Euros per experimental session (80 Euros in total) or course credit.

\subsubsection{Procedure}
The experiment was conducted in two experimental sessions for practical reasons (each of the sessions lasted approximately two hours). Sessions were separated by one to eight weeks for each participant. The procedure of both sessions was identical.

During each experimental session, the EEG was recorded while participants were
seated in a sound-proof booth. OpenSesame was used to present sentences word-by-word \citep{opensesame}. Participants were familiarized with the procedure with two practice sentences. After that, the experiment was conducted in four blocks of 50 sentences each, presenting the items in pseudorandomized order, with breaks between the blocks. Each trial started with the presentation of a fixation cross in the center of the screen for 500 ms. Next, each word of the sentence was presented in the center of the screen. Word duration for words of interest (subject, distractor, pre-pre-critical word, pre-critical word, critical verb, post-critical word) was 500 ms. Word duration of all other words was 190 ms + 20 ms per character of the specific word. The inter-stimulus interval between all words was 400 ms. After a third of the trials, participants were asked to answer yes/no comprehension questions by pressing one of two keys on a standard keyboard. The j key was always mapped to ``yes'' answers and the f key was always mapped to ``no'' answers. The correct response was counterbalanced, so that half of the time a ``no'' response was correct and half of the time a ``yes'' response was correct.

\subsubsection{EEG recording and processing}
The EEG was recorded with 24 Ag/AgCl scalp electrodes, positioned according to the international 10-20 system. During recording, an electrode at the left mastoid was used as reference and AFz as ground. The sampling rate was 500 Hz. Eye-movements were monitored with six electrodes which were positioned above, below and at the outer canthus of both eyes. Impedances of all electrodes were kept below 5k$\Omega$.

Processing of the EEG data was carried out with MNE python \citep{mne}. The EEG was offline re-referenced to the average of the left and right mastoid electrodes. Eye-movements were corrected using independent component analysis (ICA) based on bipolar electro-oculogram channels. The data was band-pass filtered between 0.1 and 30 Hz. The data was segmented into epochs starting 200 ms preceding critical verb onset and lasting until 1000 ms following critical word onset. Epochs with artifacts were excluded automatically.

\subsubsection{Statistical analyses}
The statistical analyses of the comprehension accuracy in the EEG experiment were carried out in the same manner as the analyses of the comprehension accuracy in the SPR experiment.

We used Bayesian linear mixed models to analyze the single trial EEG data in response to the critical verb. We averaged the activity of 12 centro-parietal electrodes (Cz, C3/4, CPz, CP1/2, CP5/6, Pz, P3/4, POz) in the standard N400 time window (300 to 500 ms post critical word onset) for all analyses. The models included sum-contrast coded fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. In addition to the fixed effects of interest, all models included the baseline EEG activity from 200 ms prior to critical word onset until critical word onset as a continuous predictor. This functioned as a regression-based instead of traditional baseline correction \citep{alday2019}. Random intercepts for participants and items as well as by-participant random slopes for semantic interference were included. Correlations between varying intercepts and slopes were not included. We used informative priors for all parameters of the models (see Table \ref{tab:eeg_priors}).

\begin{table}[]
    \caption{Informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}. Priors on the slope parameters were truncated to an upper bound of zero, i.e., only negative values. This corresponds to more negative brain responses to high vs.\ low interference conditions in line with the literature. The standard deviations of the slope priors were varied in order to conduct a Bayes factor sensitivity analysis. See the corresponding assumed a priori range of the difference between high vs.\ low interference.}
    \label{tab:eeg_priors}
    \centering
    \begin{tabular}{lllr}
    \toprule
    Parameter&Prior & &Assumed Range\\
    \midrule
  Intercept & Normal(2, 5)& &\\
  \cmidrule{2-4}
  \multirow{4}{1cm}{beta} & Normal(0, 0.1) & \multirow{4}{1cm}{ub=0} & -0.2 - 0\,$\mu$V\\
  &  Normal(0, 0.5)& & -1 - 0\,$\mu$V\\
  & Normal(0, 1) && -2 - 0\,$\mu$V\\
  & Normal(0, 5) && -10 - 0\,$\mu$V\\
  \cmidrule{2-4}
  sigma & Normal(10, 5)&\\
  SD & Normal(0, 2)&\\
    \bottomrule
    \end{tabular}
\end{table}

Models were run with 4 chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}. For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats}. We chose slope priors which assume a wide range of effect sizes from -0.2 to -10\,$\mu$V, but focused on smaller effect sizes from -0.2 to -2\,$\mu$V. The focus on smaller effects is sensible because the interference effect investigated here is probably less pronounced than the well-established predictability effect on the N400 amplitude analyzed by \citet{nicenboim_stats}.

\subsection{Results}
Data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.

\subsubsection{Comprehension question accuracy}
After exclusion of participants with accuracy below 70\,\%, the overall accuracy (including fillers) was 82.8\,\% (range: 71.9 -- 100\,\%). Accuracy in critical trials was 75.6\,\% (range: 55 -- 95\,\%). By-condition accuracy is presented in Table \ref{tab:eeg_acc}.

\begin{table}[]
    \caption{By-condition accuracy in critical trials in the EEG experiment.}
    \label{tab:eeg_acc}
    \centering
    \begin{tabular}{llr}
    \toprule
    syntactic & semantic & accuracy \%\\
    \midrule
        low &  low & 81.6\\
        low &  high & 74.0\\
        high &  low & 85.1\\
        high &  high & 68.1\\
    \bottomrule
    \end{tabular}
\end{table}

Table \ref{tab:eeg_acc_mod} presents the results of the generalized mixed model analyzing the comprehension accuracy in the critical trials. The log-odds estimates and 95\,\% credible intervals show reduced comprehension accuracy in high compared to low semantic interference conditions. Additionally, the results suggest an interaction, i.e., the difference between high and low semantic interference conditions was larger when syntactic interference was high vs.\ when it was low.

\begin{table}[]
    \caption{Results in log-odds from a Bayesian generalized model analyzing the comprehension accuracy in critical trials of the EEG experiment.}
    \label{tab:eeg_acc_mod}
    \centering
    \begin{tabular}{lrrr}
    \toprule
    & Estimate & Est.Error & 95\% CrI  \\
    \midrule
Intercept& 1.39 & 0.15&  [1.10, 1.68]\\
syntactic& -0.04 & 0.06& [-0.16, 0.08]\\
semantic&  -0.48 & 0.06& [-0.61, -0.36]\\
interaction& -0.17& 0.08& [-0.34, -0.01]\\
    \bottomrule
    \end{tabular}
\end{table}

\subsubsection{ERPs}
Grand average ERPs elicited by the critical verb for all conditions are presented in Figure \ref{fig:erp_all}. Table \ref{tab:eeg_mod} shows the estimates from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference with a difference up to -1\,$\mu$V. The 95\% credible intervals of syntactic interference, semantic interference and their interaction were consistent with effects for all of them. But the estimate of the semantic interference effect was the largest and its 95\% credible intervals was the furthest from zero ($\beta$ = -0.31\,$\mu$V, 95\% CrI = [-0.49, -0.13]\,$\mu$V, for the posteriors see Figure \ref{fig:eeg_posteriors}). The estimate of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).

\begin{figure}[H]
    \caption{ERPs elicited by the critical verb (word onset at 0\,ms) at electrode Cz.}
    \label{fig:erp_all}
    \centering
    \includegraphics[width=0.95\textwidth]{images/N_103_Cz_crit.png}
\end{figure}

\begin{table}[H]
    \caption{Results from the Bayesian linear mixed model analyzing the averaged brain response from 12 centro-parietal electrodes during the time window 300 to 500 ms post critical verb onset. Baseline represents the averaged activity from the 200 ms pre-verb onset and was included in the model as a covariate instead of traditional baseline correction \citep[see][]{alday2019}.}
    \label{tab:eeg_mod}
    \centering
    \begin{tabular}{lrrr}
    \toprule
    & Estimate & Est.Error & 95\% CrI  \\
    \midrule
Intercept &  -0.70 & 0.19 &  [-1.06, -0.34] \\
baseline  &  -0.22 & 0.01 &  [-0.24, -0.20] \\
syntactic &  -0.15 & 0.08 &  [-0.32, -0.02] \\
semantic  &  -0.31 & 0.09 &  [-0.49, -0.13] \\
interaction& -0.13 & 0.10 &  [-0.37, -0.01] \\
    \bottomrule
    \end{tabular}
\end{table}

\begin{figure}[H]
    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal N400 window. Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
    \label{fig:eeg_posteriors}
    \centering
    \includegraphics[width=0.6\textwidth]{images/posteriors_eeg.jpg}
\end{figure}

The semantic interference effect is further illustrated in Figure \ref{fig:erp_sem_syn} (a) and (b). Figure \ref{fig:erp_sem_syn} (a) shows that around 400 ms, the critical verb elicited a more negative ERP under high than under low semantic interference. Figure \ref{fig:erp_sem_syn} (b) shows that the topography of the effect is rather broad but with a concentration of more negative values at centro-parietal electrodes. The timing and topography of this effect suggested that it was an N400 effect, thus the N400 was modulated by semantic interference. In contrast, \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in the N400 spatio-temporal window. Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (c) does also not suggest that there was a syntactic interference effect in any other time window.

\begin{figure}[H]
    \caption{Brain responses elicited by the critical verb. (a) ERPs for semantic interference (word onset at 0\,ms) at electrode Cz. (b) Topographic map of the semantic interference effect (high semantic interference - low semantic interference) in the N400 time window. (c) ERPs for syntactic interference  (word onset at 0\,ms) at electrode Cz. (d) Topographic map of the syntactic interference effect (high syntactic interference - low syntactic interference) in the N400 time window.}
    \label{fig:erp_sem_syn}
    \centering
         \begin{subfigure}[t]{0.74\textwidth}
         \centering
        \caption{}
         \label{fig:erp_sem_ERP}
         \includegraphics[width=\textwidth]{images/ERP_animacy_Cz.png}
     \end{subfigure}
     \hfill
     \begin{subfigure}[t]{0.25\textwidth}
         \centering
        \caption{}
         \label{fig:erp_sem_topo}
         \includegraphics[width=\textwidth]{images/topo_animate-inanimate_300_500.png}
     \end{subfigure}

    \begin{subfigure}[t]{0.74\textwidth}
         \centering
        \caption{}
         \label{fig:erp_syn_ERP}
         \includegraphics[width=\textwidth]{images/ERP_subj_Cz.png}
     \end{subfigure}
     \hfill
     \begin{subfigure}[t]{0.25\textwidth}
         \centering
        \caption{}
         \label{fig:erp_syn_topo}
         \includegraphics[width=\textwidth]{images/topo_subj-nsubj_300_500.png}
     \end{subfigure}
\end{figure}

Bayes factors are shown in Figure \ref{fig:eeg_bfs}. The Bayes factors provided very strong to strong evidence for the semantic interference effect for all priors 
(Normal$_{-}$(0, 0.1): BF$_{10}$ = 27,
Normal$_{-}$(0, 0.5): BF$_{10}$ = 88,
Normal$_{-}$(0, 1): BF$_{10}$ = 52,
Normal$_{-}$(0, 5): BF$_{10}$ = 11). In contrast, the Bayes factors provided only anecdotal evidence for the syntactic interference effect under the two narrow priors (Normal$_{-}$(0, 0.1): BF$_{10}$ = 2.5,
Normal$_{-}$(0, 0.5): BF$_{10}$ = 1.2) and evidence \textit{against} it under wider priors 
(Normal$_{-}$(0, 1): BF$_{10}$ = 0.66,
Normal$_{-}$(0, 5): BF$_{10}$ = 0.14). Across all priors, the Bayes factors provided evidence \textit{against} the interaction of semantic and syntactic interference (BF$_{10}$ < 0.85).

\begin{figure}
    \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction during the time window 300 to 500 ms post critical verb onset at centro-parietal electrode sites.}
    \label{fig:eeg_bfs}
    \centering
    \includegraphics[width=0.9\textwidth]{images/BF_plot_N103_cp_300_500.jpg}
\end{figure}

\subsection{Discussion}
We presented a large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency formation. It is the first ERP experiment on this classic interference design. Strong evidence was found for an effect of semantic interference in the N400 amplitude. Bayes factors provided considerably weaker evidence in the N400 amplitude for an effect of syntactic interference under narrow priors and evidence against it under wider priors. Evidence against an interaction of syntactic and semantic interference in the N400 was found. Comprehension accuracy was reduced for high vs.\ low semantic interference conditions. A small interaction indicated that the difference in accuracy between high and low semantic interference conditions was larger, when syntactic interference was high. Syntactic interference itself did not affect accuracy.

The to date rather small body of ERP studies on interference effects during sentence comprehension gives an indeterminate picture of the nature of interference effects in ERP measures. The present study adds a large data point to the body of literature which reported negativities or more specifically N400 effects for high vs.\ low interference conditions \citep{vasishth_drenhaus_2011, martinetal2014, schoknecht2022, lee_garnsey}. Therefore, it contributes to the debate of how interference effects manifest themselves electrophysiologically (interference effects have also been reported in differences in the P600, \cite{Tanner_etal_2017, martinetal2014}). Our finding that retrieval interference modulates the N400 amplitude supports the N400 as a marker of memory retrieval account \citep{kutas&federmeier_2000, kutas_federmeier2011, brouwer2017_n4_p6, lau2008_n400}.

\section{General discussion}
The present SPR and ERP experiments consistently showed effects of semantic interference and no effects of syntactic interference. This result was not only found in the primary dependent measures of the experiments, respectively, but also in the comprehension accuracy. Given the replicability / reproducibility crises in psycholx \citep[see e.g., ][]{vasishth2018_signficancefilter}, and other (related) sciences \citep[see e.g., ][]{openscience2015_reproducibility}, the cross-methodological consistency is astonishing. It might be due to the very high sample sizes of the present experiments which granted the present study higher power than most comparable psycholinguistic studies. The higher power led to more precise estimates of the investigated effects with less uncertainty than in previous studies. This also partially explains the comparatively smaller effect sizes in the present experiments than have been previously reported.

\subsection{Quantitative evaluation of Lewis and
Vasishth (2005) model predictions}
There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \cite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}). The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}). The syntactic effect is absent in our data.

It would be implausible to argue that the absence of the syntactic interference effect suggests that the comprehender does not use syntactic information for subject-verb dependency resolution at all. We rather suggest that the \{± grammatical subject\} cue was not the right syntactic cue to investigate. In line with \citet{mertzen}, we suggest that a syntactic or structural cue along the lines of \{± same clause\} might be used during comprehension. Figure \ref{fig:model_predictions} presents the predictions of the \citet{Lewis2005} model for the \citet{vandyke07} design where additionally to the \{± animate\} cue either the canonical cue \{± grammatical subject\} or the alternative cue \{± same clause\} was implemented. The \{± same clause\} cue model does predict the syntactic interference effect to be centered around zero (95\,\% CrI [-4, 3]\,ms). This prediction is consistent with the results of the present study. However, the \{± same clause\} model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms).

\begin{figure}[H]
    \centering
    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the Lewis \& Vasishth (2005) model as implemented in R and described by Engelmann et al. (2019). The implemented model is available at https://github.com/felixengelmann/inter-act/.}
    \label{fig:model_predictions}\includegraphics[width=\textwidth]{images/PriorPredicted_LV05_engelmann_effects_syncue.png}
\end{figure}

The \citet{Lewis2005} model is generally used to predict reading times. The scaling factor \textit{F} is optimized to link the activation of items in memory to reading times \citep{Lewis2005}. But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data. The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor. Therefore, the \{± same clause\} cue instead of the \{± grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.

\subsection{Encoding and retrieval interference}
In the SPR data, we have found that the semantic interference effect originated at the distractor and persisted throughout following regions. Due to this time course, encoding interference was the best explanation for the interference effect in the reading times of the critical region (see discussion of Experiment 1). This raises the question whether the ERP effect was also caused by encoding interference, i.e., whether the ERP effect originated prior to the critical region. To answer this question, it is necessary to look at the ERPs elicited by words earlier in the sentence, although that is unusual in ERP research. Figure \ref{fig:erp_precrit} shows the ERPs for all conditions at the critical word and the two pre-critical adverbs which were identical across conditions. It is apparent from the ERPs of the pre-pre-critical and pre-critical word that the semantic interference effect found at the critical verb was not present at the words preceding it.

\begin{figure}[H]
    \caption{ERPs elicited by the pre-pre-critical adverb (word onset at -1800 ms), pre-critical adverb (word onset at -900 ms) and critical verb (word onset at 0\,ms) at electrode Cz.}
    \label{fig:erp_precrit}
    \centering
    \includegraphics[width=0.95\textwidth]{images/N_103_Cz_precrit.png}
\end{figure}

Now that we have seen that the words directly preceding the critical word did not show the semantic interference effect, one might ask which brain response was elicited by the distractor itself (see Figure \ref{fig:erp_distractor})? Visual inspection of the ERPs elicited by the distractor shows several differences between conditions, but no semantic interference effect in the N400 spatio-temporal window. We assume that the differences in the ERPs elicited by the distractor in the different conditions are caused by two factors. Firstly, the high and low syntactic interference conditions have different word orders, i.e., the distractors are embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase, see Example \ref{ex:materials}). This is also apparent from the large difference between high and low syntactic interference conditions before distractor onset. Secondly, the distractors were not identical across conditions, therefore the ERPs in Figure \ref{fig:erp_distractor} are elicited by different words in the high vs.\ low semantic interference conditions. In conclusion, the distractors induced different brain responses across conditions, but these did not linger throughout the rest of the sentence (see Figure \ref{fig:erp_precrit}). %Firstly, there is a large difference prior to distractor onset (before time 0 ms) at all electrode sites. Secondly, there is a difference between 300 and 500\,ms most at anterior (left) electrode sites. We assume that both these differences are due to the different word orders in the high vs. low syntactic interference conditions, i.e., the different phrase structures that the distractors are embedded in (simple noun phrase vs. noun phrase including an adjective within a prepositional phrase, see Example \ref{ex:materials}). Thirdly, there is a late positivity (starting at approx. 500\,ms) most pronounced at centro-parietal electrodes for high semantic vs. low semantic interference conditions. This semantic interference effect in the amplitude of a late positivity could be either due to

\begin{figure}[H]
    \centering
        \caption{ERPs elicited by the distractor (word onset at 0 ms) at eight selected electrodes.}
    \label{fig:erp_distractor}\includegraphics[width=\textwidth]{images/Distractor_N_100_some_elec.png}
\end{figure}

So, the comparison of the results of our SPR and ERP experiments yields that the SPR results were strongly affected by encoding interference, while the ERPs elicited by the critical verb were not. This difference between methods might be caused by the difference in presentation rate. In the self-paced reading experiment, as implied by the name, the participants had control over the presentation rate of the stimuli. The observed slowdown starting at the distractor  might have resulted from the probably subconscious intent to slow down the sampling rate of the incoming stimuli to gain more time to process them. In contrast, in the ERP experiment, we used rapid serial visual presentation (RSVP), which presents stimuli at a fixed rate that is beyond the control of the participant. Therefore, they must adapt to the set rate and process the stimuli as fast as the rate requires. Here, it is noteworthy that the presentation duration of 500\,ms per word and 400\,ms between words which was used in the present ERP experiment is slower than usual natural reading speed. However, while in SPR the slowdown due to encoding and maintaining the distractor in memory overshadowed potential retrieval effects, the procedure of the ERP experiment prevented that. In the ERP results, we observed an effect at the critical verb, which can be attributed to memory retrieval of the subject. This retrieval was only affected by the semantic cue \{± animate\} and not by the syntactic cue \{± grammatical subject\}.

\subsection{Relation to other sentence processing accounts}
Besides our considerations above regarding the possibility that another syntactic cue (e.g., \{± same clause\}) was used, our results highlight the importance of the semantic cue for subject-verb dependency resolution.

The evidence against a main effect of syntactic interference and hence the relative unimportance of the syntactic cue \{± grammatical subject\} for subject-verb dependency resolution in our study stands in contrast to predictions from syntax-first accounts \citep[e.g.,][]{frazier_clifton1996, frazier1979_diss, frazier1987}. These accounts predict that syntactic constraints are considered first during sentence processing. At least, our results speak against syntax as a strong filter. However, we found that a semantically similar distractor induced interference. This finding can be easily reconciled with various accounts of sentence processing. Above we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model. Our results fit predictions from constraint-based models \citep[e.g.,][]{macdonald1994, mcrae1998modeling, trueswell1993}. Under constraint-based accounts, it is assumed that multiple constraints simultaneously and immediately influence processing. These accounts predict our ERP results: Compared to an inanimate distractor, an animate distractor -- regardless of its syntactic status -- would impede processing at a verb which requires an animate subject. The lack of a syntactic interference effect could indicate that the status `grammatical subject' was not used as a syntactic constraint. As discussed above, a plausible syntactic constraint is that a verb and its arguments need to be within the same clause. Furthermore, our results are in line with the good-enough processing account \citep{ferreira2007goodenough}. This account assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete or even incorrect analyses. The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials posed high processing demands. It is reasonable to assume that the participants might have sometimes adopted a good-enough processing mode to manage resources. So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution. Animacy is an obvious choice for such a heuristic. Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject probably leads to the correct analysis (not just within the experiment context, but in everyday language use). Finally, our findings support the possibility that language processing relies predominantly on semantic associations to form (probabilistic)
representation of event structures. This view has been put forward by \citet{rabovsky_etal_2018}, when they used the sentence gestalt model \citep{mcclelland1989_sentence_gestalt}, a neural network model of sentence comprehension, to model the N400 amplitude. This model assumes that language comprehension relies on associative form-to-meaning mapping instead of syntactic rules. This is in line with our results which highlight the importance of semantic features such as animacy over syntactic functions such as grammatical subject for sentence processing.

\section{Conclusion}
%The present large-scale self-paced reading study examined the role of syntactic and semantic features during sentence comprehension. At a verb that initiated retrieval of its subject, high semantic interference led to longer reading times than low semantic interference, but only when syntactic interference was high as well. This result provided empirical support for the assumption that different types of cues are used simultaneously during memory retrieval in subject-verb dependency formation. Furthermore, it suggests that distractors must match all available retrieval cues to cause similarity-based retrieval interference. The results of the present study are similar to the predictions of cue-based retrieval models of sentence processing with non-linear combinatorics.

This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials. While our results did not directly replicate previous results from eye-tracking while reading studies \citep{vandyke07, mertzen}, our SPR and ERP results were consistent. We attributed differences between the present and previous studies to differences in statistical power and experimental methodology. Both our experiments showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor. This manifested in longer self-paced reading times starting at the distractor and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb. We discussed that these different time courses in the two experiments are due to different types of interference: Encoding interference in the SPR experiment and retrieval interference in the ERP experiment. We did not find effects of syntactic interference. We assume that the syntactic cue \{± grammatical subject\} which was manipulated in our materials, was not used during sentence processing. We proposed that another syntactic cue \{± same clause\} might be used \citep[see also][]{mertzen}. We suggest that the present findings, which are in line with various sentence processing accounts, highlight the importance of semantic cues during sentence processing.

\newpage
\printbibliography

\appendix

\section{Previous work on syntactic and semantic interference in reading}

But are both syntactic and semantic cues used during memory retrieval? \citet{vandyke07} and \citet{mertzen} investigated the effects of syntactic and semantic interference during long-distance subject-verb dependencies (see Example \ref{ex:1_vandyke07}) in a series of self-paced reading and eye-tracking while reading experiments. Overall, the original study by \citet{vandyke07} and the replication by \citet{mertzen} showed relatively consistent results. They consistently found syntactic interference but only rarely semantic interference (only in some measures of some of the experiments and sometimes in regions outside of the critical region). These results are not in line with the predictions from the standard cue-based retrieval models \citep{Lewis2005, mcelree}: While both previous studies found effects of syntactic and semantic interference, they emerged in different regions -- sometimes even prior to the retrieval site. Figure \ref{fig:previous} gives an overview of the previous findings. See the appendix for a detailed discussion of the previous results.

One possibly problematic aspect of the studies from \citet{vandyke07} and \citet{mertzen} is that effects were found in pre-critical regions. \citet{mertzen} suggested that the pre-critical effects could be due to structural confounds, encoding interference, preview effects during natural reading in the eye-tracking experiments or anticipatory retrieval at the pre-critical word because participants could predict that the verb would follow. It is unclear from the previous studies when exactly the pre-critical effects emerged, i.e., when the conditions started to differ. But it is important to know the timecourse of the pre-critical effects to determine their cause. If the structural confounds explanation is correct, then the pre-critical effects should appear only in relation to the syntactic manipulation. For the encoding interference explanation to be correct, it would be necessary that conditions start to differ at (or shortly after) the distractor. The preview and anticipatory retrieval explanations can only apply if the pre-critical effects are found only close to the critical verb.

The previous estimates had relatively wide 95\,\% confidence/credible intervals in some of the experiments, which is probably due to the relatively small sample sizes. \citet{vandyke07} had a typical number of participants and items for psycholinguistic experiments (35 - 40 participants per experiment, 36 - 48 items per experiment). \citet{mertzen} used the same number of items as \citet{vandyke07}, but larger participant sample sizes (English: N = 61, German N = 121). Large uncertainty raises the possibility of Type M errors, leading to misleading conclusions \citep{gelman_carlin}.

\begin{figure}[H]
    \caption{Overview of previous findings of syntactic and semantic interference and their interaction by \citet{vandyke07} (three experiments in English with 35, 36 or 40 participants) and \citet{mertzen} (one experiment in German with 121 participants, one experiment in English with 61 participants). Experiments used either Self-paced Reading (SPR) or Eye-tracking methodology (Eye-Tracking measures: FPRT = first-pass reading time, RPD = regression-path duration, TFT = total fixation time). Error bars are 95\% confidence intervals for \citet{vandyke07} and 95\% credible intervals for \citet{mertzen}. }
    \label{fig:previous}
    \centering
    \includegraphics[width=0.95\textwidth]{images/previous_findings_allregions_long.png}
    \figurenote{The estimates for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) data were derived form the published paper. For non-significant effects, where no statistics were reported, we used a t-value of 1.95 to compute the confidence intervals. \citet{mertzen} provided their estimates and credible intervals.}
\end{figure}

The results of the previous studies raise two questions: 1) Why were semantic interference effects found less consistently than syntactic interference effects? 2) Why were effects of retrieval interference observed in the pre-critical region? To answer question 1, we decided to run large-sample studies revisiting \citeauthor{vandyke07}'s (\citeyear{vandyke07}) design to gain more precise estimates of the effects of syntactic and semantic interference. To answer question 2 -- or to disentangle the possible causes of the pre-critical effects -- we decided it is best to use methods which provide responses which are unambiguously related to the processing of single words. The methods of our choice being self-paced reading (SPR) and electroencephalography (EEG) or event-related potentials (ERPs), respectively, with typical rapid serial visual presentation (RSVP) of single words. Both these methods provide responses directly linked to the processing of single words and, unlike eye-tracking while reading, do not grant any preview of upcoming stimuli. To reduce the structural confound at the pre-critical region identified by \citet{mertzen}, we added an additional pre-pre-critical region. These changes in the experimental design between the present and previous studies will help to distinguish between the different explanations for the pre-critical effects put forward by \citet{mertzen}.

In the previous studies, surprising effects were not only found in the pre-critical region, but also in the sentence-final region: Specifically, \citet{vandyke07} found semantic interference effects at the sentence-final region. She proposed that the effect was delayed to this late region because ``it simply takes longer for individuals to compute the semantic association between the distractor and the verb'' \citep[][p. 427]{vandyke07}. An alternative explanation for this late effect is that eye-movement control constraints can push the eye to move on before processing is finished. In contrast, ERPs are a time-sensitive direct measure of brain activity in response to the stimuli without confounds of planning and execution of motor activity (i.e., planning of saccades during natural reading in eye-tracking experiments or planning of button presses in SPR experiments). The well-studied N400 ERP component is associated with integration in the sentence context and / or memory retrieval induced by the current stimulus \citep{brouwer2017_n4_p6, lau2008_n400, kutas&federmeier_2000, kutas_federmeier2011}. The N400 is a negativity that peaks 400 ms after a word appeared on the screen in word-by-word presentation. The canonical finding is that the N400 amplitude is more negative for words which provide a poor semantic fit into the sentence context compared to words which fit the context perfectly \citep{kutas_hillyard}. Therefore, the N400 can provide insights into the immediate semantic processing of a word. This should prevent a potential semantic interference effect to be only detected at later regions in the sentence.

Although the N400 is primarily associated with semantic processing, if one follows the memory retrieval account it is reasonable to assume that retrieval of all aspects of a memory entry will modulate the N400 amplitude. The potential complications during memory retrieval due to interference should not differ dependent on the specific cue used for retrieval, i.e., \{±grammatical subject\} or \{±animate\}.  Moreover, previous ERP studies investigating interference have found negativities during the typical N400 time window (300 - 500 ms post-stimulus onset), although the scalp topographies were not typical for the N400, being either broad \citep{schoknecht2022} or left anterior \citep{vasishth_drenhaus_2011}. Therefore, ERPs and especially the N400, are suited to be used for the investigation of interference (in the design from \citet{vandyke07} and in general).  We chose to focus our EPR analyses on the N400 and expected to find interference effects -- both syntactic and semantic -- in N400 amplitude differences due to a more complex memory retrieval under high compared to low interference.

\section{I don't know why this is here}
In their \citeyear{vandyke07} paper, \citeauthor{vandyke07} reported a series of reading studies (one self-paced reading experiment and two eye-tracking experiments). For the eye-tracking experiments, \citet{vandyke07} reported three reading time measures: First-pass reading times (FPRT), regression-path durations (RPD) and total fixaton times (TFT). Reading times in multiple regions were reported, but the critical region was the verb. At the verb, \citet{vandyke07} found syntactic interference in both eye-tracking experiments and in all measures, but semantic interference was only found in Experiment 2 and only in RPD and TFT. In contrast, the spill-over region showed semantic interference more consistently than semantic interference. Surprisingly, the pre-critical region, which was only present in \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3, showed main effects of syntactic and semantic interference. On top of that, the RPDs of the sentence-final region in Experiment 2 and 3 were consistent with effects of syntactic and  semantic interference.

\citeauthor{mertzen}'s (\citeyear{mertzen}) estimates at the pre-critical region showed slowdowns for high vs.\ low  syntactic and semantic interference in English and German in RPD and TFT. In the reading times of the critical verb, \citeauthor{mertzen}'s (\citeyear{mertzen}) estimates were consistent with syntactic interference effects in all measures in English and German. Additionally, the estimates suggested a semantic interference effect in English. In the spill-over region, the estimates only suggested a semantic interference effect in German.

\end{document}

### 2023-10-07 code_data/EEG/r_scripts/acc_analysis.R

@@ -1,0 +1,201 @@
+# author: Pia Schoknecht
+# date: 30.07.2023
+
+# load packages
+library(tidyverse)
+library(brms)
+
+
+# load all .csv files from logs folder
+tbl <- list.files(path = "logs/", pattern = "*.csv", full.names = TRUE) %>% 
+                  map_df(~read_csv(., col_types = cols(.default = "c"))) 
+
+tbl$subject_nr <- ifelse(tbl$subject_nr == "136" & tbl$datetime == "Mon May  8 10:21:50 2023", "135", tbl$subject_nr)
+
+
+# select relevant columns, delete practice items
+dat <- tbl %>% select(subject_nr, session, lst, item, cond, cond_name, correct, correct_response,  response, response_time, exp_sentence, exp_question, is_question, question_type, np_in_question, live_row,experiment_file, datetime,  opensesame_version) %>%
+              drop_na(exp_sentence)
+
+
+dat2 <- dat %>% mutate(subject_nr = fct_recode(subject_nr, "1" = "200",  "3" = "7", "125" = "155"))
+
+
+# number of participants
+(N_subj <- length(unique(dat2$subject_nr)))
+
+# exclude participants who were excluded due to low accuracy or to many artefacts
+ex_s1 <- read.table("exclude_s1.txt")
+ex_s1$subj <- as.factor(ex_s1$subj)
+ex_s2 <- read.table("exclude_s2.txt")
+ex_s2$subj <- as.factor(ex_s2$subj)
+
+ex_arte <- read.csv("../TooManyArtefacts.csv")
+ex_arte$subject <- as.factor(ex_arte$subject)
+
+
+dat3 <- dat2 %>% anti_join(ex_s1, by=c("subject_nr"="subj")) %>%
+                 anti_join(ex_s2, by=c("subject_nr"="subj")) %>%
+                 anti_join(ex_arte, by=c("subject_nr"="subject"))%>%
+                 droplevels()
+
+(N_subj <- length(unique(dat3$subject_nr)))
+nrows_all <- dat3 %>% count(subject_nr)
+
+
+#### overall accuracy including fillers ####
+acc <- filter(dat3, is_question == "yes")
+
+# by-participant accuracy (per session)
+acc_s1 <- acc %>% filter(session == "1")
+acc_s2 <- acc %>% filter(session == "2")
+
[…]

### 2023-10-07 code_data/EEG/r_scripts/analyses.R

@@ -1,0 +1,437 @@
+# author: Pia Schoknecht
+# date: July 20, 2023
+# R version: 4.1.3
+
+#load packages
+library(tidyverse)
+library(dplyr)
+library(ggplot2)
+library(brms)
+library(bayesplot)
+library(truncnorm)
+library(plyr)
+library(cowplot)
+library(lattice)
+library(graphics)
+library("gridExtra")     ## for grid.arrange()
+library(stats)
+options(scipen = 999)
+
+#load data
+prestim <- read_csv("data/Pandora_mean_200_0.csv")
+n400 <- read_csv("data/Pandora_mean_300_500.csv")
+
+# remove subjects who have only one good session
+xtabs(~ session + subject, n400)
+n400.2 <- n400 %>% filter(subject != "0001" & subject != "0005" & subject != "0006" & subject != "0011" & subject != "0021"
+                      & subject != "0040" & subject != "0041" & subject != "0055" & subject != "0062" & subject != "0067" 
+                      & subject != "0069" & subject != "0079" & subject != "0085" & subject != "0089" & subject != "0094" 
+                      & subject != "0095" & subject != "0106" & subject != "0122" & subject != "0123" & subject != "0129"
+                      & subject != "0150")
+n400.2$subject <- factor(n400.2$subject)
+
+prestim.2 <- prestim %>% filter(subject != "0001" & subject != "0005" & subject != "0006" & subject != "0011" & subject != "0021"
+                                & subject != "0040" & subject != "0041" & subject != "0055" & subject != "0062" & subject != "0067" 
+                                & subject != "0069" & subject != "0079" & subject != "0085" & subject != "0089" & subject != "0094" 
+                                & subject != "0095" & subject != "0106" & subject != "0122" & subject != "0123" & subject != "0129"
+                                & subject != "0150")
+prestim.2$subject <- factor(prestim.2$subject)
+
+# how many trials were artefact-free?
+data_new <- n400.2[!duplicated(n400.2[, c("subject", "cond", "item")]), ]
+keptseg <- as.data.frame(table(data_new$subject, data_new$cond))
+names(keptseg)[names(keptseg) == "Var1"] <- "subject"
+names(keptseg)[names(keptseg) == "Var2"] <- "cond"
[…]

### 2023-10-07 code_data/SPR/r_scripts/Compute_individual_estimates.R

@@ -1,0 +1,145 @@
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(reshape2)
+library(rstan)
+library(brms)
+library(cowplot)
+
+
+load("model_fits/Fit_m_Full_RSLOPES.Rda")
+
+#variables(m_m_full)
+samples <- rstan::extract(m_m_full$fit)
+
+nsubj <- 774
+nsamples <- nrow(samples$b_Intercept)
+
+SynInt_subj <- matrix(rep(NA,nsamples*nsubj),ncol=nsubj)
+SemInt_subj <- matrix(rep(NA,nsamples*nsubj),ncol=nsubj)
+SynSemInt_subj <- matrix(rep(NA,nsamples*nsubj),ncol=nsubj)
+
+
+for(i in 1:nsubj){
+  SynInt_subj[,i] <-
+    exp((samples$b_Intercept + samples$r_2_1[,i]) + (0.5*(samples$b[,1] + samples$r_2_2[,i])))-
+    exp((samples$b_Intercept + samples$r_2_1[,i]) - (0.5*(samples$b[,1] + samples$r_2_2[,i])))
+  SemInt_subj[,i] <-
+    exp((samples$b_Intercept + samples$r_2_1[,i]) + (0.5*(samples$b[,2] + samples$r_2_3[,i])))-
+    exp((samples$b_Intercept + samples$r_2_1[,i]) - (0.5*(samples$b[,2] + samples$r_2_3[,i])))
+ SynSemInt_subj[,i] <-
+    exp((samples$b_Intercept + samples$r_2_1[,i]) + (0.5*(samples$b[,3] + samples$r_2_4[,i])))-
+    exp((samples$b_Intercept + samples$r_2_1[,i]) - (0.5*(samples$b[,3] + samples$r_2_4[,i])))
+}
+
+subj_estimates <- NULL
+
+effect_subj.m <- melt(SynInt_subj)
+colnames(effect_subj.m) <- c("sample_id","subject","effect")
+as.data.frame(effect_subj.m %>% group_by(subject) %>% summarise(Effect=mean(effect)))
+effect_subj.m$condition <- "SynInt"
+subj_estimates <- rbind(subj_estimates,effect_subj.m)
+
+effect_subj.m <- melt(SemInt_subj)
+colnames(effect_subj.m) <- c("sample_id","subject","effect")
[…]

### 2023-10-07 code_data/SPR/r_scripts/Plot_stats.R

@@ -1,0 +1,213 @@
+# author: Pia Schoknecht
+# 27.07.2023
+
+library(tidyverse)
+library(brms)
+library(bayesplot)
+library(ggplot2)
+library(ggbreak)
+library(ggimage) 
+library(cowplot)
+
+
+
+### Plot Posteriors ###
+
+# load model fits
+load("model_fits/Fit_m_Full_precrit.Rda")
+m_precrit <- m_m_full
+load("model_fits/Fit_m_Full_crit.Rda")
+m_crit <- m_m_full
+load("model_fits/Fit_m_Full_spill.Rda")
+m_spill <- m_m_full
+
+m_precrit
+m_crit
+m_spill
+
+samples_precrit <- as_draws_df(m_precrit)
+samples_crit <- as_draws_df(m_crit)
+samples_spill <- as_draws_df(m_spill)
+
+syn_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$b_syn) - exp(samples_precrit$b_Intercept - samples_precrit$b_syn)
+sem_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$b_sem) - exp(samples_precrit$b_Intercept - samples_precrit$b_sem)
+int_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit[,4]) - exp(samples_precrit$b_Intercept - samples_precrit[,4])
+
+round(c(mean = mean(syn_eff_precrit), quantile(syn_eff_precrit, probs = c(.025, .975))))
+round(c(mean = mean(sem_eff_precrit), quantile(sem_eff_precrit, probs = c(.025, .975))))
+round(c(mean = mean(int_eff_precrit[-1,]), quantile(int_eff_precrit[,1], probs = c(.025, .975))))
+
+samples_precrit$syntactic <- syn_eff_precrit
+samples_precrit$semantic <- sem_eff_precrit
+samples_precrit$interaction <- int_eff_precrit
+
+
+posteriors_precrit <- mcmc_areas(
+                          samples_precrit,
+                          pars = c("syntactic", "semantic", "interaction"),
+                          prob = 0.8, # 80% intervals
+                          #prob_outer = 0.99, # 99%
+                          point_est = "mean")+
+                          labs(title= "Pre-critical", x = "effect in ms")+
+                          vline_0(linetype="dashed")+
+                          xlim(0,35)+
[…]

### 2023-10-07 code_data/SPR/r_scripts/acc_analysis.R

@@ -1,0 +1,118 @@
+# author: Pia Schoknecht
+# date: 30.07.2023
+
+# load packages
+library(tidyverse)
+library(brms)
+
+# load data
+df1 <- read.csv("../pandora_web/data/pandora_web_data_allsessions_short.csv", sep=";")
+df2 <- read.csv("../pandora_repS1/data/pandora_repS1_data_cleaned_short.csv", sep=";")
+df3 <- read.csv("../pandora_repS2/data/pandora_repS2_data_cleaned_short.csv", sep=";")
+
+# add session info to df2 and df3 (both experiments had only one session)
+df2$session <- 1
+df3$session <- 1
+
+# delete some columns in df1 to have identical columns in all data sets
+df1$delay <- NULL
+df1$dist_len <- NULL
+
+# combine data sets
+df <- rbind(df1, df2, df3)
+
+
+# Over-all Accuracy 
+# subset with only one row per subject and item with question
+acc <- df %>% 
+  drop_na(question_rt) %>% 
+  filter(wordno == 01)# %>%
+# filter(question_rt < 20000)
+
+#calculate accuracy per participant
+acc_subj <- table(acc$correct, acc$fullname, acc$session) %>%
+  as.data.frame() %>%
+  rename(answ = "Var1", subj = "Var2", session = "Var3")
+
+#answers in percent (number of questions per subject: 32)
+acc_subj$percent <- acc_subj$Freq/32*100
+
+#how many questions were answered correctly?
+# (exclude Freq=0 because that's only the case when a subject did not participate in a session)
+acc_cor <- subset(acc_subj, acc_subj$answ == "1" & acc_subj$Freq != 0)
+round(mean(acc_cor$percent),2)
+round(range(acc_cor$percent),1)
+
+ggplot(acc_cor, aes(x=percent)) +
+  geom_density(aes()) +
+  ggtitle("accuracy") + 
+  theme_bw()
+
+# response time
+aggregate(question_rt ~ correct, data= acc, FUN=mean)
+
+# per condition
+aggregate(question_rt ~ correct + condition, data= acc, FUN=mean)
+
+# Exclude fillers
+df2 <- droplevels(subset(df, label != "filler"))
+
+# Contrast coding
+# high = 0.5, low = -0.5
+df2$syn <- ifelse(df2$condition%in%c("a","b"),-0.5,0.5)
+df2$sem <- ifelse(df2$condition%in%c("b","d"),0.5,-0.5)
+
[…]

### 2023-10-07 code_data/SPR/r_scripts/analyses_crit.R

@@ -1,0 +1,278 @@
+# author: Pia Schoknecht
+# date: 22.07.2023
+
+library(tidyverse)
+library(lme4)
+library(brms)
+library(ggplot2)
+library(bayesplot)
+library(truncnorm)
+
+
+# load data
+crit_trim <- read.csv("data/pandora_spr_774_crit.csv")
+
+
+# Inferential statistics
+# truncated varying priors 
+# (only positive effects -- longer reading times for high interference)
+
+priors_s_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+)  
+
+priors_m_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+# check priors
+round(exp(6-2*0.6));round(exp(6+2*0.6)) # intercept
+round(exp(6+0.01) - exp(6-0.01))  # beta normal-(0, 0.01)
+round(exp(6+0.05) - exp(6-0.05))  # beta normal-(0, 0.05)
+round(exp(6+0.1) - exp(6-0.1))    # beta normal-(0, 0.1)
+
+
+# models
+m_s_full <- brm(rt ~ 1 + syn * sem +
+                 (1 | participant) +
+                 (1 | item),
+               data = crit_trim,
+               family = lognormal(),
+               prior = priors_s_tr,
+               warmup = 2000,
+               iter = 20000,
+               cores = 4,
+               save_pars = save_pars(all = TRUE)
+               )
+save(m_s_full, file = paste("model_fits/Fit_s_Full_crit.Rda"))
+
+m_m_full <- brm(rt ~ 1 + syn * sem +
+                  (1 | participant) +
+                  (1 | item),
+                data = crit_trim,
+                family = lognormal(),
+                prior = priors_m_tr,
+                warmup = 2000,
[…]

### 2023-10-07 code_data/SPR/r_scripts/analyses_precrit.R

@@ -1,0 +1,265 @@
+# author: Pia Schoknecht
+# date: 30.07.2023
+
+# script to analyze reading times in the pre-critical region
+
+# load packages
+library(tidyverse)
+library(truncnorm)
+library(brms)
+library(ggplot2)
+library(bayesplot)
+
+
+# load data
+precrit_trim <- read.csv("data/pandora_spr_774_precrit.csv")
+
+
+# Inferential statistics
+# truncated varying priors 
+# (only positive effects -- longer reading times for high interference)
+
+priors_s_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+)  
+
+priors_m_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+# check priors
+round(exp(6-2*0.6));round(exp(6+2*0.6)) # intercept
+round(exp(6+0.01) - exp(6-0.01))  # beta normal-(0, 0.01)
+round(exp(6+0.05) - exp(6-0.05))  # beta normal-(0, 0.05)
+round(exp(6+0.1) - exp(6-0.1))    # beta normal-(0, 0.1)
+
+
+# models
+m_s_full <- brm(rt ~ 1 + syn * sem +
+                 (1 | participant) +
+                 (1 | item),
+               data = precrit_trim,
+               family = lognormal(),
+               prior = priors_s_tr,
+               warmup = 2000,
+               iter = 20000,
+               cores = 4,
+               save_pars = save_pars(all = TRUE)
+               )
+save(m_s_full, file = paste("model_fits/Fit_s_Full_precrit.Rda"))
+
+m_m_full <- brm(rt ~ 1 + syn * sem +
+                  (1 | participant) +
+                  (1 | item),
+                data = precrit_trim,
+                family = lognormal(),
[…]

### 2023-10-07 code_data/SPR/r_scripts/analyses_spill.R

@@ -1,0 +1,264 @@
+# author: Pia Schoknecht
+# date: 22.07.2023
+
+library(tidyverse)
+library(lme4)
+library(brms)
+library(ggplot2)
+library(bayesplot)
+library(parallel)
+library(truncnorm)
+
+
+# load data
+spill_trim <- read.csv("data/pandora_spr_774_spill.csv")
+
+
+# Inferential statistics
+# truncated varying priors 
+# (only positive effects -- longer reading times for high interference)
+
+priors_s_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+)  
+
+priors_m_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+# check priors
+round(exp(6-2*0.6));round(exp(6+2*0.6)) # intercept
+round(exp(6+0.01) - exp(6-0.01))  # beta normal-(0, 0.01)
+round(exp(6+0.05) - exp(6-0.05))  # beta normal-(0, 0.05)
+round(exp(6+0.1) - exp(6-0.1))    # beta normal-(0, 0.1)
+
+
+# models
+m_s_full <- brm(rt ~ 1 + syn * sem +
+                 (1 | participant) +
+                 (1 | item),
+               data = spill_trim,
+               family = lognormal(),
+               prior = priors_s_tr,
+               warmup = 2000,
+               iter = 20000,
+               cores = 4,
+               save_pars = save_pars(all = TRUE)
+               )
+save(m_s_full, file = paste("model_fits/Fit_s_Full_spill.Rda"))
+
+m_m_full <- brm(rt ~ 1 + syn * sem +
+                  (1 | participant) +
+                  (1 | item),
+                data = spill_trim,
+                family = lognormal(),
+                prior = priors_m_tr,
[…]

### 2023-10-08 paper/main.tex
edited up to plausibility rating

@@
 For example, \cite{dillon2013} argued, following \cite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built; instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
-\cite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference \parencite[cf.
+Using computational modeling, \cite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference, and that the reason for this is that principle A of the binding theory is the primary driver of the cue-based retrieval process \parencite[cf.
 ][]{jaeger_etal_2020,yadav2021individual}.
-The broader implication here is that the cues used during retrieval can modulate processing difficulty in possibly surprising ways.
+This was a surprising discovery because a priori, cue-based retrieval accounts do not allow for exceptions to retrieval interference in particular syntactic dependency types.
+More broadly, an important implication of the work initiated by \cite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
 
@@
 
-We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with a series of experiments using self-paced reading and EEG.
-All experiments used the 2 $\times$ 2 design by \citet{vandyke07} crossing syntactic and semantic interference.
-This is -- to our knowledge -- the first ERP study on this classic design investigating interference during subject-verb dependency resolution.
+Reading studies have been the dominant method used to investigate retrieval interference \parencite[see the review in][]{jaeger_etal_2017}.
+We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading as well as EEG, which to our knowledge has never been used to investigate interference during subject-verb dependency resolution
+
+All experiments used the 2 $\times$ 2 design developed by \citet{vandyke07}; the design crosses syntactic and semantic interference.
+
+% Wrong location for all this detail:
+%This is -- to our knowledge -- the first .
 The SPR experiments were conducted web-based via Prolific and the EEG experiments required participants to come into the lab.
@@
 \begin{figure}[H]
-    \caption{Structure of the present study.
+    \caption{The structure of the present study.
 In total, 120 items were used for the SPR (red branches) and the ERP experiments (blue branches).
@@
 \subsection{Materials}
+
 As the same materials were used for all our experiments, we describe them here once before we describe the specifics of the experiments, separately.
-The materials for the study consisted of 120 items with 4 conditions in a Latin square design shown in \ref{ex:materials}.
+The studies had 120 items with four conditions in a Latin square design shown in \ref{ex:materials}.
 The items were constructed following \citeauthor{vandyke07}'s (\citeyear{vandyke07}) items and partially re-used  \citeauthor{mertzen}'s (\citeyear{mertzen}) items.
@@
 During this retrieval, the distractor \textit{Einbrecher} (`burglar') / \textit{Verlust} (`loss') might cause interference and impede comprehension.
-Syntactic interference is manipulated via the syntactic status of the distractor, which is either subject (\ref{ex:hisyn}) or part of a prepositional phrase (\ref{ex:losyn}).
-During the retrieval of the subject, a distractor in subject position is assumed to cause more syntactic interference than a distractor inside a prepositional phrase.
+
+Syntactic interference is manipulated via the syntactic status of the distractor, which is either subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}).
+Following \cite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in any subject position causes more syntactic interference than a distractor inside a prepositional phrase.
+
 Semantic interference is manipulated via the animacy of the distractor.
 During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference.
-In contrast to the previous studies (see Introduction), we inserted a second pre-critical adverb, i.e., a pre-pre-critical region, to absorb potential spill-over effects from the embedded clause(s).
-So, potential wrap-up effects from the embedded clause(s) could spill-over into the pre-pre-critical region and the pre-critical region would be further away from clause boundaries.
-This way potential pre-critical effects are less likely to be caused by structural confounds.
+
+It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the pre-critical and critical regions.
+The spillover of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power.
+In order to attenuate spillover from the clause-final word in the embedded clause preceding the critical region, we inserted two adverbs.
+This way, the critical region will be less affected by the clause-boundary slowdown.
+This design also has the advantage that any potential effects seen at the pre-critical region (the adverb preceding the verb) are less likely to be due to spillover; it is important to look at the pre-critical region as previous work with this design found effects at the pre-critical region (we revisit this detail in the General discussion).
 
@@
 \citeauthor{vandyke07} conducted a pre-test to ensure that the distractor in high semantic interference conditions was a sufficiently probable subject of the critical verb, i.e., that it could potentially induce semantic interference.
-For the same purpose, we conducted a plausibility norming experiment.
+Following \citeauthor{vandyke07}, we also conducted a plausibility norming experiment.
 
@@
 Forty participants (18 female, 22 male, mean age: 27.3 years old, age range: 21 - 39 years old) from the SPR experiments were re-invited for the plausibility norming one month after they completed the SPR experiment.
+We used the same participants who participated in the SPR study in order to ensure that the ratings for the norming came from the same pool of participants that did the reading study.
 
 \paragraph{Materials}
-For each item quadruplet, four simple SVO sentences were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb.
-The norming materials corresponding to the item despicted in \ref{ex:materials} are shown in \ref{ex:materials_norming}.
+For each item quadruplet, four simple sentences with subject-verb-object word order were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb.
+The norming materials corresponding to the item depicted in \ref{ex:materials} are shown in \ref{ex:materials_norming}.
 
@@
 The plausibility norming experiment was implemented in PCIbex \citep{pcibex} and was conducted online.
-Participants were asked to judge the plausibility of the sentences on a 7-point scale ranging from absolute implausible (1) to absolute plausible (7).
+Participants were asked to judge the plausibility of the sentences on a 7-point scale ranging from absolutely implausible (1) to absolutely plausible (7).
 All sentences belonging to one item were shown on one screen, so that they could be judged relative to each other.
@@
 In contrast, the inanimate distractor (low semantic interference conditions) was very implausible (mean: 1.7, sd: 1.4).
-The introduction noun which was not manipulated was tested as well and it received a high plausibility rating (mean: 6.0, sd: 1.4).
-These results are illustrated in Figure \ref{fig:plausibility_rating}.
+The first noun in the target sentences, which was not part of the experimental manipulation, was tested as well and it received a high plausibility rating (mean: 6.0, sd: 1.4).
+These results are shown in Figure \ref{fig:plausibility_rating}.
 The plausibility ratings are similar to those reported for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3.

### 2023-10-08 paper/main.tex
changed title

@@
 
-\title{\hspace{1cm}Semantic interference during encoding and retrieval \newline
+\title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?\newline
 %Semantic (but no syntactic) similarity-based interference
-Evidence from self-paced reading and event-related potentials}
+Evidence from self-paced reading and event-related potentials using German}
 \authorsnames{Pia Schoknecht, Shravan Vasishth}
 
-\shorttitle{Semantic interference}
+\shorttitle{Syntactic and semantic interference effects}
 \affiliation{

### 2023-10-08 paper/main.tex
finished up to intro of spr study

@@
 \usepackage{subcaption}
+
+\usepackage{bigdelim}
 
@@
 
-\subsection{The present study}
+\section{The present study}
 
@@
 All experiments used the 2 $\times$ 2 design developed by \citet{vandyke07}; the design crosses syntactic and semantic interference.
-
-% Wrong location for all this detail:
-%This is -- to our knowledge -- the first .
-The SPR experiments were conducted web-based via Prolific and the EEG experiments required participants to come into the lab.
-All experiments were run with native speakers of German.
-Because the high signal-to-noise ratio of EEG data requires a large number of trials per condition and participant, we constructed 120 items which were used in the EEG and SPR experiments.
-It would have been very taxing for the participants to read all 120 items in one experimental session, therefore we decided to split the experiment into two sessions.
-The ERP data that we present here, comes only from participants who completed both sessions.
-In contrast, most of the SPR participants (570 out of 774) finished only one session.
-This was intentional because we found that participants' reading speed in the second session was increased, probably due to increased familiarity with the task.
-Figure \ref{fig:project_str} shows the structure of the present study, i.e., how many participants saw which subset of the items and whether these participants completed one or two experimental sessions.
-
-\begin{figure}[H]
-    \caption{The structure of the present study.
-In total, 120 items were used for the SPR (red branches) and the ERP experiments (blue branches).
-The color brightness reflects how many sessions were completed by each participant (lighter color: two sessions, darker color: one session).}
-    \label{fig:project_str}
-    \centering
-    \includegraphics[width=0.85\textwidth]{images/pandora_project_structure_figure.pdf}
-\end{figure}
 
@@
 
-Comprehension of the sentences in \ref{ex:materials} requires a memory retrieval operation at the verb \textit{trank} (`drank') to retrieve its non-adjacent subject \textit{Witwer} (`widower').
+The comprehension of the sentences in \ref{ex:materials} requires a memory retrieval operation at the verb \textit{trank} (`drank') to retrieve its non-adjacent subject \textit{Witwer} (`widower').
 During this retrieval, the distractor \textit{Einbrecher} (`burglar') / \textit{Verlust} (`loss') might cause interference and impede comprehension.
@@
 
-\paragraph{Participants}
+\paragraph{Participants in the plausibility norming study}
 Forty participants (18 female, 22 male, mean age: 27.3 years old, age range: 21 - 39 years old) from the SPR experiments were re-invited for the plausibility norming one month after they completed the SPR experiment.
-We used the same participants who participated in the SPR study in order to ensure that the ratings for the norming came from the same pool of participants that did the reading study.
-
-\paragraph{Materials}
+We used a subset of the participants who participated in the SPR study in order to ensure that the ratings for the norming came from the same pool of participants that did the reading study.
+
+\paragraph{Materials in the plausibility study}
 For each item quadruplet, four simple sentences with subject-verb-object word order were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb.
@@
 
-\paragraph{Procedure}
+\paragraph{Procedure for the plausibility study}
 The plausibility norming experiment was implemented in PCIbex \citep{pcibex} and was conducted online.
@@
 
-\paragraph{Results}
+\paragraph{Results of the plausibility study}
 The plausibility ratings showed that the semantic manipulation worked as intended.
@@
 
-\section{Experiment 1: SPR}
-\subsection{Methods}
-Initially, we planned to conduct SPR experiment in two sessions to gain data from all participants for all 120 items.
+\subsection{The structure and sample sizes of the SPR and EEG experiments}
+
+The SPR experiments were conducted web-based via Prolific and the EEG experiment was of course run in the lab.
+Because the high signal-to-noise ratio of EEG data requires a large number of trials per condition and participant, we constructed 120 items.
+In order to keep the SPR and EEG studies comparable, we also used 120 items in SPR experiments.
+In both methods, as it would have been very taxing for the participants to read all 120 items in one experimental session, we split the experiments into two sessions.
+
+The SPR study, hereafter Experiment 1a, was initially designed so that participants  (N=160) read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120).
+As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: the effects observed in the first session disappeared.
+The adaptation effect is theoretically interesting per se \parencite{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
+For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=614) were either shown items 1-60 or items 61-120 (but not both).
+In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
+As mentioned above, because the ERP method requires a large number of items, all the data come from participants who completed both sessions.
+
+Figure \ref{fig:project_str} shows the structure of the present study, i.e., how many participants saw which subset of the items and whether these participants completed one or two experimental sessions.
+
+\begin{figure}[H]
+    \caption{The structure of the present study.
+In total, 120 items were used for the SPR (red branches) and the ERP experiments (blue branches).
+The color brightness reflects how many sessions were completed by each participant (lighter color: two sessions, darker color: one session).}
+    \label{fig:project_str}
+    \centering
+    \includegraphics[width=0.85\textwidth]{images/pandora_project_structure_figure.pdf}
+\end{figure}
+
+\section{Experiments 1a and 1b: SPR}
+%\subsection{Methods}
+
+%Initially, we planned to conduct SPR experiment in two sessions to gain data from all participants for all 120 items.
 Collecting data for 120 items in one session was not considered feasible in a web-based SPR experiment, as the experiment duration would have been close to one hour.
@@
 
-Participants for the experiment were recruited via Prolific.
-In total, 908 participants took part in the self-paced reading experiment.
+Native German-speaking participants were recruited via Prolific.
+In total, 908 participants took part in the self-paced reading experiments E1a and E1b.
 117 participants were excluded due to low comprehension question accuracy (< 70\,\%).
@@
 334 participants reported to have a high-school diploma and 70 another secondary school certificate.
-Five participants replied with ``other.'' We conclude that the participants who were recruited via Prolific were slightly older and have a comparable or higher level of education than the typical undergraduate and graduate students who participate in lab-based studies.
-
-A subset of the SPR experiment was carried out in two experimental sessions per participant.
-Drop-out between sessions led to a lower sample size in the second session (204 vs.\ 160).
-In this multiple-sessions part of the experiment, compensation for the first session was \pounds 3.5 and for the second session, it was \pounds 5.
-In the single-session part of the experiment, compensation was \pounds 4.
+Five participants replied with ``other.''
+%We conclude that the participants who were recruited via Prolific were slightly older and have a comparable or higher level of education than the typical undergraduate and graduate students who participate in lab-based studies.
+
+In Experiment 1a, drop-outs between sessions led to a lower sample size in the second session (204 vs.\ 160).
+In Experiment 1a, compensation for the first session was \pounds 3.5 and for the second session, it was \pounds 5.
+In Experiment 1b, compensation was \pounds 4.
 
@@
 
-In the multiple-sessions part of the experiment, participants were re-invited for session 2 after they completed session 1.
+In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
 The experimental sessions were separated by 1 - 20 days.
@@
 \subsubsection{Statistical analyses}
+
+\paragraph{Bayesian linear mixed models}
 Comprehension question accuracy was analyzed with Bayesian generalized linear mixed models, i.e., logistic regression, in R \citep{r}, using the brms package \citep{brms}.
@@
 These were sum-contrast coded (high +0.5, low -0.5).
-Random intercepts for participants and items were included.
-For the intercept, we used a Normal(0, 1.5) prior and  for all slope parameters, a Normal(0, 0.1) prior was used.
-No other priors were specified.
+Varying intercepts for participants and items were included.
+For the intercept, we used a Normal(0, 1.5) prior and  for all fixed-effect slope parameters, a Normal(0, 0.1) prior.
+The priors for the variance components were the defaults specified in brms.
 Models were run with 4 chains and 8,000 iterations in each chain.
@@
 The critical word for analyses was the verb \textit{trank} `drank' in \ref{ex:materials}, constituting the retrieval site.
-Additionally, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed.
+Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed.
 The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
-Factors were sum-contrast coded \citep{schad_et_al_contrasts}.
-Random intercepts for participants and items were included.
-Inclusion of any random slopes led to unstable Bayes factors due to increased model complexity.
-We used informative priors for all parameters of the models (see Table \ref{tab:spr_priors}).
-We used a wide range of priors on the parameters of interest, assuming effect sizes ranging from 8 ms up to 81 ms, to carry out a Bayes factors sensitivity analysis.
-If BF$_{10}$ > 1, it provides evidence in favor of the effect.
+Varying intercepts for participants and items were included; varying slopes were excluded for reasons explained below.
+
+\paragraph{Bayes factors for model comparison}
+Model comparisons were carried out using Bayes factors. Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF}, 
+we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}).
+The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms.
+These priors are based on the observed effect sizes and their uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}.
+
+The Bayes factor is analogous to the frequentist ANOVA or likelihood ratio test: it compares the (marginal) likelihood of a full model to that of a baseline null model; this is often written as $BF_{10}$, with the subscripts indicating that marginal likelihood of the full model is in the numerator and that of the null model is in the denominator.
+One major advantage of the Bayes factor over the frequentist analogue is that it takes the uncertainty of the parameters into account; this leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
+Another important advantage of the Bayes factor---one that is very relevant for the present work---is that it is in principle possible to find evidence both for or against an effect; this is contrast to the frequentist ANOVA, which is designed to only furnish evidence against the null.
+
+Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest.
 If BF$_{10}$ < 1, it provides evidence in favor of the null model.
 The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence for the null.
-Models were run with 4 chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being the warm-up phase \citep{schad_etal_2022_BF}.
+In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
+For this reason, models were run with 4 chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.
+Bayes factors were unstable when varying slopes were included; for this reason, we fit the models only with varying intercepts by participant and by item.
 
 \begin{table}[h]
-    \caption{Informative priors for the analysis of the self-paced reading time data.
-Priors on the slope parameters were truncated to a lower bound of zero, i.e., only positive values.
-This corresponds to the assumption that high interference conditions lead to longer reading times than low interference conditions.
-The standard deviation of the slope priors was varied in order to conduct a Bayes factor sensitivity analysis following \citet{schad_etal_2022_BF}.
+    \caption{Priors used for the analysis of the self-paced reading time data.
+Priors on the fixed-effects slope parameters were truncated to a lower bound of zero, i.e., only positive values were assumed to be possible a priori.
+This corresponds to the existing evidence in the literature that high interference conditions lead to longer reading times than low interference conditions.
+The standard deviation of the fixed-effects slope priors was varied in order to conduct a Bayes factor sensitivity analysis following \citet{schad_etal_2022_BF}.
 See the corresponding assumed a priori range of the difference between reading times under high vs.\ low interference on the millisecond scale.}
@@
     \toprule
-    Parameter&Prior & &Assumed Range\\
+    Parameter&Prior & &Assumed Range in ms\\
     \midrule
-  Intercept & Normal(6, 0.6)& &\\
+  Intercept & Normal(6, 0.6)& & 125-1308\\
   \cmidrule{2-4}
-  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \multirow{3}{1cm}{lb=0} & 0 - 8 ms\\
-  &  Normal(0, 0.05)& &0 - 40 ms\\
-  & Normal(0, 0.1) & &0 - 81 ms\\
+  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \hspace{-1em}\rdelim\}{3}{*}[lower bound: 0] & 0 - 8\\
+  &  Normal(0, 0.05)& &0 - 40\\
+  & Normal(0, 0.1) & &0 - 81\\
   \cmidrule{2-4}
-  sigma & Normal(0, 0.5)&&\\
+  sigma & Normal(0, 0.5)&  \hspace{-1em}\rdelim\}{2}{*}[lower bound: 0]&\\
   SD & Normal(0, 0.1)& &\\
@@
   \end{table}
+
+ %\multirow{3}{1cm}{lb=0}
 
@@
 %\end{table}
-
-\subsection{Results}
-Data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.
 
@@
 
+\section{Data availability statement}
+
+The data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.
+
 \newpage

### 2023-10-08 paper/main.tex
minor edits

@@
 \paragraph{Bayes factors for model comparison}
-Model comparisons were carried out using Bayes factors. Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF}, 
+
+In order to quantify the uncertainty on the parameters of interest, we report 95\% credible intervals.  
+These intervals represent the range over which we can be 95\% certain that the values of the parameter lies, given the statistical model and the data.
+However, formal hypothesis testing cannot be carried out without a likelihood ratio test that compares two alternative models \parencite{schad_etal_2022_BF,Royall}.
+For this reason, formal hypothesis tests for the presence or absence of the effects of interest were carried out using Bayes factors.
+Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF},
 we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}).
@@
 In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
-For this reason, models were run with 4 chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.
+For this reason, models were run with four chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.
 Bayes factors were unstable when varying slopes were included; for this reason, we fit the models only with varying intercepts by participant and by item.
@@
 
- %\multirow{3}{1cm}{lb=0}
-
 %\bigskip
@@
 
-\subsubsection{Comprehension question accuracy}
+\subsubsection{Comprehension question accuracy (Experiments 1a and 1b combined)}
+
 After exclusion of participants with accuracy below 70\,\%, overall accuracy (including fillers) was good; on average 85.9\,\% (range: 71.9 - 100\,\%).
-Accuracy in critical trials was 81.1\,\% (range: 20 - 100\,\%).
+Accuracy for the critical items was 81.1\,\% (range: 20 - 100\,\%).
 Condition-wise accuracy is shown in Table \ref{tab:spr_acc}.
@@
 \begin{table}[]
-    \caption{By-condition accuracy in critical trials in the SPR experiment.}
+    \caption{By-condition accuracy in critical trials in the SPR experiment (E1a and E1b combined).}
     \label{tab:spr_acc}
@@
 
-Table \ref{tab:spr_acc_mod} shows the results of the generalized mixed model analyzing the comprehension accuracy in critical trials.
-The estimates and 95\,\% credible intervals in log-odds show primarily a reduction in comprehension accuracy for high compared to low semantic interference conditions.
+Table \ref{tab:spr_acc_mod} shows the results of the generalized mixed model (log-odds scale) analyzing the comprehension accuracy for the critical items (Experiments 1a and 1b combined).
+The estimates and 95\,\% credible intervals show primarily a reduction in comprehension accuracy for high compared to low semantic interference conditions.
 
 \begin{table}[H]
-    \caption{Results in log-odds from the Bayesian generalized model analyzing the comprehension accuracy in critical trials of the SPR experiment.}
+    \caption{Results in log-odds from the Bayesian generalized model analyzing the comprehension accuracy in the SPR Experiments 1a and 1b.}
     \label{tab:spr_acc_mod}
     \centering
-    \begin{tabular}{lrrr}
+    \begin{tabular}{lrr}
     \toprule
-    & Estimate & Est.Error & 95\% CrI  \\
+    & Estimate &  95\% CrI  \\
     \midrule
-Intercept& 1.66 & 0.14&  1.38 - 1.93\\
-syntactic& -0.08 & 0.04& -0.16 - -0.01\\
-semantic&  -0.61 &0.04& -0.69 - -0.54\\
-interaction& -0.10& 0.06& -0.22 - 0.02\\
+Intercept& 1.66 &   [1.38, 1.93]\\
+syntactic& -0.08 &  [-0.16, -0.01]\\
+semantic&  -0.61 & [-0.69, -0.54]\\
+interaction& -0.10&  [-0.22, 0.02]\\
     \bottomrule
@@
 
-\subsubsection{Self-paced reading times}
+Since the accuracies are not of primary interest, we do not carry out Bayes factors analyses for these.
+
+\subsubsection{Self-paced reading times (Experiments 1a and 1b combined)}
 Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structure.
@@
 
-It is apparent from Figure \ref{fig:whole_sentence} A and B that, at the distractor, regardless of the syntactic manipulation, reading times between high and low semantic interference started to differ.
-Distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
+It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation  at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
 This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions.

### 2023-10-09 paper/main.tex
edited spr study

@@
 It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the pre-critical and critical regions.
-The spillover of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power.
-In order to attenuate spillover from the clause-final word in the embedded clause preceding the critical region, we inserted two adverbs.
+The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power.
+In order to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region, we inserted two adverbs.
 This way, the critical region will be less affected by the clause-boundary slowdown.
-This design also has the advantage that any potential effects seen at the pre-critical region (the adverb preceding the verb) are less likely to be due to spillover; it is important to look at the pre-critical region as previous work with this design found effects at the pre-critical region (we revisit this detail in the General discussion).
+This design also has the advantage that any potential effects seen at the pre-critical region (the adverb preceding the verb) are less likely to be due to spill-over; it is important to look at the pre-critical region as previous work with this design found effects at the pre-critical region (we revisit this detail in the General discussion).
 
@@
 \subsubsection{Self-paced reading times (Experiments 1a and 1b combined)}
-Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structure.
-%\clearpage
+
+\paragraph{Estimated effect sizes and their uncertainty}
+
+Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structures.
 
@@
 
-It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation  at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
-This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions.
-Nonetheless, we refrained from analyzing the distractor and the immediately following regions, because they differed between conditions.
-Instead we focused our analyses on regions which were identical across conditions: The pre-critical, critical and spill-over region.
-Figure \ref{fig:whole_sentence} C shows the pooled reading times of these regions (and additionally the pre-pre-critical region which was identical across conditions as well, but might be confounded by spill-over from the syntactic manipulation).
+It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
+This pattern persisted in the following regions.
+
+Because the distractor and the immediately following regions differ by condition, we did not compute Bayes factors for these differences.
+Instead we focused our Bayes factors analyses on regions which were identical across conditions.
+These are the pre-critical, critical and spill-over regions.
+
+%This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions.
+
+Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but will be harder to interpret because of possible spill-over from the syntactic manipulation.
 The pre-pre-critical and pre-critical region show longer reading times for high vs.\ low semantic interference conditions.
@@
 
-Panel D shows the reading times of the critical and surrounding regions for each of the sub-parts of the SPR experiment.
-The sub-experiment with multiple sessions and items 1-60 (N = 204, left most sub-figure) showed results similar to the pooled results presented in Panel C.
-The sub-experiment with multiple sessions and items 61-120 (N = 160) showed a semantic interference effect at the critical verb and less clear results in the surrounding regions.
-Additionally, this sub-experiment showed shorter reading times compared to the other sub-experiments in general.
-This was probably caused by adaptation to the task, because the same participants had completed another session earlier.
-The sub-experiment with a single session and items 1-60 (N = 350) showed a semantic interference effect in all regions.
-The sub-experiment with a single session and items 61-120 (N = 220, right most sub-figure) showed a semantic interference effect in the regions surrounding the critical verb, but not at the verb itself.
+Panel D shows the reading times of the critical and surrounding regions for each of the two Experimenta 1a and 1b separately.
+In Experiment 1a, the first session (items 1-60, N = 204, left most sub-figure) showed results similar to the pooled results presented in Panel C.
+In the second session of Experiment 1a (items 61-120, N = 160) a smaller semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session.
+Additionally, the second session in Experiment 1b showed shorter reading times compared to the first session.
+As mentioned earlier, this attenuation in reading times and differences between conditions is likely due to  adaptation to the task in the second session.
+
+Panel D also shows that Experiment 1b with items 1-60 (N = 350) showed a pattern consistent with   semantic interference in all regions.
+Experiment 1b with items 61-120 (N = 220, right most sub-figure) showed a pattern consistent with semantic interference in the regions surrounding the critical verb, but not at the verb itself.
 At the verb, reading times were longest when both syntactic and semantic interference were high.
 Reading times were a bit faster when both syntactic and semantic interference were low.
-The other two conditions (high syntactic/low semantic and low syntactic/high semantic) lead to the fastest verb reading times in this sub-experiment.
-These somewhat inconsistent results, especially in the critical region, across the sub-experiments and the random assignment of participants to the sub-experiments motivated us to gain estimates of the effects under investigation for the individual participants (see Figure \ref{fig:spr_individual}).
-
-\begin{figure}[H]
-    \caption{Individual estimates of syntactic and semantic interference and their interaction at the critical verb for all 774 participants in the self-paced reading experiment.
+The other two conditions (high syntactic/low semantic and low syntactic/high semantic) lead to the fastest verb reading times in Experiment 1b with items 61-120.
+
+%These somewhat inconsistent results, especially in the critical region, across the sub-experiments and the random assignment of participants to the sub-experiments motivated us to gain estimates of the effects under investigation for the individual participants (see Figure \ref{fig:spr_individual}).
+
+%\begin{figure}[H]
+ %   \caption{Individual estimates of syntactic and semantic interference and their interaction at the critical verb for all 774 participants in the self-paced reading experiment.
 Shown are shrunken estimates from a Bayesian hierarchical linear mixed model.
@@
 Two solid red lines mark an effect size of 10 and 20 ms, respectively.}
-    \label{fig:spr_individual}
-    \centering
-    \includegraphics[width=\textwidth]{images/individual_diffs.jpg}
-\end{figure}
-
-The individual-level effects were derived from a Bayesian linear mixed model with the priors shown in Table \ref{tab:spr_priors} with the intermediate slope prior of Normal(0, 0.05) and random slopes for syntactic and semantic interference and their interaction but without correlations of random effects.
-
-Figure \ref{fig:spr_individual} shows that at the critical verb, the 95\,\% credible intervals for the syntactic interference effect of all participants were centered around zero.
+%    \label{fig:spr_individual}
+ %   \centering
+  %  \includegraphics[width=\textwidth]{images/individual_diffs.jpg}
+%\end{figure}
+
+%The individual-level effects were derived from a Bayesian linear mixed model with the priors shown in Table \ref{tab:spr_priors} with the intermediate slope prior of Normal(0, 0.05) and random slopes for syntactic and semantic interference and their interaction but without correlations of random effects.
+
+%Figure \ref{fig:spr_individual} shows that at the critical verb, the 95\,\% credible intervals for the syntactic interference effect of all participants were centered around zero.
 In contrast, participants consistently displayed a small effect of semantic interference with narrow 95\,\% credible intervals.
@@
 
-The posteriors of the pre-critical, critical and spill-over regions were very similar (see Figure \ref{fig:spr_posteriors}).
-For the syntactic interference effect, they were close to zero in all regions.
-In contrast, for the semantic interference effect, the estimates ranged from 4 ms (lower bound in the critical region) to 14 ms (upper bound in the pre-critical region).
-The posteriors for the interaction were consistent with an interaction effect between 1 ms and 13 ms in the pre-critical and spill-over region.
-They were largest in the critical region (95\,\% CrI [7, 22 ms]).
+Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
+For syntactic interference, they were close to zero in all regions.
+In contrast, for semantic interference, the estimates ranged between [8,14] ms (pre-critical), [4, 12] ms (critical), and [5, 11] ms (spill-over region).
+The posteriors for the interaction were consistent with an interaction effect.
+The interaction was the  largest in the critical region ([7, 22 ms]).
 
 \begin{figure}[H]
-    \caption{Posteriors for the syntactic and semantic interference effects and their interaction at the critical verb and surrounding regions in the self-paced reading experiment.
-Numbers are the means and 95\,\% credible intervals.
-Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
+    \caption{Posteriors for the syntactic and semantic interference effects and their interaction at the critical verb and surrounding regions in the pooled self-paced reading data (Experiments 1a and 1b combined).
+The numerical values are the means and 95\,\% credible intervals.
+The blue vertical lines represent the median and the blue shaded areas are 80\,\% credible intervals.}
     \label{fig:spr_posteriors}
@@
 
-Bayes factors, the Bayesian version of hypothesis testing, were used to evaluate whether there was evidence for the effects estimated above while taking into account the uncertainty \citep{schad_etal_2022_BF}.
+\paragraph{Hypothesis testing using Bayes factors}
+
+Was there any evidence for the effects estimated above.
 The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}.
-Bayes factors provided extremely strong evidence for the semantic interference effect in the pre-critical, critical and spill-over region (BF$_{10}$ > 200).
-Bayes factors showed no effect of syntactic interference in any of the investigated regions.
-In fact, the Bayes factor analysis actually provided evidence against the syntactic interference effect in all regions and under all priors (BF$_{10}$ < 1).
-Evidence for the interaction was only found in the critical region, where it ranged from strong (BF$_{10}$ = 40) to weak (BF$_{10}$ = 2) depending on the a priori assumed effect size.
+The analyses use all the data from Experiments 1a and 1b.
+We found extremely strong evidence for the semantic interference effect in the pre-critical, critical and spill-over regions (BF$_{10}$ > 200).
+No evidence was found for syntactic interference in any of the investigated regions; in fact, there is strong evidence against the syntactic interference effect in all regions and under all the priors used (BF$_{10}$ < 1).
+Evidence for the interaction was only found in the critical region.
+Here, the evidence for the interaction was strong (BF$_{10}$ = 40) when the prior on the parameter was N+(0, 0.01), i.e., a small a priori effect size ranging from  [0, 8] ms; the evidence for the interaction was much weaker if the prior is N+(0,0.1), i.e., a larger a priori effect size ranging from [0, 81] ms.
+In other words, only if we assume a priori that the effect size for the interaction is relatively small, then there is strong evidence for this effect.
 
 \begin{figure}[H]
-    \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times at the critical verb and surrounding regions.}
+    \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
     \label{fig:spr_bfs}
@@
 We presented a large-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-Comprehension accuracy and reading times of the critical verb showed semantic interference, but no syntactic interference.
-Importantly, in the reading times, the effect did not emerge at the critical verb, i.e., the retrieval site during subject-verb dependency formation, but actually started when the distractor was read.
-
-\subsection{Pre-critical effects}
-Animate distractors (high semantic interference) were read slower than inanimate distractors (low semantic interference, see Figure \ref{fig:whole_sentence} A and B).
-One might argue that this could be due to differences in word length.
+Reading times at the critical verb showed clear evidence for semantic interference, but evidence against syntactic interference.
+The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occured even before the retrieval event could begin.
+
+\subsection{The effects in the pre-critical region}
+
+At the distractor (which occurred well before the pre-critical region), animate distractors (high semantic interference) were read slower than inanimate distractors (low semantic interference, see Figure \ref{fig:whole_sentence} A and B).
+One might argue that this pattern could be due to differences in word length.
 The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9).
-It seems unlikely that this small difference in word length caused such a large and long-lasting effect.
-
-We assume that the semantic interference effect starting at the distractor and persisting into later regions was caused by encoding interference \citep{Oberauer_Kliegl_2006}.
-The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) caused the comprehender to slow down \citep[for similar findings, see e.g., ][]{lago_etal_2021, kush_etal_2015, gordon02}.
-
-The pre-critical effects in the previous studies \citep{vandyke07, mertzen} might have also started at the distractor.
+It seems unlikely that this small difference in word length caused such a large and long-lasting effects, but in order to be conservative we do not try to interpret this difference.
+
+At the pre-critical region, there was clear evidence for semantic interference.
+As this cannot be due to retrieval interfernce, the most likely explanation is encoding interference \citep{Oberauer_Kliegl_2006}.
+The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, kush_etal_2015, gordon02}.
+
+Indeed, previous work using the same design as in the present paper has also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
 \citeauthor{mertzen}'s (\citeyear{mertzen}) Figures 5 and 6 indicated reading time differences in earlier regions, especially for their German data (see their Figure 6).
-\citet{vandyke07} did not report reading times for the whole sentences / distractors.
-While \citet{vandyke07} assumed that the pre-critical effects were due to plausibility differences between conditions, \citet{mertzen} also discussed encoding interference as the cause of pre-critical effects in their study.
-Our results strongly support the encoding interference explanation because the effects started at the distractor.
-
-Alternative causes of the pre-critical effects discussed by \citet{mertzen} are parafoveal-on-foveal effects, sentence structure confounds and prediction.
-Our study rules out the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
+\cite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \cite{mertzen} also points out, encoding interference could be an explanation even in that study.
+Given these earlier findings, our results are consistent with the encoding interference explanation.
+
+\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. 
+They discuss several alternative explanations for their observed effects.
+In addition to encoding interference, \cite{mertzen} propose three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
+Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
 The observed difference between the high and low semantic interference conditions is independent of the syntactic manipulation.
-Therefore, the sentence structure confound is not a good explanation for the effect in our study.
-\citet{mertzen} argued that the pre-critical adverb is the first word of the verbal phrase and therefore, might trigger the retrieval of the subject.
-In our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence} A and B).
-Consequently, the prediction explanation is also ruled out by our findings, rendering the encoding interference explanation the most likely explanation of the effects found prior to the critical verb in our and previous studies.
+Therefore, the sentence structure confound is also not a good explanation for the effect in our study.
+Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming the verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region.
+However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence} A and B).
+The predictive-processing explanation would incorrectly predict that the effect begins at the second adverb, which immediately precedes the verb.
+Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference as the most likely explanation of the effects found prior to the critical verb in our data.
 Crucially this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in the critical region as well.
 The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence.
-It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region or still due to encoding and maintaining the distractor in memory.
+It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory, or a combination of the two kinds of interference.
 
@@
 Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design.
-Consequently, the estimates from the present study are the most precise estimates for this design.
-The small effect size in the present study is comparable to that of other large-scale studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
+Consequently, the estimates from the present study are the most precise ones so far for this design.
+The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
 
 An obvious difference from previous studies is the experimental methodology.
-The majority of the previous experiments employed eye-tracking while reading methodology, while we used a self-paced reading task.
-The exact relation of reading times obtained by these two methods is unclear.
-Self-paced reading could facilitate interference effects compared to natural reading (as in eye-tracking studies) due to the restricted reading format.
-Comprehenders have no opportunity to revisit previous material during conventional self-paced reading  \citep[but see][]{BSPR}.
-This is more demanding on the comprehender's memory which might enhance interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous}).
-The estimates of both main effects (syntactic and semantic interference) and the interaction in the spill-over region are very similar to the results of the present study in the critical region.
-This similarity between the two self-paced reading experiments further validates the findings of the present study.
-But it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.
-
-The contrast between our results and the results from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
-So, why did we find a semantic interference effect and \citet{mertzen} did not?
-This could be due to lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
+The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task.
+Self-paced reading (SPR) could trigger a greater tendency to cause interference effects compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading  \citep[but see][]{BSPR}.
+SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
+This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous} in the appendix).
+In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by  \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
+However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.
+
+The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
+So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
+One possible explanation is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
 While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
-This is also supported by the reduced uncertainty of the individual-level estimates in our data compared to \citeauthor{mertzen}'s (\citeyear{mertzen}).
-We will discuss the lack of syntactic interference in our data in the general discussion.
+%This is also supported by the reduced uncertainty of the individual-level estimates in our data compared to \citeauthor{mertzen}'s (\citeyear{mertzen}). 
+We will discuss the absence of syntactic interference in our data in the general discussion.
 

### 2023-10-09 code_data/LV05 modeling/interACT_subject.R
Added via the Open Science Framework

@@ -1,0 +1,340 @@
+library(tidyr)
+library(dplyr)
+
+###############################################################
+## Global parameters
+###############################################################
+reset_params <- function(){
+  lf        <<- 0.15       # latency factor
+  le        <<- 1       # latency exponent
+  rth 	    <<- -1.5    # retrieval threshold
+  bll       <<- 0.5     # decay parameter
+  ans       <<- 0.2    # activation noise
+  mas       <<- 1       # maximum associative strength 
+  mp        <<- 1       # mismatch penalty
+  ga        <<- 1       # goal source activation
+  rand_time <<- 3       # latency variability
+  blc       <<- 0       # base-level activation
+  ##
+  lp        <<- 1     # default time since last presentation (msec)
+  ldp       <<- 1       # last distractor presentation (msec)
+  ## Distractor control:
+  ndistr    <<- 1				# number of distractors
+  dbl       <<- 0       # distractor base-level
+  ##
+  ## Cue weighting, cue confusion and prominence:
+  # cueweights <<- c(1,1) # vector of cue weightings
+  cueweighting <<- 1     # Strength of structural cue as ratio structural/morphological
+  normalizeWeights <<- TRUE 
+  qcf 		  <<- 1 			# match quality correction factor
+  qco 		  <<- -2*rth 			# match quality correction offset
+  psc       <<- 1       # prominence scaling constant C1
+  pic       <<- 0       # prominence scaling constant C2
+  tprom     <<- 0       # target prominence 
+  dprom     <<- 0       # distractor prominence
+  # cl 				<<- 0 			# cue confusion level (0-100)
+  cuesim    <<- -1    # cue-feature similarity [-1..0]
+  #
+  # Fitted meta-parameters
+	meta_recent <<- 0.7
+	meta_distant <<- 1.3
+	meta_lowprom <<- -0.5
+	meta_medprom <<- 0
+	meta_highprom <<- 2.5
+	meta_conflevel <<- 0
[…]

### 2023-10-09 code_data/LV05 modeling/Generate_predictions_LV05Engelmann.R
Added via the Open Science Framework

@@ -1,0 +1,227 @@
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(parallel)
+library(LaplacesDemon)
+library(mvtnorm)
+library(truncnorm)
+library(reshape2)
+
+
+## some helper functions:
+rmsd <- function (obs, pred) {
+  sqrt(mean((obs - pred)^2, na.rm = TRUE))
+}
+
+compute_int_means <- function(d){
+  cond_means <- as.data.frame(d) %>% group_by(Set, Target, Distractor, lf, ans, mas, mp, rth, bll, psc, pic, qcf, qco, cuesim, tprom, dprom, lp, ldp, blc, dbl, ndistr, cueweighting) %>% summarise(meanRT=mean(latency), SE=sd(latency)/sqrt(length(latency)))
+  int <- reshape(as.data.frame(d[,c(2,5,40)]),direction="wide",idvar = "Iteration",timevar = "Distractor")
+  colnames(int)[2:5] <- gsub("latency.","",colnames(int)[2:5])
+  int <- int %>% mutate(SynInt=(SynMatch+FullMatch)-(NoMatch+SemMatch),SemInt=(SemMatch+FullMatch)-(NoMatch+SynMatch),SynSemInt=FullMatch-SynMatch,NoSynSemInt=SemMatch-NoMatch)
+  Effect <- unname(colMeans(int))[-1]
+  return(Effect)
+}
+
+convert2log <- function(x){
+  ifelse(x>=1, log(x), ifelse(x<=-1, -log(abs(x)), 0))
+}
+
+convert2log10 <- function(x){
+  x <- ifelse(x>-1 & x<1, 0, x)
+  x <- ifelse(x<=-1, -log10(abs(x)), x) 
+  x <- ifelse(x>=1, log10(abs(x)), x)
+}
+
+
+# Introduction 
+
+## Load interact subject VERSION (syntactic cue = +-subject position)
+
+#The following piece of code, written by Felix Engelmann and available on github, provides the main computation code for ACT-R calculations.
+
+source("interACT_subject.R")
+
+## Basic engine for generating predictions
+
+printcounts<-FALSE
+
+iterate_lf <- function(values,iterations=1000){
+  ## values is a dataframe containing either 1 row of parameters 
+  # (a combination of latency factor and cue-weighting parameter, or multiple rows i.e., multiple combinations of two parameters)
+  ## iterations is the number of iterations for that given value.
[…]

### 2023-10-10 paper/main.tex
minor edit

@@
 
-It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the pre-critical and critical regions.
+\paragraph{Regions of interest in the design}
+
+Although our principal focus was on the reading times at the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region.
+A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would cause two verbs to appear one after another ( \textit{erzählt hatte, trank}, literally `told drank').
+This is because in German, the verb is clause-final in embedded clauses.
+Such verb sequences are inherently difficult to process \parencite{VSLK11,bach1986crossed}.
+T
+
+For these reasons, we added two adverbs before the critical verb. These intervening regions had a further advantage: 
+It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region.
 The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power.
-In order to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region, we inserted two adverbs.
+Inserting two adverbs before the critical verb thus served to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region.
 This way, the critical region will be less affected by the clause-boundary slowdown.
-This design also has the advantage that any potential effects seen at the pre-critical region (the adverb preceding the verb) are less likely to be due to spill-over; it is important to look at the pre-critical region as previous work with this design found effects at the pre-critical region (we revisit this detail in the General discussion).
+This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.
+
+Thus, our focus in the present study is on the processing difficulty observed at the pre-critical, critical (verb), and post-critical regions.
 
@@
 The SPR study, hereafter Experiment 1a, was initially designed so that participants  (N=160) read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120).
+The variation between participants in the gap is due to the fact that we could not control when the participants would do the second session; however, the mean and median gaps were  about 11 days, with the first and third quartiles being 9 and 12 days.
+
 As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: the effects observed in the first session disappeared.

### 2023-10-10 code_data/plausibility_norming/norming_analysis.R
Added via the Open Science Framework

@@ -1,0 +1,56 @@
+# Plausibility Norming: PANDORA
+# author: Pia Schoknecht
+# date: 17 06 2022
+
+library(plyr)
+library(ggplot2)
+library(dplyr)
+library(tidyverse)
+library(lme4)
+library(ordinal)
+library(brms)
+
+
+# load data
+dfp <- read.csv("pandora_plausibility_rating_corrected.csv")
+
+# number of participants
+(N_subj <- length(unique(dfp$subj)))
+
+uniq <- dfp[!duplicated(dfp[1]),]
+
+#age
+(age <- table(uniq$age))
+(mean(uniq$age))
+
+#sex
+(sex <- table(uniq$sex))
+
+
+
+# Pandora Ratings
+
+# number of items
+uniq <- dfp[!duplicated(dfp[1]),]
+(N_item <- length(unique(dfp$item)))
+
+(dfp2 <- dfp %>%
+    group_by(cond) %>%
+    dplyr::summarize(mean_rating = mean(rating, na.rm=TRUE), sd_rating = sd(rating, na.rm=TRUE)))
+
+dfp$cond <- factor(dfp$cond, levels = c("intro", "subject", "distr_anim", "distr_inan"))
+
+ggplot(dfp, aes(x=cond, y=rating))+
+  geom_jitter(color="lightblue", alpha=0.2)+
+  geom_boxplot(fill="grey", alpha=0.7)+
+  stat_summary(fun=mean, geom="point", shape=18, size=4) + #adding the mean
+  expand_limits(y=c(0.5,7.5))+
+  scale_y_continuous(name="plausibility rating", breaks=seq(1, 7, 1))+
+  scale_x_discrete(name="", labels=c("distr_anim" = "animate \ndistractor", "distr_inan" = "inanimate \ndistractor",
+                                     "intro" = "introduction", "subject" = "subject"))+
+  theme_bw()+
+  labs(title = "Plausibility ratings",
+       subtitle = "1 = absolute implausible; 7 = absolute plausible")
+
+ggsave("plots/pandora_plausibility_ratings.jpg", dpi=600, width=4, height=4)
+

### 2023-10-10 code_data/plausibility_norming/norming_analysis.R
Updated via the Open Science Framework

@@ -42,7 +42,7 @@ dfp$cond <- factor(dfp$cond, levels = c("intro", "subject", "distr_anim", "distr
 
 ggplot(dfp, aes(x=cond, y=rating))+
   geom_jitter(color="lightblue", alpha=0.2)+
-  geom_boxplot(fill="grey", alpha=0.7)+
+  geom_boxplot(fill="grey", alpha=0.7, outlier.shape = NA)+
   stat_summary(fun=mean, geom="point", shape=18, size=4) + #adding the mean
   expand_limits(y=c(0.5,7.5))+
   scale_y_continuous(name="plausibility rating", breaks=seq(1, 7, 1))+

### 2023-10-10 paper/main.tex
edited up to EEG study

@@
 
-Although our principal focus was on the reading times at the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region.
-A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would cause two verbs to appear one after another ( \textit{erzählt hatte, trank}, literally `told drank').
+Although our principal focus was on the reading times at the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region (we return to this point in the discussion sections).
+A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another ( \textit{erzählt hatte, trank}, literally `told, drank').
 This is because in German, the verb is clause-final in embedded clauses.
 Such verb sequences are inherently difficult to process \parencite{VSLK11,bach1986crossed}.
-T
-
-For these reasons, we added two adverbs before the critical verb. These intervening regions had a further advantage: 
+
+For these reasons, we added two adverbs before the critical verb. Including these intervening regions had a further advantage: 
 It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region.
@@
 Inserting two adverbs before the critical verb thus served to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region.
-This way, the critical region will be less affected by the clause-boundary slowdown.
+This way, the critical region should be less affected by the clause-boundary slowdown.
 This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.
@@
 \paragraph{Materials in the plausibility study}
-For each item quadruplet, four simple sentences with subject-verb-object word order were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb.
+For each of the 120 item quadruplets, four simple sentences with subject-verb-object word order were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb.
 The norming materials corresponding to the item depicted in \ref{ex:materials} are shown in \ref{ex:materials_norming}.
@@
 
+Thus, each of the forty participants were asked to rate $120 \times 4$ items, leading to 19,200 data points.
+
 \paragraph{Procedure for the plausibility study}
@@
   \cmidrule{2-4}
-  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \hspace{-1em}\rdelim\}{3}{*}[lower bound: 0] & 0 - 8\\
-  &  Normal(0, 0.05)& &0 - 40\\
-  & Normal(0, 0.1) & &0 - 81\\
+  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \hspace{-1em}\rdelim\}{3}{*}[lower bound: 0] & [0, 8]\\
+  &  Normal(0, 0.05)& & [0, 40]\\
+  & Normal(0, 0.1) & &[ 0, 81]\\
   \cmidrule{2-4}
@@
 \section{Discussion}
-We presented a large-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
+We presented a relatively large-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
 Reading times at the critical verb showed clear evidence for semantic interference, but evidence against syntactic interference.
@@
 %This is also supported by the reduced uncertainty of the individual-level estimates in our data compared to \citeauthor{mertzen}'s (\citeyear{mertzen}). 
-We will discuss the absence of syntactic interference in our data in the general discussion.
+We will address the absence of syntactic interference in our data in the general discussion.
 
@@
 This functioned as a regression-based instead of traditional baseline correction \citep{alday2019}.
-Random intercepts for participants and items as well as by-participant random slopes for semantic interference were included.
+Varying intercepts for participants and items as well as by-participant random slopes for semantic interference were included.
 Correlations between varying intercepts and slopes were not included.
-We used informative priors for all parameters of the models (see Table \ref{tab:eeg_priors}).
+We used relatively informative priors for all parameters of the models (see Table \ref{tab:eeg_priors}).
 
 \begin{table}[]
-    \caption{Informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}.
+    \caption{Relatively informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}.
 Priors on the slope parameters were truncated to an upper bound of zero, i.e., only negative values.
@@
     \toprule
-    Parameter&Prior & &Assumed Range\\
+    Parameter&Prior & &Assumed Range ($\mu V$)\\
     \midrule
@@
   \cmidrule{2-4}
-  \multirow{4}{1cm}{beta} & Normal(0, 0.1) & \multirow{4}{1cm}{ub=0} & -0.2 - 0\,$\mu$V\\
-  &  Normal(0, 0.5)& & -1 - 0\,$\mu$V\\
-  & Normal(0, 1) && -2 - 0\,$\mu$V\\
-  & Normal(0, 5) && -10 - 0\,$\mu$V\\
+  \multirow{4}{1cm}{beta} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{3}{*}[upper bound: 0]& [-0.2, - 0]\\
+  &  Normal(0, 0.5)& & [-1, 0]\\
+  & Normal(0, 1) && [-2, 0]\\
+  & Normal(0, 5) && [-11, 0]\\
   \cmidrule{2-4}
-  sigma & Normal(10, 5)&\\
+  sigma & Normal(10, 5)& \hspace{-1em}\rdelim\}{2}{*}[lower bound: 0]\\
   SD & Normal(0, 2)&\\
@@
 
-Models were run with 4 chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}.
+Models were run with four chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}.
 For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats}.
-We chose slope priors which assume a wide range of effect sizes from -0.2 to -10\,$\mu$V, but focused on smaller effect sizes from -0.2 to -2\,$\mu$V.
-The focus on smaller effects is sensible because the interference effect investigated here is probably less pronounced than the well-established predictability effect on the N400 amplitude analyzed by \citet{nicenboim_stats}.
-
-\subsection{Results}
-Data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.
+We chose slope priors which assume a wide range of effect sizes from $-0.2$ to $-11$\,$\mu V$.
+However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$; this is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \cite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.
 
 \subsubsection{Comprehension question accuracy}
-After exclusion of participants with accuracy below 70\,\%, the overall accuracy (including fillers) was 82.8\,\% (range: 71.9 -- 100\,\%).
-Accuracy in critical trials was 75.6\,\% (range: 55 -- 95\,\%).
+After exclusion of participants with accuracy below 70\,\%, the overall accuracy (including fillers) was 82.8\,\% (range: [71.9, 100]\,\%).
+Accuracy in critical trials was 75.6\,\% ([55, 95]\,\%).
 By-condition accuracy is presented in Table \ref{tab:eeg_acc}.
@@
     \centering
-    \begin{tabular}{lrrr}
+    \begin{tabular}{lrr}
     \toprule
-    & Estimate & Est.Error & 95\% CrI  \\
+    & Estimate &  95\% CrI  \\
     \midrule
-Intercept& 1.39 & 0.15&  [1.10, 1.68]\\
-syntactic& -0.04 & 0.06& [-0.16, 0.08]\\
-semantic&  -0.48 & 0.06& [-0.61, -0.36]\\
-interaction& -0.17& 0.08& [-0.34, -0.01]\\
+Intercept& 1.39 &   [1.10, 1.68]\\
+syntactic& -0.04 &  [-0.16, 0.08]\\
+semantic&  -0.48 & [-0.61, -0.36]\\
+interaction& -0.17&  [-0.34, -0.01]\\
     \bottomrule
@@
 
-\subsubsection{ERPs}
-Grand average ERPs elicited by the critical verb for all conditions are presented in Figure \ref{fig:erp_all}.
-Table \ref{tab:eeg_mod} shows the estimates from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference with a difference up to -1\,$\mu$V.
-The 95\% credible intervals of syntactic interference, semantic interference and their interaction were consistent with effects for all of them.
-But the estimate of the semantic interference effect was the largest and its 95\% credible intervals was the furthest from zero ($\beta$ = -0.31\,$\mu$V, 95\% CrI = [-0.49, -0.13]\,$\mu$V, for the posteriors see Figure \ref{fig:eeg_posteriors}).
+\subsubsection{Event-related potentials}
+
+\paragraph{Estimated effect sizes and their uncertainty}
+
+Figure \ref{fig:erp_all} shows the 
+grand average ERPs elicited by the critical verb for all conditions are presented.
+Table \ref{tab:eeg_mod} shows the estimates from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu$V.
+The 95\% credible intervals of syntactic interference, semantic interference and their interaction were consistent with effects for all priors used.
+The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (mean: -0.31\,$\mu$V, 95\% CrI = [-0.49, -0.13]\,$\mu$V, for the posteriors see Figure \ref{fig:eeg_posteriors}).
 The estimate of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).
@@
 \begin{table}[H]
-    \caption{Results from the Bayesian linear mixed model analyzing the averaged brain response from 12 centro-parietal electrodes during the time window 300 to 500 ms post critical verb onset.
+    \caption{Results from the Bayesian linear mixed model analyzing the averaged brain response from 12 centro-parietal electrodes in the time window 300 to 500 ms post critical verb onset.
 Baseline represents the averaged activity from the 200 ms pre-verb onset and was included in the model as a covariate instead of traditional baseline correction \citep[see][]{alday2019}.}
@@
 
+\paragraph{Hypothesis testing using Bayes factors}
+
 Bayes factors are shown in Figure \ref{fig:eeg_bfs}. The Bayes factors provided very strong to strong evidence for the semantic interference effect for all priors 
@@
 Normal$_{-}$(0, 5): BF$_{10}$ = 11).
-In contrast, the Bayes factors provided only anecdotal evidence for the syntactic interference effect under the two narrow priors (Normal$_{-}$(0, 0.1): BF$_{10}$ = 2.5,
-Normal$_{-}$(0, 0.5): BF$_{10}$ = 1.2) and evidence \textit{against} it under wider priors 
+
+By contrast, the syntactic interference effect had only anecdotal evidence  under the two narrow priors (Normal$_{-}$(0, 0.1): BF$_{10}$ = 2.5,
+Normal$_{-}$(0, 0.5): BF$_{10}$ = 1.2). There was evidence \textit{against} syntactic interference under wider priors 
 (Normal$_{-}$(0, 1): BF$_{10}$ = 0.66,
-Normal$_{-}$(0, 5): BF$_{10}$ = 0.14). Across all priors, the Bayes factors provided evidence \textit{against} the interaction of semantic and syntactic interference (BF$_{10}$ < 0.85).
+Normal$_{-}$(0, 5): BF$_{10}$ = 0.14).
+What this means is that, in our data, only if we assume a priori that the effect size is relatively small, there is very weak evidence in favor of syntactic interference.
+
+Across all priors, the Bayes factors provided evidence \textit{against} the interaction of semantic and syntactic interference (BF$_{10}$ < 0.85).
 
 \begin{figure}
-    \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction during the time window 300 to 500 ms post critical verb onset at centro-parietal electrode sites.}
+    \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the time window 300 to 500 ms post critical verb onset at centro-parietal electrode sites.}
     \label{fig:eeg_bfs}
@@
 \subsection{Discussion}
-We presented a large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency formation.
-It is the first ERP experiment on this classic interference design.
-Strong evidence was found for an effect of semantic interference in the N400 amplitude.
-Bayes factors provided considerably weaker evidence in the N400 amplitude for an effect of syntactic interference under narrow priors and evidence against it under wider priors.
-Evidence against an interaction of syntactic and semantic interference in the N400 was found.
+We presented a relatively large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency completion.
+To our knowledge, this is the first ERP experiment on this classic interference design, which has only been investigated using reading studies so far \parencite{mertzen,vandyke07}.
+
 Comprehension accuracy was reduced for high vs.\ low semantic interference conditions. 
-A small interaction indicated that the difference in accuracy between high and low semantic interference conditions was larger, when syntactic interference was high.
 Syntactic interference itself did not affect accuracy.
-
-The to date rather small body of ERP studies on interference effects during sentence comprehension gives an indeterminate picture of the nature of interference effects in ERP measures.
+There was a numerical suggestion of an interaction: indicated that the difference in accuracy between low and high semantic interference conditions was larger when syntactic interference was high vs.\ when it was low (17 vs.\ 8).
+
+The Bayes factors analyses found strong evidence for semantic interference in the N400 amplitude.
+Syntactic interference had much weaker evidence in the N400 amplitude; only anecdotal evidence was found under narrow priors, and there was clear evidence against syntactic interference under wider priors.
+The evidence against an interaction under all priors was unequivocal.
+.
+
+Our finding that retrieval interference modulates the N400 amplitude supports the N400 as a marker of memory retrieval \citep{kutas&federmeier_2000, kutas_federmeier2011, brouwer2017_n4_p6, lau2008_n400}.
+
+Consistent with the self-paced reading study, there is clear evidence for semantic interference, and evidence against syntactic interference.
+How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
+It turns out that the answer to this question crucially depends on which cues one assumes are being used during retrieval.
+We turn to this point next by discussing the predictions of the \cite{Lewis2005} model.
+
+%The to date rather small body of ERP studies on interference effects during sentence comprehension gives an indeterminate picture of the nature of interference effects in ERP measures.
 The present study adds a large data point to the body of literature which reported negativities or more specifically N400 effects for high vs.\ low interference conditions \citep{vasishth_drenhaus_2011, martinetal2014, schoknecht2022, lee_garnsey}.
 Therefore, it contributes to the debate of how interference effects manifest themselves electrophysiologically (interference effects have also been reported in differences in the P600, \cite{Tanner_etal_2017, martinetal2014}).
-Our finding that retrieval interference modulates the N400 amplitude supports the N400 as a marker of memory retrieval account \citep{kutas&federmeier_2000, kutas_federmeier2011, brouwer2017_n4_p6, lau2008_n400}.
+
+\section{Quantitative predictions of the Lewis and Vasishth (2005) model for the current design}
+% stopped here xxx
+
+There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \cite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
+The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
+The syntactic effect is absent in our data.
+
+It would be implausible to argue that the absence of the syntactic interference effect suggests that the comprehender does not use syntactic information for subject-verb dependency resolution at all.
+We rather suggest that the \{± grammatical subject\} cue was not the right syntactic cue to investigate.
+In line with \citet{mertzen}, we suggest that a syntactic or structural cue along the lines of \{± same clause\} might be used during comprehension.
+Figure \ref{fig:model_predictions} presents the predictions of the \citet{Lewis2005} model for the \citet{vandyke07} design where additionally to the \{± animate\} cue either the canonical cue \{± grammatical subject\} or the alternative cue \{± same clause\} was implemented.
+The \{± same clause\} cue model does predict the syntactic interference effect to be centered around zero (95\,\% CrI [-4, 3]\,ms).
+This prediction is consistent with the results of the present study.
+However, the \{± same clause\} model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms).
+
+\begin{figure}[H]
+    \centering
+    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the Lewis \& Vasishth (2005) model as implemented in R and described by Engelmann et al.
+(2019).
+The implemented model is available at https://github.com/felixengelmann/inter-act/.}
+    \label{fig:model_predictions}\includegraphics[width=\textwidth]{images/PriorPredicted_LV05_engelmann_effects_syncue.png}
+\end{figure}
+
+The \citet{Lewis2005} model is generally used to predict reading times.
+The scaling factor \textit{F} is optimized to link the activation of items in memory to reading times \citep{Lewis2005}.
+But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
+The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor.
+Therefore, the \{± same clause\} cue instead of the \{± grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
 
@@
 This also partially explains the comparatively smaller effect sizes in the present experiments than have been previously reported.
-
-\subsection{Quantitative evaluation of Lewis and
-Vasishth (2005) model predictions}
-There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \cite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
-The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
-The syntactic effect is absent in our data.
-
-It would be implausible to argue that the absence of the syntactic interference effect suggests that the comprehender does not use syntactic information for subject-verb dependency resolution at all.
-We rather suggest that the \{± grammatical subject\} cue was not the right syntactic cue to investigate.
-In line with \citet{mertzen}, we suggest that a syntactic or structural cue along the lines of \{± same clause\} might be used during comprehension.
-Figure \ref{fig:model_predictions} presents the predictions of the \citet{Lewis2005} model for the \citet{vandyke07} design where additionally to the \{± animate\} cue either the canonical cue \{± grammatical subject\} or the alternative cue \{± same clause\} was implemented.
-The \{± same clause\} cue model does predict the syntactic interference effect to be centered around zero (95\,\% CrI [-4, 3]\,ms).
-This prediction is consistent with the results of the present study.
-However, the \{± same clause\} model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms).
-
-\begin{figure}[H]
-    \centering
-    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the Lewis \& Vasishth (2005) model as implemented in R and described by Engelmann et al.
-(2019).
-The implemented model is available at https://github.com/felixengelmann/inter-act/.}
-    \label{fig:model_predictions}\includegraphics[width=\textwidth]{images/PriorPredicted_LV05_engelmann_effects_syncue.png}
-\end{figure}
-
-The \citet{Lewis2005} model is generally used to predict reading times.
-The scaling factor \textit{F} is optimized to link the activation of items in memory to reading times \citep{Lewis2005}.
-But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
-The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor.
-Therefore, the \{± same clause\} cue instead of the \{± grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
 
@@
 The potential complications during memory retrieval due to interference should not differ dependent on the specific cue used for retrieval, i.e., \{±grammatical subject\} or \{±animate\}.
-Moreover, previous ERP studies investigating interference have found negativities during the typical N400 time window (300 - 500 ms post-stimulus onset), although the scalp topographies were not typical for the N400, being either broad \citep{schoknecht2022} or left anterior \citep{vasishth_drenhaus_2011}.
+Moreover, previous ERP studies investigating interference have found negativities in the typical N400 time window (300 - 500 ms post-stimulus onset), although the scalp topographies were not typical for the N400, being either broad \citep{schoknecht2022} or left anterior \citep{vasishth_drenhaus_2011}.
 Therefore, ERPs and especially the N400, are suited to be used for the investigation of interference (in the design from \citet{vandyke07} and in general).

### 2023-10-13 code_data/LV05_modeling/generate_priorpredictions.R
Updated via the Open Science Framework

@@ -1,227 +1,190 @@
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(parallel)
-library(LaplacesDemon)
-library(mvtnorm)
+# load packages
 library(truncnorm)
-library(reshape2)
+library(ggplot2)
+library(cowplot)
+library(stringr)
 
+#### THREE CUE MODEL FROM MERTZEN ET AL. (2023) https://osf.io/a7cg2/?view_only= ####
 
-## some helper functions:
-rmsd <- function (obs, pred) {
-  sqrt(mean((obs - pred)^2, na.rm = TRUE))
-}
-
-compute_int_means <- function(d){
-  cond_means <- as.data.frame(d) %>% group_by(Set, Target, Distractor, lf, ans, mas, mp, rth, bll, psc, pic, qcf, qco, cuesim, tprom, dprom, lp, ldp, blc, dbl, ndistr, cueweighting) %>% summarise(meanRT=mean(latency), SE=sd(latency)/sqrt(length(latency)))
-  int <- reshape(as.data.frame(d[,c(2,5,40)]),direction="wide",idvar = "Iteration",timevar = "Distractor")
-  colnames(int)[2:5] <- gsub("latency.","",colnames(int)[2:5])
-  int <- int %>% mutate(SynInt=(SynMatch+FullMatch)-(NoMatch+SemMatch),SemInt=(SemMatch+FullMatch)-(NoMatch+SynMatch),SynSemInt=FullMatch-SynMatch,NoSynSemInt=SemMatch-NoMatch)
-  Effect <- unname(colMeans(int))[-1]
-  return(Effect)
-}
-
-convert2log <- function(x){
-  ifelse(x>=1, log(x), ifelse(x<=-1, -log(abs(x)), 0))
-}
-
-convert2log10 <- function(x){
-  x <- ifelse(x>-1 & x<1, 0, x)
-  x <- ifelse(x<=-1, -log10(abs(x)), x) 
-  x <- ifelse(x>=1, log10(abs(x)), x)
-}
-
-
-# Introduction 
-
-## Load interact subject VERSION (syntactic cue = +-subject position)
-
-#The following piece of code, written by Felix Engelmann and available on github, provides the main computation code for ACT-R calculations.
-
-source("interACT_subject.R")
-
-## Basic engine for generating predictions
-
-printcounts<-FALSE
-
-iterate_lf <- function(values,iterations=1000){
[…]

### 2023-10-13 code_data/LV05_modeling/generate_priorpredictions.R
Updated via the Open Science Framework

@@ -90,7 +90,7 @@ model_4cond <- list(
   distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
   Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
   Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(0.33,0.33,0.33))
+  weights = list(c(strWeight(),semWeight()))
 )
 
 mas <<- c(1.5) # Maximum associative strength

### 2023-10-13 code_data/LV05_modeling/interACT_subject.R
Deleted via the Open Science Framework

@@ -1,340 +1,0 @@
-library(tidyr)
-library(dplyr)
-
-###############################################################
-## Global parameters
-###############################################################
-reset_params <- function(){
-  lf        <<- 0.15       # latency factor
-  le        <<- 1       # latency exponent
-  rth 	    <<- -1.5    # retrieval threshold
-  bll       <<- 0.5     # decay parameter
-  ans       <<- 0.2    # activation noise
-  mas       <<- 1       # maximum associative strength 
-  mp        <<- 1       # mismatch penalty
-  ga        <<- 1       # goal source activation
-  rand_time <<- 3       # latency variability
-  blc       <<- 0       # base-level activation
-  ##
-  lp        <<- 1     # default time since last presentation (msec)
-  ldp       <<- 1       # last distractor presentation (msec)
-  ## Distractor control:
-  ndistr    <<- 1				# number of distractors
-  dbl       <<- 0       # distractor base-level
-  ##
-  ## Cue weighting, cue confusion and prominence:
-  # cueweights <<- c(1,1) # vector of cue weightings
-  cueweighting <<- 1     # Strength of structural cue as ratio structural/morphological
-  normalizeWeights <<- TRUE 
-  qcf 		  <<- 1 			# match quality correction factor
-  qco 		  <<- -2*rth 			# match quality correction offset
-  psc       <<- 1       # prominence scaling constant C1
-  pic       <<- 0       # prominence scaling constant C2
-  tprom     <<- 0       # target prominence 
-  dprom     <<- 0       # distractor prominence
-  # cl 				<<- 0 			# cue confusion level (0-100)
-  cuesim    <<- -1    # cue-feature similarity [-1..0]
-  #
-  # Fitted meta-parameters
-	meta_recent <<- 0.7
-	meta_distant <<- 1.3
-	meta_lowprom <<- -0.5
-	meta_medprom <<- 0
-	meta_highprom <<- 2.5
-	meta_conflevel <<- 0
[…]

### 2023-10-14 paper/main.tex
Update main.tex
discussion of modeling edited

@@
   \cmidrule{2-4}
-  \multirow{4}{1cm}{beta} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{3}{*}[upper bound: 0]& [-0.2, - 0]\\
+  \multirow{4}{1cm}{slope} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{3}{*}[upper bound: 0]& [-0.2, - 0]\\
   &  Normal(0, 0.5)& & [-1, 0]\\
@@
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
-It turns out that the answer to this question crucially depends on which cues one assumes are being used during retrieval.
-We turn to this point next by discussing the predictions of the \cite{Lewis2005} model.
+We turn to this point next by discussing the quantiative predictions of the \cite{Lewis2005} model.
 
@@
 
-\section{Quantitative predictions of the Lewis and Vasishth (2005) model for the current design}
-% stopped here xxx
+\section{Quantitative predictions of the Lewis and Vasishth (2005) model for the present design}
+
+ \cite{mertzen} presented simulations that show approximately equally sized interference effects 
+for both the syntactic and semantic manipulations.
+Figure \ref{fig:model_predictions} reproduces thes simulations; these are prior predictive distributions of the effects of interest, holding all parameters in the model at fixed values and sampling from a prior distribution ($Beta(a=4,b=6)$), for the scaling parameter F in the \citet{Lewis2005}.
+This parameter scales the activation to a retrieval time.
+
+The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
 

### 2023-10-14 paper/main.tex
edited modeling section

@@
 for both the syntactic and semantic manipulations.
-Figure \ref{fig:model_predictions} reproduces thes simulations; these are prior predictive distributions of the effects of interest, holding all parameters in the model at fixed values and sampling from a prior distribution ($Beta(a=4,b=6)$), for the scaling parameter F in the \citet{Lewis2005}.
-This parameter scales the activation to a retrieval time.
-
-The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
-
-There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \cite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
+Their model assumed that three cues were used at the verb:  \{$\pm$ subject, $\pm$ same clause, $\pm$ animacy\}.
+The upper part of Figure \ref{fig:model_predictions} reproduces the \cite{mertzen}   simulations.
+These are prior predictive distributions of the effects of interest, by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter F and then generating data from the model while holding all other parameters in the model at fixed values.
+The F parameter scales the activation to a retrieval time.
+The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020}.
+
+Although the semantic interference effect estimates from the SPR study fall within the predicted range from the model, the syntactic interference effect is quite divergent: as discussed earlier, the SPR data show evidence against syntactic interference, but the model predicts a syntactic interference effect of approximately the same magnitude as semantic interference.
+Given that the EEG data also speak against syntactic interference, it is plausible that, at least in this experimental design and language, syntactic interference plays no role.
+
+The absence of (and evidence against) syntactic interference is surprising because the psycholinguistic literature has clearly established that syntactic constraints play a role in sentence processing.
+For example, \cite{dillon2013} and \cite{Sturt2003} argued that principle A of the binding theory controls the retrieval process in the antecedent-reflexive dependency, effectively eliminating interference from other overlapping features (such as number) of intervening nouns that cannot be an antecedent for the reflexive.
+For further evidence for this idea from individual differences, see \cite{yadav2021individual}.
+Even more relevant to the present study is the fact that \cite{mertzen} did find syntactic interference effects.
+
+Could the absence of syntactic interference be reconciled with a cue-based retrieval account?
+We speculated that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
+Specifically, the relevant syntactic cues that might be used at the verb is not  \{$\pm$ subject, $\pm$ same clause\} but rather a composite cue  \{$\pm$ subject-in-same-clause\} .
+In other words, in the spirit of the  \cite{dillon2013} and \cite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are   \{$\pm$ subject-in-same-clause, $\pm$ animate\}.
+The results of the prior predictive simulations are shown in
+the lower part of Figure \ref{fig:model_predictions}.
+This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
+One limitation here is that the prior predictive distribution of our post-hoc  model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR studywas [4, 12] ms at the critical region).
+This overestimation has to do with the fact that the prior on the F parameter is relatively spread out; it would be  possible to fit the parameter to the SPR data.
+
+The \citet{Lewis2005} model is generally used to predict reading times.
+The scaling factor \textit{F}  links the activation of items in memory to retrieval times \citep{Lewis2005}.
+But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
+The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor.
+Therefore, the \{$\pm$ subject-in-same-clause \} cue instead of the \{± grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
+
+Thus, one theoretical assumption that would be consistent with the observed SPR and EEG data is that the parser does not treat every subject in the sentence equally as a possible retrieval candidate; only the relevant subjects (those in the same clause) are considered.
+This is of course a speculative claim that would need to be further investigated empirically.
+
+%There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \cite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
 The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
@@
 
-It would be implausible to argue that the absence of the syntactic interference effect suggests that the comprehender does not use syntactic information for subject-verb dependency resolution at all.
+%It would be implausible to argue that the absence of the syntactic interference effect suggests that the comprehender does not use syntactic information for subject-verb dependency resolution at all.
 We rather suggest that the \{± grammatical subject\} cue was not the right syntactic cue to investigate.
@@
     \centering
-    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the Lewis \& Vasishth (2005) model as implemented in R and described by Engelmann et al.
-(2019).
-The implemented model is available at https://github.com/felixengelmann/inter-act/.}
-    \label{fig:model_predictions}\includegraphics[width=\textwidth]{images/PriorPredicted_LV05_engelmann_effects_syncue.png}
+    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the Lewis \& Vasishth (2005) model as implemented in R by Engelmann et al.\ (2019).
+The upper panel shows the predictions using the three retrieval cues used in Mertzen et al.\ 2022, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause, $\pm$ animate\}.
+    The implemented model is available with the code and data provided with this paper.}
+    \label{fig:model_predictions}\includegraphics[width=\textwidth]{images/PriorPredicted_interACTversions_2vs3.png}
 \end{figure}
-
-The \citet{Lewis2005} model is generally used to predict reading times.
-The scaling factor \textit{F} is optimized to link the activation of items in memory to reading times \citep{Lewis2005}.
-But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
-The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor.
-Therefore, the \{± same clause\} cue instead of the \{± grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
 

### 2023-10-14 paper/main.tex
minor edits to gen disc

@@
 The absence of (and evidence against) syntactic interference is surprising because the psycholinguistic literature has clearly established that syntactic constraints play a role in sentence processing.
-For example, \cite{dillon2013} and \cite{Sturt2003} argued that principle A of the binding theory controls the retrieval process in the antecedent-reflexive dependency, effectively eliminating interference from other overlapping features (such as number) of intervening nouns that cannot be an antecedent for the reflexive.
+For example, \cite{dillon2013} and \cite{Sturt2003} argued that principle A of the binding theory controls the retrieval process in the antecedent-reflexive dependency, effectively eliminating or attenuating interference from other overlapping features (such as number marking) of intervening nouns that cannot be an antecedent for the reflexive.
 For further evidence for this idea from individual differences, see \cite{yadav2021individual}.
@@
 \section{General discussion}
-The present SPR and ERP experiments consistently showed effects of semantic interference and no effects of syntactic interference.
-This result was not only found in the primary dependent measures of the experiments, respectively, but also in the comprehension accuracy.
-Given the replicability / reproducibility crises in psycholx \citep[see e.g., ][]{vasishth2018_signficancefilter}, and other (related) sciences \citep[see e.g., ][]{openscience2015_reproducibility}, the cross-methodological consistency is astonishing.
+
+Considered together, the present SPR and ERP experiments consistently showed effects of semantic interference and no effects of syntactic interference.
+This pattern was not only found in the primary dependent measures of the experiments, respectively, but also in the comprehension accuracy.
+The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval accounts is by assuming that the parser intelligently searches for a subject that is within the same clause.
+A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \cite[e.g.,][]{Sturt2003,dillon2013}.
+
+%Given the replicability / reproducibility crises in psycholinguistics \citep[see e.g., ][]{vasishth2018_signficancefilter,jaeger_etal_2020,nicenboim,}, and other (related) sciences \citep[see e.g., ][]{openscience2015_reproducibility,nicenboim_etal_2020,nieuwland_etal_2018}, the cross-methodological consistency is aston.
 It might be due to the very high sample sizes of the present experiments which granted the present study higher power than most comparable psycholinguistic studies.
 The higher power led to more precise estimates of the investigated effects with less uncertainty than in previous studies.
-This also partially explains the comparatively smaller effect sizes in the present experiments than have been previously reported.
+%This also partially explains the comparatively smaller effect sizes in the present experiments than have been previously reported.
+
+In the remainder of this section, we discuss several important issues that relate to previous experiments that investigated syntactic and semantic interference.
 
 \subsection{Encoding and retrieval interference}
-In the SPR data, we have found that the semantic interference effect originated at the distractor and persisted throughout following regions.
-Due to this time course, encoding interference was the best explanation for the interference effect in the reading times of the critical region (see discussion of Experiment 1).
+In the SPR data, we found that the semantic interference effect originated at the distractor and persisted throughout the following regions.
+Due to this time course, encoding interference \parencite{Yadavetal2022,hammerly2019grammaticality} is the best explanation for the interference effect in the reading times at the critical region.
 This raises the question whether the ERP effect was also caused by encoding interference, i.e., whether the ERP effect originated prior to the critical region.
-To answer this question, it is necessary to look at the ERPs elicited by words earlier in the sentence, although that is unusual in ERP research.
+To answer this question, it is necessary to look at the ERPs elicited by words earlier in the sentence, This is an unusual step in ERP research but makes sense as an exploratory step in the present context.
 Figure \ref{fig:erp_precrit} shows the ERPs for all conditions at the critical word and the two pre-critical adverbs which were identical across conditions.
@@
 Now that we have seen that the words directly preceding the critical word did not show the semantic interference effect, one might ask which brain response was elicited by the distractor itself (see Figure \ref{fig:erp_distractor})?
-Visual inspection of the ERPs elicited by the distractor shows several differences between conditions, but no semantic interference effect in the N400 spatio-temporal window.
-We assume that the differences in the ERPs elicited by the distractor in the different conditions are caused by two factors.
-Firstly, the high and low syntactic interference conditions have different word orders, i.e., the distractors are embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase, see Example \ref{ex:materials}).
+An exploratory visual inspection of the ERPs elicited by the distractor shows several differences between conditions, but no semantic interference effect in the N400 spatio-temporal window.
+We believe that the differences in the ERPs elicited by the distractor in the different conditions are caused by two factors.
+First, the high and low syntactic interference conditions have different word orders, i.e., the distractors are embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase, see Example \ref{ex:materials}).
 This is also apparent from the large difference between high and low syntactic interference conditions before distractor onset.

### 2023-10-14 paper/main.tex
minor edits to gen disc

@@
 
-Now that we have seen that the words directly preceding the critical word did not show the semantic interference effect, one might ask which brain response was elicited by the distractor itself (see Figure \ref{fig:erp_distractor})?
+Thus, the words directly preceding the critical verb did not show the semantic interference effect.
+A further question that arises is: which brain response was elicited by the distractor itself (see Figure \ref{fig:erp_distractor})?
 An exploratory visual inspection of the ERPs elicited by the distractor shows several differences between conditions, but no semantic interference effect in the N400 spatio-temporal window.
 We believe that the differences in the ERPs elicited by the distractor in the different conditions are caused by two factors.
-First, the high and low syntactic interference conditions have different word orders, i.e., the distractors are embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase, see Example \ref{ex:materials}).
+First, the high and low syntactic interference conditions have different word orders, i.e., the distractors are embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase; see Example \ref{ex:materials}).
 This is also apparent from the large difference between high and low syntactic interference conditions before distractor onset.
-Secondly, the distractors were not identical across conditions, therefore the ERPs in Figure \ref{fig:erp_distractor} are elicited by different words in the high vs.\ low semantic interference conditions.
-In conclusion, the distractors induced different brain responses across conditions, but these did not linger throughout the rest of the sentence (see Figure \ref{fig:erp_precrit}).
-%Firstly, there is a large difference prior to distractor onset (before time 0 ms) at all electrode sites.
-Secondly, there is a difference between 300 and 500\,ms most at anterior (left) electrode sites.
-We assume that both these differences are due to the different word orders in the high vs.
-low syntactic interference conditions, i.e., the different phrase structures that the distractors are embedded in (simple noun phrase vs.
-noun phrase including an adjective within a prepositional phrase, see Example \ref{ex:materials}).
-Thirdly, there is a late positivity (starting at approx.
-500\,ms) most pronounced at centro-parietal electrodes for high semantic vs.
-low semantic interference conditions.
-This semantic interference effect in the amplitude of a late positivity could be either due to
+Second, the distractors were not identical across conditions, therefore the ERPs in Figure \ref{fig:erp_distractor} are elicited by different words in the high vs.\ low semantic interference conditions.
+In summary, this exploratory analysis suggests that the distractors did induce different brain responses across conditions, but these could be due to factors other than encoding intererence.
+More importantly, these effects did not linger throughout the rest of the sentence (see Figure \ref{fig:erp_precrit}).
 
@@
 
-So, the comparison of the results of our SPR and ERP experiments yields that the SPR results were strongly affected by encoding interference, while the ERPs elicited by the critical verb were not.
+The comparison of the results of our SPR and ERP experiments suggests that the SPR results were strongly affected by encoding interference, while the ERPs elicited by the critical verb were not.
 This difference between methods might be caused by the difference in presentation rate.
@@
 In the ERP results, we observed an effect at the critical verb, which can be attributed to memory retrieval of the subject.
-This retrieval was only affected by the semantic cue \{± animate\} and not by the syntactic cue \{± grammatical subject\}.
+This retrieval was only affected by the experimental manipulation.
 

### 2023-10-15 paper/main.tex
finished one pass through paper

@@
 We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition.
+We are grateful to Dario Paape for helpful discussions, and Himanshu Yadav for assistance with the computational modeling.
 Correspondence concerning this article should be addressed to Pia Schoknecht, e-mail: pia.schoknecht@uni-potsdam.de
@@
 % What we did:
-We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07}, investigating interference due to syntactic and semantic cues in German.
+We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07,mertzen}, to investigate interference due to syntactic and semantic cues in German.
 % Not needed for an abstract, plus sounds boring:
@@
 Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to syntactic cues.
-This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play a role in retrieval.
-We show through computational modeling that cue-based retrieval also shows no syntactic interference if the parser is assumed to keep track of which clause a noun occurs in.
-Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model are consistent with the observed patterns in our data.
-
-In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that although both syntactic and semantic cues drive retrieval at the verb, syntactic cues include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
+This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play an equal role in retrieval.
+We show through computational modeling that cue-based retrieval also shows no syntactic interference if the parser is assumed to keep track of which clause the subject occurs in.
+Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model would be consistent with the observed patterns in our data.
+
+In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
 
@@
 
-Thus, one theoretical assumption that would be consistent with the observed SPR and EEG data is that the parser does not treat every subject in the sentence equally as a possible retrieval candidate; only the relevant subjects (those in the same clause) are considered.
+Thus, one theoretical assumption that would be consistent with the observed SPR and EEG data is that the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval; however, the parser uses the syntactic cue while keeping track of the syntactic location of the subject.
 This is of course a speculative claim that would need to be further investigated empirically.
@@
 However, while in SPR the slowdown due to encoding and maintaining the distractor in memory overshadowed potential retrieval effects, the procedure of the ERP experiment prevented that.
-In the ERP results, we observed an effect at the critical verb, which can be attributed to memory retrieval of the subject.
-This retrieval was only affected by the experimental manipulation.
+In the ERP results, we observed an effect at the critical verb, which can be attributed to the retrieval process hypothesized by cue-based retrieval accounts.
 
 \subsection{Relation to other sentence processing accounts}
-Besides our considerations above regarding the possibility that another syntactic cue (e.g., \{± same clause\}) was used, our results highlight the importance of the semantic cue for subject-verb dependency resolution.
-
-The evidence against a main effect of syntactic interference and hence the relative unimportance of the syntactic cue \{± grammatical subject\} for subject-verb dependency resolution in our study stands in contrast to predictions from syntax-first accounts \citep[e.g.,][]{frazier_clifton1996, frazier1979_diss, frazier1987}.
-These accounts predict that syntactic constraints are considered first during sentence processing.
-At least, our results speak against syntax as a strong filter.
-However, we found that a semantically similar distractor induced interference.
-This finding can be easily reconciled with various accounts of sentence processing.
-Above we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model.
-Our results fit predictions from constraint-based models \citep[e.g.,][]{macdonald1994, mcrae1998modeling, trueswell1993}.
+
+%Besides our considerations above regarding the possibility that another syntactic cue (e.g., \{± same clause\}) was used, our results highlight the importance of the semantic cue for subject-verb dependency resolution.
+
+The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{± grammatical subject\}.
+  However, the semantic interference effect we found can be easily reconciled with previous work in  sentence processing.
+In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model.
+  %Our results may also be consisted with constraint-based models \citep[e.g.,][]{macdonald1994, mcrae1998modeling, trueswell1993}.
 Under constraint-based accounts, it is assumed that multiple constraints simultaneously and immediately influence processing.
@@
 As discussed above, a plausible syntactic constraint is that a verb and its arguments need to be within the same clause.
-Furthermore, our results are in line with the good-enough processing account \citep{ferreira2007goodenough}.
-This account assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete or even incorrect analyses.
-The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials posed high processing demands.
-It is reasonable to assume that the participants might have sometimes adopted a good-enough processing mode to manage resources.
+ Our results could also be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
+The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands.
+It is reasonable to assume that the participants might have--at least occasionally--adopted a good-enough processing mode when faced with high task demands \parencite{swets2008underspecification,LogacevMultiple} or due to working memory capacity limitations \parencite{MalsburgVasishth2013}, or both.
 So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution.
 Animacy is an obvious choice for such a heuristic.
-Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject probably leads to the correct analysis (not just within the experiment context, but in everyday language use).
-Finally, our findings support the possibility that language processing relies predominantly on semantic associations to form (probabilistic)
+Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject leads to the correct analysis with high probability (not just within the experiment context, but in everyday language use).
+Finally, our findings are also consistent with the possibility that language processing relies predominantly on semantic associations to form (probabilistic)
 representation of event structures.
-This view has been put forward by \citet{rabovsky_etal_2018}, when they used the sentence gestalt model \citep{mcclelland1989_sentence_gestalt}, a neural network model of sentence comprehension, to model the N400 amplitude.
+This view has been put forward by \cite{rabovsky_etal_2018}, when they used the neural-network sentence gestalt model \parencite{mcclelland1989_sentence_gestalt}, to model the N400 amplitude.
 This model assumes that language comprehension relies on associative form-to-meaning mapping instead of syntactic rules.
-This is in line with our results which highlight the importance of semantic features such as animacy over syntactic functions such as grammatical subject for sentence processing.
+This assumption is consistent with our results.
+
+Without the modeling results reported above, it would be easy to conclude that our data show that  syntactic cues play no role in sentence processing.
+However, this conclusion would be obviously wrong.
+First, as we showed above, the retrieval model suggests that--at least in our data--the subject cue is used, but differently than previously assumed: the relevant cue is subject-in-same-clause.
+Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite{ALV2019,HusainEtAl2014}, syntactic constraints like Principle A of the binding theory \parencite{Sturt2003,dillon2013,yadav2021individual},  and
+the importance of word order \parencite{meng2000mode}.
+There are of course important examples of the parser ignoring syntactic information.
+Three dramatic examples are the grammaticality illusion in multiple embeddings \parencite{gibsonthomas99}, the agreement attraction phenomenon \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}, and local coherence effects \parencite{taboretal04}.
+However, there are languages (such as German and Dutch) in which the grammaticality illusion does not occur \parencite{VSLK11,FrankEtAl2015}, and a large-sample study \parencite{lcpaape2023} presents evidence against the syntactic local coherence effect that was reported by \cite{taboretal04}.
+Thus, it seems that the role of syntax in building incremental structure remains as important as ever.
 
@@
 This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials.
-While our results did not directly replicate previous results from eye-tracking while reading studies \citep{vandyke07, mertzen}, our SPR and ERP results were consistent.
-We attributed differences between the present and previous studies to differences in statistical power and experimental methodology.
-Both our experiments showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor.
-This manifested in longer self-paced reading times starting at the distractor and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb.
-We discussed that these different time courses in the two experiments are due to different types of interference: Encoding interference in the SPR experiment and retrieval interference in the ERP experiment.
-We did not find effects of syntactic interference.
-We assume that the syntactic cue \{± grammatical subject\} which was manipulated in our materials, was not used during sentence processing.
-We proposed that another syntactic cue \{± same clause\} might be used \citep[see also][]{mertzen}.
-We suggest that the present findings, which are in line with various sentence processing accounts, highlight the importance of semantic cues during sentence processing.
+Both our experiments consistently showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor.
+This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb.
+These different time courses in the two experiments are likely to be due to different types of interference: Encoding interference in the SPR experiment and retrieval interference in the ERP experiment.
+These differences are likely to be due to the methods used (self-paced vs.\ auto-paced reading).Overall, we found evidence against syntactic interference; computational modeling shows that the most likely explanation for this is that the parser uses the syntactic cue  \{± subject-in-same-clause\}  to identify the correct target for retrieval.
+A broader implication is that the parser may be able to use hierarchical structure in the input to intelligently target the correct syntactic dependent from memory.
 

### 2023-10-15 paper/main.tex
minor edits

@@
 
-Considered together, the present SPR and ERP experiments consistently showed effects of semantic interference and no effects of syntactic interference.
-This pattern was not only found in the primary dependent measures of the experiments, respectively, but also in the comprehension accuracy.
-The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval accounts is by assuming that the parser intelligently searches for a subject that is within the same clause.
+Taken together, the SPR and ERP experiments consistently showed effects of semantic interference and no effects of syntactic interference.
+This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
+The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is by assuming that the parser intelligently searches for a subject that is within the same clause, and uses the animacy cue without reference to the clause that a noun appears in.
 A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \cite[e.g.,][]{Sturt2003,dillon2013}.
@@
 
-In the remainder of this section, we discuss several important issues that relate to previous experiments that investigated syntactic and semantic interference.
+In the remainder of this section, we discuss two important issues that relate to previous experiments that investigated syntactic and semantic interference.
+The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.
 
@@
 
-Thus, the words directly preceding the critical verb did not show the semantic interference effect.
+Thus, the words directly preceding the critical verb do not show a pattern consistent with semantic interference.
 A further question that arises is: which brain response was elicited by the distractor itself (see Figure \ref{fig:erp_distractor})?
@@
 Here, it is noteworthy that the presentation duration of 500\,ms per word and 400\,ms between words which was used in the present ERP experiment is slower than usual natural reading speed.
-However, while in SPR the slowdown due to encoding and maintaining the distractor in memory overshadowed potential retrieval effects, the procedure of the ERP experiment prevented that.
-In the ERP results, we observed an effect at the critical verb, which can be attributed to the retrieval process hypothesized by cue-based retrieval accounts.
+However, while in SPR the slowdown due to encoding and maintaining the distractor in memory overshadowed potential retrieval effects, the procedure of the ERP experiment might have prevented that.
+In the ERP results, we observed an effect at the critical verb, which could be attributed exclusively to the retrieval process hypothesized by cue-based retrieval accounts.
+Having said that, we cannot rule out the possibility that--due to the auto-paced nature of the ERP experiment--any encoding interference that started at or after the distractor is mixed in with the cost of retrieval interference.
+Indeed, recent modeling work on agreement attraction in reading \parencite{Yadavetal2022} has shown conclusively that some kind of feature distortion as well as cue-based retrieval are needed to explain the benchmark data-sets that exist so far on agreement attraction in reading.
+If the  \cite{Yadavetal2022} extends to interference effects in general,
+it is reasonable to assume that the effect observed at the verb in the ERP study is driven by both encoding and retrieval interference.
+However, we cannot resolve the relative roles of these latent processes in the present study.
 
@@
 The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands.
-It is reasonable to assume that the participants might have--at least occasionally--adopted a good-enough processing mode when faced with high task demands \parencite{swets2008underspecification,LogacevMultiple} or due to working memory capacity limitations \parencite{MalsburgVasishth2013}, or both.
+It is reasonable to assume that the participants might have--at least occasionally--adopted a good-enough processing mode when faced with high task demands \parencite{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \parencite{MalsburgVasishth2013}, or both.
 So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution.
@@
 However, this conclusion would be obviously wrong.
-First, as we showed above, the retrieval model suggests that--at least in our data--the subject cue is used, but differently than previously assumed: the relevant cue is subject-in-same-clause.
-Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite{ALV2019,HusainEtAl2014}, syntactic constraints like Principle A of the binding theory \parencite{Sturt2003,dillon2013,yadav2021individual},  and
-the importance of word order \parencite{meng2000mode}.
+First, as we showed above, the retrieval model suggests that--at least in our data--the subject cue is used, but differently than previously assumed: the relevant cue may be subject-in-same-clause.
+Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite[e.g.,][]{ALV2019,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \parencite[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and
+the importance of word order \parencite[e.g.,][]{meng2000mode}.
 There are of course important examples of the parser ignoring syntactic information.
-Three dramatic examples are the grammaticality illusion in multiple embeddings \parencite{gibsonthomas99}, the agreement attraction phenomenon \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}, and local coherence effects \parencite{taboretal04}.
+Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, the agreement attraction phenomenon across different languages \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}, and local coherence effects \parencite{taboretal04}.
 However, there are languages (such as German and Dutch) in which the grammaticality illusion does not occur \parencite{VSLK11,FrankEtAl2015}, and a large-sample study \parencite{lcpaape2023} presents evidence against the syntactic local coherence effect that was reported by \cite{taboretal04}.
-Thus, it seems that the role of syntax in building incremental structure remains as important as ever.
+Also, there seems to be at least one language that has rich case marking (Czech) that does not show any evidence of agreement attraction \parencite{chromy2023number}.
+Thus, it seems that, cross-linguistically, the role of syntax in building incremental structure and in completing dependencies remains as important as ever.
 
@@
 This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb.
-These different time courses in the two experiments are likely to be due to different types of interference: Encoding interference in the SPR experiment and retrieval interference in the ERP experiment.
-These differences are likely to be due to the methods used (self-paced vs.\ auto-paced reading).Overall, we found evidence against syntactic interference; computational modeling shows that the most likely explanation for this is that the parser uses the syntactic cue  \{± subject-in-same-clause\}  to identify the correct target for retrieval.
-A broader implication is that the parser may be able to use hierarchical structure in the input to intelligently target the correct syntactic dependent from memory.
+These different time courses in the two experiments are likely to be due to different types of interference: Encoding interference in the SPR experiment and retrieval (and possibly also encoding) interference in the ERP experiment.
+These differences in the time course are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading).
+Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that the most likely explanation for the observed data is that the parser uses the syntactic cue  \{± subject-in-same-clause\}  to identify the correct target for retrieval.
+A broader implication is that the parser may be able to use hierarchical structure in the input to intelligently target the correct syntactic dependent from memory, leading to no interference from a distractor that is in subject position within an embedded clause.
 

### 2023-10-16 code_data/LV05_modeling/generate_posteriorpredictions.R
Updated via the Open Science Framework

@@ -27,31 +27,32 @@ VERBOSE <<- FALSE
 
 # Load simulated data
 load("simdat_2cues_sameclause.Rda")
-# values used to generate prior predictions: 
-# nsims<-5000
-# lfs<-rbeta(nsims,a,b)
 
 # Run ABC with rejection sampling
-# results from Schoknecht & Vasishth (critical region): 
-# syntactic mean effect 2ms CrI [0,4] --> Normal(2,1) --> SE: 1
-# semantic mean effect 8ms CrI [4,12] --> Normal(8,2) --> SE: 2
-# interaction mean effect 14ms, CrI [7, 22] --> Normal(14,4) --> SE: 4
+# results from Schoknecht & Vasishth (N400 elicited by critical verb): 
+# syntactic mean effect -0.2 muV, CrI [-0.4,0.0] --> Normal(-0.2,0.1) --> SE: 0.1
+# semantic mean effect -0.3 muV, CrI [-0.6,-0.1] --> Normal(-0.3,0.15) --> SE: 0.15
+# interaction mean effect -0.1 muV, CrI [-0.4, 0.0] --> Normal(-0.1,0.1) --> SE: 0.1
 
-synlower <- 2-1 # mean effect - 1*SE
-synupper <- 2+1 # mean effect + 1*SE
+# I'm changing the sign of the effect here (sem: -0.3 --> 0.3)
+# to match the sign of the latency factor
+# --> a positive effect represents are more negative N400 
 
-semlower <- 8-2 # mean effect - 1*SE
-semupper <- 8+2 # mean effect + 1*SE
+synlower <- 0.2-(2*0.1) # mean effect - 2*SE
+synupper <- 0.2+(2*0.1) # mean effect + 2*SE
 
-synsemlower <- 14-4 # mean effect - 1*SE
-synsemupper <- 14+4 # mean effect + 1*SE
+semlower <- 0.3-(2*0.15) # mean effect - 2*SE
+semupper <- 0.3+(2*0.15) # mean effect + 2*SE
+
+synsemlower <- 0.1-(2*0.1) # mean effect - 2*SE
+synsemupper <- 0.1+(2*0.1) # mean effect + 2*SE
 
 # posterior latency factor (we base this only on the semantic interference result)
 SEM<-semlower<=simdat_2cues_sameclause$MESem & simdat_2cues_sameclause$MESem<=semupper
 table(SEM)
 posterior_lf_sem <-simdat_2cues_sameclause[SEM,]$lfs
 
[…]

### 2023-10-16 code_data/LV05_modeling/brain_response/generate_priorpredictions.R
Added via the Open Science Framework

@@ -1,0 +1,121 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+# Load model 
+source("interACT.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate 20,000 simulations from model:
+nsims<-20000
+
+## Initial hyperparameters for latency factor
+## Visualization of priors
+
+#Truncated Normal
+latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
+                                                    a=0.003, b=0.04, mean=0.01, sd=0.005))
+
+prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
+  geom_histogram(aes(y=..density..),
+                 position="identity",fill="gray",
+                 binwidth=0.001, center = 0.0005)+
+  theme_bw()+
+  theme(strip.text.x = element_text(size = 16, colour = "black", angle = 0))+
+  xlab("latency factor")+
+  ggtitle("Prior on scaling factor")
+prior_lf
+ggsave("LF_Prior.png", width=5, height=3, dpi=600)
+
+lfs<-rtruncnorm(nsims, a=0.003, b=0.04, mean=0.01, sd=0.005)
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
[…]

### 2023-10-16 code_data/LV05_modeling/brain_response/generate_posteriorpredictions.R
Added via the Open Science Framework

@@ -1,0 +1,175 @@
+library(truncnorm)
+library(tidyverse)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+# Load model
+source("interACT.R")
+reset_params()
+
+# TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Load simulated data
+load("simdat_2cues_sameclause.Rda")
+
+# Run ABC with rejection sampling
+# results from Schoknecht & Vasishth (N400 elicited by critical verb): 
+# syntactic mean effect -0.2 muV, CrI [-0.4,0.0] --> Normal(-0.2,0.1) --> SE: 0.1
+# semantic mean effect -0.3 muV, CrI [-0.6,-0.1] --> Normal(-0.3,0.15) --> SE: 0.15
+# interaction mean effect -0.1 muV, CrI [-0.4, 0.0] --> Normal(-0.1,0.1) --> SE: 0.1
+
+# I'm changing the sign of the effect here (sem: -0.3 --> 0.3)
+# to match the sign of the latency factor
+# --> a positive effect represents are more negative N400 
+
+synlower <- 0.2-(2*0.1) # mean effect - 2*SE
+synupper <- 0.2+(2*0.1) # mean effect + 2*SE
+
+semlower <- 0.3-(2*0.15) # mean effect - 2*SE
+semupper <- 0.3+(2*0.15) # mean effect + 2*SE
+
+synsemlower <- 0.1-(2*0.1) # mean effect - 2*SE
+synsemupper <- 0.1+(2*0.1) # mean effect + 2*SE
+
+# posterior latency factor (we base this only on the semantic interference result)
[…]

### 2023-10-16 code_data/LV05_modeling/retrieval_times/generate_priorpredictions.R
Added via the Open Science Framework

@@ -1,0 +1,265 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+#### THREE CUE MODEL FROM MERTZEN ET AL. (2023) https://osf.io/a7cg2/?view_only= ####
+
+# Load model 
+source("interACT-threecues.R")
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
+  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(0.33,0.33,0.33))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate 20,000 simulations from model:
+nsims<-20000
+
+## Initial hyperparameters for latency factor
+## from Methodsx paper:
+a<-4
+b<-6
+lfs<-rbeta(nsims,a,b)
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=5000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
[…]

### 2023-10-16 code_data/LV05_modeling/retrieval_times/generate_posteriorpredictions.R
Added via the Open Science Framework

@@ -1,0 +1,174 @@
+library(truncnorm)
+library(tidyverse)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+# Load model
+source("interACT.R")
+reset_params()
+
+# TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Load simulated data
+load("simdat_2cues_sameclause.Rda")
+# values used to generate prior predictions: 
+# nsims<-20000
+# lfs<-rbeta(nsims,a,b)
+
+# Run ABC with rejection sampling
+# results from Schoknecht & Vasishth (critical region): 
+# syntactic mean effect 2ms CrI [0,4] --> Normal(2,1) --> SE: 1
+# semantic mean effect 8ms CrI [4,12] --> Normal(8,2) --> SE: 2
+# interaction mean effect 14ms, CrI [7, 22] --> Normal(14,4) --> SE: 4
+
+synlower <- 2-(2*1) # mean effect - 2*SE
+synupper <- 2+(2*1) # mean effect + 2*SE
+
+semlower <- 8-(2*2) # mean effect - 2*SE
+semupper <- 8+(2*2) # mean effect + 2*SE
+
+synsemlower <- 14-(2*4) # mean effect - 2*SE
+synsemupper <- 14+(2*4) # mean effect + 2*SE
+
+# posterior latency factor (we base this only on the semantic interference result)
+SEM<-semlower<=simdat_2cues_sameclause$MESem & simdat_2cues_sameclause$MESem<=semupper
+table(SEM)
[…]

### 2023-10-16 code_data/LV05_modeling/generate_priorpredictions.R
Delete code_data/LV05_modeling/generate_priorpredictions.R

@@ -1,190 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-#### THREE CUE MODEL FROM MERTZEN ET AL. (2023) https://osf.io/a7cg2/?view_only= ####
-
-# Load model 
-source("interACT-threecues.R")
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
-  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(0.33,0.33,0.33))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Generate 5,000 simulations from model:
-nsims<-5000
-
-## Initial hyperparameters for latency factor
-## from Methodsx paper:
-a<-4
-b<-6
-lfs<-rbeta(nsims,a,b)
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
-  lf<<- lfs[i]  
-  sims <- create_param_matrix(model_4cond, iterations=5000)
-  results <- run(sims)
-  condMeans <- compute_cond_means(results)
-  ## compute main effect of Syntactic manipulation
-  ## HiSyn-LoSyn:
-  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
-    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
-  ## save the estimate:
-  maineffectSyn[i]<-meSyn
-  ## compute main effect of Semantic manipulation
-  ## HiSem-LoSem:
[…]

### 2023-10-16 paper/main.tex
fixed typo
changed all \cite to \textcite commands

@@
 %\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage[natbib,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
+\usepackage[natbib=true,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
 \usepackage{csquotes}
@@
 This situation can occur when the relative clause  in \ref{ex:1_vandyke07} that modifies \textit{the resident} were to be changed to \textit{who said that the warehouse was dangerous}.
-Now, the distractor matches the \{+grammatical subject\}.
+Now, the distractor matches the \{+grammatical subject\} cue.
 
@@
 However, two important questions remain unanswered.
-First, as discussed in \cite{jaeger_etal_2017}, almost all the published studies on interference effects are severely underpowered.
+First, as discussed in \textcite{jaeger_etal_2017}, almost all the published studies on interference effects are severely underpowered.
 Low power has the consequence that published estimates will tend to be overestimates or misestimates of the effect size \parencite{vasishth2018_signficancefilter}.
 Over-/misestimates have important implications for theory evaluation.
-One example is discussed in \cite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \parencite{levy&keller_2013} could not be validated when a larger-sample study was conducted; by contrast, the original small-sample study  \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory.
+One example is discussed in \textcite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \parencite{levy&keller_2013} could not be validated when a larger-sample study was conducted; by contrast, the original small-sample study  \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory.
 Another example is the recent heated debate on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
 Second, although it is clear and uncontroversial  \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
-For example, \cite{dillon2013} argued, following \cite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built; instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
-Using computational modeling, \cite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference, and that the reason for this is that principle A of the binding theory is the primary driver of the cue-based retrieval process \parencite[cf.
+For example, \textcite{dillon2013} argued, following \textcite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built; instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
+Using computational modeling, \textcite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference, and that the reason for this is that principle A of the binding theory is the primary driver of the cue-based retrieval process \parencite[cf.
 ][]{jaeger_etal_2020,yadav2021individual}.
 This was a surprising discovery because a priori, cue-based retrieval accounts do not allow for exceptions to retrieval interference in particular syntactic dependency types.
-More broadly, an important implication of the work initiated by \cite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
+More broadly, an important implication of the work initiated by \textcite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
 
@@
 
-All experiments used the 2 $\times$ 2 design developed by \citet{vandyke07}; the design crosses syntactic and semantic interference.
+All experiments used the 2 $\times$ 2 design developed by \textcite{vandyke07}; the design crosses syntactic and semantic interference.
 
@@
 Syntactic interference is manipulated via the syntactic status of the distractor, which is either subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}).
-Following \cite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in any subject position causes more syntactic interference than a distractor inside a prepositional phrase.
+Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in any subject position causes more syntactic interference than a distractor inside a prepositional phrase.
 
@@
 
-Indeed, previous work using the same design as in the present paper has also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
+Indeed, previous work using the same design as in the present paper has also found semantic interference effects at the pre-critical region: both \textcite{vandyke07} and \textcite{mertzen} found such effects.
 \citeauthor{mertzen}'s (\citeyear{mertzen}) Figures 5 and 6 indicated reading time differences in earlier regions, especially for their German data (see their Figure 6).
-\cite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \cite{mertzen} also points out, encoding interference could be an explanation even in that study.
+\textcite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \textcite{mertzen} also points out, encoding interference could be an explanation even in that study.
 Given these earlier findings, our results are consistent with the encoding interference explanation.
 
-\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. 
+\textcite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region. 
 They discuss several alternative explanations for their observed effects.
-In addition to encoding interference, \cite{mertzen} propose three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
+In addition to encoding interference, \textcite{mertzen} propose three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
 Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
@@
 \citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively.
-\cite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
+\textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
 In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
@@
 We chose slope priors which assume a wide range of effect sizes from $-0.2$ to $-11$\,$\mu V$.
-However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$; this is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \cite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.
+However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$; this is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \textcite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.
 
@@
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
-We turn to this point next by discussing the quantiative predictions of the \cite{Lewis2005} model.
+We turn to this point next by discussing the quantiative predictions of the \textcite{Lewis2005} model.
 
@@
 The present study adds a large data point to the body of literature which reported negativities or more specifically N400 effects for high vs.\ low interference conditions \citep{vasishth_drenhaus_2011, martinetal2014, schoknecht2022, lee_garnsey}.
-Therefore, it contributes to the debate of how interference effects manifest themselves electrophysiologically (interference effects have also been reported in differences in the P600, \cite{Tanner_etal_2017, martinetal2014}).
+Therefore, it contributes to the debate of how interference effects manifest themselves electrophysiologically (interference effects have also been reported in differences in the P600, \textcite{Tanner_etal_2017, martinetal2014}).
 
@@
 
- \cite{mertzen} presented simulations that show approximately equally sized interference effects 
+ \textcite{mertzen} presented simulations that show approximately equally sized interference effects 
 for both the syntactic and semantic manipulations.
 Their model assumed that three cues were used at the verb:  \{$\pm$ subject, $\pm$ same clause, $\pm$ animacy\}.
-The upper part of Figure \ref{fig:model_predictions} reproduces the \cite{mertzen}   simulations.
+The upper part of Figure \ref{fig:model_predictions} reproduces the \textcite{mertzen}   simulations.
 These are prior predictive distributions of the effects of interest, by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter F and then generating data from the model while holding all other parameters in the model at fixed values.
@@
 The absence of (and evidence against) syntactic interference is surprising because the psycholinguistic literature has clearly established that syntactic constraints play a role in sentence processing.
-For example, \cite{dillon2013} and \cite{Sturt2003} argued that principle A of the binding theory controls the retrieval process in the antecedent-reflexive dependency, effectively eliminating or attenuating interference from other overlapping features (such as number marking) of intervening nouns that cannot be an antecedent for the reflexive.
-For further evidence for this idea from individual differences, see \cite{yadav2021individual}.
-Even more relevant to the present study is the fact that \cite{mertzen} did find syntactic interference effects.
+For example, \textcite{dillon2013} and \textcite{Sturt2003} argued that principle A of the binding theory controls the retrieval process in the antecedent-reflexive dependency, effectively eliminating or attenuating interference from other overlapping features (such as number marking) of intervening nouns that cannot be an antecedent for the reflexive.
+For further evidence for this idea from individual differences, see \textcite{yadav2021individual}.
+Even more relevant to the present study is the fact that \textcite{mertzen} did find syntactic interference effects.
 
@@
 Specifically, the relevant syntactic cues that might be used at the verb is not  \{$\pm$ subject, $\pm$ same clause\} but rather a composite cue  \{$\pm$ subject-in-same-clause\} .
-In other words, in the spirit of the  \cite{dillon2013} and \cite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are   \{$\pm$ subject-in-same-clause, $\pm$ animate\}.
+In other words, in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are   \{$\pm$ subject-in-same-clause, $\pm$ animate\}.
 The results of the prior predictive simulations are shown in
@@
 
-%There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \cite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
+%There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \textcite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
 The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
@@
 Indeed, recent modeling work on agreement attraction in reading \parencite{Yadavetal2022} has shown conclusively that some kind of feature distortion as well as cue-based retrieval are needed to explain the benchmark data-sets that exist so far on agreement attraction in reading.
-If the  \cite{Yadavetal2022} extends to interference effects in general,
+If the  \textcite{Yadavetal2022} extends to interference effects in general,
 it is reasonable to assume that the effect observed at the verb in the ERP study is driven by both encoding and retrieval interference.
@@
 
-The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{± grammatical subject\}.
+The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{± grammatical subject\}.
   However, the semantic interference effect we found can be easily reconciled with previous work in  sentence processing.
@@
 representation of event structures.
-This view has been put forward by \cite{rabovsky_etal_2018}, when they used the neural-network sentence gestalt model \parencite{mcclelland1989_sentence_gestalt}, to model the N400 amplitude.
+This view has been put forward by \textcite{rabovsky_etal_2018}, when they used the neural-network sentence gestalt model \parencite{mcclelland1989_sentence_gestalt}, to model the N400 amplitude.
 This model assumes that language comprehension relies on associative form-to-meaning mapping instead of syntactic rules.
@@
 Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, the agreement attraction phenomenon across different languages \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}, and local coherence effects \parencite{taboretal04}.
-However, there are languages (such as German and Dutch) in which the grammaticality illusion does not occur \parencite{VSLK11,FrankEtAl2015}, and a large-sample study \parencite{lcpaape2023} presents evidence against the syntactic local coherence effect that was reported by \cite{taboretal04}.
+However, there are languages (such as German and Dutch) in which the grammaticality illusion does not occur \parencite{VSLK11,FrankEtAl2015}, and a large-sample study \parencite{lcpaape2023} presents evidence against the syntactic local coherence effect that was reported by \textcite{taboretal04}.
 Also, there seems to be at least one language that has rich case marking (Czech) that does not show any evidence of agreement attraction \parencite{chromy2023number}.

### 2023-10-16 paper/main.tex
minor edits up to Exp2: EEG

@@
 Over-/misestimates have important implications for theory evaluation.
-One example is discussed in \textcite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \parencite{levy&keller_2013} could not be validated when a larger-sample study was conducted; by contrast, the original small-sample study  \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory.
+One example is discussed in \textcite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \parencite{levy&keller_2013} could not be validated when a larger-sample study was conducted.
+By contrast, the original small-sample study \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory.
 Another example is the recent heated debate on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
 Second, although it is clear and uncontroversial \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
-For example, \textcite{dillon2013} argued, following \textcite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built; instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
+For example, \textcite{dillon2013} argued, following \textcite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built.
+Instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
 Using computational modeling, \textcite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference, and that the reason for this is that principle A of the binding theory is the primary driver of the cue-based retrieval process \parencite[cf.
@@
 Reading studies have been the dominant method used to investigate retrieval interference \parencite[see the review in][]{jaeger_etal_2017}.
-We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading as well as EEG, which to our knowledge has never been used to investigate interference during subject-verb dependency resolution
+We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading as well as EEG, which to our knowledge has never been used to investigate interference during subject-verb dependency resolution.
 
@@
 
-Although our principal focus was on the reading times at the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region (we return to this point in the discussion sections).
-A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another ( \textit{erzählt hatte, trank}, literally `told, drank').
+Although our principal focus was on the processing of the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region (we return to this point in the discussion sections).
+A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another (\textit{erzählt hatte, trank}, literally `told, drank').
 This is because in German, the verb is clause-final in embedded clauses.
@@
 
-Thus, each of the forty participants were asked to rate $120 \times 4$ items, leading to 19,200 data points.
+Thus, each of the forty participants were asked to rate $120 \times 4$ sentences, leading to 19,200 data points.
 
@@
 \caption{Plausibility ratings for the animate (high semantic interference conditions) and inanimate distractor (low semantic interference conditions), the introduction noun (\textit{X believed that \dots}) and the subject.
-The diamonds represent the mean rating.}\label{fig:plausibility_rating}
+The diamonds represent the mean rating.
+Blue dots show the individual data points.
+}\label{fig:plausibility_rating}
 \centering
@@
 Because the high signal-to-noise ratio of EEG data requires a large number of trials per condition and participant, we constructed 120 items.
-In order to keep the SPR and EEG studies comparable, we also used 120 items in SPR experiments.
+In order to keep the SPR and EEG studies comparable, we also used 120 items in the SPR experiments.
 In both methods, as it would have been very taxing for the participants to read all 120 items in one experimental session, we split the experiments into two sessions.
@@
 In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
-As mentioned above, because the ERP method requires a large number of items, all the data come from participants who completed both sessions.
+As mentioned above, because the ERP method requires a large number of items, all the ERP data come from participants who completed both sessions.
 
@@
 The Bayes factor is analogous to the frequentist ANOVA or likelihood ratio test: it compares the (marginal) likelihood of a full model to that of a baseline null model; this is often written as $BF_{10}$, with the subscripts indicating that marginal likelihood of the full model is in the numerator and that of the null model is in the denominator.
-One major advantage of the Bayes factor over the frequentist analogue is that it takes the uncertainty of the parameters into account; this leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
-Another important advantage of the Bayes factor---one that is very relevant for the present work---is that it is in principle possible to find evidence both for or against an effect; this is contrast to the frequentist ANOVA, which is designed to only furnish evidence against the null.
+One major advantage of the Bayes factor over the frequentist analogue is that it takes the uncertainty of the parameters into account.
+This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
+Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is in principle possible to find evidence both for or against an effect.
+This stands in contrast to the frequentist ANOVA, which is designed to only furnish evidence against the null.
 
@@
     \midrule
-  Intercept & Normal(6, 0.6)& & 125-1308\\
+  Intercept & Normal(6, 0.6)& & [125, 1308]\\
   \cmidrule{2-4}
@@
 Panel C shows the pooled reading times of all conditions in the critical and surrounding regions.
-Panel D shows the reading times of all conditions in the critical and surrounding regions in the different parts of the self-paced reading experiment.}
+Panels D and E show the reading times of all conditions in the critical and surrounding regions in the different parts of the self-paced reading experiment.}
     \label{fig:whole_sentence}
@@
 
-Because the distractor and the immediately following regions differ by condition, we did not compute Bayes factors for these differences.
+Because the distractor and the immediately following regions differ between conditions, we did not compute Bayes factors for the observed reading time differences.
 Instead we focused our Bayes factors analyses on regions which were identical across conditions.
@@
 
-Panel D shows the reading times of the critical and surrounding regions for each of the two Experimenta 1a and 1b separately.
-In Experiment 1a, the first session (items 1-60, N = 204, left most sub-figure) showed results similar to the pooled results presented in Panel C.
-In the second session of Experiment 1a (items 61-120, N = 160) a smaller semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session.
+Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b.
+In Experiment 1a, the first session (items 1-60, N = 204, left sub-figure) showed results similar to the pooled results presented in Panel C.
+In the second session of Experiment 1a (items 61-120, N = 160, right sub-figure) a smaller semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session.
 Additionally, the second session in Experiment 1b showed shorter reading times compared to the first session.
@@
 
-Panel D also shows that Experiment 1b with items 1-60 (N = 350) showed a pattern consistent with   semantic interference in all regions.
-Experiment 1b with items 61-120 (N = 220, right most sub-figure) showed a pattern consistent with semantic interference in the regions surrounding the critical verb, but not at the verb itself.
+Panel E shows that Experiment 1b with items 1-60 (N = 350, left sub-figure) showed a pattern consistent with semantic interference in all regions.
+Experiment 1b with items 61-120 (N = 220, right sub-figure) showed a pattern consistent with semantic interference in the regions surrounding the critical verb, but not at the verb itself.
 At the verb, reading times were longest when both syntactic and semantic interference were high.
@@
 
-Was there any evidence for the effects estimated above.
+Was there any evidence for the effects estimated above?
 The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}.
@@
 Evidence for the interaction was only found in the critical region.
-Here, the evidence for the interaction was strong (BF$_{10}$ = 40) when the prior on the parameter was N+(0, 0.01), i.e., a small a priori effect size ranging from  [0, 8] ms; the evidence for the interaction was much weaker if the prior is N+(0,0.1), i.e., a larger a priori effect size ranging from [0, 81] ms.
+Here, the evidence for the interaction was strong (BF$_{10}$ = 40) when the prior on the parameter was N+(0, 0.01), i.e., a small a priori effect size ranging from  [0, 8] ms; the evidence for the interaction was much weaker (BF$_{10}$ = 1.8) when the prior was N+(0,0.1), i.e., a larger a priori effect size ranging from [0, 81] ms.
 In other words, only if we assume a priori that the effect size for the interaction is relatively small, then there is strong evidence for this effect.
@@
 
-At the distractor (which occurred well before the pre-critical region), animate distractors (high semantic interference) were read slower than inanimate distractors (low semantic interference, see Figure \ref{fig:whole_sentence} A and B).
+At the distractor (which occurred well before the pre-critical region), high semantic interference (animate distractors) lead to longer reading times than low semantic interference (inanimate distractors, see Figure \ref{fig:whole_sentence} A and B).
 One might argue that this pattern could be due to differences in word length.
 The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9).
-It seems unlikely that this small difference in word length caused such a large and long-lasting effects, but in order to be conservative we do not try to interpret this difference.
+It seems unlikely that this small difference in word length caused such large and long-lasting effects, but in order to be conservative we do not try to interpret this difference.
 
 At the pre-critical region, there was clear evidence for semantic interference.
-As this cannot be due to retrieval interfernce, the most likely explanation is encoding interference \citep{Oberauer_Kliegl_2006}.
+As this cannot be due to retrieval interference, the most likely explanation is encoding interference \citep{Oberauer_Kliegl_2006}.
 The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, kush_etal_2015, gordon02}.
@@
 Therefore, the sentence structure confound is also not a good explanation for the effect in our study.
-Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming the verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region.
+Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region.
 However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence} A and B).
-The predictive-processing explanation would incorrectly predict that the effect begins at the second adverb, which immediately precedes the verb.
+The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase.
 Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference as the most likely explanation of the effects found prior to the critical verb in our data.
@@
 SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous} in the appendix).
-In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by  \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
+This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous} in the appendix):  In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by  \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
 However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.

### 2023-10-17 paper/main.tex
minor edits until end of manuscript

@@
   \cmidrule{2-4}
-  \multirow{4}{1cm}{slope} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{3}{*}[upper bound: 0]& [-0.2, - 0]\\
+  \multirow{4}{1cm}{slope} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{4}{*}[upper bound: 0]& [-0.2, 0]\\
   & Normal(0, 0.5)& & [-1, 0]\\
@@
 We chose slope priors which assume a wide range of effect sizes from $-0.2$ to $-11$\,$\mu V$.
-However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$; this is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \textcite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.
+However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$.
+This assessment is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \textcite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.
 
@@
 Figure \ref{fig:erp_all} shows the 
-grand average ERPs elicited by the critical verb for all conditions are presented.
-Table \ref{tab:eeg_mod} shows the estimates from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu$V.
-The 95\% credible intervals of syntactic interference, semantic interference and their interaction were consistent with effects for all priors used.
-The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (mean: -0.31\,$\mu$V, 95\% CrI = [-0.49, -0.13]\,$\mu$V, for the posteriors see Figure \ref{fig:eeg_posteriors}).
+grand average ERPs elicited by the critical verb for all conditions.
+Table \ref{tab:eeg_mod} shows the estimates from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu V$.
+The 95\% credible intervals for the fixed effects are consistent with effects of syntactic interference, semantic interference and their interaction.
+The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (mean: -0.31\,$\mu V$, 95\% CrI = [-0.49, -0.13]\,$\mu V$, for the posteriors see Figure \ref{fig:eeg_posteriors}).
 The estimate of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{images/N_103_Cz_crit.png}
+    \includegraphics[width=\textwidth]{images/N_103_Cz_crit.png}
 \end{figure}
@@
 The timing and topography of this effect suggested that it was an N400 effect, thus the N400 was modulated by semantic interference.
-In contrast, \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in the N400 spatio-temporal window.
+In contrast, Figure \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in the N400 spatio-temporal window.
 Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (c) does also not suggest that there was a syntactic interference effect in any other time window.
@@
 Syntactic interference itself did not affect accuracy.
-There was a numerical suggestion of an interaction: indicated that the difference in accuracy between low and high semantic interference conditions was larger when syntactic interference was high vs.\ when it was low (17 vs.\ 8).
+There was a numerical suggestion of an interaction: the difference in accuracy between low and high semantic interference conditions was larger when syntactic interference was high vs.\ when it was low (17 vs.\ 8 \% difference).
 
@@
 The evidence against an interaction under all priors was unequivocal.
-.
 
@@
 for both the syntactic and semantic manipulations.
-Their model assumed that three cues were used at the verb:  \{$\pm$ subject, $\pm$ same clause, $\pm$ animacy\}.
+Their model assumed that three cues were used at the verb:  \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}.
 The upper part of Figure \ref{fig:model_predictions} reproduces the \textcite{mertzen}   simulations.
@@
 We speculated that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
-Specifically, the relevant syntactic cues that might be used at the verb is not  \{$\pm$ subject, $\pm$ same clause\} but rather a composite cue  \{$\pm$ subject-in-same-clause\} .
-In other words, in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are   \{$\pm$ subject-in-same-clause, $\pm$ animate\}.
+Specifically, the relevant syntactic cues that might be used at the verb are not  \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} but rather a composite cue \{$\pm$ subject-in-same-clause\}.
+In other words, in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
 The results of the prior predictive simulations are shown in
@@
 This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
-One limitation here is that the prior predictive distribution of our post-hoc  model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR studywas [4, 12] ms at the critical region).
+One limitation here is that the prior predictive distribution of our post-hoc  model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
 This overestimation has to do with the fact that the prior on the F parameter is relatively spread out; it would be  possible to fit the parameter to the SPR data.
@@
 The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor.
-Therefore, the \{$\pm$ subject-in-same-clause \} cue instead of the \{± grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
-
-Thus, one theoretical assumption that would be consistent with the observed SPR and EEG data is that the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval; however, the parser uses the syntactic cue while keeping track of the syntactic location of the subject.
+Therefore, the \{$\pm$ subject-in-same-clause\} cue instead of the \{$\pm$ grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
+
+Thus, one theoretical assumption that would be consistent with the observed SPR and ERP data is that the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval.
+However, the parser uses the syntactic cue while keeping track of the syntactic location of the subject.
 This is of course a speculative claim that would need to be further investigated empirically.
@@
     \centering
-    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the Lewis \& Vasishth (2005) model as implemented in R by Engelmann et al.\ (2019).
-The upper panel shows the predictions using the three retrieval cues used in Mertzen et al.\ 2022, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause, $\pm$ animate\}.
+    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the \textcite{Lewis2005} model as implemented in R by \textcite{engelmann_etal_2019}.
+The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
     The implemented model is available with the code and data provided with this paper.}
@@
 This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
-The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is by assuming that the parser intelligently searches for a subject that is within the same clause, and uses the animacy cue without reference to the clause that a noun appears in.
-A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \cite[e.g.,][]{Sturt2003,dillon2013}.
+The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser intelligently searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
+A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013}.
 
@@
 In the ERP results, we observed an effect at the critical verb, which could be attributed exclusively to the retrieval process hypothesized by cue-based retrieval accounts.
-Having said that, we cannot rule out the possibility that--due to the auto-paced nature of the ERP experiment--any encoding interference that started at or after the distractor is mixed in with the cost of retrieval interference.
+Having said that, we cannot rule out the possibility that -- due to the auto-paced nature of the ERP experiment -- any encoding interference that started at or after the distractor is mixed in with the cost of retrieval interference.
 Indeed, recent modeling work on agreement attraction in reading \parencite{Yadavetal2022} has shown conclusively that some kind of feature distortion as well as cue-based retrieval are needed to explain the benchmark data-sets that exist so far on agreement attraction in reading.
-If the  \textcite{Yadavetal2022} extends to interference effects in general,
-it is reasonable to assume that the effect observed at the verb in the ERP study is driven by both encoding and retrieval interference.
+If the  \textcite{Yadavetal2022} account extends to interference effects in general,
+it is reasonable to assume that the semantic interference effect observed at the verb in the present ERP study is driven by both encoding and retrieval interference.
 However, we cannot resolve the relative roles of these latent processes in the present study.
@@
 
-The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{± grammatical subject\}.
+The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
   However, the semantic interference effect we found can be easily reconciled with previous work in  sentence processing.
@@
 The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands.
-It is reasonable to assume that the participants might have--at least occasionally--adopted a good-enough processing mode when faced with high task demands \parencite{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \parencite{MalsburgVasishth2013}, or both.
+It is reasonable to assume that the participants might have -- at least occasionally -- adopted a good-enough processing mode when faced with high task demands \parencite{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \parencite{MalsburgVasishth2013}, or both.
 So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution.
@@
 However, this conclusion would be obviously wrong.
-First, as we showed above, the retrieval model suggests that--at least in our data--the subject cue is used, but differently than previously assumed: the relevant cue may be subject-in-same-clause.
+First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}.
 Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite[e.g.,][]{ALV2019,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \parencite[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and
@@
 These differences in the time course are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading).
-Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that the most likely explanation for the observed data is that the parser uses the syntactic cue  \{± subject-in-same-clause\}  to identify the correct target for retrieval.
-A broader implication is that the parser may be able to use hierarchical structure in the input to intelligently target the correct syntactic dependent from memory, leading to no interference from a distractor that is in subject position within an embedded clause.
+Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that the most likely explanation for the observed data is that the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+A broader implication is that the parser may be able to use hierarchical structure in the input to intelligently target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that is in subject position within an embedded clause.
 
@@
 Although the N400 is primarily associated with semantic processing, if one follows the memory retrieval account it is reasonable to assume that retrieval of all aspects of a memory entry will modulate the N400 amplitude.
-The potential complications during memory retrieval due to interference should not differ dependent on the specific cue used for retrieval, i.e., \{±grammatical subject\} or \{±animate\}.
+The potential complications during memory retrieval due to interference should not differ dependent on the specific cue used for retrieval, i.e., \{$\pm$ grammatical subject\} or \{$\pm$ animate\}.
 Moreover, previous ERP studies investigating interference have found negativities in the typical N400 time window (300 - 500 ms post-stimulus onset), although the scalp topographies were not typical for the N400, being either broad \citep{schoknecht2022} or left anterior \citep{vasishth_drenhaus_2011}.

### 2023-10-17 paper/main.tex
tidied up headline structure

@@
 
+\subsection{Methods}
 \subsubsection{Participants}
@@
 
+\subsection{Results}
 \subsubsection{Comprehension question accuracy (Experiments 1a and 1b combined)}
@@
 
-\section{Discussion}
+\subsection{Discussion}
 We presented a relatively large-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
@@
 
-\subsection{The effects in the pre-critical region}
+\subsubsection{The effects in the pre-critical region}
 
@@
 
-\subsection{Comparison with previous findings}
+\subsubsection{Comparison with previous findings}
 The results of the present study stand in contrast to previous findings \citep{vandyke07, mertzen}.
@@
 
+\subsection{Results}
 \subsubsection{Comprehension question accuracy}

### 2023-10-17 paper/main.tex
added dois

@@
 
-\appendix
-
-\section{Previous work on syntactic and semantic interference in reading}
-
-But are both syntactic and semantic cues used during memory retrieval?
+%\appendix
+
+%\section{Previous work on syntactic and semantic interference in reading}
+
+%Are both syntactic and semantic cues used during memory retrieval?
 \citet{vandyke07} and \citet{mertzen} investigated the effects of syntactic and semantic interference during long-distance subject-verb dependencies (see Example \ref{ex:1_vandyke07}) in a series of self-paced reading and eye-tracking while reading experiments.
@@
 Figure \ref{fig:previous} gives an overview of the previous findings.
-See the appendix for a detailed discussion of the previous results.
-
-One possibly problematic aspect of the studies from \citet{vandyke07} and \citet{mertzen} is that effects were found in pre-critical regions.
+
+%One possibly problematic aspect of the studies from \citet{vandyke07} and \citet{mertzen} is that effects were found in pre-critical regions.
 \citet{mertzen} suggested that the pre-critical effects could be due to structural confounds, encoding interference, preview effects during natural reading in the eye-tracking experiments or anticipatory retrieval at the pre-critical word because participants could predict that the verb would follow.
@@
 
-The previous estimates had relatively wide 95\,\% confidence/credible intervals in some of the experiments, which is probably due to the relatively small sample sizes.
+%The previous estimates had relatively wide 95\,\% confidence/credible intervals in some of the experiments, which is probably due to the relatively small sample sizes.
 \citet{vandyke07} had a typical number of participants and items for psycholinguistic experiments (35 - 40 participants per experiment, 36 - 48 items per experiment).
@@
 
-\begin{figure}[H]
-    \caption{Overview of previous findings of syntactic and semantic interference and their interaction by \citet{vandyke07} (three experiments in English with 35, 36 or 40 participants) and \citet{mertzen} (one experiment in German with 121 participants, one experiment in English with 61 participants).
+%\begin{figure}[H]
+%    \caption{Overview of previous findings of syntactic and semantic interference and their interaction by \citet{vandyke07} (three experiments in English with 35, 36 or 40 participants) and \citet{mertzen} (one experiment in German with 121 participants, one experiment in English with 61 participants).
 Experiments used either Self-paced Reading (SPR) or Eye-tracking methodology (Eye-Tracking measures: FPRT = first-pass reading time, RPD = regression-path duration, TFT = total fixation time).
@@
 }
-    \label{fig:previous}
-    \centering
-    \includegraphics[width=0.95\textwidth]{images/previous_findings_allregions_long.png}
-    \figurenote{The estimates for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) data were derived form the published paper.
+%    \label{fig:previous}
+%    \centering
+%    \includegraphics[width=0.95\textwidth]{images/previous_findings_allregions_long.png}
+%    \figurenote{The estimates for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) data were derived form the published paper.
 For non-significant effects, where no statistics were reported, we used a t-value of 1.95 to compute the confidence intervals.
 \citet{mertzen} provided their estimates and credible intervals.}
-\end{figure}
-
-The results of the previous studies raise two questions: 1) Why were semantic interference effects found less consistently than syntactic interference effects?
+%\end{figure}
+
+%The results of the previous studies raise two questions: 1) Why were semantic interference effects found less consistently than syntactic interference effects?
 2) Why were effects of retrieval interference observed in the pre-critical region?
@@
 
-In the previous studies, surprising effects were not only found in the pre-critical region, but also in the sentence-final region: Specifically, \citet{vandyke07} found semantic interference effects at the sentence-final region.
+%In the previous studies, surprising effects were not only found in the pre-critical region, but also in the sentence-final region: Specifically, \citet{vandyke07} found semantic interference effects at the sentence-final region.
 She proposed that the effect was delayed to this late region because ``it simply takes longer for individuals to compute the semantic association between the distractor and the verb'' \citep[][p.
@@
 
-Although the N400 is primarily associated with semantic processing, if one follows the memory retrieval account it is reasonable to assume that retrieval of all aspects of a memory entry will modulate the N400 amplitude.
+%Although the N400 is primarily associated with semantic processing, if one follows the memory retrieval account it is reasonable to assume that retrieval of all aspects of a memory entry will modulate the N400 amplitude.
 The potential complications during memory retrieval due to interference should not differ dependent on the specific cue used for retrieval, i.e., \{$\pm$ grammatical subject\} or \{$\pm$ animate\}.
@@
 
-\section{I don't know why this is here}
-In their \citeyear{vandyke07} paper, \citeauthor{vandyke07} reported a series of reading studies (one self-paced reading experiment and two eye-tracking experiments).
+%\section{I don't know why this is here}
+%In their \citeyear{vandyke07} paper, \citeauthor{vandyke07} reported a series of reading studies (one self-paced reading experiment and two eye-tracking experiments).
 For the eye-tracking experiments, \citet{vandyke07} reported three reading time measures: First-pass reading times (FPRT), regression-path durations (RPD) and total fixaton times (TFT).
@@
 
-\citeauthor{mertzen}'s (\citeyear{mertzen}) estimates at the pre-critical region showed slowdowns for high vs.\ low  syntactic and semantic interference in English and German in RPD and TFT.
+%\citeauthor{mertzen}'s (\citeyear{mertzen}) estimates at the pre-critical region showed slowdowns for high vs.\ low  syntactic and semantic interference in English and German in RPD and TFT.
 In the reading times of the critical verb, \citeauthor{mertzen}'s (\citeyear{mertzen}) estimates were consistent with syntactic interference effects in all measures in English and German.

### 2023-10-17 paper/main.tex
fixed bib

@@
 SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous} in the appendix):  In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
+This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design: %(see Figure \ref{fig:previous} in the appendix):
+In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
 However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.

### 2023-12-15 paper/main.tex
fixed bib
fixed typo
minor changes up until exp2

--- 2023-10-17 paper/main.tex
@@
 
-\title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?\newline
-%Semantic (but no syntactic) similarity-based interference
+% Which clause are you from? Evidence for interference effects from SPR and ERPs using German
+\title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?
+\newline
 Evidence from self-paced reading and event-related potentials using German}
@@
 \begin{itemize}
-\item Self-paced reading (SPR) and event-related potentials (ERP) experiments on interference
+\item Self-paced reading and event-related potentials experiments on interference
+\item Largest studies on interference during subject-verb integration (N=774, N=103)
 \item Bayes factors show evidence for semantic, but against syntactic interference
-\item Modeling shows that the parser tracks hierarchical information, attenuating syntactic interference.
+\item Modeling shows that tracking of syntactic structure attenuates syntactic interference
 % Wasting highlights on low information content
@@
 
-\keywords{Cue-based Retrieval, Interference, Syntax, Semantics, Replication, Self-paced Reading, Event-related Potentials, Computational Modeling}
+\keywords{Cue-based Retrieval, Syntactic Interference, Semantic Interference, Self-paced Reading, Event-related Potentials, Computational Modeling}
 
@@
 Another example is the recent heated debate on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
-Second, although it is clear and uncontroversial \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
+Second, although it has been established \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
 For example, \textcite{dillon2013} argued, following \textcite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built.
@@
 Reading studies have been the dominant method used to investigate retrieval interference \parencite[see the review in][]{jaeger_etal_2017}.
-We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading as well as EEG, which to our knowledge has never been used to investigate interference during subject-verb dependency resolution.
+We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading as well as electroencephalography (EEG), which to our knowledge has never been used to investigate interference during subject-verb dependency resolution.
 
@@
     \begin{xlist}   
-    \ex {High syntactic interference with and high / low semantic interference:}\label{ex:hisyn} 
+    \ex {High syntactic interference with high / low semantic interference:}\label{ex:hisyn} 
     \gll Die  Nachbarin glaubte,	dass	der Witwer,  der  erzählt hatte, dass der \textit{Einbrecher} / \textit{Verlust}  schrecklich war, abends regelmäßig \textbf{trank}, um zu vergessen.\\ 
@@
 
-The SPR study, hereafter Experiment 1a, was initially designed so that participants  (N=160) read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120).
+The SPR study, hereafter Experiment 1a, was initially designed so that participants read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120).
 The variation between participants in the gap is due to the fact that we could not control when the participants would do the second session; however, the mean and median gaps were  about 11 days, with the first and third quartiles being 9 and 12 days.
 
-As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: the effects observed in the first session disappeared.
+As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: Reading times decreased overall.
 The adaptation effect is theoretically interesting per se \parencite{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
-For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=614) were either shown items 1-60 or items 61-120 (but not both).
+For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both).
 In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
@@
     \centering
-    \includegraphics[width=0.85\textwidth]{images/pandora_project_structure_figure.pdf}
+    \includegraphics[width=\textwidth]{images/pandora_project_structure_figure.pdf}
 \end{figure}
@@
 The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms.
-These priors are based on the observed effect sizes and their uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}.
+These priors are based on the observed sign and sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}.
 
@@
 Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b.
-In Experiment 1a, the first session (items 1-60, N = 204, left sub-figure) showed results similar to the pooled results presented in Panel C.
-In the second session of Experiment 1a (items 61-120, N = 160, right sub-figure) a smaller semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session.
+Panel D shows the results from Experiment 1a.
+The first session of Experiment 1a (items 1-60, N = 204, left sub-figure) showed results similar to the pooled results presented in Panel C.
+In the second session of Experiment 1a (items 61-120, N = 160, right sub-figure) a semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session.
 Additionally, the second session in Experiment 1b showed shorter reading times compared to the first session.
-As mentioned earlier, this attenuation in reading times and differences between conditions is likely due to  adaptation to the task in the second session.
+As mentioned earlier, this attenuation in reading times and differences between conditions are likely due to  adaptation to the SPR task in the second session.
 
@@
 \subsection{Discussion}
-We presented a relatively large-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
+We presented the to date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
 Reading times at the critical verb showed clear evidence for semantic interference, but evidence against syntactic interference.
@@
 As this cannot be due to retrieval interference, the most likely explanation is encoding interference \citep{Oberauer_Kliegl_2006}.
-The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, kush_etal_2015, gordon02}.
+The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
 
@@
 The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase.
-Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference as the most likely explanation of the effects found prior to the critical verb in our data.
+Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data.
 Crucially this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in the critical region as well.
 The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence.
-It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory, or a combination of the two kinds of interference.
+It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference.
+Due to this unclear status of the effects at the critical verb, we refrain from interpreting the interaction which was observed in this region.
 
@@
 SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other self-paced reading experiment using this design (see Figure \ref{fig:previous} in the appendix):  In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
+This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design: %(see Figure \ref{fig:previous} in the appendix):
+In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
 However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.
@@
 The experiment was conducted in two experimental sessions for practical reasons (each of the sessions lasted approximately two hours).
-Sessions were separated by one to eight weeks for each participant.
+Sessions were separated by one to eight weeks for each participant (the median gap was 2 weeks, the first and third quartiles being 1 and 3 weeks).
 The procedure of both sessions was identical.
@@
     \midrule
-  Intercept & Normal(2, 5)& &\\
+  Intercept & Normal(2, 5)& & [-8, 12]\\
   \cmidrule{2-4}
@@
 
-\section{Data availability statement}
-
-The data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.
+%\section{Data availability statement}
+
+%The data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.
 

### 2023-12-18 paper/main.tex
updated prior predictive modeling

@@
 
-Although our principal focus was on the processing of the verb (\textit{trank}, `drank'), and the immediately following spill-over region, the pre-critical region was also of interest because previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in this region (we return to this point in the discussion sections).
+Our principal focus was on the processing of the verb (\textit{trank}, `drank').
+The reading times of the immediately following spill-over region and of the pre-critical region were also of interest.
+Previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in the pre-critical region (we return to this point in the discussion sections).
 A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another (\textit{erzählt hatte, trank}, literally `told, drank').
@@
 
-Thus, our focus in the present study is on the processing difficulty observed at the pre-critical, critical (verb), and post-critical regions.
+Thus, our focus in the present study is on the processing difficulty observed at the critical verb and the surrounding regions.
 
@@
 It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference.
-Due to this unclear status of the effects at the critical verb, we refrain from interpreting the interaction which was observed in this region.
+Because of the unclear status of the effects at the critical verb, we refrain from interpreting the interaction which was only observed in this region.
 
@@
 Priors on the slope parameters were truncated to an upper bound of zero, i.e., only negative values.
-This corresponds to more negative brain responses to high vs.\ low interference conditions in line with the literature.
+This corresponds to more negative brain responses to high vs.\ low interference conditions in line with the literature \citep{lee_garnsey, martinetal2014, schoknecht2022, vasishth_drenhaus_2011}.
 The standard deviations of the slope priors were varied in order to conduct a Bayes factor sensitivity analysis.
@@
 Additionally, the results suggest an interaction, i.e., the difference between high and low semantic interference conditions was larger when syntactic interference was high vs.\ when it was low.
+Because the accuracies are not of primary interest, we did not carry out Bayes factors analyses for these.
 
@@
 grand average ERPs elicited by the critical verb for all conditions.
-Table \ref{tab:eeg_mod} shows the estimates from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu V$.
+Table \ref{tab:eeg_mod} shows the estimates for the brain activity in the spatio-temporal N400 window from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu V$.
 The 95\% credible intervals for the fixed effects are consistent with effects of syntactic interference, semantic interference and their interaction.
@@
 
-\begin{figure}
+\begin{figure}[H]
     \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the time window 300 to 500 ms post critical verb onset at centro-parietal electrode sites.}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{images/BF_plot_N103_cp_300_500.jpg}
+    \includegraphics[width=0.8\textwidth]{paper/images/BF_plot_N103_cp_300_500.jpg}
 \end{figure}
@@
 \subsection{Discussion}
-We presented a relatively large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency completion.
+We presented a large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency completion.
 To our knowledge, this is the first ERP experiment on this classic interference design, which has only been investigated using reading studies so far \parencite{mertzen,vandyke07}.
@@
 Our finding that retrieval interference modulates the N400 amplitude supports the N400 as a marker of memory retrieval \citep{kutas&federmeier_2000, kutas_federmeier2011, brouwer2017_n4_p6, lau2008_n400}.
+The more complex memory retrieval under high semantic interference resulted in a more negative N400.
 
@@
 Their model assumed that three cues were used at the verb: \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}.
-The upper part of Figure \ref{fig:model_predictions} reproduces the \textcite{mertzen}   simulations.
-These are prior predictive distributions of the effects of interest, by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter F and then generating data from the model while holding all other parameters in the model at fixed values.
+The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations.
+\ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter F and then generating data from the model while holding all other parameters in the model at fixed values.
 The F parameter scales the activation to a retrieval time.
 The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020}.
+Figure \ref{fig:model_predictions_effects} shows the effects of interest computed based on the condition means shown in \ref{fig:model_predictions_rts}.
+
+\begin{figure}[htbp]
+    \centering
+    \caption{Prior predictions from the \textcite{Lewis2005} model as implemented in R by \textcite{engelmann_etal_2019}. 
+    The implemented models are available with the code and data provided with this paper.}
+    \label{fig:model_predictions}
+    \centering
+         \begin{subfigure}[t]{\textwidth}
+         \centering
+        \caption{Prior predicted by-condition retrieval time means.
+The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+Diamonds represent the means (numbers to the right) and blue dots show each simulated data point.}
+         \label{fig:model_predictions_rts}
+         \includegraphics[width=\textwidth]{paper/images/PriorPredictedMeans_interACTversions.png}
+     \end{subfigure}
+
+     \begin{subfigure}[t]{\textwidth}
+         \centering
+        \caption{Prior predicted effect sizes, i.e., retrieval time differences, for syntactic and semantic interference and their interaction.
+The upper panels show the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panels show the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.}
+         \label{fig:model_predictions_effects}
+         \includegraphics[width=\textwidth]{paper/images/PriorPredicted_interACTversions.png}
+     \end{subfigure}
+\end{figure}
 
 Although the semantic interference effect estimates from the SPR study fall within the predicted range from the model, the syntactic interference effect is quite divergent: as discussed earlier, the SPR data show evidence against syntactic interference, but the model predicts a syntactic interference effect of approximately the same magnitude as semantic interference.
-Given that the EEG data also speak against syntactic interference, it is plausible that, at least in this experimental design and language, syntactic interference plays no role.
+Given the evidence against syntactic interference in the SPR data and the flip between evidence for and evidence against it depending on the priors in the ERP data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
 
@@
 The results of the prior predictive simulations are shown in
-the lower part of Figure \ref{fig:model_predictions}.
+the lower parts of Figure \ref{fig:model_predictions} a and b.
 This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
@@
 However, the \{± same clause\} model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms).
-
-\begin{figure}[H]
-    \centering
-    \caption{Prior predicted reading times for syntactic and semantic interference and their interaction from the \textcite{Lewis2005} model as implemented in R by \textcite{engelmann_etal_2019}.
-The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-    The implemented model is available with the code and data provided with this paper.}
-    \label{fig:model_predictions}\includegraphics[width=\textwidth]{images/PriorPredicted_interACTversions_2vs3.png}
-\end{figure}
 

### 2023-12-18 paper/main.tex
added SPR posterior predictions

@@
 The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations.
-\ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter F and then generating data from the model while holding all other parameters in the model at fixed values.
-The F parameter scales the activation to a retrieval time.
+\ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter \textit{F} and then generating data from the model while holding all other parameters in the model at fixed values.
+The \textit{F} parameter scales the activation to a retrieval time.
 The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020}.
@@
 In other words, in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-The results of the prior predictive simulations are shown in
-the lower parts of Figure \ref{fig:model_predictions} a and b.
+The results of the prior predictive simulations are shown in the lower parts of Figure \ref{fig:model_predictions} a and b.
 This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
 One limitation here is that the prior predictive distribution of our post-hoc  model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
-This overestimation has to do with the fact that the prior on the F parameter is relatively spread out; it would be  possible to fit the parameter to the SPR data.
+This overestimation has to do with the fact that the prior on the \textit{F} parameter is relatively spread out.
+
+To reduce the overestimation of the effect size of the semantic interference effect, we generated posterior predictive distributions for the effects of interest by fitting the \textit{F} parameter to the SPR data.
+New data was simulated using only the values of F which produced a semantic interference effect in the range of the semantic interference effect in our SPR data (mean $\pm \,2\times$SE).
+This applied to only 355 out of 20,000 sampled values of \textit{F}.
+We fitted \textit{F} only to the semantic interference effect (and not the syntactic interference effect or interaction) to avoid overfitting.
+Figure \ref{fig:model_predictions_posterior} shows the posterior predictive distributions of the effects of interest in comparison to the prior predictive distribution and our SPR data.
+The posterior predictions match the syntactic and semantic interference effects in our data, but underestimate the interaction.
+
+\begin{figure}[h]
+    \caption{Prior and posterior predicted effect sizes, i.e., retrieval time differences, for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+The data with 95 \% credible intervals is shown in red.}
+    \label{fig:model_predictions_posterior}
+    \centering
+    \includegraphics[width=\textwidth]{paper/images/priorpred_vs_postpred_vs_data.png}
+\end{figure}
 

### 2023-12-20 paper/main.tex
ERP modeling

@@
 To reduce the overestimation of the effect size of the semantic interference effect, we generated posterior predictive distributions for the effects of interest by fitting the \textit{F} parameter to the SPR data.
-New data was simulated using only the values of F which produced a semantic interference effect in the range of the semantic interference effect in our SPR data (mean $\pm \,2\times$SE).
+New data was simulated using only the values of \textit{F} which produced a semantic interference effect in the range of the semantic interference effect in our SPR data (mean $\pm \,2\times$SE).
 This applied to only 355 out of 20,000 sampled values of \textit{F}.
@@
 
-\begin{figure}[h]
-    \caption{Prior and posterior predicted effect sizes, i.e., retrieval time differences, for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-The data with 95 \% credible intervals is shown in red.}
-    \label{fig:model_predictions_posterior}
+\begin{figure}[H]
+    \caption{Prior and posterior predicted effect sizes for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+The data with 95 \% credible intervals are shown in red.}
+\begin{subfigure}[t]{\textwidth}
+\caption{On the millisecond scale, i.e., for reading times.}\label{fig:model_predictions_posterior}
     \centering
     \includegraphics[width=\textwidth]{paper/images/priorpred_vs_postpred_vs_data.png}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+         \caption{On the microvolt scale, i.e., for ERPs.}\label{fig:model_predictions_ERP}
+         \includegraphics[width=\textwidth]{paper/images/ERP_priorpred_vs_postpred_vs_data.png}
+     \end{subfigure}
 \end{figure}
@@
 The scaling factor \textit{F} links the activation of items in memory to retrieval times \citep{Lewis2005}.
-But this scaling factor could be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
-The general effect pattern (no syntactic interference effect, but a semantic interference effect) would not be affected by an adjustment of the scaling factor.
+But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
+The here reported modeling is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been applied to ERP data.
+To generate prior predictions on the microvolt scale, we sampled the free scaling parameter \textit{F} from a truncated normal prior distribution ($Normal_+(0.01, 0.005)$) with the lower and upper bounds at 0.003 and 0.04, respectively.
+For posterior predictions, we fitted \textit{F} only to the semantic interference effect to avoid overfitting, like we did when modeling the reading data.
+For the ERP modeling, 3,146 out of 20,000 sampled values of \textit{F} lead to simulated effects in the range of the observed effect and were thus used for posterior predictions.
+The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}.
+The prior and posterior predictions of the model overlap with the effects of interest in the data.
+The general effect pattern (no syntactic interference effect, but a semantic interference effect) is not affected by the adjustment of the scaling factor.
 Therefore, the \{$\pm$ subject-in-same-clause\} cue instead of the \{$\pm$ grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.

### 2023-12-20 paper/main.tex
updated ERP plots

@@
 This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
-The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser intelligently searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
+The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
 A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013}.
@@
 In contrast, in the ERP experiment, we used rapid serial visual presentation (RSVP), which presents stimuli at a fixed rate that is beyond the control of the participant.
-Therefore, they must adapt to the set rate and process the stimuli as fast as the rate requires.
+Therefore, they must adapt to the set rate and process the stimuli at the speed that the rate requires.
 Here, it is noteworthy that the presentation duration of 500\,ms per word and 400\,ms between words which was used in the present ERP experiment is slower than usual natural reading speed.

### 2023-12-20 paper/main.tex
minor text edits

@@
 Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role in sentence processing.
-However, this conclusion would be obviously wrong.
+However, this conclusion would be problematic.
 First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}.
@@
 Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that the most likely explanation for the observed data is that the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
-A broader implication is that the parser may be able to use hierarchical structure in the input to intelligently target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that is in subject position within an embedded clause.
+A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that is in subject position within an embedded clause.
 

### 2023-12-20 code_data/LV05_modeling/retrieval_times/generate_posteriorpredictions.R
Add files via upload

@@ -47,9 +47,9 @@ synsemlower <- 14-(2*4) # mean effect - 2*SE
 synsemupper <- 14+(2*4) # mean effect + 2*SE
 
 # posterior latency factor (we base this only on the semantic interference result)
-SEM<-semlower<=simdat_2cues_sameclause$MESem & simdat_2cues_sameclause$MESem<=semupper
+SEM<-semlower<=simdat_2cues_sameclause_complete$MESem & simdat_2cues_sameclause_complete$MESem<=semupper
 table(SEM)
-posterior_lf_sem <-simdat_2cues_sameclause[SEM,]$lfs
+posterior_lf_sem <-simdat_2cues_sameclause_complete[SEM,]$lfs
 
 length(posterior_lf_sem)
 quantile(posterior_lf_sem,probs=c(0.025,0.975))
@@ -106,6 +106,7 @@ postpreds <-data.frame(lfs=lf,syntactic=maineffectSyn,semantic=maineffectSem, in
 postpreds_longer <- pivot_longer(postpreds, cols = c("syntactic", "semantic", "interaction"),
                                   names_to = "effect", values_to = "rts")
 postpreds_longer$effect <- factor(postpreds_longer$effect, levels = c("syntactic", "semantic", "interaction"))
+postpreds_longer$pred <- "posterior predictive"
 
 # plot
 ggplot(postpreds_longer, aes(x=rts)) +
@@ -118,57 +119,37 @@ ggplot(postpreds_longer, aes(x=rts)) +
 ggsave("posteriors.png", width=6, height=3, dpi=600)
 
 
-#### compare posteriors and data ####
-## compute mean and 95% CI of each posterior
-## combine all posteriors in a list
-vectors <- list(maineffectSyn, maineffectSem, interactionSynSem)
-
-## length of list vectors
-vec_len <- length(vectors)
-## initialize empty data.frame to store values
-names <- c("mean", "lower", "upper")
-posteriors <- as.data.frame(matrix(numeric(),nrow = vec_len, ncol = 3))
-colnames(posteriors) <- names
-
-for (i in 1:vec_len){
-  posteriors[i,1] <- mean(vectors[[i]])
-  posteriors[i,2] <- unname(quantile(vectors[[i]], probs=0.025))
-  posteriors[i,3] <- unname(quantile(vectors[[i]], probs=0.975))
-}
-
-## add info regarding experiment, measure and cue (interference) to posteriors dataframe
-real_model <- rep(c("model 2cues.sameclause"), times=3)
[…]

### 2023-12-20 code_data/LV05_modeling/retrieval_times/generate_priorpredictions_2models.R

@@ -1,0 +1,233 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+#### THREE CUE MODEL FROM MERTZEN ET AL. (2023) https://osf.io/a7cg2/?view_only= ####
+
+# Load model 
+source("interACT-threecues.R")
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
+  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(0.33,0.33,0.33))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate 20,000 simulations from model:
+nsims<-20000
+
+## Initial hyperparameters for latency factor
+## from Methodsx paper:
+a<-4
+b<-6
+lfs<-rbeta(nsims,a,b)
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+# empty df for saving means
+generated_means <- data.frame(matrix(ncol=9,nrow=nsims))
+colnames(generated_means) <- c("sample.id","RT_LoSynLoSem","RT_LoSynHiSem","RT_HiSynLoSem","RT_HiSynHiSem", "Acc_LoSynLoSem","Acc_LoSynHiSem","Acc_HiSynLoSem","Acc_HiSynHiSem")
+
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=5000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  # save means and acc
[…]

### 2023-12-20 code_data/LV05_modeling/retrieval_times/generate_priorpredictions.R
Deleted via the Open Science Framework

@@ -1,265 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-#### THREE CUE MODEL FROM MERTZEN ET AL. (2023) https://osf.io/a7cg2/?view_only= ####
-
-# Load model 
-source("interACT-threecues.R")
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
-  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(0.33,0.33,0.33))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Generate 20,000 simulations from model:
-nsims<-20000
-
-## Initial hyperparameters for latency factor
-## from Methodsx paper:
-a<-4
-b<-6
-lfs<-rbeta(nsims,a,b)
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
-  lf<<- lfs[i]  
-  sims <- create_param_matrix(model_4cond, iterations=5000)
-  results <- run(sims)
-  condMeans <- compute_cond_means(results)
-  ## compute main effect of Syntactic manipulation
-  ## HiSyn-LoSyn:
-  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
-    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
-  ## save the estimate:
-  maineffectSyn[i]<-meSyn
-  ## compute main effect of Semantic manipulation
-  ## HiSem-LoSem:
[…]

### 2023-12-20 code_data/LV05_modeling/brain_response/generate_posteriorpredictions.R
Deleted via the Open Science Framework

@@ -1,175 +1,0 @@
-library(truncnorm)
-library(tidyverse)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-# Load model
-source("interACT.R")
-reset_params()
-
-# TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Load simulated data
-load("simdat_2cues_sameclause.Rda")
-
-# Run ABC with rejection sampling
-# results from Schoknecht & Vasishth (N400 elicited by critical verb): 
-# syntactic mean effect -0.2 muV, CrI [-0.4,0.0] --> Normal(-0.2,0.1) --> SE: 0.1
-# semantic mean effect -0.3 muV, CrI [-0.6,-0.1] --> Normal(-0.3,0.15) --> SE: 0.15
-# interaction mean effect -0.1 muV, CrI [-0.4, 0.0] --> Normal(-0.1,0.1) --> SE: 0.1
-
-# I'm changing the sign of the effect here (sem: -0.3 --> 0.3)
-# to match the sign of the latency factor
-# --> a positive effect represents are more negative N400 
-
-synlower <- 0.2-(2*0.1) # mean effect - 2*SE
-synupper <- 0.2+(2*0.1) # mean effect + 2*SE
-
-semlower <- 0.3-(2*0.15) # mean effect - 2*SE
-semupper <- 0.3+(2*0.15) # mean effect + 2*SE
-
-synsemlower <- 0.1-(2*0.1) # mean effect - 2*SE
-synsemupper <- 0.1+(2*0.1) # mean effect + 2*SE
-
-# posterior latency factor (we base this only on the semantic interference result)
[…]

### 2023-12-20 code_data/LV05_modeling/brain_response/generate_priorpredictions.R
Deleted via the Open Science Framework

@@ -1,121 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-
-#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-# Load model 
-source("interACT.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Generate 20,000 simulations from model:
-nsims<-20000
-
-## Initial hyperparameters for latency factor
-## Visualization of priors
-
-#Truncated Normal
-latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
-                                                    a=0.003, b=0.04, mean=0.01, sd=0.005))
-
-prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
-  geom_histogram(aes(y=..density..),
-                 position="identity",fill="gray",
-                 binwidth=0.001, center = 0.0005)+
-  theme_bw()+
-  theme(strip.text.x = element_text(size = 16, colour = "black", angle = 0))+
-  xlab("latency factor")+
-  ggtitle("Prior on scaling factor")
-prior_lf
-ggsave("LF_Prior.png", width=5, height=3, dpi=600)
-
-lfs<-rtruncnorm(nsims, a=0.003, b=0.04, mean=0.01, sd=0.005)
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
[…]

### 2023-12-20 code_data/LV05_modeling/brain_response/generate_predictions.R
Added via the Open Science Framework

@@ -1,0 +1,248 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+# Load model 
+source("interACT.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+#### Generate prior predictions #####
+
+# Generate 20,000 simulations from model:
+nsims<-20000
+
+## Initial hyperparameters for latency factor
+## Visualization of priors
+
+#Truncated Normal
+latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
+                                                    a=0.003, b=0.04, mean=0.01, sd=0.005))
+
+prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
+  geom_histogram(aes(y=..density..),
+                 position="identity",fill="gray",
+                 binwidth=0.001, center = 0.0005)+
+  theme_bw()+
+  theme(strip.text.x = element_text(size = 16, colour = "black", angle = 0))+
+  xlab("latency factor")+
+  ggtitle("Prior on scaling factor")
+prior_lf
+ggsave("LF_Prior.png", width=5, height=3, dpi=600)
+
+lfs<-rtruncnorm(nsims, a=0.003, b=0.04, mean=0.01, sd=0.005)
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
[…]

### 2023-12-22 paper/main.tex

@@
     \centering
-    \includegraphics[width=0.8\textwidth]{paper/images/BF_plot_N103_cp_300_500.jpg}
+    \includegraphics[width=0.8\textwidth]{images/BF_plot_N103_cp_300_500.jpg}
 \end{figure}
@@
          \label{fig:model_predictions_rts}
-         \includegraphics[width=\textwidth]{paper/images/PriorPredictedMeans_interACTversions.png}
+         \includegraphics[width=\textwidth]{images/PriorPredictedMeans_interACTversions.png}
      \end{subfigure}
@@
          \label{fig:model_predictions_effects}
-         \includegraphics[width=\textwidth]{paper/images/PriorPredicted_interACTversions.png}
+         \includegraphics[width=\textwidth]{images/PriorPredicted_interACTversions.png}
      \end{subfigure}
@@
     \centering
-    \includegraphics[width=\textwidth]{paper/images/priorpred_vs_postpred_vs_data.png}
+    \includegraphics[width=\textwidth]{images/priorpred_vs_postpred_vs_data.png}
     \end{subfigure}
@@
          \caption{On the microvolt scale, i.e., for ERPs.}\label{fig:model_predictions_ERP}
-         \includegraphics[width=\textwidth]{paper/images/ERP_priorpred_vs_postpred_vs_data.png}
+         \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
      \end{subfigure}

### 2023-12-22 paper/main.tex
fixed typo

--- 2023-12-20 paper/main.tex
@@
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
-We turn to this point next by discussing the quantiative predictions of the \textcite{Lewis2005} model.
+We turn to this point next by discussing the quantitative predictions of the \textcite{Lewis2005} model.
 
@@
 The \textit{F} parameter scales the activation to a retrieval time.
-The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020}.
+The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020,vasishth2020_abc}.
 Figure \ref{fig:model_predictions_effects} shows the effects of interest computed based on the condition means shown in \ref{fig:model_predictions_rts}.
@@
          \centering
-        \caption{Prior predicted by-condition retrieval time means.
+        \caption{Prior predicted by-condition retrieval times.
 The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.

### 2023-12-25 paper/main.tex
fixed sentence
minor edits

@@
 The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations.
-\ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter \textit{F} and then generating data from the model while holding all other parameters in the model at fixed values.
-The \textit{F} parameter scales the activation to a retrieval time.
+\ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter $F$ and then generating data from the model while holding all other parameters in the model at fixed values.
+The $F$ parameter scales the activation to a retrieval time.
 The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020,vasishth2020_abc}.
-Figure \ref{fig:model_predictions_effects} shows the effects of interest computed based on the condition means shown in \ref{fig:model_predictions_rts}.
+Figure \ref{fig:model_predictions_effects} shows the prior predictive distributions of the effects of interest.
 
@@
 Specifically, the relevant syntactic cues that might be used at the verb are not  \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} but rather a composite cue \{$\pm$ subject-in-same-clause\}.
-In other words, in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} proposal, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval, we evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+This proposal is in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
+
+We evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
 The results of the prior predictive simulations are shown in the lower parts of Figure \ref{fig:model_predictions} a and b.
 This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
-One limitation here is that the prior predictive distribution of our post-hoc  model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
-This overestimation has to do with the fact that the prior on the \textit{F} parameter is relatively spread out.
-
-To reduce the overestimation of the effect size of the semantic interference effect, we generated posterior predictive distributions for the effects of interest by fitting the \textit{F} parameter to the SPR data.
-New data was simulated using only the values of \textit{F} which produced a semantic interference effect in the range of the semantic interference effect in our SPR data (mean $\pm \,2\times$SE).
-This applied to only 355 out of 20,000 sampled values of \textit{F}.
-We fitted \textit{F} only to the semantic interference effect (and not the syntactic interference effect or interaction) to avoid overfitting.
+One limitation here is that the prior predictive distribution of our model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
+This overestimation has to do with the fact that the prior on the $F$ parameter ($Beta(a=4,b=6)$) is relatively spread out.
+This leads to a relatively wide range of predicted reading times.
+
+It is informative to fit the model to the data and then examine the posterior predictive distribution of the reading times for each effect. 
+We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the SPR data.
+This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the  scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ (close being defined as $d \pm \,2\times SE$ ms) in the SPR data.
+For details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
+
+%New data was simulated using only the values of $F$ which produced a semantic interference effect in the range of the semantic interference effect in our SPR data.
+This applied to only 355 out of 20,000 sampled values of $F$.
+To avoid overfitting, we fitted the scaling parameter $F$ only to the semantic interference effect (and not the syntactic interference effect or interaction). 
+We obtained only 355 samples of $F$ (from 20,000 iterations of the ABC algorithm); this is not surprising given the relatively wide prior on $F$.
 Figure \ref{fig:model_predictions_posterior} shows the posterior predictive distributions of the effects of interest in comparison to the prior predictive distribution and our SPR data.
@@
 The \citet{Lewis2005} model is generally used to predict reading times.
-The scaling factor \textit{F} links the activation of items in memory to retrieval times \citep{Lewis2005}.
+The scaling factor $F$ links the activation of items in memory to retrieval times \citep{Lewis2005}.
 But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
-The here reported modeling is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been applied to ERP data.
-To generate prior predictions on the microvolt scale, we sampled the free scaling parameter \textit{F} from a truncated normal prior distribution ($Normal_+(0.01, 0.005)$) with the lower and upper bounds at 0.003 and 0.04, respectively.
-For posterior predictions, we fitted \textit{F} only to the semantic interference effect to avoid overfitting, like we did when modeling the reading data.
-For the ERP modeling, 3,146 out of 20,000 sampled values of \textit{F} lead to simulated effects in the range of the observed effect and were thus used for posterior predictions.
+The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+To generate prior predictions on the microvolt scale, we sampled the free scaling parameter $F$ from a truncated normal prior distribution ($Normal_+(0.01, 0.005)$) with the lower and upper bounds at 0.003 and 0.04, respectively.
+As in the reading time modeling, to avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects.
+For the ERP modeling, 3,146 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions.
 The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}.
 The prior and posterior predictions of the model overlap with the effects of interest in the data.
-The general effect pattern (no syntactic interference effect, but a semantic interference effect) is not affected by the adjustment of the scaling factor.
-Therefore, the \{$\pm$ subject-in-same-clause\} cue instead of the \{$\pm$ grammatical subject\} cue explanation can account for the estimates of the present ERP study as well.
-
-Thus, one theoretical assumption that would be consistent with the observed SPR and ERP data is that the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval.
-However, the parser uses the syntactic cue while keeping track of the syntactic location of the subject.
+The general effect pattern (no syntactic interference effect, but a semantic interference effect) is similar in both the prior and posterior predictive distributions.
+Therefore, the assumption that the relevant syntactic cue is \{$\pm$ subject-in-same-clause\} cue, rather than  the \{$\pm$ grammatical subject\} cue originally assumed in previous modeling work, seems reasonable for the present ERP study as well.
+
+Thus, the theoretical assumption that is consistent with the observed SPR and ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
 This is of course a speculative claim that would need to be further investigated empirically.
@@
 To answer this question, it is necessary to look at the ERPs elicited by words earlier in the sentence, This is an unusual step in ERP research but makes sense as an exploratory step in the present context.
+
 Figure \ref{fig:erp_precrit} shows the ERPs for all conditions at the critical word and the two pre-critical adverbs which were identical across conditions.
@@
 An exploratory visual inspection of the ERPs elicited by the distractor shows several differences between conditions, but no semantic interference effect in the N400 spatio-temporal window.
+
 We believe that the differences in the ERPs elicited by the distractor in the different conditions are caused by two factors.
@@
 This is also apparent from the large difference between high and low syntactic interference conditions before distractor onset.
-Second, the distractors were not identical across conditions, therefore the ERPs in Figure \ref{fig:erp_distractor} are elicited by different words in the high vs.\ low semantic interference conditions.
-In summary, this exploratory analysis suggests that the distractors did induce different brain responses across conditions, but these could be due to factors other than encoding intererence.
+Second, the distractors were not identical across conditions; therefore, the ERPs in Figure \ref{fig:erp_distractor} are elicited by different words in the high vs.\ low semantic interference conditions.
+In summary, this exploratory analysis suggests that the distractors did induce different brain responses across conditions, but these could be due to factors other than encoding interference.
 More importantly, these effects did not linger throughout the rest of the sentence (see Figure \ref{fig:erp_precrit}).
@@
 Here, it is noteworthy that the presentation duration of 500\,ms per word and 400\,ms between words which was used in the present ERP experiment is slower than usual natural reading speed.
-However, while in SPR the slowdown due to encoding and maintaining the distractor in memory overshadowed potential retrieval effects, the procedure of the ERP experiment might have prevented that.
+However, while in SPR the slowdown due to encoding and maintaining the distractor in memory overshadowed potential retrieval effects, the procedure of the ERP experiment might have prevented encoding interference from co-occurring with retrieval interference effects.
 In the ERP results, we observed an effect at the critical verb, which could be attributed exclusively to the retrieval process hypothesized by cue-based retrieval accounts.
 Having said that, we cannot rule out the possibility that -- due to the auto-paced nature of the ERP experiment -- any encoding interference that started at or after the distractor is mixed in with the cost of retrieval interference.
-Indeed, recent modeling work on agreement attraction in reading \parencite{Yadavetal2022} has shown conclusively that some kind of feature distortion as well as cue-based retrieval are needed to explain the benchmark data-sets that exist so far on agreement attraction in reading.
+Indeed, recent modeling work on agreement attraction in reading \parencite{Yadavetal2022} has shown that some kind of feature distortion as well as cue-based retrieval are needed to explain the data-sets that were publicly available at the time that the modeling was done.
 If the  \textcite{Yadavetal2022} account extends to interference effects in general,
@@
 Animacy is an obvious choice for such a heuristic.
-Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject leads to the correct analysis with high probability (not just within the experiment context, but in everyday language use).
+Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject leads to the correct analysis with high probability (not just within the experiment context, but also in everyday language use).
 Finally, our findings are also consistent with the possibility that language processing relies predominantly on semantic associations to form (probabilistic)
@@
 the importance of word order \parencite[e.g.,][]{meng2000mode}.
+
 There are of course important examples of the parser ignoring syntactic information.
-Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, the agreement attraction phenomenon across different languages \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}, and local coherence effects \parencite{taboretal04}.
+Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, local coherence effects \parencite{taboretal04}, and the agreement attraction phenomenon across different languages \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}.
 However, there are languages (such as German and Dutch) in which the grammaticality illusion does not occur \parencite{VSLK11,FrankEtAl2015}, and a large-sample study \parencite{lcpaape2023} presents evidence against the syntactic local coherence effect that was reported by \textcite{taboretal04}.
 Also, there seems to be at least one language that has rich case marking (Czech) that does not show any evidence of agreement attraction \parencite{chromy2023number}.
-Thus, it seems that, cross-linguistically, the role of syntax in building incremental structure and in completing dependencies remains as important as ever.
+Thus, it seems that, cross-linguistically, the role of syntax in building incremental structure and in completing dependencies remains an open question.
 
@@
 This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb.
-These different time courses in the two experiments are likely to be due to different types of interference: Encoding interference in the SPR experiment and retrieval (and possibly also encoding) interference in the ERP experiment.
-These differences in the time course are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading).
-Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that the most likely explanation for the observed data is that the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+These different time courses in the two experiments are likely to be due to a mix of encoding and retrieval interference.
+%different types of interference: In SPR, the effects at the pre-critical region are likely to be due to encoding interference; and at the critical region, the effects could be a mix of encoding and retrieval interference.
+In the ERP experiment, the patterns seen could be due to encoding interference but they are   in the pre-critical regions in the SPR experiment, and retrieval (and possibly also encoding) interference in the ERP experiment.
+The observed differences in the time course in the SPR and ERP studies are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading).
+Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that--at least given the present data--the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
 A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that is in subject position within an embedded clause.

### 2023-12-25 paper/main.tex
minor edit to abstract and removed comments

@@
 This difficulty in identifying the correct dependent is called similarity-based interference.
-% We are explaining what we did, not summarizing every detail in the paper:
-%Previously, behavioral studies have consistently found syntactic interference at the verb, but semantic interference was only found in later regions or not at all. 
-% What we did:
 We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07,mertzen}, to investigate interference due to syntactic and semantic cues in German.
-% Not needed for an abstract, plus sounds boring:
-%The aim of our experiments was to gain precise estimates of the effects of syntactic and semantic interference to inform theory building.
-% What we found:
 A Bayes factor analysis showed that 
@@
 In the self-paced reading experiment, interference due to the semantic cues caused a reading time slow-down starting at the distractor.
-% The reader does not even know what that is:
-%The best explanation for this effect is encoding interference. 
 The event-related potentials showed a more negative N400 amplitude at the critical verb, i.e., the retrieval site, for high vs.\ low semantic interference.
-% Irrelevant in an abstract:
- %and no such effects at preceding words. 
- % No new information in this sentence, a waste:
- %Therefore, the event-related potentials were modulated by retrieval interference.
-% Said this more compactly already:
-%Bayes factors for both reading times and event-related potentials provided evidence for the semantic interference effect. 
 Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to syntactic cues.
@@
 
-In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
-
-% This cannot be the main point of this paper! 
-%The present findings, which are in line with various sentence processing accounts, highlight the importance of semantic cues during sentence processing.
-
-% Highlights must be in a separate file (not in the paper itself)
+In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, in our data, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
+
 \clearpage
@@
 \item Modeling shows that tracking of syntactic structure attenuates syntactic interference
-% Wasting highlights on low information content
-%\item SPR: Encoding interference lead to longer reading times starting at a distractor
-%\item ERP: Time course of effect clearly indicated an effect of retrieval interference
 \end{itemize}

### 2023-12-27 paper/main.tex
minor edits

@@
 It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval. As an example, consider 
-\ref{ex:1_vandyke07}.
+(\ref{ex:1_vandyke07}).
 
@@
 Here, we consider two kinds of cues used in retrieval: semantic and  syntactic cues.
-In a sentence like \ref{ex:1_vandyke07}, the retrieval cues generated at the verb are standardly assumed to include at least  the semantic cue \{+animate\} and the syntactic cue \{+grammatical subject\}.
+In a sentence like (\ref{ex:1_vandyke07}), the retrieval cues generated at the verb are standardly assumed \parencite[e.g.,][]{vandyke07} to include at least  the semantic cue \{+animate\} and the syntactic cue \{+grammatical subject\}.
 These cues pick out the target noun \textit{the resident} as the subject.
@@
 However, if the distractor were \textit{neighbor} instead of \textit{warehouse}, then it would match the \{+animate\} cue at the verb.
-This overlap in cues between the target noun and the distractor is called cue-overload in cue-based retrieval accounts, and is predicted to trigger semantic interference, which expresses itself as greater processing time at the verb compared to the baseline condition \ref{ex:1_vandyke07}.
+This overlap in cues between the target noun and the distractor is called cue-overload in cue-based retrieval accounts, and is predicted to trigger semantic interference, which expresses itself as greater processing time at the verb compared to the baseline condition (\ref{ex:1_vandyke07}).
 
 By contrast, syntactic interference is expected to be triggered if the distractor matches the \{+grammatical subject\} cue.
-This situation can occur when the relative clause  in \ref{ex:1_vandyke07} that modifies \textit{the resident} were to be changed to \textit{who said that the warehouse was dangerous}.
+This situation can occur when the relative clause  in (\ref{ex:1_vandyke07}) that modifies \textit{the resident} were to be changed to \textit{who said that the warehouse was dangerous}.
 Now, the distractor matches the \{+grammatical subject\} cue.
@@
 By contrast, the original small-sample study \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory.
-Another example is the recent heated debate on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
+Another example is the recent heated debate \parencite{nieuwland_etal_2018} on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
 Second, although it has been established \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
@@
 ][]{jaeger_etal_2020,yadav2021individual}.
-This was a surprising discovery because a priori, cue-based retrieval accounts do not allow for exceptions to retrieval interference in particular syntactic dependency types.
-More broadly, an important implication of the work initiated by \textcite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
+This was a surprising discovery because the default assumption in cue-based retrieval accounts was that all cues were equally weighted, regardless of the dependency type; this had the consequence that retrieval interference was expected whenever a cue-overload configuration was present in a sentence.
+More broadly, an important insight in the work of \textcite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
+We build on this idea in the present paper.
 
@@
 As the same materials were used for all our experiments, we describe them here once before we describe the specifics of the experiments, separately.
-The studies had 120 items with four conditions in a Latin square design shown in \ref{ex:materials}.
+The studies had 120 items with four conditions in a Latin square design shown in (\ref{ex:materials}).
 The items were constructed following \citeauthor{vandyke07}'s (\citeyear{vandyke07}) items and partially re-used  \citeauthor{mertzen}'s (\citeyear{mertzen}) items.
@@
 
-The comprehension of the sentences in \ref{ex:materials} requires a memory retrieval operation at the verb \textit{trank} (`drank') to retrieve its non-adjacent subject \textit{Witwer} (`widower').
+The comprehension of the sentences in (\ref{ex:materials}) requires a memory retrieval operation at the verb \textit{trank} (`drank') to retrieve its non-adjacent subject \textit{Witwer} (`widower').
 During this retrieval, the distractor \textit{Einbrecher} (`burglar') / \textit{Verlust} (`loss') might cause interference and impede comprehension.
 
-Syntactic interference is manipulated via the syntactic status of the distractor, which is either subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}).
-Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in any subject position causes more syntactic interference than a distractor inside a prepositional phrase.
+Syntactic interference is manipulated via the syntactic status of the distractor, which is either the subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}) and hence a non-subject.
+Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase.
 
@@
 For each of the 120 item quadruplets, four simple sentences with subject-verb-object word order were constructed, so that each of the nouns in the item once functioned as the subject of the critical verb.
-The norming materials corresponding to the item depicted in \ref{ex:materials} are shown in \ref{ex:materials_norming}.
+The norming materials corresponding to the item depicted in (\ref{ex:materials}) are shown in (\ref{ex:materials_norming}).
 
@@
 Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}.
-The critical word for analyses was the verb \textit{trank} `drank' in \ref{ex:materials}, constituting the retrieval site.
+The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site.
 Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed.
@@
 The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms.
-These priors are based on the observed sign and sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}.
-
-The Bayes factor is analogous to the frequentist ANOVA or likelihood ratio test: it compares the (marginal) likelihood of a full model to that of a baseline null model; this is often written as $BF_{10}$, with the subscripts indicating that marginal likelihood of the full model is in the numerator and that of the null model is in the denominator.
+These priors are based on the observed sign and sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}, and on meta-analyses relating to interference effects  \parencite{jaeger_etal_2017}.
+For the syntactic and semantic interference effects, we use directional priors so that any evidence for the effect of interest from Bayes factors can be interpreted as evidence of replication success \parencite{verhagen2014bayesian}.
+
+The Bayes factor is analogous to the frequentist ANOVA or likelihood ratio test: it compares the marginal likelihood of a full model to that of a baseline null model; this is often written as $BF_{10}$, with the subscripts indicating that marginal likelihood of the full model is in the numerator and that of the null model is in the denominator.
 One major advantage of the Bayes factor over the frequentist analogue is that it takes the uncertainty of the parameters into account.
@@
 Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is in principle possible to find evidence both for or against an effect.
-This stands in contrast to the frequentist ANOVA, which is designed to only furnish evidence against the null.
+This stands in contrast to the frequentist ANOVA which, in its standard usage, is designed to only furnish evidence against the null.
 

### 2023-12-27 code_data/LV05_modeling/brain_response/generate_predictions.R
updated prior

@@ -38,8 +38,15 @@ nsims<-20000
 ## Visualization of priors
 
 #Truncated Normal
-latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
+## the first two are not reasonable, they cause bimodality.
+#latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
                                                     a=0.003, b=0.04, mean=0.01, sd=0.005))
+
+#latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
+                                                    a=0, mean=0.01, sd=0.005))
+
+latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
+                                                    a=0.005, b=0.04, mean=0.02, sd=0.005))
 
 prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
   geom_histogram(aes(y=..density..),
@@ -52,7 +59,15 @@ prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
 prior_lf
 ggsave("LF_Prior.png", width=5, height=3, dpi=600)
 
-lfs<-rtruncnorm(nsims, a=0.003, b=0.04, mean=0.01, sd=0.005)
+## not reasonable, cause bimodality:
+#lfs<-rtruncnorm(nsims, a=0.003, b=0.04, mean=0.01, sd=0.005)
+#lfs<-rtruncnorm(nsims, a=0, mean=0.01, sd=0.005)
+
+lfs<-rtruncnorm(nsims, a=0.005, b=0.04, mean=0.02, sd=0.005)
+
+## check:
+hist(lfs,freq=FALSE)
+
 maineffectSyn<-rep(NA,nsims)
 maineffectSem<-rep(NA,nsims)
 interactionSynSem<-rep(NA,nsims)

### 2023-12-27 code_data/LV05_modeling/brain_response/generate_predictions.R
fine-tuned figures

@@ -38,13 +38,6 @@ nsims<-20000
 ## Visualization of priors
 
 #Truncated Normal
-## the first two are not reasonable, they cause bimodality.
-#latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
-                                                    a=0.003, b=0.04, mean=0.01, sd=0.005))
-
-#latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
-                                                    a=0, mean=0.01, sd=0.005))
-
 latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
                                                     a=0.005, b=0.04, mean=0.02, sd=0.005))
 
@@ -226,7 +219,7 @@ postpreds_longer$effsize_neg <- -1 * postpreds_longer$effsize
 ggplot(postpreds_longer, aes(x=effsize_neg)) +
   geom_density(aes(),fill="darkgrey")+
   ggh4x::facet_grid2(. ~ effect, scale="free_y", independent = "y")+
-  #xlim(c(-0.2,2))+
+  xlim(c(-0.7,0.1))+
   xlab("Predicted effect (ms)")+
   geom_vline(xintercept=0, linetype="dashed")+
   theme_bw(base_size = 12)
@@ -239,7 +232,7 @@ prior_predictions$pred <- "prior predictions"
 postpreds_longer$pred <- "posterior predictions"
 preds <- rbind(postpreds_longer, prior_predictions)
 
-# add data from Schoknecht & Vasishth SPR experiment
+# add data from Schoknecht & Vasishth ERP experiment
 dat <- data.frame(pred   = rep("data",3),
                   effect    = c("syntactic", "semantic", "interaction"), 
                   mean  = c(-0.2,-0.3,-0.1),
@@ -260,4 +253,4 @@ ggplot(data=preds, x=effsize_neg) +
   guides(fill = guide_legend(reverse=TRUE))+
   theme(legend.title=element_blank(), legend.position = c(0.83, 0.8))
 
-ggsave("ERP_priorpred_vs_postpred_vs_data.png", width=7, height=3, dpi=600)
+ggsave("ERP_priorpred_vs_postpred_vs_data.png", width=7.5, height=3, dpi=600)

### 2023-12-29 paper/main.tex
minor edits

@@
 \subsection{Discussion}
-We presented the to date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
+We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
 Reading times at the critical verb showed clear evidence for semantic interference, but evidence against syntactic interference.
-The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occured even before the retrieval event could begin.
+The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occurred even before the retrieval event could begin.
+The effects in the pre-critical region are discussed next.
 
@@
 The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9).
-It seems unlikely that this small difference in word length caused such large and long-lasting effects, but in order to be conservative we do not try to interpret this difference.
+It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.
 
@@
 Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
-The observed difference between the high and low semantic interference conditions is independent of the syntactic manipulation.
-Therefore, the sentence structure confound is also not a good explanation for the effect in our study.
+
+Regarding the possibility that sentence structure confounds as a possible explanation for the effects in the pre-critical region, this also seems implausible because the observed difference between the high and low semantic interference conditions is independent of the syntactic manipulation; it is the syntactic manipulation that has the confound.
+Therefore, the sentence structure confound is also not a good explanation for the semantic interference effect observed in our study.
+
 Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region.
@@
 Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data.
-Crucially this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in the critical region as well.
+Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in the critical region as well.
 The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence.
-It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference.
+It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \parencite{Yadavetal2022}.
 Because of the unclear status of the effects at the critical verb, we refrain from interpreting the interaction which was only observed in this region.
@@
 
-The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
+There is one further possible reason for the diverging effects in our studies compared to previous work: the theoretically-predicted effects we are investigating here may be too subtle to consistently be detectable across studies.
+In several other attempts to replicate previously observed findings, we have often found it impossible to replicate the originally claimed effects, despite spending four or more years collecting data to obtain sufficient evidence.
+For example, contrary to published claims \parencite[e.g.,][]{vandyke_mcelree06}, in a large cross-linguistic study, we found evidence against proactive interference from sentence-external distractors \parencite{mertzen_repVDM2006}.
+A second example is the well-known syntactic local coherence effect in English \parencite{taboretal04}; in a recent large-sample study, we have found evidence against syntactic local coherence \parencite{lcpaape2023}.
+A third example is the interaction of locality and surprisal effects reported in \textcite{levy&keller_2013}; we attempted and failed to replicate the key claims in seven experiments \parencite{vasishth2018_signficancefilter}.
+A fourth example is the processing of antecedent-reflexive dependencies; \textcite{dillon2013} argued that reflexives are exempt from interference, but a large-scale direct replication attempt  \parencite{jaeger_etal_2020} found evidence that reflexives do show interference effects \parencite[cf.][]{yadav2021individual}.
+Finally, we have also failed to replicate the surprising claim  \parencite{gibsonwu,hsiao03} that Chinese object relatives are easier to process than subject relatives \parencite{VasishthetalPLoSOne2013,VasishthetalPLoSOne2013}.
+It is in principle possible that all the originally claimed effects actually represent true effects but, due to the inherent variability of data, the effects may be impossible to detect consistently in replications.
+Independent of any power issues in all these different studies, it is possible that it is unreasonable to expect consistent replications of effects in behavioral data, especially when investigating subtle effects.
+Certain other effects in psycholinguistics, such as strong garden-path effects \parencite{frazierrayner82}, are easily and consistently replicable \parencite[e.g.,][]{huang2023surprisal}.
+By contrast, subtle manipulations such as interference effects and local coherence may not show up consistently in replication attempts.
+One strong possibility is that there are systematic and important individual differences between participants' behavior; an increased focus away from investigating average effects and toward individual differences may reveal important insights about the conditions under which subtle effects are replicable; we demonstrate this point for the antecedent-reflexive dependency in \parencite{yadav2021individual}.
+
+Nevertheless, the contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
 So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
-One possible explanation is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
+Apart from the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
 While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
-%This is also supported by the reduced uncertainty of the individual-level estimates in our data compared to \citeauthor{mertzen}'s (\citeyear{mertzen}). 
-We will address the absence of syntactic interference in our data in the general discussion.
+We will revisit the absence of syntactic interference in our data in the general discussion, where we investigate the question using computational modeling.
+
+We turn next to the event-related potentials experiment.
 
@@
 
-\begin{table}[]
+\begin{table}[!htbp]
     \caption{Relatively informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}.
@@
 This corresponds to more negative brain responses to high vs.\ low interference conditions in line with the literature \citep{lee_garnsey, martinetal2014, schoknecht2022, vasishth_drenhaus_2011}.
+The directional priors also serve to allow us to quantify the extent to which a predicted effect replicates; the Bayes factors can be interpreted as indicating the evidence supporting or not supporting replication success \citep{verhagen2014bayesian}.
     The standard deviations of the slope priors were varied in order to conduct a Bayes factor sensitivity analysis.
@@
 
-Models were run with four chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}.
-For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats}.
+Bayesian models were run with four chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}.
+For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats,schad_etal_2022_BF}.
 We chose slope priors which assume a wide range of effect sizes from $-0.2$ to $-11$\,$\mu V$.

### 2023-12-30 paper/main.tex
minor edits

@@
 The 95\% credible intervals for the fixed effects are consistent with effects of syntactic interference, semantic interference and their interaction.
-The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (mean: -0.31\,$\mu V$, 95\% CrI = [-0.49, -0.13]\,$\mu V$, for the posteriors see Figure \ref{fig:eeg_posteriors}).
+The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (mean: -0.31\,$\mu V$, 95\% CrI = [-0.49, -0.13]\,$\mu V$, for the posteriors, see Figure \ref{fig:eeg_posteriors}).
 The estimate of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).
@@
     \caption{Results from the Bayesian linear mixed model analyzing the averaged brain response from 12 centro-parietal electrodes in the time window 300 to 500 ms post critical verb onset.
-Baseline represents the averaged activity from the 200 ms pre-verb onset and was included in the model as a covariate instead of traditional baseline correction \citep[see][]{alday2019}.}
+The predictor baseline represents the averaged activity from the 200 ms pre-verb onset and was included in the model as a covariate instead of traditional baseline correction \citep[see][]{alday2019}.}
     \label{tab:eeg_mod}
@@
 Figure \ref{fig:erp_sem_syn} (b) shows that the topography of the effect is rather broad but with a concentration of more negative values at centro-parietal electrodes.
-The timing and topography of this effect suggested that it was an N400 effect, thus the N400 was modulated by semantic interference.
+The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference.
 In contrast, Figure \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in the N400 spatio-temporal window.
@@
 Comprehension accuracy was reduced for high vs.\ low semantic interference conditions.
-Syntactic interference itself did not affect accuracy.
+There was no indication that syntactic interference affected accuracy, although the sign of the effect was in the expected direction, but  amounting to a 1\% difference on the probability scale between the high and low syntactic interference conditions.
 There was a numerical suggestion of an interaction: the difference in accuracy between low and high semantic interference conditions was larger when syntactic interference was high vs.\ when it was low (17 vs.\ 8 \% difference).
@@
 This proposal is in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
+In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
 
@@
 This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
-One limitation here is that the prior predictive distribution of our model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
-This overestimation has to do with the fact that the prior on the $F$ parameter ($Beta(a=4,b=6)$) is relatively spread out.
+One limitation here is that the prior predictive distribution of our model generates a very widely distributed semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
+This broad prior predictive distribution has to do with the fact that the prior on the $F$ parameter ($Beta(a=4,b=6)$) is relatively spread out.
 This leads to a relatively wide range of predicted reading times.
+In future work, the posterior distribution, with range  [4, 12] ms, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.
 
@@
 We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the SPR data.
-This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the  scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ (close being defined as $d \pm \,2\times SE$ ms) in the SPR data.
-For details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
+This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the  scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ in the SPR data.
+`Close' is defined as $d \pm \,2\times SE$ ms; for details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
 
@@
 The \citet{Lewis2005} model is generally used to predict reading times.
-The scaling factor $F$ links the activation of items in memory to retrieval times \citep{Lewis2005}.
+The scaling factor $F$ links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R}.
 But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.

### 2023-12-30 paper/main.tex
removed commented out text

@@
 \documentclass[a4paper, man, floatsintext]{apa7}
-%dvipsnames
 
@@
 
-% Which clause are you from? Evidence for interference effects from SPR and ERPs using German
 \title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?
@@
 More generally, interference that is induced by such an overlap between retrieval cues and features of items in memory is called retrieval interference.
-%In grammatical sentences, where the retrieval cues match with the features of the target of the memory retrieval (e.g., the subject \textit{the resident} in Example \ref{ex:1_vandyke07}), interference from distractors is assumed to cause processing difficulty.
 
@@
 \section{Experiments 1a and 1b: SPR}
-%\subsection{Methods}
-
-%Initially, we planned to conduct SPR experiment in two sessions to gain data from all participants for all 120 items.
-Collecting data for 120 items in one session was not considered feasible in a web-based SPR experiment, as the experiment duration would have been close to one hour.
-So, the first batch of participants (160 out of 774) completed two experimental sessions.
-This approach was dropped after we realized, that participants adapted to the task quickly which led to a considerable reading time speed-up between sessions (see Figure \ref{fig:whole_sentence} D).
-After this realization, all further participants took part in a single session during which they read half of the 120 items (for a visualization of the project structure, see Figure \ref{fig:project_str}).
 
@@
 \subsubsection{Participants}
-%two sessions, session 1
-% 224 participants completed
-% 17 low acc
-% 2 inconsistent demographic info
-% 1 technical error
-% final 204
-
-% two sessions, session 2
-% 178 participants completed
-% 18 low acc
-% final 160
-
-% rep s1
-% 387 participants completed
-% 31 low acc
-% 6 inconsistent demographic info
-% final 350
-
-% rep s2
-% 279 participants completed
-% 51 low acc
-% 8 inconsistent demographic info
-% final 220
-
-% 224 + 18  + 387 + 279 = 908
 
@@
 Five participants replied with ``other.''
-%We conclude that the participants who were recruited via Prolific were slightly older and have a comparable or higher level of education than the typical undergraduate and graduate students who participate in lab-based studies.
 
@@
 
-%\bigskip
-
-%\begin{table}[ht]
-%    \centering
-%        \caption{Contrast coding.}
-%    \label{tab:contrasts}
-%    \begin{tabular}{cccc}
-%\toprule
-%condition & syntactic interference & semantic interference & interaction\\
-%\midrule
-%a)  LoSyn LoSem &  $-1$& $-1$ & $+1$ \\
-%b)  LoSyn HiSem & $-1$ & $+1$ & $-1$  \\
-%c)  HiSyn LoSem &  $+1$ &$-1$ & $-1$    \\ 
-%d)  HiSyn HiSem &  $+1$ &$+1$  & $+1$   \\
-%\bottomrule
-%    \end{tabular}
-%\end{table}
-
 \subsection{Results}
@@
 
-%This is probably due to encoding interference \citep{mertzen, jaeger_etal_2015_retrieval_encoding} and persisted in the following regions.
-
 Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but will be harder to interpret because of possible spill-over from the syntactic manipulation.
@@
 The other two conditions (high syntactic/low semantic and low syntactic/high semantic) lead to the fastest verb reading times in Experiment 1b with items 61-120.
-
-%These somewhat inconsistent results, especially in the critical region, across the sub-experiments and the random assignment of participants to the sub-experiments motivated us to gain estimates of the effects under investigation for the individual participants (see Figure \ref{fig:spr_individual}).
-
-%\begin{figure}[H]
- %   \caption{Individual estimates of syntactic and semantic interference and their interaction at the critical verb for all 774 participants in the self-paced reading experiment.
-Shown are shrunken estimates from a Bayesian hierarchical linear mixed model.
-Circles show individual’s mean effects, errors bars are 95\,\% credible intervals.
-The red dashed line shows an effect of zero (difference of zero milliseconds).
-Two solid red lines mark an effect size of 10 and 20 ms, respectively.}
-%    \label{fig:spr_individual}
- %   \centering
-  %  \includegraphics[width=\textwidth]{images/individual_diffs.jpg}
-%\end{figure}
-
-%The individual-level effects were derived from a Bayesian linear mixed model with the priors shown in Table \ref{tab:spr_priors} with the intermediate slope prior of Normal(0, 0.05) and random slopes for syntactic and semantic interference and their interaction but without correlations of random effects.
-
-%Figure \ref{fig:spr_individual} shows that at the critical verb, the 95\,\% credible intervals for the syntactic interference effect of all participants were centered around zero.
-In contrast, participants consistently displayed a small effect of semantic interference with narrow 95\,\% credible intervals.
-The individual-level estimates of the interaction were a bit larger but they were associated with more uncertainty (wider 95\,\% credible intervals).
-The individual-level results help to understand the population-level posteriors presented in Figure \ref{fig:spr_posteriors} and the population-level Bayes factor analysis (see Figure \ref{fig:spr_bfs}).
 
@@
 We turn next to the event-related potentials experiment.
-
-%We derived quantitative predictions from four versions of an adapted \citet{Lewis2005} model for the investigated 2 $\times$ 2 interference design \citep{mertzen}.
-The versions of the model differed in their cue-combinatorics (linear, non-linear) and cue-weighting (equal weighting, higher weighting of syntactic/structural cues).
-We described in our Predictions section that previous findings \citep{vandyke07, mertzen} were most compatible with linear cue-combinatorics with higher weighing of syntactic/structural cues.
-This was mainly driven by the lack of an interaction in the previous studies.
-In contrast, the results of the present study are only compatible with non-linear cue-combinatorics.
-As predicted by the non-linear cue-combinatorics model, only distractors which matched all retrieval cues induced interference; see the interaction / nested semantic interference effect.
-The predictions of non-linear cue-combinatorics models changed only slightly when the cue-weighting was varied (equal weighting or higher weighting of syntactic/structural cues).
-Therefore, we do not claim that our results provide evidence for one type of cue-weighting over the other \citep[for a detailed discussion of cue-weighting, see e.g.,][]{yadav2021individual}.
-
-%The present findings can inform the debate on cue-combinatorics in cue-based retrieval models as they provide evidence in favor of non-linear cue-weighting \citep{parker2019cue,vandyke07, raaijmakers_shiffrin_1981, hintzman1984minerva}.
-Non-linear cue-weighting is an alternative to linear cue-weighting; the latter assumes a simple additive process in cue combination.
-In contrast, non-linear cue-weighting denotes that cues are combined via multiplication.
-One mismatching cue multiplies all other cues, even if they match, with zero.
-Therefore, a distractor which mismatches at least one of the retrieval cues will be dismissed.
-The present study provides evidence that cues are non-linearly combined.
-Distractors which only partially matched retrieval cues did not induce interference.
-
-%The observed effect sizes are smaller than the predicted effect sizes, but that might just be due to the latency parameter in the models (the scaling parameter used to link activation levels to reading times).
-We used the latency parameter distribution which was previously used \citep[][]{jaeger_etal_2020, engelmann_etal_2019, vasishth2020_abc}.
-In future research, it might be adjusted to better match the smaller effect sizes found in large-scale self-paced reading studies of interference \citep[see][]{nicenboim}.
-
-%The observed interaction at the critical verb, i.e., a semantic interference effect only within the high syntactic interference conditions, was predicted by syntax-first models of sentence processing.
-But as described above no main effect of syntactic interference was observed.
-Such an effect would be a prerequisite to claim primacy or dominance of syntactic features.
-The observed interaction and the corresponding nested comparisons rather suggest that syntactic and semantic cues are used simultaneously to guide retrieval (see constraint-based sentence processing models, \cite[e.g.,][]{macdonald1994}).
 
@@
 
-%The to date rather small body of ERP studies on interference effects during sentence comprehension gives an indeterminate picture of the nature of interference effects in ERP measures.
-The present study adds a large data point to the body of literature which reported negativities or more specifically N400 effects for high vs.\ low interference conditions \citep{vasishth_drenhaus_2011, martinetal2014, schoknecht2022, lee_garnsey}.
-Therefore, it contributes to the debate of how interference effects manifest themselves electrophysiologically (interference effects have also been reported in differences in the P600, \textcite{Tanner_etal_2017, martinetal2014}).
-
 \section{Quantitative predictions of the Lewis and Vasishth (2005) model for the present design}
@@
 
-%New data was simulated using only the values of $F$ which produced a semantic interference effect in the range of the semantic interference effect in our SPR data.
-This applied to only 355 out of 20,000 sampled values of $F$.
 To avoid overfitting, we fitted the scaling parameter $F$ only to the semantic interference effect (and not the syntactic interference effect or interaction). 
@@
 
-%There is another point to make regarding the small size of the effects observed in the present study: The observed semantic effect in self-paced reading times (8 ms, 95\,\% CrI [4, 12]) only slightly overlaps with the quantitative predictions from the \citet{Lewis2005} model of cue-based retrieval for the \citet{vandyke07} design (95\,\% CrI [9, 49]\,ms, implementation of the model by \textcite{engelmann_etal_2019}, adapted to the current design with the retrieval cues \{± animate\} and \{± grammatical subject\}).
-The \citet{Lewis2005} model does not only overestimate the effect size of semantic interference, but it also predicts equally sized semantic and syntactic interference effects (see Figure \ref{fig:model_predictions}).
-The syntactic effect is absent in our data.
-
-%It would be implausible to argue that the absence of the syntactic interference effect suggests that the comprehender does not use syntactic information for subject-verb dependency resolution at all.
-We rather suggest that the \{± grammatical subject\} cue was not the right syntactic cue to investigate.
-In line with \citet{mertzen}, we suggest that a syntactic or structural cue along the lines of \{± same clause\} might be used during comprehension.
-Figure \ref{fig:model_predictions} presents the predictions of the \citet{Lewis2005} model for the \citet{vandyke07} design where additionally to the \{± animate\} cue either the canonical cue \{± grammatical subject\} or the alternative cue \{± same clause\} was implemented.
-The \{± same clause\} cue model does predict the syntactic interference effect to be centered around zero (95\,\% CrI [-4, 3]\,ms).
-This prediction is consistent with the results of the present study.
-However, the \{± same clause\} model still overestimates the semantic interference effect (95\,\% CrI [10, 67]\,ms).
-
 \section{General discussion}
@@
 A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013}.
-
-%Given the replicability / reproducibility crises in psycholinguistics \citep[see e.g., ][]{vasishth2018_signficancefilter,jaeger_etal_2020,nicenboim,}, and other (related) sciences \citep[see e.g., ][]{openscience2015_reproducibility,nicenboim_etal_2020,nieuwland_etal_2018}, the cross-methodological consistency is aston.
-It might be due to the very high sample sizes of the present experiments which granted the present study higher power than most comparable psycholinguistic studies.
-The higher power led to more precise estimates of the investigated effects with less uncertainty than in previous studies.
-%This also partially explains the comparatively smaller effect sizes in the present experiments than have been previously reported.
 
@@
 
-%Besides our considerations above regarding the possibility that another syntactic cue (e.g., \{± same clause\}) was used, our results highlight the importance of the semantic cue for subject-verb dependency resolution.
-
 The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
@@
 In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model.
-  %Our results may also be consisted with constraint-based models \citep[e.g.,][]{macdonald1994, mcrae1998modeling, trueswell1993}.
-Under constraint-based accounts, it is assumed that multiple constraints simultaneously and immediately influence processing.
-These accounts predict our ERP results: Compared to an inanimate distractor, an animate distractor -- regardless of its syntactic status -- would impede processing at a verb which requires an animate subject.
-The lack of a syntactic interference effect could indicate that the status `grammatical subject' was not used as a syntactic constraint.
-As discussed above, a plausible syntactic constraint is that a verb and its arguments need to be within the same clause.
  Our results could also be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
@@
 \section{Conclusion}
-%The present large-scale self-paced reading study examined the role of syntactic and semantic features during sentence comprehension.
-At a verb that initiated retrieval of its subject, high semantic interference led to longer reading times than low semantic interference, but only when syntactic interference was high as well.
-This result provided empirical support for the assumption that different types of cues are used simultaneously during memory retrieval in subject-verb dependency formation.
-Furthermore, it suggests that distractors must match all available retrieval cues to cause similarity-based retrieval interference.
-The results of the present study are similar to the predictions of cue-based retrieval models of sentence processing with non-linear combinatorics.
-
 This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials.
@@
 These different time courses in the two experiments are likely to be due to a mix of encoding and retrieval interference.
-%different types of interference: In SPR, the effects at the pre-critical region are likely to be due to encoding interference; and at the critical region, the effects could be a mix of encoding and retrieval interference.
-In the ERP experiment, the patterns seen could be due to encoding interference but they are   in the pre-critical regions in the SPR experiment, and retrieval (and possibly also encoding) interference in the ERP experiment.
+
 The observed differences in the time course in the SPR and ERP studies are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading).
@@
 
-%\appendix
-
-%\section{Previous work on syntactic and semantic interference in reading}
-
-%Are both syntactic and semantic cues used during memory retrieval?
-\citet{vandyke07} and \citet{mertzen} investigated the effects of syntactic and semantic interference during long-distance subject-verb dependencies (see Example \ref{ex:1_vandyke07}) in a series of self-paced reading and eye-tracking while reading experiments.
-Overall, the original study by \citet{vandyke07} and the replication by \citet{mertzen} showed relatively consistent results.
-They consistently found syntactic interference but only rarely semantic interference (only in some measures of some of the experiments and sometimes in regions outside of the critical region).
-These results are not in line with the predictions from the standard cue-based retrieval models \citep{Lewis2005, mcelree}: While both previous studies found effects of syntactic and semantic interference, they emerged in different regions -- sometimes even prior to the retrieval site.
-Figure \ref{fig:previous} gives an overview of the previous findings.
-
-%One possibly problematic aspect of the studies from \citet{vandyke07} and \citet{mertzen} is that effects were found in pre-critical regions.
-\citet{mertzen} suggested that the pre-critical effects could be due to structural confounds, encoding interference, preview effects during natural reading in the eye-tracking experiments or anticipatory retrieval at the pre-critical word because participants could predict that the verb would follow.
-It is unclear from the previous studies when exactly the pre-critical effects emerged, i.e., when the conditions started to differ.
-But it is important to know the timecourse of the pre-critical effects to determine their cause.
-If the structural confounds explanation is correct, then the pre-critical effects should appear only in relation to the syntactic manipulation.
-For the encoding interference explanation to be correct, it would be necessary that conditions start to differ at (or shortly after) the distractor.
-The preview and anticipatory retrieval explanations can only apply if the pre-critical effects are found only close to the critical verb.
-
-%The previous estimates had relatively wide 95\,\% confidence/credible intervals in some of the experiments, which is probably due to the relatively small sample sizes.
-\citet{vandyke07} had a typical number of participants and items for psycholinguistic experiments (35 - 40 participants per experiment, 36 - 48 items per experiment).
-\citet{mertzen} used the same number of items as \citet{vandyke07}, but larger participant sample sizes (English: N = 61, German N = 121).
-Large uncertainty raises the possibility of Type M errors, leading to misleading conclusions \citep{gelman_carlin}.
-
-%\begin{figure}[H]
-%    \caption{Overview of previous findings of syntactic and semantic interference and their interaction by \citet{vandyke07} (three experiments in English with 35, 36 or 40 participants) and \citet{mertzen} (one experiment in German with 121 participants, one experiment in English with 61 participants).
-Experiments used either Self-paced Reading (SPR) or Eye-tracking methodology (Eye-Tracking measures: FPRT = first-pass reading time, RPD = regression-path duration, TFT = total fixation time).
-Error bars are 95\% confidence intervals for \citet{vandyke07} and 95\% credible intervals for \citet{mertzen}.
-}
-%    \label{fig:previous}
-%    \centering
-%    \includegraphics[width=0.95\textwidth]{images/previous_findings_allregions_long.png}
-%    \figurenote{The estimates for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) data were derived form the published paper.
-For non-significant effects, where no statistics were reported, we used a t-value of 1.95 to compute the confidence intervals.
-\citet{mertzen} provided their estimates and credible intervals.}
-%\end{figure}
-
-%The results of the previous studies raise two questions: 1) Why were semantic interference effects found less consistently than syntactic interference effects?
-2) Why were effects of retrieval interference observed in the pre-critical region?
-To answer question 1, we decided to run large-sample studies revisiting \citeauthor{vandyke07}'s (\citeyear{vandyke07}) design to gain more precise estimates of the effects of syntactic and semantic interference.
-To answer question 2 -- or to disentangle the possible causes of the pre-critical effects -- we decided it is best to use methods which provide responses which are unambiguously related to the processing of single words.
-The methods of our choice being self-paced reading (SPR) and electroencephalography (EEG) or event-related potentials (ERPs), respectively, with typical rapid serial visual presentation (RSVP) of single words.
-Both these methods provide responses directly linked to the processing of single words and, unlike eye-tracking while reading, do not grant any preview of upcoming stimuli.
-To reduce the structural confound at the pre-critical region identified by \citet{mertzen}, we added an additional pre-pre-critical region.
-These changes in the experimental design between the present and previous studies will help to distinguish between the different explanations for the pre-critical effects put forward by \citet{mertzen}.
-
-%In the previous studies, surprising effects were not only found in the pre-critical region, but also in the sentence-final region: Specifically, \citet{vandyke07} found semantic interference effects at the sentence-final region.
-She proposed that the effect was delayed to this late region because ``it simply takes longer for individuals to compute the semantic association between the distractor and the verb'' \citep[][p.
-427]{vandyke07}.
-An alternative explanation for this late effect is that eye-movement control constraints can push the eye to move on before processing is finished.
-In contrast, ERPs are a time-sensitive direct measure of brain activity in response to the stimuli without confounds of planning and execution of motor activity (i.e., planning of saccades during natural reading in eye-tracking experiments or planning of button presses in SPR experiments).
-The well-studied N400 ERP component is associated with integration in the sentence context and / or memory retrieval induced by the current stimulus \citep{brouwer2017_n4_p6, lau2008_n400, kutas&federmeier_2000, kutas_federmeier2011}.
-The N400 is a negativity that peaks 400 ms after a word appeared on the screen in word-by-word presentation.
-The canonical finding is that the N400 amplitude is more negative for words which provide a poor semantic fit into the sentence context compared to words which fit the context perfectly \citep{kutas_hillyard}.
-Therefore, the N400 can provide insights into the immediate semantic processing of a word.
-This should prevent a potential semantic interference effect to be only detected at later regions in the sentence.
-
-%Although the N400 is primarily associated with semantic processing, if one follows the memory retrieval account it is reasonable to assume that retrieval of all aspects of a memory entry will modulate the N400 amplitude.
-The potential complications during memory retrieval due to interference should not differ dependent on the specific cue used for retrieval, i.e., \{$\pm$ grammatical subject\} or \{$\pm$ animate\}.
-Moreover, previous ERP studies investigating interference have found negativities in the typical N400 time window (300 - 500 ms post-stimulus onset), although the scalp topographies were not typical for the N400, being either broad \citep{schoknecht2022} or left anterior \citep{vasishth_drenhaus_2011}.
-Therefore, ERPs and especially the N400, are suited to be used for the investigation of interference (in the design from \citet{vandyke07} and in general).
-We chose to focus our EPR analyses on the N400 and expected to find interference effects -- both syntactic and semantic -- in N400 amplitude differences due to a more complex memory retrieval under high compared to low interference.
-
-%\section{I don't know why this is here}
-%In their \citeyear{vandyke07} paper, \citeauthor{vandyke07} reported a series of reading studies (one self-paced reading experiment and two eye-tracking experiments).
-For the eye-tracking experiments, \citet{vandyke07} reported three reading time measures: First-pass reading times (FPRT), regression-path durations (RPD) and total fixaton times (TFT).
-Reading times in multiple regions were reported, but the critical region was the verb.
-At the verb, \citet{vandyke07} found syntactic interference in both eye-tracking experiments and in all measures, but semantic interference was only found in Experiment 2 and only in RPD and TFT.
-In contrast, the spill-over region showed semantic interference more consistently than semantic interference.
-Surprisingly, the pre-critical region, which was only present in \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3, showed main effects of syntactic and semantic interference.
-On top of that, the RPDs of the sentence-final region in Experiment 2 and 3 were consistent with effects of syntactic and  semantic interference.
-
-%\citeauthor{mertzen}'s (\citeyear{mertzen}) estimates at the pre-critical region showed slowdowns for high vs.\ low  syntactic and semantic interference in English and German in RPD and TFT.
-In the reading times of the critical verb, \citeauthor{mertzen}'s (\citeyear{mertzen}) estimates were consistent with syntactic interference effects in all measures in English and German.
-Additionally, the estimates suggested a semantic interference effect in English.
-In the spill-over region, the estimates only suggested a semantic interference effect in German.
-
 \end{document}

### 2023-12-30 code_data/LV05_modeling/brain_response/generate_predictions.R
polished uniform

@@ -36,26 +36,13 @@ nsims<-20000
 
 ## Initial hyperparameters for latency factor
 ## Visualization of priors
-
-#Truncated Normal
-## the first two are not reasonable, they cause bimodality.
-#latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
-                                                    a=0.003, b=0.04, mean=0.01, sd=0.005))
-
-#latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
-                                                    a=0, mean=0.01, sd=0.005))
-
-# latency_factor_prior <-data.frame(x_temp=rtruncnorm(1000000, 
-                                                    a=0.005, b=0.04, mean=0.0001, sd=0.05))
-
 latency_factor_prior <-data.frame(x_temp=runif(1000000,min=0.005,max=0.04))
-                                    
-hist(latency_factor_prior$x_temp)
+hist(latency_factor_prior$x_temp, freq=FALSE)
 
 prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
   geom_histogram(aes(y=..density..),
-                 position="identity",fill="gray",
-                 binwidth=0.001, center = 0.0005)+
+                 position="identity",fill="gray", color="black",
+                 binwidth=0.003)+
   theme_bw()+
   theme(strip.text.x = element_text(size = 16, colour = "black", angle = 0))+
   xlab("latency factor")+
@@ -63,19 +50,14 @@ prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
 prior_lf
 ggsave("LF_Prior.png", width=5, height=3, dpi=600)
 
-
-
-## not reasonable, cause bimodality:
-#lfs<-rtruncnorm(nsims, a=0.003, b=0.04, mean=0.01, sd=0.005)
-#lfs<-rtruncnorm(nsims, a=0, mean=0.01, sd=0.005)
-
-lfs<-rtruncnorm(nsims, a=0.005, b=0.04, mean=0.0001, sd=0.0005)
-
 ## trying a uniform:
 lfs<-runif(nsims,min=0.005,max=0.04)
+lfs<-runif(nsims,min=0.005,max=0.025)
+
 
 ## check:
 hist(lfs,freq=FALSE)
+
 
 maineffectSyn<-rep(NA,nsims)
 maineffectSem<-rep(NA,nsims)
[…]

### 2023-12-31 paper/main.tex
removed paragraph

--- 2023-12-30 paper/main.tex
@@
 SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design: %(see Figure \ref{fig:previous} in the appendix):
+This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design:
 In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
@@
 
-There is one further possible reason for the diverging effects in our studies compared to previous work: the theoretically-predicted effects we are investigating here may be too subtle to consistently be detectable across studies.
-In several other attempts to replicate previously observed findings, we have often found it impossible to replicate the originally claimed effects, despite spending four or more years collecting data to obtain sufficient evidence.
-For example, contrary to published claims \parencite[e.g.,][]{vandyke_mcelree06}, in a large cross-linguistic study, we found evidence against proactive interference from sentence-external distractors \parencite{mertzen_repVDM2006}.
-A second example is the well-known syntactic local coherence effect in English \parencite{taboretal04}; in a recent large-sample study, we have found evidence against syntactic local coherence \parencite{lcpaape2023}.
-A third example is the interaction of locality and surprisal effects reported in \textcite{levy&keller_2013}; we attempted and failed to replicate the key claims in seven experiments \parencite{vasishth2018_signficancefilter}.
-A fourth example is the processing of antecedent-reflexive dependencies; \textcite{dillon2013} argued that reflexives are exempt from interference, but a large-scale direct replication attempt  \parencite{jaeger_etal_2020} found evidence that reflexives do show interference effects \parencite[cf.][]{yadav2021individual}.
-Finally, we have also failed to replicate the surprising claim  \parencite{gibsonwu,hsiao03} that Chinese object relatives are easier to process than subject relatives \parencite{VasishthetalPLoSOne2013,VasishthetalPLoSOne2013}.
-It is in principle possible that all the originally claimed effects actually represent true effects but, due to the inherent variability of data, the effects may be impossible to detect consistently in replications.
-Independent of any power issues in all these different studies, it is possible that it is unreasonable to expect consistent replications of effects in behavioral data, especially when investigating subtle effects.
-Certain other effects in psycholinguistics, such as strong garden-path effects \parencite{frazierrayner82}, are easily and consistently replicable \parencite[e.g.,][]{huang2023surprisal}.
-By contrast, subtle manipulations such as interference effects and local coherence may not show up consistently in replication attempts.
-One strong possibility is that there are systematic and important individual differences between participants' behavior; an increased focus away from investigating average effects and toward individual differences may reveal important insights about the conditions under which subtle effects are replicable; we demonstrate this point for the antecedent-reflexive dependency in \parencite{yadav2021individual}.
-
-Nevertheless, the contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
+The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
 So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?

### 2023-12-31 paper/main.tex
minor edits

@@
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
-A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013}.
+A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
 
@@
 Animate entities are proto-typical agents \citep{dowty1991thematic} and therefore, the primary use of the \{+ animate\} cue to retrieve a subject leads to the correct analysis with high probability (not just within the experiment context, but also in everyday language use).
+
  Finally, our findings are also consistent with the possibility that language processing relies predominantly on semantic associations to form (probabilistic)
@@
 There are of course important examples of the parser ignoring syntactic information.
-Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, local coherence effects \parencite{taboretal04}, and the agreement attraction phenomenon across different languages \parencite{wagersetal,lago_etal_2021,jaeger_etal_2017}.
-However, there are languages (such as German and Dutch) in which the grammaticality illusion does not occur \parencite{VSLK11,FrankEtAl2015}, and a large-sample study \parencite{lcpaape2023} presents evidence against the syntactic local coherence effect that was reported by \textcite{taboretal04}.
-Also, there seems to be at least one language that has rich case marking (Czech) that does not show any evidence of agreement attraction \parencite{chromy2023number}.
-Thus, it seems that, cross-linguistically, the role of syntax in building incremental structure and in completing dependencies remains an open question.
+Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, syntactic local coherence effects \parencite{taboretal04}, and the agreement attraction phenomenon across different languages \parencite[e.g.,][]{wagersetal,lago_etal_2021,tucker2015representing}.
+However, there exists  evidence inconsistent with these findings.
+Regarding grammaticality illusions in double-center embeddings, the grammaticality illusion does not occur in German and Dutch \parencite{VSLK11,FrankEtAl2015}.
+Regarding syntactic local coherence effects, a large-sample study \parencite{lcpaape2023} presents evidence against syntactic local coherence.
+Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of agreement attraction \parencite{chromy2023number}.
+Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.
+Of course, this does not rule out the possibility that, under certain conditions, other factors dominate.
 

### 2024-01-04 paper/main.tex
minor edits in conclusion

@@
 Both our experiments consistently showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor.
-This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb.
-These different time courses in the two experiments are likely to be due to a mix of encoding and retrieval interference.
-
-The observed differences in the time course in the SPR and ERP studies are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading).
-Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that--at least given the present data--the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
-A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that is in subject position within an embedded clause.
+This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions at the critical verb.
+The observed differences in the time course of the effect are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
+Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
 

### 2024-01-05 paper/main.tex
updated uniform prior

@@
         \centering
-         \caption{On the microvolt scale, i.e., for ERPs.}\label{fig:model_predictions_ERP}
+         \caption{On the microvolt scale, i.e., for negative ERPs.}\label{fig:model_predictions_ERP}
          \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
@@
 The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
-To generate prior predictions on the microvolt scale, we sampled the free scaling parameter $F$ from a truncated normal prior distribution ($Normal_+(0.01, 0.005)$) with the lower and upper bounds at 0.003 and 0.04, respectively.
+We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$).
+This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal probability for all effect sizes in this range.
+We are modeling N400 amplitude differences, so larger predicted effect sizes correspond to a more negative N400.
+To reflect this, the predicted effects were multiplied with $-1$.
 As in the reading time modeling, to avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects.
-For the ERP modeling, 3,146 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions.
+For the ERP modeling, 473 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions.
 The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}.

### 2024-01-05 code_data/LV05_modeling/brain_response/generate_predictions.R

@@ -35,7 +35,7 @@ VERBOSE <<- FALSE
 nsims<-20000
 
 ## Initial hyperparameters for latency factor
-## Visualization of priors
+## Visualization of uniform prior
 latency_factor_prior <-data.frame(x_temp=runif(1000000,min=0.005,max=0.04))
 hist(latency_factor_prior$x_temp, freq=FALSE)
 
@@ -50,10 +50,8 @@ prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
 prior_lf
 ggsave("LF_Prior.png", width=5, height=3, dpi=600)
 
-## trying a uniform:
+## uniform prior (effect between 0.5 and 4 muV):
 lfs<-runif(nsims,min=0.005,max=0.04)
-lfs<-runif(nsims,min=0.005,max=0.025)
-
 
 ## check:
 hist(lfs,freq=FALSE)
@@ -117,7 +115,7 @@ simdat_2cues_sameclause_long$effsize_neg <- -1 * simdat_2cues_sameclause_long$ef
 
 
 ggplot(data=simdat_2cues_sameclause_long, x=effsize_neg) +
-  geom_density(aes(x=effsize_neg), fill="lightgrey", bw=0.1) +
+  geom_density(aes(x=effsize_neg), fill="lightgrey") +
   ggh4x::facet_grid2(. ~ effect, scale="free_y", independent = "y")+
   theme_bw(base_size = 10)+
   geom_vline(xintercept=0, linetype="dashed")+
@@ -127,13 +125,10 @@ ggsave("PriorPredicted.png", width=6, height=4, dpi=600)
 
 #### Generate posterior predictions ####
 
-# Load simulated data
-load("simdat_2cues_sameclause.Rda")
-
 # Run ABC with rejection sampling
 # results from Schoknecht & Vasishth (N400 elicited by critical verb): 
 # syntactic mean effect -0.2 muV, CrI [-0.4,0.0] --> Normal(-0.2,0.1) --> SE: 0.1
-# semantic mean effect -0.3 muV, CrI [-0.6,-0.1] --> Normal(-0.3,0.15) --> SE should be 0.125 but set slightly wider at SE: 0.15
+# semantic mean effect -0.3 muV, CrI [-0.6,-0.1] --> Normal(-0.3,0.125) --> SE: 0.125
 # interaction mean effect -0.1 muV, CrI [-0.4, 0.0] --> Normal(-0.1,0.1) --> SE: 0.1
 
[…]

### 2024-01-05 paper/main.tex
explained choice of prior
added osf link

@@
 This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal probability for all effect sizes in this range.
+This uninformative prior was chosen because of limited prior knowledge about interference effects during subject-verb dependency resolution in ERPs \citep[but see e.g.,][]{lee_garnsey, vasishth_drenhaus_2011}.
 We are modeling N400 amplitude differences, so larger predicted effect sizes correspond to a more negative N400.
@@
 
-%\section{Data availability statement}
-
-%The data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=.
+\vspace{2em}
+\section{Data availability statement}
+
+The data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.
 

### 2024-01-05 paper/main.tex
removed highlights and keywords for submission

@@
 
-\section{Highlights} 
-% 3-5 highlights
-% only 85 characters including spaces per highlight
-\begin{itemize}
-\item Self-paced reading and event-related potentials experiments on interference
-\item Largest studies on interference during subject-verb integration (N=774, N=103)
-\item Bayes factors show evidence for semantic, but against syntactic interference
-\item Modeling shows that tracking of syntactic structure attenuates syntactic interference
-\end{itemize}
-
 }
-
-\keywords{Cue-based Retrieval, Syntactic Interference, Semantic Interference, Self-paced Reading, Event-related Potentials, Computational Modeling}
 

### 2024-01-08 paper/main.tex
moved ethics statement
edited position of fig
removed ethics statement
edited uniform prior explanation

@@
 \subsection{Methods}
-This experiment was conducted with the approval of the ethics committee of the University of Potsdam, Germany.
-
 \subsubsection{Participants}
@@
 
-\begin{figure}[H]
+\begin{figure}[!htbp]
     \caption{Prior and posterior predicted effect sizes for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
@@
 We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$).
-This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal probability for all effect sizes in this range.
-This uninformative prior was chosen because of limited prior knowledge about interference effects during subject-verb dependency resolution in ERPs \citep[but see e.g.,][]{lee_garnsey, vasishth_drenhaus_2011}.
+This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal likelihood for all effect sizes in this range.
+This uninformative prior was chosen because there exists no previously published ERP data on the Van Dyke (2007) design.
 We are modeling N400 amplitude differences, so larger predicted effect sizes correspond to a more negative N400.

### 2024-01-19 code_data/EEG/r_scripts/acc_analysis.R
Updated via the Open Science Framework

@@ -7,7 +7,7 @@ library(brms)
 
 
 # load all .csv files from logs folder
-tbl <- list.files(path = "logs/", pattern = "*.csv", full.names = TRUE) %>% 
+tbl <- list.files(path = "../data/logs/", pattern = "*.csv", full.names = TRUE) %>% 
                   map_df(~read_csv(., col_types = cols(.default = "c"))) 
 
 tbl$subject_nr <- ifelse(tbl$subject_nr == "136" & tbl$datetime == "Mon May  8 10:21:50 2023", "135", tbl$subject_nr)
@@ -25,12 +25,12 @@ dat2 <- dat %>% mutate(subject_nr = fct_recode(subject_nr, "1" = "200",  "3" = "
 (N_subj <- length(unique(dat2$subject_nr)))
 
 # exclude participants who were excluded due to low accuracy or to many artefacts
-ex_s1 <- read.table("exclude_s1.txt")
+ex_s1 <- read.table("../data/exclude/exclude_s1.txt")
 ex_s1$subj <- as.factor(ex_s1$subj)
-ex_s2 <- read.table("exclude_s2.txt")
+ex_s2 <- read.table("../data/exclude/exclude_s2.txt")
 ex_s2$subj <- as.factor(ex_s2$subj)
 
-ex_arte <- read.csv("../TooManyArtefacts.csv")
+ex_arte <- read.csv("../data/exclude/TooManyArtefacts.csv")
 ex_arte$subject <- as.factor(ex_arte$subject)
 
 

### 2024-01-19 code_data/EEG/r_scripts/analyses.R
Updated via the Open Science Framework

@@ -18,8 +18,8 @@ library(stats)
 options(scipen = 999)
 
 #load data
-prestim <- read_csv("data/Pandora_mean_200_0.csv")
-n400 <- read_csv("data/Pandora_mean_300_500.csv")
+prestim <- read_csv("../data/Pandora_mean_200_0.csv")
+n400 <- read_csv("../data/Pandora_mean_300_500.csv")
 
 # remove subjects who have only one good session
 xtabs(~ session + subject, n400)
@@ -49,7 +49,7 @@ plot(keptseg$subj, keptseg$num_seg, keptseg$cond)
 low10 <- subset(keptseg, keptseg$num_seg < 20)
 lows <- low10[!duplicated(low10[1]),]
 
-write.csv(lows, file="TooManyArtefacts.csv", row.names = FALSE)
+write.csv(lows, file="../data/exclude/TooManyArtefacts.csv", row.names = FALSE)
 
 # exclude subjects with less than 20 trials per condition
 n400.3 <- anti_join(n400.2, lows, by="subject")
@@ -128,13 +128,6 @@ priors_xl_tr <- c(
   prior(normal(10, 5), class = sigma),
   prior(normal(0, 2), class = sd)
 )  
-
-# check priors
-round(2-2*5);round(2+2*5) # intercept
-(2+(-0.1))-(2-(-0.1)) # beta normal-(0, 0.1)
-(2+(-0.5))-(2-(-0.5)) # beta normal-(0, 0.5)
-(2+(-1))-(2-(-1))     # beta normal-(0, 1)
-(2+(-5))-(2-(-5))     # beta normal-(0, 5)
 
 # full models with truncated priors
 m_s_tr <- brm(amplitude ~ 1 + prestim_amp + syn * sem  + (1 + sem ||subject) + (1|item),

### 2024-01-19 code_data/plausibility_norming/norming_analysis.R
Updated via the Open Science Framework

@@ -42,7 +42,7 @@ dfp$cond <- factor(dfp$cond, levels = c("intro", "subject", "distr_anim", "distr
 
 ggplot(dfp, aes(x=cond, y=rating))+
   geom_jitter(color="lightblue", alpha=0.2)+
-  geom_boxplot(fill="grey", alpha=0.7, outlier.shape = NA)+
+  geom_boxplot(fill="grey", alpha=0.7)+
   stat_summary(fun=mean, geom="point", shape=18, size=4) + #adding the mean
   expand_limits(y=c(0.5,7.5))+
   scale_y_continuous(name="plausibility rating", breaks=seq(1, 7, 1))+
@@ -50,7 +50,7 @@ ggplot(dfp, aes(x=cond, y=rating))+
                                      "intro" = "introduction", "subject" = "subject"))+
   theme_bw()+
   labs(title = "Plausibility ratings",
-       subtitle = "1 = absolute implausible; 7 = absolute plausible")
+       subtitle = "1 = absolutely implausible; 7 = absolutely plausible")
 
-ggsave("plots/pandora_plausibility_ratings.jpg", dpi=600, width=4, height=4)
+ggsave("pandora_plausibility_ratings.jpg", dpi=600, width=4, height=4)
 

### 2024-01-19 code_data/SPR/r_scripts/acc_analysis.R
Updated via the Open Science Framework

@@ -6,9 +6,9 @@ library(tidyverse)
 library(brms)
 
 # load data
-df1 <- read.csv("../pandora_web/data/pandora_web_data_allsessions_short.csv", sep=";")
-df2 <- read.csv("../pandora_repS1/data/pandora_repS1_data_cleaned_short.csv", sep=";")
-df3 <- read.csv("../pandora_repS2/data/pandora_repS2_data_cleaned_short.csv", sep=";")
+df1 <- read.csv("../data/pandora_web_data_allsessions_short.csv", sep=";")
+df2 <- read.csv("../data/pandora_repS1_data_cleaned_short.csv", sep=";")
+df3 <- read.csv("../data/pandora_repS2_data_cleaned_short.csv", sep=";")
 
 # add session info to df2 and df3 (both experiments had only one session)
 df2$session <- 1

### 2024-01-19 code_data/SPR/r_scripts/analyses_crit.R
Updated via the Open Science Framework

@@ -10,7 +10,7 @@ library(truncnorm)
 
 
 # load data
-crit_trim <- read.csv("data/pandora_spr_774_crit.csv")
+crit_trim <- read.csv("../data/pandora_spr_774_crit.csv")
 
 
 # Inferential statistics
@@ -37,13 +37,6 @@ priors_l_tr <- c(
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
-
-# check priors
-round(exp(6-2*0.6));round(exp(6+2*0.6)) # intercept
-round(exp(6+0.01) - exp(6-0.01))  # beta normal-(0, 0.01)
-round(exp(6+0.05) - exp(6-0.05))  # beta normal-(0, 0.05)
-round(exp(6+0.1) - exp(6-0.1))    # beta normal-(0, 0.1)
-
 
 # models
 m_s_full <- brm(rt ~ 1 + syn * sem +
@@ -260,19 +253,3 @@ df.bf[nrow(df.bf)+1,] <- c("critical", "interaction","Normal(0, 0.05)","yes", ro
 df.bf[nrow(df.bf)+1,] <- c("critical", "interaction","Normal(0, 0.1)","yes", round(bf_l_int1,2), round(bf_l_int2,2))
 
 save(df.bf, file = paste("BFs_spr_pooled_774_crit.Rda"))
-
-
-#### model with random slopes for individual estimates ####
-
-m_m_full_slopes <- brm(rt ~ 1 + syn * sem +
-                  (1 + syn * sem|| participant) +
-                  (1 + syn * sem|| item),
-                data = crit_trim,
-                family = lognormal(),
-                prior = priors_m,
-                warmup = 2000,
-                iter = 20000,
-                cores = 4,
-                save_pars = save_pars(all = TRUE)
-)
-save(m_m_full_slopes, file = paste("model_fits/Fit_m_Full_RSLOPES_crit.Rda"))

### 2024-01-19 code_data/SPR/r_scripts/analyses_precrit.R
Updated via the Open Science Framework

@@ -12,7 +12,7 @@ library(bayesplot)
 
 
 # load data
-precrit_trim <- read.csv("data/pandora_spr_774_precrit.csv")
+precrit_trim <- read.csv("../data/pandora_spr_774_precrit.csv")
 
 
 # Inferential statistics
@@ -39,13 +39,6 @@ priors_l_tr <- c(
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
-
-# check priors
-round(exp(6-2*0.6));round(exp(6+2*0.6)) # intercept
-round(exp(6+0.01) - exp(6-0.01))  # beta normal-(0, 0.01)
-round(exp(6+0.05) - exp(6-0.05))  # beta normal-(0, 0.05)
-round(exp(6+0.1) - exp(6-0.1))    # beta normal-(0, 0.1)
-
 
 # models
 m_s_full <- brm(rt ~ 1 + syn * sem +

### 2024-01-19 code_data/SPR/r_scripts/analyses_spill.R
Updated via the Open Science Framework

@@ -11,7 +11,7 @@ library(truncnorm)
 
 
 # load data
-spill_trim <- read.csv("data/pandora_spr_774_spill.csv")
+spill_trim <- read.csv("../data/pandora_spr_774_spill.csv")
 
 
 # Inferential statistics
@@ -38,13 +38,6 @@ priors_l_tr <- c(
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
-
-# check priors
-round(exp(6-2*0.6));round(exp(6+2*0.6)) # intercept
-round(exp(6+0.01) - exp(6-0.01))  # beta normal-(0, 0.01)
-round(exp(6+0.05) - exp(6-0.05))  # beta normal-(0, 0.05)
-round(exp(6+0.1) - exp(6-0.1))    # beta normal-(0, 0.1)
-
 
 # models
 m_s_full <- brm(rt ~ 1 + syn * sem +

### 2024-01-19 code_data/SPR/r_scripts/Plot_ReadingTimes.R
Updated via the Open Science Framework

@@ -9,9 +9,9 @@ library(tidyverse)
 library(cowplot)
 
 # load data
-df1 <- read.csv("../pandora_web/data/pandora_web_data_allsessions_short.csv", sep=";")
-df2 <- read.csv("../pandora_repS1/data/pandora_repS1_data_cleaned_short.csv", sep=";")
-df3 <- read.csv("../pandora_repS2/data/pandora_repS2_data_cleaned_short.csv", sep=";")
+df1 <- read.csv("../data/pandora_web_data_allsessions_short.csv", sep=";")
+df2 <- read.csv("../data/pandora_repS1_data_cleaned_short.csv", sep=";")
+df3 <- read.csv("../data/pandora_repS2_data_cleaned_short.csv", sep=";")
 
 # delete some columns in df1 to have identical columns in all data sets
 df1$delay <- NULL

### 2024-01-19 code_data/SPR/r_scripts/create_regionwise_datasets.R
Updated via the Open Science Framework

@@ -8,9 +8,9 @@ library(tidyverse)
 
 
 # load data
-df1 <- read.csv("../pandora_web/data/pandora_web_data_allsessions_short.csv", sep=";")
-df2 <- read.csv("../pandora_repS1/data/pandora_repS1_data_cleaned_short.csv", sep=";")
-df3 <- read.csv("../pandora_repS2/data/pandora_repS2_data_cleaned_short.csv", sep=";")
+df1 <- read.csv("../data/pandora_web_data_allsessions_short.csv", sep=";")
+df2 <- read.csv("../data/pandora_repS1_data_cleaned_short.csv", sep=";")
+df3 <- read.csv("../data/pandora_repS2_data_cleaned_short.csv", sep=";")
 
 # delete some columns in df1 to have identical columns in all data sets
 df1$delay <- NULL

### 2024-01-19 code_data/EEG/r_scripts/analyses.R
fixed x-axis labels in BF plot

@@ -397,23 +397,30 @@ df.bf$BF10 <- as.numeric(df.bf$BF10.1)
 df.bf$effect <- factor(df.bf$effect, levels=c("semantic", "syntactic", "interaction"))
 
 
+df.bf$prior2 <- ifelse(df.bf$prior == "Normal(0, 0.1)", "Normal_(0, 0.1)", (
+                          ifelse(df.bf$prior == "Normal(0, 0.5)", "Normal_(0, 0.5)",(
+                               ifelse(df.bf$prior == "Normal(0, 1)", "Normal_(0, 1)",
+                                  "Normal_(0, 5)")))))
+
 
 # Plot different BFs
 library(ggbreak)
 library(ggimage) 
 
 ## original plot
-plot_BF<- ggplot(df.bf, aes(x = prior, y = BF10, group = interaction(effect, truncated))) +
+plot_BF<- ggplot(df.bf, aes(x = prior2, y = BF10, group = interaction(effect, truncated))) +
   geom_point(aes(color=effect)) +
   geom_line(aes(color=effect)) +
   geom_hline(yintercept = 1, linetype="dashed") +
-  #coord_cartesian(ylim = c(0, 50)) +
-  #annotate("text", x =3,y= 1.5, label = "Evidence in favor of H1", size =3)+
-  #annotate("text", x =3,y= 2/3, label = "Evidence in favor of H0", size =3) +
-  theme_bw(base_size = 8)+
-  theme(legend.position = c(0.8, 0.8), legend.text=element_text(size=6))+
-  #scale_y_break(c(12, 21), scales=2) + #scale_y_continuous(limits = c(0, 90), breaks =  c(1, 3, 10, 30, 50, 80, 90))+
-  #scale_y_break(c(12, 21), scales=2) +
+#  annotate("text", x =4,y= 2, label = "evidence for effect ")+
+#  geom_segment(aes(x = 3.35, y = 1.5, xend = 3.35, yend = 2.75), color="black",
+#               arrow = arrow(length = unit(0.15, "cm")), inherit.aes = FALSE)+
+#  annotate("text", x =3.82,y= -0.75, label = "evidence against effect") +
+#  geom_segment(aes(x = 3.05, y = -0.25, xend = 3.05, yend = -1.5), color="black",
+#               arrow = arrow(length = unit(0.15, "cm")), inherit.aes = FALSE)+
[…]

### 2024-01-19 paper/main.tex
Updated via the Open Science Framework

@@
 Reading studies have been the dominant method used to investigate retrieval interference \parencite[see the review in][]{jaeger_etal_2017}.
-We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading as well as electroencephalography (EEG), which to our knowledge has never been used to investigate interference during subject-verb dependency resolution.
+We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading (SPR) as well as electroencephalography (EEG), which to our knowledge has never been used to investigate interference during subject-verb dependency resolution.
 
@@
 The SPR experiments were conducted web-based via Prolific and the EEG experiment was of course run in the lab.
-Because the high signal-to-noise ratio of EEG data requires a large number of trials per condition and participant, we constructed 120 items.
+Because the signal-to-noise ratio of EEG data requires a large number of trials per condition and participant, we constructed 120 items.
 In order to keep the SPR and EEG studies comparable, we also used 120 items in the SPR experiments.
@@
 The SPR study, hereafter Experiment 1a, was initially designed so that participants read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120).
-The variation between participants in the gap is due to the fact that we could not control when the participants would do the second session; however, the mean and median gaps were  about 11 days, with the first and third quartiles being 9 and 12 days.
+The variation in the gap is due to the fact that we could not control when the participants would do the second session; however, the mean and median gaps were  about 11 days, with the first and third quartiles being 9 and 12 days.
 
@@
 This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
-Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is in principle possible to find evidence both for or against an effect.
+Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is possible to find evidence for or against an effect.
 This stands in contrast to the frequentist ANOVA which, in its standard usage, is designed to only furnish evidence against the null.
@@
 The results of the present study stand in contrast to previous findings \citep{vandyke07, mertzen}.
-The present study is the only one to find semantic and no syntactic interference effect.
+The present study is the only one to find semantic and no syntactic interference effects.
 The difference in observed effects could be explained by differences in statistical power or by methodological differences.
@@
 The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task.
-Self-paced reading (SPR) could trigger a greater tendency to cause interference effects compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}.
+Self-paced reading (SPR) might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}.
 SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
@@
 
-The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even identical items.
+The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even the same linguistic material.
 So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
@@
 The experiment was conducted in two experimental sessions for practical reasons (each of the sessions lasted approximately two hours).
-Sessions were separated by one to eight weeks for each participant (the median gap was 2 weeks, the first and third quartiles being 1 and 3 weeks).
+Sessions were separated by one to eight weeks for each participant (the median gap was two weeks, the first and third quartiles being one and three weeks).
 The procedure of both sessions was identical.
@@
 Eye-movements were monitored with six electrodes which were positioned above, below and at the outer canthus of both eyes.
-Impedances of all electrodes were kept below 5k$\Omega$.
+Impedances of all electrodes were kept below 5\,k$\Omega$.
 
@@
 Comprehension accuracy was reduced for high vs.\ low semantic interference conditions.
-There was no indication that syntactic interference affected accuracy, although the sign of the effect was in the expected direction, but  amounting to a 1\% difference on the probability scale between the high and low syntactic interference conditions.
+There was no indication that syntactic interference affected accuracy.
+The sign of the effect was in the expected direction, but  amounting to a 1\% difference on the probability scale between the high and low syntactic interference conditions.
 There was a numerical suggestion of an interaction: the difference in accuracy between low and high semantic interference conditions was larger when syntactic interference was high vs.\ when it was low (17 vs.\ 8 \% difference).
@@
 The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations.
-\ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter $F$ and then generating data from the model while holding all other parameters in the model at fixed values.
+Figure \ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter $F$ and then generating data from the model while holding all other parameters in the model at fixed values.
 The $F$ parameter scales the activation to a retrieval time.
@@
 Could the absence of syntactic interference be reconciled with a cue-based retrieval account?
-We speculated that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
+We speculate that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
 Specifically, the relevant syntactic cues that might be used at the verb are not  \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} but rather a composite cue \{$\pm$ subject-in-same-clause\}.
@@
 The general effect pattern (no syntactic interference effect, but a semantic interference effect) is similar in both the prior and posterior predictive distributions.
-Therefore, the assumption that the relevant syntactic cue is \{$\pm$ subject-in-same-clause\} cue, rather than the \{$\pm$ grammatical subject\} cue originally assumed in previous modeling work, seems reasonable for the present ERP study as well.
+Therefore, the assumption that the relevant syntactic cue is the \{$\pm$ subject-in-same-clause\} cue -- rather than the previously assumed \{$\pm$ grammatical subject\} cue --  seems reasonable for the present ERP study as well.
 
@@
 Regarding syntactic local coherence effects, a large-sample study \parencite{lcpaape2023} presents evidence against syntactic local coherence.
-Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of agreement attraction \parencite{chromy2023number}.
+Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of number agreement attraction \parencite{chromy2023number}.
 Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.

### 2024-05-29 code_data/previous_studies/PreviousFindings.Rmd
Added via the Open Science Framework

---
title: "Plot previous findings"
author: "Pia Schoknecht"
output: pdf_document
---

```{r}
library(ggplot2)
library(tidyverse)
library(cowplot)
library(ggbreak)

```

# Van Dyke (2007) estimates from published paper:

## E1

### critical

```{r}
## E1 critical

## not sig. Syn effect in E1 crit:
VD07E1critSynmn <- (875+871)/2 - (912+858)/2
VD07E1critSynSE <- 30
VD07E1critSynlower <- round(VD07E1critSynmn - 2*VD07E1critSynSE)
VD07E1critSynupper <- round(VD07E1critSynmn + 2*VD07E1critSynSE)

## sig. Sem effect in E1 crit:
VD07E1critSemmn <- (875+912)/2 - (871+858)/2
VD07E1critSemF1 <- 6.33
VD07E1critSemSE <- VD07E1critSemmn/sqrt(VD07E1critSemF1)
VD07E1critSemlower <- round(VD07E1critSemmn - 2*VD07E1critSemSE)
VD07E1critSemupper <- round(VD07E1critSemmn + 2*VD07E1critSemSE)

## not sig. Interaction effect in E1 crit:
VD07E1critIntmn <- (875-871) - (912-858)
VD07E1critIntF1 <- 2.36
VD07E1critIntSE <- VD07E1critIntmn/sqrt(VD07E1critIntF1)
VD07E1critIntlower <- round(VD07E1critIntmn - 2*VD07E1critIntSE)
VD07E1critIntupper <- round(VD07E1critIntmn + 2*VD07E1critIntSE)
```

### spill-over

```{r}
## E1 spill-over

## not sig. Syn effect in E1 spill-over:
VD07E1spillSynmn <- (568+551)/2 - (571+566)/2
VD07E1spillSynSE <- 18 #highest SE in spill-over region, Table 3
VD07E1spillSynlower <- round(VD07E1spillSynmn - 2*VD07E1spillSynSE)
VD07E1spillSynupper <- round(VD07E1spillSynmn + 2*VD07E1spillSynSE)

## sig. Sem effect in  E1 spill-over:
VD07E1spillSemmn <- (568+571)/2 - (551+566)/2
VD07E1spillSemF1 <- 18 #highest SE in spill-over region, Table 3
VD07E1spillSemSE <- VD07E1spillSemmn/sqrt(VD07E1spillSemF1)
VD07E1spillSemlower <- round(VD07E1spillSemmn - 2*VD07E1spillSemSE)
VD07E1spillSemupper <- round(VD07E1spillSemmn + 2*VD07E1spillSemSE)

## not sig. Interaction effect in E1 spill-over:
VD07E1spillIntmn <- (568-551) - (571-566)
VD07E1spillIntF1 <- 18 #highest SE in spill-over region, Table 3
VD07E1spillIntSE <- VD07E1spillIntmn/sqrt(VD07E1spillIntF1)
VD07E1spillIntlower <- round(VD07E1spillIntmn - 2*VD07E1spillIntSE)
VD07E1spillIntupper <- round(VD07E1spillIntmn + 2*VD07E1spillIntSE)
```

### Sentence-Final

```{r}
## E1 sentence-final

##  Syn effect in E1 final:
VD07E1finalSynmn <- (822+667)/2 - (697+723)/2
VD07E1finalSyntval <- sqrt(0.99) # F1 < 1 (Table 4)
VD07E1finalSynSE <- VD07E1finalSynmn/VD07E1finalSyntval
VD07E1finalSynlower <- round(VD07E1finalSynmn - 2*VD07E1finalSynSE)
VD07E1finalSynupper <- round(VD07E1finalSynmn + 2*VD07E1finalSynSE)

## Sem effect in  E1 final:
VD07E1finalSemmn <- (822+697)/2 - (667+723)/2
VD07E1finalSemtval <- sqrt(3.06) # F1 (Table 4)
VD07E1finalSemSE <- VD07E1finalSemmn/VD07E1finalSemtval
VD07E1finalSemlower <- round(VD07E1finalSemmn - 2*VD07E1finalSemSE)
VD07E1finalSemupper <- round(VD07E1finalSemmn + 2*VD07E1finalSemSE)

## Interaction effect in E1 final:
VD07E1finalIntmn <- (822-667) - (697-723)
VD07E1finalInttval <- sqrt(5.39) # F1 (Table 4)
VD07E1finalIntSE <- VD07E1finalIntmn/VD07E1finalInttval
VD07E1finalIntlower <- round(VD07E1finalIntmn - 2*VD07E1finalIntSE)
VD07E1finalIntupper <- round(VD07E1finalIntmn + 2*VD07E1finalIntSE)
```

## Experiment 2

### critical: SYN

```{r}
# Experiment 2 SYN in critical

## sig. Syn effect in E2 crit region FPRT:
VD07E2critSynFPRTmn <- (413+418)/2-(376+382)/2
VD07E2critSynFPRTtval <- sqrt(8.32)
VD07E2critSynFPRTSE <- VD07E2critSynFPRTmn/VD07E2critSynFPRTtval
VD07E2critSynFPRTlower <- round(VD07E2critSynFPRTmn - 2*VD07E2critSynFPRTSE)
VD07E2critSynFPRTupper <- round(VD07E2critSynFPRTmn + 2*VD07E2critSynFPRTSE)

# sig. Syn effect in E2 crit region RPD:
VD07E2critSynRPDmn <- (594+663)/2-(454+495)/2
VD07E2critSynRPDtval <- sqrt(27.21)
VD07E2critSynRPDSE <- VD07E2critSynRPDmn/VD07E2critSynRPDtval
VD07E2critSynRPDlower <- round(VD07E2critSynRPDmn - 2*VD07E2critSynRPDSE)
VD07E2critSynRPDupper <- round(VD07E2critSynRPDmn + 2*VD07E2critSynRPDSE)

## sig. Syn effect in E2 crit region TFT:
VD07E2critSynTFTmn <- (738+761)/2-(630+653)/2
VD07E2critSynTFTtval <- sqrt(29.95)
VD07E2critSynTFTSE <- VD07E2critSynTFTmn/VD07E2critSynTFTtval
VD07E2critSynTFTlower <- round(VD07E2critSynTFTmn - 2*VD07E2critSynTFTSE)
VD07E2critSynTFTupper <- round(VD07E2critSynTFTmn + 2*VD07E2critSynTFTSE)
```

### critical: SEM

```{r}
# Experiment 2 Sem in critical

## non sig. Sem effect in E2 crit region FPRT:
VD07E2critSemFPRTmn <- (418+382)/2-(413+376)/2
VD07E2critSemFPRTtval <- sqrt(1.13)
VD07E2critSemFPRTSE <- VD07E2critSemFPRTmn/VD07E2critSemFPRTtval
VD07E2critSemFPRTlower <- round(VD07E2critSemFPRTmn - 2*VD07E2critSemFPRTSE)
VD07E2critSemFPRTupper <- round(VD07E2critSemFPRTmn + 2*VD07E2critSemFPRTSE)

# non sig. Sem effect in E2 crit region RPD:
VD07E2critSemRPDmn <- (663+495)/2-(594+454)/2
VD07E2critSemRPDtval <- sqrt(3.24)
VD07E2critSemRPDSE <- VD07E2critSemRPDmn/VD07E2critSemRPDtval
VD07E2critSemRPDlower <- round(VD07E2critSemRPDmn - 2*VD07E2critSemRPDSE)
VD07E2critSemRPDupper <- round(VD07E2critSemRPDmn + 2*VD07E2critSemRPDSE)

## non sig. Sem effect in E2 crit region TFT:
VD07E2critSemTFTmn <- (761+653)/2-(738+630)/2
VD07E2critSemTFTtval <- 1.95 # reported as non-sig.
VD07E2critSemTFTSE <- VD07E2critSemTFTmn/VD07E2critSemTFTtval
VD07E2critSemTFTlower <- round(VD07E2critSemTFTmn - 2*VD07E2critSemTFTSE)
VD07E2critSemTFTupper <- round(VD07E2critSemTFTmn + 2*VD07E2critSemTFTSE)
```

### critical: Int

```{r}
# Experiment 2 Int in critical

## non sig. Int effect in E2 crit region FPRT:
VD07E2critIntFPRTmn <- (418-413)-(382-376)
VD07E2critIntFPRTtval <- 1.95 # reported as non-sig.
VD07E2critIntFPRTSE <- VD07E2critIntFPRTmn/VD07E2critIntFPRTtval
VD07E2critIntFPRTlower <- round(VD07E2critIntFPRTmn - 2*VD07E2critIntFPRTSE)
VD07E2critIntFPRTupper <- round(VD07E2critIntFPRTmn + 2*VD07E2critIntFPRTSE)

# non sig. Int effect in E2 crit region RPD:
VD07E2critIntRPDmn <- (663-594)-(495-454)
VD07E2critIntRPDtval <- 1.95 # reported as non-sig.
VD07E2critIntRPDSE <- VD07E2critIntRPDmn/VD07E2critIntRPDtval
VD07E2critIntRPDlower <- round(VD07E2critIntRPDmn - 2*VD07E2critIntRPDSE)
VD07E2critIntRPDupper <- round(VD07E2critIntRPDmn + 2*VD07E2critIntRPDSE)

## non sig. Int effect in E2 crit region TFT:
VD07E2critIntTFTmn <- (761-738)-(653-630)
VD07E2critIntTFTtval <- 1.95 # reported as non-sig.
VD07E2critIntTFTSE <- VD07E2critIntTFTmn/VD07E2critIntTFTtval
VD07E2critIntTFTlower <- round(VD07E2critIntTFTmn - 2*VD07E2critIntTFTSE)
VD07E2critIntTFTupper <- round(VD07E2critIntTFTmn + 2*VD07E2critIntTFTSE)
```

### spill-over: SYN

```{r}
# Experiment 2 SYN in spill-over

## almost sig. Syn effect in E2 post-crit region FPRT (negative sign)
VD07E2postcritSynFPRTmn<-(325+296)/2-(320+364)/2
VD07E2postcritSynFPRTtval<-sqrt(6.58)
VD07E2postcritSynFPRTSE<-VD07E2postcritSynFPRTmn/VD07E2postcritSynFPRTtval
VD07E2postcritSynFPRTlower<-VD07E2postcritSynFPRTmn-2*VD07E2postcritSynFPRTSE
VD07E2postcritSynFPRTupper<-VD07E2postcritSynFPRTmn+2*VD07E2postcritSynFPRTSE

## almost sig. Syn effect in E2 post-crit region RPD:
VD07E2postcritSynRPDmn <- (1365+1295)/2-(970+1205)/2
VD07E2postcritSynRPDtval <- sqrt(12.17)
VD07E2postcritSynRPDSE <- VD07E2postcritSynRPDmn/VD07E2postcritSynRPDtval
VD07E2postcritSynRPDlower <- round(VD07E2postcritSynRPDmn - 2*VD07E2postcritSynRPDSE)
VD07E2postcritSynRPDupper <- round(VD07E2postcritSynRPDmn + 2*VD07E2postcritSynRPDSE)

## non sig. Syn effect in E2 post-crit region TFT:
VD07E2postcritSynTFTmn <- (491+493)/2-(502+540)/2
VD07E2postcritSynTFTtval <- 1.95 # reported in paper as non-sig
VD07E2postcritSynTFTSE<- VD07E2postcritSynTFTmn/VD07E2postcritSynTFTtval
VD07E2postcritSynTFTlower <- round(VD07E2postcritSynTFTmn - 2*VD07E2postcritSynTFTSE)
VD07E2postcritSynTFTupper <- round(VD07E2postcritSynTFTmn + 2*VD07E2postcritSynTFTSE)
```

### spill-over: Sem

```{r}
# Experiment 2 Sem in spill-over

## almost sig. Sem effect in E2 post-crit region FPRT (negative sign)
VD07E2postcritSemFPRTmn<- (296+364)/2-(325+320)/2
VD07E2postcritSemFPRTtval<- 1.95 # reported in paper as non-sig
VD07E2postcritSemFPRTSE<-VD07E2postcritSemFPRTmn/VD07E2postcritSemFPRTtval
VD07E2postcritSemFPRTlower<-VD07E2postcritSemFPRTmn-2*VD07E2postcritSemFPRTSE
VD07E2postcritSemFPRTupper<-VD07E2postcritSemFPRTmn+2*VD07E2postcritSemFPRTSE

## almost sig. Sem effect in E2 post-crit region RPD:
VD07E2postcritSemRPDmn <- (1295+1205)/2-(1365+970)/2
VD07E2postcritSemRPDtval <- 1.95 # reported in paper as non-sig
VD07E2postcritSemRPDSE <- VD07E2postcritSemRPDmn/VD07E2postcritSemRPDtval
VD07E2postcritSemRPDlower <- round(VD07E2postcritSemRPDmn - 2*VD07E2postcritSemRPDSE)
VD07E2postcritSemRPDupper <- round(VD07E2postcritSemRPDmn + 2*VD07E2postcritSemRPDSE)

## almost sig. Sem effect in E2 post-crit region TFT:
VD07E2postcritSemTFTmn <- (493+540)/2-(491+502)/2
VD07E2postcritSemTFTtval <- 1.95 # reported in paper as non-sig
VD07E2postcritSemTFTSE <- VD07E2postcritSemTFTmn/VD07E2postcritSemTFTtval
VD07E2postcritSemTFTlower <- round(VD07E2postcritSemTFTmn - 2*VD07E2postcritSemTFTSE)
VD07E2postcritSemTFTupper <- round(VD07E2postcritSemTFTmn + 2*VD07E2postcritSemTFTSE)
```

### spill-over: Int

```{r}
# Experiment 2 Int in spill-over

## almost sig. Int effect in E2 post-crit region FPRT (negative sign)
VD07E2postcritIntFPRTmn<- (296-325)-(364-320)
VD07E2postcritIntFPRTtval<- sqrt(5.22)
VD07E2postcritIntFPRTSE<-VD07E2postcritIntFPRTmn/VD07E2postcritIntFPRTtval
VD07E2postcritIntFPRTlower<-VD07E2postcritIntFPRTmn-2*VD07E2postcritIntFPRTSE
VD07E2postcritIntFPRTupper<-VD07E2postcritIntFPRTmn+2*VD07E2postcritIntFPRTSE

## almost sig. Int effect in E2 post-crit region RPD:
VD07E2postcritIntRPDmn <- (1295-1365)-(1205-970)
VD07E2postcritIntRPDtval <- sqrt(3.45)
VD07E2postcritIntRPDSE <- VD07E2postcritIntRPDmn/VD07E2postcritIntRPDtval
VD07E2postcritIntRPDlower <- round(VD07E2postcritIntRPDmn - 2*VD07E2postcritIntRPDSE)
VD07E2postcritIntRPDupper <- round(VD07E2postcritIntRPDmn + 2*VD07E2postcritIntRPDSE)

## almost sig. Int effect in E2 post-crit region TFT:
VD07E2postcritIntTFTmn <- (493-491)-(540-502)
VD07E2postcritIntTFTtval <- 1.95 # reported in paper as non-sig
VD07E2postcritIntTFTSE <-  VD07E2postcritIntTFTmn/VD07E2postcritIntTFTtval
VD07E2postcritIntTFTlower <- round(VD07E2postcritIntTFTmn - 2*VD07E2postcritIntTFTSE)
VD07E2postcritIntTFTupper <- round(VD07E2postcritIntTFTmn + 2*VD07E2postcritIntTFTSE)
```

### Sentence-Final: SYN

```{r}
# Experiment 2 SYN in sentence-final region (Table 7)

## Syn effect in E2 sentence-final region FPRT 
VD07E2finalSynFPRTmn<-(259+271)/2-(286+274)/2
VD07E2finalSynFPRTtval<- 1.95 # reported in paper as non-sig
VD07E2finalSynFPRTSE<-VD07E2finalSynFPRTmn/VD07E2finalSynFPRTtval
VD07E2finalSynFPRTlower<-VD07E2finalSynFPRTmn-2*VD07E2finalSynFPRTSE
VD07E2finalSynFPRTupper<-VD07E2finalSynFPRTmn+2*VD07E2finalSynFPRTSE

## Syn effect in E2 sentence-final region RPD:
VD07E2finalSynRPDmn <- (1925+2131)/2-(1695+1806)/2 
VD07E2finalSynRPDtval <- sqrt(4.74) # F1 Table 10
VD07E2finalSynRPDSE <- VD07E2finalSynRPDmn/VD07E2finalSynRPDtval
VD07E2finalSynRPDlower <- round(VD07E2finalSynRPDmn - 2*VD07E2finalSynRPDSE)
VD07E2finalSynRPDupper <- round(VD07E2finalSynRPDmn + 2*VD07E2finalSynRPDSE)

## Syn effect in E2 sentence-final region TFT:
VD07E2finalSynTFTmn <- (349+360)/2-(362+373)/2
VD07E2finalSynTFTtval <- 1.95 # reported in paper as non-sig
VD07E2finalSynTFTSE<- VD07E2finalSynTFTmn/VD07E2finalSynTFTtval
VD07E2finalSynTFTlower <- round(VD07E2finalSynTFTmn - 2*VD07E2finalSynTFTSE)
VD07E2finalSynTFTupper <- round(VD07E2finalSynTFTmn + 2*VD07E2finalSynTFTSE)
```

### sentence-final: Sem

```{r}
# Experiment 2 Sem in sentence-final (Table 7)

## Sem effect in E2 sentence-final region FPRT
VD07E2finalSemFPRTmn<- (271+274)/2-(259+286)/2
VD07E2finalSemFPRTtval<- 1.95 # reported in paper as non-sig
VD07E2finalSemFPRTSE<-VD07E2finalSemFPRTmn/VD07E2finalSemFPRTtval
VD07E2finalSemFPRTlower<-VD07E2finalSemFPRTmn-2*VD07E2finalSemFPRTSE
VD07E2finalSemFPRTupper<-VD07E2finalSemFPRTmn+2*VD07E2finalSemFPRTSE

## Sem effect in E2 sentence-final region RPD:
VD07E2finalSemRPDmn <- (2131+1806)/2-(1925+1695)/2
VD07E2finalSemRPDtval <- sqrt(1.69) # F1 Table 10
VD07E2finalSemRPDSE <- VD07E2finalSemRPDmn/VD07E2finalSemRPDtval
VD07E2finalSemRPDlower <- round(VD07E2finalSemRPDmn - 2*VD07E2finalSemRPDSE)
VD07E2finalSemRPDupper <- round(VD07E2finalSemRPDmn + 2*VD07E2finalSemRPDSE)

## Sem effect in E2 sentence-final region TFT:
VD07E2finalSemTFTmn <- (360+373)/2-(349+362)/2
VD07E2finalSemTFTtval <- 1.95 # reported in paper as non-sig
VD07E2finalSemTFTSE <- VD07E2finalSemTFTmn/VD07E2finalSemTFTtval
VD07E2finalSemTFTlower <- round(VD07E2finalSemTFTmn - 2*VD07E2finalSemTFTSE)
VD07E2finalSemTFTupper <- round(VD07E2finalSemTFTmn + 2*VD07E2finalSemTFTSE)
```

### sentence-final: Int

```{r}
# Experiment 2 Int in sentence-final (Table 7)

## Int effect in E2 sentence-final region FPRT
VD07E2finalIntFPRTmn<- (271-259)-(274-286)
VD07E2finalIntFPRTtval<- 1.95 # reported in paper as non-sig
VD07E2finalIntFPRTSE<-VD07E2finalIntFPRTmn/VD07E2finalIntFPRTtval
VD07E2finalIntFPRTlower<-VD07E2finalIntFPRTmn-2*VD07E2finalIntFPRTSE
VD07E2finalIntFPRTupper<-VD07E2finalIntFPRTmn+2*VD07E2finalIntFPRTSE

## Int effect in E2 sentence-final region RPD:
VD07E2finalIntRPDmn <- (2131-1925)-(1806-1695)
VD07E2finalIntRPDtval <- sqrt(0.99) # F1 < 1, Table 10
VD07E2finalIntRPDSE <- VD07E2finalIntRPDmn/VD07E2finalIntRPDtval
VD07E2finalIntRPDlower <- round(VD07E2finalIntRPDmn - 2*VD07E2finalIntRPDSE)
VD07E2finalIntRPDupper <- round(VD07E2finalIntRPDmn + 2*VD07E2finalIntRPDSE)

## Int effect in E2 sentence-final region TFT:
VD07E2finalIntTFTmn <- (360-349)-(373-362)
VD07E2finalIntTFTtval <- 1.95 # reported in paper as non-sig
VD07E2finalIntTFTSE <-  VD07E2finalIntTFTmn/VD07E2finalIntTFTtval
VD07E2finalIntTFTlower <- round(VD07E2finalIntTFTmn - 2*VD07E2finalIntTFTSE)
VD07E2finalIntTFTupper <- round(VD07E2finalIntTFTmn + 2*VD07E2finalIntTFTSE)
```

## Experiment 3  
### precrit: SYN

```{r}
# Experiment 3  SYN in pre-crit

## almost sig. Syn effect in E3 precrit region FPRT:
VD07E3precritSynFPRTmn <- (462+447)/2-(449+455)/2
VD07E3precritSynFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3precritSynFPRTSE <- VD07E3precritSynFPRTmn/VD07E3precritSynFPRTtval
VD07E3precritSynFPRTlower <- round(VD07E3precritSynFPRTmn-2*VD07E3precritSynFPRTSE)
VD07E3precritSynFPRTupper <- round(VD07E3precritSynFPRTmn+2*VD07E3precritSynFPRTSE)

## almost sig. Syn effect in E3 precrit region RPD:
VD07E3precritSynRPDmn <- (490+551)/2-(472+493)/2
VD07E3precritSynRPDtval <- sqrt(6.59)
VD07E3precritSynRPDSE <- VD07E3precritSynRPDmn/VD07E3precritSynRPDtval
VD07E3precritSynRPDlower <- round(VD07E3precritSynRPDmn-2*VD07E3precritSynRPDSE)
VD07E3precritSynRPDupper <- round(VD07E3precritSynRPDmn+2*VD07E3precritSynRPDSE)

## almost sig. effect in E3 precrit region TFT:
VD07E3precritSynTFTmn <- (622+710)/2-(586+640)/2
VD07E3precritSynTFTtval <- sqrt(7.29)
VD07E3precritSynTFTSE <- VD07E3precritSynTFTmn/VD07E3precritSynTFTtval
VD07E3precritSynTFTlower <- round(VD07E3precritSynTFTmn-2*VD07E3precritSynTFTSE)
VD07E3precritSynTFTupper <- round(VD07E3precritSynTFTmn+2*VD07E3precritSynTFTSE)
```

### precrit: SEM

```{r}
# Experiment 3  SEM in pre-crit

## almost sig. Sem effect in E3 precrit region FPRT:
VD07E3precritSemFPRTmn <- (455+447)/2-(449+462)/2
VD07E3precritSemFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3precritSemFPRTSE <- VD07E3precritSemFPRTmn/VD07E3precritSemFPRTtval
VD07E3precritSemFPRTlower <- round(VD07E3precritSemFPRTmn-2*VD07E3precritSemFPRTSE)
VD07E3precritSemFPRTupper <- round(VD07E3precritSemFPRTmn+2*VD07E3precritSemFPRTSE)

## almost sig. Sem effect in E3 precrit region RPD:
VD07E3precritSemRPDmn <- (493+551)/2-(472+490)/2
VD07E3precritSemRPDtval <- sqrt(4.89)
VD07E3precritSemRPDSE <- VD07E3precritSemRPDmn/VD07E3precritSemRPDtval
VD07E3precritSemRPDlower <- round(VD07E3precritSemRPDmn-2*VD07E3precritSemRPDSE)
VD07E3precritSemRPDupper <- round(VD07E3precritSemRPDmn+2*VD07E3precritSemRPDSE)

## almost sig. Sem effect in E3 precrit region TFT:
VD07E3precritSemTFTmn <- (640+710)/2-(586+622)/2
VD07E3precritSemTFTtval <- sqrt(10.73)
VD07E3precritSemTFTSE <- VD07E3precritSemTFTmn/VD07E3precritSemTFTtval
VD07E3precritSemTFTlower <- round(VD07E3precritSemTFTmn-2*VD07E3precritSemTFTSE)
VD07E3precritSemTFTupper <- round(VD07E3precritSemTFTmn+2*VD07E3precritSemTFTSE)
```

### precrit: INT

```{r}
# Experiment 3  Int in pre-crit

## almost sig. Int effect in E3 precrit region FPRT:
VD07E3precritIntFPRTmn <- (447-462)-(455-449)
VD07E3precritIntFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3precritIntFPRTSE <- VD07E3precritIntFPRTmn/VD07E3precritIntFPRTtval
VD07E3precritIntFPRTlower <- round(VD07E3precritIntFPRTmn-2*VD07E3precritIntFPRTSE)
VD07E3precritIntFPRTupper <- round(VD07E3precritIntFPRTmn+2*VD07E3precritIntFPRTSE)

## almost sig. Int effect in E3 precrit region RPD:
VD07E3precritIntRPDmn <- (551-490)-(493-472)
VD07E3precritIntRPDtval <- 1.95 # reported in paper as non-sig
VD07E3precritIntRPDSE <- VD07E3precritIntRPDmn/VD07E3precritIntRPDtval
VD07E3precritIntRPDlower <- round(VD07E3precritIntRPDmn-2*VD07E3precritIntRPDSE)
VD07E3precritIntRPDupper <- round(VD07E3precritIntRPDmn+2*VD07E3precritIntRPDSE)

## almost sig. Int effect in E3 precrit region TFT:
VD07E3precritIntTFTmn <- (710-622)-(640-586)
VD07E3precritIntTFTtval <- 1.95 # reported in paper as non-sig
VD07E3precritIntTFTSE <- VD07E3precritIntTFTmn/VD07E3precritIntTFTtval
VD07E3precritIntTFTlower <- round(VD07E3precritIntTFTmn-2*VD07E3precritIntTFTSE)
VD07E3precritIntTFTupper <- round(VD07E3precritIntTFTmn+2*VD07E3precritIntTFTSE)
```

### crit: SYN
```{r}
# Experiment 3  SYN in crit

## almost sig. Syn effect in E3 crit region FPRT:
VD07E3critSynFPRTmn <- (294+280)/2-(274+282)/2
VD07E3critSynFPRTtval <- sqrt(3.95) # reported as <3.96
VD07E3critSynFPRTSE <- VD07E3critSynFPRTmn/VD07E3critSynFPRTtval
VD07E3critSynFPRTlower <- round(VD07E3critSynFPRTmn-2*VD07E3critSynFPRTSE)
VD07E3critSynFPRTupper <- round(VD07E3critSynFPRTmn+2*VD07E3critSynFPRTSE)

## almost sig. Syn effect in E3 crit region RPD:
VD07E3critSynRPDmn <- (354+344)/2-(314+315)/2
VD07E3critSynRPDtval <- sqrt(6.15)
VD07E3critSynRPDSE <- VD07E3critSynRPDmn/VD07E3critSynRPDtval
VD07E3critSynRPDlower <- round(VD07E3critSynRPDmn-2*VD07E3critSynRPDSE)
VD07E3critSynRPDupper <- round(VD07E3critSynRPDmn+2*VD07E3critSynRPDSE)

## almost sig. Syn effect in E3 crit region TFT:
VD07E3critSynTFTmn <- (451+467)/2-(414+421)/2
VD07E3critSynTFTtval <- 1.95 # reported in paper as non-sig
VD07E3critSynTFTSE <- VD07E3critSynTFTmn/VD07E3critSynTFTtval
VD07E3critSynTFTlower <- round(VD07E3critSynTFTmn-2*VD07E3critSynTFTSE)
VD07E3critSynTFTupper <- round(VD07E3critSynTFTmn+2*VD07E3critSynTFTSE)

```

### crit: Sem 
```{r}
# Experiment 3  SEM in crit

## almost sig. Sem effect in E3 crit region FPRT:
VD07E3critSemFPRTmn <- (282+280)/2-(274+294)/2
VD07E3critSemFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3critSemFPRTSE <- VD07E3critSemFPRTmn/VD07E3critSemFPRTtval
VD07E3critSemFPRTlower <- round(VD07E3critSemFPRTmn-2*VD07E3critSemFPRTSE)
VD07E3critSemFPRTupper <- round(VD07E3critSemFPRTmn+2*VD07E3critSemFPRTSE)

## almost sig. Sem effect in E3 crit region RPD:
VD07E3critSemRPDmn <- (315+344)/2-(314+354)/2
VD07E3critSemRPDtval <- 1.95 # reported in paper as non-sig
VD07E3critSemRPDSE <- VD07E3critSemRPDmn/VD07E3critSemRPDtval
VD07E3critSemRPDlower <- round(VD07E3critSemRPDmn-2*VD07E3critSemRPDSE)
VD07E3critSemRPDupper <- round(VD07E3critSemRPDmn+2*VD07E3critSemRPDSE)

## almost sig. Sem effect in E3 crit region TFT:
VD07E3critSemTFTmn <- (421+467)/2-(414+451)/2
VD07E3critSemTFTtval <- 1.95 # reported in paper as non-sig
VD07E3critSemTFTSE <- VD07E3critSemTFTmn/VD07E3critSemTFTtval
VD07E3critSemTFTlower <- round(VD07E3critSemTFTmn-2*VD07E3critSemTFTSE)
VD07E3critSemTFTupper <- round(VD07E3critSemTFTmn+2*VD07E3critSemTFTSE)
```

### crit: Int 
```{r}
# Experiment 3  Int in crit

## almost sig. Int effect in E3 crit region FPRT:
VD07E3critIntFPRTmn <- (280-294)-(282-274)
VD07E3critIntFPRTtval <- sqrt(3.34) # reported in paper as <3.35
VD07E3critIntFPRTSE <- VD07E3critIntFPRTmn/VD07E3critIntFPRTtval
VD07E3critIntFPRTlower <- round(VD07E3critIntFPRTmn-2*VD07E3critIntFPRTSE)
VD07E3critIntFPRTupper <- round(VD07E3critIntFPRTmn+2*VD07E3critIntFPRTSE)

## almost sig. Int effect in E3 crit region RPD:
VD07E3critIntRPDmn <- (344-354)-(315-314)
VD07E3critIntRPDtval <- 1.95 # reported in paper as non-sig
VD07E3critIntRPDSE <- VD07E3critIntRPDmn/VD07E3critIntRPDtval
VD07E3critIntRPDlower <- round(VD07E3critIntRPDmn-2*VD07E3critIntRPDSE)
VD07E3critIntRPDupper <- round(VD07E3critIntRPDmn+2*VD07E3critIntRPDSE)

## almost sig. Int effect in E3 crit region TFT:
VD07E3critIntTFTmn <- (467-451)-(421-414)
VD07E3critIntTFTtval <- 1.95 # reported in paper as non-sig
VD07E3critIntTFTSE <- VD07E3critIntTFTmn/VD07E3critIntTFTtval
VD07E3critIntTFTlower <- round(VD07E3critIntTFTmn-2*VD07E3critIntTFTSE)
VD07E3critIntTFTupper <- round(VD07E3critIntTFTmn+2*VD07E3critIntTFTSE)
```

### spill-over: SYN

```{r}
# Experiment 3  SYN in postcrit

## almost sig. Syn effect in E3 postcrit region FPRT:
VD07E3postcritSynFPRTmn <- (424+442)/2-(437+448)/2
VD07E3postcritSynFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3postcritSynFPRTSE <- VD07E3postcritSynFPRTmn/VD07E3postcritSynFPRTtval
VD07E3postcritSynFPRTlower <- round(VD07E3postcritSynFPRTmn-2*VD07E3postcritSynFPRTSE)
VD07E3postcritSynFPRTupper <- round(VD07E3postcritSynFPRTmn+2*VD07E3postcritSynFPRTSE)

## almost sig. Syn effect in E3 postcrit region RPD:
VD07E3postcritSynRPDmn <- (615+625)/2-(553+607)/2
VD07E3postcritSynRPDtval <- sqrt(2.78) # reported as <2.79
VD07E3postcritSynRPDSE <- VD07E3postcritSynRPDmn/VD07E3postcritSynRPDtval
VD07E3postcritSynRPDlower <- round(VD07E3postcritSynRPDmn-2*VD07E3postcritSynRPDSE)
VD07E3postcritSynRPDupper <- round(VD07E3postcritSynRPDmn+2*VD07E3postcritSynRPDSE)

## almost sig. Syn effect in E3 postcrit region TFT:
VD07E3postcritSynTFTmn <- (766+731)/2-(707+711)/2
VD07E3postcritSynTFTtval <- 1.95 # reported in paper as non-sig
VD07E3postcritSynTFTSE <- VD07E3postcritSynTFTmn/VD07E3postcritSynTFTtval
VD07E3postcritSynTFTlower <- round(VD07E3postcritSynTFTmn-2*VD07E3postcritSynTFTSE)
VD07E3postcritSynTFTupper <- round(VD07E3postcritSynTFTmn+2*VD07E3postcritSynTFTSE)

```

### spill-over: Sem 
```{r}
# Experiment 3  SEM in postcrit

## almost sig. Sem effect in E3 postcrit region FPRT:
VD07E3postcritSemFPRTmn <- (437+424)/2-(448+442)/2
VD07E3postcritSemFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3postcritSemFPRTSE <- VD07E3postcritSemFPRTmn/VD07E3postcritSemFPRTtval
VD07E3postcritSemFPRTlower <- round(VD07E3postcritSemFPRTmn-2*VD07E3postcritSemFPRTSE)
VD07E3postcritSemFPRTupper <- round(VD07E3postcritSemFPRTmn+2*VD07E3postcritSemFPRTSE)

## almost sig. Sem effect in E3 postcrit region RPD:
VD07E3postcritSemRPDmn <- (553+615)/2-(607+625)/2
VD07E3postcritSemRPDtval <- sqrt(2.78) # reported as <2.79
VD07E3postcritSemRPDSE <- VD07E3postcritSemRPDmn/VD07E3postcritSemRPDtval
VD07E3postcritSemRPDlower <- round(VD07E3postcritSemRPDmn-2*VD07E3postcritSemRPDSE)
VD07E3postcritSemRPDupper <- round(VD07E3postcritSemRPDmn+2*VD07E3postcritSemRPDSE)

## almost sig. Sem effect in E3 postcrit region TFT:
VD07E3postcritSemTFTmn <- (707+766)/2-(711+731)/2
VD07E3postcritSemTFTtval <- sqrt(6.15)
VD07E3postcritSemTFTSE <- VD07E3postcritSemTFTmn/VD07E3postcritSemTFTtval
VD07E3postcritSemTFTlower <- round(VD07E3postcritSemTFTmn-2*VD07E3postcritSemTFTSE)
VD07E3postcritSemTFTupper <- round(VD07E3postcritSemTFTmn+2*VD07E3postcritSemTFTSE)
```

### spill-over: Int 
```{r}
# Experiment 3  Int in postcrit

## almost sig. Int effect in E3 postcrit region FPRT:
VD07E3postcritIntFPRTmn <- (424-442)-(437-448)
VD07E3postcritIntFPRTtval <- 1.95 # reported in paper as non-sig
VD07E3postcritIntFPRTSE <- VD07E3postcritIntFPRTmn/VD07E3postcritIntFPRTtval
VD07E3postcritIntFPRTlower <- round(VD07E3postcritIntFPRTmn-2*VD07E3postcritIntFPRTSE)
VD07E3postcritIntFPRTupper <- round(VD07E3postcritIntFPRTmn+2*VD07E3postcritIntFPRTSE)

## almost sig. Int effect in E3 postcrit region RPD:
VD07E3postcritIntRPDmn <- (615-625)-(553-607)
VD07E3postcritIntRPDtval <- 1.95 # reported in paper as non-sig
VD07E3postcritIntRPDSE <- VD07E3postcritIntRPDmn/VD07E3postcritIntRPDtval
VD07E3postcritIntRPDlower <- round(VD07E3postcritIntRPDmn-2*VD07E3postcritIntRPDSE)
VD07E3postcritIntRPDupper <- round(VD07E3postcritIntRPDmn+2*VD07E3postcritIntRPDSE)

## almost sig. Int effect in E3 postcrit region TFT:
VD07E3postcritIntTFTmn <- (766-731)-(707-711)
VD07E3postcritIntTFTtval <- 1.95 # reported in paper as non-sig
VD07E3postcritIntTFTSE <- VD07E3postcritIntTFTmn/VD07E3postcritIntTFTtval
VD07E3postcritIntTFTlower <- round(VD07E3postcritIntTFTmn-2*VD07E3postcritIntTFTSE)
VD07E3postcritIntTFTupper <- round(VD07E3postcritIntTFTmn+2*VD07E3postcritIntTFTSE)
```

### sentence final: SYN

```{r}
# Experiment 3  SYN in sentence final region (Table 16 in VD07)

## almost sig. Syn effect in E3 final region FPRT:
VD07E3finalSynFPRTmn <- (463+483)/2-(480+464)/2
VD07E3finalSynFPRTtval <- sqrt(1.06) # reported in paper as below 1.07
VD07E3finalSynFPRTSE <- VD07E3finalSynFPRTmn/VD07E3finalSynFPRTtval
VD07E3finalSynFPRTlower <- round(VD07E3finalSynFPRTmn-2*VD07E3finalSynFPRTSE)
VD07E3finalSynFPRTupper <- round(VD07E3finalSynFPRTmn+2*VD07E3finalSynFPRTSE)

## almost sig. Syn effect in E3 final region RPD:
VD07E3finalSynRPDmn <- (2474+2068)/2-(2147+1875)/2
VD07E3finalSynRPDtval <- sqrt(2.79) # F1
VD07E3finalSynRPDSE <- VD07E3finalSynRPDmn/VD07E3finalSynRPDtval
VD07E3finalSynRPDlower <- round(VD07E3finalSynRPDmn-2*VD07E3finalSynRPDSE)
VD07E3finalSynRPDupper <- round(VD07E3finalSynRPDmn+2*VD07E3finalSynRPDSE)

## almost sig. Syn effect in E3 final region TFT:
VD07E3finalSynTFTmn <- (663+675)/2-(705+649)/2
VD07E3finalSynTFTtval <- sqrt(1.53) # reported in paper as below 1.54
VD07E3finalSynTFTSE <- VD07E3finalSynTFTmn/VD07E3finalSynTFTtval
VD07E3finalSynTFTlower <- round(VD07E3finalSynTFTmn-2*VD07E3finalSynTFTSE)
VD07E3finalSynTFTupper <- round(VD07E3finalSynTFTmn+2*VD07E3finalSynTFTSE)

```

### Sentence-final: Sem 
```{r}
# Experiment 3  SEM in final (Table 16 in VD07)

## Sem effect in E3 final region FPRT:
VD07E3finalSemFPRTmn <- (480+463)/2-(464+483)/2
VD07E3finalSemFPRTtval <- sqrt(1.06) # reported in paper as below 1.07
VD07E3finalSemFPRTSE <- VD07E3finalSemFPRTmn/VD07E3finalSemFPRTtval
VD07E3finalSemFPRTlower <- round(VD07E3finalSemFPRTmn-2*VD07E3finalSemFPRTSE)
VD07E3finalSemFPRTupper <- round(VD07E3finalSemFPRTmn+2*VD07E3finalSemFPRTSE)

## Sem effect in E3 final region RPD:
VD07E3finalSemRPDmn <- (2147+2474)/2-(1875+2068)/2
VD07E3finalSemRPDtval <- sqrt(9.26) # F1 (Table 19)
VD07E3finalSemRPDSE <- VD07E3finalSemRPDmn/VD07E3finalSemRPDtval
VD07E3finalSemRPDlower <- round(VD07E3finalSemRPDmn-2*VD07E3finalSemRPDSE)
VD07E3finalSemRPDupper <- round(VD07E3finalSemRPDmn+2*VD07E3finalSemRPDSE)

## Sem effect in E3 final region TFT:
VD07E3finalSemTFTmn <- (705+663)/2-(649+675)/2
VD07E3finalSemTFTtval <- sqrt(1.53) # reported in paper as below 1.54
VD07E3finalSemTFTSE <- VD07E3finalSemTFTmn/VD07E3finalSemTFTtval
VD07E3finalSemTFTlower <- round(VD07E3finalSemTFTmn-2*VD07E3finalSemTFTSE)
VD07E3finalSemTFTupper <- round(VD07E3finalSemTFTmn+2*VD07E3finalSemTFTSE)
```

### Sentence-Final: Int 
```{r}
# Experiment 3  Int in final (Table 16 in VD07)

## Int effect in E3 final region FPRT:
VD07E3finalIntFPRTmn <- (463-483)-(480-464)
VD07E3finalIntFPRTtval <- sqrt(1.06) # reported in paper as below 1.07
VD07E3finalIntFPRTSE <- VD07E3finalIntFPRTmn/VD07E3finalIntFPRTtval
VD07E3finalIntFPRTlower <- round(VD07E3finalIntFPRTmn-2*VD07E3finalIntFPRTSE)
VD07E3finalIntFPRTupper <- round(VD07E3finalIntFPRTmn+2*VD07E3finalIntFPRTSE)

## Int effect in E3 final region RPD:
VD07E3finalIntRPDmn <- (2474-2068)-(2147-1875)
VD07E3finalIntRPDtval <- sqrt(0.99) # reported in paper as below 1
VD07E3finalIntRPDSE <- VD07E3finalIntRPDmn/VD07E3finalIntRPDtval
VD07E3finalIntRPDlower <- round(VD07E3finalIntRPDmn-2*VD07E3finalIntRPDSE)
VD07E3finalIntRPDupper <- round(VD07E3finalIntRPDmn+2*VD07E3finalIntRPDSE)

## Int effect in E3 final region TFT:
VD07E3finalIntTFTmn <- (663-675)-(705-649)
VD07E3finalIntTFTtval <- sqrt(1.53) # reported in paper as below 1.54
VD07E3finalIntTFTSE <- VD07E3finalIntTFTmn/VD07E3finalIntTFTtval
VD07E3finalIntTFTlower <- round(VD07E3finalIntTFTmn-2*VD07E3finalIntTFTSE)
VD07E3finalIntTFTupper <- round(VD07E3finalIntTFTmn+2*VD07E3finalIntTFTSE)
```

# Prepare data frame
```{r}
VD07summary<-rbind(c(VD07E1critSynmn, VD07E1critSynlower, VD07E1critSynupper),
                   c(VD07E1critSemmn, VD07E1critSemlower, VD07E1critSemupper),
                   c(VD07E1critIntmn, VD07E1critIntlower, VD07E1critIntupper),

                   c(VD07E1spillSynmn, VD07E1spillSynlower, VD07E1spillSynupper),
                   c(VD07E1spillSemmn, VD07E1spillSemlower, VD07E1spillSemupper),
                   c(VD07E1spillIntmn, VD07E1spillIntlower, VD07E1spillIntupper),

                   c(VD07E1finalSynmn, VD07E1finalSynlower, VD07E1finalSynupper),
                   c(VD07E1finalSemmn, VD07E1finalSemlower, VD07E1finalSemupper),
                   c(VD07E1finalIntmn, VD07E1finalIntlower, VD07E1finalIntupper),

                   c(VD07E2critSynFPRTmn, VD07E2critSynFPRTlower, VD07E2critSynFPRTupper),
                   c(VD07E2critSynRPDmn, VD07E2critSynRPDlower, VD07E2critSynRPDupper),
                   c(VD07E2critSynTFTmn, VD07E2critSynTFTlower, VD07E2critSynTFTupper),

                   c(VD07E2critSemFPRTmn, VD07E2critSemFPRTlower, VD07E2critSemFPRTupper),
                   c(VD07E2critSemRPDmn, VD07E2critSemRPDlower, VD07E2critSemRPDupper),
                   c(VD07E2critSemTFTmn, VD07E2critSemTFTlower, VD07E2critSemTFTupper),

                   c(VD07E2critIntFPRTmn, VD07E2critIntFPRTlower, VD07E2critIntFPRTupper),
                   c(VD07E2critIntRPDmn, VD07E2critIntRPDlower, VD07E2critIntRPDupper),
                   c(VD07E2critIntTFTmn, VD07E2critIntTFTlower, VD07E2critIntTFTupper),

                   c(VD07E2postcritSynFPRTmn, VD07E2postcritSynFPRTlower, VD07E2postcritSynFPRTupper),
                   c(VD07E2postcritSynRPDmn, VD07E2postcritSynRPDlower, VD07E2postcritSynRPDupper),
                   c(VD07E2postcritSynTFTmn, VD07E2postcritSynTFTlower, VD07E2postcritSynTFTupper),

                   c(VD07E2postcritSemFPRTmn, VD07E2postcritSemFPRTlower, VD07E2postcritSemFPRTupper),
                   c(VD07E2postcritSemRPDmn, VD07E2postcritSemRPDlower, VD07E2postcritSemRPDupper),
                   c(VD07E2postcritSemTFTmn, VD07E2postcritSemTFTlower, VD07E2postcritSemTFTupper),

                   c(VD07E2postcritIntFPRTmn, VD07E2postcritIntFPRTlower, VD07E2postcritIntFPRTupper),
                   c(VD07E2postcritIntRPDmn, VD07E2postcritIntRPDlower, VD07E2postcritIntRPDupper),
                   c(VD07E2postcritIntTFTmn, VD07E2postcritIntTFTlower, VD07E2postcritIntTFTupper),

                   c(VD07E2finalSynFPRTmn, VD07E2finalSynFPRTlower, VD07E2finalSynFPRTupper),
                   c(VD07E2finalSynRPDmn, VD07E2finalSynRPDlower, VD07E2finalSynRPDupper),
                   c(VD07E2finalSynTFTmn, VD07E2finalSynTFTlower, VD07E2finalSynTFTupper),

                   c(VD07E2finalSemFPRTmn, VD07E2finalSemFPRTlower, VD07E2finalSemFPRTupper),
                   c(VD07E2finalSemRPDmn, VD07E2finalSemRPDlower, VD07E2finalSemRPDupper),
                   c(VD07E2finalSemTFTmn, VD07E2finalSemTFTlower, VD07E2finalSemTFTupper),

                   c(VD07E2finalIntFPRTmn, VD07E2finalIntFPRTlower, VD07E2finalIntFPRTupper),
                   c(VD07E2finalIntRPDmn, VD07E2finalIntRPDlower, VD07E2finalIntRPDupper),
                   c(VD07E2finalIntTFTmn, VD07E2finalIntTFTlower, VD07E2finalIntTFTupper),

                   c(VD07E3precritSynFPRTmn, VD07E3precritSynFPRTlower, VD07E3precritSynFPRTupper),
                   c(VD07E3precritSynRPDmn, VD07E3precritSynRPDlower, VD07E3precritSynRPDupper),
                   c(VD07E3precritSynTFTmn, VD07E3precritSynTFTlower, VD07E3precritSynTFTupper),

                   c(VD07E3precritSemFPRTmn, VD07E3precritSemFPRTlower, VD07E3precritSemFPRTupper),
                   c(VD07E3precritSemRPDmn, VD07E3precritSemRPDlower, VD07E3precritSemRPDupper),
                   c(VD07E3precritSemTFTmn, VD07E3precritSemTFTlower, VD07E3precritSemTFTupper),

                   c(VD07E3precritIntFPRTmn, VD07E3precritIntFPRTlower, VD07E3precritIntFPRTupper),
                   c(VD07E3precritIntRPDmn, VD07E3precritIntRPDlower, VD07E3precritIntRPDupper),
                   c(VD07E3precritIntTFTmn, VD07E3precritIntTFTlower, VD07E3precritIntTFTupper),

                   c(VD07E3critSynFPRTmn, VD07E3critSynFPRTlower, VD07E3critSynFPRTupper),
                   c(VD07E3critSynRPDmn, VD07E3critSynRPDlower, VD07E3critSynRPDupper),
                   c(VD07E3critSynTFTmn, VD07E3critSynTFTlower, VD07E3critSynTFTupper),

                   c(VD07E3critSemFPRTmn, VD07E3critSemFPRTlower, VD07E3critSemFPRTupper),
                   c(VD07E3critSemRPDmn, VD07E3critSemRPDlower, VD07E3critSemRPDupper),
                   c(VD07E3critSemTFTmn, VD07E3critSemTFTlower, VD07E3critSemTFTupper),

                   c(VD07E3critIntFPRTmn, VD07E3critIntFPRTlower, VD07E3critIntFPRTupper),
                   c(VD07E3critIntRPDmn, VD07E3critIntRPDlower, VD07E3critIntRPDupper),
                   c(VD07E3critIntTFTmn, VD07E3critIntTFTlower, VD07E3critIntTFTupper),

                   c(VD07E3postcritSynFPRTmn, VD07E3postcritSynFPRTlower, VD07E3postcritSynFPRTupper),
                   c(VD07E3postcritSynRPDmn, VD07E3postcritSynRPDlower, VD07E3postcritSynRPDupper),
                   c(VD07E3postcritSynTFTmn, VD07E3postcritSynTFTlower, VD07E3postcritSynTFTupper),

                   c(VD07E3postcritSemFPRTmn, VD07E3postcritSemFPRTlower, VD07E3postcritSemFPRTupper),
                   c(VD07E3postcritSemRPDmn, VD07E3postcritSemRPDlower, VD07E3postcritSemRPDupper),
                   c(VD07E3postcritSemTFTmn, VD07E3postcritSemTFTlower, VD07E3postcritSemTFTupper),

                   c(VD07E3postcritIntFPRTmn, VD07E3postcritIntFPRTlower, VD07E3postcritIntFPRTupper),
                   c(VD07E3postcritIntRPDmn, VD07E3postcritIntRPDlower, VD07E3postcritIntRPDupper),
                   c(VD07E3postcritIntTFTmn, VD07E3postcritIntTFTlower, VD07E3postcritIntTFTupper),

                   c(VD07E3finalSynFPRTmn, VD07E3finalSynFPRTlower, VD07E3finalSynFPRTupper),
                   c(VD07E3finalSynRPDmn, VD07E3finalSynRPDlower, VD07E3finalSynRPDupper),
                   c(VD07E3finalSynTFTmn, VD07E3finalSynTFTlower, VD07E3finalSynTFTupper),

                   c(VD07E3finalSemFPRTmn, VD07E3finalSemFPRTlower, VD07E3finalSemFPRTupper),
                   c(VD07E3finalSemRPDmn, VD07E3finalSemRPDlower, VD07E3finalSemRPDupper),
                   c(VD07E3finalSemTFTmn, VD07E3finalSemTFTlower, VD07E3finalSemTFTupper),

                   c(VD07E3finalIntFPRTmn, VD07E3finalIntFPRTlower, VD07E3finalIntFPRTupper),
                   c(VD07E3finalIntRPDmn, VD07E3finalIntRPDlower, VD07E3finalIntRPDupper),
                   c(VD07E3finalIntTFTmn, VD07E3finalIntTFTlower, VD07E3finalIntTFTupper)
            )

VD07expt <- c(rep("VD07.E1 (N=35)", 9), rep("VD07.E2 (N=36)", 27), rep("VD07.E3 (N=40)", 36))

VD07region <- c(rep("critical", 3), rep("spill-over", 3), rep("sentence-final", 3),
                rep("critical", 9), rep("spill-over", 9), rep("sentence-final", 9),
                rep("pre-critical", 9), rep("critical", 9), rep("spill-over", 9), rep("sentence-final", 9)
                )

VD07int_type <- c(rep(c("syntactic", "semantic", "interaction"), 3),

                  rep(c("syntactic", "syntactic", "syntactic", 
                        "semantic", "semantic", "semantic",
                        "interaction", "interaction", "interaction"), 7)
                  )

VD07measure <- c(rep("SPR", 9), rep(c("FPRT", "RPD","TFT"), 21))

VD07estimates <- as.data.frame(cbind(VD07summary, 
                                     VD07expt,
                                     VD07region,
                                     VD07int_type,
                                     VD07measure))

colnames(VD07estimates)<-c("mean", "lower" ,"upper",  "experiment", "region", "int_type","measure")

VD07estimates$mean <- as.numeric(VD07estimates$mean)
VD07estimates$lower <- as.numeric(VD07estimates$lower)
VD07estimates$upper <- as.numeric(VD07estimates$upper)

VD07estimates$experiment <- factor(VD07estimates$experiment,levels=c("VD07.E1 (N=35)", "VD07.E2 (N=36)","VD07.E3 (N=40)"))

VD07estimates$measure <- factor(VD07estimates$measure,levels=c("SPR","FPRT","RPD","TFT"))

VD07estimates$region <- factor(VD07estimates$region,levels=c("pre-critical","critical", "spill-over", "sentence-final"))

VD07estimates$int_type <- factor(VD07estimates$int_type,levels=c("syntactic","semantic", "interaction"))

str(VD07estimates)
```

# Mertzen et al 2021:

```{r}
load("M21_byregionsummaries_de.rda")

M21.G <- d
M21.G$experiment <- "M21.G (N=121)"

load("M21_byregionsummaries_en.rda")
M21.E <- d
M21.E$experiment <- "M21.E (N=61)"

M21estimates <- rbind(M21.G, M21.E[,-7])

M21estimates$int_type <- ifelse(M21estimates$effect%in%c("Subject"),"syntactic",
                              ifelse(M21estimates$effect%in%c("Animacy"),"semantic", 
                                   "interaction"))

M21estimates$region <- as.factor(M21estimates$region)
levels(M21estimates$region)[3] <- "spill-over"

M21estimates <- M21estimates[,-4]
M21estimates <- M21estimates[,-5]
M21estimates <- M21estimates[,-6]

```

# Plot previous findings
```{r}
my_colors <- RColorBrewer::brewer.pal(11, "RdBu")

previous <- rbind(VD07estimates, M21estimates)
save(previous, file="VD07_M21_estimates.rda")

previous_crit.plot <- ggplot(subset(previous, previous$region%in%c("pre-critical","critical")),
                   aes(x=experiment, y=mean, ymin=lower, ymax=upper )) +
  geom_pointrange(aes(color=experiment, shape=measure), 
             size=0.5, position=position_dodge(0.9))+
  scale_shape_manual(values = c(18,16,17,15))+
  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9]))+
  geom_hline(yintercept=0, linetype="solid", alpha=.4)+
  ggh4x::facet_grid2(region ~ int_type, scales="free_y", independent = "y") +
  theme_bw(base_size = 10)+
  scale_x_discrete(breaks = NULL)+
  xlab("")+
  ylab("reading time difference (ms)")+
  #ggtitle("Previous findings")+
  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))

previous_crit.plot
ggsave("plots/previous_findings.jpg", width=6, height=5, dpi=600)
```

# Plot present and previous findings
```{r, eval=FALSE}
my_colors <- RColorBrewer::brewer.pal(11, "RdBu")

# prepare Schoknecht & Vasishth estimates (numbers from script "Plot_stats.R")
SVestimates <- data.frame(mean   = c(1, 11, 7, 2, 8, 14),
                          lower  = c(0, 8, 1, 0, 4, 7),
                          upper = c(2, 14, 13, 4, 12, 22),
                           region = c(rep(c("pre-critical"),3), rep(c("critical"), 3)),
                           measure  = c(rep(c("SPR"), 6)),
                           experiment = c(rep(c("present (N=774)"), 6)),
                           int_type = c(rep(c("syntactic", "semantic", "interaction"), 2))
                          )

# combine with previous estimates
estimates <- rbind(previous, SVestimates)
#estimates$experiment <- factor(estimates$experiment,levels=c("VD07.E1", "VD07.E2","VD07.E3", "present", "M21.G","M21.E"))

pp.plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical")),
                   aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
    geom_pointrange(aes(color=experiment, shape=measure), 
             size=0.8, position=position_dodge(1.45))+
  #scale_y_continuous(breaks = c(-50,-40,-30,-20, -10, 0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100))+
  scale_shape_manual(values = c(18,16,17,15))+
  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
  geom_hline(yintercept=0, linetype="solid", alpha=.5)+
  facet_grid(region~int_type, scales="free", space="free") +
  theme_bw(base_size = 15)+
  scale_x_discrete(breaks = NULL)+
  theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
  xlab("")+
  ylab("reading time difference (ms)")+
  ggtitle("Present and previous findings")+
  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))

pp.plot + coord_cartesian(ylim=c(-10, 30))  
ggsave("plots/present_previous_findings.jpg", width=8, height=8, dpi=600)

```

### 2024-05-29 code_data/previous_studies/compute_BFs_PreviousStudies.R
Added via the Open Science Framework

@@ -1,0 +1,379 @@
+# author: Pia Schoknecht
+# date: 21.05.2024
+
+# script to compute Bayes Factors for the previous studies
+
+# Mertzen et al., (2024) English and German eye-tracking data
+# Van Dyke (2007) English eye-tracking and self-paced reading data
+# for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
+
+
+library(tidyverse)
+library(lme4)
+library(brms)
+library(ggplot2)
+library(bayesplot)
+library(truncnorm)
+library(ggbreak)
+library(ggimage) 
+library(cowplot)
+
+
+# load Mertzen et al. GERMAN data
+df <- read.table("dataRetroDE_processed.txt")
+
+df_crit <- df %>% filter(region == "crit") %>%
+                  select(subject, trial, list, item, condition, roi, region, TFT) %>%
+                  filter(TFT != 0)
+
+# contrast coding
+df_crit$syn <- ifelse(df_crit$condition%in%c("a","b"),-0.5,0.5)
+df_crit$sem <- ifelse(df_crit$condition%in%c("b","d"),0.5,-0.5)
+
+# trim all rts below 150 and above 3000 ms
+range(df_crit$TFT)
+crit_trim <- subset(df_crit, df_crit$TFT > 80)
+range(crit_trim$TFT)
+
+
+# Inferential statistics
+# truncated varying priors 
+# (only positive effects -- longer reading times for high interference)
+
+priors_s_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd),
+  prior(lkj(2), class = cor)
+)  
+
+priors_m_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd),
+  prior(lkj(2), class = cor)
+) 
+
+priors_l_tr <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.5), class = sigma),
[…]

### 2024-05-29 code_data/previous_studies/compute_BFs_PreviousStudies.R
Updated via the Open Science Framework

@@ -301,16 +301,16 @@ df.bf_mertzen$Prior2 <- ifelse(df.bf_mertzen$Prior == "Normal(0, 0.01)", "N+(0, 
 df.bf_mertzen$BF10 <- as.numeric(df.bf_mertzen$BF10)
 df.bf_mertzen$effect <- factor(df.bf_mertzen$Effect, levels=c("syntactic", "semantic", "interaction"))
 
-ggplot(df.bf_mertzen, aes(x = Prior, y = BF10, group = effect)) +
+ggplot(df.bf_mertzen, aes(x = Prior2, y = BF10, group = effect)) +
   geom_point(aes(color=effect)) +
   geom_line(aes(color=effect)) +
   facet_grid(.~Language)+
   geom_hline(yintercept = 1, linetype="dashed") +
-  theme_bw(base_size = 12)+
+  theme_bw(base_size = 10)+
   xlab("Truncated Prior")+
   theme(legend.position = "top")
 
-ggsave("BF_plot_mertzen.png", width = 14, height = 8, units = "cm", dpi=300)
+ggsave("BF_plot_mertzen.png", width = 11, height = 6, units = "cm", dpi=300)
 
 
 #####################################################################################
@@ -323,6 +323,8 @@ load("VD07_M21_estimates.rda")
 
 # as a test case, we first try this approach with the Mertzen data, for which we properly computed the BFs above
 # note: This is done on the ms scale!
+
+#### German ####
 (MG_tft <- previous %>% filter(experiment == "M21.G (N=121)") %>% filter(region == "critical") %>% filter(measure == "TFT"))
 
 # simulate data for a reasonable prior (effect between 0 and 40 ms) and the estimates of Mertzen
@@ -377,3 +379,57 @@ df.bf_mertzen %>% filter(Prior == "Normal(0, 0.05)")
 
 # --> Evidence for/against an effect is the same (BFs above or below 1), 
 # but the magnitude of the BF for the syntactic effect is very off
+
+#### English ####
+(ME_tft <- previous %>% filter(experiment == "M21.E (N=61)") %>% filter(region == "critical") %>% filter(measure == "TFT"))
+
+# simulate data for a reasonable prior (effect between 0 and 40 ms) and the estimates of Mertzen
+set.seed(2)
+prior_samples <- data.frame(b = round(rtruncnorm(n = 100000, a=0, mean = 15, sd = 5.5),1))
[…]

### 2024-06-11 code_data/plausibility_norming/plausibility_report_HighvsLowSem.Rmd
Add files via upload

---
title: "Plausibility Norming"
author: "Pia Schoknecht"
date: "17 06 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lme4)
library(ordinal)
library(brms)
```

# Load and check
```{r load}
# load data
df <- read.csv("norming_data_HighvsLowSem.csv")

#dfp_ex <- dfp  %>% filter(subject !="plausib_anim19" & subject !="plausib_anim1")

uniq <- df[!duplicated(df[2]),]
(N_subj <- length(unique(df$subject)))

xtabs(~ latin_square_list, data=df)/160
```

# Ratings
```{r}
df$condition <- factor(df$condition, levels = c("c", "d", "f"))

df %>% group_by(condition) %>%
             dplyr::summarize(mean_rating = mean(value, na.rm=TRUE), 
                              sd_rating = sd(value, na.rm=TRUE))

ggplot(df, aes(x=condition, y=value))+
  geom_jitter(color="lightblue", alpha=0.2)+
  geom_boxplot(fill="grey", alpha=0.7)+
  stat_summary(fun=mean, geom="point", shape=18, size=4) + #adding the mean
  expand_limits(y=c(0.5,7.5))+
  scale_y_continuous(name="plausibility rating", breaks=seq(1, 7, 1))+
  scale_x_discrete(name="", labels=c("c" = "low semantic\ninterference condition", "d" = "high semantic\ninterference condition", "f" = "implausible filler"))+
  theme_bw()+
  labs(title = "Plausibility ratings",
      subtitle = "1 = absolutely implausible; 7 = absolutely plausible")

ggsave("plausibility_ratings.jpg", dpi=600, width=4, height=4)

```

# t-test
```{r}

test <- df %>% filter(condition != "f") %>%
             group_by(subject, condition) %>%
             dplyr::summarize(mean_rating = mean(value))

test_wider <- test %>% pivot_wider(names_from = condition, values_from = mean_rating)

t.test(test_wider$c, test_wider$d, paired=TRUE)

```

# Ratings per participant and per item
```{r scale_p}
N <- length(df$value)

p_scale_use <- df %>%
             group_by(subject, condition) %>%
             dplyr::summarize(mean_rating = mean(value, na.rm=TRUE), 
                              sd_rating = sd(value, na.rm=TRUE), 
                              se_rating = sd_rating / sqrt(N))

ggplot(p_scale_use, aes(x=condition, y=mean_rating, colour=subject)) +
  geom_point(size=1, position = position_dodge(width = 0.3))+
geom_errorbar(aes(ymin=mean_rating-se_rating, ymax=mean_rating+se_rating),
              width=1, position= position_dodge(width = 0.3)) +  
  facet_wrap(.~subject, nrow=4)+
  xlab("condition")+
  ylab("mean_rating")+
  theme_bw()+
  theme(legend.position="none")+
  ggtitle("Pausibility Ratings per subject \n 7 = absolutely plausible, 1 = absolutely implausible \n errorbars = SE")
ggsave("scale_use.jpg", dpi=600, width=5)

#### PER ITEM
p_scale_use <- df %>% filter(condition != "f") %>%
             group_by(item, condition) %>%
             dplyr::summarize(mean_rating = mean(value, na.rm=TRUE), 
                              sd_rating = sd(value, na.rm=TRUE), 
                              se_rating = sd_rating / sqrt(N))

write.csv(p_scale_use, "plausib_rating_semantics.csv")

p_scale_use$condition <- factor(p_scale_use$condition, levels = c("d", "c"))

items_wide <- p_scale_use %>% select(-sd_rating, -se_rating) %>%
                          pivot_wider(names_from = condition, values_from = mean_rating) %>%
                          mutate(diff = c-d)

ggplot(p_scale_use, aes(x=item, y=mean_rating, colour=condition, group=item)) +
  geom_point(size=1, position = position_dodge(width = 0.3))+
  geom_line(color="grey")+
geom_errorbar(aes(ymin=mean_rating-se_rating, ymax=mean_rating+se_rating),
              width=1, position= position_dodge(width = 0.3)) +  
  xlab("item")+
  ylab("mean_rating")+
  scale_color_discrete(name="", labels=c("c" = "low semantic\ninterference condition", "d" = "high semantic\ninterference condition", "f" = "implausible filler"))+
  theme_bw()+
  theme(legend.position="bottom")+
  ggtitle("Pausibility Ratings per item \n 7 = absolutely plausible, 1 = absolutely implausible \n errorbars = SE")

```

# Does the plausibility affect reading times in the pre-critical region?
```{r}
precrit <- read.csv("../SPR/data/pandora_spr_774_precrit.csv")

precrit_plausib <- inner_join(precrit, p_scale_use, by=c("item", "condition"))

# null model
m_null <- lmer(log(rt) ~ 1 +
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

# raw plausibility rating as predictor
m_plausib <- lmer(log(rt) ~ mean_rating +
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_plausib)

# semantic interference as predictor
m_sem <- lmer(log(rt) ~ sem +
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_sem)

# semantic interference AND plausibility as predictors
m_sem.plausib <- lmer(log(rt) ~ sem + scale(mean_rating)+
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_sem.plausib)

# do they interact?
m_interaction <- lmer(log(rt) ~ sem*scale(mean_rating)+
                       (1 |fullname)+  
                       (1 |item), 
                     data=precrit_plausib, REML=FALSE)

summary(m_interaction)
```

# look at reading times for high vs. low plausibility and high vs. low semantic interference

```{r}
# dichotomize plausibility ratings
precrit_plausib$plausibility_cat <- ifelse(precrit_plausib$mean_rating > 5.9, "high",
                                                ifelse(precrit_plausib$mean_rating < 5.5, "low", "med"))

table(precrit_plausib$plausibility_cat)

precrit_plausib %>% filter(plausibility_cat != "med") %>%
                      group_by(plausibility_cat) %>%
                      dplyr::summarize(mean_rt = mean(rt, na.rm=TRUE))
```

```{r}
precrit_plausib %>% filter(plausibility_cat != "med") %>%
                      group_by(sem) %>%
                      dplyr::summarize(mean_rt = mean(rt, na.rm=TRUE))

```

```{r}
precrit_plausib %>% filter(plausibility_cat != "med") %>%
                      group_by(sem, plausibility_cat) %>%
                      dplyr::summarize(mean_rt = mean(rt, na.rm=TRUE))

```

### 2024-06-11 code_data/plausibility_norming/plausibility_report_HighvsLowSem.Rmd
edits to plausibility code

@@
 
+head(df)
+
 #dfp_ex <- dfp  %>% filter(subject !="plausib_anim19" & subject !="plausib_anim1")
@@
 (N_subj <- length(unique(df$subject)))
+
+## SV: isn't this simpler than your obscure code above?
+length(unique(df$md5_hash_participant_IP))
+
+colnames(df)
 
@@
 
+## SV: this calculation is incorrect as it is not aggregated first:
 df %>% group_by(condition) %>%
@@
 
+## SV: redone so I can understand it
+xtabs(~subject+condition,df_crit)
+
+df_crit<-subset(df,condition!="f")
+
+xtabs(~subject+condition,df_crit)
+
+xtabs(~item+condition,df_crit)
+
+## c is low, d is high int:
+df_crit$cond<-ifelse(df_crit$condition=="c",0.5,-0.5)
+
+library(lme4)
+m<-lmer(value~cond + (1+cond|subject) + (1+cond|item),df_crit)
+## low semantic int conditions are more plausible than high:
+summary(m)
+library(car)
+## normal approximation seems OK:
+qqPlot(residuals(m))
+
+## looks OK:
+acf(residuals(m))
+
+## SV: why are we doing this t-test at all?
 test_wider <- test %>% pivot_wider(names_from = condition, values_from = mean_rating)
 
+head(test_wider)
+
+## SV: You can see that this t-test is over-enthusiastic because it
+## hides significant by-item variance by averaging it away (see the lmer
+## results above--by item variability is remarkably high)
 t.test(test_wider$c, test_wider$d, paired=TRUE)
@@
 # raw plausibility rating as predictor
+library(MASS)
+m_lm<-lm(rt~mean_rating,precrit_plausib)
+
 m_plausib <- lmer(log(rt) ~ mean_rating +
@@
                      data=precrit_plausib, REML=FALSE)
+```
+
+```{r}
+summary(m_plausib)
+## strong skew:
+qqPlot(residuals(m_plausib)
+### massive autocorrelation:       
+acf(residuals(m_plausib))
+
+m_plausib <- lmer(log(rt) ~ scale(mean_rating,scale=FALSE) +
+                       (1 + scale(mean_rating,scale=FALSE)|fullname)+  
+                       (1 + scale(mean_rating,scale=FALSE)|item), 
+                     data=precrit_plausib, REML=FALSE)
 
 summary(m_plausib)
+### massive autocorrelation remains. Why?       
+acf(residuals(m_plausib))
+```
+
+SV: here, instead of fitting to mean ratings only, can you store the mean by-item ratings as well as the SE in the data frame used for the inner join above? Then center the mean ratings. 
+So, the data frame used in the model m_plausib should then have not just the centered mean_rating but also the SE. Then the measurement error model would be (possible in brms only):
+
+```
+mbrms<-brm(formula = rt ~ me(c_mean_rating, se_mean_rating) +
+(1 | fullname) + (1|item),
+data = precrit_plausib, family = lognormal(), prior = some_uninf_priors,
+iter = 2000, chains = 4,
+control = list(adapt_delta = 0.999,
+max_treedepth = 15))
+```
+
+SV: Then use the measurement error syntax in the model below where you compare sem with ratings.
 

### 2024-06-22 paper/main.tex
Updated via the Open Science Framework

@@
 \usepackage{bigdelim}
+
+\usepackage{clipboard}
+\newclipboard{output-reviews}
+\openclipboard{output-reviews}
 
@@
 
-All experiments used the 2 $\times$ 2 design developed by \textcite{vandyke07}; the design crosses syntactic and semantic interference.
+\Copy{vd_design}{All experiments used the 2 $\times$ 2 design developed by \textcite{vandyke07}; the design crosses syntactic and semantic interference.} \label{vd_design}
 
@@
 During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference.
+\Copy{semantic_features_encoding}{\textcolor{blue}{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained.
+In \ref{ex:materials}, the semantic retrieval cue might be \{+animate\} or \{+human\} but it might as well be \{+can drink\}.
+While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant.
+The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval.
+A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
+}}\label{semantic_features_encoding}
 
 \paragraph{Regions of interest in the design}
-
-Our principal focus was on the processing of the verb (\textit{trank}, `drank').
-The reading times of the immediately following spill-over region and of the pre-critical region were also of interest.
+\textcolor{blue}{For both our SPR and EEG experiments, o}ur principal focus was on the processing of the verb (\textit{trank}, `drank').
+\textcolor{blue}{In the SPR experiment, t}he reading times of the immediately following spill-over region and of the pre-critical region were also of interest.
 Previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in the pre-critical region (we return to this point in the discussion sections).
@@
 
-Thus, our focus in the present study is on the processing difficulty observed at the critical verb and the surrounding regions.
-
-\subsubsection{Plausibility norming}
+Thus, our focus in the \textcolor{blue}{SPR experiment} is on the processing difficulty observed at the critical verb and the surrounding regions.
+
+\Copy{why_n400}{\textcolor{blue}{In the EEG experiment, the region of interest is the critical verb.
+Due to the direct measurement of brain activity with excellent temporal resolution event-related potential (ERP) effects do not tend to spill-over into later regions.
+We focus specifically on the N400 elicited by the verb.
+The N400 is a negative deflection in the ERP signal that peaks approximately 400 ms after any word was encountered.
+Traditionally, it is associated with lexical-semantic processing, but the finding of multi-modal N400 effects suggests that it rather indexes differences in the processing of meaningful stimuli in general \citep{kutas_federmeier2011}.
+The N400 amplitude is greater for words that deviate from the meaning activated by its context compared to words that fit its context \citep[see, e.g.,][]{kutas_hillyard, Kutas_Iragui_1998}.
+Consequently, the amplitude of the N400 to a specific word is a negative function of its predictability within the given context \citep{kutas_hillyard1984}.
+Additionally, the N400 elicited by isolated words is negatively correlated with their frequency \citep{kutas&federmeier_2000}.
+Therefore, the N400 amplitude is thought to index a) the increased difficulty to integrate a semantically anomalous or simply unexpected word into the sentence context \citep[e.g.,][]{hagoort2004integration} and/or b) the necessity to retrieve lexical information from memory which was not already pre-activated before the stimulus was perceived \citep[e.g.,][]{brouwer2017_n4_p6}.
+These two accounts of the N400 have been extensively discussed in the ERP literature on prediction \citep[see e.g.,][]{nieuwland_etal_2018, nicenboim_etal_2020, Freunberger_Roehm_2017, Delong_etal_2005, mantegna2019, nieuwland2019_integrationprediction}.
+As a likely marker of memory retrieval during sentence comprehension, it is fair to assume that the N400 would show modulations due to retrieval interference.
+The few ERP studies that have investigated retrieval interference so far have indeed found negativities \citep{lee_garnsey, martinetal2014, schoknecht2022, vasishth_drenhaus_2011}, although the topography and latency of these negativities was mostly not typical for the N400.
+It is important to highlight here that the previous ERP studies on interference either increased the number of distractors in a sentences or used grammatical gender or number as the critical retrieval cue.
+So, interference due to non-semantic features elicited negativities which were more or less similiar to the N400.
+Therefore, we expect the interference manipulation regardless of the type of retrieval cue (syntactic or semantic) to elicit N400 or similar effects.
+For a discussion of the relation between the N400 and other negativities, see \textcite{ina_n400}.}
+
+\textcolor{blue}{Alternately, interference -- especially syntactic interference -- might modulate the P600.
+The P600 is a positive deflection in the ERP wave form that is typically observed at least 600 ms after stimulus onset.
+It is associated with syntactic processing, e.g., morpho-syntactic errors and reanalysis \citep[see, e.g.,][]{osterhout&holcomb_1992, kaan_etal_2000_P600}.
+High interference could increase difficulty during syntactic integration of subject and verb in our materials.
+Potential misretrievals of the distractor might lead to reanalysis processes which could manifest as an increased P600 amplitude compared to low interference conditions.
+To anticipate our results here, visual inspection of the ERPs in Figure \ref{fig:erp_syn_ERP} does not indicate syntactic interference effects in the P600 time window.}} \label{why_n400}
+
+\subsubsection{Plausibility norming} \label{plausib_normin_orig}
 \citeauthor{vandyke07} conducted a pre-test to ensure that the distractor in high semantic interference conditions was a sufficiently probable subject of the critical verb, i.e., that it could potentially induce semantic interference.
@@
 
+\textcolor{red}{\section{Data availability statement}
+All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}\label{statement}
+
 \section{Experiments 1a and 1b: SPR}
@@
 
-In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
+\Copy{fillers1}{In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
 The experimental sessions were separated by 1 - 20 days.
 The procedure of the sessions was identical.
-Each session lasted approximately 25 minutes.
+\textcolor{blue}{In session 1, the experimental items 1-60 were presented interspersed with 40 filler sentences.
+In session 2, the experimental items 61-120 were presented with another set of 40 filler sentences.
+The fillers were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers1} Each session lasted approximately 25 minutes.
 
@@
 
-The Bayes factor is analogous to the frequentist ANOVA or likelihood ratio test: it compares the marginal likelihood of a full model to that of a baseline null model; this is often written as $BF_{10}$, with the subscripts indicating that marginal likelihood of the full model is in the numerator and that of the null model is in the denominator.
-One major advantage of the Bayes factor over the frequentist analogue is that it takes the uncertainty of the parameters into account.
+\textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method.
+A Bayesian hypothesis test for which the posterior density is devided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$.
+\textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account.
 This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
@@
 Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest.
-If BF$_{10}$ < 1, it provides evidence in favor of the null model.
-The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence for the null.
+If BF$_{10}$ < 1, it provides evidence \textcolor{blue}{against the effect of interest}.
+The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence \textcolor{blue}{against the effect of interest}.
 In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
 For this reason, models were run with four chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.
-Bayes factors were unstable when varying slopes were included; for this reason, we fit the models only with varying intercepts by participant and by item.
 
@@
 Sessions were separated by one to eight weeks for each participant (the median gap was two weeks, the first and third quartiles being one and three weeks).
-The procedure of both sessions was identical.
-
-During each experimental session, the EEG was recorded while participants were
-seated in a sound-proof booth.
+\Copy{fillers2}{The procedure of both sessions was identical \textcolor{blue}{and the same lists were used as in Experiment 1, with 60 critical sentences and 40 fillers per session}.
+
+During each experimental session, the EEG was recorded while participants were seated in a sound-proof booth.
 OpenSesame was used to present sentences word-by-word \citep{opensesame}.
 Participants were familiarized with the procedure with two practice sentences.
-After that, the experiment was conducted in four blocks of 50 sentences each, presenting the items in pseudorandomized order, with breaks between the blocks.
-Each trial started with the presentation of a fixation cross in the center of the screen for 500 ms.
+After that, the experiment was conducted in four blocks of \textcolor{blue}{25} sentences each, presenting the items in pseudorandomized order, with breaks between the blocks.}\label{fillers2} Each trial started with the presentation of a fixation cross in the center of the screen for 500 ms.
 Next, each word of the sentence was presented in the center of the screen.
@@
 This functioned as a regression-based instead of traditional baseline correction \citep{alday2019}.
-Varying intercepts for participants and items as well as by-participant random slopes for semantic interference were included.
-Correlations between varying intercepts and slopes were not included.
+Varying intercepts for participants and items as well as by-participant \textcolor{blue}{and by-item} random slopes for \textcolor{blue}{syntactic interference,} semantic interference \textcolor{blue}{and their interaction} were included.
 We used relatively informative priors for all parameters of the models (see Table \ref{tab:eeg_priors}).
@@
   SD & Normal(0, 2)&\\
+ \textcolor{blue}{correlations} & \textcolor{blue}{LKJcorr(2)} & & \\
     \bottomrule
@@
 The 95\% credible intervals for the fixed effects are consistent with effects of syntactic interference, semantic interference and their interaction.
-The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (mean: -0.31\,$\mu V$, 95\% CrI = [-0.49, -0.13]\,$\mu V$, for the posteriors, see Figure \ref{fig:eeg_posteriors}).
-The estimate of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).
+The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (\textcolor{blue}{mean: -0.32\,$\mu V$, 95\% CrI = [-0.59, -0.10]\,$\mu V$}, for the posteriors, see Figure \ref{fig:eeg_posteriors}).
+The estimates of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).
 
@@
 baseline  &  -0.22 & 0.01 &  [-0.24, -0.20] \\
-syntactic &  -0.15 & 0.08 &  [-0.32, -0.02] \\
-semantic  &  -0.31 & 0.09 &  [-0.49, -0.13] \\
-interaction& -0.13 & 0.10 &  [-0.37, -0.01] \\
+syntactic &  \textcolor{blue}{-0.16}& \textcolor{blue}{0.08} &  \textcolor{blue}{[-0.33, -0.02]} \\
+semantic  &  \textcolor{blue}{-0.30} & \textcolor{blue}{0.10} &  \textcolor{blue}{[-0.50, -0.11]} \\
+interaction& \textcolor{blue}{-0.14} & \textcolor{blue}{0.11} &  \textcolor{blue}{[-0.40, -0.01]} \\
     \bottomrule
@@
     \centering
-    \includegraphics[width=0.6\textwidth]{images/posteriors_eeg.jpg}
+    \includegraphics[width=0.7\textwidth]{images/posteriors_eeg.jpg}
 \end{figure}
@@
 
-Bayes factors are shown in Figure \ref{fig:eeg_bfs}. The Bayes factors provided very strong to strong evidence for the semantic interference effect for all priors 
-(Normal$_{-}$(0, 0.1): BF$_{10}$ = 27,
-Normal$_{-}$(0, 0.5): BF$_{10}$ = 88,
-Normal$_{-}$(0, 1): BF$_{10}$ = 52,
-Normal$_{-}$(0, 5): BF$_{10}$ = 11).
-
-By contrast, the syntactic interference effect had only anecdotal evidence  under the two narrow priors (Normal$_{-}$(0, 0.1): BF$_{10}$ = 2.5,
-Normal$_{-}$(0, 0.5): BF$_{10}$ = 1.2). There was evidence \textit{against} syntactic interference under wider priors 
-(Normal$_{-}$(0, 1): BF$_{10}$ = 0.66,
-Normal$_{-}$(0, 5): BF$_{10}$ = 0.14).
-What this means is that, in our data, only if we assume a priori that the effect size is relatively small, there is very weak evidence in favor of syntactic interference.
+Bayes factors are shown in Figure \ref{fig:eeg_bfs}. The Bayes factors provided very strong to \textcolor{blue}{moderate} evidence for the semantic interference effect for all priors 
+\textcolor{blue}{(Normal$_{-}$(0, 0.1): BF$_{10}$ = 11,
+Normal$_{-}$(0, 0.5): BF$_{10}$ = 80,
+Normal$_{-}$(0, 1): BF$_{10}$ = 9,
+Normal$_{-}$(0, 5): BF$_{10}$ = 5)}.
+
+By contrast, the syntactic interference effect had only anecdotal evidence  under the \textcolor{blue}{narrowest prior (Normal$_{-}$(0, 0.1): BF$_{10}$ = \textcolor{blue}{2.18}) and under the Normal$_{-}$(0, 0.5) prior there was no evidence for or against it (BF$_{10}$ = 1.05).} There was evidence \textit{against} syntactic interference under wider priors
+(Normal$_{-}$(0, 1): BF$_{10}$ = \textcolor{blue}{0.53},
+Normal$_{-}$(0, 5): BF$_{10}$ = \textcolor{blue}{0.1}).
+What this means is that, in our data, only if we assume a priori that the effect size is \textcolor{blue}{very small, i.e., maximally $-$0.2\,$\mu V$,} there is very weak evidence in favor of syntactic interference.
 
@@
 
-The Bayes factors analyses found strong evidence for semantic interference in the N400 amplitude.
-Syntactic interference had much weaker evidence in the N400 amplitude; only anecdotal evidence was found under narrow priors, and there was clear evidence against syntactic interference under wider priors.
+The Bayes factors analyses found \textcolor{blue}{very strong to moderate evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($-$0.2 to $-11$\,$\mu V$)}.
+Syntactic interference had much weaker evidence in the N400 amplitude; only anecdotal evidence was found under \textcolor{blue}{the narrowest prior, which assumed effect sizes smaller than $-$0.2\,$\mu V$.
+No evidence for or against syntactic interference was found under a prior which assumed a reasonable effect size of up to $-1$\,$\mu V$ ($BF_{10}$=1.05).
+There was clear evidence against syntactic interference under wider priors (effect sizes of up to $-$2 or up to $-11$\,$\mu V$)}.
 The evidence against an interaction under all priors was unequivocal.
+\textcolor{blue}{In sum, the ERP results showed clear evidence for semantic interference and clear evidence against an interaction of syntactic and semantic interference.
+The evidence for or against syntactic interference was dependent on the a priori assumed effect size.
+However, even when very small effects were assumed a priori, the evidence for syntactic interference was weak at best.
+Consequently, we conclude that the ERP data showed no syntactic interference.} \textcolor{red}{Add conclusion after SPR results are revised.}
 
@@
 
-Consistent with the self-paced reading study, there is clear evidence for semantic interference, and evidence against syntactic interference.
+Consistent with the self-paced reading study, there is clear evidence for semantic interference, and \textcolor{blue}{little to no evidence for} syntactic interference.
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
@@
 
-\vspace{2em}
-\section{Data availability statement}
-
-The data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.
-
 \newpage

### 2024-06-22 paper/ResponseToReviewers.tex
Added via the Open Science Framework

\documentclass[12pt]{article}
\usepackage[a4paper,top=2cm,bottom=2cm,left=2.5cm,right=2.5cm,marginparwidth=1.75cm]{geometry}
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{color}

%\usepackage[backend=biber,useprefix=true,style=apa,url=false,sorting=nyt,eprint=false]{biblatex}

%\addbibresource{interferencebib.bib}

\usepackage[natbib=true,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
\usepackage{csquotes}
\DeclareLanguageMapping{american}{american-apa}
\bibliography{interferencebib}

\usepackage{marginnote}
\setcounter{section}{-1}

\usepackage{lineno,xcolor,clipboard,graphicx}

\openclipboard{output-reviews}

\renewcommand{\qedsymbol}{\rule{0.7em}{0.7em}}

\usepackage{epstopdf}
\newcommand{\revised}[1]{{\color{black}{#1}}}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Response to editor and reviewers (JML-24-17)}
\author{Pia Schoknecht and Shravan Vasishth}
%\date{}                                           % Activate to display a given date or no date

\usepackage{xr}

\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
\typeout{(#1)}% latexmk will find this if $recorder=0
% however, in that case, it will ignore #1 if it is a .aux or 
% .pdf file etc and it exists! If it doesn't exist, it will appear 
% in the list of dependents regardless)
%
% Write the following if you want it to appear in \listfiles 
% --- although not really necessary and latexmk doesn't use this
%
\@addtofilelist{#1}
%
% latexmk will find this message if #1 doesn't exist (yet)
\IfFileExists{#1}{}{\typeout{No file #1.}}
}\makeatother

\newcommand*{\myexternaldocument}[1]{%
\externaldocument{#1}%
\addFileDependency{#1.tex}%
\addFileDependency{#1.aux}%
}

\myexternaldocument{main}

\begin{document}
\maketitle
%\section{}
%\subsection{}

We want to thank the reviewers and editor for their detailed and constructive comments; these have significantly improved the paper. We believe we have addressed every point raised by the editor and reviewers; the comments with our responses and a record of the changes in the paper are listed below. When relevant, we provide references to pages from the revised paper.

All the changes in the paper are marked in blue.

%\Paste{vd_design} from page \pageref{vd_design}.

\section*{Editor's comments}

\begin{quote}
``Dear Dr. Schoknecht,

Thank you for submitting your manuscript to our special issue on the Language-Memory Interface at the Journal of Memory and Language. Your paper has now been reviewed by three experts in the field and your will find their comments below. I also read the manuscript carefully myself before turning to the reviews.

You will see that all three of the reviews are thorough, well-informed, and detailed and I am grateful to the reviewers for the care and time they have clearly put in to reading the manuscript. My own reading aligned with Reviewer 1's assessment in that I found your paper to be very clearly written and enjoyable to read. Your very good sample sizes and sophisticated analyses and modelling suggest to me that this work has the potential to make a strong contribution to our special issue.

However, I am far from being an expert in this area, and the (therefore much more knowledgeable) reviewers have raised a number of issues with the paper, some of which I judge to be substantive. For this reason I have come to a 'revise and resubmit' decision. If you do revise and resubmit your paper, as I hope you will, I will approach the same set of reviewers to see if they are able to re-review (and obviously there is no guarantee of acceptance at that stage).

In revising your paper you should take into account all of the reviewer's points. However, there are two themes that I picked out as particularly important.

First, \textbf{Reviewer 2 raises a potentially serious concern that your particular manipulation of syntactic dis/similarity might not be expected to lead to substantial syntactic interference (their point 1)}\marginpar{E.1}. In other words, the questions are whether one would ever expect to see much of this effect with your design and whether a more appropriate design would lead to stronger findings. Some of Reviewer 3's points (see their paragraph starting 'The authors attribute ...' and their final summary comment) are perhaps in the same vein.

Second, a couple of Reviewer 1's comments (e.g., their point 2, and their final point about priors) relate to the fact that \textbf{there actually is some evidence for a degree of syntactic interference in aspects of the data, particularly perhaps in terms of the interaction and the combination with high semantic similarity, that you risk underplaying}\marginpar{E.2} (see also Reviewer 2's point 5).

Finally, I would also add that while I agree with Reviewer 1's assessment that your paper would fit well in the special issue, \textbf{as someone more versed in the memory than the language literature I thought you could be more explicit about how your findings do integrate both of these two areas.}\marginpar{E.3}

The reviewers have made suggestions which the Editors feel will improve your manuscript. We encourage you to consider these comments and make an appropriate revision of your manuscript.

We are trying to improve the reproducibility of work published in the Journal. \textbf{Please ensure that you have included a "Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code} \marginpar{E.4}(see our 'guide for authors' for instructions). Please make sure that you have updated the data and analysis code following any revisions to your manuscript. Please confirm in your next covering letter that all results presented in your manuscript can be reproduced using the materials that you have made available.''
\end{quote}

\subsection*{Response to E.1}
We address this concern in our response to Reviewer \#2 (2.\textcolor{red}{X}).
\subsection*{Response to E.2}
We address this concern in our response to Reviewer \#1 (1.\textcolor{red}{X}).
\subsection*{Response to E.3}
\textcolor{red}{write more about memory during sentence processing}
\subsection*{Response to E.4}
We have moved our data availability statement to page \pageref{statement}, before Experiments 1a and 1b.

\section*{Reviewer \#1's comments} 
\subsection*{Comment R1.1}
\begin{quote}
``In the manuscript "Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German" the authors investigate the effect of syntactic and semantic interference in a large-sample self-paced reading and EEG studies that follow in design Van Dyke (2007) and Mertzen et al. (2023).

This is a very interesting paper with solid findings, well-written and relevant for the JML special issue on language and memory. I really enjoyed reading it - I very much liked the rigorous combination of experimental work with a Bayesian analysis and computational cognitive modeling. The data, while largely replicating previous studies, are important for a discussion on memory and language and would deserve to be available to scientific community. All in all, I can see a path forward to publication for the manuscript. At the same time, though, I do have several issues that in my opinion should be resolved - mainly regarding the interpretation of the results and several modeling choices. I discuss these issues, listed by topic, below.

Interpretation of the results\\
================

One of the main findings is that the studies show a convincing, strong effect of semantic interference, but the effect of syntactic interference (of syntax-semantics interaction) is not present. This is in contrast to several previous studies, which did show syntactic interference, in addition to semantic interference.

In Discussion, the authors propose the interpretation of the results in a cue-based retrieval model, in which two cues are relevant for retrieval: [+subject in-same-clause], [+animate].

I have several problems with this interpretation:

1. Encoding vs.\ retrieval: As far as I can see, \textbf{the interpretation cast in the Lewis and Vasishth (2005) model means that the interference is present at retrieval, not at encoding. But that goes against the proposed informal interpretation of the findings of the self-paced reading study. There it is argued that the observed effect of semantic interference is due to *encoding*, not (just) retrieval. That is, there is a discrepancy between the informal interpretation of the self-paced reading study and the computational model.} ''\end{quote}

\subsubsection*{Response to R1.1}
The reviewer is right. The Lewis and Vasishth (2005) model is a model of memory retrieval during sentence processing and should therefore not be used to model effects of encoding interference. We have removed the modeling of the SPR effects. In the revised manuscript, we now only present modeling of retrieval interference in ERPs.

\textcolor{red}{Remove SPR modeling from manuscript}

\subsection*{Comment R1.2}
\begin{quote}
``2. Match of the model and data: The interpretation is set in such a way that only semantic interference, but no syntactic interference or interaction are predicted. This is achieved by hard-coding that the syntactic cue selects the target and nothing else. I find this problematic. For the modeling results of the current experiment, in Figure 11 it seems that the syntactic interference and interaction are underpredicted by the model - not by much, but they are. While the model predicts no effect, the data do show an effect, be it small (btw, it is hard to read this figure because of the chosen scale; \textbf{it would be much better to zoom in and show for example 0-30 ms scale for the reading study 0 to -1 microvolt scale for the EEG study)}.''\end{quote}

\subsubsection*{Response to R1.2}
We changed the scale of the x-axis in Figure 11 to improve readability.

\subsection*{Comment R1.3}
\begin{quote}
``In my reading, the discrepancy mainly comes from the fact that the decision to neglect syntactic interference is based on Bayes factors (see also below), but when matching the model predictions to actual results (p. 39-42), the match is done on posterior distributions, which do show a positive effect of interaction (the self-paced reading study) and an effect of syntactic interference (the self-paced reading study and the EEG study). \textbf{I think what the authors could do to at least partially address this issue is to be more explicit about model comparisons. They could leverage their use of Bayes factors to compare different computational models - the one proposed in this paper, which hard-codes that only the syntactic target is selected, the one based on Mertzen et al., possibly with varying priors, and, possibly, a null model which does not predict any interference (e.g., it hard-codes that only the target is retrieved).}''\end{quote}

\subsubsection*{Response to R1.3}
\textcolor{red}{Do k-fold crossvalidation of computational models of ERP data}

\subsection*{Comment R1.4}
\begin{quote}
``3. Comparison to previous findings: Syntactic interference has been observed previously. The authors discuss in detail Van Dyke (2007) and Mertzen et al. (2023) as two studies that argue for syntactic interference. This is in contrast to the current study, which is claimed not to show syntactic interference. One problem for me is that we are not making a fair comparison: the current study bases its interpretation on Bayes factors, which were not used in Van Dyke or Mertzen et al. To be sure, my point is not that Bayes factors should not be used (of course, I find it good that they are used to compare hypotheses), just that the comparison between this study and the other findings is far from straightforward. If we only check posterior distributions, which were explored in Mertzen et al. for their conclusions, we do see some effect of syntactic interference - e.g., the predominantly negative credible interval in the EEG study. As far as I can see, \textbf{to really make a sensible comparison, one would either have to reanalyze Mertzen and Van Dyke's data using Bayes factors or be more precise in explaining how the posterior distributions in this study differ from previous studies} - just looking at the posteriors, they do not seem to paint such different pictures between the current study and Mertzen et al. ''\end{quote}

\subsubsection*{Response to R1.4}
In the revised manuscript, we are now more explicit about the comparison to previous findings. We added a new figure (Figure \textcolor{red}{X}) to the General Discussion showing the estimates of the previous studies along with our own estimates. Additionally, we computed Bayes Factors for the two previous studies. Because the Van Dyke (2007) data is not available to us, we computed Bayes Factors using the Savage-Dickey method based on the estimates derived from their published paper. To maximize comparability between studies, we have recomputed the Bayes Factors of our study with the Savage-Dickey method (previously we had used bridgesampling).

\subsection*{Comment R1.5}
\begin{quote}
``Another problem I have is more general and has to do with progress in science. The authors in my reading find the current findings more convincing than, say, Mertzen et al. study. I get that the current study is convincing (I do find it convincing) but that does not disqualify what Mertzen et al. found in any way - both results co-exist. Ideally, I would expect that a model would be able to explain the current data, but also the past data. I get it that this is not always possible but then, \textbf{it would be good if the authors make their position explicit - either by admitting that only some data can be captured, or, if the authors believe that the past data should be ignored, it should be explained why this is so.}

All in all, I think the interpretation of the data and the model have to be strengthened to meet requirements 1-3. Alternatively, the manuscript should explicitly admit issues in the proposed interpretation - that is, explain that the model fails to capture encoding interference (point 1 if I am right), explain why the model underpredicts interference in syntax/interaction and explain why it is enough, at least for now, to only model the current findings and to put previous findings aside (assuming they actually differ - see point 3 about Bayes factors above).''\end{quote}

\subsubsection*{Response to R1.5}
\textcolor{red}{Write something about the co-existence of contradictory results}

\subsection*{Comment R1.6}
\begin{quote}
``Semantic encoding interference vs.\ acceptability\\
===========================

It is proposed that the semantic interference in the self-paced reading study is due to encoding. This interpretation is supported by the fact that we observe increased reading times from the moment that the +animate distractor is read (nicely shown in Figure 3). I wonder whether an even more straightforward interpretation is possible here. Could the effect of increased reading times simply be due to the fact that the +animate distractor is not as good as the -animate distractor in this context? That is, the effect has nothing to do with encoding interference, it's just a consequence of the fact that, e.g., in (2a) the statement that "the burglar was awful" is harder to make sense of than that "the loss was awful". \textbf{I think one easy way to check it would be to run an acceptability/plausibility study which checks whether the +animate version and the -animate version are comparable in their acceptability.} I know that the manuscript mentions a possible confound due to plausibility (p. 22, 23) but as far as I could understand it is assumed that this confound would only affect the syntactic manipulation, not the semantic one. However, I think the semantic manipulation could just as well result in more or less plausible sentences.''\end{quote}

\subsubsection*{Response to R1.6}
We ran the suggested plausibility judgement experiment. The -animate distractor is indeed slightly more plausible in the context than the +animate distractor (\textcolor{red}{add numbers}). We discuss this confound in the revised manuscript and ran additional analyses which show that the reading times slowdown for high vs.\ low semantic interference can not be solely explained by this plausibility difference.

\subsection*{Comment R1.7}
\begin{quote}
``Intercept-only models\\
=============

Models with intercept-only random-effect structures are used for the data analysis. I see why this has to be done for Bayes factors. I wonder, though, whether this justifies the use of intercept-only models to find posterior distributions of parameters. I would worry that given that all Bayes models have intercept-only random-effect structures, we probably overestimate the size of fixed effects (e.g., in Figure 4). \textbf{Why not doing the following: one could use models with full random effect structure for estimation of posterior distribution of parameters, and then switch to intercept-only models to calculate Bayes factors,} since more complex models lead to unstable Bayes factors.''\end{quote}

\subsubsection*{Response to R1.7}
As mentioned in our response to R1.4, we recomputed our Bayes Factors using the Savage-Dickey method. With this method we were able to compute Bayes Factors for models with complex random effect structures. All results in the manuscript have been updated accordingly.

\subsection*{Comment R1.8}
\begin{quote}
``Priors for Bayes factors\\
=============

Various priors are considered for Bayes factors. Of course, it is good to explore several values but I wonder whether it is justified to conclude that syntactic interference or interaction should be excluded if they are not supported in models with unreasonably large priors (e.g., N+(0, 0.1)). Would anybody expect that interference is in range of 0.1 log-ms? As one of the authors of this manuscript nicely showed in Jager et al. (2017) and Schad et al. (2022), interference effects are small (at least for subject-verb agreement and ungrammatical illusions). \textbf{Given that, I don't understand why we require a BF in favor of interference for a prior that is of size N+(0, 0.1). Shouldn't a smaller prior distribution suffice as evidence? This brings me again to my confusion in what the experiments actually show. The authors seem to argue that the only interference is the semantic one and they base the conclusion on large priors. But focusing on small priors (N+(0, 0.05) for spr, N-(0, 0.5) for EEG)), one could conclude that the interaction/syntactic interference might be present after all. Is there an independent reason to select the former priors over the latter for one's interpretation? To be sure, I don't want the authors to change the interpretation, but I would like to see either a more explicit argumentation why we should base our conclusion on such and such priors, or a more reflection in the text that explicitly states that the results are not equivocal and the current study is still compatible with a position that syntactic interference and/or interaction are real.}''\end{quote}

\subsubsection*{Response to R1.8}
\textcolor{red}{Write something about sensitivity analysis}

\section*{Reviewer \#2} 
\subsection*{Comment R2.1}
\begin{quote}
``This paper reports two high-powered studies (SPR and ERP) in German aiming to investigate the role of both syntactic and semantic interference in sentence comprehension. The authors investigate syntactic interference by manipulating the syntactic status of the distractor (subject vs. non-subject), and semantic interference by manipulating the semantic status of the distractor (animate vs. inanimate). Results show evidence for semantic interference and evidence against syntactic interference. With the aim to reconcile their results with the vast literature attesting to a clear role of syntactic interference, the authors suggest that the parser is driven by structural information allowing it to discard syntactically implausible elements as potential retrieval targets, thereby mitigating their potential for intervention.

While I am favorably impressed by the large-sample size tested by the authors and the meticulous statistical analyses, I also have concerns regarding the paper's overall lack of theoretical depth, the operationalization of syntactic interference, as well as several decisions made by the authors in generating materials.

I honestly don't know if these problems can be resolved in a major revision, or whether they will require so much new experimentation and/or theorizing that the manuscript which results would be essentially a new manuscript. I'll leave that to the editor.

I discuss my concerns in turn.

The authors identify 2 major problems with the previous literature I agree with: 1) studies lacking sufficient statistical power, and 2) ambiguity regarding which cues are pertinent for retrieval.

They address the 1st concern by increasing the number of participants and items tested, which is commendable. Regarding the second issue, they address it by testing both syntactic and semantic cues. As a syntactic cue, the authors tested "being a subject". In the high syntactic interference condition, the distractor functions as the subject of an embedded clause that serves as the complement of a non-restrictive relative clause (who had told her that the loss/ burglar was awful). In the low syntactic interference condition, the distractor is within a prepositional phrase serving as the complement of a non-restrictive relative clause (who had told her about the awful loss/ burglar).

1. From a theoretical point of view, "being a subject" is not a primitive. Subjects are identified by a cluster of features. One of these features is c-command (the subject c-commands the verb). There is a whole literature (in agreement attraction, for instance) showing that distractors that c-command the verb generate strong syntactic interference even when they do not linearly intervene between the subject and the verb (and even when they are not subject themselves). So, we know, as a matter of fact, that syntactic interference is real.
The authors are concerned about underpowered studies, a concern I totally agree with. But it's a real shame they used the full force of Bayesian analysis, EEG, and a very good-sized subject pool to test conditions that do not instantiate the basic ingredients I would expect to generate syntactic interference, leading the authors to reinvent the wheel: the parser is not considering an element that could never serve as the potential subject of the matrix clause during subject retrieval.

Unsurprisingly, the authors reach this very same conclusion: "We speculate that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory." Yes, exactly, but this isn't new.

I also agree with this statement (p.41): "[…] the parser intelligently targets only the syntactically relevant noun during antecedent retrieval. In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb." Once again, yes, that's accurate. However, \textbf{this implies that the syntactic manipulation the authors went for is just not doing the job the authors would like it to do: the syntactic distractor is not a syntactically relevant noun. Being a subject of whichever clause at whichever level of embedding is not a sensible manipulation to test syntactic interference.}''
\end{quote}

\subsubsection*{Response to R2.1}
\textcolor{red}{??}

\subsection*{Comment R2.2}
\begin{quote}
``2. The syntactic manipulation does not lead to minimally different conditions (i.e., conditions that only differ on the critical manipulation). These are the glosses for the two syntactic conditions:\\

who told had that the burglar/loss awful was\\
who about the awful burglar/loss told had\\

These are completely different structures with even different word orders. As a result, any difference in RTs observed at the critical region might ensue from a number of different reasons that might have nothing to do with burglar/loss being or not the subject of the embedded clause.

3. The semantic manipulation is not minimal either. In this case, not only is the number of letters different (as acknowledged by the authors themselves), but even the number of syllables isn't the same. Once more, this is not a minimal comparison.

4. The difference in RTs starts at the distractor and continues through the critical verb region (in the high syntactic interference condition at least). Therefore, we cannot exclude the possibility that the effect observed at the critical verb is the same effect that originates at the distractor that is lingering through the sentence, rather than being a different effect altogether. Consequently, \textbf{I don't think that the effect at the critical region can be interpreted.}''
\end{quote}

\subsubsection*{Response to R2.2}
\textcolor{red}{??}

\subsection*{Comment R2.3}
\begin{quote}
``\textbf{(Furthermore, what's happening in regions 8-9? There is an increase in RTs in these regions in the high syntactic interference condition and a decrease in the low syntactic interference condition, while the trend for the remaining regions is quite similar. It's almost as if participants were able to discern the two syntactic conditions as early as region 8).}''
\end{quote}

\subsubsection*{Response to R2.3}
\textcolor{red}{??}

\subsection*{Comment R2.4}
\begin{quote}
``5. \textbf{Another puzzling aspect is why the effect of the semantic manipulation only manifests in the high syntactic interference condition. The authors fail to address this interaction altogether.} Instead, they repeatedly emphasize finding evidence for semantic interference and against syntactic interference, without discussing the interaction. Once more, I believe these results are difficult to interpret, but it is surprising to see such a limited discussion of the findings.''
\end{quote}

\subsubsection*{Response to R2.4}
\textcolor{red}{??}

\subsection*{Comment R2.5}
\begin{quote}
``6. \textbf{The N400 is traditionally considered to be the hallmark of semantic effects. However, what is the linking hypothesis between syntactic interference and N400? The paper fails to address this crucial point.}''
\end{quote}

\subsubsection*{Response to R2.5}
The linking hypothesis is that memory retrieval might modulate the N400. Any type of retrieval interference, may it be due to syntactic or semantic retrieval cues, might affect memory retrieval and therefore the N400. We added the following text to the paragraph about the regions of interest on page \pageref{why_n400}:

\Paste{why_n400}

\begin{quote}
``7. \textbf{It is noteworthy that the general discussion does not mention or discuss the works showing that hierarchical intervention matters in agreement attraction and that pre-date all the studies mentioned in the several parentheticals. I have in mind Franck et al. 2002, 2006 and 2011, for instance. These works offer a treatment of attraction in terms of intervention and, as far as my knowledge goes, these are the first studies highlighting the crucial role of hierarchical interference effects. While authors mention other papers about agreement attraction, they seem to overlook these particular ones.}''
\end{quote}

\section*{Reviewer \#3} 
\subsection*{Comment R3.1}
\begin{quote}
``This article describes two experiments that test whether comprehension of subject-verb relations in complex sentences (in German) is disrupted by the presence of irrelevant-but-tempting nearby noun phrases. The temptation is due either to semantic properties ("good fit to the verb") or syntactic properties ("looks like a subject"). The key finding from both experiments is that semantic temptation effects are clear, while syntactic temptation effects are largely absent. The authors interpret this contrast as evidence for specific types of features used in sentence encoding and retrieval operations that should prevent effects of syntactic interference, i.e., if encountering a verb induces retrieval of "subject of the same clause", then the subject of a different clause shouldn't interfere.

The design of the current studies repeats a design used in previous studies in German and other languages. The more novel aspects of the current studies are the use of the ERP measure, and the inclusion of a very large number of participants, allowing for more precise estimates of the syntactic and semantic interference effects. The quantitative analyses are characteristically thorough and careful.

I would support publication in JML. But I think that the paper would be stronger if there was fuller engagement with the range of possible reasons for the observed effects. This covers (i) mechanisms that could capture the lack of a syntactic interference effect, and (ii) reasons for differences between the current study and others where a syntactic interference effect was found.

The authors attribute the lack of a syntactic interference effect to the use of a syntactic feature like "subject of the same clause", which has the effect that a subject of a different clause does not count as similar to the retrieval cue. This gives a special status to the correct subject via the choice of retrieval cues. \textbf{There may be other ways of giving special status to the correct subject, and thereby capturing the lack of syntactic interference. For example, if the subject of the current clause is encoded in a special status, e.g., with boosted activation, or with its needs prioritized, then this could have a similar effect.}''
\end{quote}

\subsubsection*{Response to R3.1}
\textcolor{red}{??}

\subsection*{Comment R3.2}
\begin{quote}
``The use of a feature like "subject of the same clause" calls for some further discussion, since, taken literally, this is not a likely feature of a subject phrase in memory. Features like "animate" or "plural" are inherent properties of an item in memory, but "subject of the same clause" is not. \textbf{I imagine that what the authors have in mind is that comprehenders assign a clause index to each individual piece of structure as they parse a sentence. Then, when they encounter a verb, they could identify the current clause index, e.g., "3", and then embed that in a retrieval cue, e.g., "subject of clause 3". This is a fine way to introduce a limited amount of relational information into a structural encoding. But note that it means that a good deal of work is now done by the mechanism that continually tracks clause indices and keeps track of the index of the current clause. So, it amounts to a mechanism that does more than simply elaborate the choice of retrieval cues at the verb.}''
\end{quote}

\subsubsection*{Response to R3.2}
\textcolor{red}{??}

\subsection*{Comment R3.3}
\begin{quote}
``\textbf{I also have some questions or reservations about the account of semantic interference effects in terms of features like [+animate]. This is a straightforward enough feature to invoke, and it is a plausible content feature for noun phrases. But I am less sure how well it generalizes. There are many situations in which verb-noun combinations are intuitively a good fit, but where it is less easily cast in terms of features and retrieval cues. For example, we share the intuition that "the teacher drove …" is a better combination than "the little boy drove …", because teachers are plausible drivers, whereas small children are not. This is less easily captured in terms of features and retrieval cues. Maybe the verb "drive" wants a subject with a [+driver] feature. But is this plausibly something that is part of the memory encoding of "teacher" but not "little boy"? Does this require something more than a toy theory of semantic encoding? Or does it predict that only simple features like [+animate] are able to cause semantic interference?}''
\end{quote}

\subsubsection*{Response to R3.3}
It has been shown in the literature that fine-grained semantic features, such as [+fixable] \citep{vandyke_mcelree06} and [+shatterable] \citep{cunnings_sturt_2018retrieval}, are used during cue-based retrieval. The materials in the design employed here were constructed so that the potentially fine-grained features relevant for the situation at hand (e.g., [+can drive], [+can drink] or [+can smile]) could be summed under the umbrella feature [+animate] or actually [+human]. To resolve the critical dependency it was always sufficient to use the umbrella feature [+animate] (or [+human]). The subject and the animate distractor always shared the [+animate] (or more precisely [+human]) feature, while the inanimate distractor was always [-animate] and hence [-human]. There were no animate distractors which mismatched a fine-grained verb-specific retrieval cue like [+can drive]. The results of the plausibility norming, see Figure \ref{fig:plausibility_rating} on page \pageref{fig:plausibility_rating} of the manuscript illustrate that whatever semantic feature determines a good fit between subject and verb matched with the features of the subject and the animate distractor, but not with the features of the inanimate distractor.

As the reviewer suggested it seems unlikely that all potential fine-grained features are part of an initial memory encoding of each word during sentence processing. But there needs to be a way to retrospectively enrich that memory encoding as the sentence unfolds. There might be a hierarchy of semantic features. A feature like [+can smile] might be a daughter of the feature [+animate/+human]. The degree to which that tree of semantic features becomes activated during sentence processing might differ between individuals. The present study can not provide insight into the specifics of memory encoding.

We have added the following text to the description of the materials on page \pageref{semantic_features_encoding} of the manuscript:

\Paste{semantic_features_encoding}

\subsection*{Comment R3.4}
\begin{quote}
``\textbf{The authors offer only limited discussion of why their results differed from those of other studies. To some degree, the paper seems to take the position that since the current studies had huge numbers of participants and hence high statistical power, this study reveals the true state of the world, and other studies should be taken less seriously because their results were less reliable. This may be an over-simplistic approach to the findings.}''
\end{quote}

\subsubsection*{Response to R3.4}
We think that we addressed these concerns in our response to Reviewer \#1 in R1.5.

\subsection*{Comment R3.5}
\begin{quote}
``First, \textbf{could there have been an impact of the fact that the study included an unusually large number of experimental items with identical structures, leading to practice effects?}''
\end{quote}

\subsubsection*{Response to R3.5}
To account for practice effects, we added the centered trial id as a fixed effect as well as by-participant and by-item random slopes to the models analyzing the SPR data. \textcolor{red}{The results did not change considerably.}

\subsection*{Comment R3.6}
\begin{quote}
``\textbf{(One thing that was less clear from the manuscript was the nature and number of any filler items. This receives little discussion in the paper.)}''
\end{quote}

\subsubsection*{Response to R3.6}
We added the following regarding the filler items to the Procedure section of Experiment 1, page \pageref{fillers1}:

\Paste{fillers1}

\noindent We also added a brief statement on fillers to the Procedure section of Experiment 2 (page \pageref{fillers2}) and corrected a typo regarding the number of trials per experimental block:

\Paste{fillers2}

\subsection*{Comment R3.7}
\begin{quote}
``\textbf{Second, could there have been a contribution from the fact that the correct subject NP appeared in a very prominent position, delimited by commas and appearing at a line break?}''
\end{quote}

\subsubsection*{Response to R3.7}
\textcolor{red}{prominence felix, reduced interference effects, linebreaks VD07?}

\subsection*{Comment R3.8}
\begin{quote}
``\textbf{Third, in looking through the full set of experimental items, it appears that the inappropriate subject NP is always the subject of an Aux + Adj predicate, rather than, say, the agent of an action verb. Could this different status of that subject have contributed to the lack of a syntactic interference effect?}''
\end{quote}

\subsubsection*{Response to R3.8}
\textcolor{red}{see items VD07}

\begin{quote}
``Summarizing, I am overall enthusiastic about this study. But I am less sure of how we should generalize from this specific finding using one specific manipulation. There are multiple possible ways of capturing the lack of a syntactic interference effect, and there could be multiple reasons why the current study might have found different results than other similar studies in the literature.''
\end{quote}

\end{document}

### 2024-07-03 paper/main.tex
Updated via the Open Science Framework

@@
 
-\textcolor{red}{\section{Data availability statement}
+\textcolor{blue}{\section{Data availability statement}
 All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}\label{statement}
@@
 
+\Copy{figwhole_sentence}{
 \begin{sidewaysfigure}[h]
@@
     \includegraphics[width=0.85\textwidth]{images/Pandora_all_wholesentence_pooled_zoom_exp.jpg}
-\end{sidewaysfigure}
+\end{sidewaysfigure}}
 \clearpage
@@
 The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even the same linguistic material.
+\Copy{items_mertzen}{\textcolor{blue}{However, \citet{mertzen} changed the structure of the relative clause modifying the subject by adding an additional animate distractor to all conditions (see an example of one of their English items in \ref{ex:mertzen}; their items had this structure cross-linguistically).
+This was done to ``increase the strength of the manipulation.'' \citep[][p.
+9]{mertzen}.
+The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
+But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
+While this additional distractor might have increased syntactic interference compared to our study (and \citeauthor{vandyke07}'s study as well actually), it is not obvious why it would decrease semantic interference compared to our study.}\label{items_mertzen}
+
+\begin{exe}[ht]
+\ex \label{ex:mertzen} 
+\textcolor{blue}{It turned out that the attorney whose secretary had forgotten that the visitor was important frequently complained about the salary at the firm. \parencite{mertzen}}
+\end{exe}
+}
+
 So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
-Apart from the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
+Apart from \textcolor{blue}{the additional distractor and} the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
 While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
-We will revisit the absence of syntactic interference in our data in the general discussion, where we investigate the question using computational modeling.
+We will revisit the absence of syntactic interference in our data in the general discussion.
 
@@
 
+\Copy{figpriorposteriordata}{
 \begin{figure}[!htbp]
@@
 \end{figure}
+}
 

### 2024-07-03 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \usepackage{graphicx}
+\usepackage{subfigure}
+\usepackage{rotating}
 \usepackage{amssymb}
@@
 \usepackage{color}
+\usepackage{gb4e}
+\noautomath
 
@@
 All the changes in the paper are marked in blue.
-
-%\Paste{vd_design} from page \pageref{vd_design}.
 
@@
 \subsubsection*{Response to R1.2}
-We changed the scale of the x-axis in Figure 11 to improve readability.
+We have changed the scale of the x-axis in Figure 11 to improve readability. For convenience, we paste the Figure here:
+
+\Paste{figpriorposteriordata}
 
@@
 Additionally, we computed Bayes Factors for the two previous studies.
-Because the Van Dyke (2007) data is not available to us, we computed Bayes Factors using the Savage-Dickey method based on the estimates derived from their published paper.
+
+Because the Van Dyke (2007) data is not available to us, we have used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) reported in the Van Dyke (2007) paper.
+Then using these posteriors and a reasonable prior (a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors using the Savage-Dickey method.
 To maximize comparability between studies, we have recomputed the Bayes Factors of our study with the Savage-Dickey method (previously we had used bridgesampling).
@@
 The -animate distractor is indeed slightly more plausible in the context than the +animate distractor (\textcolor{red}{add numbers}).
-We discuss this confound in the revised manuscript and ran additional analyses which show that the reading times slowdown for high vs.\ low semantic interference can not be solely explained by this plausibility difference.
+We discuss this potential confound in the revised manuscript.
+However, we carried out additional analyses which included plausibility as a co-predictor and find that the evidence is still strong for semantic interference.
 
@@
 Various priors are considered for Bayes factors.
-Of course, it is good to explore several values but I wonder whether it is justified to conclude that syntactic interference or interaction should be excluded if they are not supported in models with unreasonably large priors (e.g., N+(0, 0.1)).
+\textbf{Of course, it is good to explore several values but I wonder whether it is justified to conclude that syntactic interference or interaction should be excluded if they are not supported in models with unreasonably large priors (e.g., N+(0, 0.1)).
 Would anybody expect that interference is in range of 0.1 log-ms?
@@
 (2022), interference effects are small (at least for subject-verb agreement and ungrammatical illusions).
-\textbf{Given that, I don't understand why we require a BF in favor of interference for a prior that is of size N+(0, 0.1).
+Given that, I don't understand why we require a BF in favor of interference for a prior that is of size N+(0, 0.1).
 Shouldn't a smaller prior distribution suffice as evidence?
@@
 \subsubsection*{Response to R1.8}
-\textcolor{red}{Write something about sensitivity analysis}
+BFs do not to need to be strongly in favor of an effect across all investigated priors to conclude that an effect is supported.
+However, to draw decisive conclusions in favor of an effect, the evidence for that effect should be stronger than anecdotal for at least a subset of reasonable priors.
+
+The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}).
+Because such large interference effects have been found previously (see also our new Figure \textcolor{red}{X on page X} showing all estimates of \cite{vandyke07} and \cite{mertzen}), we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from 8 to 80 ms for the reading times.
+
+We have revised the description of the SPR results to include more nuance regarding the BFs under the different priors:
+
+\textcolor{red}{add text from the SPR results after new analyses are finished}
+
+For the analysis of the ERP data, we changed the largest prior from N(0,5) to N(0,2), so from effects ranging up to 11 $\mu V$ to effects ranging up to 4 $\mu V$, to accommodate the concern that we might focus on too large effect sizes.
+
+We revised the description of the BF analysis for the ERP data as follows (see page \textcolor{red}{X}):
+
+\textcolor{red}{add description of ERP BF analysis}
 
@@
 \subsubsection*{Response to R2.1}
-\textcolor{red}{??}
+We used a well-studied design and the previous studies using it, specifically the same syntactic retrieval cue, found effects of syntactic interference \citep{vandyke07, mertzen}.
+Therefore, while our results might not be surprising for some researchers because the design might be flawed, others might be convinced that the design is very appropriate to investigate syntactic interference, because it produced such effects in the past.
+We believe that our results add a valuable contribution to the literature on syntactic interference as a counterpoint to the previous studies.
+
+\textcolor{red}{We added the following text to the manuscript to highlight that our conclusions only apply to the design used in the present study:}
 
@@
 \subsubsection*{Response to R2.2}
-\textcolor{red}{??}
+\textcolor{red}{We agree with this statement and therefore refrained from interpreting the results at the critical region.\\
+Maybe add sth to manuscript}
 
@@
 \subsubsection*{Response to R2.3}
-\textcolor{red}{??}
+The participants were able to discern the two syntactic conditions at region 8 because this is the region were the conditions started to differ.
+We have added the words of an example item as x-axis labels to Figure \ref{fig:whole_sentence} on page \pageref{fig:whole_sentence} to illustrate clearly.
+For convenience, we paste the Figure here:
+\newpage
+\Paste{figwhole_sentence}
 
@@
 \subsubsection*{Response to R2.4}
-\textcolor{red}{??}
+\textcolor{red}{Emphasize BF results of the interaction}
 
@@
 \Paste{why_n400}
+
+\vspace{1em}
+\noindent Following, the argumentation that interference, especially syntactic interference, might influence the P600, we have added analyses of the spatio-temporal P600 window.
+These new analyses provide evidence against an effect of syntactic interference on the P600 and moderate bordering weak evidence for an effect of semantic interference.
+We added the following text to the Results and Discussion.
+
+\textcolor{red}{add p600 results and discussion}
 
@@
 ``\textbf{The authors offer only limited discussion of why their results differed from those of other studies.
-To some degree, the paper seems to take the position that since the current studies had huge numbers of participants and hence high statistical power, this study reveals the true state of the world, and other studies should be taken less seriously because their results were less reliable.
-This may be an over-simplistic approach to the findings.}''
+To some degree, the paper seems to take the position that since the current studies had huge numbers of participants and hence high statistical power, this study reveals the true state of the world, and other studies should be taken less seriously because their results were less reliable.}''
 \end{quote}
@@
 \begin{quote}
-``First, \textbf{could there have been an impact of the fact that the study included an unusually large number of experimental items with identical structures, leading to practice effects?}''
+``This may be an over-simplistic approach to the findings.
+First, \textbf{could there have been an impact of the fact that the study included an unusually large number of experimental items with identical structures, leading to practice effects?}''
 \end{quote}
@@
 \subsubsection*{Response to R3.6}
-We added the following regarding the filler items to the Procedure section of Experiment 1, page \pageref{fillers1}:
+We have added the following regarding the filler items to the Procedure section of Experiment 1, page \pageref{fillers1}:
 
@@
 
-\noindent We also added a brief statement on fillers to the Procedure section of Experiment 2 (page \pageref{fillers2}) and corrected a typo regarding the number of trials per experimental block:
+\noindent We have also added a brief statement on fillers to the Procedure section of Experiment 2 (page \pageref{fillers2}) and corrected a typo regarding the number of trials per experimental block:
 
@@
 \subsubsection*{Response to R3.7}
-\textcolor{red}{prominence felix, reduced interference effects, linebreaks VD07?}
+The difference in methodology lead to another difference between our study and the previous studies.
+In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line.
+Whether \citeauthor{vandyke07}'s \citeyear{vandyke07} SPR experiment used line breaks or not is not reported.
+In our SPR experiments, it was necessary to introduce line breaks into the sentences.
+These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb.
+The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor.
+It has been proposed by \textcite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability.
+Therefore, increased prominence of the subject could have decreased interference.
+Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference.
+For this point to hold, one would need to assume that the line break provided additional structural information by emphasizing the syntactic structure of the items in our SPR experiment.
+However, the EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor.
+So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
+Nevertheless, neither our SPR nor our EEG experiments showed evidence for syntactic interference.
+
+All our experiments (SPR and EEG) displayed the subject with a comma.
+Commas are mandatory in German to separate the subordinate clauses from the main clause.
+Similar to the line break discussed above, the comma could have emphasized the syntactic structure of the sentences and therefore decreased syntactic interference from the distractor which occurred inside a relative clause and not in the clause of the critical verb.
+But if this comma provided information about the syntactic structure of the items then it should have so as well in \citeauthor{mertzen}'s \citeyear{mertzen} German study because they also presented the subject with the comma.
+So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.
 
@@
 ``\textbf{Third, in looking through the full set of experimental items, it appears that the inappropriate subject NP is always the subject of an Aux + Adj predicate, rather than, say, the agent of an action verb.
-Could this different status of that subject have contributed to the lack of a syntactic interference effect?}''
-\end{quote}
+Could this different status of that subject have contributed to the lack of a syntactic interference effect?}
 
-\subsubsection*{Response to R3.8}
-\textcolor{red}{see items VD07}
-
-\begin{quote}
-``Summarizing, I am overall enthusiastic about this study.
+Summarizing, I am overall enthusiastic about this study.
 But I am less sure of how we should generalize from this specific finding using one specific manipulation.
@@
 
+\subsubsection*{Response to R3.8}
+We used the same design as \textcite{vandyke07} and \textcite{mertzen}.
+We manipulated syntactic interference in the same way as they did.
+Nevertheless, their data showed syntactic interference and our data do not.
+Apart from the reasons that we discuss in the manuscript (cross-linguistic differences, power, partially different methods), it is unclear why the results differ between the studies.
+
+However, \textcite{mertzen} added an additional distractor to the items which is relevant regarding the subject status of the distractor(s).
+We had previously not described this in our manuscript, in the revision we have added the following paragraph to page \pageref{items_mertzen}:
+
+\setcounter{exx}{4}
+\Paste{items_mertzen}
+
 \end{document}

### 2024-07-12 code_data/EEG/r_scripts/analyses.R
Updated via the Open Science Framework

@@ -1,5 +1,5 @@
 # author: Pia Schoknecht
-# date: June 14, 2024
+# date: June 21, 2024
 # R version: 4.3.3
 
 #load packages
@@ -7,143 +7,162 @@ library(tidyverse)
 library(brms)
 library(bayesplot)
 library(ggplot2)
+library(cowplot)
 
+#### prepare ####
 #load data
 prestim <- read_csv("data/Pandora_mean_200_0.csv")
 n400 <- read_csv("data/Pandora_mean_300_500.csv")
-
+p600 <- read_csv("data/Pandora_mean_600_900.csv")
+ 
 # remove subjects who have only one good session
 xtabs(~ session + subject, n400)
 n400.2 <- n400 %>% filter(subject != "0001" & subject != "0005" & subject != "0006" & subject != "0011" & subject != "0021"
-                      & subject != "0040" & subject != "0041" & subject != "0055" & subject != "0062" & subject != "0067"
-                      & subject != "0069" & subject != "0079" & subject != "0085" & subject != "0089" & subject != "0094"
-                      & subject != "0095" & subject != "0106" & subject != "0122" & subject != "0123" & subject != "0129"
-                      & subject != "0150")
+                           & subject != "0040" & subject != "0041" & subject != "0055" & subject != "0062" & subject != "0067"
+                           & subject != "0069" & subject != "0079" & subject != "0085" & subject != "0089" & subject != "0094"
+                           & subject != "0095" & subject != "0106" & subject != "0122" & subject != "0123" & subject != "0129"
+                           & subject != "0150")
 n400.2$subject <- factor(n400.2$subject)
-
+xtabs(~ session + subject, n400.2)
+ 
+p600.2 <- p600 %>% filter(subject != "0001" & subject != "0005" & subject != "0006" & subject != "0011" & subject != "0021"
+                       & subject != "0040" & subject != "0041" & subject != "0055" & subject != "0062" & subject != "0067"
+                       & subject != "0069" & subject != "0079" & subject != "0085" & subject != "0089" & subject != "0094"
[…]

### 2024-07-12 code_data/EEG/r_scripts/analyses.R
Updated via the Open Science Framework

@@ -266,10 +266,10 @@ posteriors_n400 <- mcmc_areas(
                       labs(title= "N400", x = "effect in microvolts")+
                       vline_0(linetype="dashed")+
                       scale_x_continuous(breaks=(seq(-1, 1, 0.2)), limits = c(-1, 1)) +
-                      theme_bw(base_size=9)+
-                      annotate("text", x = 0.4, y = 3.75, label = "-0.33 — -0.14 — 0.05", size=2) +
-                      annotate("text", x = 0.4, y = 2.65, label = "-0.5 — -0.3 — -0.1", size=2) +
-                      annotate("text", x = 0.4, y = 1.7, label = "-0.33 — 0.03 — 0.38", size=2)
+                      theme_bw(base_size=11)+
+                      annotate("text", x = 0.55, y = 3.75, label = "-0.33 — -0.14 — 0.05", size=3) +
+                      annotate("text", x = 0.55, y = 2.65, label = "-0.5 — -0.3 — -0.1", size=3)+
+                      annotate("text", x = 0.55, y = 1.7, label = "-0.33 — 0.03 — 0.38", size=3)
 ggsave("plots/posteriors_eeg_n400.jpg", width = 8, height = 6, units = "cm", dpi=300)
 
 # bfs
@@ -351,10 +351,10 @@ posteriors_p600 <- mcmc_areas(
                         labs(title= "P600", x = "effect in microvolts")+
                         vline_0(linetype="dashed")+
                         scale_x_continuous(breaks=(seq(-1, 1, 0.2)), limits = c(-1, 1)) +
-                        theme_bw(base_size=9)+
-                        annotate("text", x = 0.4, y = 3.75, label = "-0.22 — -0.08 — 0.06", size=2) +
-                        annotate("text", x = 0.4, y = 2.65, label = "-0.37 — -0.21 — -0.05", size=2) +
-                        annotate("text", x = 0.4, y = 1.7, label = "-0.48 — -0.2 — 0.08", size=2)
+                        theme_bw(base_size=11)+
+                        annotate("text", x = 0.55, y = 3.75, label = "-0.22 — -0.08 — 0.06", size=3) +
+                        annotate("text", x = 0.55, y = 2.65, label = "-0.37 — -0.21 — -0.05", size=3) +
+                        annotate("text", x = 0.55, y = 1.7, label = "-0.48 — -0.2 — 0.08", size=3)
[…]

### 2024-07-12 code_data/SPR/r_scripts/compute_BFs.R
Added via the Open Science Framework

@@ -1,0 +1,203 @@
+# author: Pia Schoknecht
+# date: 09.07.2024
+
+# script to compute Bayes Factors with the Savage-Dickey method
+library(tidyverse)
+library(lme4)
+library(brms)
+library(ggplot2)
+library(bayesplot)
+library(truncnorm)
+library(ggbreak)
+library(ggimage) 
+library(cowplot)
+
+# varying priors 
+priors_s <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, coef="syn"),
+  prior(normal(0, 0.01), class = b, coef="sem"),
+  prior(normal(0, 0.01), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+)  
+
+priors_m <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, coef="syn"),
+  prior(normal(0, 0.05), class = b, coef="sem"),
+  prior(normal(0, 0.05), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+)
+
+priors_l <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, coef="syn"),
+  prior(normal(0, 0.1), class = b, coef="sem"),
+  prior(normal(0, 0.1), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+load("E://model_fits/Fit_s_Full_precrit_slopes_symmetrical.Rda")
+load("E://model_fits/Fit_m_Full_precrit_slopes_symmetrical.Rda")
[…]

### 2024-07-12 code_data/SPR/r_scripts/Plot_stats.R
Updated via the Open Science Framework

@@ -1,47 +1,41 @@
 # author: Pia Schoknecht
-# 27.07.2023
+# 10.07.2024
 
 library(tidyverse)
 library(brms)
 library(bayesplot)
 library(ggplot2)
-library(ggbreak)
-library(ggimage) 
 library(cowplot)
-
-
 
 ### Plot Posteriors ###
 
+# Pre-critical region
+
 # load model fits
-load("model_fits/Fit_m_Full_precrit.Rda")
+load("E://model_fits/Fit_m_Full_precrit_slopes_symmetrical.Rda")
 m_precrit <- m_m_full
-load("model_fits/Fit_m_Full_crit.Rda")
-m_crit <- m_m_full
-load("model_fits/Fit_m_Full_spill.Rda")
-m_spill <- m_m_full
+rm(list = c("m_m_full"))
 
-m_precrit
-m_crit
-m_spill
-
+# extract samples
 samples_precrit <- as_draws_df(m_precrit)
-samples_crit <- as_draws_df(m_crit)
-samples_spill <- as_draws_df(m_spill)
+rm(list = c("m_precrit"))
 
+# backtransform estimates to ms scale
 syn_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$b_syn) - exp(samples_precrit$b_Intercept - samples_precrit$b_syn)
 sem_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$b_sem) - exp(samples_precrit$b_Intercept - samples_precrit$b_sem)
-int_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit[,4]) - exp(samples_precrit$b_Intercept - samples_precrit[,4])
+int_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$`b_syn:sem`) - exp(samples_precrit$b_Intercept - samples_precrit$`b_syn:sem`)
 
+# credible intervals
 round(c(mean = mean(syn_eff_precrit), quantile(syn_eff_precrit, probs = c(.025, .975))))
 round(c(mean = mean(sem_eff_precrit), quantile(sem_eff_precrit, probs = c(.025, .975))))
-round(c(mean = mean(int_eff_precrit[-1,]), quantile(int_eff_precrit[,1], probs = c(.025, .975))))
+round(c(mean = mean(int_eff_precrit), quantile(int_eff_precrit, probs = c(.025, .975))))
 
+# add back-transformed values to samples df
[…]

### 2024-07-12 code_data/SPR/r_scripts/Plot_ReadingTimes.R
Updated via the Open Science Framework

@@ -33,10 +33,6 @@ df_nofill <- droplevels(subset(df, label != "filler"))
 # trim all rts below 150 and above 3000 ms
 df_trim<- subset(df_nofill, df_nofill$rt > 150 & df_nofill$rt < 3000)
 
-#extract words from example item for x-axis labels
-allwords_losyn<-unique(df_trim[df_trim$item=="008" & df_trim$condition=="a" ,c("wordno")])
-allwords_hisyn<-unique(df_trim[df_trim$item=="008" & df_trim$condition=="c" ,c("wordno")])
-
 # subsets for hisyn and losyn (because they have different lengths)
 df_hisyn <- df_trim %>% filter(condition == "c" | condition == "d")
 df_losyn <- df_trim %>% filter(condition == "a" | condition == "b")
@@ -46,13 +42,13 @@ df_losyn <- df_trim %>% filter(condition == "a" | condition == "b")
 #calculate mean reading time per subject per word per condition
 hisyn_by_participants <- summarize(
   group_by(
-    filter(df_hisyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19")),
+    filter(df_hisyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "21")),
     wordno,condition,subj, exp),
   logrt=mean(log(rt)))
 
 losyn_by_participants <- summarize(
   group_by(
-    filter(df_losyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18")),
+    filter(df_losyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "20")),
     wordno,condition,subj, exp),
   logrt=mean(log(rt)))
 
@@ -88,60 +84,50 @@ hisyn_summ_by_participants$region<- ifelse(hisyn_summ_by_participants$wordno == 
                                                          ifelse(hisyn_summ_by_participants$wordno == 18, "spill-over",        
                                                                 "other"))))
 
-hisyn_summ_by_participants$word <- hisyn_summ_by_participants$wordno
-losyn_summ_by_participants$word <- losyn_summ_by_participants$wordno
[…]

### 2024-07-12 code_data/SPR/r_scripts/analyses_crit.R
Updated via the Open Science Framework

@@ -1,255 +1,88 @@
 # author: Pia Schoknecht
-# date: 22.07.2023
+# date: 30.07.2023
 
-library(tidyverse)
-library(lme4)
+# script to analyze reading times in the critical region
+
+# load packages
 library(brms)
-library(ggplot2)
-library(bayesplot)
-library(truncnorm)
 
 
 # load data
-crit_trim <- read.csv("../data/pandora_spr_774_crit.csv")
+crit_trim <- read.csv("pandora_spr_774_crit.csv")
+crit_trim$c_trial <- crit_trim$trial/100
 
 
-# Inferential statistics
-# truncated varying priors 
-# (only positive effects -- longer reading times for high interference)
-
-priors_s_tr <- c(
+# priors
+priors_s <- c(
   prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.01), class = b, lb=0),
-  prior(normal(0, 0.5), class = sigma),
-  prior(normal(0, 0.1), class = sd)
-)  
-
-priors_m_tr <- c(
-  prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.01), class = b, coef="syn"),
+  prior(normal(0, 0.01), class = b, coef="sem"),
+  prior(normal(0, 0.01), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
 
-priors_l_tr <- c(
+priors_m <- c(
   prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.05), class = b, coef="syn"),
+  prior(normal(0, 0.05), class = b, coef="sem"),
+  prior(normal(0, 0.05), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
+
+priors_l <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, coef="syn"),
+  prior(normal(0, 0.1), class = b, coef="sem"),
[…]

### 2024-07-12 code_data/SPR/r_scripts/analyses_crit.R
Updated via the Open Science Framework

@@ -1,5 +1,5 @@
 # author: Pia Schoknecht
-# date: 30.07.2023
+# date: 12.07.2024
 
 # script to analyze reading times in the critical region
 

### 2024-07-12 code_data/SPR/r_scripts/analyses_precrit.R
Updated via the Open Science Framework

@@ -1,258 +1,86 @@
 # author: Pia Schoknecht
-# date: 30.07.2023
+# date: 12.07.2024
 
 # script to analyze reading times in the pre-critical region
 
 # load packages
-library(tidyverse)
-library(truncnorm)
 library(brms)
-library(ggplot2)
-library(bayesplot)
 
 
 # load data
-precrit_trim <- read.csv("../data/pandora_spr_774_precrit.csv")
+precrit_trim <- read.csv("pandora_spr_774_precrit.csv")
+precrit_trim$c_trial <- precrit_trim$trial/100
 
-
-# Inferential statistics
-# truncated varying priors 
-# (only positive effects -- longer reading times for high interference)
-
-priors_s_tr <- c(
+# priors
+priors_s <- c(
   prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.01), class = b, lb=0),
-  prior(normal(0, 0.5), class = sigma),
-  prior(normal(0, 0.1), class = sd)
-)  
-
-priors_m_tr <- c(
-  prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.01), class = b, coef="syn"),
+  prior(normal(0, 0.01), class = b, coef="sem"),
+  prior(normal(0, 0.01), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
 
-priors_l_tr <- c(
+priors_m <- c(
   prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.05), class = b, coef="syn"),
+  prior(normal(0, 0.05), class = b, coef="sem"),
+  prior(normal(0, 0.05), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, coef="syn"),
+  prior(normal(0, 0.1), class = b, coef="sem"),
[…]

### 2024-07-12 code_data/SPR/r_scripts/analyses_spill.R
Updated via the Open Science Framework

@@ -1,257 +1,87 @@
 # author: Pia Schoknecht
-# date: 22.07.2023
+# date: 12.07.2024
 
-library(tidyverse)
-library(lme4)
+# script to analyze reading times in the spill-over region
+
+# load packages
 library(brms)
-library(ggplot2)
-library(bayesplot)
-library(parallel)
-library(truncnorm)
 
 
 # load data
-spill_trim <- read.csv("../data/pandora_spr_774_spill.csv")
+spill_trim <- read.csv("pandora_spr_774_spill.csv")
+spill_trim$c_trial <- spill_trim$trial/100
 
 
-# Inferential statistics
-# truncated varying priors 
-# (only positive effects -- longer reading times for high interference)
-
-priors_s_tr <- c(
+# priors
+priors_s <- c(
   prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.01), class = b, lb=0),
-  prior(normal(0, 0.5), class = sigma),
-  prior(normal(0, 0.1), class = sd)
-)  
-
-priors_m_tr <- c(
-  prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.05), class = b, lb=0),
+  prior(normal(0, 0.01), class = b, coef="syn"),
+  prior(normal(0, 0.01), class = b, coef="sem"),
+  prior(normal(0, 0.01), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
   prior(normal(0, 0.5), class = sigma),
   prior(normal(0, 0.1), class = sd)
 ) 
 
-priors_l_tr <- c(
+priors_m <- c(
   prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.1), class = b, lb=0),
+  prior(normal(0, 0.05), class = b, coef="syn"),
+  prior(normal(0, 0.05), class = b, coef="sem"),
+  prior(normal(0, 0.05), class = b, coef="syn:sem"),
+  prior(normal(0, 0.1),  class = b, coef="c_trial"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, coef="syn"),
+  prior(normal(0, 0.1), class = b, coef="sem"),
[…]

### 2024-07-12 code_data/previous_studies/PreviousFindings.Rmd
Updated via the Open Science Framework

@@
 
-M21.G <- d
-M21.G$experiment <- "M21.G (N=121)"
+M23.G <- d
+M23.G$experiment <- "M23.G (N=121)"
 
 load("M21_byregionsummaries_en.rda")
-M21.E <- d
-M21.E$experiment <- "M21.E (N=61)"
+M23.E <- d
+M23.E$experiment <- "M23.E (N=61)"
 
-M21estimates <- rbind(M21.G, M21.E[,-7])
+M23estimates <- rbind(M23.G, M23.E[,-7])
 
-M21estimates$int_type <- ifelse(M21estimates$effect%in%c("Subject"),"syntactic",
-                              ifelse(M21estimates$effect%in%c("Animacy"),"semantic", 
+M23estimates$int_type <- ifelse(M23estimates$effect%in%c("Subject"),"syntactic",
+                              ifelse(M23estimates$effect%in%c("Animacy"),"semantic", 
                                    "interaction"))
 
-M21estimates$region <- as.factor(M21estimates$region)
-levels(M21estimates$region)[3] <- "spill-over"
+M23estimates$region <- as.factor(M23estimates$region)
+levels(M23estimates$region)[3] <- "spill-over"
 
-M21estimates <- M21estimates[,-4]
-M21estimates <- M21estimates[,-5]
-M21estimates <- M21estimates[,-6]
+M23estimates <- M23estimates[,-4]
+M23estimates <- M23estimates[,-5]
+M23estimates <- M23estimates[,-6]
 
@@
 
-previous <- rbind(VD07estimates, M21estimates)
+previous <- rbind(VD07estimates, M23estimates)
 save(previous, file="VD07_M21_estimates.rda")
@@
 # prepare Schoknecht & Vasishth estimates (numbers from script "Plot_stats.R")
-SVestimates <- data.frame(mean   = c(1, 11, 7, 2, 8, 14),
-                          lower  = c(0, 8, 1, 0, 4, 7),
-                          upper = c(2, 14, 13, 4, 12, 22),
-                           region = c(rep(c("pre-critical"),3), rep(c("critical"), 3)),
-                           measure  = c(rep(c("SPR"), 6)),
-                           experiment = c(rep(c("present (N=774)"), 6)),
-                           int_type = c(rep(c("syntactic", "semantic", "interaction"), 2))
+SVestimates <- data.frame(mean   = c(3, 14, 8, -2, 10, 14, 8, 14, 8),
+                          lower  = c(-4, 7, -5, -11, 1, -3, 1, 7, -6),
+                          upper = c(10, 21, 22, 7, 19, 31, 15, 21, 21),
+                           region = c(rep(c("pre-critical"),3), rep(c("critical"), 3), rep(c("spill-over"), 3)),
+                           measure  = c(rep(c("SPR"), 9)),
+                           experiment = c(rep(c("present (N=774)"), 9)),
+                           int_type = c(rep(c("syntactic", "semantic", "interaction"), 3))
                           )
@@
 estimates <- rbind(previous, SVestimates)
-#estimates$experiment <- factor(estimates$experiment,levels=c("VD07.E1", "VD07.E2","VD07.E3", "present", "M21.G","M21.E"))
 
-pp.plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical")),
+pp.plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over")),
                    aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
@@
              size=0.8, position=position_dodge(1.45))+
-  #scale_y_continuous(breaks = c(-50,-40,-30,-20, -10, 0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100))+
-  scale_shape_manual(values = c(18,16,17,15))+
+  scale_shape_manual(values = c(18,16,17,15), breaks=c("SPR", "FPRT", "RPD", "TFT"),
+                         labels=c("self-paced reading time", "first-pass reading time", "regression path duration", "total fixation time"))+
   scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
@@
   ylab("reading time difference (ms)")+
-  ggtitle("Present and previous findings")+
+  #ggtitle("Present and previous findings")+
   guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
 
-pp.plot + coord_cartesian(ylim=c(-10, 30))  
-ggsave("plots/present_previous_findings.jpg", width=8, height=8, dpi=600)
+pp.plot + coord_cartesian(ylim=c(-25, 75))  
+ggsave("plots/present_previous_findings.jpg", width=9, height=8, dpi=600)
 

### 2024-07-13 code_data/previous_studies/PreviousFindings.Rmd
Updated via the Open Science Framework

@@
 VD07E2critSemTFTmn <- (761+653)/2-(738+630)/2
-VD07E2critSemTFTtval <- 1.95 # reported as non-sig.
-VD07E2critSemTFTSE <- VD07E2critSemTFTmn/VD07E2critSemTFTtval
+VD07E2critSemTFTSE <- 42 #highest SE in critical region, Table 7
 VD07E2critSemTFTlower <- round(VD07E2critSemTFTmn - 2*VD07E2critSemTFTSE)
@@
 VD07E2critIntFPRTmn <- (418-413)-(382-376)
-VD07E2critIntFPRTtval <- 1.95 # reported as non-sig.
-VD07E2critIntFPRTSE <- VD07E2critIntFPRTmn/VD07E2critIntFPRTtval
+VD07E2critIntFPRTSE <- 21 #highest SE in critical region, Table 7
 VD07E2critIntFPRTlower <- round(VD07E2critIntFPRTmn - 2*VD07E2critIntFPRTSE)
@@
 VD07E2critIntRPDmn <- (663-594)-(495-454)
-VD07E2critIntRPDtval <- 1.95 # reported as non-sig.
-VD07E2critIntRPDSE <- VD07E2critIntRPDmn/VD07E2critIntRPDtval
+VD07E2critIntRPDSE <- 44 #highest SE in critical region, Table 7
 VD07E2critIntRPDlower <- round(VD07E2critIntRPDmn - 2*VD07E2critIntRPDSE)
@@
 VD07E2critIntTFTmn <- (761-738)-(653-630)
-VD07E2critIntTFTtval <- 1.95 # reported as non-sig.
-VD07E2critIntTFTSE <- VD07E2critIntTFTmn/VD07E2critIntTFTtval
+VD07E2critIntTFTSE <- 42 #highest SE in critical region, Table 7
 VD07E2critIntTFTlower <- round(VD07E2critIntTFTmn - 2*VD07E2critIntTFTSE)
@@
 VD07E2postcritSynTFTmn <- (491+493)/2-(502+540)/2
-VD07E2postcritSynTFTtval <- 1.95 # reported in paper as non-sig
-VD07E2postcritSynTFTSE<- VD07E2postcritSynTFTmn/VD07E2postcritSynTFTtval
+VD07E2postcritSynTFTSE <- 30 #highest SE in spill-over region, Table 7
 VD07E2postcritSynTFTlower <- round(VD07E2postcritSynTFTmn - 2*VD07E2postcritSynTFTSE)
@@
 VD07E2postcritSemFPRTmn<- (296+364)/2-(325+320)/2
-VD07E2postcritSemFPRTtval<- 1.95 # reported in paper as non-sig
-VD07E2postcritSemFPRTSE<-VD07E2postcritSemFPRTmn/VD07E2postcritSemFPRTtval
+VD07E2postcritSemFPRTSE<- 21 #highest SE in spill-over region, Table 7
 VD07E2postcritSemFPRTlower<-VD07E2postcritSemFPRTmn-2*VD07E2postcritSemFPRTSE
@@
 VD07E2postcritSemRPDmn <- (1295+1205)/2-(1365+970)/2
-VD07E2postcritSemRPDtval <- 1.95 # reported in paper as non-sig
-VD07E2postcritSemRPDSE <- VD07E2postcritSemRPDmn/VD07E2postcritSemRPDtval
+VD07E2postcritSemRPDSE <- 140 #highest SE in spill-over region, Table 7
 VD07E2postcritSemRPDlower <- round(VD07E2postcritSemRPDmn - 2*VD07E2postcritSemRPDSE)
@@
 VD07E2postcritSemTFTmn <- (493+540)/2-(491+502)/2
-VD07E2postcritSemTFTtval <- 1.95 # reported in paper as non-sig
-VD07E2postcritSemTFTSE <- VD07E2postcritSemTFTmn/VD07E2postcritSemTFTtval
+VD07E2postcritSemTFTSE <-  30 #highest SE in spill-over region, Table 7
 VD07E2postcritSemTFTlower <- round(VD07E2postcritSemTFTmn - 2*VD07E2postcritSemTFTSE)
@@
 VD07E2postcritIntTFTmn <- (493-491)-(540-502)
-VD07E2postcritIntTFTtval <- 1.95 # reported in paper as non-sig
-VD07E2postcritIntTFTSE <-  VD07E2postcritIntTFTmn/VD07E2postcritIntTFTtval
+VD07E2postcritIntTFTSE <- 30 #highest SE in spill-over region, Table 7
 VD07E2postcritIntTFTlower <- round(VD07E2postcritIntTFTmn - 2*VD07E2postcritIntTFTSE)
@@
 VD07E2finalSynFPRTmn<-(259+271)/2-(286+274)/2
-VD07E2finalSynFPRTtval<- 1.95 # reported in paper as non-sig
-VD07E2finalSynFPRTSE<-VD07E2finalSynFPRTmn/VD07E2finalSynFPRTtval
+VD07E2finalSynFPRTSE<- 22 #highest SE in sentence-final region, Table 7
 VD07E2finalSynFPRTlower<-VD07E2finalSynFPRTmn-2*VD07E2finalSynFPRTSE
@@
 VD07E2finalSynTFTmn <- (349+360)/2-(362+373)/2
-VD07E2finalSynTFTtval <- 1.95 # reported in paper as non-sig
-VD07E2finalSynTFTSE<- VD07E2finalSynTFTmn/VD07E2finalSynTFTtval
+VD07E2finalSynTFTSE<- 40 #highest SE in sentence-final region, Table 7
 VD07E2finalSynTFTlower <- round(VD07E2finalSynTFTmn - 2*VD07E2finalSynTFTSE)
@@
 VD07E2finalSemFPRTmn<- (271+274)/2-(259+286)/2
-VD07E2finalSemFPRTtval<- 1.95 # reported in paper as non-sig
-VD07E2finalSemFPRTSE<-VD07E2finalSemFPRTmn/VD07E2finalSemFPRTtval
+VD07E2finalSemFPRTSE<- 22 #highest SE in sentence-final region, Table 7
 VD07E2finalSemFPRTlower<-VD07E2finalSemFPRTmn-2*VD07E2finalSemFPRTSE
@@
 VD07E2finalSemTFTmn <- (360+373)/2-(349+362)/2
-VD07E2finalSemTFTtval <- 1.95 # reported in paper as non-sig
-VD07E2finalSemTFTSE <- VD07E2finalSemTFTmn/VD07E2finalSemTFTtval
+VD07E2finalSemTFTSE <- 40 #highest SE in sentence-final region, Table 7
 VD07E2finalSemTFTlower <- round(VD07E2finalSemTFTmn - 2*VD07E2finalSemTFTSE)
@@
 VD07E2finalIntFPRTmn<- (271-259)-(274-286)
-VD07E2finalIntFPRTtval<- 1.95 # reported in paper as non-sig
-VD07E2finalIntFPRTSE<-VD07E2finalIntFPRTmn/VD07E2finalIntFPRTtval
+VD07E2finalIntFPRTSE<- 22 #highest SE in sentence-final region, Table 7
 VD07E2finalIntFPRTlower<-VD07E2finalIntFPRTmn-2*VD07E2finalIntFPRTSE
@@
 VD07E2finalIntTFTmn <- (360-349)-(373-362)
-VD07E2finalIntTFTtval <- 1.95 # reported in paper as non-sig
-VD07E2finalIntTFTSE <-  VD07E2finalIntTFTmn/VD07E2finalIntTFTtval
+VD07E2finalIntTFTSE <-  40 #highest SE in sentence-final region, Table 7
 VD07E2finalIntTFTlower <- round(VD07E2finalIntTFTmn - 2*VD07E2finalIntTFTSE)
@@
 VD07E3precritSynFPRTmn <- (462+447)/2-(449+455)/2
-VD07E3precritSynFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3precritSynFPRTSE <- VD07E3precritSynFPRTmn/VD07E3precritSynFPRTtval
+VD07E3precritSynFPRTSE <- 18 #highest SE in pre-crit region, Table 16
 VD07E3precritSynFPRTlower <- round(VD07E3precritSynFPRTmn-2*VD07E3precritSynFPRTSE)
@@
 VD07E3precritSemFPRTmn <- (455+447)/2-(449+462)/2
-VD07E3precritSemFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3precritSemFPRTSE <- VD07E3precritSemFPRTmn/VD07E3precritSemFPRTtval
+VD07E3precritSemFPRTSE <- 18 #highest SE in pre-crit region, Table 16
 VD07E3precritSemFPRTlower <- round(VD07E3precritSemFPRTmn-2*VD07E3precritSemFPRTSE)
@@
 VD07E3precritIntFPRTmn <- (447-462)-(455-449)
-VD07E3precritIntFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3precritIntFPRTSE <- VD07E3precritIntFPRTmn/VD07E3precritIntFPRTtval
+VD07E3precritIntFPRTSE <- 18 #highest SE in pre-crit region, Table 16
 VD07E3precritIntFPRTlower <- round(VD07E3precritIntFPRTmn-2*VD07E3precritIntFPRTSE)
@@
 VD07E3precritIntRPDmn <- (551-490)-(493-472)
-VD07E3precritIntRPDtval <- 1.95 # reported in paper as non-sig
-VD07E3precritIntRPDSE <- VD07E3precritIntRPDmn/VD07E3precritIntRPDtval
+VD07E3precritIntRPDSE <- 36 #highest SE in pre-crit region, Table 16
 VD07E3precritIntRPDlower <- round(VD07E3precritIntRPDmn-2*VD07E3precritIntRPDSE)
@@
 VD07E3precritIntTFTmn <- (710-622)-(640-586)
-VD07E3precritIntTFTtval <- 1.95 # reported in paper as non-sig
-VD07E3precritIntTFTSE <- VD07E3precritIntTFTmn/VD07E3precritIntTFTtval
+VD07E3precritIntTFTSE <- 47 #highest SE in pre-crit region, Table 16
 VD07E3precritIntTFTlower <- round(VD07E3precritIntTFTmn-2*VD07E3precritIntTFTSE)
@@
 VD07E3critSynTFTmn <- (451+467)/2-(414+421)/2
-VD07E3critSynTFTtval <- 1.95 # reported in paper as non-sig
-VD07E3critSynTFTSE <- VD07E3critSynTFTmn/VD07E3critSynTFTtval
+VD07E3critSynTFTSE <- 38 #highest SE in crit region, Table 16
 VD07E3critSynTFTlower <- round(VD07E3critSynTFTmn-2*VD07E3critSynTFTSE)
@@
 VD07E3critSemFPRTmn <- (282+280)/2-(274+294)/2
-VD07E3critSemFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3critSemFPRTSE <- VD07E3critSemFPRTmn/VD07E3critSemFPRTtval
+VD07E3critSemFPRTSE <- 11 #highest SE in crit region, Table 16
 VD07E3critSemFPRTlower <- round(VD07E3critSemFPRTmn-2*VD07E3critSemFPRTSE)
@@
 VD07E3critSemRPDmn <- (315+344)/2-(314+354)/2
-VD07E3critSemRPDtval <- 1.95 # reported in paper as non-sig
-VD07E3critSemRPDSE <- VD07E3critSemRPDmn/VD07E3critSemRPDtval
+VD07E3critSemRPDSE <- 23 #highest SE in crit region, Table 16
 VD07E3critSemRPDlower <- round(VD07E3critSemRPDmn-2*VD07E3critSemRPDSE)
@@
 VD07E3critSemTFTmn <- (421+467)/2-(414+451)/2
-VD07E3critSemTFTtval <- 1.95 # reported in paper as non-sig
-VD07E3critSemTFTSE <- VD07E3critSemTFTmn/VD07E3critSemTFTtval
+VD07E3critSemTFTSE <- 38 #highest SE in crit region, Table 16
 VD07E3critSemTFTlower <- round(VD07E3critSemTFTmn-2*VD07E3critSemTFTSE)
@@
 VD07E3critIntRPDmn <- (344-354)-(315-314)
-VD07E3critIntRPDtval <- 1.95 # reported in paper as non-sig
-VD07E3critIntRPDSE <- VD07E3critIntRPDmn/VD07E3critIntRPDtval
+VD07E3critIntRPDSE <- 23 #highest SE in crit region, Table 16
 VD07E3critIntRPDlower <- round(VD07E3critIntRPDmn-2*VD07E3critIntRPDSE)
@@
 VD07E3critIntTFTmn <- (467-451)-(421-414)
-VD07E3critIntTFTtval <- 1.95 # reported in paper as non-sig
-VD07E3critIntTFTSE <- VD07E3critIntTFTmn/VD07E3critIntTFTtval
+VD07E3critIntTFTSE <- 38 #highest SE in crit region, Table 16
 VD07E3critIntTFTlower <- round(VD07E3critIntTFTmn-2*VD07E3critIntTFTSE)
@@
 VD07E3postcritSynFPRTmn <- (424+442)/2-(437+448)/2
-VD07E3postcritSynFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3postcritSynFPRTSE <- VD07E3postcritSynFPRTmn/VD07E3postcritSynFPRTtval
+VD07E3postcritSynFPRTSE <- 24 #highest SE in spill-over region, Table 16
 VD07E3postcritSynFPRTlower <- round(VD07E3postcritSynFPRTmn-2*VD07E3postcritSynFPRTSE)
@@
 VD07E3postcritSynTFTmn <- (766+731)/2-(707+711)/2
-VD07E3postcritSynTFTtval <- 1.95 # reported in paper as non-sig
-VD07E3postcritSynTFTSE <- VD07E3postcritSynTFTmn/VD07E3postcritSynTFTtval
+VD07E3postcritSynTFTSE <- 59 #highest SE in spill-over region, Table 16
 VD07E3postcritSynTFTlower <- round(VD07E3postcritSynTFTmn-2*VD07E3postcritSynTFTSE)
@@
 VD07E3postcritSemFPRTmn <- (437+424)/2-(448+442)/2
-VD07E3postcritSemFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3postcritSemFPRTSE <- VD07E3postcritSemFPRTmn/VD07E3postcritSemFPRTtval
+VD07E3postcritSemFPRTSE <- 24 #highest SE in spill-over region, Table 16
 VD07E3postcritSemFPRTlower <- round(VD07E3postcritSemFPRTmn-2*VD07E3postcritSemFPRTSE)
@@
 VD07E3postcritIntFPRTmn <- (424-442)-(437-448)
-VD07E3postcritIntFPRTtval <- 1.95 # reported in paper as non-sig
-VD07E3postcritIntFPRTSE <- VD07E3postcritIntFPRTmn/VD07E3postcritIntFPRTtval
+VD07E3postcritIntFPRTSE <-24 #highest SE in spill-over region, Table 16
 VD07E3postcritIntFPRTlower <- round(VD07E3postcritIntFPRTmn-2*VD07E3postcritIntFPRTSE)
@@
 VD07E3postcritIntRPDmn <- (615-625)-(553-607)
-VD07E3postcritIntRPDtval <- 1.95 # reported in paper as non-sig
-VD07E3postcritIntRPDSE <- VD07E3postcritIntRPDmn/VD07E3postcritIntRPDtval
+VD07E3postcritIntRPDSE <- 41 #highest SE in spill-over region, Table 16
 VD07E3postcritIntRPDlower <- round(VD07E3postcritIntRPDmn-2*VD07E3postcritIntRPDSE)
@@
 VD07E3postcritIntTFTmn <- (766-731)-(707-711)
-VD07E3postcritIntTFTtval <- 1.95 # reported in paper as non-sig
-VD07E3postcritIntTFTSE <- VD07E3postcritIntTFTmn/VD07E3postcritIntTFTtval
+VD07E3postcritIntTFTSE <- 59 #highest SE in spill-over region, Table 16
 VD07E3postcritIntTFTlower <- round(VD07E3postcritIntTFTmn-2*VD07E3postcritIntTFTSE)
@@
 estimates <- rbind(previous, SVestimates)
+estimates$experiment <- factor(estimates$experiment, levels=c("VD07.E1 (N=35)", "VD07.E2 (N=36)", "VD07.E3 (N=40)", "M23.E (N=61)", "M23.G (N=121)", "present (N=774)"))
 
@@
 
-pp.plot + coord_cartesian(ylim=c(-25, 75))  
+RTs.plot <- pp.plot + coord_cartesian(ylim=c(-75, 150)) + ggtitle("Reading time differences") 
 ggsave("plots/present_previous_findings.jpg", width=9, height=8, dpi=600)
 
+pp2.plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over") & measure %in%c("SPR", "TFT")),
+                   aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
+    geom_pointrange(aes(color=experiment, shape=measure), 
+             size=0.8, position=position_dodge(1.45))+
+  scale_shape_manual(values = c(18,16,17,15), breaks=c("SPR", "FPRT", "RPD", "TFT"),
+                         labels=c("self-paced reading time", "first-pass reading time", "regression path duration", "total fixation time"))+
+  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
+  geom_hline(yintercept=0, linetype="solid", alpha=.5)+
+  facet_grid(region~int_type, scales="free", space="free") +
+  theme_bw(base_size = 15)+
+  scale_x_discrete(breaks = NULL)+
+  theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
+  xlab("")+
+  ylab("reading time difference (ms)")+
+  #ggtitle("Present and previous findings")+
+  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
+
+pp2.plot + coord_cartesian(ylim=c(-50, 100))  
+ggsave("plots/present_previous_findings_SPR_TFT.jpg", width=9, height=8, dpi=600)
+
 ```
+
+# Compute BF10 for Van Dyke (2007) data
+```{r}
+# function to compute BF10 repeatedly
+compute_BF <- function(mean_effect, se){
+
+  #universal values for a reasonable prior: N(0,0.5), on the ms scale [-40, 40]
+    xbar <- mean_effect
+    SE <- se
+    mu_prior<-0
+    prior_variance<-40^2
+    a <- 1/prior_variance
+    b <- 1/SE^2
+    mu_posterior <- (a*mu_prior+b*xbar)/(a+b)
+    mu_posterior
+
+    var_posterior <- 1/(a+b)
+    sqrt(var_posterior)
+
+    mu_posterior <- (a*mu_prior+b*mean_effect)/(a+1/SE^2)
+    var_posterior <- 1/(a+(1/SE^2))
+    # BF10 (1/(density under posterior at 0 / density under prior at 0):
+    1/(dnorm(0,mu_posterior,sqrt(var_posterior))/dnorm(0,0,40))
+}
+
+##### E1 critical ####
+bf.E1.syn.crit <- compute_BF(VD07E1critSynmn, VD07E1critSynSE)
+bf.E1.sem.crit <- compute_BF(VD07E1critSemmn, VD07E1critSemSE)
+bf.E1.int.crit <- compute_BF(VD07E1critIntmn, VD07E1critIntSE)
+
+##### E1 spill-over ####
+bf.E1.syn.spill <- compute_BF(VD07E1spillSynmn, VD07E1spillSynSE)
+bf.E1.sem.spill <- compute_BF(VD07E1spillSemmn, VD07E1spillSemSE)
+bf.E1.int.spill <- compute_BF(VD07E1spillIntmn, VD07E1spillIntSE)
+
+##### E2 critical #### 
+# for the eye-tracking experiments, we just compute BF10 for TFT 
+bf.E2.syn.crit <- compute_BF(VD07E2critSynTFTmn, VD07E2critSynTFTSE)
+bf.E2.sem.crit <- compute_BF(VD07E2critSemTFTmn, VD07E2critSemTFTSE)
+bf.E2.int.crit <- compute_BF(VD07E2critIntTFTmn, VD07E2critIntTFTSE)
+
+##### E2 spill-over #### 
+bf.E2.syn.spill <- compute_BF(VD07E2postcritSynTFTmn, VD07E2postcritSynTFTSE)
+bf.E2.sem.spill <- compute_BF(VD07E2postcritSemTFTmn, VD07E2postcritSemTFTSE)
+bf.E2.int.spill <- compute_BF(VD07E2postcritIntTFTmn, VD07E2postcritIntTFTSE)
+
+##### E3 pre-critical #### 
+bf.E3.syn.precrit <- compute_BF(VD07E3precritSynTFTmn, VD07E3precritSynTFTSE)
+bf.E3.sem.precrit <- compute_BF(VD07E3precritSemTFTmn, VD07E3precritSemTFTSE)
+bf.E3.int.precrit <- compute_BF(VD07E3precritIntTFTmn, VD07E3precritIntTFTSE)
+
+##### E3 critical #### 
+bf.E3.syn.crit <- compute_BF(VD07E3critSynTFTmn, VD07E3critSynTFTSE)
+bf.E3.sem.crit <- compute_BF(VD07E3critSemTFTmn, VD07E3critSemTFTSE)
+bf.E3.int.crit <- compute_BF(VD07E3critIntTFTmn, VD07E3critIntTFTSE)
+
+##### E3 spill-over #### 
+bf.E3.syn.spill <- compute_BF(VD07E3postcritSynTFTmn, VD07E3postcritSynTFTSE)
+bf.E3.sem.spill <- compute_BF(VD07E3postcritSemTFTmn, VD07E3postcritSemTFTSE)
+bf.E3.int.spill <- compute_BF(VD07E3postcritIntTFTmn, VD07E3postcritIntTFTSE)
+
+# save BFs in df
+df.bf_vd <- data.frame(matrix(ncol=8,nrow=0))
+colnames(df.bf_vd) <- c("Experiment", "BF_method", "Language", "Region", "Effect","Prior", "truncated", "BF10")
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "critical", "syntactic","Normal(0, 0.05)","no", round(bf.E1.syn.crit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "critical", "semantic","Normal(0, 0.05)","no", round(bf.E1.sem.crit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "critical", "interaction","Normal(0, 0.05)","no", round(bf.E1.int.crit,2))
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "spill-over", "syntactic","Normal(0, 0.05)","no", round(bf.E1.syn.spill,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "spill-over", "semantic","Normal(0, 0.05)","no", round(bf.E1.sem.spill,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E1", "S-D", "English", "spill-over", "interaction","Normal(0, 0.05)","no", round(bf.E1.int.spill,2))
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "critical", "syntactic","Normal(0, 0.05)","no", round(bf.E2.syn.crit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "critical", "semantic","Normal(0, 0.05)","no", round(bf.E2.sem.crit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "critical", "interaction","Normal(0, 0.05)","no", round(bf.E2.int.crit,2))
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "spill-over", "syntactic","Normal(0, 0.05)","no", round(bf.E2.syn.spill,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "spill-over", "semantic","Normal(0, 0.05)","no", round(bf.E2.sem.spill,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E2", "S-D", "English", "spill-over", "interaction","Normal(0, 0.05)","no", round(bf.E2.int.spill,2))
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "pre-critical", "syntactic","Normal(0, 0.05)","no", round(bf.E3.syn.precrit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "pre-critical", "semantic","Normal(0, 0.05)","no", round(bf.E3.sem.precrit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "pre-critical", "interaction","Normal(0, 0.05)","no", round(bf.E3.int.precrit,2))
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "critical", "syntactic","Normal(0, 0.05)","no", round(bf.E3.syn.crit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "critical", "semantic","Normal(0, 0.05)","no", round(bf.E3.sem.crit,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "critical", "interaction","Normal(0, 0.05)","no", round(bf.E3.int.crit,2))
+
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "spill-over", "syntactic","Normal(0, 0.05)","no", round(bf.E3.syn.spill,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "spill-over", "semantic","Normal(0, 0.05)","no", round(bf.E3.sem.spill,2))
+df.bf_vd[nrow(df.bf_vd)+1,] <- c("VD07.E3", "S-D", "English", "spill-over", "interaction","Normal(0, 0.05)","no", round(bf.E3.int.spill,2))
+```
+
+# Plot BFs
+```{r}
+# load other BFs
+load("BFs_mertzen23.Rda")
+df.bf_m23$Experiment <- ifelse(df.bf_m23$Language == "English", "M23.E", "M23.G")
+
+load("BFs_schoknecht.Rda")
+df.bf_schoknecht$Experiment <- "present"
+
+df.bf <- rbind(df.bf_vd, df.bf_m23, df.bf_schoknecht)
+
+# prepare
+df.bf$Prior2 <- ifelse(df.bf$Prior == "Normal(0, 0.01)", "N(0, 0.01)", 
+                        ifelse(df.bf$Prior == "Normal(0, 0.05)", "N(0, 0.05)", 
+                              ("N(0, 0.1)")))
+df.bf$BF10 <- as.numeric(df.bf$BF10)
+df.bf$effect <- factor(df.bf$Effect, levels=c("syntactic","semantic", "interaction"))
+df.bf$region <- factor(df.bf$Region, levels=c("pre-critical", "critical", "spill-over"))
+df.bf$experiment <- factor(df.bf$Experiment, levels=c("VD07.E1", "VD07.E2", "VD07.E3", "M23.E", "M23.G", "present"))
+
+bfs_plot_1prior <- ggplot(data=filter(df.bf, Prior2 == "N(0, 0.05)"),
+                                 aes(x=experiment, y=BF10)) +
+                      geom_line(aes(),color="grey", group=1, linewidth=1) +
+                    geom_point(aes(color=experiment, fill=experiment, shape=experiment), size=4) +
+                    scale_shape_manual(values = c(18,15,15,15,15,18))+
+                scale_fill_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
+                scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
+                geom_hline(yintercept=1, linetype="dashed")+
+                facet_grid(region~effect, scales="free", space="free") +
+                theme_bw(base_size = 15)+
+                scale_x_discrete(breaks = NULL)+
+                theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
+                xlab("")+
+              scale_y_log10("Bayes factor (BF10)",
+                             breaks =  c(0.05, 0.1, 0.5, 1, 3, 10, 50, 100, 200, 500, 75000),
+                             labels = c(0.05, 0.1, 0.5, 1,  3, 10, 50, 100, 200, 500, 75000)
+                            )+
+                guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
+                ggtitle("Bayes factors under prior N(0, 0.05)")
+ggsave("plots/present_previous_BFs_1Prior.jpg", width=9, height=8, dpi=600)
+
+ggplot(data=df.bf,aes(x=Prior2, y=BF10, group=experiment)) +
+      geom_line(aes(color=experiment)) +
+      geom_point(aes(color=experiment), size=2) +
+  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9], "black"))+
+  geom_hline(yintercept=1, linetype="dashed")+
+  facet_grid(region~effect, scales="free", space="free") +
+  theme_bw(base_size = 15)+
+  #scale_x_discrete(breaks = NULL)+
+  theme(legend.position="right", legend.box="vertical", legend.direction="vertical")+
+  xlab("")+
+scale_y_log10("Bayes factor (BF10)",
+               breaks =  c(0.05, 0.1, 0.5, 1, 3, 10, 50, 100, 200, 500, 75000),
+               labels = c(0.05, 0.1, 0.5, 1,  3, 10, 50, 100, 200, 500, 75000)
+              )+
+  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
+ggsave("plots/present_previous_BFs_allPriors.jpg", width=11, height=8, dpi=600)
+
+# combined plot
+plot_grid(RTs.plot, bfs_plot_1prior + theme(legend.position = "none"), rel_widths =c(1,0.6), nrow=1, labels=c("A", "B"))
+ggsave("plots/present_previous_allplots.jpg", width=16, height=10, dpi=600)
+
+```

### 2024-07-13 code_data/previous_studies/compute_BFs_PreviousStudies.R
Deleted via the Open Science Framework

@@ -1,624 +1,0 @@
-# author: Pia Schoknecht
-# date: 21.05.2024
-
-# script to compute Bayes Factors for the previous studies
-
-# Mertzen et al., (2024) English and German eye-tracking data
-# Van Dyke (2007) English eye-tracking and self-paced reading data
-# for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
-
-
-library(tidyverse)
-library(lme4)
-library(brms)
-library(ggplot2)
-library(bayesplot)
-library(truncnorm)
-library(ggbreak)
-library(ggimage) 
-library(cowplot)
-
-
-# load Mertzen et al. GERMAN data
-df <- read.table("dataRetroDE_processed.txt")
-
-df_crit <- df %>% filter(region == "crit") %>%
-                  select(subject, trial, list, item, condition, roi, region, TFT) %>%
-                  filter(TFT != 0)
-
-# contrast coding
-df_crit$syn <- ifelse(df_crit$condition%in%c("a","b"),-0.5,0.5)
-df_crit$sem <- ifelse(df_crit$condition%in%c("b","d"),0.5,-0.5)
-
-# trim all rts below 150 and above 3000 ms
-range(df_crit$TFT)
-crit_trim <- subset(df_crit, df_crit$TFT > 80)
-range(crit_trim$TFT)
-
-
-# Inferential statistics
-# truncated varying priors 
-# (only positive effects -- longer reading times for high interference)
-
-priors_s_tr <- c(
-  prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.01), class = b, lb=0),
-  prior(normal(0, 0.5), class = sigma),
-  prior(normal(0, 0.1), class = sd),
-  prior(lkj(2), class = cor)
-)  
-
-priors_m_tr <- c(
-  prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.05), class = b, lb=0),
-  prior(normal(0, 0.5), class = sigma),
-  prior(normal(0, 0.1), class = sd),
-  prior(lkj(2), class = cor)
-) 
-
-priors_l_tr <- c(
-  prior(normal(6, 0.6), class = Intercept),
-  prior(normal(0, 0.1), class = b, lb=0),
[…]

### 2024-07-13 code_data/previous_studies/compute_BFs.R
Added via the Open Science Framework

@@ -1,0 +1,79 @@
+# author: Pia Schoknecht
+# date: 09.07.2024
+
+# script to compute Bayes Factors with the Savage-Dickey method
+library(tidyverse)
+library(brms)
+library(ggplot2)
+library(bayesplot)
+
+
+load("model_fits_E/Fit_s_Full_precrit.Rda")
+load("model_fits_E/Fit_m_Full_precrit.Rda")
+load("model_fits_E/Fit_l_Full_precrit.Rda")
+
+load("model_fits_E/Fit_s_Full_crit.Rda")
+load("model_fits_E/Fit_m_Full_crit.Rda")
+load("model_fits_E/Fit_l_Full_crit.Rda")
+
+load("model_fits_E/Fit_s_Full_postcrit.Rda")
+load("model_fits_E/Fit_m_Full_postcrit.Rda")
+load("model_fits_E/Fit_l_Full_postcrit.Rda")
+
+#bfs 
+
+compute_bfs <- function(model){
+  h <- hypothesis(model, "syn = 0")
+  bf10.syn <- 1 / h$hypothesis$Evid.Ratio
+  
+  h <- hypothesis(model, "sem = 0")
+  bf10.sem <- 1 / h$hypothesis$Evid.Ratio
+  
+  h <- hypothesis(model, "syn:sem = 0")
+  bf10.int <- 1 / h$hypothesis$Evid.Ratio
+  
+  return(c(bf10.syn, bf10.sem, bf10.int))
+}
+
+# save BFs in df
+df.bf_m23.E <- data.frame(matrix(ncol=8,nrow=0))
+colnames(df.bf_m23.E) <- c("Experiment", "BF_method", "Language", "Region", "Effect","Prior","truncated", "BF10")
+
+df.bf_m23.E[nrow(df.bf_m23.E)+1,] <- c("M23", "S-D", "English", "pre-critical", "syntactic","Normal(0, 0.01)","no", round(compute_bfs(m_s_full_precrit)[1],2))
+df.bf_m23.E[nrow(df.bf_m23.E)+1,] <- c("M23", "S-D", "English", "pre-critical", "semantic","Normal(0, 0.01)","no", round(compute_bfs(m_s_full_precrit)[2],2))
[…]

### 2024-07-13 code_data/previous_studies/models_E.R
Added via the Open Science Framework

@@ -1,0 +1,190 @@
+# author: Pia Schoknecht
+# date: 21.05.2024
+
+# script to compute Bayes Factors for the previous studies
+
+# Mertzen et al., (2024) English and German eye-tracking data
+# Van Dyke (2007) English eye-tracking and self-paced reading data
+# for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
+
+
+#library(tidyverse)
+#library(lme4)
+library(brms)
+#library(ggplot2)
+#library(bayesplot)
+#library(truncnorm)
+#library(cowplot)
+
+
+# load Mertzen et al. ENGLISH data
+df <- read.table("dataRetroEN_processed.txt")
+df2 <- subset(df, df$region %in% c("precrit", "crit", "postcrit"))
+
+# contrast coding
+df2$syn <- ifelse(df2$condition%in%c("a","b"),-0.5,0.5)
+df2$sem <- ifelse(df2$condition%in%c("b","d"),0.5,-0.5)
+
+# trim all rts below 80
+range(df2$TFT)
+df2_trim <- subset(df2, df2$TFT > 80)
+range(df2_trim$TFT)
+
+
+# same priors as for the present study
+priors_s <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, coef="syn"),
+  prior(normal(0, 0.01), class = b, coef="sem"),
+  prior(normal(0, 0.01), class = b, coef="syn:sem"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_m <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, coef="syn"),
+  prior(normal(0, 0.05), class = b, coef="sem"),
+  prior(normal(0, 0.05), class = b, coef="syn:sem"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, coef="syn"),
+  prior(normal(0, 0.1), class = b, coef="sem"),
+  prior(normal(0, 0.1), class = b, coef="syn:sem"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+##### models ####
+##### small prior ####
[…]

### 2024-07-13 code_data/previous_studies/models_G.R
Added via the Open Science Framework

@@ -1,0 +1,651 @@
+# author: Pia Schoknecht
+# date: 21.05.2024
+
+# script to compute Bayes Factors for the previous studies
+
+# Mertzen et al., (2024) English and German eye-tracking data
+# Van Dyke (2007) English eye-tracking and self-paced reading data
+# for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
+
+
+#library(tidyverse)
+#library(lme4)
+library(brms)
+#library(ggplot2)
+#library(bayesplot)
+#library(truncnorm)
+#library(cowplot)
+
+
+#### load Mertzen et al. GERMAN data ####
+df <- read.table("dataRetroDE_processed.txt")
+df2 <- subset(df, df$region %in% c("precrit", "crit", "postcrit"))
+
+# contrast coding
+df2$syn <- ifelse(df2$condition%in%c("a","b"),-0.5,0.5)
+df2$sem <- ifelse(df2$condition%in%c("b","d"),0.5,-0.5)
+
+# trim all rts below 80
+range(df2$TFT)
+df2_trim <- subset(df2, df2$TFT > 80)
+range(df2_trim$TFT)
+
+
+# same priors as for the present study
+priors_s <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.01), class = b, coef="syn"),
+  prior(normal(0, 0.01), class = b, coef="sem"),
+  prior(normal(0, 0.01), class = b, coef="syn:sem"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_m <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.05), class = b, coef="syn"),
+  prior(normal(0, 0.05), class = b, coef="sem"),
+  prior(normal(0, 0.05), class = b, coef="syn:sem"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
+priors_l <- c(
+  prior(normal(6, 0.6), class = Intercept),
+  prior(normal(0, 0.1), class = b, coef="syn"),
+  prior(normal(0, 0.1), class = b, coef="sem"),
+  prior(normal(0, 0.1), class = b, coef="syn:sem"),
+  prior(normal(0, 0.5), class = sigma),
+  prior(normal(0, 0.1), class = sd)
+) 
+
[…]

### 2024-07-13 code_data/previous_studies/models_G.R
Updated via the Open Science Framework

@@ -1,20 +1,13 @@
 # author: Pia Schoknecht
 # date: 21.05.2024
 
-# script to compute Bayes Factors for the previous studies
+# script to run models (to later compute Bayes Factors) for the previous studies
 
 # Mertzen et al., (2024) English and German eye-tracking data
-# Van Dyke (2007) English eye-tracking and self-paced reading data
 # for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
 
 
-#library(tidyverse)
-#library(lme4)
 library(brms)
-#library(ggplot2)
-#library(bayesplot)
-#library(truncnorm)
-#library(cowplot)
 
 
 #### load Mertzen et al. GERMAN data ####
@@ -188,464 +181,3 @@ m_l_full_postcrit <- brm(TFT ~ 1 + syn * sem +
                          sample_prior="yes"
 )
 save(m_l_full_postcrit, file = paste("model_fits_G/Fit_l_Full_postcrit.Rda"))
-
-
-###########################################################################################
-
-
-# load Mertzen et al. ENGLISH data
-df <- read.table("dataRetroEN_processed.txt")
-
-df_crit <- df %>% filter(region == "crit") %>%
-  select(subject, trial, list, item, condition, roi, region, TFT) %>%
-  filter(TFT != 0)
-
-# contrast coding
-df_crit$syn <- ifelse(df_crit$condition%in%c("a","b"),-0.5,0.5)
-df_crit$sem <- ifelse(df_crit$condition%in%c("b","d"),0.5,-0.5)
-
-# trim all rts below 150 and above 3000 ms
-range(df_crit$TFT)
-crit_trim <- subset(df_crit, df_crit$TFT > 80)
-range(crit_trim$TFT)
-
-
-# models
-#directional priors
-m_s_full <- brm(TFT ~ 1 + syn * sem +
-                  (1 + syn * sem| subject) +
-                  (1 + syn * sem| item),
-                data = crit_trim,
-                family = lognormal(),
-                prior = priors_s_tr,
-                warmup = 2000,
-                iter = 20000,
-                cores = 4,
-                save_pars = save_pars(all = TRUE),
[…]

### 2024-07-13 code_data/previous_studies/models_E.R
Updated via the Open Science Framework

@@ -1,21 +1,13 @@
 # author: Pia Schoknecht
 # date: 21.05.2024
 
-# script to compute Bayes Factors for the previous studies
+# script to run models (to later compute Bayes Factors) for the previous studies
 
 # Mertzen et al., (2024) English and German eye-tracking data
-# Van Dyke (2007) English eye-tracking and self-paced reading data
 # for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
 
 
-#library(tidyverse)
-#library(lme4)
 library(brms)
-#library(ggplot2)
-#library(bayesplot)
-#library(truncnorm)
-#library(cowplot)
-
 
 # load Mertzen et al. ENGLISH data
 df <- read.table("dataRetroEN_processed.txt")

### 2024-07-13 code_data/previous_studies/PreviousFindings.Rmd
Updated via the Open Science Framework

@@
 library(cowplot)
-library(ggbreak)
-
 ```
@@
 # Van Dyke (2007) estimates from published paper:
-
 ## E1 
-
 ### critical
@@
 ### spill-over
-
 ```{r}
 ## E1 spill-over
-
 ## not sig. Syn effect in E1 spill-over:
@@
 ### Sentence-Final
-
 ```{r}
@@
 ## Experiment 2
-
 ### critical: SYN 
-
 ```{r}
 # Experiment 2 SYN in critical
-
 ## sig. Syn effect in E2 crit region FPRT:
@@
 ### critical: SEM
-
 ```{r}
 # Experiment 2 Sem in critical
-
 ## non sig. Sem effect in E2 crit region FPRT:
@@
 ### critical: Int
-
 ```{r}
 # Experiment 2 Int in critical
-
 ## non sig. Int effect in E2 crit region FPRT:
@@
 ### spill-over: SYN 
-
 ```{r}
 # Experiment 2 SYN in spill-over
-
 ## almost sig. Syn effect in E2 post-crit region FPRT (negative sign)
@@
 ### spill-over: Sem 
-
 ```{r}
 # Experiment 2 Sem in spill-over
-
 ## almost sig. Sem effect in E2 post-crit region FPRT (negative sign)
@@
 ### spill-over: Int 
-
 ```{r}
 # Experiment 2 Int in spill-over
-
 ## almost sig. Int effect in E2 post-crit region FPRT (negative sign)
@@
 ### Sentence-Final: SYN 
-
 ```{r}
 # Experiment 2 SYN in sentence-final region (Table 7)
-
 ## Syn effect in E2 sentence-final region FPRT 
@@
 ### sentence-final: Sem 
-
 ```{r}
 # Experiment 2 Sem in sentence-final (Table 7)
-
 ## Sem effect in E2 sentence-final region FPRT
@@
 ### sentence-final: Int 
-
 ```{r}
 # Experiment 2 Int in sentence-final (Table 7)
-
 ## Int effect in E2 sentence-final region FPRT
@@
 ### precrit: SYN
-
 ```{r}
 # Experiment 3  SYN in pre-crit
-
 ## almost sig. Syn effect in E3 precrit region FPRT:
@@
 ### precrit: SEM
-
 ```{r}
 # Experiment 3  SEM in pre-crit
-
 ## almost sig. Sem effect in E3 precrit region FPRT:
@@
 ### precrit: INT
-
 ```{r}
 # Experiment 3  Int in pre-crit
-
 ## almost sig. Int effect in E3 precrit region FPRT:
@@
 VD07E3critSynTFTupper <- round(VD07E3critSynTFTmn+2*VD07E3critSynTFTSE)
-
 ```
@@
 # Experiment 3  SEM in crit
-
 ## almost sig. Sem effect in E3 crit region FPRT:
@@
 # Experiment 3  Int in crit
-
 ## almost sig. Int effect in E3 crit region FPRT:
@@
 ### spill-over: SYN
-
 ```{r}
 # Experiment 3  SYN in postcrit
-
 ## almost sig. Syn effect in E3 postcrit region FPRT:
@@
 VD07E3postcritSynTFTupper <- round(VD07E3postcritSynTFTmn+2*VD07E3postcritSynTFTSE)
-
 ```
@@
 # Experiment 3  SEM in postcrit
-
 ## almost sig. Sem effect in E3 postcrit region FPRT:
@@
 # Experiment 3  Int in postcrit
-
 ## almost sig. Int effect in E3 postcrit region FPRT:
@@
 ### sentence final: SYN
-
 ```{r}
 # Experiment 3  SYN in sentence final region (Table 16 in VD07)
-
 ## almost sig. Syn effect in E3 final region FPRT:
@@
 VD07E3finalSynTFTupper <- round(VD07E3finalSynTFTmn+2*VD07E3finalSynTFTSE)
-
 ```
@@
 # Experiment 3  Int in final (Table 16 in VD07)
-
 ## Int effect in E3 final region FPRT:
@@
 VD07estimates$int_type <- factor(VD07estimates$int_type,levels=c("syntactic","semantic", "interaction"))
-
-str(VD07estimates)
 ```
 
-# Mertzen et al 2021:
-
+# Mertzen et al 2023: 
 ```{r}
 load("M21_byregionsummaries_de.rda")
-
 M23.G <- d
@@
 
-M23estimates <- M23estimates[,-4]
-M23estimates <- M23estimates[,-5]
-M23estimates <- M23estimates[,-6]
-
-```
-
-# Plot previous findings
-```{r}
-my_colors <- RColorBrewer::brewer.pal(11, "RdBu")
-
-previous <- rbind(VD07estimates, M23estimates)
-save(previous, file="VD07_M21_estimates.rda")
-
-previous_crit.plot <- ggplot(subset(previous, previous$region%in%c("pre-critical","critical")),
-                   aes(x=experiment, y=mean, ymin=lower, ymax=upper )) +
-  geom_pointrange(aes(color=experiment, shape=measure), 
-             size=0.5, position=position_dodge(0.9))+
-  scale_shape_manual(values = c(18,16,17,15))+
-  scale_color_manual(values = c(my_colors[3], my_colors[4],my_colors[5],my_colors[8],my_colors[9]))+
-  geom_hline(yintercept=0, linetype="solid", alpha=.4)+
-  ggh4x::facet_grid2(region ~ int_type, scales="free_y", independent = "y") +
-  theme_bw(base_size = 10)+
-  scale_x_discrete(breaks = NULL)+
-  xlab("")+
-  ylab("reading time difference (ms)")+
-  #ggtitle("Previous findings")+
-  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
-
-previous_crit.plot
-ggsave("plots/previous_findings.jpg", width=6, height=5, dpi=600)
+M23estimates <- M23estimates %>% select(-lang, -region.label, -effect)
 ```
@@
 # Plot present and previous findings
-```{r, eval=FALSE}
+```{r}
 my_colors <- RColorBrewer::brewer.pal(11, "RdBu")
@@
 # combine with previous estimates
-estimates <- rbind(previous, SVestimates)
+estimates <- rbind(VD07estimates, M23estimates, SVestimates)
 estimates$experiment <- factor(estimates$experiment, levels=c("VD07.E1 (N=35)", "VD07.E2 (N=36)", "VD07.E3 (N=40)", "M23.E (N=61)", "M23.G (N=121)", "present (N=774)"))
 
-pp.plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over")),
+RT_plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over")),
                        aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
@@
                   #ggtitle("Present and previous findings")+
-  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
-
-RTs.plot <- pp.plot + coord_cartesian(ylim=c(-75, 150)) + ggtitle("Reading time differences") 
+                  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
+                  coord_cartesian(ylim=c(-75, 150)) +
+                  ggtitle("Reading time differences") 
 ggsave("plots/present_previous_findings.jpg", width=9, height=8, dpi=600)
 
-pp2.plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over") & measure %in%c("SPR", "TFT")),
-                   aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
+TFT_plot <- ggplot(data=filter(estimates, region %in% c("pre-critical", "critical", "spill-over") & measure %in%c("SPR", "TFT")), aes(x=int_type, y=mean, ymin=lower, ymax=upper )) +
                   geom_pointrange(aes(color=experiment, shape=measure), 
@@
                 #ggtitle("Present and previous findings")+
-  guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))
-
-pp2.plot + coord_cartesian(ylim=c(-50, 100))  
+                guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
+                coord_cartesian(ylim=c(-50, 100))  
 ggsave("plots/present_previous_findings_SPR_TFT.jpg", width=9, height=8, dpi=600)
@@
 # combined plot
-plot_grid(RTs.plot, bfs_plot_1prior + theme(legend.position = "none"), rel_widths =c(1,0.6), nrow=1, labels=c("A", "B"))
-ggsave("plots/present_previous_allplots.jpg", width=16, height=10, dpi=600)
+plot_grid(RT_plot, bfs_plot_1prior + theme(legend.position = "none"), rel_widths =c(1,0.6), nrow=1, labels=c("A", "B"))
+ggsave("plots/present_previous_allplots.jpg", width=15, height=8, dpi=600)
 

### 2024-07-15 paper/main.tex
Updated via the Open Science Framework

@@
 The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
-Varying intercepts for participants and items were included; varying slopes were excluded for reasons explained below.
-
-\paragraph{Bayes factors for model comparison}
+\textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task.
+The trial id of the 100 trials per session was recoded to span from 0 to 1.
+This was done to bring all fixed effects to a comparable scale which decreases the run time of the models.
+The models were run with full random effects (varying intercepts and slopes of all fixed effects and their interactions for participants and items).}
+
+\paragraph{Bayes factors for model comparison}\label{BF_analysis_SPR}
 
@@
 we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}).
-The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms.
-These priors are based on the observed sign and sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}, and on meta-analyses relating to interference effects \parencite{jaeger_etal_2017}.
-For the syntactic and semantic interference effects, we use directional priors so that any evidence for the effect of interest from Bayes factors can be interpreted as evidence of replication success \parencite{verhagen2014bayesian}.
+The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, \textcolor{blue}{ranging from -8 to 8 ms, -40 to 40 ms, or -81 to 81 ms.
+These priors are based on the observed sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}, and on meta-analyses relating to interference effects \parencite{jaeger_etal_2017}.}
 
@@
     \caption{Priors used for the analysis of the self-paced reading time data.
-Priors on the fixed-effects slope parameters were truncated to a lower bound of zero, i.e., only positive values were assumed to be possible a priori.
-This corresponds to the existing evidence in the literature that high interference conditions lead to longer reading times than low interference conditions.
 The standard deviation of the fixed-effects slope priors was varied in order to conduct a Bayes factor sensitivity analysis following \citet{schad_etal_2022_BF}.
@@
     \centering
-     \begin{tabular}{lllr}
+     \begin{tabular}{llr}
     \toprule
-    Parameter&Prior & &Assumed Range in ms\\
+    Parameter&Prior &Assumed Range in ms\\
     \midrule
-  Intercept & Normal(6, 0.6)& & [125, 1308]\\
-  \cmidrule{2-4}
-  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \hspace{-1em}\rdelim\}{3}{*}[lower bound: 0] & [0, 8]\\
-  &  Normal(0, 0.05)& & [0, 40]\\
-  & Normal(0, 0.1) & &[ 0, 81]\\
-  \cmidrule{2-4}
-  sigma & Normal(0, 0.5)&  \hspace{-1em}\rdelim\}{2}{*}[lower bound: 0]&\\
-  SD & Normal(0, 0.1)& &\\
+  Intercept & Normal(6, 0.6)& [125, 1308]\\
+  \cmidrule{2-3}
+  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \textcolor{blue}{[-8, 8]}\\
+  &  Normal(0, 0.05)&  \textcolor{blue}{[-40, 40]}\\
+  & Normal(0, 0.1) & \textcolor{blue}{[-81, 81]}\\
+  \cmidrule{2-3}
+  sigma & Normal(0, 0.5)&\\
+  SD & Normal(0, 0.1) &\\
   \bottomrule
@@
 Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
-For syntactic interference, they were close to zero in all regions.
-In contrast, for semantic interference, the estimates ranged between [8,14] ms (pre-critical), [4, 12] ms (critical), and [5, 11] ms (spill-over region).
-The posteriors for the interaction were consistent with an interaction effect.
-The interaction was the  largest in the critical region ([7, 22 ms]).
+\textcolor{blue}{The 95 \% credible intervals for syntactic interference included zero in the pre-critical and critical regions.
+In the spill-over region, it included only positive values ([1, 15] ms).
+In contrast, for semantic interference, the 95 \% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms).
+The posteriors for the interaction were mostly positive, but the 95 \% credible intervals included zero in all regions.
+The interaction was the  largest in the critical region ([-3, 31 ms]).}
 
@@
 The analyses use all the data from Experiments 1a and 1b.
-We found extremely strong evidence for the semantic interference effect in the pre-critical, critical and spill-over regions (BF$_{10}$ > 200).
-No evidence was found for syntactic interference in any of the investigated regions; in fact, there is strong evidence against the syntactic interference effect in all regions and under all the priors used (BF$_{10}$ < 1).
-Evidence for the interaction was only found in the critical region.
-Here, the evidence for the interaction was strong (BF$_{10}$ = 40) when the prior on the parameter was N+(0, 0.01), i.e., a small a priori effect size ranging from  [0, 8] ms; the evidence for the interaction was much weaker (BF$_{10}$ = 1.8) when the prior was N+(0,0.1), i.e., a larger a priori effect size ranging from [0, 81] ms.
-In other words, only if we assume a priori that the effect size for the interaction is relatively small, then there is strong evidence for this effect.
+\textcolor{blue}{We found very to extremely strong evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
+In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5).
+There was strong evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ < 0.5).
+Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was weak evidence for syntactic interference (BF$_{10}$ = 2.8).
+Similarly, there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ < 1).
+Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}
+
+\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but inconclusive results in the critical region.
+In contrast, the Bayes factor analysis provided either at best weak evidence for syntactic interference and the interaction or even evidence against them.}
 
@@
 We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-Reading times at the critical verb showed clear evidence for semantic interference, but evidence against syntactic interference.
-The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occurred even before the retrieval event could begin.
-The effects in the pre-critical region are discussed next.
+\textcolor{blue}{Reading times at the critical verb were inconclusive for semantic interference and showed evidence against syntactic interference.
+We will discuss the absence of syntactic interference in our data in the general discussion.
+The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occurred even before the retrieval event could begin.} The effects in the pre-critical region are discussed next.
 
@@
 It seems unlikely that this small difference of one letter in word length caused such large and long-lasting effects, but of course one cannot rule out this possibility with complete certainty.
+
+\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
+To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}.
+Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}).
+Forty-four native speakers of German (mean age XX, age range: , w, m), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible).
+The experiment included 40 implausible filler sentences which were created by exchanging nouns between the fillers used in the SPR experiment (e.g., Der Sohn des Kaisers, der den Feldzug gewonnen hatte, plante äußerst strategisch, weil er die Gießkanne an sich reißen wollte., `The son of the emperor who had won the campaign planned extremely strategically because he wanted to take the watering can by force.').
+Participants spent approximately 37 minutes to complete the task and received 6.50 \pounds \, as compensation.
+The results of an ordinal brms model analyzing the ratings are shown in Figure X.
+Sentences with an inanimate distractor had a slightly higher probability to gain a high rating on the plausibility scale ($\beta$ = XX), i.e., the low semantic interference conditions are more plausible.
+To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran an additional model which used the same specifications as described in the Statistical Analyses section with prior N(0,0.5) and additionally, the centered plausibility ratings as a fixed effect.
+Figure X presents the posteriors from this model.
+}
 
@@
 Because of the unclear status of the effects at the critical verb, we refrain from interpreting the interaction which was only observed in this region.
-
-\subsubsection{Comparison with previous findings}
-The results of the present study stand in contrast to previous findings \citep{vandyke07, mertzen}.
-The present study is the only one to find semantic and no syntactic interference effects.
-The difference in observed effects could be explained by differences in statistical power or by methodological differences.
-The present study has higher statistical power than the previous studies.
-\citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively.
-\textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
-In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
-Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design.
-Consequently, the estimates from the present study are the most precise ones so far for this design.
-The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
-
-An obvious difference from previous studies is the experimental methodology.
-The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task.
-Self-paced reading (SPR) might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}.
-SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design:
-In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
-However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.
-
-The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even the same linguistic material.
-\Copy{items_mertzen}{\textcolor{blue}{However, \citet{mertzen} changed the structure of the relative clause modifying the subject by adding an additional animate distractor to all conditions (see an example of one of their English items in \ref{ex:mertzen}; their items had this structure cross-linguistically).
-This was done to ``increase the strength of the manipulation.'' \citep[][p.
-9]{mertzen}.
-The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
-But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
-While this additional distractor might have increased syntactic interference compared to our study (and \citeauthor{vandyke07}'s study as well actually), it is not obvious why it would decrease semantic interference compared to our study.}\label{items_mertzen}
-
-\begin{exe}[ht]
-\ex \label{ex:mertzen} 
-\textcolor{blue}{It turned out that the attorney whose secretary had forgotten that the visitor was important frequently complained about the salary at the firm. \parencite{mertzen}}
-\end{exe}
-}
-
-So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
-Apart from \textcolor{blue}{the additional distractor and} the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
-While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
-We will revisit the absence of syntactic interference in our data in the general discussion.
 
@@
     \caption{Relatively informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}.
-Priors on the slope parameters were truncated to an upper bound of zero, i.e., only negative values.
-This corresponds to more negative brain responses to high vs.\ low interference conditions in line with the literature \citep{lee_garnsey, martinetal2014, schoknecht2022, vasishth_drenhaus_2011}.
-The directional priors also serve to allow us to quantify the extent to which a predicted effect replicates; the Bayes factors can be interpreted as indicating the evidence supporting or not supporting replication success \citep{verhagen2014bayesian}.
 The standard deviations of the slope priors were varied in order to conduct a Bayes factor sensitivity analysis.
@@
     \centering
-    \begin{tabular}{lllr}
+    \begin{tabular}{llr}
     \toprule
-    Parameter&Prior & &Assumed Range ($\mu V$)\\
+    Parameter&Prior &Assumed Range ($\mu V$)\\
     \midrule
-  Intercept & Normal(2, 5)& & [-8, 12]\\
-  \cmidrule{2-4}
-  \multirow{4}{1cm}{slope} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{4}{*}[upper bound: 0]& [-0.2, 0]\\
-  & Normal(0, 0.5)& & [-1, 0]\\
-  & Normal(0, 1) && [-2, 0]\\
-  & Normal(0, 5) && [-11, 0]\\
-  \cmidrule{2-4}
-  sigma & Normal(10, 5)& \hspace{-1em}\rdelim\}{2}{*}[lower bound: 0]\\
+  Intercept & Normal(2, 5)& [-8, 12]\\
+  \cmidrule{2-3}
+  \multirow{4}{1cm}{slope} & Normal(0, 0.1) & [-0.2, 0]\\
+  & Normal(0, 0.5)& [-1, 1]\\
+  & Normal(0, 1) & [-2, 2]\\
+  & Normal(0, 2) & [-4, 4]\\
+  \cmidrule{2-3}
+  sigma & Normal(10, 5)&\\
   SD & Normal(0, 2)&\\
- \textcolor{blue}{correlations} & \textcolor{blue}{LKJcorr(2)} & & \\
     \bottomrule
@@
 For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats,schad_etal_2022_BF}.
-We chose slope priors which assume a wide range of effect sizes from $-0.2$ to $-11$\,$\mu V$.
-However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$.
-This assessment is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \textcite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.
+We chose slope priors which assume a wide range of effect sizes from [$-0.2$, 0.2] to [$-4$, 4]\,$\mu V$.
 
@@
 grand average ERPs elicited by the critical verb for all conditions.
-Table \ref{tab:eeg_mod} shows the estimates for the brain activity in the spatio-temporal N400 window from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu V$.
-The 95\% credible intervals for the fixed effects are consistent with effects of syntactic interference, semantic interference and their interaction.
-The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (\textcolor{blue}{mean: -0.32\,$\mu V$, 95\% CrI = [-0.59, -0.10]\,$\mu V$}, for the posteriors, see Figure \ref{fig:eeg_posteriors}).
-The estimates of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).
+\textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as -1\,$\mu V$.
+The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600.
+The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals include zero for both the N400 and the P600.
+The 95\% credible interval of the interaction was centered around zero for the N400.
+For the P600, it included mostly negative values but crossed zero.}
 
@@
 
-\begin{table}[H]
-    \caption{Results from the Bayesian linear mixed model analyzing the averaged brain response from 12 centro-parietal electrodes in the time window 300 to 500 ms post critical verb onset.
-The predictor baseline represents the averaged activity from the 200 ms pre-verb onset and was included in the model as a covariate instead of traditional baseline correction \citep[see][]{alday2019}.}
-    \label{tab:eeg_mod}
-    \centering
-    \begin{tabular}{lrrr}
-    \toprule
-    & Estimate & Est.Error & 95\% CrI  \\
-    \midrule
-Intercept &  -0.70 & 0.19 &  [-1.06, -0.34] \\
-baseline  &  -0.22 & 0.01 &  [-0.24, -0.20] \\
-syntactic &  \textcolor{blue}{-0.16}& \textcolor{blue}{0.08} &  \textcolor{blue}{[-0.33, -0.02]} \\
-semantic  &  \textcolor{blue}{-0.30} & \textcolor{blue}{0.10} &  \textcolor{blue}{[-0.50, -0.11]} \\
-interaction& \textcolor{blue}{-0.14} & \textcolor{blue}{0.11} &  \textcolor{blue}{[-0.40, -0.01]} \\
-    \bottomrule
-    \end{tabular}
-\end{table}
-
 \begin{figure}[H]
-    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal N400 window.
+    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal \textcolor{blue}{windows of the N400 (A) and P600 (B)}.
 Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{images/posteriors_eeg.jpg}
+    \includegraphics[width=\textwidth]{images/posteriors_eeg.jpg}
 \end{figure}
@@
 \section{Quantitative predictions of the Lewis and Vasishth (2005) model for the present design}
+
+% root mean square deviation
 
@@
 
-In the remainder of this section, we discuss two important issues that relate to previous experiments that investigated syntactic and semantic interference.
+\textcolor{red}{Rewrite this section to reflect the new discussion.} In the remainder of this section, we discuss two important issues that relate to previous experiments that investigated syntactic and semantic interference.
 The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.
+
+\subsection{Comparison with previous findings}
+\textcolor{blue}{The present study used the same design as two previous studies \parencite{vandyke07, mertzen}.
+Figure \ref{fig:previous_vs_present} A shows the estimated reading times differences from the previous studies and our self-paced reading experiment.
+Because the \textcite{vandyke07} data was not available to us, we derived estimates for their study from the summary statistics in the published paper.
+\textcite{mertzen} made their data available.
+In comparison to the previous estimates, the effects of the present study are relatively small and they are associated with less uncertainty.
+To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies.
+For \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE).
+Then using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors using the Savage-Dickey method.
+Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from models which we run on their data along to our own analyses described in Statistical Analyses for Experiments 1a and b.
+The Bayes factors of the effects of interest in self-paced reading times and total fixation times of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}
+
+\begin{sidewaysfigure}[h]
+    \caption{A) Reading time differences in the regions of interest from the previous and present studies using the design by \textcite{vandyke07}.
+VD07.E1, VD07.E2 and VD07.E3 stand for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiments 1-3.
+M23.E and M23.G stand for \citeauthor{mertzen}'s (\citeyear{mertzen}) English and German experiment, respectively.
+The intervals for their study and the present study are Bayesian 95\% credible intervals.
+There are less estimates for the pre-critical region because \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1 and 2 did not have a pre-critical region.
+B) Bayes factors for the effects of interest in for the present and previous studies under prior Normal(0,0.5).}
+    \label{fig:previous_vs_present}
+    \centering
+    \includegraphics[width=0.9\linewidth]{images/present_previous_allplots.jpg}
+\end{sidewaysfigure}
+\clearpage
+
+\textcolor{blue}{The Bayes factors for the syntactic interference effect show that the present study is the only one to provide evidence against the effect in the pre-critical region.
+In the critical region, both self-paced reading experiments provide evidence against the effect.
+The Bayes factors of three of the eye-tracking studies show weak or no evidence for the effect.
+The Bayes factor of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provides extremely strong evidence for the syntactic interference effect in the critical region.
+In the spill-over region, the Bayes factors for syntactic interference rather consistently provide evidence against the effect.
+Only \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment provides moderate evidence for it.}
+
+\textcolor{blue}{The Bayes factors for the semantic interference effect consistently provide evidence for the effect in the pre-critical region.
+In the critical region, most experiments provide evidence against semantic interference, the two self-paced reading experiments, being the exceptions, provide evidence for the effect.
+In the spill-over region, the Bayes factor results for semantic interference are mixed: the two self-paced reading experiments provide extremely strong evidence for the effect, the total fixation times for two of the eye-tracking experiments provide weak evidence for the effect and the total fixation times of two other eye-tracking experiments provide evidence against the effect.
+Importantly, the clustering in the total fixation time results does not align with the authors of the studies.}
+
+\textcolor{blue}{The Bayes factors for the interaction in all regions are rather consistent between studies and are either inconclusive or provide evidence against the interaction.
+The only exception is the very large Bayes factor of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1.
+In sum, while the present and previous studies show comparable region-wise results for semantic interference and the interaction, they differ remarkably regarding syntactic interference in the pre-critical and critical region.
+In the pre-critical region, the present study is the obvious outlier, being the only study to provide evidence against the effect.
+In the critical region, four of the investigated experiments show weak evidence for or against syntactic interference with two outliers: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 providing extreme evidence for the effect and the present study providing strong evidence against it.
+In conclusion, the most remarkable difference between the present study and previous findings concerns the syntactic interference in the pre-critical region.}
+
+%The difference in observed effects \textcolor{blue}{between our SPR experiment and the other reading time studies} could be explained by differences in statistical power or by methodological differences.
+
+The present study has higher statistical power than the previous studies.
+\citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively.
+\textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
+In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
+Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design.
+Consequently, the estimates from the present study are the most precise ones so far for this design.\textcolor{blue}{This is reflected by the tighter estimates in Figure \ref{fig:previous_vs_present}}.
+The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
+
+An obvious difference from previous studies is the experimental methodology.
+The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task.
+Self-paced reading (SPR) might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}.
+SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
+This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design:
+In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
+However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.
+
+\Copy{syn_confounds}{\textcolor{blue}{The difference in methodology lead to another difference between our study and the previous studies.
+In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line.
+Whether \citeauthor{vandyke07}'s \citeyear{vandyke07} SPR experiment used line breaks or not is not reported.
+In our SPR experiments, it was necessary to introduce line breaks into the sentences.
+These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb.
+The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor.
+It has been proposed by \textcite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability.
+Therefore, increased prominence of the subject could have decreased interference.
+Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference.
+For this point to hold, one would need to assume that the line break provided additional structural information by emphasizing the syntactic structure of the items in our SPR experiment.
+However, the EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor.
+So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
+Nevertheless, neither our SPR nor our EEG experiments showed evidence for syntactic interference.}
+
+\textcolor{blue}{All our experiments (SPR and EEG) displayed the subject with a comma.
+Commas are mandatory in German to separate the subordinate clauses from the main clause.
+Similar to the line break discussed above, the comma could have emphasized the syntactic structure of the sentences and therefore decreased syntactic interference from the distractor which occurred inside a relative clause and not in the clause of the critical verb.
+But if this comma provided information about the syntactic structure of the items then it should have so as well in \citeauthor{mertzen}'s \citeyear{mertzen} German study because they also presented the subject with the comma.
+So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.}}\label{syn_confounds}
+
+The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even the same linguistic material.
+\Copy{items_mertzen}{\textcolor{blue}{However, \citet{mertzen} changed the structure of the relative clause modifying the subject by adding an additional animate distractor to all conditions (see an example of one of their English items in \ref{ex:mertzen}; their items had this structure cross-linguistically).
+This was done to ``increase the strength of the manipulation.'' \citep[][p.
+9]{mertzen}.
+The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
+But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
+While this additional distractor might have increased syntactic interference compared to our study (and \citeauthor{vandyke07}'s study as well actually), it is not obvious why it would decrease semantic interference compared to our study.}\label{items_mertzen}
+
+\begin{exe}[ht]
+\ex \label{ex:mertzen} 
+\textcolor{blue}{It turned out that the attorney whose secretary had forgotten that the visitor was important frequently complained about the salary at the firm. \parencite{mertzen}}
+\end{exe}
+}
+
+So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
+Apart from \textcolor{blue}{the additional distractor and} the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
+While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
 

### 2024-07-15 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-Because the Van Dyke (2007) data is not available to us, we have used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) reported in the Van Dyke (2007) paper.
+Because the Van Dyke (2007) data was not available to us, we have used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) reported in the Van Dyke (2007) paper.
 Then using these posteriors and a reasonable prior (a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors using the Savage-Dickey method.
@@
 \subsubsection*{Response to R2.1}
-We used a well-studied design and the previous studies using it, specifically the same syntactic retrieval cue, found effects of syntactic interference \citep{vandyke07, mertzen}.
-Therefore, while our results might not be surprising for some researchers because the design might be flawed, others might be convinced that the design is very appropriate to investigate syntactic interference, because it produced such effects in the past.
+We used a well-studied design and the previous studies using the design, specifically the same syntactic retrieval cue, found effects of syntactic interference \citep{vandyke07, mertzen}.
+Therefore, while our results are not surprising for some researchers because the syntactic manipulation might be flawed, others might be convinced that the design is very appropriate to investigate syntactic interference, because it produced such effects in the past.
 We believe that our results add a valuable contribution to the literature on syntactic interference as a counterpoint to the previous studies.
@@
 \subsubsection*{Response to R2.2}
-\textcolor{red}{We agree with this statement and therefore refrained from interpreting the results at the critical region.\\
-Maybe add sth to manuscript}
+We agree with this statement and therefore refrained from interpreting the results at the critical region.\\
+\textcolor{red}{Maybe add sth to manuscript}
 
@@
 \end{quote}
+
+\textcolor{red}{add references}
 
@@
 \subsubsection*{Response to R3.4}
-We think that we addressed these concerns in our response to Reviewer \#1 in R1.5.
+We think that we have addressed the concerns regarding high- vs.\ lower-powered studies in our Response to R1.5.
+
+We have created a new subsection in our General Discussion providing an extended discussion of the potential reasons for the differences between our and previous results.
+We provide the relevant new text passages in our responses below (Response to R3.7 and R3.8).
 
@@
 \subsubsection*{Response to R3.5}
-To account for practice effects, we added the centered trial id as a fixed effect as well as by-participant and by-item random slopes to the models analyzing the SPR data.
+To account for practice effects, we have added the centered trial id as a fixed effect as well as by-participant and by-item random slopes to the models analyzing the SPR data.
 \textcolor{red}{The results did not change considerably.}
@@
 \subsubsection*{Response to R3.7}
-The difference in methodology lead to another difference between our study and the previous studies.
-In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line.
-Whether \citeauthor{vandyke07}'s \citeyear{vandyke07} SPR experiment used line breaks or not is not reported.
-In our SPR experiments, it was necessary to introduce line breaks into the sentences.
-These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb.
-The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor.
-It has been proposed by \textcite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability.
-Therefore, increased prominence of the subject could have decreased interference.
-Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference.
-For this point to hold, one would need to assume that the line break provided additional structural information by emphasizing the syntactic structure of the items in our SPR experiment.
-However, the EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor.
-So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-Nevertheless, neither our SPR nor our EEG experiments showed evidence for syntactic interference.
+We have added the following text addressing this concern to page \pageref{syn_confounds} of the manuscript:
 
-All our experiments (SPR and EEG) displayed the subject with a comma.
-Commas are mandatory in German to separate the subordinate clauses from the main clause.
-Similar to the line break discussed above, the comma could have emphasized the syntactic structure of the sentences and therefore decreased syntactic interference from the distractor which occurred inside a relative clause and not in the clause of the critical verb.
-But if this comma provided information about the syntactic structure of the items then it should have so as well in \citeauthor{mertzen}'s \citeyear{mertzen} German study because they also presented the subject with the comma.
-So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.
+\Paste{syn_confounds}
 

### 2024-07-24 code_data/plausibility_norming/analysis_ordinal.R
Added via the Open Science Framework

@@ -1,0 +1,181 @@
+# author: Pia Schoknecht
+# date: 17.07.24
+
+# load packages
+library(ggplot2)
+library(tidyverse)
+library(ordinal)
+library(brms)
+
+# load data
+df <- read.csv("norming_data_HighvsLowSem.csv")
+
+# number of participants
+length(unique(df$subject))
+
+# number of items
+length(unique(df$item))
+
+# demographics
+uniq <- df[!duplicated(df[2]),]
+table(uniq$Sex)
+range(uniq$Age)
+round(mean(uniq$Age))
+ 
+# did the latin square design work? -yes
+xtabs(~ condition + item, data=filter(df, latin_square_list == 0))
+xtabs(~ condition + item, data=filter(df, latin_square_list == 1))
+ 
+# factorize
+df$condition <- factor(df$condition, levels = c("c", "d", "f"))
+#df$value <- factor(df$value, levels = c("1", "2", "3", "4", "5", "6", "7"))
+ 
+# overview table
+n <- df %>% group_by(condition, value) %>% count()
+pivot_wider(n, values_from = "n", names_from = "value")
+
+# filler ratings 
+rating_byfilleritem <- aggregate(value ~ item, FUN=mean, data= filter(df, condition == "f"))
+round(range(rating_byfilleritem$value),1)
+ 
+# exclude fillers 
+df_crit<-subset(df,condition!="f")
+ 
+# contrast coding 
+# keep the default dummy coding. We compare conditions c and d
+# sum contrast d is high semantic interference, c is low semantic interference
+df_crit$sem <- ifelse(df_crit$condition == "d", 0.5, -0.5)
+df_crit$sem <- as.factor(df_crit$sem)
+ 
+# priors
+priors <- c(prior(normal(0, 3), class = Intercept),
+             prior(normal(0, 1), class = b),
+             prior(normal(0, 2), class = sd))
+ 
+# check priors
+extraDistr::qtnorm(0.025,mean=0, sd=3); extraDistr::qtnorm(0.975,mean=0, sd=3)# intercept normal(0, 3) upper and lower
+extraDistr::qtnorm(0.025,mean=0, sd=1); extraDistr::qtnorm(0.975,mean=0, sd=1) # normal(0, 1) upper and lower
+extraDistr::qtnorm(0.025,mean=0, sd=2); extraDistr::qtnorm(0.975,mean=0, sd=2) # normal(0, 2) upper and lower
+ 
+# cumulative model
+m <- brm(value ~ 1 + sem +
[…]

### 2024-07-24 code_data/plausibility_norming/analysis_ordinal.R
Updated via the Open Science Framework

@@ -6,6 +6,7 @@ library(ggplot2)
 library(tidyverse)
 library(ordinal)
 library(brms)
+library(bayesplot)
 
 # load data
 df <- read.csv("norming_data_HighvsLowSem.csv")
@@ -165,6 +166,43 @@ pp_check(m_m_full, ndraws = 100, type = "stat", stat = "max")
  
 summary(m_m_full)
 plot(m_m_full)
+
+# plot posteriors
+
+# extract samples
+samples_precrit <- as_draws_df(m_m_full)
+
+# backtransform estimates to ms scale
+sem_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$b_sem) - exp(samples_precrit$b_Intercept - samples_precrit$b_sem)
+rating_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$b_c_rating) - exp(samples_precrit$b_Intercept - samples_precrit$b_c_rating)
+int_eff_precrit <- exp(samples_precrit$b_Intercept + samples_precrit$`b_sem:c_rating`) - exp(samples_precrit$b_Intercept - samples_precrit$`b_sem:c_rating`)
+
+# credible intervals
+round(c(mean = mean(sem_eff_precrit), quantile(sem_eff_precrit, probs = c(.025, .975))))
+round(c(mean = mean(rating_eff_precrit), quantile(rating_eff_precrit, probs = c(.025, .975))))
+round(c(mean = mean(int_eff_precrit), quantile(int_eff_precrit, probs = c(.025, .975))))
+
+# add back-transformed values to samples df
+samples_precrit$semantic <- sem_eff_precrit
+samples_precrit$plausibility <- rating_eff_precrit
+samples_precrit$interaction <- int_eff_precrit
+
+# Plot
+posteriors_precrit <- mcmc_areas(
+  samples_precrit,
+  pars = c("semantic", "plausibility", "interaction"),
+  prob = 0.8, # 80% intervals
+  #prob_outer = 0.99, # 99%
+  point_est = "mean")+
+  labs(title= "Pre-critical region", x = "effect in ms")+
+  vline_0(linetype="dashed")+
+  scale_x_continuous(breaks=(seq(-50, 40, 10)), limits = c(-50, 40)) +
+  theme_bw(base_size=12) +
+  annotate("text", x = 35, y = 3.8, label = "2 — 13 — 24", size=3) +
[…]

### 2024-07-24 paper/main.tex
Updated via the Open Science Framework

@@
 Syntactic interference is manipulated via the syntactic status of the distractor, which is either the subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}) and hence a non-subject.
-Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase.
+\Copy{franck_refs}{Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase \textcolor{blue}{ -- even if the distractor occurs in an embedded clause {\parencite[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary}.}}}\label{franck_refs}
 
@@
 \textcolor{blue}{In the SPR experiment, t}he reading times of the immediately following spill-over region and of the pre-critical region were also of interest.
-Previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in the pre-critical region (we return to this point in the discussion sections).
+Previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in the pre-critical region (we return to this point in the discussion sections\textcolor{blue}{, see Figure \ref{fig:previous_vs_present} A}).
 A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another (\textit{erzählt hatte, trank}, literally `told, drank').
@@
 Traditionally, it is associated with lexical-semantic processing, but the finding of multi-modal N400 effects suggests that it rather indexes differences in the processing of meaningful stimuli in general \citep{kutas_federmeier2011}.
-The N400 amplitude is greater for words that deviate from the meaning activated by its context compared to words that fit its context \citep[see, e.g.,][]{kutas_hillyard, Kutas_Iragui_1998}.
-Consequently, the amplitude of the N400 to a specific word is a negative function of its predictability within the given context \citep{kutas_hillyard1984}.
+The N400 amplitude is greater for words that deviate from the meaning activated by its context compared to words that fit its context \citep[see e.g.,][]{kutas_hillyard, Kutas_Iragui_1998}.
+Consequently, the amplitude of the N400 to a specific word is a negative function of its predictability within the given context \citep[see e.g.,][]{kutas_hillyard1984, frank2015_erp_surprisal}.
 Additionally, the N400 elicited by isolated words is negatively correlated with their frequency \citep{kutas&federmeier_2000}.
@@
 
-\textcolor{blue}{\section{Data availability statement}
-All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}\label{statement}
+\Copy{osf}{\textcolor{blue}{\section{Data availability statement}
+All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}}\label{statement}
 
@@
 The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
-\textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task.
+\textcolor{blue}{In addition to the predictors of interest, we included the trial id to account for potential effects of fatigue or adaptation to the task.
 The trial id of the 100 trials per session was recoded to span from 0 to 1.
 This was done to bring all fixed effects to a comparable scale which decreases the run time of the models.
-The models were run with full random effects (varying intercepts and slopes of all fixed effects and their interactions for participants and items).}
+The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items).}
 
@@
 Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
-\textcolor{blue}{The 95 \% credible intervals for syntactic interference included zero in the pre-critical and critical regions.
+\Copy{SPR_results}{\textcolor{blue}{The 95 \% credible intervals for syntactic interference included zero in the pre-critical and critical regions.
 In the spill-over region, it included only positive values ([1, 15] ms).
@@
 \end{figure}
+}\label{SPR_results}
+
+\Copy{adaptation}{\textcolor{blue}{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last, i.e., 100th, trial, reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
+This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment.
+While this adaption is interesting, it is not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}
 
@@
 
-Was there any evidence for the effects estimated above?
+Was there any evidence for the effects \textcolor{blue}{of interest }estimated above?
 The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}.
 The analyses use all the data from Experiments 1a and 1b.
-\textcolor{blue}{We found very to extremely strong evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
+\Copy{SPR_bfs}{\textcolor{blue}{We found very to extremely strong evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
 In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5).
-There was strong evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ < 0.5).
+There was strong evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
 Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was weak evidence for syntactic interference (BF$_{10}$ = 2.8).
-Similarly, there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ < 1).
-Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}
+Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
+Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}
 
 \textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but inconclusive results in the critical region.
-In contrast, the Bayes factor analysis provided either at best weak evidence for syntactic interference and the interaction or even evidence against them.}
-
-\begin{figure}[H]
+In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}
+
+\Copy{fig_bf_spr}{\begin{figure}[htpb]
     \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
@@
 \end{figure}
+}}
 
@@
 We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-\textcolor{blue}{Reading times at the critical verb were inconclusive for semantic interference and showed evidence against syntactic interference.
-We will discuss the absence of syntactic interference in our data in the general discussion.
-The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occurred even before the retrieval event could begin.} The effects in the pre-critical region are discussed next.
+\textcolor{blue}{Reading times of the pre-critical word and critical verb showed evidence against syntactic interference.
+In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms.
+We will discuss the almost complete absence of syntactic interference in our data in the general discussion.
+The reading time results at the critical verb were inconclusive for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
+This means that  semantic interference occurred even before the retrieval event could begin.
+Reading times at the critical verb showed very weak support for a small effect of a syntactic $\times$ semantic interference interaction.
+There was evidence against the interaction under broader priors and in the surrounding regions.}\Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
+Therefore, we refrain from discussing effects occurring in these regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.
 
@@
 
-\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
+\Copy{plausib_anim_inan}{\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
 To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}.
-Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}).
-Forty-four native speakers of German (mean age XX, age range: , w, m), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible).
+Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience).
+}
+
+\begin{exe}  
+\ex \label{ex:materials_repeat} Example item (critical word in bold, distractor in italics) of the present study:
+    \begin{xlist}   
+    \ex {High syntactic interference with high / low semantic interference:}\label{ex:hisyn_repeat} 
+    \gll Die  Nachbarin glaubte,	dass	der Witwer,  der  erzählt hatte, dass der \textit{Einbrecher} / \textit{Verlust}  schrecklich war, abends regelmäßig \textbf{trank}, um zu vergessen.\\ 
+    The\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} believed that the widower who told had that the burglar / loss awful was in.the.evening regularly drank in.order to forget \\
+    \trans `The neighbor believed that the widower, who had told her that the loss was awful, drank regularly in the evenings to forget.' \\
+    \end{xlist}
+\end{exe}
+
+\textcolor{blue}{Forty-four native speakers of German (mean age: 26 years old, age range: 18 - 35 years, 21 female, 23 male), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible).
+The mean ratings per item ranged from 3.6 to 6.6.
 The experiment included 40 implausible filler sentences which were created by exchanging nouns between the fillers used in the SPR experiment (e.g., Der Sohn des Kaisers, der den Feldzug gewonnen hatte, plante äußerst strategisch, weil er die Gießkanne an sich reißen wollte., `The son of the emperor who had won the campaign planned extremely strategically because he wanted to take the watering can by force.').
-Participants spent approximately 37 minutes to complete the task and received 6.50 \pounds \, as compensation.
-The results of an ordinal brms model analyzing the ratings are shown in Figure X.
-Sentences with an inanimate distractor had a slightly higher probability to gain a high rating on the plausibility scale ($\beta$ = XX), i.e., the low semantic interference conditions are more plausible.
-To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran an additional model which used the same specifications as described in the Statistical Analyses section with prior N(0,0.5) and additionally, the centered plausibility ratings as a fixed effect.
-Figure X presents the posteriors from this model.
+The mean rating of the implausible fillers ranged from 1.1 to 3.2.
+Participants spent approximately 37 minutes to complete the task and received \pounds 6.50 as compensation.
+The results of an ordinal brms model analyzing the ratings of the critical items are shown in Figure \ref{fig:plausibility_anim_inanim}.}
+
+\begin{figure}[htpb]
+    \centering
+        \caption{Posterior probability with 95\% credible intervals of each rating for items with inanimate and animate distractors.}
+    \label{fig:plausibility_anim_inanim}
+    \includegraphics[width=0.8\linewidth]{images/plausibility_anim_inan.jpg}
+\end{figure}
+
+\textcolor{blue}{Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible.
+To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
+This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings.
+For the slopes, we used a Normal(0,0.5) prior; the priors for all other parameters were the same as presented in Table \ref{tab:spr_priors}.
+The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.
+Figure \ref{fig:posteriors_plausibility} presents the posteriors from this model.
+Although an increase in plausibility lead to faster reading times (CrI [-30, -6] ms), there was still an effect of semantic interference (CrI [2, 24] ms).
+The Bayes factor provided weak evidence for semantic interference in this analysis (BF$_{10}$ = 2.4).
+The interaction of semantic interference and plausibility was centered around zero.
+In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
+However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference lead to slower reading times independent of plausibility.}\label{plausib_anim_inan}
+
+\begin{figure}
+    \centering
+        \caption{Posteriors of the semantic interference and plausibility effect and their interaction on the reading times in the pre-critical region.
+The numerical values are the means and 95\,\% credible intervals.
+The blue vertical lines represent the median and the blue shaded areas are 80\,\% credible intervals.}\label{fig:posteriors_plausibility}
+    \includegraphics[width=0.8\linewidth]{images/posteriors_spr_pooled_774_plausibility.png}
+\end{figure}
 }
 
-At the pre-critical region, there was clear evidence for semantic interference.
-As this cannot be due to retrieval interference, the most likely explanation is encoding interference \citep{Oberauer_Kliegl_2006}.
-The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
+At the pre-critical region, there was evidence for semantic interference.
+As this cannot be due to retrieval interference, the most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}\textcolor{blue}{, but see the potential confounds discussed above).} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
 
@@
 We used Bayesian linear mixed models to analyze the single trial EEG data in response to the critical verb.
-We averaged the activity of 12 centro-parietal electrodes (Cz, C3/4, CPz, CP1/2, CP5/6, Pz, P3/4, POz) in the standard N400 time window (300 to 500 ms post critical word onset) for all analyses.
+We averaged the activity of 12 centro-parietal electrodes (Cz, C3/4, CPz, CP1/2, CP5/6, Pz, P3/4, POz) in the standard \textcolor{blue}{time windows of the  N400 (300 to 500 ms post critical word onset) and P600 (600 to 900 ms post critical word onset)} for all analyses.
 The models included sum-contrast coded fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
@@
     \midrule
-  Intercept & Normal(2, 5)& [-8, 12]\\
+  Intercept & \textcolor{blue}{Normal(0, 5)}& \textcolor{blue}{[-10, 10]}\\
   \cmidrule{2-3}
-  \multirow{4}{1cm}{slope} & Normal(0, 0.1) & [-0.2, 0]\\
-  & Normal(0, 0.5)& [-1, 1]\\
-  & Normal(0, 1) & [-2, 2]\\
-  & Normal(0, 2) & [-4, 4]\\
+  \multirow{4}{1cm}{slope} & \textcolor{blue}{Normal(0, 0.1)} & \textcolor{blue}{[-0.2, 0.2}]\\
+  & \textcolor{blue}{Normal(0, 0.5)}& \textcolor{blue}{[-1, 1]}\\
+  & \textcolor{blue}{Normal(0, 1)} & \textcolor{blue}{[-2, 2]}\\
+  & \textcolor{blue}{Normal(0, 2)} & \textcolor{blue}{[-4, 4]}\\
   \cmidrule{2-3}
@@
 For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats,schad_etal_2022_BF}.
-We chose slope priors which assume a wide range of effect sizes from [$-0.2$, 0.2] to [$-4$, 4]\,$\mu V$.
+We chose slope priors which assume a wide range of effect sizes from \textcolor{blue}{[$-0.2$, 0.2] to [$-4$, 4]\,$\mu V$}.
 
@@
 grand average ERPs elicited by the critical verb for all conditions.
-\textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as -1\,$\mu V$.
-The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600.
+\textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as $\pm$1\,$\mu V$.
+}\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600.
 The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals include zero for both the N400 and the P600.
 The 95\% credible interval of the interaction was centered around zero for the N400.
-For the P600, it included mostly negative values but crossed zero.}
+For the P600, it included mostly negative values but crossed zero.}}\label{ERP_results1}
 
@@
 The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference.
-In contrast, Figure \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in the N400 spatio-temporal window.
-Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (c) does also not suggest that there was a syntactic interference effect in any other time window.
+\textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a reduced P600 amplitude for high vs.\ low semantic interference.
+See Figure X for the topographic distribution of P600 effect.} \Copy{ERP_results2}{In contrast, Figure \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in \textcolor{blue}{neither the N400 nor P600 spatio-temporal windows}.}\label{ERP_results2} Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (c) does also not suggest that there was a syntactic interference effect in any other time window.
 
@@
 
-Bayes factors are shown in Figure \ref{fig:eeg_bfs}. The Bayes factors provided very strong to \textcolor{blue}{moderate} evidence for the semantic interference effect for all priors 
-\textcolor{blue}{(Normal$_{-}$(0, 0.1): BF$_{10}$ = 11,
-Normal$_{-}$(0, 0.5): BF$_{10}$ = 80,
-Normal$_{-}$(0, 1): BF$_{10}$ = 9,
-Normal$_{-}$(0, 5): BF$_{10}$ = 5)}.
-
-By contrast, the syntactic interference effect had only anecdotal evidence  under the \textcolor{blue}{narrowest prior (Normal$_{-}$(0, 0.1): BF$_{10}$ = \textcolor{blue}{2.18}) and under the Normal$_{-}$(0, 0.5) prior there was no evidence for or against it (BF$_{10}$ = 1.05).} There was evidence \textit{against} syntactic interference under wider priors
-(Normal$_{-}$(0, 1): BF$_{10}$ = \textcolor{blue}{0.53},
-Normal$_{-}$(0, 5): BF$_{10}$ = \textcolor{blue}{0.1}).
-What this means is that, in our data, only if we assume a priori that the effect size is \textcolor{blue}{very small, i.e., maximally $-$0.2\,$\mu V$,} there is very weak evidence in favor of syntactic interference.
-
-Across all priors, the Bayes factors provided evidence \textit{against} the interaction of semantic and syntactic interference (BF$_{10}$ < 0.85).
-
+Bayes factors are shown in Figure \ref{fig:eeg_bfs}.
+\Copy{ERP_results3}{\textcolor{blue}{In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect for all priors
+(Normal(0, 0.1): BF$_{10}$ = 6,
+Normal(0, 0.5): BF$_{10}$ = 17.5,
+Normal(0, 1): BF$_{10}$ = 9.4,
+Normal(0, 2): BF$_{10}$ = 5).
+In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and no evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$.
+By contrast, Bayes factors unequivocally provided evidence against the syntactic interference effect in both spatio-temporal windows (BF$_{10}$ $<$ 0.9).
+The only exception was under the narrowest prior (Normal(0, 0.5)) in the N400 window, there was either no evidence for syntactic interference or even evidence against it (BF$_{10}$ $<$ 0.9).
+Similarly, Bayes factors unequivocally provided evidence against the interaction in both spatio-temporal windows.
+The only exception was under the narrowest prior (Normal(0, 0.5)) in the P600 window, there was no evidence for or against the interaction (BF$_{10}$ = 1.2).
+In sum, in our ERP data, there was decisive evidence for the semantic interference effect and no evidence or even evidence against syntactic interference and the interaction of syntactic and semantic interference.}}\label{ERP_results3}
+
+\Copy{fig_bf_eeg}{
 \begin{figure}[H]
-    \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the time window 300 to 500 ms post critical verb onset at centro-parietal electrode sites.}
-    \label{fig:eeg_bfs}
-    \centering
-    \includegraphics[width=0.8\textwidth]{images/BF_plot_N103_cp_300_500.jpg}
-\end{figure}
+    \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the N400 (300 - 500 ms post critical verb onset) and P600 (600 - 900 ms post critical verb onset) time windows at centro-parietal electrode sites.}\label{fig:eeg_bfs}
+    \centering
+    \includegraphics[width=1\textwidth]{images/BF_plot_eeg.jpg}
+\end{figure} }
 
@@
 
-The Bayes factors analyses found \textcolor{blue}{very strong to moderate evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($-$0.2 to $-11$\,$\mu V$)}.
-Syntactic interference had much weaker evidence in the N400 amplitude; only anecdotal evidence was found under \textcolor{blue}{the narrowest prior, which assumed effect sizes smaller than $-$0.2\,$\mu V$.
-No evidence for or against syntactic interference was found under a prior which assumed a reasonable effect size of up to $-1$\,$\mu V$ ($BF_{10}$=1.05).
-There was clear evidence against syntactic interference under wider priors (effect sizes of up to $-$2 or up to $-11$\,$\mu V$)}.
-The evidence against an interaction under all priors was unequivocal.
-\textcolor{blue}{In sum, the ERP results showed clear evidence for semantic interference and clear evidence against an interaction of syntactic and semantic interference.
-The evidence for or against syntactic interference was dependent on the a priori assumed effect size.
-However, even when very small effects were assumed a priori, the evidence for syntactic interference was weak at best.
-Consequently, we conclude that the ERP data showed no syntactic interference.} \textcolor{red}{Add conclusion after SPR results are revised.}
+\Copy{ERP_discussion}{The Bayes factors analyses found \textcolor{blue}{moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$).
+In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference.
+Both the N400 and P600, were more negative under high vs.
+low semantic interference.
+In contrast, there was no evidence or even evidence against effects of syntactic interference and the interaction on the N400 and P600.
+In sum, the ERP results showed clear evidence for semantic interference and no evidence or evidence against syntactic interference and the interaction of syntactic and semantic interference.
+Consequently, we conclude that the ERP data showed no syntactic interference and no interaction effect.}}\label{ERP_discussion}
 
@@
 The more complex memory retrieval under high semantic interference resulted in a more negative N400.
-
-Consistent with the self-paced reading study, there is clear evidence for semantic interference, and \textcolor{blue}{little to no evidence for} syntactic interference.
+\textcolor{blue}{The modulation of the P600 might be due to facilitatory interference.
+The P600 is typically affected by syntactic processing, with more positive responses for more complex syntactic processing, e.g., reanalysis \citep{osterhout&holcomb_1992}.
+In the present design, the animate distractor might have facilitated integration of the verb with its (false) subject.
+Integration with one of two potentially good candidates could be easier than integration when there is only one good candidate.
+The smaller size of the P600 effect compared to the N400 amplitude modulation could indicate that this facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration.
+}
+
+Consistent with the self-paced reading study, there is clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}.
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
@@
 
-\section{Quantitative predictions of the Lewis and Vasishth (2005) model for the present design}
-
-% root mean square deviation
-
-\textcite{mertzen} presented simulations that show approximately equally sized interference effects 
-for both the syntactic and semantic manipulations.
-Their model assumed that three cues were used at the verb: \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}.
-The upper parts of Figure \ref{fig:model_predictions} a and b reproduce the \textcite{mertzen} simulations.
-Figure \ref{fig:model_predictions_rts} shows the prior predictive distributions of by-condition retrieval times that were generated by sampling from a prior distribution ($Beta(a=4,b=6)$) for the free scaling parameter $F$ and then generating data from the model while holding all other parameters in the model at fixed values.
-The $F$ parameter scales the activation to a retrieval time.
-The prior $Beta(a=4,b=6)$ was chosen in earlier work \parencite{jaeger_etal_2020,vasishth2020_abc}.
-Figure \ref{fig:model_predictions_effects} shows the prior predictive distributions of the effects of interest.
-
-\begin{figure}[htbp]
-    \centering
-    \caption{Prior predictions from the \textcite{Lewis2005} model as implemented in R by \textcite{engelmann_etal_2019}. 
-    The implemented models are available with the code and data provided with this paper.}
-    \label{fig:model_predictions}
-    \centering
-         \begin{subfigure}[t]{\textwidth}
-         \centering
-        \caption{Prior predicted by-condition retrieval times.
-The upper panel shows the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panel shows the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-Diamonds represent the means (numbers to the right) and blue dots show each simulated data point.}
-         \label{fig:model_predictions_rts}
-         \includegraphics[width=\textwidth]{images/PriorPredictedMeans_interACTversions.png}
-     \end{subfigure}
-
-     \begin{subfigure}[t]{\textwidth}
-         \centering
-        \caption{Prior predicted effect sizes, i.e., retrieval time differences, for syntactic and semantic interference and their interaction.
-The upper panels show the predictions using the three retrieval cues used in \textcite{mertzen}, and the lower panels show the predictions using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.}
-         \label{fig:model_predictions_effects}
-         \includegraphics[width=\textwidth]{images/PriorPredicted_interACTversions.png}
-     \end{subfigure}
+\section{Quantitative predictions of the Lewis and Vasishth (2005) model for \textcolor{blue}{ERPs elicited in} the present design}
+
+\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times.
+Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized interference effects
+for both the syntactic and semantic manipulations at the critical verb.
+Since we have found that the reading times at the critical verb, i.e., the retrieval site, were confounded by pre-critical effects, which were best explained by \textit{encoding} interference, it makes no sense to compare the results from the present SPR experiment to predictions from the Lewis and Vasishth (2005) model of sentence processing
+as skilled memory retrieval.
+Instead we use the Lewis and Vasishth (2005) model to generate predictions on the microvolt scale for ERP data by adjusting the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.}
+
+The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$).
+This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal likelihood for all effect sizes in this range.
+This uninformative prior was chosen because there exists no previously published ERP data on the Van Dyke (2007) design.
+After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
+\textcolor{blue}{We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study.
+To reflect that larger predicted effect sizes correspond to more negative N400 amplitudes, the predicted effects from the model} were multiplied with $-1$.
+To avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects.
+
+\textcolor{blue}{
+For their recent simulations, \textcite{mertzen} used a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}.
+Although it is clear that our data does not support equally sized effects of syntactic and semantic interference, we decided to use their version of the model for our modeling because it has been established as a good model for the present design.
+We have implemented two alternative versions of the model.
+A model which puts more emphasis on hierarchical syntactic structure.
+Given the evidence against syntactic interference in our data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
+We speculate that this could be due to the parser tracking} hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
+Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} \textcolor{blue}{-- as implemented by \textcite{mertzen}} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
+\textcolor{blue}{This idea actually goes back to the roots of the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005}.
+The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.} This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
+In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
+\textcolor{blue}{Lastly, we implemented a model with one cue which unambiguously matches the subject.
+This cue would be the master composite cue \{$\pm$ animate-subject-in-same-clause\}.
+This model naturally does not predict any interference effects.
+}
+
+It is informative to fit the models to the data and then examine the posterior predictive distribution of the \textcolor{blue}{ERPs} for each effect. 
+We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the \textcolor{blue}{ERP} data.
+This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the  scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ in the \textcolor{blue}{ERP} data.
+`Close' is defined as $d \pm \,2\times SE$ ms; for details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
+The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}.
+
+\Copy{figpriorposteriordata}{
+\begin{figure}[htpb]
+    \caption{Prior and posterior predicted effect sizes on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval mode.}
+    %using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. 
+    The data with 95 \% credible intervals are shown in red.}\label{fig:model_predictions_ERP}
+        \centering
+         \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
 \end{figure}
-
-Although the semantic interference effect estimates from the SPR study fall within the predicted range from the model, the syntactic interference effect is quite divergent: as discussed earlier, the SPR data show evidence against syntactic interference, but the model predicts a syntactic interference effect of approximately the same magnitude as semantic interference.
-Given the evidence against syntactic interference in the SPR data and the flip between evidence for and evidence against it depending on the priors in the ERP data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
-
-The absence of (and evidence against) syntactic interference is surprising because the psycholinguistic literature has clearly established that syntactic constraints play a role in sentence processing.
-For example, \textcite{dillon2013} and \textcite{Sturt2003} argued that principle A of the binding theory controls the retrieval process in the antecedent-reflexive dependency, effectively eliminating or attenuating interference from other overlapping features (such as number marking) of intervening nouns that cannot be an antecedent for the reflexive.
-For further evidence for this idea from individual differences, see \textcite{yadav2021individual}.
-Even more relevant to the present study is the fact that \textcite{mertzen} did find syntactic interference effects.
-
-Could the absence of syntactic interference be reconciled with a cue-based retrieval account?
-We speculate that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
-Specifically, the relevant syntactic cues that might be used at the verb are not  \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} but rather a composite cue \{$\pm$ subject-in-same-clause\}.
-This proposal is in the spirit of the  \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
-In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
-
-We evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+}
+
+\textcolor{red}{To do (Pia): Describe figure\\
+model comparison (cross-validation)}
+%For the ERP modeling, 473 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions.
+The prior and posterior predictions of the model overlap with the effects of interest in the data.
+The general effect pattern (no syntactic interference effect \textcolor{blue}{and no interaction}, but a semantic interference effect) is similar in both the prior and posterior predictive distributions.
+Therefore, the assumption that the relevant syntactic cue is the \{$\pm$ subject-in-same-clause\} cue -- rather than the previously assumed \{$\pm$ grammatical subject\} cue --  seems reasonable for the present ERP study.
+
+%We evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
 The results of the prior predictive simulations are shown in the lower parts of Figure \ref{fig:model_predictions} a and b.
@@
 
-It is informative to fit the model to the data and then examine the posterior predictive distribution of the reading times for each effect. 
-We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the SPR data.
-This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the  scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ in the SPR data.
-`Close' is defined as $d \pm \,2\times SE$ ms; for details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
-
-To avoid overfitting, we fitted the scaling parameter $F$ only to the semantic interference effect (and not the syntactic interference effect or interaction). 
-We obtained only 355 samples of $F$ (from 20,000 iterations of the ABC algorithm); this is not surprising given the relatively wide prior on $F$.
-Figure \ref{fig:model_predictions_posterior} shows the posterior predictive distributions of the effects of interest in comparison to the prior predictive distribution and our SPR data.
-The posterior predictions match the syntactic and semantic interference effects in our data, but underestimate the interaction.
-
-\Copy{figpriorposteriordata}{
-\begin{figure}[!htbp]
-    \caption{Prior and posterior predicted effect sizes for syntactic and semantic interference and their interaction using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-The data with 95 \% credible intervals are shown in red.}
-\begin{subfigure}[t]{\textwidth}
-\caption{On the millisecond scale, i.e., for reading times.}\label{fig:model_predictions_posterior}
-    \centering
-    \includegraphics[width=\textwidth]{images/priorpred_vs_postpred_vs_data.png}
-    \end{subfigure}
-
-    \begin{subfigure}[t]{\textwidth}
-        \centering
-         \caption{On the microvolt scale, i.e., for negative ERPs.}\label{fig:model_predictions_ERP}
-         \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
-     \end{subfigure}
-\end{figure}
-}
-
-The \citet{Lewis2005} model is generally used to predict reading times.
-The scaling factor $F$ links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R}.
-But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data.
-The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
-We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$).
-This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal likelihood for all effect sizes in this range.
-This uninformative prior was chosen because there exists no previously published ERP data on the Van Dyke (2007) design.
-We are modeling N400 amplitude differences, so larger predicted effect sizes correspond to a more negative N400.
-To reflect this, the predicted effects were multiplied with $-1$.
-As in the reading time modeling, to avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects.
-For the ERP modeling, 473 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions.
-The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}.
-The prior and posterior predictions of the model overlap with the effects of interest in the data.
-The general effect pattern (no syntactic interference effect, but a semantic interference effect) is similar in both the prior and posterior predictive distributions.
-Therefore, the assumption that the relevant syntactic cue is the \{$\pm$ subject-in-same-clause\} cue -- rather than the previously assumed \{$\pm$ grammatical subject\} cue --  seems reasonable for the present ERP study as well.
-
-Thus, the theoretical assumption that is consistent with the observed SPR and ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
+Thus, the theoretical assumption that is consistent with the observed ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
 This is of course a speculative claim that would need to be further investigated empirically.
@@
 
-Taken together, the SPR and ERP experiments consistently showed effects of semantic interference and no effects of syntactic interference.
-This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
+\Copy{only_this_design}{Taken together, the SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and no effects of syntactic interference.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
@@
 
-\textcolor{red}{Rewrite this section to reflect the new discussion.} In the remainder of this section, we discuss two important issues that relate to previous experiments that investigated syntactic and semantic interference.
+\textcolor{blue}{In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design.
+Then we} discuss two important issues that relate to {the broader interference literature}.
 The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.
@@
 \subsection{Comparison with previous findings}
-\textcolor{blue}{The present study used the same design as two previous studies \parencite{vandyke07, mertzen}.
-Figure \ref{fig:previous_vs_present} A shows the estimated reading times differences from the previous studies and our self-paced reading experiment.
+\Copy{comparison}{\textcolor{blue}{The present study used the same design as two previous studies \parencite{vandyke07, mertzen}.
+Figure \ref{fig:previous_vs_present} A shows the estimated reading times differences from the previous studies and the present self-paced reading experiment.
 Because the \textcite{vandyke07} data was not available to us, we derived estimates for their study from the summary statistics in the published paper.
 \textcite{mertzen} made their data available.
-In comparison to the previous estimates, the effects of the present study are relatively small and they are associated with less uncertainty.
-To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies.
+In comparison to the previous estimates, the effects of the present study are relatively small and associated with less uncertainty.
+It seems like the present estimates of syntactic interference are smaller, i.e., closer to zero, than the previous estimates and larger, i.e., further from zero in the positive direction, for the interaction.
+The present and previous estimates of semantic interference seem relatively consistent: They indicate a positive effect which is largest in the pre-critical region and smallest in the critical region.
+To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies to compare them to the present Bayes factors.
+For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
 For \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE).
-Then using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors using the Savage-Dickey method.
-Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from models which we run on their data along to our own analyses described in Statistical Analyses for Experiments 1a and b.
-The Bayes factors of the effects of interest in self-paced reading times and total fixation times of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}
+Then, using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors with the Savage-Dickey method.
+Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we run on their data analog to our own analyses described in Statistical Analyses for Experiments 1a and b.
+The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}
 
@@
 In the critical region, most experiments provide evidence against semantic interference, the two self-paced reading experiments, being the exceptions, provide evidence for the effect.
-In the spill-over region, the Bayes factor results for semantic interference are mixed: the two self-paced reading experiments provide extremely strong evidence for the effect, the total fixation times for two of the eye-tracking experiments provide weak evidence for the effect and the total fixation times of two other eye-tracking experiments provide evidence against the effect.
-Importantly, the clustering in the total fixation time results does not align with the authors of the studies.}
+In the spill-over region, the Bayes factor results for semantic interference are mixed: the two self-paced reading experiments provide extremely strong evidence for the effect, two of the eye-tracking experiments provide weak evidence for the effect and two other eye-tracking experiments provide evidence against the effect.
+Importantly, the clustering in the results does neither align with the investigated language nor with the authors of the studies.}
 
 \textcolor{blue}{The Bayes factors for the interaction in all regions are rather consistent between studies and are either inconclusive or provide evidence against the interaction.
-The only exception is the very large Bayes factor of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1.
+The only exception is the very large Bayes factor in favor of the effect for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1.
 In sum, while the present and previous studies show comparable region-wise results for semantic interference and the interaction, they differ remarkably regarding syntactic interference in the pre-critical and critical region.
 In the pre-critical region, the present study is the obvious outlier, being the only study to provide evidence against the effect.
-In the critical region, four of the investigated experiments show weak evidence for or against syntactic interference with two outliers: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 providing extreme evidence for the effect and the present study providing strong evidence against it.
-In conclusion, the most remarkable difference between the present study and previous findings concerns the syntactic interference in the pre-critical region.}
-
-%The difference in observed effects \textcolor{blue}{between our SPR experiment and the other reading time studies} could be explained by differences in statistical power or by methodological differences.
-
-The present study has higher statistical power than the previous studies.
-\citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively.
-\textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
-In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
-Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design.
-Consequently, the estimates from the present study are the most precise ones so far for this design.\textcolor{blue}{This is reflected by the tighter estimates in Figure \ref{fig:previous_vs_present}}.
-The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
-
-An obvious difference from previous studies is the experimental methodology.
-The majority of the previous experiments employed eye-tracking while reading, while we used a self-paced reading task.
-Self-paced reading (SPR) might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format in SPR: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR}.
-SPR is likely to be more demanding on the comprehender's memory, which might make it easier to detect interference effects.
-This argumentation is supported by our estimation of the results of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1, the only other SPR experiment using this design:
-In the critical region, the estimates of both main effects (syntactic and semantic interference) are very similar to the results of the present study; the interaction effect observed by \citeauthor{vandyke07} in the spill-over region is also similar to our estimate.
-However, it remains unclear why semantic interference was especially prominent in our study compared to the previous ones.
-
-\Copy{syn_confounds}{\textcolor{blue}{The difference in methodology lead to another difference between our study and the previous studies.
+In the critical region, four of the investigated experiments show either weak evidence for or weak evidence against syntactic interference, but there are two outliers: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extreme evidence for the effect and the present study provided strong evidence against it.
+In conclusion, the most remarkable difference between the present study and previous findings concerns syntactic interference in the pre-critical region.}}
+
+\textcolor{blue}{The lack of syntactic interference in the pre-critical region in the present study compared to the previous studies could be explained by differences in i) methodology, ii) items and/or iii) statistical power.
+We discuss each of these differences in turn.}
+
+An obvious difference between the studies is the experimental methodology.
+The majority of the previous experiments employed eye-tracking while reading, while we used self-paced reading \textcolor{blue}{and EEG / ERPs.
+However, word-by-word} self-paced reading (SPR) \textcolor{blue}{and ERPs elicited by single words presented in the typical rapid serial presentation (RSVP) format, if anything} might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR} \textcolor{blue}{and RSVP.
+Word-by-word presentation} is likely to be more demanding on the comprehender's memory, which \textcolor{blue}{should} make it easier to detect interference effects.
+\textcolor{blue}{So, the differences in methodology do not offer a straightforward explanation for the lack of syntactic interference in our data compared to the previous studies.}
+
+\Copy{linebreak}{\textcolor{blue}{The difference in methodology lead to another difference between our study and the previous studies.
 In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line.
-Whether \citeauthor{vandyke07}'s \citeyear{vandyke07} SPR experiment used line breaks or not is not reported.
+Whether \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment used line breaks or not is not reported.
 In our SPR experiments, it was necessary to introduce line breaks into the sentences.
@@
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-Nevertheless, neither our SPR nor our EEG experiments showed evidence for syntactic interference.}
-
-\textcolor{blue}{All our experiments (SPR and EEG) displayed the subject with a comma.
-Commas are mandatory in German to separate the subordinate clauses from the main clause.
-Similar to the line break discussed above, the comma could have emphasized the syntactic structure of the sentences and therefore decreased syntactic interference from the distractor which occurred inside a relative clause and not in the clause of the critical verb.
-But if this comma provided information about the syntactic structure of the items then it should have so as well in \citeauthor{mertzen}'s \citeyear{mertzen} German study because they also presented the subject with the comma.
-So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.}}\label{syn_confounds}
+Nevertheless, neither our SPR nor our EEG experiments showed decisive evidence for syntactic interference.}
+}\label{linebreak}
+
+\textcolor{blue}{We now turn to the differences regarding the experimental items.
+The three studies investigating syntactic and semantic interference with the same design used either English or German items.
+The manipulated features, i.e., subjecthood and animacy, are rather universal and are expressed in the same way across these languages.
+Cue-based retrieval as a theory on the memory processes during sentence comprehension implicitly states that these memory processes do not differ cross-linguistically \citep{lewis06, Lewis2005}.
+This is in line with no apparent clustering of the English vs.\, German results (see Figure \ref{fig:previous_vs_present} A).
+So, the use of materials in different languages cannot account for the differences in results between the studies.}
+
+\textcolor{blue}{The previous and present items differed in the pre-critical material.
+\citet{vandyke07} used one region consisting of two words (e.g.
+`yesterday afternoon') which were analyzed together.
+\textcite{mertzen} used a one-word pre-critical region (e.g.
+`tatsächlich', indeed).
+In the present study, we added a pre-pre-critical region to help us differentiate between the potential sources of the pre-critical effects in the literature as discussed in the Discussion of the SPR experiments.
+So, the present study was the only one which had a pre-critical region that did not directly follow a clause boundary and it was also the only one to not find syntactic interference in the pre-critical region.
+These two observations combined suggest that the pre-critical syntactic interference effect in the previous studies might be a spill-over effect from the syntactic manipulation.
+The syntactic interference manipulation in all the studies using this design was confounded with syntactic complexity of the material intervening between the subject and the (pre-)critical region.
+In the low syntactic interference conditions, it was a simple relative clause (who VP PP).
+In the high syntactic interference conditions, it was a relative clause with an embedded complement clause (who VP that NP VP).
+This difference in syntactic complexity could have ``spilled over'' into the pre-critical region of the previous studies causing the syntactic (interference) effect.
+While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region.
+However, Figure \ref{fig:whole_sentence} C shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.}
+
+\Copy{comma}{\textcolor{blue}{Another item-related difference is that all our experiments (SPR and EEG) displayed the subject with a comma.
+Commas are mandatory in German to separate subordinate clauses from the main clause.
+Similar to the line break discussed above, the comma could have emphasized the syntactic structure of the sentences and therefore decreased syntactic interference from the distractor which occurred in another clause than the critical verb.
+But if this comma provided information about the syntactic structure of the items then it should have done so in \citeauthor{mertzen}'s (\citeyear{mertzen}) German study as well because they also presented the subject with a comma.
+But \citet{mertzen} found syntactic interference and we did not.
+So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.}}\label{comma}
 
@@
 9]{mertzen}.
-The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
-But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
-While this additional distractor might have increased syntactic interference compared to our study (and \citeauthor{vandyke07}'s study as well actually), it is not obvious why it would decrease semantic interference compared to our study.}\label{items_mertzen}
+}
 
@@
 
-So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region?
-Apart from \textcolor{blue}{the additional distractor and} the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study.
-While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
+\textcolor{blue}{The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
+But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
+This additional distractor might have increased syntactic interference compared to our study.
+But it would have increased syntactic interference in comparison to \citeauthor{vandyke07}'s study as well and this was not the case (see Figure \ref{fig:previous_vs_present} A).}\label{items_mertzen} \textcolor{blue}{All in all, differences in the items provide no clear explanation why the previous studies found syntactic interference and the present study did not.}
+
+\textcolor{blue}{A plausible explanation for the divergence between studies is a difference in statistical power:} The present study has higher statistical power than the previous studies.
+\citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively.
+\textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
+In contrast, \textcolor{blue}{in our SPR experiments,} we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
+Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design.
+Consequently, the estimates from the present study are the most precise ones so far for this design.
+\textcolor{blue}{This is reflected by the tighter estimates in Figure \ref{fig:previous_vs_present} A}.
+The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
+\textcolor{blue}{In contrast, the large estimates of the previous studies are likely Type M errors, i.e., exaggerations, which are common in low-powered studies \parencite{gelman_carlin}.
+That at least some of the previous studies had low power is further emphasized by the large uncertainty associated with some of the previous estimates (see Figure \ref{fig:previous_vs_present} A).
+In sum, while there are differences in the employed methodology and items between the present and previous studies, the most likely cause of the different results is the difference in statistical power.
+However, an alternative explanation could be that syntactic interference might differ inter-individually \parencite{yadav2021individual} and that the previous studies by chance sampled disproportionally many participants who show syntactic interference, while we by chance sampled participants who do not.
+}
 

### 2024-07-24 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \usepackage{graphicx}
+\usepackage{float}
 \usepackage{subfigure}
@@
 \begin{document}
-\maketitle
+\hfill \today
 %\section{}
@@
 
-We want to thank the reviewers and editor for their detailed and constructive comments; these have significantly improved the paper.
-We believe we have addressed every point raised by the editor and reviewers; the comments with our responses and a record of the changes in the paper are listed below.
-When relevant, we provide references to pages from the revised paper.
+\noindent \textbf{Prof. Chris Jarrold}\\
+\noindent \textbf{Associate Editor}\\
+\noindent \textbf{Journal of Memory and Language}
+\vskip 1em
 
-All the changes in the paper are marked in blue.
+\noindent \textbf{Response letter for Submission JML-24-17}
+\vskip 2em
+\noindent Dear Prof. Jarrold, 
+\vskip 1em
+together with my co-author Shravan Vasishth, I would like to submit our
+revision of the article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.
+We are grateful for thoughtful and constructive comments and suggestions from you and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
+
+You noted in your decision letter that the manuscript could potentially be a valuable contribution to the special issue.
+This view was largely shared by the Reviewers.
+Reviewer 1 was concerned with some decisions that we had made in our computational modeling and statistical analyses and the interpretation of the results.
+In response to the concerns about our statistical analyses, we have completely re-done the analyses of the reading times and ERP data and added analyses of the previous data.
+Additionally, we have added one computational model and use k-fold cross-validation for model comparisons.
+We believe that these new analyses strengthen our initial interpretation that our study shows semantic interference, but no syntactic interference and no interaction.
+
+Reviewer 2 primarily raised concerns regarding our experimental design.
+In response to the concerns about our experimental design, we have emphasized that we used a well-established design.
+This point is further underlined by the direct comparison of the posteriors and Bayes factors from the present and previous studies using the same design.
+
+Reviewer 3 suggested a more in-depth discussion of the mechanisms of cue-based retrieval and had some concerns about our methodology.
+We have added a more detailed discussion of the memory mechanisms which are assumed to underlie sentence processing in the present design.
+We have added a thorough discussion of all concerns which were raised regarding our methodology.
+
+Below, we separately address each comment from the reviewers. All changes have been highlighted
+in the revised manuscript.
+\vskip 1em
+\noindent Sincerely,
+\vskip 1em
+\noindent Pia Schoknecht\\
+Postdoctoral researcher\\
+Department of Linguistics\\
+University of Potsdam, Germany
+\thispagestyle{empty}
+\newpage
 
 \section*{Editor's comments} 
-
+\subsection*{Comment E.1}
 \begin{quote}
@@
 
-First, \textbf{Reviewer 2 raises a potentially serious concern that your particular manipulation of syntactic dis/similarity might not be expected to lead to substantial syntactic interference (their point 1)}\marginpar{E.1}.
+First, \textbf{Reviewer 2 raises a potentially serious concern that your particular manipulation of syntactic dis/similarity might not be expected to lead to substantial syntactic interference (their point 1)}.
 In other words, the questions are whether one would ever expect to see much of this effect with your design and whether a more appropriate design would lead to stronger findings.
-Some of Reviewer 3's points (see their paragraph starting 'The authors attribute ...' and their final summary comment) are perhaps in the same vein.
+Some of Reviewer 3's points (see their paragraph starting 'The authors attribute ...' and their final summary comment) are perhaps in the same vein.''
+\end{quote}
 
-Second, a couple of Reviewer 1's comments (e.g., their point 2, and their final point about priors) relate to the fact that \textbf{there actually is some evidence for a degree of syntactic interference in aspects of the data, particularly perhaps in terms of the interaction and the combination with high semantic similarity, that you risk underplaying}\marginpar{E.2} (see also Reviewer 2's point 5).
+\subsection*{Response to E.1}
+We used a well-studied design and the previous studies using the design, specifically the same syntactic retrieval cue, found effects of syntactic interference \citep{vandyke07, mertzen}.
+Therefore, while some researchers might not expect the design to lead to decisive syntactic interference effects because the syntactic manipulation might be flawed, others might be convinced that the design is very appropriate to investigate syntactic interference, because it produced such effects in the past.
+We believe that our results add a valuable contribution to the literature on syntactic interference as a counterpoint to the previous studies \citep{vandyke07, mertzen}.
 
-Finally, I would also add that while I agree with Reviewer 1's assessment that your paper would fit well in the special issue, \textbf{as someone more versed in the memory than the language literature I thought you could be more explicit about how your findings do integrate both of these two areas.}\marginpar{E.3}
+We added the following text to our General Discussion on page \pageref{only_this_design} to emphasize that our findings and conclusions only apply to the present design:
 
-The reviewers have made suggestions which the Editors feel will improve your manuscript. We encourage you to consider these comments and make an appropriate revision of your manuscript.
+\Paste{only_this_design}
+
+\subsection*{Comment E.2}
+
+\begin{quote}
+``Second, a couple of Reviewer 1's comments (e.g., their point 2, and their final point about priors) relate to the fact that \textbf{there actually is some evidence for a degree of syntactic interference in aspects of the data, particularly perhaps in terms of the interaction and the combination with high semantic similarity, that you risk underplaying} (see also Reviewer 2's point 5).''
+\end{quote}
+
+\subsection*{Response to E.2}
+We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide evidence for syntactic interference or the interaction.
+
+As an overview of the revised Bayes factor analyses, we here include the Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs}.
+
+\setcounter{figure}{4}
+\Paste{fig_bf_spr}
+
+\setcounter{figure}{10}
+\Paste{fig_bf_eeg}
+
+\subsection*{Comment E.3}
+\begin{quote}
+``Finally, I would also add that while I agree with Reviewer 1's assessment that your paper would fit well in the special issue, \textbf{as someone more versed in the memory than the language literature I thought you could be more explicit about how your findings do integrate both of these two areas.}''
+\end{quote}
+
+\subsection*{Response to E.3}
+\textcolor{red}{??}
+
+\subsection*{Comment E.4}
+\begin{quote}
+``The reviewers have made suggestions which the Editors feel will improve your manuscript. We encourage you to consider these comments and make an appropriate revision of your manuscript.
 
 We are trying to improve the reproducibility of work published in the Journal.
-\textbf{Please ensure that you have included a "Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code} \marginpar{E.4}(see our 'guide for authors' for instructions).
+\textbf{Please ensure that you have included a ``Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code}(see our 'guide for authors' for instructions).
 Please make sure that you have updated the data and analysis code following any revisions to your manuscript.
@@
 
-\subsection*{Response to E.1}
-We address this concern in our response to Reviewer \#2 (2.\textcolor{red}{X}).
-\subsection*{Response to E.2}
-We address this concern in our response to Reviewer \#1 (1.\textcolor{red}{X}).
-\subsection*{Response to E.3}
-\textcolor{red}{write more about memory during sentence processing}
 \subsection*{Response to E.4}
-We have moved our data availability statement to page \pageref{statement}, before Experiments 1a and 1b.
+We have moved our data availability statement to page \pageref{statement}, before Experiments 1a and 1b:
+
+\Paste{osf}
 
@@
 
-\textcolor{red}{Remove SPR modeling from manuscript}
-
 \subsection*{Comment R1.2}
@@
 \subsubsection*{Response to R1.2}
-We have changed the scale of the x-axis in Figure 11 to improve readability. For convenience, we paste the Figure here:
+We have re-done the Figure to reflect our revised modeling. We hope that the scale of the new Figure is easier to read:
 
@@
 In the revised manuscript, we are now more explicit about the comparison to previous findings.
-We added a new figure (Figure \textcolor{red}{X}) to the General Discussion showing the estimates of the previous studies along with our own estimates.
-Additionally, we computed Bayes Factors for the two previous studies.
+We have added a new figure (Figure \ref{fig:previous_vs_present}) to the General Discussion.
+The new figure shows the estimates of the previous studies and Bayes factors which we have computed for the previous studies, along with our own estimates and Bayes factors.
+The following was added to the manuscript:
+\setcounter{figure}{11}
 
-Because the Van Dyke (2007) data was not available to us, we have used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) reported in the Van Dyke (2007) paper.
-Then using these posteriors and a reasonable prior (a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors using the Savage-Dickey method.
-To maximize comparability between studies, we have recomputed the Bayes Factors of our study with the Savage-Dickey method (previously we had used bridgesampling).
+\Paste{comparison}
 
@@
 We ran the suggested plausibility judgement experiment.
-The -animate distractor is indeed slightly more plausible in the context than the +animate distractor (\textcolor{red}{add numbers}).
-We discuss this potential confound in the revised manuscript.
-However, we carried out additional analyses which included plausibility as a co-predictor and find that the evidence is still strong for semantic interference.
+The -animate distractor is indeed slightly more plausible in the context than the +animate distractor.
+We discuss this potential confound in the revised manuscript starting on page \pageref{plausib_anim_inan}:
+
+\Paste{plausib_anim_inan}
 
@@
 As mentioned in our response to R1.4, we recomputed our Bayes Factors using the Savage-Dickey method.
-With this method we were able to compute Bayes Factors for models with complex random effect structures.
-All results in the manuscript have been updated accordingly.
+With this method, we were able to compute Bayes Factors for models with complex random effect structures.
+We have re-run all models with full random effect structures and updated all results in the manuscript accordingly.
 
@@
 The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}).
-Because such large interference effects have been found previously (see also our new Figure \textcolor{red}{X on page X} showing all estimates of \cite{vandyke07} and \cite{mertzen}), we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from 8 to 80 ms for the reading times.
+Because such large interference effects have been found previously (see also our new Figure 12 showing all estimates of \cite{vandyke07} and \cite{mertzen}), we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times.
 
@@
 
-\textcolor{red}{add text from the SPR results after new analyses are finished}
+\setcounter{figure}{4}
+\Paste{SPR_bfs}
 
 For the analysis of the ERP data, we changed the largest prior from N(0,5) to N(0,2), so from effects ranging up to 11 $\mu V$ to effects ranging up to 4 $\mu V$, to accommodate the concern that we might focus on too large effect sizes.
+We have revised the description of the BF analysis for the ERP data as follows (see page \pageref{ERP_results3}):
 
-We revised the description of the BF analysis for the ERP data as follows (see page \textcolor{red}{X}):
-
-\textcolor{red}{add description of ERP BF analysis}
+\Paste{ERP_results3}
 
@@
 However, \textbf{this implies that the syntactic manipulation the authors went for is just not doing the job the authors would like it to do: the syntactic distractor is not a syntactically relevant noun.
-Being a subject of whichever clause at whichever level of embedding is not a sensible manipulation to test syntactic interference.}''
-\end{quote}
+Being a subject of whichever clause at whichever level of embedding is not a sensible manipulation to test syntactic interference.}
 
-\subsubsection*{Response to R2.1}
-We used a well-studied design and the previous studies using the design, specifically the same syntactic retrieval cue, found effects of syntactic interference \citep{vandyke07, mertzen}.
-Therefore, while our results are not surprising for some researchers because the syntactic manipulation might be flawed, others might be convinced that the design is very appropriate to investigate syntactic interference, because it produced such effects in the past.
-We believe that our results add a valuable contribution to the literature on syntactic interference as a counterpoint to the previous studies.
-
-\textcolor{red}{We added the following text to the manuscript to highlight that our conclusions only apply to the design used in the present study:}
-
-\subsection*{Comment R2.2}
-\begin{quote}
-``2.
+2.
 The syntactic manipulation does not lead to minimally different conditions (i.e., conditions that only differ on the critical manipulation).
@@
 In this case, not only is the number of letters different (as acknowledged by the authors themselves), but even the number of syllables isn't the same.
-Once more, this is not a minimal comparison.
+Once more, this is not a minimal comparison.''
+\end{quote}
 
-4.
+\subsubsection*{Response to R2.1}
+We believe that we have addressed this concern in our Response to E.1.
+
+\subsection*{Comment R2.2}
+\begin{quote}
+``4.
 The difference in RTs starts at the distractor and continues through the critical verb region (in the high syntactic interference condition at least).
@@
 \subsubsection*{Response to R2.2}
-We agree with this statement and therefore refrained from interpreting the results at the critical region.\\
-\textcolor{red}{Maybe add sth to manuscript}
+We agree with this statement and therefore refrained from interpreting the results at the critical region.
+To emphasize this point, we added the following to the manuscript on page \pageref{only_precritical}:\\
+
+\Paste{only_precritical}
 
@@
 
+\newpage
+
 \subsection*{Comment R2.4}
@@
 \subsubsection*{Response to R2.4}
-\textcolor{red}{Emphasize BF results of the interaction}
+We agree with the Reviewer that visual inspection of Figure \ref{fig:whole_sentence} suggests an interaction, but the Bayes factor analysis, i.e., hypothesis testing, provided no evidence for the interaction.
+See the following text in the manuscript on page \pageref{insig_interaction}:
+
+\Paste{insig_interaction}
+
+We are not discussing the small interaction in the critical region because we refrained from interpreting effects after the pre-critical region due to pre-critical effects (see Response to R2.2).
 
@@
 These new analyses provide evidence against an effect of syntactic interference on the P600 and moderate bordering weak evidence for an effect of semantic interference.
-We added the following text to the Results and Discussion.
+We added the following text to the Results on page \pageref{ERP_results1}:
 
-\textcolor{red}{add p600 results and discussion}
+\Paste{ERP_results1}
 
+\noindent We added the following text to the Results on page \pageref{ERP_results2}:
+
+\Paste{ERP_results2}
+
+\noindent We added the following text to the Results on page \pageref{ERP_results3}:
+
+\Paste{ERP_results3}
+
+\noindent We added the following text to the Discussion on page \pageref{ERP_discussion}:
+
+\Paste{ERP_discussion}
+
+\subsection*{Comment R2.6}
 \begin{quote}
@@
 
-\textcolor{red}{add references}
+We have added the suggested references to page \pageref{franck_refs}:
+
+\Paste{franck_refs}
 
@@
 To account for practice effects, we have added the centered trial id as a fixed effect as well as by-participant and by-item random slopes to the models analyzing the SPR data.
-\textcolor{red}{The results did not change considerably.}
+The result pattern did not change (no syntactic interference or interaction, but a semantic interference effect).
 
@@
 \subsubsection*{Response to R3.7}
-We have added the following text addressing this concern to page \pageref{syn_confounds} of the manuscript:
+We have added the following text addressing the concern regarding the commas to page \pageref{comma} of the manuscript:
 
-\Paste{syn_confounds}
+\Paste{comma}
+
+\noindent Regarding the line break, we have added the following text to page \pageref{linebreak}:
+
+\Paste{linebreak}
 
@@
 Nevertheless, their data showed syntactic interference and our data do not.
-Apart from the reasons that we discuss in the manuscript (cross-linguistic differences, power, partially different methods), it is unclear why the results differ between the studies.
+We have extended the discussion of the potential reasons for the differences in results in the manuscript (cross-linguistic differences, different methods, differences in the items, statistical power, inter-individual differences, see below).
+it is unclear why the results differ between the studies.
 

### 2024-07-25 code_data/LV05_modeling/retrieval_times/generate_priorpredictions_2models.R
Deleted via the Open Science Framework

@@ -1,233 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-#### THREE CUE MODEL FROM MERTZEN ET AL. (2023) https://osf.io/a7cg2/?view_only= ####
-
-# Load model 
-source("interACT-threecues.R")
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
-  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(0.33,0.33,0.33))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Generate 20,000 simulations from model:
-nsims<-20000
-
-## Initial hyperparameters for latency factor
-## from Methodsx paper:
-a<-4
-b<-6
-lfs<-rbeta(nsims,a,b)
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-# empty df for saving means
-generated_means <- data.frame(matrix(ncol=9,nrow=nsims))
-colnames(generated_means) <- c("sample.id","RT_LoSynLoSem","RT_LoSynHiSem","RT_HiSynLoSem","RT_HiSynHiSem", "Acc_LoSynLoSem","Acc_LoSynHiSem","Acc_HiSynLoSem","Acc_HiSynHiSem")
-
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
-  lf<<- lfs[i]  
-  sims <- create_param_matrix(model_4cond, iterations=5000)
-  results <- run(sims)
-  condMeans <- compute_cond_means(results)
-  # save means and acc
[…]

### 2024-07-25 code_data/SPR/r_scripts/Plot_stats.R
Updated via the Open Science Framework

@@ -148,10 +148,10 @@ round(c(mean = mean(trial_eff_spill), quantile(trial_eff_spill, probs = c(.025, 
 
 ### Plot BFs ###
 
-# load region-wise BFs
+# load BFs
 load("BFs_schoknecht.Rda")
 
-# combine
+# prepare
 df.bf_schoknecht$Prior2 <- ifelse(df.bf_schoknecht$Prior == "Normal(0, 0.01)", "N(0, 0.01)", 
                         ifelse(df.bf_schoknecht$Prior == "Normal(0, 0.05)", "N(0, 0.05)", 
                               ("N(0, 0.1)")))

### 2024-07-26 paper/ResponseToReviewers.tex
minor wording changes

@@
 
-You noted in your decision letter that the manuscript could potentially be a valuable contribution to the special issue.
-This view was largely shared by the Reviewers.
+You noted in your decision letter that the manuscript could potentially be a valuable contribution to the special issue, a view largely shared by the reviewers.
+
 Reviewer 1 was concerned with some decisions that we had made in our computational modeling and statistical analyses and the interpretation of the results.
@@
 Reviewer 2 primarily raised concerns regarding our experimental design.
-In response to the concerns about our experimental design, we have emphasized that we used a well-established design.
-This point is further underlined by the direct comparison of the posteriors and Bayes factors from the present and previous studies using the same design.
+In response to the concerns about our experimental design, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textit{Journal of Memory and Language}.
+We now stress in the paper that we are building on previously published work by directly comparing the statistical inferences, using Bayes factors, from the present and previous studies that use exactly the same design.
 

### 2024-07-26 paper/ResponseToReviewers.tex
minor wording changes

@@
 \vskip 1em
-together with my co-author Shravan Vasishth, I would like to submit our
+Together with my co-author Shravan Vasishth, I would like to submit our
 revision of the article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.
@@
 
-Below, we separately address each comment from the reviewers. All changes have been highlighted
-in the revised manuscript.
+Below, we separately address each comment from the reviewers.
+All changes have been highlighted in the revised manuscript.
+The actionable parts of the comments from the reviewers and yourself are highlighted in bold.
+
 \vskip 1em

### 2024-07-26 paper/ResponseToReviewers.tex
minor wording changes to E1.1 response

@@
 \subsection*{Response to E.1}
-We used a well-studied design and the previous studies using the design, specifically the same syntactic retrieval cue, found effects of syntactic interference \citep{vandyke07, mertzen}.
-Therefore, while some researchers might not expect the design to lead to decisive syntactic interference effects because the syntactic manipulation might be flawed, others might be convinced that the design is very appropriate to investigate syntactic interference, because it produced such effects in the past.
-We believe that our results add a valuable contribution to the literature on syntactic interference as a counterpoint to the previous studies \citep{vandyke07, mertzen}.
+We have used a well-studied design that was previously published in the Journal of Memory and Language, among other places.
+The previously publisheed studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
+It is possible that the reviewer is right to not expect the design to lead to decisive syntactic interference effects because the syntactic manipulation might not be appropriate.
+However, we are building the previously published claim that this design leads to syntactic interference.
+We believe that our results are a valuable contribution to the literature on syntactic interference, and serves as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
+The reviewer's objection can and should be pursued in a future study that uses a design that the reviewer would consider appropriate, but this would require a completely new paper with a different research focus (namely, what is an appropriate syntactic manipulation that would lead to interference?).
+We now mention in the General Discussion
+that such a future investigation is needed.
 

### 2024-07-26 paper/ResponseToReviewers.tex
minor wording changes to E1.3 response
fixed opening quote

@@
 
-As an overview of the revised Bayes factor analyses, we here include the Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs}.
+As an overview of the revised Bayes factor analyses, we show below the Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs}.
 
@@
 We are trying to improve the reproducibility of work published in the Journal.
-\textbf{Please ensure that you have included a ``Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code}(see our 'guide for authors' for instructions).
+\textbf{Please ensure that you have included a ``Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code}(see our `guide for authors' for instructions).
 Please make sure that you have updated the data and analysis code following any revisions to your manuscript.

### 2024-07-26 paper/ResponseToReviewers.tex
fixed spacing

@@
 We are trying to improve the reproducibility of work published in the Journal.
-\textbf{Please ensure that you have included a ``Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code}(see our `guide for authors' for instructions).
+\textbf{Please ensure that you have included a ``Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code} (see our `guide for authors' for instructions).
 Please make sure that you have updated the data and analysis code following any revisions to your manuscript.

### 2024-07-26 code_data/LV05_modeling/brain_response/predictive-density-functions.R
Added via the Open Science Framework

@@ -1,0 +1,121 @@
+# adapted script from Himanshu Yadav
+# downloaded from https://osf.io/f4tkd/?view_only=
+
+#####################################
+# Using only mean of test data
+
+abc_elpd <- function(posterior_samples,kfold_data,tol_scaler){
+  ytest <- as.data.frame(kfold_data %>% group_by(set_id,dataset) %>% 
+                           summarise(mean_test_syntactic=mean(test.syntactic),
+                                     mean_test_semantic=mean(test.semantic),
+                                     mean_test_interaction=mean(test.interaction),
+                                     tol_syn=sd(test.syntactic)/tol_scaler,
+                                     tol_sem=sd(test.semantic)/tol_scaler,
+                                     tol_int=sd(test.interaction)/tol_scaler))
+  
+  df.lpd <- data.frame(matrix(ncol=3,nrow=nrow(ytest)))
+  colnames(df.lpd) <- c("set_id","testset","ELPD")
+  
+  
+  
+  for(k in 1:nrow(ytest)){
+    ypred_mk.syntactic <- subset(posterior_samples,set_id==ytest[k,]$set_id)$xsim.syntactic
+    ypred_mk.semantic <- subset(posterior_samples,set_id==ytest[k,]$set_id)$xsim.semantic
+    ypred_mk.interaction <- subset(posterior_samples,set_id==ytest[k,]$set_id)$xsim.interaction
+    
+    lpd_k <-   log(mean(dnorm((ytest[k,]$mean_test_syntactic-ypred_mk.syntactic),0,ytest[k,]$tol_syn)*
+                        dnorm((ytest[k,]$mean_test_semantic-ypred_mk.semantic),0,ytest[k,]$tol_sem))*
+                        dnorm((ytest[k,]$mean_test_interaction-ypred_mk.interaction),0,ytest[k,]$tol_int))
+    
+    df.lpd[k,] <- c(ytest[k,]$set_id,ytest[k,]$dataset,lpd_k)
+  }
+  df.lpd
+}
+
+
+
+abc_elpd2 <- function(posterior_samples,kfold_data,tol){
+  ytest <- as.data.frame(kfold_data %>% group_by(set_id,dataset) %>% 
[…]

### 2024-07-26 code_data/LV05_modeling/brain_response/model_comparison.R
Added via the Open Science Framework

@@ -1,0 +1,136 @@
+# author: Pia Schoknecht
+# date: 23.07.2024
+
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+library(tidyverse)
+
+# load posterior predictions from all 3 models
+
+load("posterior_predictions_3cues_kfold.Rda")
+load("posterior_predictions_2cues_kfold.Rda")
+load("posterior_predictions_1cue_kfold.Rda")
+
+unique(cues3_postpreds_wo[[1]]$lfs)
+unique(cues3_postpreds_wo[[2]]$lfs)
+unique(cue1_postpreds_wo[[2]]$lfs)
+unique(cue1_postpreds_wo[[1]]$lfs)
+
+# prepare data
+
+# combine all kfold sets
+cues2_postpreds_wo[[1]]$set_id <- 1
+cues2_postpreds_wo[[2]]$set_id <- 2
+cues2_postpreds_wo[[3]]$set_id <- 3
+cues2_postpreds_wo[[4]]$set_id <- 4
+cues2_postpreds_wo[[5]]$set_id <- 5
+cues2_postpreds_wo[[6]]$set_id <- 6
+cues2_postpreds_wo[[7]]$set_id <- 7
+cues2_postpreds_wo[[8]]$set_id <- 8
+cues2_postpreds_wo[[9]]$set_id <- 9
+cues2_postpreds_wo[[10]]$set_id <- 10
+
+cues2_postpreds_wo[[9]]$heldout <- 9
+cues2_postpreds_wo[[10]]$heldout <- 10
+
+postpreds2 <- rbind(cues2_postpreds_wo[[1]], cues2_postpreds_wo[[2]], cues2_postpreds_wo[[3]], cues2_postpreds_wo[[4]], cues2_postpreds_wo[[5]],
+                    cues2_postpreds_wo[[6]], cues2_postpreds_wo[[7]], cues2_postpreds_wo[[8]],cues2_postpreds_wo[[9]], cues2_postpreds_wo[[10]])
+
+
+#rename columns and add dataset column
+postpreds2 <- postpreds2 %>% rename(test.syntactic = "syntactic", test.semantic = "semantic", test.interaction = "interaction")
+postpreds2$dataset <- 1
+
+# save complete data (no heldout data) as test_data and create the same columns as in the postpreds df
+test_data <- cues2_postpreds_wo[[11]]
[…]

### 2024-07-26 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 We have used a well-studied design that was previously published in the Journal of Memory and Language, among other places.
-The previously publisheed studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
+The previously published studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
 It is possible that the reviewer is right to not expect the design to lead to decisive syntactic interference effects because the syntactic manipulation might not be appropriate.
-However, we are building the previously published claim that this design leads to syntactic interference.
-We believe that our results are a valuable contribution to the literature on syntactic interference, and serves as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
+However, we are building on the previously published claim that this design leads to syntactic interference.
+We believe that our results are a valuable contribution to the literature on syntactic interference, and serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
 The reviewer's objection can and should be pursued in a future study that uses a design that the reviewer would consider appropriate, but this would require a completely new paper with a different research focus (namely, what is an appropriate syntactic manipulation that would lead to interference?).
-We now mention in the General Discussion
-that such a future investigation is needed.
+We now mention in the General Discussion on page \pageref{future_work}
+that such a future investigation is needed:
+
+\Paste{future_work}
 
@@
 We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide evidence for syntactic interference or the interaction.
-
 As an overview of the revised Bayes factor analyses, we show below the Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs}.
+Figure \ref{fig:spr_bfs} shows the Bayes factor analysis of the self-paced reading data.
+There was evidence against syntactic interference under all priors in the pre-critical and critical region.
+Only in the spill-over region, there was anecdotal evidence (BF$_{10}$ = 2.8) evidence for a small syntactic interference effect ($\pm$ 8 ms).
+Similarly, there was evidence against the interaction under all priors in the pre-critical and spill-over regions.
+Only in the critical region, there was anecdotal evidence (BF$_{10}$ = 1.4) for a small interaction effect ($\pm$ 8 ms).
 
@@
 \Paste{fig_bf_spr}
+
+Figure \ref{fig:eeg_bfs} shows the Bayes factor analysis of the ERP data.
+There was anecdotal evidence for a very small syntactic interference effect ($\pm$ 0.2\,$\mu V$) on the N400 and evidence against larger effects ($>$ 1\,$\mu V$).
+There was evidence against an interaction under all priors in the N400 spatio-temporal window.
+In a new analysis of the P600 spatio-temporal window, there was anecdotal evidence for a very small interaction ($\pm$ 0.2\,$\mu V$) and evidence against larger interaction effects ($>$ 1\,$\mu V$).
+There was unequivocal evidence against an effect of syntactic interference in the P600.
+
+Taken these Bayes factor results together (at best anecdotal evidence for small effects in the size of an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), we are confident that in our data, there was no evidence for an effect of considerable size of neither syntactic interference nor the interaction.
 
@@
 \subsubsection*{Response to R1.2}
-We have re-done the Figure to reflect our revised modeling. We hope that the scale of the new Figure is easier to read:
+We have re-done the Figure to reflect our revised modeling. We hope that the scale of the new Figure \ref{fig:model_predictions_ERP} is easier to read:
 
@@
 The new figure shows the estimates of the previous studies and Bayes factors which we have computed for the previous studies, along with our own estimates and Bayes factors.
-The following was added to the manuscript:
-\setcounter{figure}{11}
+The Bayes factors of the present and previous studies align well with one exception: The previous studies found evidence for syntactic interference in the pre-critical
+region, while our study did not. The following description of the estimates and Bayes factors of the present and previous studies and their interpretation was added to the manuscript:
+\setcounter{figure}{12}
 
@@
 \subsubsection*{Response to R1.6}
-We ran the suggested plausibility judgement experiment.
-The -animate distractor is indeed slightly more plausible in the context than the +animate distractor.
-We discuss this potential confound in the revised manuscript starting on page \pageref{plausib_anim_inan}:
+We ran the suggested plausibility judgement experiment. The -animate distractor is indeed slightly more plausible in the context than the +animate distractor. To investigate whether this plausibility
+difference is driving the semantic interference effect in our SPR data, we ran a Bayesian hierarchical mixed model on the reading times of the pre-critical region which included plausibility as a
+co-predictor and find that there is still evidence for semantic interference.
+We discuss the plausibility confound and the additional analyses in the revised manuscript starting on page \pageref{plausib_anim_inan}:
 
@@
 We have re-run all models with full random effect structures and updated all results in the manuscript accordingly.
+The results did not change qualitatively.
 
@@
 We believe that we have addressed this concern in our Response to E.1.
+We summarize our response here: We have used a well-studied design that was previously published in a couple of journals, e.g., JEP:LMC.
+The previously published studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
+While the syntactic manipulation in the present design might not be appropriate, it was previously claimed in publications that this design leads to syntactic interference.
+Our results are a valuable contribution to the literature on syntactic interference as they serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
+The question which exact syntactic manipulation leads to interference could be the research focus of future work, but it is outside of the scope of the present manuscript.
 
@@
 \subsubsection*{Response to R2.3}
-The participants were able to discern the two syntactic conditions at region 8 because this is the region were the conditions started to differ.
-We have added the words of an example item as x-axis labels to Figure \ref{fig:whole_sentence} on page \pageref{fig:whole_sentence} to illustrate clearly.
+The participants were able to discern the two syntactic conditions at region 8 because this is the region were the conditions started to differ (`\dots der erzählt hatte \dots' vs.\ `\dots der von dem \dots').
+We have added the words of an example item as x-axis labels to Figure \ref{fig:whole_sentence} on page \pageref{fig:whole_sentence} to illustrate this clearly.
 For convenience, we paste the Figure here:
@@
 We agree with the Reviewer that visual inspection of Figure \ref{fig:whole_sentence} suggests an interaction, but the Bayes factor analysis, i.e., hypothesis testing, provided no evidence for the interaction.
+Therefore, we do not discuss this visual impression.
 See the following text in the manuscript on page \pageref{insig_interaction}:
@@
 These new analyses provide evidence against an effect of syntactic interference on the P600 and moderate bordering weak evidence for an effect of semantic interference.
-We added the following text to the Results on page \pageref{ERP_results1}:
+We added the following text describing the estimates to the Results on page \pageref{ERP_results1}:
 
@@
 
-\noindent We added the following text to the Results on page \pageref{ERP_results2}:
+\noindent We added the following text describing the ERPs for high vs.\ low syntactic interference, which show no difference, to the Results on page \pageref{ERP_results2}:
 
@@
 
-\noindent We added the following text to the Results on page \pageref{ERP_results3}:
+\noindent We added the following description of the Bayes factor analysis of the ERP data to the Results on page \pageref{ERP_results3}:
 
@@
 
-\noindent We added the following text to the Discussion on page \pageref{ERP_discussion}:
+\noindent We added the following summary of the results to the Discussion on page \pageref{ERP_discussion}:
 
@@
 
-We have added the suggested references to page \pageref{franck_refs}:
+We agree that this line of work is highly relevant for our paper.
+We have added the suggested references (and some more by the same and other authors) to the introduction (page) \pageref{franck_refs} and General Discussion (page \pageref{future_work}):
 
 \Paste{franck_refs}
+
+\Paste{future_work}
 
@@
 \subsubsection*{Response to R3.4}
-We think that we have addressed the concerns regarding high- vs.\ lower-powered studies in our Response to R1.5.
+As described in our Response to R1.4, we have added a detailed comparison of our results to the previous results.
+The main difference is that our data showed no evidence for syntactic interference in the pre-critical region, while the previous studies did show support for it.
+We have created a new subsection in our General Discussion providing an extended discussion of the potential reasons for the differences between our and previous results.
+This discussion did not lead to a definite answer.
+We provide the relevant new text passages in our responses below (Responses to R3.7 and R3.8).
 
-We have created a new subsection in our General Discussion providing an extended discussion of the potential reasons for the differences between our and previous results.
-We provide the relevant new text passages in our responses below (Response to R3.7 and R3.8).
+We think that we have addressed the concerns regarding high- vs.\ lower-powered studies and contradictory results between studies in our Response to R1.5.
+We do not claim that our results are correct and previous claims have been false.
+Progress in science is achieved by carrying out adequately powered studies in the future that can
+distinguish between theoretical claims.
+The replication crisis \citep{openscience2015_reproducibility} has shown that contradiction between results even when using the same design is not an anomaly; if anything, it's quite the opposite.
 
@@
 \subsubsection*{Response to R3.6}
-We have added the following regarding the filler items to the Procedure section of Experiment 1, page \pageref{fillers1}:
+We have added the following description of the filler items to the Procedure section of Experiment 1, page \pageref{fillers1}:
 

### 2024-07-26 paper/main.tex
Updated via the Open Science Framework

@@
 The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference.
-\textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a reduced P600 amplitude for high vs.\ low semantic interference.
-See Figure X for the topographic distribution of P600 effect.} \Copy{ERP_results2}{In contrast, Figure \ref{fig:erp_sem_syn} (c) and (d) show that syntactic interference did not considerably affect the brain response to the critical verb in \textcolor{blue}{neither the N400 nor P600 spatio-temporal windows}.}\label{ERP_results2} Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (c) does also not suggest that there was a syntactic interference effect in any other time window.
+\textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced P600 amplitude for high vs.\ low semantic interference.
+See Figure \ref{fig:erp_sem_syn} (c) for the topographic distribution of this effect.} \Copy{ERP_results2}{In contrast, Figure \ref{fig:erp_sem_syn} (d), (e) and (f) show that syntactic interference did not considerably affect the brain response to the critical verb in \textcolor{blue}{neither the N400 nor P600 spatio-temporal windows}.}\label{ERP_results2} Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (d) does also not suggest that there was a syntactic interference effect in any other time window.
 
@@
 (a) ERPs for semantic interference (word onset at 0\,ms) at electrode Cz.
-(b) Topographic map of the semantic interference effect (high semantic interference - low semantic interference) in the N400 time window.
-(c) ERPs for syntactic interference  (word onset at 0\,ms) at electrode Cz.
-(d) Topographic map of the syntactic interference effect (high syntactic interference - low syntactic interference) in the N400 time window.}
+Topographic maps of the semantic interference effect (high semantic interference - low semantic interference) in the N400 time window (b) and P600 time window (c).
+(d) ERPs for syntactic interference  (word onset at 0\,ms) at electrode Cz.
+Topographic maps of the syntactic interference effect (high syntactic interference - low syntactic interference) in the N400 time window (e) and P600 time window (f).}
     \label{fig:erp_sem_syn}
-    \centering
-         \begin{subfigure}[t]{0.74\textwidth}
-         \centering
-        \caption{}
-         \label{fig:erp_sem_ERP}
-         \includegraphics[width=\textwidth]{images/ERP_animacy_Cz.png}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[t]{0.25\textwidth}
-         \centering
-        \caption{}
-         \label{fig:erp_sem_topo}
-         \includegraphics[width=\textwidth]{images/topo_animate-inanimate_300_500.png}
-     \end{subfigure}
-
-    \begin{subfigure}[t]{0.74\textwidth}
-         \centering
-        \caption{}
-         \label{fig:erp_syn_ERP}
-         \includegraphics[width=\textwidth]{images/ERP_subj_Cz.png}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[t]{0.25\textwidth}
-         \centering
-        \caption{}
-         \label{fig:erp_syn_topo}
-         \includegraphics[width=\textwidth]{images/topo_subj-nsubj_300_500.png}
-     \end{subfigure}
+\begin{minipage}{.74\linewidth}
+\centering
+\subfloat[]{\label{fig:erp_sem_ERP}
+\includegraphics[scale=.62]{images/ERP_animacy_Cz.png}}
+\end{minipage}%
+\begin{minipage}{.25\linewidth}
+\centering
+\subfloat[]{\label{fig:erp_sem_topo}\includegraphics[scale=.65]{images/topo_animate-inanimate_300_500.png}}
+\par\medskip
+\centering
+\subfloat[]{\label{fig:erp_sem_topo_p600}\includegraphics[scale=.35]{images/topo_animate-inanimate_600_900.png}}
+\end{minipage}
+\begin{minipage}{.74\linewidth}
+\centering
+\subfloat[]{\label{fig:erp_syn_ERP}
+\includegraphics[scale=.62]{images/ERP_subj_Cz.png}}
+\end{minipage}%
+\begin{minipage}{.25\linewidth}
+\centering
+\subfloat[]{\label{fig:erp_syn_topo}\includegraphics[scale=.65]{images/topo_subj-nsubj_300_500.png}}
+\par\medskip
+\centering
+\subfloat[]{\label{fig:erp_syn_topo_p600}\includegraphics[scale=.35]{images/topo_subj-nsubj_600_900.png}}
+\end{minipage}
 \end{figure}
@@
 \begin{figure}[htpb]
-    \caption{Prior and posterior predicted effect sizes on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval mode.}
+    \caption{Prior and posterior predicted effect sizes on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
     %using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. 
@@
 In the critical region, four of the investigated experiments show either weak evidence for or weak evidence against syntactic interference, but there are two outliers: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extreme evidence for the effect and the present study provided strong evidence against it.
-In conclusion, the most remarkable difference between the present study and previous findings concerns syntactic interference in the pre-critical region.}}
+In conclusion, the most remarkable difference between the present study and previous findings concerns syntactic interference in the pre-critical region.}
+}
 
@@
 The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
+\Copy{future_work}{\textcolor{blue}{A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
+This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.}}\label{future_work}
+
 However, the semantic interference effect we found can be easily reconciled with previous work in  sentence processing.

### 2024-07-26 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \subsubsection*{Response to R1.5}
-\textcolor{red}{Write something about the co-existence of contradictory results}
+About why our results diverge from previous studies' and whether we
+think our results are correct and the previous work is not, we have no
+position on this. We are simply reporting what we found. Contradictory
+results have been reported for many phenomena: e.g., reflexives
+processing, main verb/reduced relative garden paths, predictive
+parsing effects, local coherence \textcolor{red}{add refs}. The
+only way to resolve these many contradictions in the literature is by
+carrying out adequately powered studies in the future that can
+distinguish between the theoretical claims.  Our work does not stand
+out as an anomaly; if anything, it's quite the opposite.
+
+\textcolor{red}{add something to paper}
 

### 2024-07-31 paper/main.tex
Updated via the Open Science Framework

@@
     \includegraphics[width=0.85\textwidth]{images/Pandora_all_wholesentence_pooled_zoom_exp.jpg}
-\end{sidewaysfigure}}
-\clearpage
+\end{sidewaysfigure}
+\clearpage}
 
@@
 
-\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times.
-Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized interference effects
-for both the syntactic and semantic manipulations at the critical verb.
-Since we have found that the reading times at the critical verb, i.e., the retrieval site, were confounded by pre-critical effects, which were best explained by \textit{encoding} interference, it makes no sense to compare the results from the present SPR experiment to predictions from the Lewis and Vasishth (2005) model of sentence processing
-as skilled memory retrieval.
-Instead we use the Lewis and Vasishth (2005) model to generate predictions on the microvolt scale for ERP data by adjusting the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.}
-
-The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+\Copy{modeling}{\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times.
+Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
+We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference.
+Therefore, it makes no sense to compare the results from the present SPR experiment to predictions from the Lewis and Vasishth (2005) model of sentence processing
+as skilled memory \textit{retrieval}.
+Instead we use the Lewis and Vasishth (2005) model to generate predictions on the microvolt scale for ERP data.
+For this we adjust the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.} The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
 We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$).
@@
 To reflect that larger predicted effect sizes correspond to more negative N400 amplitudes, the predicted effects from the model} were multiplied with $-1$.
-To avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects.
+To avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects \textcolor{blue}{and their interaction.}
 
 \textcolor{blue}{
-For their recent simulations, \textcite{mertzen} used a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ grammatical subject\}, \{$\pm$ same clause\} and \{$\pm$ animacy\}.
-Although it is clear that our data does not support equally sized effects of syntactic and semantic interference, we decided to use their version of the model for our modeling because it has been established as a good model for the present design.
+For their recent simulations, \textcite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}.
+The \{$\pm$ same clause\} cue matches only with nouns that occurred in the same clause as the verb.
+``The addition of the \{$\pm$ same clause\} cue [was] necessary to identify the correct subject'' and as a means for the model to ``to achieve correct retrieval'' in the condition with high syntactic and high semantic interference \citep[][p.
+14]{mertzen}.
+This model predicted approximately equally sized effects of syntactic and semantic interference.
+Although it is clear that our data does not support equally sized effects of syntactic and semantic interference, we decided to include their version of the model here because it has previously been claimed to be a good model for the present design.
 We have implemented two alternative versions of the model.
-A model which puts more emphasis on hierarchical syntactic structure.
 Given the evidence against syntactic interference in our data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
-We speculate that this could be due to the parser tracking} hierarchical structure and that this hierarchical information is used when searching for a noun in memory.
-Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} \textcolor{blue}{-- as implemented by \textcite{mertzen}} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
-\textcolor{blue}{This idea actually goes back to the roots of the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005}.
+We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.
+So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} \textcolor{blue}{-- as implemented by \textcite{mertzen}} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
+\textcolor{blue}{This idea goes back to the roots of the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005}.
 The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.} This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
-In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
-\textcolor{blue}{Lastly, we implemented a model with one cue which unambiguously matches the subject.
+In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.\textcolor{blue}{This model using two cues predicts no syntactic interference and no interaction, but semantic interference effects.
+Lastly, we implemented a model with one cue which unambiguously matches the subject.
 This cue would be the master composite cue \{$\pm$ animate-subject-in-same-clause\}.
 This model naturally does not predict any interference effects.
-}
-
-It is informative to fit the models to the data and then examine the posterior predictive distribution of the \textcolor{blue}{ERPs} for each effect. 
+It can be regarded the null model or no-interference model.}
+
+It is informative to fit the model\textcolor{blue}{s} to the data and then examine the posterior predictive distribution of the \textcolor{blue}{ERPs} for each effect. 
 We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the \textcolor{blue}{ERP} data.
+To avoid overfitting, we fitted $F$ only to the semantic interference effect (and not the syntactic interference effect or interaction).
 This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ in the \textcolor{blue}{ERP} data.
 `Close' is defined as $d \pm \,2\times SE$ ms; for details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
-The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}.
+
+\textcolor{blue}{Out of 100,000 iterations of ABC algorithm, we obtained only 6467 samples of $F$ for the 3-cues model, 1010 for the 2-cues model and 869 for the 1-cue model.
+This is not surprising given
+the relatively wide prior on $F$.
+In future work, the posterior distribution, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.
+Figure \ref{fig:model_predictions_ERP} shows the  posterior predictive distributions
+of the effects of interest in comparison to the prior predictive distribution and our ERP
+data.
+The prior and posterior predictions from the model with three cues show equally large effects of syntactic and semantic interference and an interaction.
+In contrast, the model using two cues, including a syntactic cue that utilizes syntactic hierarchy, predicts no effect of syntactic interference and no interaction, but a semantic interference effect.
+The model using one cue predicts no interference effects.
+The prior and posterior predictions of all models overlap with the effects of interest in the data.
+To evaluate the predictive fit of the three models to our data quantitatively, we used the likelihood-free k-fold cross-validation model comparison approach developed by \textcite{Yadavetal2022}.}
 
 \Copy{figpriorposteriordata}{
-\begin{figure}[htpb]
+\begin{sidewaysfigure}[htpb]
     \caption{Prior and posterior predicted effect sizes on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
@@
          \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
+\end{sidewaysfigure}
+}
+
+\textcolor{blue}{For this k-fold cross-validation model comparison, we repeatedly computed how well each of the three models performed on a held-out small subset of the data after being trained on the remaining large subset of the data.
+We repeatedly left out 1/10$^{th}$ of the data which was then used as the test data, while the other 9/10$^{th}$ functioned as the training data.
+It was made sure that each of the 10 subsets included data from all participants, all items and all conditions.
+The models were trained by generating posterior predictive distributions by fitting the $F$ parameter to the posterior distribution of the effects of interest in the training data.
+Then these posterior distributions of $F$ were used to compute the likelihood of generating the held-out test data.
+The log of this likelihood is the predictive accuracy of the model.
+This was done 10 times.
+Then, we computed the sum and standard error of sum of the predictive accuracies of the models of each of the training and test datasets to gain the overall predictive accuracy, the expected log predictive density $\widehat{elpd}$, of each model.
+To evaluate the performances of the models, we compare the difference in the $\widehat{elpd}$ values between each model and the other two models (see Figure \ref{fig:elpd}).
+}
+
+\begin{figure}
+    \centering
+    \caption{Comparison of the predictive accuracies of the three versions of the Lewis and Vasishth (2005) cue-based retrieval model on our ERP data.
+Each facet shows how the model in the facet's title performs in comparison to the models on the x-axis (positive $\widehat{elpd}$ differences correspond to better performance).
+Error bars show two times the standard error of the difference in $\widehat{elpd}$ values.}
+    \label{fig:elpd}
+    \includegraphics[width=\linewidth]{images/ELPD_difference.jpg}
 \end{figure}
-}
-
-\textcolor{red}{To do (Pia): Describe figure\\
-model comparison (cross-validation)}
-%For the ERP modeling, 473 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions.
-The prior and posterior predictions of the model overlap with the effects of interest in the data.
-The general effect pattern (no syntactic interference effect \textcolor{blue}{and no interaction}, but a semantic interference effect) is similar in both the prior and posterior predictive distributions.
-Therefore, the assumption that the relevant syntactic cue is the \{$\pm$ subject-in-same-clause\} cue -- rather than the previously assumed \{$\pm$ grammatical subject\} cue --  seems reasonable for the present ERP study.
-
-%We evaluated the predictions of the model assuming that the cues are \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-The results of the prior predictive simulations are shown in the lower parts of Figure \ref{fig:model_predictions} a and b.
-This model now shows the observed pattern: a semantic interference effect but no syntactic interference.
-One limitation here is that the prior predictive distribution of our model generates a very widely distributed semantic interference effect (95\,\% CrI [10, 67]\,ms; the largest effect observed the SPR study was [4, 12] ms at the critical region).
-This broad prior predictive distribution has to do with the fact that the prior on the $F$ parameter ($Beta(a=4,b=6)$) is relatively spread out.
-This leads to a relatively wide range of predicted reading times.
-In future work, the posterior distribution, with range  [4, 12] ms, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.
-
-Thus, the theoretical assumption that is consistent with the observed ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
+
+\textcolor{red}{add description of figure and conclusion of the modeling section}
+
+%Thus, the theoretical assumption that is consistent with the observed ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
 This is of course a speculative claim that would need to be further investigated empirically.
+
+}\label{modeling}%end copy
 
@@
 In sum, while there are differences in the employed methodology and items between the present and previous studies, the most likely cause of the different results is the difference in statistical power.
-However, an alternative explanation could be that syntactic interference might differ inter-individually \parencite{yadav2021individual} and that the previous studies by chance sampled disproportionally many participants who show syntactic interference, while we by chance sampled participants who do not.
-}
+However, an alternative explanation could be that syntactic interference might differ inter-individually \parencite{yadav2021individual} and that the previous studies by chance sampled disproportionally many participants who showed syntactic interference, while we by chance sampled participants who did not.}
+
+\Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature \citep[see e.g.,][for a high-powered failed replication attempt of a classic result in the area of predictive processing]{nieuwland_etal_2018}.
+The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.}}\label{not_an_anomaly}
 

### 2024-07-31 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \subsection*{Response to E.3}
-\textcolor{red}{??}
+Our work is situated in the cue-based retrieval framework which has been built over the last decades \citep{Lewis2005, nairne, mcelree}.
+Its main assumption is that sentence processing relies on skilled memory processes, with a focus on memory retrieval.
+We built on this assumption and focus on the question which specific retrieval cues are being used during subject-verb dependency processing.
+Our findings showed that the retrieval cues generated at the verb are highly specific (including hierarchical syntactic information) and can successfully identify the subject as the retrieval target.
+Therefore, the present findings further support the assumption that sentence processing relies on skilled memory retrieval.
+We have added the following text to the General Discussion on page \textcolor{red}{X} to emphasize this point:
+
+\textcolor{red}{Add something to GenDisc after modeling is done}
 
@@
 \subsubsection*{Response to R1.2}
-We have re-done the Figure to reflect our revised modeling. We hope that the scale of the new Figure \ref{fig:model_predictions_ERP} is easier to read:
-
-\Paste{figpriorposteriordata}
+We have re-done the Figure to reflect our revised modeling. We hope that the scale of the new Figure \ref{fig:model_predictions_ERP} is easier to read (see below in our Response to R1.3).
 
@@
 \subsubsection*{Response to R1.3}
-\textcolor{red}{Do k-fold crossvalidation of computational models of ERP data}
+As suggested, we have implemented a null model (no interference), leaving us with a total of three computational models (the null model, the three-cues model by \textcite{mertzen} and our newly proposed two-cues model).
+To quantitatively compare the performance of these models, we have used a likelihood-free k-fold cross-validation approach developed by \textcite{Yadavetal2022}.
+The result of this model comparison is that \textcolor{red}{XXXX}.
+We have described the new modeling and model comparisons starting page \pageref{modeling}:
+
+\Paste{modeling}
+
+\newpage
 
@@
 
+\newpage
+
 \subsection*{Comment R1.5}
@@
 results have been reported for many phenomena: e.g., reflexives
-processing, main verb/reduced relative garden paths, predictive
-parsing effects, local coherence \textcolor{red}{add refs}. The
+processing \citep{jaeger_etal_2020}, main verb/reduced relative garden paths (\textcolor{red}{REF?}), predictive
+parsing effects \citep{nieuwland_etal_2018}, local coherence \citep{lcpaape2023}. The
 only way to resolve these many contradictions in the literature is by
@@
 distinguish between the theoretical claims.  Our work does not stand
-out as an anomaly; if anything, it's quite the opposite.
+out as an anomaly; if anything, it's quite the opposite. We have added a discussion of divergent results in the literature to page \pageref{not_an_anomaly} in the General Discussion:
 
-\textcolor{red}{add something to paper}
+\Paste{not_an_anomaly}
 
@@
 \Paste{plausib_anim_inan}
+
+\newpage
 
@@
 We summarize our response here: We have used a well-studied design that was previously published in a couple of journals, e.g., JEP:LMC.
-The previously published studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
+The previously published studies used exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
 While the syntactic manipulation in the present design might not be appropriate, it was previously claimed in publications that this design leads to syntactic interference.
@@
 For convenience, we paste the Figure here:
-\newpage
+
+\setcounter{figure}{2}
+
 \Paste{figwhole_sentence}
@@
 ``This article describes two experiments that test whether comprehension of subject-verb relations in complex sentences (in German) is disrupted by the presence of irrelevant-but-tempting nearby noun phrases.
-The temptation is due either to semantic properties ("good fit to the verb") or syntactic properties ("looks like a subject").
+The temptation is due either to semantic properties (``good fit to the verb") or syntactic properties (``looks like a subject").
 The key finding from both experiments is that semantic temptation effects are clear, while syntactic temptation effects are largely absent.
-The authors interpret this contrast as evidence for specific types of features used in sentence encoding and retrieval operations that should prevent effects of syntactic interference, i.e., if encountering a verb induces retrieval of "subject of the same clause", then the subject of a different clause shouldn't interfere.
+The authors interpret this contrast as evidence for specific types of features used in sentence encoding and retrieval operations that should prevent effects of syntactic interference, i.e., if encountering a verb induces retrieval of ``subject of the same clause", then the subject of a different clause shouldn't interfere.
 
@@
 
-The authors attribute the lack of a syntactic interference effect to the use of a syntactic feature like "subject of the same clause", which has the effect that a subject of a different clause does not count as similar to the retrieval cue.
+The authors attribute the lack of a syntactic interference effect to the use of a syntactic feature like ``subject of the same clause", which has the effect that a subject of a different clause does not count as similar to the retrieval cue.
 This gives a special status to the correct subject via the choice of retrieval cues.
@@
 \begin{quote}
-``The use of a feature like "subject of the same clause" calls for some further discussion, since, taken literally, this is not a likely feature of a subject phrase in memory.
-Features like "animate" or "plural" are inherent properties of an item in memory, but "subject of the same clause" is not.
+``The use of a feature like ``subject of the same clause" calls for some further discussion, since, taken literally, this is not a likely feature of a subject phrase in memory.
+Features like ``animate" or ``plural" are inherent properties of an item in memory, but ``subject of the same clause" is not.
 \textbf{I imagine that what the authors have in mind is that comprehenders assign a clause index to each individual piece of structure as they parse a sentence.
-Then, when they encounter a verb, they could identify the current clause index, e.g., "3", and then embed that in a retrieval cue, e.g., "subject of clause 3".
+Then, when they encounter a verb, they could identify the current clause index, e.g., ``3", and then embed that in a retrieval cue, e.g., ``subject of clause 3".
 This is a fine way to introduce a limited amount of relational information into a structural encoding.
@@
 There are many situations in which verb-noun combinations are intuitively a good fit, but where it is less easily cast in terms of features and retrieval cues.
-For example, we share the intuition that "the teacher drove …" is a better combination than "the little boy drove …", because teachers are plausible drivers, whereas small children are not.
+For example, we share the intuition that ``the teacher drove …" is a better combination than ``the little boy drove …", because teachers are plausible drivers, whereas small children are not.
 This is less easily captured in terms of features and retrieval cues.
-Maybe the verb "drive" wants a subject with a [+driver] feature.
-But is this plausibly something that is part of the memory encoding of "teacher" but not "little boy"?
+Maybe the verb ``drive" wants a subject with a [+driver] feature.
+But is this plausibly something that is part of the memory encoding of "teacher" but not ``little boy"?
 Does this require something more than a toy theory of semantic encoding?

### 2024-07-31 paper/main.tex
Updated via the Open Science Framework

@@
 
-It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval. As an example, consider 
+\Copy{intro}{It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval \textcolor{blue}{\parencite[see e.g., ][]{lewis06}}.
+As an example, consider
 (\ref{ex:1_vandyke07}).
@@
 There is clear evidence in the literature  \parencite[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
-A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues.
+A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues \textcolor{blue}{\citep{Lewis2005, mcelree}}.}\label{intro}
 
@@
 
-\textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method.
-A Bayesian hypothesis test for which the posterior density is devided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$.
-\textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account.
-This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
-Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is possible to find evidence for or against an effect.
-This stands in contrast to the frequentist ANOVA which, in its standard usage, is designed to only furnish evidence against the null.
-
-Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest.
-If BF$_{10}$ < 1, it provides evidence \textcolor{blue}{against the effect of interest}.
-The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence \textcolor{blue}{against the effect of interest}.
-In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
-For this reason, models were run with four chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.
-
 \begin{table}[h]
@@
   SD & Normal(0, 0.1) &\\
+  \bottomrule
+  \end{tabular}
+  \end{table}
+
+\textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method.
+A Bayesian hypothesis test for which the posterior density is devided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$.
+\textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account.
+This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
+Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is possible to find evidence for or against an effect.
+This stands in contrast to the frequentist ANOVA which, in its standard usage, is designed to only furnish evidence against the null.
+
+Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest.
+If BF$_{10}$ < 1, it provides evidence \textcolor{blue}{against the effect of interest}.
+The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence \textcolor{blue}{against the effect of interest (see Table \ref{tab:bf_interpretation})}.
+In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
+For this reason, models were run with four chains and 20,000 iterations in each chain, with the first 2,000 iterations in each chain being discarded as the warm-up phase.
+
+\begin{table}[h]
+    \centering
+    \caption{Interpretation of Bayes factors \parencite{lee2014bayesian}.}
+    \label{tab:bf_interpretation}
+    \begin{tabular}{cc}
+    \toprule
+    BF$_{10}$ & Interpretation \\
+    \midrule
+$>$ 100 & extreme evidence for the effect\\
+30 - 100 & very strong evidence for the effect\\
+10 - 30 & strong evidence for the effect\\
+3 - 10 & moderate evidence for the effect\\
+1 - 3 &	anecdotal evidence for the effect\\
+1  &	no evidence\\
+1 - 0.3 & anecdotal evidence against the effect\\
+0.3 - 0.1 & moderate evidence against the effect\\
+0.1 - 0.03 & strong evidence against the effect\\
+0.03 - 0.001 & very strong evidence against the effect\\
+$<$ 0.001 & extreme evidence against the effect\\
 \bottomrule

### 2024-07-31 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \subsection*{Response to E.3}
-Our work is situated in the cue-based retrieval framework which has been built over the last decades \citep{Lewis2005, nairne, mcelree}.
+Our work is situated in the cue-based retrieval framework which has been built over the last decades \citep{Lewis2005, mcelree}.
 Its main assumption is that sentence processing relies on skilled memory processes, with a focus on memory retrieval.
@@
 Therefore, the present findings further support the assumption that sentence processing relies on skilled memory retrieval.
-We have added the following text to the General Discussion on page \textcolor{red}{X} to emphasize this point:
+We have emphasized the theoretical grounding of our work by adding more references to the introduction on page \pageref{intro}:
+
+\Paste{intro}
+
+Additionally, we have added the following text to the General Discussion on page \textcolor{red}{X} to emphasize this point:
 

### 2024-08-01 paper/main.tex
Updated via the Open Science Framework

@@
 We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference.
-Therefore, it makes no sense to compare the results from the present SPR experiment to predictions from the Lewis and Vasishth (2005) model of sentence processing
-as skilled memory \textit{retrieval}.
+Therefore, it makes no sense to compare the results from the present SPR experiment to predictions from the Lewis and Vasishth (2005) model of sentence processing as skilled memory \textit{retrieval}.
 Instead we use the Lewis and Vasishth (2005) model to generate predictions on the microvolt scale for ERP data.
@@
 
-\textcolor{blue}{
-For their recent simulations, \textcite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}.
+\textcolor{blue}{For their recent simulations, \textcite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}.
 The \{$\pm$ same clause\} cue matches only with nouns that occurred in the same clause as the verb.
@@
 Given the evidence against syntactic interference in our data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
-We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.
-So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} \textcolor{blue}{-- as implemented by \textcite{mertzen}} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
-\textcolor{blue}{This idea goes back to the roots of the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005}.
-The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.} This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
-In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.\textcolor{blue}{This model using two cues predicts no syntactic interference and no interaction, but semantic interference effects.
+We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
+\textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, we can easily assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}.
+This idea goes back to the roots of the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and entails a simple clause tracking mechanism.} The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
+In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
+\textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}}\label{clause_tracker} \textcolor{blue}{The here proposed model using two cues predicts no syntactic interference and no interaction, but semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.

### 2024-08-01 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-Below, we separately address each comment from the reviewers.
-All changes have been highlighted in the revised manuscript.
-The actionable parts of the comments from the reviewers and yourself are highlighted in bold.
+Below, we separately address each comment from the reviewers and yourself. All changes have been highlighted in the revised manuscript. The actionable parts of the comments are highlighted in bold.
 
@@
 \subsubsection*{Response to R3.1}
-\textcolor{red}{??}
+It is possible that the subject of the current clause receives an activation boost but the mechanism for that is not trivial.
+Let's assume that the subject is encoded with boosted activation compared to for example an object.
+As the sentence progresses the subject's activation decays (we could even assume that it decays more slowly than other memory representations due to its special status) and other nouns are processed / encoded.
+At one point within a long sentence, the subject's activation will be lower than the activation of other nouns in memory.
+If we assume that a long sentence has multiple embedded sentences, we would need to assume that either the matrix sentence subject activation is downgraded when an embedded sentence is entered or that the embedded sentence subject gets an even higher activation boost than the matrix sentence subject.
+Then when the embedded sentence is finished and we return back to the matrix sentence, the matrix sentence subject would need to receive another activation boost to return to its special status.
+Essentially, this proposal condenses down to a clause tracking mechanism with a modulation of activation levels.
+We think it is more parsimonious to implement the clause tracking in the memory encoding and retrieval operations than the activation levels.
 
@@
 \subsubsection*{Response to R3.2}
-\textcolor{red}{??}
+The Reviewer is correct, our proposal that hierarchical syntactic structure is used during retrieval relies on the general assumption that hierarchical syntactic structure is being tracked.
+We think that this can easily be represented in encodings and then be used during retrieval.
+We have added a more explicit description of this underlying mechanism to page \pageref{clause_tracker}:
+
+\Paste{clause_tracker}
 
@@
 To account for practice effects, we have added the centered trial id as a fixed effect as well as by-participant and by-item random slopes to the models analyzing the SPR data.
-The result pattern did not change (no syntactic interference or interaction, but a semantic interference effect).
+Theoretically, participants could either speed up or slow down (or do not change their reading speed) over the course of the experiment.
+Therefore, a symmetrical prior on the slope of trial id (effect could be positive, negative or zero) is the most reasonable.
+We decided to use a Normal(0, 0.1) prior on the trial id slope in all models, a priori assuming effects up to $\pm$ 81\,ms.
+For the first trial in the experiment to the last trial, participants actually sped up drastically: Reading times in the regions of interest decreased by on average around 250\,ms.
+
+It is not possible to use symmetrical priors (for trial id) and directional priors (which we had used originally for the effects of interest to reflect longer reading times in high interference conditions) within the same brms model \citep{brms}.
+So in the revised analyses, we now use symmetrical priors on all slopes which are a priori agnostic regarding the sign of the effect.
+To ensure consistency in our analyses and comparability between the analyses, we changed all our analyses (SPR and ERP data) to use symmetrical priors.
+These revised analyses have produced the same result pattern as our previous analyses (no syntactic interference or interaction, but a semantic interference effect).
 

### 2024-08-09 paper/main.tex
Updated via the Open Science Framework

@@
 Evidence from self-paced reading and event-related potentials using German}
-\authorsnames{Pia Schoknecht, Shravan Vasishth}
+
+\authorsnames[1,{2},1]{Pia Schoknecht, Himanshu Yadav, Shravan Vasishth}
+\authorsaffiliations{{Department of Linguistics, University of Potsdam},
+{Department of Cognitive Science, Indian Institute of Technology Kanpur}}
 
 \shorttitle{Syntactic and semantic interference effects}
-\affiliation{
-Department of Linguistics, University of Potsdam}
 
 \authornote{This research was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), Project ID 317633480, SFB 1287.
-We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition.
-We are grateful to Dario Paape for helpful discussions, and Himanshu Yadav for assistance with the computational modeling.
+We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition. We are grateful to Dario Paape for helpful discussions.
 Correspondence concerning this article should be addressed to Pia Schoknecht, e-mail: pia.schoknecht@uni-potsdam.de
@@
 Syntactic interference is manipulated via the syntactic status of the distractor, which is either the subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}) and hence a non-subject.
-\Copy{franck_refs}{Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase \textcolor{blue}{ -- even if the distractor occurs in an embedded clause {\parencite[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary}.}}}\label{franck_refs}
+\Copy{franck_refs}{Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase \textcolor{blue}{ -- even if the distractor occurs in an embedded clause {\parencite[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary, franck2020hierarchical}.}}}\label{franck_refs}
 
@@
 While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant.
-The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval.
+The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval (for work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \cite{vandyke_mcelree06, cunnings_sturt_2018retrieval}).
 A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
@@
 This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.
-
 Thus, our focus in the \textcolor{blue}{SPR experiment} is on the processing difficulty observed at the critical verb and the surrounding regions.
@@
 
-\textcolor{blue}{Alternately, interference -- especially syntactic interference -- might modulate the P600.
+\textcolor{blue}{Alternately, retrieval interference -- especially syntactic retrieval interference -- might modulate the P600.
 The P600 is a positive deflection in the ERP wave form that is typically observed at least 600 ms after stimulus onset.
@@
 High interference could increase difficulty during syntactic integration of subject and verb in our materials.
-Potential misretrievals of the distractor might lead to reanalysis processes which could manifest as an increased P600 amplitude compared to low interference conditions.
-To anticipate our results here, visual inspection of the ERPs in Figure \ref{fig:erp_syn_ERP} does not indicate syntactic interference effects in the P600 time window.}} \label{why_n400}
+Potential misretrievals of the distractor might lead to reanalysis which could manifest as an increased P600 amplitude in high compared to low interference conditions.}} \label{why_n400}
 
@@
 The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
-\textcolor{blue}{In addition to the predictors of interest, we included the trial id to account for potential effects of fatigue or adaptation to the task.
+\textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task.
 The trial id of the 100 trials per session was recoded to span from 0 to 1.
@@
 \textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method.
-A Bayesian hypothesis test for which the posterior density is devided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$.
+A Bayesian hypothesis test for which the posterior density is divided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$.
 \textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account.
@@
 Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
-\Copy{SPR_results}{\textcolor{blue}{The 95 \% credible intervals for syntactic interference included zero in the pre-critical and critical regions.
+\Copy{SPR_results}{\textcolor{blue}{The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions.
 In the spill-over region, it included only positive values ([1, 15] ms).
-In contrast, for semantic interference, the 95 \% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms).
-The posteriors for the interaction were mostly positive, but the 95 \% credible intervals included zero in all regions.
-The interaction was the  largest in the critical region ([-3, 31 ms]).}
+In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms).
+The posteriors for the interaction were mostly positive, but the 95\% credible intervals included zero in all regions.
+The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}
 
@@
 
-\Copy{adaptation}{\textcolor{blue}{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last, i.e., 100th, trial, reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
+\Copy{adaptation}{\textcolor{blue}{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last trial (100th), reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
 This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment.
-While this adaption is interesting, it is not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}
+While this adaption is interesting in itself, it is not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}
 
@@
 The analyses use all the data from Experiments 1a and 1b.
-\Copy{SPR_bfs}{\textcolor{blue}{We found very to extremely strong evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
+\Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
 In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5).
-There was strong evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
-Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was weak evidence for syntactic interference (BF$_{10}$ = 2.8).
+There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
+Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).
 Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
@@
 This means that semantic interference occurred even before the retrieval event could begin.
-Reading times at the critical verb showed very weak support for a small effect of a syntactic $\times$ semantic interference interaction.
-There was evidence against the interaction under broader priors and in the surrounding regions.}\Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
+Reading times at the critical verb showed very weak support for a small interaction of syntactic and semantic interference.
+There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
 Therefore, we refrain from discussing effects occurring in these regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.
@@
 
-\Copy{plausib_anim_inan}{\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
+\label{plausib_anim_inan}\Copy{plausib_anim_inan}{\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
 To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}.
@@
 Participants spent approximately 37 minutes to complete the task and received \pounds 6.50 as compensation.
-The results of an ordinal brms model analyzing the ratings of the critical items are shown in Figure \ref{fig:plausibility_anim_inanim}.}
+The results of an ordinal brms model analyzing the ratings of the critical items are shown in Figure \ref{fig:plausibility_anim_inanim}.
+Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible.}
 
@@
 
-\textcolor{blue}{Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible.
-To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
+\textcolor{blue}{To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
 This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings.
-For the slopes, we used a Normal(0,0.5) prior; the priors for all other parameters were the same as presented in Table \ref{tab:spr_priors}.
+We used the same priors as presented in Table \ref{tab:spr_priors}.
 The model was run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.
-Figure \ref{fig:posteriors_plausibility} presents the posteriors from this model.
-Although an increase in plausibility lead to faster reading times (CrI [-30, -6] ms), there was still an effect of semantic interference (CrI [2, 24] ms).
-The Bayes factor provided weak evidence for semantic interference in this analysis (BF$_{10}$ = 2.4).
+Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes.
+Although plausibility influenced reading times (higher plausibility lead to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms).
+The Bayes factors provided evidence for semantic interference in this analysis under all priors (Normal(0, 0.01): BF$_{10}$ = 4.2,
+Normal(0, 0.05): BF$_{10}$ = 2.4,
+Normal(0, 0.1): BF$_{10}$ = 1.3).
 The interaction of semantic interference and plausibility was centered around zero.
 In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
-However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference lead to slower reading times independent of plausibility.}\label{plausib_anim_inan}
+However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference lead to slower reading times independent of plausibility.}
 
@@
 }\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600.
-The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals include zero for both the N400 and the P600.
+The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals included zero for both the N400 and the P600.
 The 95\% credible interval of the interaction was centered around zero for the N400.
@@
 \centering
-\subfloat[]{\label{fig:erp_sem_topo_p600}\includegraphics[scale=.35]{images/topo_animate-inanimate_600_900.png}}
+\subfloat[]{\label{fig:erp_sem_topo_p600}\includegraphics[scale=.65]{images/topo_animate-inanimate_600_900.png}}
 \end{minipage}
@@
 \centering
-\subfloat[]{\label{fig:erp_syn_topo_p600}\includegraphics[scale=.35]{images/topo_subj-nsubj_600_900.png}}
+\subfloat[]{\label{fig:erp_syn_topo_p600}\includegraphics[scale=.65]{images/topo_subj-nsubj_600_900.png}}
 \end{minipage}
@@
 Bayes factors are shown in Figure \ref{fig:eeg_bfs}.
-\Copy{ERP_results3}{\textcolor{blue}{In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect for all priors
+\Copy{ERP_results3}{\textcolor{blue}{In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect under all priors
 (Normal(0, 0.1): BF$_{10}$ = 6,
@@
 Normal(0, 2): BF$_{10}$ = 5).
-In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and no evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$.
-By contrast, Bayes factors unequivocally provided evidence against the syntactic interference effect in both spatio-temporal windows (BF$_{10}$ $<$ 0.9).
-The only exception was under the narrowest prior (Normal(0, 0.5)) in the N400 window, there was either no evidence for syntactic interference or even evidence against it (BF$_{10}$ $<$ 0.9).
-Similarly, Bayes factors unequivocally provided evidence against the interaction in both spatio-temporal windows.
-The only exception was under the narrowest prior (Normal(0, 0.5)) in the P600 window, there was no evidence for or against the interaction (BF$_{10}$ = 1.2).
-In sum, in our ERP data, there was decisive evidence for the semantic interference effect and no evidence or even evidence against syntactic interference and the interaction of syntactic and semantic interference.}}\label{ERP_results3}
+In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and very weak evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$.
+By contrast, Bayes factors provided evidence against the syntactic interference effect in both spatio-temporal windows (BF$_{10}$ $<$ 0.9).
+The only exception was under the narrowest prior in the N400 window, there was very weak evidence for syntactic interference (BF$_{10}$ = 1.2).
+Similarly, Bayes factors provided evidence against the interaction in both spatio-temporal windows.
+The only exception was under the narrowest prior in the P600 window, there was very weak evidence for the interaction (BF$_{10}$ = 1.2).
+In sum, in our ERP data, there was decisive evidence for the semantic interference effect and mainly evidence against syntactic interference and the interaction.}}\label{ERP_results3}
 
@@
 In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference.
-Both the N400 and P600, were more negative under high vs.
+Both the N400 and P600 were more negative under high vs.
 low semantic interference.
-In contrast, there was no evidence or even evidence against effects of syntactic interference and the interaction on the N400 and P600.
-In sum, the ERP results showed clear evidence for semantic interference and no evidence or evidence against syntactic interference and the interaction of syntactic and semantic interference.
-Consequently, we conclude that the ERP data showed no syntactic interference and no interaction effect.}}\label{ERP_discussion}
+In contrast, there was only very weak evidence for very small effects of syntactic interference and the interaction or even evidence against them in both the N400 and P600 windows.
+In sum, the ERP results showed clear evidence for semantic interference and mostly evidence against syntactic interference and the interaction of syntactic and semantic interference.
+Consequently, we conclude that the ERP data showed no syntactic interference and no interaction.}}\label{ERP_discussion}
 
@@
 The more complex memory retrieval under high semantic interference resulted in a more negative N400.
-\textcolor{blue}{The modulation of the P600 might be due to facilitatory interference.
-The P600 is typically affected by syntactic processing, with more positive responses for more complex syntactic processing, e.g., reanalysis \citep{osterhout&holcomb_1992}.
-In the present design, the animate distractor might have facilitated integration of the verb with its (false) subject.
-Integration with one of two potentially good candidates could be easier than integration when there is only one good candidate.
-The smaller size of the P600 effect compared to the N400 amplitude modulation could indicate that this facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration.
-}
+\textcolor{blue}{This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}.
+The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference lead to a reduced P600 amplitude compared to low semantic interference.
+Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}.
+The reduced P600 amplitude under high semantic interference in the present study might be explained by facilitatory interference \citep[see e.g.,][]{jaeger_etal_2017}.
+In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to act as the subject.
+A similar effect was found by \textcite{Tanner_etal_2017}.
+In their study, the P600 elicited by ungrammatical verbs which were incongruent in number was reduced when the sentence included a matching distractor.
+However, the present study investigated only grammatical subject-verb integration and the size of the P600 effect was rather small.
+The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration.
+Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.}
 
@@
 
-\Copy{only_this_design}{Taken together, the SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and no effects of syntactic interference.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
+\Copy{only_this_design}{Taken together, the SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and no effects of syntactic interference \textcolor{blue}{or  an interaction}.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
@@
 \textcolor{blue}{In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design.
-Then we} discuss two important issues that relate to {the broader interference literature}.
+Then we discuss two important issues that relate to the broader interference literature}.
 The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.
 
-\subsection{Comparison with previous findings}
+\subsection{\textcolor{blue}{Comparison with previous findings}}\label{comparison}
 \Copy{comparison}{\textcolor{blue}{The present study used the same design as two previous studies \parencite{vandyke07, mertzen}.
@@
 \textcite{mertzen} made their data available.
-In comparison to the previous estimates, the effects of the present study are relatively small and associated with less uncertainty.
-It seems like the present estimates of syntactic interference are smaller, i.e., closer to zero, than the previous estimates and larger, i.e., further from zero in the positive direction, for the interaction.
+In comparison to the previous estimates, the effects of the present study are relatively small and associated with less uncertainty in general.
 The present and previous estimates of semantic interference seem relatively consistent: They indicate a positive effect which is largest in the pre-critical region and smallest in the critical region.
+It seems like the present estimates of syntactic interference are smaller, i.e., closer to zero, than the previous estimates.
+For the interaction, the present estimates are larger, i.e., further from zero in the positive direction, than the previous ones.
 To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies to compare them to the present Bayes factors.
 For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
-For \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE).
-Then, using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors with the Savage-Dickey method.
+To compute Bayes factors for \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) in the published paper.
+Then, using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method.
 Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we run on their data analog to our own analyses described in Statistical Analyses for Experiments 1a and b.
@@
 \begin{sidewaysfigure}[h]
-    \caption{A) Reading time differences in the regions of interest from the previous and present studies using the design by \textcite{vandyke07}.
+    \caption{A) Reading time differences in the regions of interest from the present and previous studies using the design by \textcite{vandyke07}.
 VD07.E1, VD07.E2 and VD07.E3 stand for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiments 1-3.
+The intervals for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) estimates are 95\% confidence intervals.
 M23.E and M23.G stand for \citeauthor{mertzen}'s (\citeyear{mertzen}) English and German experiment, respectively.
@@
 There are less estimates for the pre-critical region because \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1 and 2 did not have a pre-critical region.
-B) Bayes factors for the effects of interest in for the present and previous studies under prior Normal(0,0.5).}
+B) Bayes factors for the effects of interest in the present and previous studies under prior Normal(0,0.5).}
     \label{fig:previous_vs_present}
@@
 
-\textcolor{blue}{The Bayes factors for the syntactic interference effect show that the present study is the only one to provide evidence against the effect in the pre-critical region.
-In the critical region, both self-paced reading experiments provide evidence against the effect.
-The Bayes factors of three of the eye-tracking studies show weak or no evidence for the effect.
-The Bayes factor of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provides extremely strong evidence for the syntactic interference effect in the critical region.
-In the spill-over region, the Bayes factors for syntactic interference rather consistently provide evidence against the effect.
-Only \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment provides moderate evidence for it.}
-
-\textcolor{blue}{The Bayes factors for the semantic interference effect consistently provide evidence for the effect in the pre-critical region.
-In the critical region, most experiments provide evidence against semantic interference, the two self-paced reading experiments, being the exceptions, provide evidence for the effect.
-In the spill-over region, the Bayes factor results for semantic interference are mixed: the two self-paced reading experiments provide extremely strong evidence for the effect, two of the eye-tracking experiments provide weak evidence for the effect and two other eye-tracking experiments provide evidence against the effect.
-Importantly, the clustering in the results does neither align with the investigated language nor with the authors of the studies.}
-
-\textcolor{blue}{The Bayes factors for the interaction in all regions are rather consistent between studies and are either inconclusive or provide evidence against the interaction.
-The only exception is the very large Bayes factor in favor of the effect for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1.
-In sum, while the present and previous studies show comparable region-wise results for semantic interference and the interaction, they differ remarkably regarding syntactic interference in the pre-critical and critical region.
-In the pre-critical region, the present study is the obvious outlier, being the only study to provide evidence against the effect.
-In the critical region, four of the investigated experiments show either weak evidence for or weak evidence against syntactic interference, but there are two outliers: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extreme evidence for the effect and the present study provided strong evidence against it.
-In conclusion, the most remarkable difference between the present study and previous findings concerns syntactic interference in the pre-critical region.}
+\textcolor{blue}{In the pre-critical region, the Bayes factors from the present and previous studies showed no clear pattern regarding syntactic interference.
+The present study was the only one to provide evidence against the effect, while the previous studies supported it.
+The Bayes factors aligned qualitatively across studies for the semantic interference effect and the interaction -- consistently providing evidence for semantic interference and against the interaction.}
+
+\textcolor{blue}{In the critical region, the Bayes factors across studies showed a mixed picture for syntactic interference: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extremely strong evidence for it, while \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3 provided weak evidence against it.
+\citeauthor{mertzen}'s (\citeyear{mertzen}) experiments showed weak evidence for the effect.
+The two self-paced reading experiments provided evidence against it.
+Similarly, there is no clear pattern for semantic interference: Most experiments provided evidence against it, but the two self-paced reading experiments (present and \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1) provided evidence for semantic interference.
+The Bayes factors across studies provided evidence against the interaction.}
+
+\textcolor{blue}{In the spill-over region, the Bayes factors for syntactic interference rather consistently provided evidence against the effect.
+Only \citeauthor{mertzen}'s (\citeyear{mertzen}) English experiment provided moderate evidence for it.
+The Bayes factor results for semantic interference were mixed: the two self-paced reading experiments provided extremely strong evidence for the effect, two of the eye-tracking experiments (\citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3, \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment) provided weak evidence for the effect and the two other eye-tracking experiments (\citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2, \citeauthor{mertzen}'s (\citeyear{mertzen}) English experiment) provided evidence against the effect.
+Importantly, the clustering in the results did neither align with the investigated language nor with the authors of the studies.
+The Bayes factor results for the interaction showed evidence against it from all experiments but one.
+The outlier was \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment, which showed extreme evidence for the interaction.}
+
+\textcolor{blue}{In sum, while the present and previous studies mostly showed comparable region-wise results for semantic interference and the interaction, they differed remarkably regarding syntactic interference in the pre-critical and critical region.
+In the pre-critical region, the present study was the obvious outlier, being the only study to provide evidence against the effect.
+In the critical region, the Bayes factors for the syntactic interference were very mixed with no clear clustering.
+In conclusion, the most remarkable difference between the present study and previous findings concerns syntactic interference in the pre-critical region.
+The lack of syntactic interference in the pre-critical region in the present study compared to the previous studies could be explained by differences in i) methodology, ii) items and/or iii) statistical power.
+We discuss each of these differences in turn.}
 }
-
-\textcolor{blue}{The lack of syntactic interference in the pre-critical region in the present study compared to the previous studies could be explained by differences in i) methodology, ii) items and/or iii) statistical power.
-We discuss each of these differences in turn.}
 
@@
 Whether \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment used line breaks or not is not reported.
-In our SPR experiments, it was necessary to introduce line breaks into the sentences.
+In our SPR experiments, it was necessary to introduce line breaks.
 These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb.
@@
 Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference.
-For this point to hold, one would need to assume that the line break provided additional structural information by emphasizing the syntactic structure of the items in our SPR experiment.
-However, the EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor.
+For this point to hold, one would need to make the reasonable assumption that the line break provided additional structural information by emphasizing the syntactic structure of our sentences.
+However, our EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor.
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-Nevertheless, neither our SPR nor our EEG experiments showed decisive evidence for syntactic interference.}
+Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.}
 }\label{linebreak}
@@
 The three studies investigating syntactic and semantic interference with the same design used either English or German items.
-The manipulated features, i.e., subjecthood and animacy, are rather universal and are expressed in the same way across these languages.
 Cue-based retrieval as a theory on the memory processes during sentence comprehension implicitly states that these memory processes do not differ cross-linguistically \citep{lewis06, Lewis2005}.
@@
 \textcolor{blue}{The previous and present items differed in the pre-critical material.
-\citet{vandyke07} used one region consisting of two words (e.g.
-`yesterday afternoon') which were analyzed together.
-\textcite{mertzen} used a one-word pre-critical region (e.g.
-`tatsächlich', indeed).
+\citet{vandyke07} used one region consisting of two words (e.g., `yesterday afternoon') which were analyzed together.
+\textcite{mertzen} used a one-word pre-critical region (e.g., `tatsächlich', indeed).
 In the present study, we added a pre-pre-critical region to help us differentiate between the potential sources of the pre-critical effects in the literature as discussed in the Discussion of the SPR experiments.
@@
 \textcolor{blue}{In contrast, the large estimates of the previous studies are likely Type M errors, i.e., exaggerations, which are common in low-powered studies \parencite{gelman_carlin}.
-That at least some of the previous studies had low power is further emphasized by the large uncertainty associated with some of the previous estimates (see Figure \ref{fig:previous_vs_present} A).
-In sum, while there are differences in the employed methodology and items between the present and previous studies, the most likely cause of the different results is the difference in statistical power.
+That at least some of the previous studies had low power is further emphasized by the large uncertainty associated with some of the previous estimates (see Figure \ref{fig:previous_vs_present} A).}
+
+\textcolor{blue}{In sum, while there are differences in the employed methodology and items between the present and previous studies, the most likely cause of the different results is the difference in statistical power.
 However, an alternative explanation could be that syntactic interference might differ inter-individually \parencite{yadav2021individual} and that the previous studies by chance sampled disproportionally many participants who showed syntactic interference, while we by chance sampled participants who did not.}
@@
 \Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature \citep[see e.g.,][for a high-powered failed replication attempt of a classic result in the area of predictive processing]{nieuwland_etal_2018}.
-The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.}}\label{not_an_anomaly}
+The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}
 
@@
 In the SPR data, we found that the semantic interference effect originated at the distractor and persisted throughout the following regions.
-Due to this time course, encoding interference \parencite{Yadavetal2022,hammerly2019grammaticality} is the best explanation for the interference effect in the reading times at the critical region.
+Due to this time course, encoding interference \parencite{Yadavetal2022,hammerly2019grammaticality} is the best explanation for the interference effect\textcolor{blue}{s in our SPR data}.
 This raises the question whether the ERP effect was also caused by encoding interference, i.e., whether the ERP effect originated prior to the critical region.

### 2024-08-09 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \vskip 1em
-Together with my co-author Shravan Vasishth, I would like to submit our
-revision of the article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.
-We are grateful for thoughtful and constructive comments and suggestions from you and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
+Together with my co-authors Shravan Vasishth and Himanshu Yadav, who was added during the revision, I would like to submit our revision of the article titled ``Do syntactic and semantic similarity lead to interference effects?
+Evidence from self-paced reading and event-related potentials using German''.
+We are grateful for thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
 
@@
 In response to the concerns about our statistical analyses, we have completely re-done the analyses of the reading times and ERP data and added analyses of the previous data.
-Additionally, we have added one computational model and use k-fold cross-validation for model comparisons.
+We brought Himanshu Yadav into the project to refine our modeling, including the addition of a third model, and carry out model comparisons.
 We believe that these new analyses strengthen our initial interpretation that our study shows semantic interference, but no syntactic interference and no interaction.
@@
 Reviewer 2 primarily raised concerns regarding our experimental design.
-In response to the concerns about our experimental design, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textit{Journal of Memory and Language}.
+In response to these concerns, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textcolor{red}{\textit{Journal of Memory and Language}}.
 We now stress in the paper that we are building on previously published work by directly comparing the statistical inferences, using Bayes factors, from the present and previous studies that use exactly the same design.
@@
 In other words, the questions are whether one would ever expect to see much of this effect with your design and whether a more appropriate design would lead to stronger findings.
-Some of Reviewer 3's points (see their paragraph starting 'The authors attribute ...' and their final summary comment) are perhaps in the same vein.''
+Some of Reviewer 3's points (see their paragraph starting `The authors attribute ...' and their final summary comment) are perhaps in the same vein.''
 \end{quote}
@@
 \subsection*{Response to E.1}
-We have used a well-studied design that was previously published in the Journal of Memory and Language, among other places.
+We have used a well-studied design that was previously published in the \textcolor{red}{Journal of Memory and Language}, among other places.
 The previously published studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
@@
 We believe that our results are a valuable contribution to the literature on syntactic interference, and serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
+We added the following text to our General Discussion on page \pageref{only_this_design} to emphasize that our findings and conclusions only apply to the present design:
+
+\begin{quote}
+    \Paste{only_this_design}
+\end{quote}
+
 The reviewer's objection can and should be pursued in a future study that uses a design that the reviewer would consider appropriate, but this would require a completely new paper with a different research focus (namely, what is an appropriate syntactic manipulation that would lead to interference?).
@@
 
+\begin{quote}
     \Paste{future_work}
-
-We added the following text to our General Discussion on page \pageref{only_this_design} to emphasize that our findings and conclusions only apply to the present design:
-
-\Paste{only_this_design}
+\end{quote}
 
@@
 We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide evidence for syntactic interference or the interaction.
-As an overview of the revised Bayes factor analyses, we show below the Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs}.
+As an overview of the revised Bayes factor analyses, we show Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs} below.
 Figure \ref{fig:spr_bfs} shows the Bayes factor analysis of the self-paced reading data.
 There was evidence against syntactic interference under all priors in the pre-critical and critical region.
-Only in the spill-over region, there was anecdotal evidence (BF$_{10}$ = 2.8) evidence for a small syntactic interference effect ($\pm$ 8 ms).
+Only in the spill-over region, there was anecdotal evidence (BF$_{10}$ = 2.8) for a small syntactic interference effect ($\pm$ 8 ms).
 Similarly, there was evidence against the interaction under all priors in the pre-critical and spill-over regions.
@@
 Its main assumption is that sentence processing relies on skilled memory processes, with a focus on memory retrieval.
-We built on this assumption and focus on the question which specific retrieval cues are being used during subject-verb dependency processing.
+We built on this assumption and focus on the question which specific retrieval cues are being used during subject-verb dependency resolution.
 Our findings showed that the retrieval cues generated at the verb are highly specific (including hierarchical syntactic information) and can successfully identify the subject as the retrieval target.
@@
 
+\begin{quote}
     \Paste{intro}
+\end{quote}
 
-Additionally, we have added the following text to the General Discussion on page \textcolor{red}{X} to emphasize this point:
+Additionally, we have added the following text to the General Discussion on page \textcolor{red}{X} to emphasize the theoretical grounding of our work:
 
@@
 
+\begin{quote}
     \Paste{osf}
+\end{quote}
 
@@
 As suggested, we have implemented a null model (no interference), leaving us with a total of three computational models (the null model, the three-cues model by \textcite{mertzen} and our newly proposed two-cues model).
-To quantitatively compare the performance of these models, we have used a likelihood-free k-fold cross-validation approach developed by \textcite{Yadavetal2022}.
-The result of this model comparison is that \textcolor{red}{XXXX}.
+We use Bayes factors to quantitatively compare these models.
+The result of this model comparison is that the two-cues and three-cues models outperform the null model.
+Additionally, our new two-cues model performs slightly better than the three-cues model by \textcite{mertzen}.
 We have described the new modeling and model comparisons starting page \pageref{modeling}:
 
+\begin{quote}
     \Paste{modeling}
-
+\end{quote}
 \newpage
@@
 The Bayes factors of the present and previous studies align well with one exception: The previous studies found evidence for syntactic interference in the pre-critical
-region, while our study did not. The following description of the estimates and Bayes factors of the present and previous studies and their interpretation was added to the manuscript:
+region, while our study did not.
+The following description of the estimates and Bayes factors of the present and previous studies and their interpretation was added to the manuscript, starting page \pageref{comparison}:
 \setcounter{figure}{12}
 
+\begin{quote}
     \Paste{comparison}
+\end{quote}
 
@@
 \subsubsection*{Response to R1.5}
-About why our results diverge from previous studies' and whether we
-think our results are correct and the previous work is not, we have no
-position on this. We are simply reporting what we found. Contradictory
-results have been reported for many phenomena: e.g., reflexives
-processing \citep{jaeger_etal_2020}, main verb/reduced relative garden paths (\textcolor{red}{REF?}), predictive
-parsing effects \citep{nieuwland_etal_2018}, local coherence \citep{lcpaape2023}. The
-only way to resolve these many contradictions in the literature is by
-carrying out adequately powered studies in the future that can
-distinguish between the theoretical claims.  Our work does not stand
-out as an anomaly; if anything, it's quite the opposite. We have added a discussion of divergent results in the literature to page \pageref{not_an_anomaly} in the General Discussion:
+About why our results diverge from previous studies' and whether we think our results are correct and the previous work is not, we have no position on this.
+We are simply reporting what we found.
+Contradictory results have been reported for many phenomena: e.g., reflexives processing \citep{jaeger_etal_2020}, main verb/reduced relative garden paths (\textcolor{red}{REF?}), predictive parsing effects \citep{nieuwland_etal_2018}, local coherence \citep{lcpaape2023}.
+The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.
+Our work does not stand out as an anomaly; if anything, it's quite the opposite.
+We have added a discussion of divergent results in the literature to page \pageref{not_an_anomaly} in the General Discussion:
 
+\begin{quote}
     \Paste{not_an_anomaly}
+\end{quote}
 
@@
 \subsubsection*{Response to R1.6}
-We ran the suggested plausibility judgement experiment. The -animate distractor is indeed slightly more plausible in the context than the +animate distractor. To investigate whether this plausibility
-difference is driving the semantic interference effect in our SPR data, we ran a Bayesian hierarchical mixed model on the reading times of the pre-critical region which included plausibility as a
-co-predictor and find that there is still evidence for semantic interference.
+We ran the suggested plausibility judgement experiment.
+The -animate distractor is indeed slightly more plausible in the context than the +animate distractor.
+To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian hierarchical mixed model on the reading times of the pre-critical region which included plausibility as a co-predictor and find that there is still evidence for an independent semantic interference effect.
 We discuss the plausibility confound and the additional analyses in the revised manuscript starting on page \pageref{plausib_anim_inan}:
 
+\begin{quote}
     \Paste{plausib_anim_inan}
+\end{quote}
 
@@
 \subsubsection*{Response to R1.7}
-As mentioned in our response to R1.4, we recomputed our Bayes Factors using the Savage-Dickey method.
-With this method, we were able to compute Bayes Factors for models with complex random effect structures.
+We have recomputed all our Bayes factors using the Savage-Dickey method.
+With this method, we were able to compute Bayes factors for models with complex random effect structures.
 We have re-run all models with full random effect structures and updated all results in the manuscript accordingly.
@@
 The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}).
-Because such large interference effects have been found previously (see also our new Figure 12 showing all estimates of \cite{vandyke07} and \cite{mertzen}), we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times.
+Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}), we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times.
 
@@
 \setcounter{figure}{4}
+\begin{quote}
     \Paste{SPR_bfs}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{ERP_results3}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{only_precritical}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{insig_interaction}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{why_n400} 
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{ERP_results1}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{ERP_results2}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{ERP_discussion}
+\end{quote}
 
@@
 We agree that this line of work is highly relevant for our paper.
-We have added the suggested references (and some more by the same and other authors) to the introduction (page) \pageref{franck_refs} and General Discussion (page \pageref{future_work}):
+We have added the suggested references (and some more by the same and other authors) to the introduction (page \pageref{franck_refs}) and General Discussion (page \pageref{future_work}):
 
+\begin{quote}
     \Paste{franck_refs}
+\end{quote}
 
+\begin{quote}
     \Paste{future_work}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{clause_tracker}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{semantic_features_encoding}    
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{fillers1}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{fillers2}    
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{comma}
+\end{quote}
 
@@
 
+\begin{quote}
     \Paste{linebreak}
+\end{quote}
 
@@
 \setcounter{exx}{4}
+\begin{quote}
     \Paste{items_mertzen}
+\end{quote}
 

### 2024-08-11 code_data/Model-comparison-using-Bayes-factors/Simulate_data_two_cues_model.R
Added via the Open Science Framework

@@ -1,0 +1,105 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+# Load model 
+source("Models/interACT.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+#### Simulate data from the proposal density #####
+
+nsims <- 200000
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf <<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
+  meSem<-mean(c(condMeans$Latency[2],condMeans$Latency[4]))-
[…]

### 2024-08-11 code_data/Model-comparison-using-Bayes-factors/Compute-Bayes-factors.R
Added via the Open Science Framework

@@ -1,0 +1,296 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+library(dplyr)
+
+#############
+SVdat <- data.frame(pred   = rep("data",3),
+                    effect    = c("syntactic", "semantic", "interaction"), 
+                    mean  = c(-0.14,-0.3,-0.03),
+                    upper = c(0.05,-0.1,0.38),
+                    lower = c(-0.33,-0.5,-0.33))
+SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
+################
+
+tol_scaler <- 9
+delta <- (abs(SVdat[2,]$upper-SVdat[2,]$lower)/(4*tol_scaler))
+
+## Prior 1: normal_{lb=0,ub=0.05}(0.01,0.01)
+
+load("Simulated-data/Simulated_data_two_cues_model.Rda")
+df.lkl <- simdat_2cues_sameclause
+head(df.lkl)
+df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
+df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
+df.lkl$proposal.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0,sd=0.025)
+ML <- mean((df.lkl$lkl)*(df.lkl$prior.dens)/(df.lkl$proposal.dens))
+df.lkl$ML <- ML
+df.lkl$prior <- "normal(0.01,0.01)"
+df.bf <- as.data.frame(df.lkl %>% group_by(prior) %>% 
+                         summarize(MargLik=mean(ML)))
+df.bf
+df.bf$Model <- "Two cues model"
+
+load("Simulated-data/Simulated_data_three_cues_model.Rda")
+df.lkl <- simdat_3cues
+df.lkl$Model <- "Three cues model"
+df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
+df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
[…]

### 2024-08-11 code_data/Model-comparison-using-Bayes-factors/Models/interACT-threecues.R
Added via the Open Science Framework

@@ -1,0 +1,240 @@
+library(tidyr)
+library(dplyr)
+
+# Original code by Felix Engelmann (https://github.com/felixengelmann/inter-act)
+
+###############################################################
+## Global parameters
+###############################################################
+
+reset_params <- function(){
+  lf        <<- 0.15       # latency factor
+  le        <<- 1       # latency exponent
+  rth 	    <<- -1.5    # retrieval threshold
+  bll       <<- 0.5     # decay parameter
+  ans       <<- 0.2    # activation noise
+  mas       <<- 1       # maximum associative strength 
+  mp        <<- 1       # mismatch penalty
+  ga        <<- 1       # goal source activation
+  rand_time <<- 3       # latency variability
+  blc       <<- 0       # base-level activation
+  ##
+  lp        <<- 1     # default time since last presentation (msec)
+  ldp       <<- 1       # last distractor presentation (msec)
+  ## Distractor control:
+  ndistr    <<- 1				# number of distractors
+  dbl       <<- 0       # distractor base-level
+  ##
+  ## Cue weighting, cue confusion and prominence:
+  cueweighting <<- 1     # Strength of structural cue as ratio str/sem
+  normalizeWeights <<- TRUE 
+  qcf 		  <<- 1 			# match quality correction factor
+  qco 		  <<- -2*rth 			# match quality correction offset
+  psc       <<- 1       # prominence scaling constant C1
+  pic       <<- 0       # prominence scaling constant C2
+  tprom     <<- 0       # target prominence 
+  dprom     <<- 0       # distractor prominence
+  # cl 				<<- 0 			# cue confusion level (0-100)
+  cuesim    <<- -1    # cue-feature similarity [-1..0]
+  #
+  VERBOSE <<- TRUE
+}
+reset_params()
+
+idnames <- c("Set","Iteration","Condition","Target","Distractor")
+actrnames <- c("weights","bl1","bl2","times1","times2","noise1","noise2","blact1","blact2","act1","act2","activation","latency","retrieved","acc","miss","fail")
[…]

### 2024-08-11 code_data/Model-comparison-using-Bayes-factors/Simulate_data_one_cue_model.R
Added via the Open Science Framework

@@ -1,0 +1,113 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+################
+
+#### One CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+
+# Load model 
+source("Models/interACT-onecue.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1), c(1), c(1), c(1)),
+  distractor_match = list(c(0), c(0), c(0), c(0)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(1))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate prior predictions #####
+
+# Generate 200,000 simulations from model:
+nsims<-200000
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
+
+## check:
+hist(lfs,freq=FALSE)
+
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
[…]

### 2024-08-11 code_data/Model-comparison-using-Bayes-factors/Simulate_data_three_cues_model.R
Added via the Open Science Framework

@@ -1,0 +1,105 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+################
+
+
+# Load model 
+source("Models/interACT-threecues.R")
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
+  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(0.33,0.33,0.33))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate prior predictions #####
+# Generate 100,000 simulations from model:
+nsims<-200000
+
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
+  meSem<-mean(c(condMeans$Latency[2],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[3]))
[…]

### 2024-08-11 code_data/LV05_modeling/Compute-Bayes-factors.R
Added via the Open Science Framework

@@ -1,0 +1,319 @@
+# author: Himanshu Yadav
+# date: 11.08.2024
+
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+library(dplyr)
+
+#############
+SVdat <- data.frame(pred   = rep("data",3),
+                    effect    = c("syntactic", "semantic", "interaction"), 
+                    mean  = c(-0.14,-0.3,-0.03),
+                    upper = c(0.05,-0.1,0.38),
+                    lower = c(-0.33,-0.5,-0.33))
+SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
+################
+
+tol_scaler <- 9
+delta <- (abs(SVdat[2,]$upper-SVdat[2,]$lower)/(4*tol_scaler))
+
+## Prior 1: normal_{lb=0,ub=0.05}(0.01,0.01)
+
+load("Simulated-data/Simulated_data_two_cues_model.Rda")
+df.lkl <- simdat_2cues_sameclause
+head(df.lkl)
+df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
+df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
+df.lkl$proposal.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0,sd=0.025)
+ML <- mean((df.lkl$lkl)*(df.lkl$prior.dens)/(df.lkl$proposal.dens))
+df.lkl$ML <- ML
+df.lkl$prior <- "normal(0.01,0.01)"
+df.bf <- as.data.frame(df.lkl %>% group_by(prior) %>% 
+                         summarize(MargLik=mean(ML)))
+df.bf
+df.bf$Model <- "Two cues model"
+
+load("Simulated-data/Simulated_data_three_cues_model.Rda")
+df.lkl <- simdat_3cues
+df.lkl$Model <- "Three cues model"
+df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
[…]

### 2024-08-11 code_data/LV05_modeling/Generate_posterior_predictions.R
Added via the Open Science Framework

@@ -1,0 +1,380 @@
+# author: Pia Schoknecht
+# date: 11.08.2024
+
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(ggbreak)
+library(cowplot)
+library(stringr)
+library(tidyverse)
+
+# save Schoknecht & Vasishth mean, lower and upper separately to add them to plots later
+# add data from Schoknecht & Vasishth ERP experiment
+SVdat <- data.frame(pred   = rep("data",3),
+                  effect    = c("syntactic", "semantic", "interaction"), 
+                  mean  = c(-0.14,-0.3,-0.03),
+                  upper = c(0.05,-0.1,0.38),
+                  lower = c(-0.33,-0.5,-0.33))
+SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
+
+# load simulated data from all 3 models
+load("Simulated-data/Simulated_data_one_cue_model.Rda")
+load("Simulated-data/Simulated_data_two_cues_model.Rda")
+load("Simulated-data/Simulated_data_three_cues_model.Rda")
+
+
+#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+# Load model 
+source("Models/interACT.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+
+# multiply by -1 for negative effects
+simdat_2cues_sameclause$MESyn <- simdat_2cues_sameclause$MESyn * -1
[…]

### 2024-08-11 code_data/LV05_modeling/Simulate_data_one_cue_model.R
Added via the Open Science Framework

@@ -1,0 +1,116 @@
+# author: Himanshu Yadav
+# date: 11.08.2024
+
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+################
+
+#### One CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+
+# Load model 
+source("Models/interACT-onecue.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1), c(1), c(1), c(1)),
+  distractor_match = list(c(0), c(0), c(0), c(0)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(1))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate prior predictions #####
+
+# Generate 200,000 simulations from model:
+nsims<-200000
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
+
+## check:
+hist(lfs,freq=FALSE)
+
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
[…]

### 2024-08-11 code_data/LV05_modeling/Simulate_data_three_cues_model.R
Added via the Open Science Framework

@@ -1,0 +1,108 @@
+# author: Himanshu Yadav
+# date: 11.08.2024
+
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+################
+
+
+# Load model 
+source("Models/interACT-threecues.R")
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
+  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(0.33,0.33,0.33))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+# Generate prior predictions #####
+# Generate 100,000 simulations from model:
+nsims<-200000
+
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
+  meSem<-mean(c(condMeans$Latency[2],condMeans$Latency[4]))-
[…]

### 2024-08-11 code_data/LV05_modeling/Simulate_data_two_cues_model.R
Added via the Open Science Framework

@@ -1,0 +1,108 @@
+# author: Himanshu Yadav
+# date: 11.08.2024
+
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+
+
+#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+# Load model 
+source("Models/interACT.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+VERBOSE <<- FALSE
+
+#### Simulate data from the proposal density #####
+
+nsims <- 200000
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf <<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
[…]

### 2024-08-11 paper/main.tex
Updated via the Open Science Framework

@@
 
-\Copy{modeling}{\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times.
+\label{modeling}\Copy{modeling}{\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times.
 Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
@@
 For this we adjust the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.} The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
-We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$).
-This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal likelihood for all effect sizes in this range.
-This uninformative prior was chosen because there exists no previously published ERP data on the Van Dyke (2007) design.
-After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
+We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal(0, 0.025)$) with a lower bound of 0 and an upper bound of 0.05.
+This prior reflects the a priori belief that effects will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 1 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
+This relatively uninformative prior was chosen because although there exists no previously published ERP data on the Van Dyke (2007) design it reflects the reasonable range of effect sizes in psycholinguistic ERP research.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
 \textcolor{blue}{We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study.
 To reflect that larger predicted effect sizes correspond to more negative N400 amplitudes, the predicted effects from the model} were multiplied with $-1$.
-To avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects \textcolor{blue}{and their interaction.}
 
@@
 
-\textcolor{blue}{Out of 100,000 iterations of ABC algorithm, we obtained only 6467 samples of $F$ for the 3-cues model, 1010 for the 2-cues model and 869 for the 1-cue model.
+\textcolor{blue}{Out of 200,000 iterations of ABC algorithm, we obtained only 40,748 samples of $F$ for the 3-cues model, 28,718 for the 2-cues model and 6,615 for the 1-cue model.
 This is not surprising given
@@
 The prior and posterior predictions of all models overlap with the effects of interest in the data.
-To evaluate the predictive fit of the three models to our data quantitatively, we used the likelihood-free k-fold cross-validation model comparison approach developed by \textcite{Yadavetal2022}.}
+}
 
@@
 
-\textcolor{blue}{For this k-fold cross-validation model comparison, we repeatedly computed how well each of the three models performed on a held-out small subset of the data after being trained on the remaining large subset of the data.
-We repeatedly left out 1/10$^{th}$ of the data which was then used as the test data, while the other 9/10$^{th}$ functioned as the training data.
-It was made sure that each of the 10 subsets included data from all participants, all items and all conditions.
-The models were trained by generating posterior predictive distributions by fitting the $F$ parameter to the posterior distribution of the effects of interest in the training data.
-Then these posterior distributions of $F$ were used to compute the likelihood of generating the held-out test data.
-The log of this likelihood is the predictive accuracy of the model.
-This was done 10 times.
-Then, we computed the sum and standard error of sum of the predictive accuracies of the models of each of the training and test datasets to gain the overall predictive accuracy, the expected log predictive density $\widehat{elpd}$, of each model.
-To evaluate the performances of the models, we compare the difference in the $\widehat{elpd}$ values between each model and the other two models (see Figure \ref{fig:elpd}).
+\textcolor{blue}{To evaluate whether our new 2-cues model quantitatively outperforms the two other models in its predictive fit, we computed Bayes factors for the relevant model pairs under six different priors for the $F$ parameter
+($Normal$(0.01,0.01), 
+$Normal$(0.01,0.02),
+$Normal$(0.01,0.005),
+$Normal$(0.02,0.01),
+$Normal$(0.02,0.02),
+$Normal$(0.02,0.005)).
+First, we computed the likelihood of the sampled $F$ parameters under each of the priors (LKL$_{F-prior}$) and the sampling prior (LKL$_{F-model}$).
+Then, the likelihood of the predicted semantic interference effect from the model was computed given the size of the observed semantic interference effect (LKL$_{predicted-effect}$).
+To obtain the marginal likelihood of the model, the product of LKL$_{predicted-effect}$ and LKL$_{F-prior}$ was divided by LKL$_{F-model}$.
+The marginal likelihood of the 2-cues model was then divided by the marginal likelihood of the three-cues and one-cue model, respectively.
+The resulting Bayes factors are shown in Figure \ref{fig:BF_compmods}.
+Compared to the null model with one cue, the Bayes factors provide extreme evidence for the new two-cues model.
+Therefore, we can conclude that the computational modeling shows that our data shows interference.
+Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide moderate evidence for the new two-cues model.
+Therefore, the assumption that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue – is moderately supported by these modeling results.
 }
@@
     \caption{Comparison of the predictive accuracies of the three versions of the Lewis and Vasishth (2005) cue-based retrieval model on our ERP data.
-Each facet shows how the model in the facet's title performs in comparison to the models on the x-axis (positive $\widehat{elpd}$ differences correspond to better performance).
-Error bars show two times the standard error of the difference in $\widehat{elpd}$ values.}
-    \label{fig:elpd}
-    \includegraphics[width=\linewidth]{images/ELPD_difference.jpg}
+The facets show how the new two-cues model performs compared the the three-cues model and the one-cue model, respectively.
+Larger Bayes factors correspond to better performance of the two-cues model.}
+    \label{fig:BF_compmods}
+    \includegraphics[width=\linewidth]{images/BF_plot_compmodels.png}
 \end{figure}
 
-\textcolor{red}{add description of figure and conclusion of the modeling section}
-
-%Thus, the theoretical assumption that is consistent with the observed ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
+Thus, the theoretical assumption that is consistent with the observed ERP data is that although the parser uses the animacy cue exactly as proposed in previous work on cue-based retrieval, the syntactic cue keeps track of the clausal location of the subject.
 This is of course a speculative claim that would need to be further investigated empirically.
 
-}\label{modeling}%end copy
+}%end copy
 

### 2024-08-11 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-Additionally, we have added the following text to the General Discussion on page \textcolor{red}{X} to emphasize the theoretical grounding of our work:
-
-\textcolor{red}{Add something to GenDisc after modeling is done}
-
 \subsection*{Comment E.4}
@@
 We use Bayes factors to quantitatively compare these models.
-The result of this model comparison is that the two-cues and three-cues models outperform the null model.
-Additionally, our new two-cues model performs slightly better than the three-cues model by \textcite{mertzen}.
+The result of this model comparison is that our new two-cues model outperforms the other two models.
 We have described the new modeling and model comparisons starting page \pageref{modeling}:
@@
 \subsubsection*{Response to R2.3}
-The participants were able to discern the two syntactic conditions at region 8 because this is the region were the conditions started to differ (`\dots der erzählt hatte \dots' vs.\ `\dots der von dem \dots').
-We have added the words of an example item as x-axis labels to Figure \ref{fig:whole_sentence} on page \pageref{fig:whole_sentence} to illustrate this clearly.
+The participants were able to discern the two syntactic conditions at region 8 because this is the region were the conditions started to differ (`\dots erzählt hatte \dots' vs.\ `\dots von dem \dots').
+We have added the words of an example item as x-axis labels to Figure \ref{fig:whole_sentence} to illustrate this clearly.
 For convenience, we paste the Figure here:

### 2024-08-12 paper/main.tex
Updated via the Open Science Framework

@@
 We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07,mertzen}, to investigate interference due to syntactic and semantic cues in German.
-A Bayes factor analysis showed that 
-both the self-paced reading times as well as event-related potentials provided clear evidence for a semantic interference effect.
+A Bayes factor analysis showed that both the self-paced reading times as well as event-related potentials provided clear evidence for a semantic interference effect.
 In the self-paced reading experiment, interference due to the semantic cues caused a reading time slow-down starting at the distractor.
@@
 For this we adjust the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.} The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
-We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal(0, 0.025)$) with a lower bound of 0 and an upper bound of 0.05.
-This prior reflects the a priori belief that effects will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 1 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
+We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05.
+This prior reflects the a priori belief that effect on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
 This relatively uninformative prior was chosen because although there exists no previously published ERP data on the Van Dyke (2007) design it reflects the reasonable range of effect sizes in psycholinguistic ERP research.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
@@
 
-It is informative to fit the model\textcolor{blue}{s} to the data and then examine the posterior predictive distribution of the \textcolor{blue}{ERPs} for each effect. 
+%It is informative to fit the model\textcolor{blue}{s} to the data and then examine the posterior predictive distribution of the \textcolor{blue}{ERPs} for each effect.
 We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the \textcolor{blue}{ERP} data.
@@
 
-\textcolor{blue}{Out of 200,000 iterations of ABC algorithm, we obtained only 40,748 samples of $F$ for the 3-cues model, 28,718 for the 2-cues model and 6,615 for the 1-cue model.
-This is not surprising given
-the relatively wide prior on $F$.
+%\textcolor{blue}{Out of 200,000 iterations of ABC algorithm, we obtained only 40,748 samples of $F$ for the 3-cues model, 28,718 for the 2-cues model and 6,615 for the 1-cue model.
+This is not surprising given the relatively wide prior on $F$.
 In future work, the posterior distribution, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.
-Figure \ref{fig:model_predictions_ERP} shows the  posterior predictive distributions
-of the effects of interest in comparison to the prior predictive distribution and our ERP
-data.
+Figure \ref{fig:model_predictions_ERP} shows the  posterior predictive distributions of the effects of interest in comparison to the prior predictive distribution and our ERP data.
 The prior and posterior predictions from the model with three cues show equally large effects of syntactic and semantic interference and an interaction.
@@
 \begin{sidewaysfigure}[htpb]
-    \caption{Prior and posterior predicted effect sizes on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
+    \caption{The prior predictions on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
     %using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. 
@@
 
-\textcolor{blue}{To evaluate whether our new 2-cues model quantitatively outperforms the two other models in its predictive fit, we computed Bayes factors for the relevant model pairs under six different priors for the $F$ parameter
-($Normal$(0.01,0.01), 
+\textcolor{blue}{We evaluate whether our new two-cues model quantitatively outperforms the three-cues model and the one-cue model in its predictive performance on the semantic interference data.
+To evaluate the relative performance of the models, we use the Bayes factor analysis.
+We compute Bayes factors for the two-cues model compared with the three-cues model and the one cue model.
+The Bayes factor obtained for the two-cues model would imply the evidence in favor of this model when it is compared against the other two models.
+Since Bayes factors are sensitive to the choice of priors on the parameters of interest. We compute Bayes factors under six different priors for the latency factor parameter $F$,
+$Normal$(0.01,0.01), 
 $Normal$(0.01,0.02),
@@
 $Normal$(0.02,0.01),
-$Normal$(0.02,0.02),
-$Normal$(0.02,0.005)).
-First, we computed the likelihood of the sampled $F$ parameters under each of the priors (LKL$_{F-prior}$) and the sampling prior (LKL$_{F-model}$).
-Then, the likelihood of the predicted semantic interference effect from the model was computed given the size of the observed semantic interference effect (LKL$_{predicted-effect}$).
-To obtain the marginal likelihood of the model, the product of LKL$_{predicted-effect}$ and LKL$_{F-prior}$ was divided by LKL$_{F-model}$.
-The marginal likelihood of the 2-cues model was then divided by the marginal likelihood of the three-cues and one-cue model, respectively.
-The resulting Bayes factors are shown in Figure \ref{fig:BF_compmods}.
-Compared to the null model with one cue, the Bayes factors provide extreme evidence for the new two-cues model.
+$Normal$(0.02,0.02), and 
+$Normal$(0.02,0.005).
+
+The procedure for computing Bayes factors is described as follows.
+The Bayes factor computation requires us to estimate the marginal likelihood of the competing models.
+We used the Monte Carlo integration method to estimate the marginal likelihood of each model.
+First, we chose an importance density for the parameter $F$ from which we draw $0.2$ million samples of $F$.
+Then, for each of these proposed parameter values, we compute the likelihood and prior density.
+The estimated marginal likelihood of a model will be given by
+
+\begin{equation}
+ML = \frac{1}{n} \sum_{i=1}{n} \mathcal{L}{F_i|y} p(F_i)  \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
+\end{equation}
+
+\noindent where $y$ is the observed data, i.e., observed semantic interference effect in our case, $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$ and $p(F_i)$ gives the prior density of $F_i$.
+We compute marginal likelihoods under different prior assumptions $p(F)$.
+
+The LV05 model is a complex, non-deterministic model and it is difficult the express the model's likelihood anallytically without drastically simplifying the model cite[e.g., in ][]{NicenboimRetrieval2018,lisson2020computational}.
+In absence of an analytically-expressed likelihood function, we use Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood} approach to estimate approximate likelihood for a given value of $F$.
+Under this approach, we compare the model generated data conditioned on $F=F_i$ and the actual observed data $y$.
+If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
+The likelihood term can be rewritten as $\mathcal{L}{F_i|y} = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represent a density kernel that weighs a proposal $F_i$ based on ditance between model-generated data and actual data, $d(x_i,y)$.
+
+Using the above method, we estimate marginal likelihood for each candidate model under six different prior assumptions. The Bayes factor in favor of a model M1
+compared to a model M2, i.e., BF12 is computed as the ratio
+of the marginal likelihoods of M1 and M2. The Bayes factor BF12 represents the extent to which the model M1 is more
+likely than M2 given the data. Following the convention from
+Jeffreys (1939/1998), a Bayes factor value of larger than 10
+is interpreted as strong evidence in favor of M1 and a value
+between 3 and 10 is interpreted as moderate evidence in favor
+of M1.
+
+From the marginal likelihoods, we compute Bayes factors for the two-cues model compared against the three-cues model and the one-cue model.
+That is, the marginal likelihood of the 2-cues model $\mathrm{M1}$ was divided by the marginal likelihood of the three-cues model $\mathrm{M2}$ and one-cue model $\mathrm{M3}$ to obtain the Bayes factors $BF_{12}$ and $BF_{13}$ respectively.
+Following the convention from \citep{jeffreys1998theory}, if the Bayes factor value is larger than $10$.
+it would be interpreted as strong evidence in favor of the two-cues model and if the Bayes factor is between $3$ and $10$, it is interpreted as moderate evidence in favor of the two-cues model compared against the competing model.
+
+The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
+Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model.
 Therefore, we can conclude that the computational modeling shows that our data shows interference.
@@
 Therefore, the assumption that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue – is moderately supported by these modeling results.
-}
+This result holds for prior assumptions that allows higher probability mass for small values of $F$.
+i.e., for $F<0.01$.
 

### 2024-08-12 paper/main.tex
Updated via the Open Science Framework

--- 2024-08-11 paper/main.tex
@@
 For this we adjust the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.} The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
-We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal(0, 0.025)$) with a lower bound of 0 and an upper bound of 0.05.
-This prior reflects the a priori belief that effects will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 1 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
-This relatively uninformative prior was chosen because although there exists no previously published ERP data on the Van Dyke (2007) design it reflects the reasonable range of effect sizes in psycholinguistic ERP research.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
+We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05.
+This prior reflects the a priori belief that effect on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
+This relatively uninformative prior was chosen because although there exists no previously published ERP data on the Van Dyke (2007) design it reflects the reasonable range of effect sizes in psycholinguistic ERP research.
+In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
 \textcolor{blue}{We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study.
@@
 The \{$\pm$ same clause\} cue matches only with nouns that occurred in the same clause as the verb.
-``The addition of the \{$\pm$ same clause\} cue [was] necessary to identify the correct subject'' and as a means for the model to ``to achieve correct retrieval'' in the condition with high syntactic and high semantic interference \citep[][p.
+``The addition of the \{$\pm$ same clause\} cue [was] necessary to identify the correct subject'' and as a means for the model ``to achieve correct retrieval'' in the condition with high syntactic and high semantic interference \citep[][p.
 14]{mertzen}.
@@
 
-It is informative to fit the model\textcolor{blue}{s} to the data and then examine the posterior predictive distribution of the \textcolor{blue}{ERPs} for each effect. 
-We generated posterior predictive distributions for the effects of interest by fitting the $F$ parameter to the \textcolor{blue}{ERP} data.
-To avoid overfitting, we fitted $F$ only to the semantic interference effect (and not the syntactic interference effect or interaction).
-This was done by implementing an Approximate Bayesian Computation approach \parencite{SissonABC}: the model's predictions are generated iteratively, and only those parameter values for the scaling parameter $F$ are retained that predict values for the semantic interference effect that are close to the observed effect size $d$ in the \textcolor{blue}{ERP} data.
-`Close' is defined as $d \pm \,2\times SE$ ms; for details of this approach as applied to the cue-based retrieval model, see  \textcite{vasishth2020_abc}.
-
-\textcolor{blue}{Out of 200,000 iterations of ABC algorithm, we obtained only 40,748 samples of $F$ for the 3-cues model, 28,718 for the 2-cues model and 6,615 for the 1-cue model.
-This is not surprising given
-the relatively wide prior on $F$.
-In future work, the posterior distribution, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.
-Figure \ref{fig:model_predictions_ERP} shows the  posterior predictive distributions
-of the effects of interest in comparison to the prior predictive distribution and our ERP
-data.
-The prior and posterior predictions from the model with three cues show equally large effects of syntactic and semantic interference and an interaction.
-In contrast, the model using two cues, including a syntactic cue that utilizes syntactic hierarchy, predicts no effect of syntactic interference and no interaction, but a semantic interference effect.
-The model using one cue predicts no interference effects.
-The prior and posterior predictions of all models overlap with the effects of interest in the data.
+\textcolor{blue}{Figure \ref{fig:model_predictions_ERP} shows the  prior predictive distributions of the effects of interest for the three competing models discussed above in comparison to our ERP data.
+The null model model using one cue predicts no interference effects.
+The new two-cues model, including a syntactic cue that utilizes syntactic hierarchy, predicts no effect of syntactic interference and no interaction, but a semantic interference effect.
+The prior predictions from the three-cues model show equally large effects of syntactic and semantic interference and an interaction.
+The predictions of all models overlap with the effects of interest in the data.
 }
@@
 \Copy{figpriorposteriordata}{
-\begin{sidewaysfigure}[htpb]
-    \caption{Prior and posterior predicted effect sizes on the microvolt scale, i.e., for negative ERPs, for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
+\begin{figure}[htpb]
+    \caption{The prior predictions on the microvolt scale for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
     %using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. 
@@
         \centering
-         \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
-\end{sidewaysfigure}
+         \includegraphics[width=\textwidth]{images/compmod_predictions.png}
+\end{figure}
 }
 
-\textcolor{blue}{To evaluate whether our new 2-cues model quantitatively outperforms the two other models in its predictive fit, we computed Bayes factors for the relevant model pairs under six different priors for the $F$ parameter
-($Normal$(0.01,0.01), 
+\textcolor{blue}{To quantitatively evaluate the relative predictive performances of the models on the semantic interference data, we used a Bayes factor analysis.
+We computed Bayes factors for the two-cues model compared to the three-cues model and to the one cue model, respectively.
+Since Bayes factors are sensitive to the choice of priors on the parameters of interest, we computed Bayes factors under six different priors for the latency factor parameter $F$,
+$Normal$(0.01,0.01), 
 $Normal$(0.01,0.02),
@@
 $Normal$(0.02,0.01),
-$Normal$(0.02,0.02),
-$Normal$(0.02,0.005)).
-First, we computed the likelihood of the sampled $F$ parameters under each of the priors (LKL$_{F-prior}$) and the sampling prior (LKL$_{F-model}$).
-Then, the likelihood of the predicted semantic interference effect from the model was computed given the size of the observed semantic interference effect (LKL$_{predicted-effect}$).
-To obtain the marginal likelihood of the model, the product of LKL$_{predicted-effect}$ and LKL$_{F-prior}$ was divided by LKL$_{F-model}$.
-The marginal likelihood of the 2-cues model was then divided by the marginal likelihood of the three-cues and one-cue model, respectively.
-The resulting Bayes factors are shown in Figure \ref{fig:BF_compmods}.
-Compared to the null model with one cue, the Bayes factors provide extreme evidence for the new two-cues model.
+$Normal$(0.02,0.02), and 
+$Normal$(0.02,0.005). }
+
+\textcolor{blue}{The procedure for computing Bayes factors is described as follows.
+The Bayes factor computation required us to estimate the marginal likelihood of the competing models.
+We used the Monte Carlo integration method to estimate the marginal likelihood of each model.
+First, we chose an importance density for the parameter $F$ from which we draw $0.2$ million samples of $F$.
+Then, for each of these proposed parameter values, we compute the likelihood and prior density.
+The estimated marginal likelihood of a model will be given by}
+
+\begin{equation}
+ML = \frac{1}{n} \sum_{i=1}{n} \mathcal{L}{F_i|y} p(F_i)  \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
+\end{equation}
+
+\noindent \textcolor{blue}{where $y$ is the observed data, i.e., the observed semantic interference effect in our case, $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$ and $p(F_i)$ gives the prior density of $F_i$.
+We compute marginal likelihoods under different prior assumptions $p(F)$.}
+
+\textcolor{blue}{The Lewis and Vasishth (2005) model is a complex, non-deterministic model and it is difficult to express the model's likelihood analytically without drastically simplifying the model \parencite[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}.
+In absence of an analytically-expressed likelihood function, we used an Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood} approach to estimate approximate likelihood for a given value of $F$.
+Under this approach, we compared the model generated data conditioned on $F=F_i$ and the actual observed data $y$.
+If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ was close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
+The likelihood term can be rewritten as $\mathcal{L}{F_i|y} = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represent a density kernel that weighs a proposal $F_i$ based on distance between model-generated data and actual data, $d(x_i,y)$.}
+
+%\textcolor{blue}{Using the above method, we estimate marginal likelihood for each candidate model under six different prior assumptions.
+The Bayes factor in favor of a model M$_1$ compared to a model M$_2$, i.e., BF$_{12}$, is computed as the ratio of the marginal likelihoods of M$_1$ and M$_2$.
+The Bayes factor BF$_{12}$ represents the extent to which the model M$_1$ is more likely than M$_2$ given the data.
+Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above 10 is interpreted as strong evidence in favor of M$_1$ and a value between 3 and 10 is interpreted as moderate evidence in favor of M$_1$ (see also Table \ref{tab:bf_interpretation}).}
+
+\textcolor{blue}{From the marginal likelihoods, we computed Bayes factors for the two-cues model compared to the three-cues model and the one-cue model.
+That is, the marginal likelihood of the two-cues model $\mathrm{M_1}$ was divided by the marginal likelihood of the three-cues model $\mathrm{M_2}$ and one-cue model $\mathrm{M_3}$ to obtain the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
+Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above 10 is interpreted as strong evidence in favor of the two-cues model and a value between 3 and 10 is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).}
+
+\textcolor{blue}{The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
+Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model.
 Therefore, we can conclude that the computational modeling shows that our data shows interference.
@@
 Therefore, the assumption that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue – is moderately supported by these modeling results.
-}
+This result holds for prior assumptions that allows higher probability mass for small values of $F$., i.e., for $F<0.01$.}
 

### 2024-08-12 code_data/LV05_modeling/Generate_posterior_predictions.R
Deleted via the Open Science Framework

@@ -1,380 +1,0 @@
-# author: Pia Schoknecht
-# date: 11.08.2024
-
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(ggbreak)
-library(cowplot)
-library(stringr)
-library(tidyverse)
-
-# save Schoknecht & Vasishth mean, lower and upper separately to add them to plots later
-# add data from Schoknecht & Vasishth ERP experiment
-SVdat <- data.frame(pred   = rep("data",3),
-                  effect    = c("syntactic", "semantic", "interaction"), 
-                  mean  = c(-0.14,-0.3,-0.03),
-                  upper = c(0.05,-0.1,0.38),
-                  lower = c(-0.33,-0.5,-0.33))
-SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
-
-# load simulated data from all 3 models
-load("Simulated-data/Simulated_data_one_cue_model.Rda")
-load("Simulated-data/Simulated_data_two_cues_model.Rda")
-load("Simulated-data/Simulated_data_three_cues_model.Rda")
-
-
-#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-# Load model 
-source("Models/interACT.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-
-# multiply by -1 for negative effects
-simdat_2cues_sameclause$MESyn <- simdat_2cues_sameclause$MESyn * -1
[…]

### 2024-08-12 code_data/LV05_modeling/Prior_predictions-all-models.R
Added via the Open Science Framework

@@ -1,0 +1,391 @@
+# load packages
+library(truncnorm)
+library(ggplot2)
+library(cowplot)
+library(stringr)
+library(tidyverse)
+
+#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
+
+# Load model 
+source("Models/interACT.R")
+
+reset_params()
+
+# Parameters to be passed to run() function for the simulations, including cue weights
+model_4cond <- list(
+  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
+  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
+  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
+  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
+  weights = list(c(strWeight(),semWeight()))
+)
+
+mas <<- c(1.5) # Maximum associative strength
+mp <<- 0.5 # Mismatch penalty
+psc <<- 0 # Prominence scaling
+qcf <<- 1 # Match quality correction factor
+cuesim <<- -1 # Cue-feature similarity [-1..0]
+cueweighting <<- 1 # Cue weighting
+ans <<- 0.2
+VERBOSE <<- FALSE
+
+
+#### Generate prior predictions #####
+
+nsims<-2000
+
+
+lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0.01,sd=0.01)
+
+maineffectSyn<-rep(NA,nsims)
+maineffectSem<-rep(NA,nsims)
+interactionSynSem<-rep(NA,nsims)
+
+for(i in 1:nsims){
+  # Run model
+  print(paste("run: ",i,sep=""))  
+  lf<<- lfs[i]  
+  sims <- create_param_matrix(model_4cond, iterations=2000)
+  results <- run(sims)
+  condMeans <- compute_cond_means(results)
+  ## compute main effect of Syntactic manipulation
+  ## HiSyn-LoSyn:
+  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
+    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
+  ## save the estimate:
+  maineffectSyn[i]<-meSyn
+  ## compute main effect of Semantic manipulation
+  ## HiSem-LoSem:
+  meSem<-mean(c(condMeans$Latency[2],condMeans$Latency[4]))-
[…]

### 2024-08-12 paper/main.tex
Updated via the Open Science Framework

@@
 
-\Copy{figpriorposteriordata}{
+\Copy{fig:model_predictions_ERP}{
 \begin{figure}[htpb]
@@
 
-%\textcolor{blue}{Using the above method, we estimate marginal likelihood for each candidate model under six different prior assumptions.
-The Bayes factor in favor of a model M$_1$ compared to a model M$_2$, i.e., BF$_{12}$, is computed as the ratio of the marginal likelihoods of M$_1$ and M$_2$.
-The Bayes factor BF$_{12}$ represents the extent to which the model M$_1$ is more likely than M$_2$ given the data.
-Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above 10 is interpreted as strong evidence in favor of M$_1$ and a value between 3 and 10 is interpreted as moderate evidence in favor of M$_1$ (see also Table \ref{tab:bf_interpretation}).}
-
 \textcolor{blue}{From the marginal likelihoods, we computed Bayes factors for the two-cues model compared to the three-cues model and the one-cue model.
-That is, the marginal likelihood of the two-cues model $\mathrm{M_1}$ was divided by the marginal likelihood of the three-cues model $\mathrm{M_2}$ and one-cue model $\mathrm{M_3}$ to obtain the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
+That is, the marginal likelihood of the two-cues model $M_1$ was divided by the marginal likelihood of the three-cues model $M_2$ and one-cue model $M_3$ to obtain the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
 Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above 10 is interpreted as strong evidence in favor of the two-cues model and a value between 3 and 10 is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).}

### 2024-08-18 paper/ResponseToReviewers.tex
Updated via the Open Science Framework
fixed journal name error

@@
 Reviewer 2 primarily raised concerns regarding our experimental design.
-In response to these concerns, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textcolor{red}{\textit{Journal of Memory and Language}}.
+In response to these concerns, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textcolor{red}{\textit{Journal of Experimental Psychology: Learning, Memory, and Cognition}}.
 We now stress in the paper that we are building on previously published work by directly comparing the statistical inferences, using Bayes factors, from the present and previous studies that use exactly the same design.
@@
 The following description of the estimates and Bayes factors of the present and previous studies and their interpretation was added to the manuscript, starting page \pageref{comparison}:
-\setcounter{figure}{12}
+\setcounter{figure}{13}
 
@@
 We discuss the plausibility confound and the additional analyses in the revised manuscript starting on page \pageref{plausib_anim_inan}:
+
+\setcounter{figure}{5}
 

### 2024-08-18 paper/ResponseToReviewers.tex
fixed journal name error

@@
 Reviewer 2 primarily raised concerns regarding our experimental design.
-In response to these concerns, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textcolor{red}{\textit{Journal of Experimental Psychology: Learning, Memory, and Cognition}}.
+In response to these concerns, we have emphasized that we used a well-established design that was previously published in, among other journals, the \textit{Journal of Experimental Psychology: Learning, Memory, and Cognition}.
 We now stress in the paper that we are building on previously published work by directly comparing the statistical inferences, using Bayes factors, from the present and previous studies that use exactly the same design.

### 2024-08-18 paper/ResponseToReviewers.tex
fixed journal name error

@@
 \subsection*{Response to E.1}
-We have used a well-studied design that was previously published in the \textcolor{red}{Journal of Memory and Language}, among other places.
+We have used a well-studied design that was previously published in the \textit{Journal of Experimental Psychology: Learning, Memory, and Cognition}, among other places.
 The previously published studies use exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.

### 2024-08-18 paper/ResponseToReviewers.tex
minor rephrasing

@@
 However, we are building on the previously published claim that this design leads to syntactic interference.
-We believe that our results are a valuable contribution to the literature on syntactic interference, and serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
+We believe that our results are a valuable contribution to the literature on syntactic interference, because our work serves as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
 We added the following text to our General Discussion on page \pageref{only_this_design} to emphasize that our findings and conclusions only apply to the present design:
@@
 There was evidence against syntactic interference under all priors in the pre-critical and critical region.
-Only in the spill-over region, there was anecdotal evidence (BF$_{10}$ = 2.8) for a small syntactic interference effect ($\pm$ 8 ms).
+Only in the spill-over region was there anecdotal evidence (BF$_{10}$ = 2.8) for a small syntactic interference effect ($\pm$ 8 ms).
 Similarly, there was evidence against the interaction under all priors in the pre-critical and spill-over regions.
-Only in the critical region, there was anecdotal evidence (BF$_{10}$ = 1.4) for a small interaction effect ($\pm$ 8 ms).
+Only in the critical region was there anecdotal evidence (BF$_{10}$ = 1.4) for a small interaction effect ($\pm$ 8 ms).
 
@@
 
-Taken these Bayes factor results together (at best anecdotal evidence for small effects in the size of an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), we are confident that in our data, there was no evidence for an effect of considerable size of neither syntactic interference nor the interaction.
+Taken these Bayes factor results together (at best anecdotal evidence for small effects in the size of an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), we don't feel that it is defensible to argue that, in our data, there is some evidence for syntactic interference (or for an interaction).
 

### 2024-08-20 paper/ResponseToReviewers.tex
minor wording change

@@
 
-Taken these Bayes factor results together (at best anecdotal evidence for small effects in the size of an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), we don't feel that it is defensible to argue that, in our data, there is some evidence for syntactic interference (or for an interaction).
+Given these Bayes factor results taken together (at best anecdotal evidence for small effects with magnitudes like an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), at least in our data, we don't feel that it is defensible to argue that, there is some evidence for syntactic interference (or for an interaction).
 
@@
 \subsubsection*{Response to R1.3}
-As suggested, we have implemented a null model (no interference), leaving us with a total of three computational models (the null model, the three-cues model by \textcite{mertzen} and our newly proposed two-cues model).
+As suggested, we have implemented three computational models: a null model that assumes no interference, the three-cues model by \textcite{mertzen}, and our newly proposed two-cues model.
 We use Bayes factors to quantitatively compare these models.

### 2024-08-20 paper/main.tex
rephrased text on modeling

@@
 We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference.
-Therefore, it makes no sense to compare the results from the present SPR experiment to predictions from the Lewis and Vasishth (2005) model of sentence processing as skilled memory \textit{retrieval}.
-Instead we use the Lewis and Vasishth (2005) model to generate predictions on the microvolt scale for ERP data.
-For this we adjust the scaling factor $F$ which usually links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R} to link memory activation to brain responses.} The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+Because  Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions.
+However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale.
+In order to model the ERP data, we therefore adjust the scaling factor $F$, which recales the activation of items in memory to the relevant dependent measure.
+The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+
 We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05.

### 2024-08-20 paper/main.tex
added missing commas

@@
 This prior reflects the a priori belief that effect on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
-This relatively uninformative prior was chosen because although there exists no previously published ERP data on the Van Dyke (2007) design it reflects the reasonable range of effect sizes in psycholinguistic ERP research.
+This relatively uninformative prior was chosen because, although there exists no previously published ERP data on the Van Dyke (2007) design, it reflects the reasonable range of effect sizes in psycholinguistic ERP research.
 In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.

### 2024-08-20 paper/main.tex
fixed some wording problems

@@
 We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
-\textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, we can easily assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}.
-This idea goes back to the roots of the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and entails a simple clause tracking mechanism.} The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+\textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, as a simplification, we assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}.
+This idea goes back to an assumption in the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and implies a simple clause tracking mechanism.} The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
 This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.

### 2024-08-20 paper/main.tex
fixed some wording problems

@@
 In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
-\textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}}\label{clause_tracker} \textcolor{blue}{The here proposed model using two cues predicts no syntactic interference and no interaction, but semantic interference effects.
+\textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}}\label{clause_tracker} \textcolor{blue}{The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.

### 2024-08-20 paper/main.tex
added ref

@@
 We computed Bayes factors for the two-cues model compared to the three-cues model and to the one cue model, respectively.
-Since Bayes factors are sensitive to the choice of priors on the parameters of interest, we computed Bayes factors under six different priors for the latency factor parameter $F$,
+Since Bayes factors are sensitive to the choice of priors on the parameters of interest \citep{kass1995bayes}, we computed Bayes factors under six different priors for the latency factor parameter $F$,
 $Normal$(0.01,0.01), 
@@
 $Normal$(0.02,0.02), and 
-$Normal$(0.02,0.005). }
+$Normal$(0.02,0.005).}
 

### 2024-08-20 paper/main.tex
corrected notation and wording

@@
 
-\textcolor{blue}{The procedure for computing Bayes factors is described as follows.
+\textcolor{blue}{The procedure for computing Bayes factors is as follows.
 The Bayes factor computation required us to estimate the marginal likelihood of the competing models.
@@
 \begin{equation}
-ML = \frac{1}{n} \sum_{i=1}{n} \mathcal{L}{F_i|y} p(F_i)  \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
+ML = \frac{1}{n} \sum_{i=1}{n} \mathcal{L}(F_i|y) p(F_i)  \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
 \end{equation}
 
-\noindent \textcolor{blue}{where $y$ is the observed data, i.e., the observed semantic interference effect in our case, $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+\noindent \textcolor{blue}{where $y$ is the observed data, i.e., the observed semantic interference effect in our case, and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
 The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$ and $p(F_i)$ gives the prior density of $F_i$.
@@
 
-\textcolor{blue}{The Lewis and Vasishth (2005) model is a complex, non-deterministic model and it is difficult to express the model's likelihood analytically without drastically simplifying the model \parencite[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}.
-In absence of an analytically-expressed likelihood function, we used an Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood} approach to estimate approximate likelihood for a given value of $F$.
-Under this approach, we compared the model generated data conditioned on $F=F_i$ and the actual observed data $y$.
-If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ was close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
-The likelihood term can be rewritten as $\mathcal{L}{F_i|y} = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represent a density kernel that weighs a proposal $F_i$ based on distance between model-generated data and actual data, $d(x_i,y)$.}
+\textcolor{blue}{The Lewis and Vasishth (2005) model is a complex, non-deterministic model which can be simplified such that a likelihood function represents the model's generative process  \parencite[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}.
+However, it is also possible to carry out model comparison without recasting the model as a likelihood function.
+In this case, model comparison is carried out using Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood}.
+Under this approach, the model-generated data conditional on $F=F_i$ is compared with the actually observed data $y$.
+If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
+The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weighs a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.}
 
@@
 That is, the marginal likelihood of the two-cues model $M_1$ was divided by the marginal likelihood of the three-cues model $M_2$ and one-cue model $M_3$ to obtain the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
-Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above 10 is interpreted as strong evidence in favor of the two-cues model and a value between 3 and 10 is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).}
+Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above $10$ is interpreted as strong evidence in favor of the two-cues model, and a value between $3$ and $10$ is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).}
 

### 2024-08-20 paper/main.tex
minor edit

@@
 \textcolor{blue}{The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
-Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model.
-Therefore, we can conclude that the computational modeling shows that our data shows interference.
+Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model; this comparison shows that there is strong evidence for interference effects.
 Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide moderate evidence for the new two-cues model.
-Therefore, the assumption that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue – is moderately supported by these modeling results.
+This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue.
 This result holds for prior assumptions that allows higher probability mass for small values of $F$., i.e., for $F<0.01$.}

### 2024-08-21 paper/main.tex
fixed english

@@
 Then, using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method.
-Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we run on their data analog to our own analyses described in Statistical Analyses for Experiments 1a and b.
+Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we run on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
 The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}

### 2024-08-21 paper/main.tex
fixed english

@@
 \textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task.
-The trial id of the 100 trials per session was recoded to span from 0 to 1.
+The trial id of the 100 trials per session was rescaled to span from 0 to 1.
 This was done to bring all fixed effects to a comparable scale which decreases the run time of the models.
@@
 In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
-However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference lead to slower reading times independent of plausibility.}
+However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference leads to slower reading times independent of plausibility.}
 

### 2024-08-21 paper/ResponseToReviewers.tex
rephrased response 2.1

@@
 The previously published studies used exactly the same design, specifically the same syntactic retrieval cue, and found effects of syntactic interference \citep{vandyke07,mertzen}.
-While the syntactic manipulation in the present design might not be appropriate, it was previously claimed in publications that this design leads to syntactic interference.
-Our results are a valuable contribution to the literature on syntactic interference as they serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
-The question which exact syntactic manipulation leads to interference could be the research focus of future work, but it is outside of the scope of the present manuscript.
+The reviewer may well be right that the syntactic manipulation in the present design may not be appropriate; however, it was previously claimed in publications that this design leads to syntactic interference.
+Our results contribute to the literature on syntactic interference as they serve as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
+The reviewer's question--which exact syntactic manipulation leads to interference--is an important concern that could be the research focus of future work, but it is outside of the scope of our investigation in the present manuscript.
 

### 2024-08-21 paper/ResponseToReviewers.tex
edited wording in response to R2.2
minor changes in wording to 3.1 response

@@
 \subsubsection*{Response to R2.2}
-We agree with this statement and therefore refrained from interpreting the results at the critical region.
-To emphasize this point, we added the following to the manuscript on page \pageref{only_precritical}:\\
+We agree with the reviewer. We now refrain from interpreting the results at the critical region. To emphasize this point, we added the following to the manuscript on page \pageref{only_precritical}:\\
 
@@
 \subsubsection*{Response to R2.4}
-We agree with the Reviewer that visual inspection of Figure \ref{fig:whole_sentence} suggests an interaction, but the Bayes factor analysis, i.e., hypothesis testing, provided no evidence for the interaction.
+We agree with the reviewer that visual inspection of Figure \ref{fig:whole_sentence} suggests an interaction, but the Bayes factor analysis, i.e., a formal model comparison, provided no evidence for the interaction.
 Therefore, we do not discuss this visual impression.
@@
 \subsubsection*{Response to R3.1}
-It is possible that the subject of the current clause receives an activation boost but the mechanism for that is not trivial.
+The reviewer's proposal could well be right, that the subject of the current clause receives an activation boost.
+However, the mechanism for implementing this idea computationally might not be trivial.
 Let's assume that the subject is encoded with boosted activation compared to for example an object.
@@
 At one point within a long sentence, the subject's activation will be lower than the activation of other nouns in memory.
-If we assume that a long sentence has multiple embedded sentences, we would need to assume that either the matrix sentence subject activation is downgraded when an embedded sentence is entered or that the embedded sentence subject gets an even higher activation boost than the matrix sentence subject.
+If we assume that a long sentence has multiple embedded clauses, we would need to assume that either the matrix sentence subject activation is downgraded when an embedded sentence is entered or that the embedded sentence subject gets an even higher activation boost than the matrix sentence subject.
 Then when the embedded sentence is finished and we return back to the matrix sentence, the matrix sentence subject would need to receive another activation boost to return to its special status.
 Essentially, this proposal condenses down to a clause tracking mechanism with a modulation of activation levels.
-We think it is more parsimonious to implement the clause tracking in the memory encoding and retrieval operations than the activation levels.
+In our opinion, it is more parsimonious to implement the clause tracking in the memory encoding and retrieval operations than the activation levels.
 

### 2024-08-21 paper/ResponseToReviewers.tex
wording change to R3.4 response

@@
 We do not claim that our results are correct and previous claims have been false.
-Progress in science is achieved by carrying out adequately powered studies in the future that can
+We feel that progress in sentence processing research will be greatly facilitated by carrying out adequately powered studies in the future that can
 distinguish between theoretical claims.
-The replication crisis \citep{openscience2015_reproducibility} has shown that contradiction between results even when using the same design is not an anomaly; if anything, it's quite the opposite.
+Our work aims to be a contribution in this direction.
+As discussed earlier in this response letter, the contradiction between published results even when using the same design (and even the same items) is not an anomaly in sentence processing research.
 

### 2024-08-21 paper/main.tex
fixed formatting
Resolved Shravan's comments on modeling section

@@
 We used the Monte Carlo integration method to estimate the marginal likelihood of each model.
-First, we chose an importance density for the parameter $F$ from which we draw $0.2$ million samples of $F$.
+For Monte Carlo integration, we use the importance sampling estimator where a large number of number proposals are obtained from an importance density $q(F)$ and the average $likelihood \times prior$ is calculated across all these proposals to obtain the approximate marginal likelihood.
+The importance density should be chosen such that it covers the entire parameter space where $likelihood \times prior$ is non-zero and does not draw too many samples from the space where $likelihood \times prior$ is zero.
+We estimated the marginal likelihoods using Monte Carlo importance sampling as follows.
+
+First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we use a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
+This importance density allows us to obtain more proposals for smaller values of $F$ (for F<0.03) while maintaining fatter tails for the larger values of $F$ ($F>0.03$); effectively, the whole parameter space of $F$ between $0$ and $0.05$ is explored to draw $0.2$ million proposals.
 Then, for each of these proposed parameter values, we compute the likelihood and prior density.
@@
 \begin{equation}
-ML = \frac{1}{n} \sum_{i=1}{n} \mathcal{L}(F_i|y) p(F_i)  \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
+ML = \frac{1}{n} \sum_{i=1}{n} \frac{\mathcal{L}(F_i|y) p(F_i)}{q(F_i)} \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
 \end{equation}
 
-\noindent \textcolor{blue}{where $y$ is the observed data, i.e., the observed semantic interference effect in our case, and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
-The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$ and $p(F_i)$ gives the prior density of $F_i$.
+\noindent \textcolor{blue}{where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case, and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.
 We compute marginal likelihoods under different prior assumptions $p(F)$.}
@@
 If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
-The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weighs a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.}
+The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weighs a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.
+We choose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
+The $\delta$ parameter determines how good is the approximation: the lower is the value of the $delta$, the better is the approximation of the likelihood.
+However, if $\delta$ is too low, the computational demand increases because too many proposals would end up with zero likelihood.
+We choose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising the computational efficiency.}
 

### 2024-08-22 paper/ResponseToReviewers.tex
added ending to the response

@@
 
+We agree with the reviewer regarding their overall conclusion, that there may be multiple ways to explain the absence of syntactic interference, and multiple reasons why our results differ from previous studies.
+Our opinion is that the present work still has value because it attempts to replicate  well-known results in sentence processing, uses a method different than self-paced reading and eye-tracking (EEG) to investigate these effects, and proposes one computational implementation that could explain the EEG results.
+
 \end{document}

### 2024-08-22 paper/main.tex
minor wording change

@@
 If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
-The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weighs a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.
+The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weights a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.
 We choose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
-The $\delta$ parameter determines how good is the approximation: the lower is the value of the $delta$, the better is the approximation of the likelihood.
+The $\delta$ parameter determines how good the approximation is: the lower the value of the $delta$, the better the approximation of the likelihood.
 However, if $\delta$ is too low, the computational demand increases because too many proposals would end up with zero likelihood.

### 2024-08-22 paper/main.tex
minor wording change

@@
 
-\noindent \textcolor{blue}{where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case, and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+\noindent \textcolor{blue}{where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
 The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.

### 2024-08-22 paper/main.tex
Updated via the Open Science Framework

@@
 However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale.
-In order to model the ERP data, we therefore adjust the scaling factor $F$, which recales the activation of items in memory to the relevant dependent measure.
+In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure.
 The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+}
 
@@
 We estimated the marginal likelihoods using Monte Carlo importance sampling as follows.
-
-First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we use a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
+}
+
+\textcolor{blue}{First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we use a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
 This importance density allows us to obtain more proposals for smaller values of $F$ (for F<0.03) while maintaining fatter tails for the larger values of $F$ ($F>0.03$); effectively, the whole parameter space of $F$ between $0$ and $0.05$ is explored to draw $0.2$ million proposals.

### 2024-08-23 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-\noindent \textbf{Prof. Chris Jarrold}\\
+\noindent \textbf{Prof. Jarrold}\\
 \noindent \textbf{Associate Editor}\\
@@
 We are simply reporting what we found.
-Contradictory results have been reported for many phenomena: e.g., reflexives processing \citep{jaeger_etal_2020}, main verb/reduced relative garden paths (\textcolor{red}{REF?}), predictive parsing effects \citep{nieuwland_etal_2018}, local coherence \citep{lcpaape2023}.
+Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.

### 2024-08-24 paper/main.tex

@@
 Lastly, we implemented a model with one cue which unambiguously matches the subject.
-This cue would be the master composite cue \{$\pm$ animate-subject-in-same-clause\}.
+This cue could be the composite cue \{$\pm$ animate-subject-in-same-clause\} or it could be even more specific.
 This model naturally does not predict any interference effects.
@@
 
-\Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature \citep[see e.g.,][for a high-powered failed replication attempt of a classic result in the area of predictive processing]{nieuwland_etal_2018}.
+\Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature.
+Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}

### 2024-08-24 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

--- 2024-08-22 paper/ResponseToReviewers.tex
@@
 \usepackage{rotating}
+\usepackage{booktabs}
 \usepackage{amssymb}
@@
 We are simply reporting what we found.
-Contradictory results have been reported for many phenomena: e.g., reflexives processing \citep{jaeger_etal_2020}, main verb/reduced relative garden paths (\textcolor{red}{REF?}), predictive parsing effects \citep{nieuwland_etal_2018}, local coherence \citep{lcpaape2023}.
+Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.
@@
 \subsubsection*{Response to R1.8}
-BFs do not to need to be strongly in favor of an effect across all investigated priors to conclude that an effect is supported.
+We agree that syntactic interference and/or the interaction could be real, but our data does not provide conclusive evidence for them.
+Because it can be difficult to discern small differences in the Bayes factor plots (Figures 5 and 11 in Response to E.2), the Table below shows all Bayes factors for syntactic interference and the interaction.
+All Bayes factors above 1 are highlighted in bold.
+Following \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we regard Bayes factors around 1 as providing no evidence.
+The only Bayes factor remarkably above 1 is BF$_{10}$ = 2.8 for the syntactic interference effect in the reading times of the spill-over region.
+Because of the pre-critical effects in the reading times, this effect in the spill-over region is confounded.
+Therefore, we think that although the Bayes factor theoretically provided anecdotal evidence for syntactic interference in the spill-over region, there is no \textit{decisive} evidence for syntactic interference in our data.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{ccccc}
+\toprule
+Experiment &      Region   &   Effect  &         Prior & BF10\\
+\midrule
+SPR & pre-critical&   syntactic &Normal(0, 0.01)&0.5\\
+SPR & pre-critical&   syntactic &Normal(0, 0.05)& 0.11\\
+SPR & pre-critical&   syntactic  &Normal(0, 0.1)& 0.06\\
+SPR & critical   &syntactic &Normal(0, 0.01)&0.5\\
+SPR & critical   &syntactic &Normal(0, 0.05)& 0.11\\
+SPR & critical   &syntactic  &Normal(0, 0.1)&0.05\\
+SPR & spill-over  & syntactic &Normal(0, 0.01)& \textbf{2.8}\\
+SPR & spill-over  & syntactic &Normal(0, 0.05)& 0.8\\
+SPR & spill-over  & syntactic  &Normal(0, 0.1)&0.42\\
+
+EEG & N400   &syntactic &Normal(0, 0.1)&\textbf{1.18}\\
+EEG & N400   &syntactic &Normal(0, 0.5)&0.52\\
+EEG &  N400  & syntactic &  Normal(0, 1)&0.27\\
+EEG & N400   &syntactic  & Normal(0, 2)&0.14\\
+EEG & P600  & syntactic &Normal(0, 0.1) & 0.91\\
+EEG & P600  & syntactic &Normal(0, 0.5) &0.27\\
+EEG &  P600 &  syntactic &  Normal(0, 1)&0.14\\
+EEG & P600  & syntactic  & Normal(0, 2) & 0.07\\
+
+SPR & pre-critical &interaction &Normal(0, 0.01) & 1\\
+SPR & pre-critical &interaction &Normal(0, 0.05) & 0.36\\
+SPR & pre-critical &interaction & Normal(0, 0.1) & 0.18\\
+SPR & critical &interaction &Normal(0, 0.01) & \textbf{1.36}\\
+SPR & critical &interaction &Normal(0, 0.05) & 0.71\\
+SPR & critical &interaction & Normal(0, 0.1) & 0.37\\
+SPR & spill-over& interaction& Normal(0, 0.01) &0.92\\
+SPR & spill-over &interaction &Normal(0, 0.05) & 0.3\\
+SPR & spill-over &interaction  &Normal(0, 0.1) & 0.15\\
+
+EEG & N400 &interaction &Normal(0, 0.1)&0.88\\
+EEG & N400 &interaction &Normal(0, 0.5)&0.36\\
+EEG & N400 &interaction  & Normal(0, 1)&0.19\\
+EEG & N400 &interaction  & Normal(0, 2) & 0.1\\
+EEG & P600 &interaction &Normal(0, 0.1)& \textbf{1.16}\\
+EEG & P600 &interaction &  Normal(0, 1) &0.43\\
+EEG & P600 &interaction &  Normal(0, 2) &0.22\\
+\bottomrule
+\end{tabular}
+
+\end{table}
+
+%BFs do not to need to be strongly in favor of an effect across all investigated priors to conclude that an effect is supported.
 However, to draw decisive conclusions in favor of an effect, the evidence for that effect should be stronger than anecdotal for at least a subset of reasonable priors.
 
+Regarding the choice of priors, we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times.
 The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}).
-Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}), we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times.
+Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}).
 
@@
 The results of the plausibility norming, see Figure \ref{fig:plausibility_rating} on page \pageref{fig:plausibility_rating} of the manuscript illustrate that whatever semantic feature determines a good fit between subject and verb matched with the features of the subject and the animate distractor, but not with the features of the inanimate distractor.
+For a principled approach to feature selection within the cue-based retrieval framework, see \textcite{smith2020principled}.
 

### 2024-08-24 paper/main.tex
Updated via the Open Science Framework

@@
 While more generic features like \{+animate\} or \{+human\} (all animate distractors were human) might be created during memory encoding, more specific features like \{+can drink\} are probably only activated when relevant.
-The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval (for work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \cite{vandyke_mcelree06, cunnings_sturt_2018retrieval}).
+The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval.
+For experimental work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \textcite{vandyke_mcelree06, cunnings_sturt_2018retrieval}.
+For a principled approach to feature selection using word embeddings, see \textcite{smith2020principled}.
 A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
@@
 \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
-In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5).
+In the critical region, the evidence for semantic interference was weak (Normal(0, 0.01): BF$_{10}$ = 3.5,
+Normal(0, 0.05): BF$_{10}$ = 1.3,
+Normal(0, 0.1): BF$_{10}$ = 0.7).
 There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
@@
 
-\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but inconclusive results in the critical region.
-In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}
+\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
+In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}}
 
@@
 \end{figure}
-}}
+}
 
@@
 We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-\textcolor{blue}{Reading times of the pre-critical word and critical verb showed evidence against syntactic interference.
+\textcolor{blue}{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
+This means that semantic interference occurred even before the retrieval event could begin.
+Reading times of the pre-critical word and critical verb showed evidence against syntactic interference.
 In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms.
-We will discuss the almost complete absence of syntactic interference in our data in the general discussion.
-The reading time results at the critical verb were inconclusive for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
-This means that semantic interference occurred even before the retrieval event could begin.
 Reading times at the critical verb showed very weak support for a small interaction of syntactic and semantic interference.
 There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
-Therefore, we refrain from discussing effects occurring in these regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.
+This means that although it is interesting that small effects of syntactic interference and the interaction emerged only in these regions  any interpretation of these effects would be purely speculative.
+Therefore, we refrain from discussing the effects occurring in the critical and spill-over regions.}}\label{only_precritical} \textcolor{blue}{We will discuss the absence of decisive evidence for syntactic interference in our data in the general discussion.} The effects in the pre-critical region are discussed next.
 
@@
 Although plausibility influenced reading times (higher plausibility lead to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms).
-The Bayes factors provided evidence for semantic interference in this analysis under all priors (Normal(0, 0.01): BF$_{10}$ = 4.2,
+The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
 Normal(0, 0.05): BF$_{10}$ = 2.4,
@@
 The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference.
-\textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced P600 amplitude for high vs.\ low semantic interference.
-See Figure \ref{fig:erp_sem_syn} (c) for the topographic distribution of this effect.} \Copy{ERP_results2}{In contrast, Figure \ref{fig:erp_sem_syn} (d), (e) and (f) show that syntactic interference did not considerably affect the brain response to the critical verb in \textcolor{blue}{neither the N400 nor P600 spatio-temporal windows}.}\label{ERP_results2} Visual inspection of the ERPs in \ref{fig:erp_sem_syn} (d) does also not suggest that there was a syntactic interference effect in any other time window.
+\textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced amplitude for high vs.\ low semantic interference in the P600 time window.
+Figure \ref{fig:erp_sem_syn} (c) shows that this effect had a broad distribution.} \Copy{ERP_results2}{\textcolor{blue}{Figure \ref{fig:erp_sem_syn} (d) shows that syntactic interference affected the brain response to the critical verb in the N400 and P600 spatio-temporal windows to a smaller extend than semantic interference did.
+Figures \ref{fig:erp_sem_syn} (d) and (e) show that high vs.\ low syntactic interference lead to a slightly more negative ERP response with a broad distribution in the N400 time window.
+Figure  \ref{fig:erp_sem_syn} (f) shows a slightly reduced P600 amplitude for high vs.\ low syntactic interference with a centro-parietal distribution.
+}}\label{ERP_results2}
 
@@
 Normal(0, 2): BF$_{10}$ = 5).
-In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and very weak evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$.
-By contrast, Bayes factors provided evidence against the syntactic interference effect in both spatio-temporal windows (BF$_{10}$ $<$ 0.9).
-The only exception was under the narrowest prior in the N400 window, there was very weak evidence for syntactic interference (BF$_{10}$ = 1.2).
-Similarly, Bayes factors provided evidence against the interaction in both spatio-temporal windows.
-The only exception was under the narrowest prior in the P600 window, there was very weak evidence for the interaction (BF$_{10}$ = 1.2).
+In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and no evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$.
+By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatio-temporal windows (BF$_{10}$ $<$ 1.2).
+Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF$_{10}$ $<$ 1.2).
 In sum, in our ERP data, there was decisive evidence for the semantic interference effect and mainly evidence against syntactic interference and the interaction.}}\label{ERP_results3}
@@
 low semantic interference.
-In contrast, there was only very weak evidence for very small effects of syntactic interference and the interaction or even evidence against them in both the N400 and P600 windows.
+In contrast, there was no evidence for effects of syntactic interference and the interaction or even evidence against them in both the N400 and P600 windows.
 In sum, the ERP results showed clear evidence for semantic interference and mostly evidence against syntactic interference and the interaction of syntactic and semantic interference.
@@
 The reduced P600 amplitude under high semantic interference in the present study might be explained by facilitatory interference \citep[see e.g.,][]{jaeger_etal_2017}.
-In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to act as the subject.
+In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to function as the subject.
 A similar effect was found by \textcite{Tanner_etal_2017}.
@@
 
-Consistent with the self-paced reading study, there is clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}.
+\textcolor{blue}{Similar to the self-paced reading results,} there is clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}.
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
@@
 Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
-We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference.
+We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
 Because  Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions.
@@
 In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure.
-The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
-}
+The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.}
 
 We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05.
-This prior reflects the a priori belief that effect on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
-This relatively uninformative prior was chosen because, although there exists no previously published ERP data on the Van Dyke (2007) design, it reflects the reasonable range of effect sizes in psycholinguistic ERP research.
-In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
+This prior reflects the a priori belief that effects on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
+This relatively uninformative prior was chosen because, although there exists no previously published ERP data on the Van Dyke (2007) design, it reflects the reasonable range of effect sizes in psycholinguistic ERP research.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
 \textcolor{blue}{We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study.
-To reflect that larger predicted effect sizes correspond to more negative N400 amplitudes, the predicted effects from the model} were multiplied with $-1$.
+Since the scaling parameter $F$ cannot be negative, an additional step was needed to reflect that larger predicted effect sizes correspond to more negative N400 amplitudes: The predicted effects from the model} were multiplied with $-1$.
+\textcolor{blue}{In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.}
 
@@
 This model predicted approximately equally sized effects of syntactic and semantic interference.
-Although it is clear that our data does not support equally sized effects of syntactic and semantic interference, we decided to include their version of the model here because it has previously been claimed to be a good model for the present design.
+Although it is clear that our ERP data does not support equally sized effects of syntactic and semantic interference, we decided to include their version of the model here because it has previously been claimed to be a good model for the present design.
 We have implemented two alternative versions of the model.
-Given the evidence against syntactic interference in our data, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
+Given that in our ERP data, there was no evidence for syntactic interference or depending on the a priori assumed effect size even evidence against it, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
 We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
@@
 \textcolor{blue}{To quantitatively evaluate the relative predictive performances of the models on the semantic interference data, we used a Bayes factor analysis.
-We computed Bayes factors for the two-cues model compared to the three-cues model and to the one cue model, respectively.
-Since Bayes factors are sensitive to the choice of priors on the parameters of interest \citep{kass1995bayes}, we computed Bayes factors under six different priors for the latency factor parameter $F$,
+We computed Bayes factors for the two-cues model compared to the three-cues model and to the one-cue model, respectively.
+Since Bayes factors are sensitive to the choice of priors on the parameters of interest \citep{kass1995bayes}, we computed Bayes factors under six different priors for the scaling parameter $F$,
 $Normal$(0.01,0.01), 
@@
 Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model; this comparison shows that there is strong evidence for interference effects.
-Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide moderate evidence for the new two-cues model.
+Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior.
 This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue.
@@
 
-\Copy{only_this_design}{Taken together, the SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and no effects of syntactic interference \textcolor{blue}{or an interaction}.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
+\Copy{only_this_design}{Taken together, \textcolor{blue}{present} the SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and \textcolor{blue}{little or no effects of syntactic interference or an interaction}.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
@@
 Then, using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method.
-Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we run on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
+Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
 The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}
@@
 \citeauthor{mertzen}'s (\citeyear{mertzen}) experiments showed weak evidence for the effect.
-The two self-paced reading experiments provided evidence against it.
-Similarly, there is no clear pattern for semantic interference: Most experiments provided evidence against it, but the two self-paced reading experiments (present and \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1) provided evidence for semantic interference.
+The two self-paced reading experiments (present and \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1) provided evidence against it.
+Similarly, there is no clear pattern for semantic interference: Most experiments provided evidence against it, but the two self-paced reading experiments provided evidence for semantic interference.
 The Bayes factors across studies provided evidence against the interaction.}
@@
 \Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature.
-Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
+Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{taboretal04, lcpaape2023}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}
@@
 
-The evidence against a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
+The \textcolor{blue}{lack of decisive evidence for} a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
 \Copy{future_work}{\textcolor{blue}{A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
@@
 First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}.
-Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite[e.g.,][]{ALV2019,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \parencite[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and
+Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite[e.g.,][]{Avetisyan_etal_2020_agree,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \parencite[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and
 the importance of word order \parencite[e.g.,][]{meng2000mode}.
@@
 The observed differences in the time course of the effect are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+Overall, we found \textcolor{blue}{no decisive evidence for} the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
 A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.

### 2024-08-26 code_data/LV05_modeling/Compute-Bayes-factors.R
Updated via the Open Science Framework

@@ -7,6 +7,7 @@ library(ggplot2)
 library(cowplot)
 library(stringr)
 library(dplyr)
+library(tidyr)
 
 #############
 SVdat <- data.frame(pred   = rep("data",3),
@@ -34,11 +35,11 @@ df.lkl$prior <- "normal(0.01,0.01)"
 df.bf <- as.data.frame(df.lkl %>% group_by(prior) %>% 
                          summarize(MargLik=mean(ML)))
 df.bf
-df.bf$Model <- "Two cues model"
+df.bf$Model <- "Two-cues model"
 
 load("Simulated-data/Simulated_data_three_cues_model.Rda")
 df.lkl <- simdat_3cues
-df.lkl$Model <- "Three cues model"
+df.lkl$Model <- "Three-cues model"
 df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
 df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
 df.lkl$proposal.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0,sd=0.025)
@@ -51,7 +52,7 @@ df.bf
 
 load("Simulated-data/Simulated_data_one_cue_model.Rda")
 df.lkl <- simdat_1cue
-df.lkl$Model <- "One cue model"
+df.lkl$Model <- "One-cue model"
 df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
 df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
 df.lkl$proposal.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0,sd=0.025)
@@ -67,7 +68,7 @@ df.bf
 
 load("Simulated-data/Simulated_data_two_cues_model.Rda")
 df.lkl <- simdat_2cues_sameclause
-df.lkl$Model <- "Two cues model"
+df.lkl$Model <- "Two-cues model"
 df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
 df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.02)
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/generate_kfold_posterior_predictions.R
Deleted via the Open Science Framework

@@ -1,433 +1,0 @@
-# author: Pia Schoknecht
-# date: 23.07.2024
-
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(ggbreak)
-library(cowplot)
-library(stringr)
-library(tidyverse)
-
-# load k-fold model fits
-kfold.estimates <- read.csv("kfold_N400_estimates.csv")
-estimates <- kfold.estimates
-
-# save Schoknecht & Vasishth mean, lower and upper separately to add them to plots later
-# add data from Schoknecht & Vasishth ERP experiment
-SVdat <- data.frame(pred   = rep("data",3),
-                  effect    = c("syntactic", "semantic", "interaction"), 
-                  mean  = c(-0.14,-0.3,-0.03),
-                  lower = c(0.05,-0.1,0.38),
-                  upper = c(-0.33,-0.5,-0.33))
-SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
-
-
-#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-# Load model 
-source("interACT.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-load("simdat_2cues_sameclause.Rda")
-
-# effect sizes in simulated data
-round(c(mean = mean(simdat_2cues_sameclause$MESyn), quantile(simdat_2cues_sameclause$MESyn, probs = c(.025, .975))),2)
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/generate_posterior_predictions.R

@@ -1,381 +1,0 @@
-# author: Pia Schoknecht
-# date: 23.07.2024
-
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(ggbreak)
-library(cowplot)
-library(stringr)
-library(tidyverse)
-
-# save Schoknecht & Vasishth mean, lower and upper separately to add them to plots later
-# add data from Schoknecht & Vasishth ERP experiment
-SVdat <- data.frame(pred   = rep("data",3),
-                  effect    = c("syntactic", "semantic", "interaction"), 
-                  mean  = c(-0.14,-0.3,-0.03),
-                  lower = c(0.05,-0.1,0.38),
-                  upper = c(-0.33,-0.5,-0.33))
-SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
-
-
-#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-# Load model 
-source("interACT.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-load("simdat_2cues_sameclause.Rda")
-
-# multiply by -1 for negative effects
-simdat_2cues_sameclause$MESyn <- simdat_2cues_sameclause$MESyn * -1
-simdat_2cues_sameclause$MESem <- simdat_2cues_sameclause$MESem * -1
-simdat_2cues_sameclause$InteractionSynSem <- simdat_2cues_sameclause$InteractionSynSem * -1
-
-# effect sizes in simulated data
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/generate_prior_predictions.R

@@ -1,331 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-library(tidyverse)
-
-
-#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-# Load model 
-source("interACT.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-#### Generate prior predictions #####
-
-# Generate 100,000 simulations from model:
-nsims<-100000
-
-## Initial hyperparameters for latency factor
-## Visualization of uniform prior
-latency_factor_prior <-data.frame(x_temp=runif(1000000,min=0.005,max=0.04))
-hist(latency_factor_prior$x_temp, freq=FALSE)
-
-prior_lf <-ggplot(latency_factor_prior,aes(x=x_temp))+
-  geom_histogram(aes(y=..density..),
-                 position="identity",fill="gray", color="black",
-                 binwidth=0.003)+
-  theme_bw()+
-  theme(strip.text.x = element_text(size = 16, colour = "black", angle = 0))+
-  xlab("latency factor")+
-  ggtitle("Prior on scaling factor")
-prior_lf
-ggsave("LF_Prior.png", width=5, height=3, dpi=600)
-
-## uniform prior (effect between 0.5 and 4 muV):
-lfs<-runif(nsims,min=0.005,max=0.04)
-
-## check:
-hist(lfs,freq=FALSE)
-
-
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/interACT-threecues.R

@@ -1,240 +1,0 @@
-library(tidyr)
-library(dplyr)
-
-# Original code by Felix Engelmann (https://github.com/felixengelmann/inter-act)
-
-###############################################################
-## Global parameters
-###############################################################
-
-reset_params <- function(){
-  lf        <<- 0.15       # latency factor
-  le        <<- 1       # latency exponent
-  rth 	    <<- -1.5    # retrieval threshold
-  bll       <<- 0.5     # decay parameter
-  ans       <<- 0.2    # activation noise
-  mas       <<- 1       # maximum associative strength 
-  mp        <<- 1       # mismatch penalty
-  ga        <<- 1       # goal source activation
-  rand_time <<- 3       # latency variability
-  blc       <<- 0       # base-level activation
-  ##
-  lp        <<- 1     # default time since last presentation (msec)
-  ldp       <<- 1       # last distractor presentation (msec)
-  ## Distractor control:
-  ndistr    <<- 1				# number of distractors
-  dbl       <<- 0       # distractor base-level
-  ##
-  ## Cue weighting, cue confusion and prominence:
-  cueweighting <<- 1     # Strength of structural cue as ratio str/sem
-  normalizeWeights <<- TRUE 
-  qcf 		  <<- 1 			# match quality correction factor
-  qco 		  <<- -2*rth 			# match quality correction offset
-  psc       <<- 1       # prominence scaling constant C1
-  pic       <<- 0       # prominence scaling constant C2
-  tprom     <<- 0       # target prominence 
-  dprom     <<- 0       # distractor prominence
-  # cl 				<<- 0 			# cue confusion level (0-100)
-  cuesim    <<- -1    # cue-feature similarity [-1..0]
-  #
-  VERBOSE <<- TRUE
-}
-reset_params()
-
-idnames <- c("Set","Iteration","Condition","Target","Distractor")
-actrnames <- c("weights","bl1","bl2","times1","times2","noise1","noise2","blact1","blact2","act1","act2","activation","latency","retrieved","acc","miss","fail")
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/model_comparison.R

@@ -1,325 +1,0 @@
-# author: Pia Schoknecht
-# date: 30.07.2024
-
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-library(tidyverse)
-
-# load posterior predictions from all 3 models
-
-load("posterior_predictions_3cues_kfold.Rda")
-load("posterior_predictions_2cues_kfold.Rda")
-load("posterior_predictions_1cue_kfold.Rda")
-
-####################
-
-#### 3 cues model ####
-
-# prepare data
-# combine all kfold sets
-cues3_postpreds_wo[[1]]$set_id <- 1
-cues3_postpreds_wo[[2]]$set_id <- 2
-cues3_postpreds_wo[[3]]$set_id <- 3
-cues3_postpreds_wo[[4]]$set_id <- 4
-cues3_postpreds_wo[[5]]$set_id <- 5
-cues3_postpreds_wo[[6]]$set_id <- 6
-cues3_postpreds_wo[[7]]$set_id <- 7
-cues3_postpreds_wo[[8]]$set_id <- 8
-cues3_postpreds_wo[[9]]$set_id <- 9
-cues3_postpreds_wo[[10]]$set_id <- 10
-
-postpreds_cues3 <- rbind(cues3_postpreds_wo[[1]], cues3_postpreds_wo[[2]], cues3_postpreds_wo[[3]], cues3_postpreds_wo[[4]], cues3_postpreds_wo[[5]],
-                         cues3_postpreds_wo[[6]], cues3_postpreds_wo[[7]], cues3_postpreds_wo[[8]], cues3_postpreds_wo[[9]], cues3_postpreds_wo[[10]])
-
-#rename columns
-postpreds_cues3 <- postpreds_cues3 %>% rename(ypred.syntactic = "syntactic", ypred.semantic = "semantic", ypred.interaction = "interaction")
-
-# multiply by -1 for negative effects
-postpreds_cues3$ypred.syntactic <- postpreds_cues3$ypred.syntactic * -1
-postpreds_cues3$ypred.semantic <- postpreds_cues3$ypred.semantic * -1
-postpreds_cues3$ypred.interaction <- postpreds_cues3$ypred.interaction * -1
-
-# load test data
-kfold.estimates <- read.csv("kfold_N400_estimates.csv")
-#test_data <- kfold.estimates[,c(1,8:13)]
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/predictive-density-functions.R

@@ -1,121 +1,0 @@
-# adapted script from Himanshu Yadav
-# downloaded from https://osf.io/f4tkd/?view_only=
-
-#####################################
-# Using only mean of test data
-
-abc_elpd <- function(posterior_samples,kfold_data,tol_scaler){
-  ytest <- as.data.frame(kfold_data %>% group_by(set_id,dataset) %>% 
-                           summarise(mean_test_syntactic=mean(test.syntactic),
-                                     mean_test_semantic=mean(test.semantic),
-                                     mean_test_interaction=mean(test.interaction),
-                                     tol_syn=sd(test.syntactic)/tol_scaler,
-                                     tol_sem=sd(test.semantic)/tol_scaler,
-                                     tol_int=sd(test.interaction)/tol_scaler))
-  
-  df.lpd <- data.frame(matrix(ncol=3,nrow=nrow(ytest)))
-  colnames(df.lpd) <- c("set_id","testset","ELPD")
-  
-  
-  
-  for(k in 1:nrow(ytest)){
-    ypred_mk.syntactic <- subset(posterior_samples,set_id==ytest[k,]$set_id)$xsim.syntactic
-    ypred_mk.semantic <- subset(posterior_samples,set_id==ytest[k,]$set_id)$xsim.semantic
-    ypred_mk.interaction <- subset(posterior_samples,set_id==ytest[k,]$set_id)$xsim.interaction
-    
-    lpd_k <-   log(mean(dnorm((ytest[k,]$mean_test_syntactic-ypred_mk.syntactic),0,ytest[k,]$tol_syn)*
-                        dnorm((ytest[k,]$mean_test_semantic-ypred_mk.semantic),0,ytest[k,]$tol_sem))*
-                        dnorm((ytest[k,]$mean_test_interaction-ypred_mk.interaction),0,ytest[k,]$tol_int))
-    
-    df.lpd[k,] <- c(ytest[k,]$set_id,ytest[k,]$dataset,lpd_k)
-  }
-  df.lpd
-}
-
-
-
-abc_elpd2 <- function(posterior_samples,kfold_data,tol){
-  ytest <- as.data.frame(kfold_data %>% group_by(set_id,dataset) %>% 
[…]

### 2024-08-26 code_data/LV05_modeling/brain_response/run_kfold_models.R

@@ -1,105 +1,0 @@
-library(tidyverse)
-library(brms)
-
-# load data
-eeg <- read.csv("../eeg/data_eeg.csv")
-
-# create 10 folds
-# repeat the numbers 1:0 and add it as set_id to data frame
-id <- rep(c(1:10), times=1197)
-id2 <- id[1:11967]
-eeg$set_id <- id2
-
-# checks
-eeg %>% group_by(set_id) %>%
-  distinct(subject) %>%
-  count()
-# --> each set includes all subjects
-
-eeg %>% group_by(set_id) %>%
-  distinct(item) %>%
-  count()
-# --> each set includes all items
-
-eeg %>% group_by(set_id) %>%
-  distinct(cond) %>%
-  count()
-# --> each set includes all conditions
-
-
-# run model for each fold
-
-# symmetrical priors 
-priors_m <- c(
-  prior(normal(0, 5), class = Intercept),
-  prior(normal(0, 2), class = b, coef="amplitude_prestim"),
-  prior(normal(0, 0.5), class = b, coef="syn"),
-  prior(normal(0, 0.5), class = b, coef="sem"),
-  prior(normal(0, 0.5), class = b, coef="syn:sem"),
-  prior(normal(10, 5), class = sigma),
-  prior(normal(0, 2), class = sd)
-)  
-
-for(k in 1:10){
-  m_wo_train <- brm(amplitude_n400 ~ 1 + amplitude_prestim + syn*sem + (1+syn*sem||subject) + (1+syn*sem||item),
-                     prior = priors_m,
-                     warmup = 2000,
-                     iter = 10000,
-                     cores = 4,
-                     control = list(adapt_delta = 0.9),
-                     save_pars = save_pars(all = TRUE),
-                     data = subset(eeg, set_id != k),
-                     sample_prior="yes")
-  save(m_wo_train, file = paste0("kfold_model_fits/Fit_wo",as.character(k),"_training.Rda"))
-  
-  m_wo_test <- brm(amplitude_n400 ~ 1 + amplitude_prestim + syn*sem + (1+syn*sem||subject) + (1+syn*sem||item),
-                     prior = priors_m,
-                     warmup = 2000,
-                     iter = 10000,
-                     cores = 4,
-                     control = list(adapt_delta = 0.9),
-                     save_pars = save_pars(all = TRUE),
[…]

### 2024-08-26 code_data/Model-comparison-using-Bayes-factors/Compute-Bayes-factors.R
Deleted via the Open Science Framework

@@ -1,296 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-library(dplyr)
-
-#############
-SVdat <- data.frame(pred   = rep("data",3),
-                    effect    = c("syntactic", "semantic", "interaction"), 
-                    mean  = c(-0.14,-0.3,-0.03),
-                    upper = c(0.05,-0.1,0.38),
-                    lower = c(-0.33,-0.5,-0.33))
-SVdat$effect <- factor(SVdat$effect, levels=c("syntactic", "semantic", "interaction"))
-################
-
-tol_scaler <- 9
-delta <- (abs(SVdat[2,]$upper-SVdat[2,]$lower)/(4*tol_scaler))
-
-## Prior 1: normal_{lb=0,ub=0.05}(0.01,0.01)
-
-load("Simulated-data/Simulated_data_two_cues_model.Rda")
-df.lkl <- simdat_2cues_sameclause
-head(df.lkl)
-df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
-df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
-df.lkl$proposal.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0,sd=0.025)
-ML <- mean((df.lkl$lkl)*(df.lkl$prior.dens)/(df.lkl$proposal.dens))
-df.lkl$ML <- ML
-df.lkl$prior <- "normal(0.01,0.01)"
-df.bf <- as.data.frame(df.lkl %>% group_by(prior) %>% 
-                         summarize(MargLik=mean(ML)))
-df.bf
-df.bf$Model <- "Two cues model"
-
-load("Simulated-data/Simulated_data_three_cues_model.Rda")
-df.lkl <- simdat_3cues
-df.lkl$Model <- "Three cues model"
-df.lkl$lkl <- dnorm(df.lkl$MESem-(-SVdat[2,]$mean),0,delta)
-df.lkl$prior.dens <- dtruncnorm(df.lkl$lfs,a=0,b=0.05,mean=0.01,sd=0.01)
[…]

### 2024-08-26 code_data/Model-comparison-using-Bayes-factors/Models/interACT-onecue.R

@@ -1,240 +1,0 @@
-library(tidyr)
-library(dplyr)
-
-# Original code by Felix Engelmann (https://github.com/felixengelmann/inter-act)
-
-###############################################################
-## Global parameters
-###############################################################
-
-reset_params <- function(){
-  lf        <<- 0.15       # latency factor
-  le        <<- 1       # latency exponent
-  rth 	    <<- -1.5    # retrieval threshold
-  bll       <<- 0.5     # decay parameter
-  ans       <<- 0.2    # activation noise
-  mas       <<- 1       # maximum associative strength 
-  mp        <<- 1       # mismatch penalty
-  ga        <<- 1       # goal source activation
-  rand_time <<- 3       # latency variability
-  blc       <<- 0       # base-level activation
-  ##
-  lp        <<- 1     # default time since last presentation (msec)
-  ldp       <<- 1       # last distractor presentation (msec)
-  ## Distractor control:
-  ndistr    <<- 1				# number of distractors
-  dbl       <<- 0       # distractor base-level
-  ##
-  ## Cue weighting, cue confusion and prominence:
-  cueweighting <<- 1     # Strength of structural cue as ratio str/sem
-  normalizeWeights <<- TRUE 
-  qcf 		  <<- 1 			# match quality correction factor
-  qco 		  <<- -2*rth 			# match quality correction offset
-  psc       <<- 1       # prominence scaling constant C1
-  pic       <<- 0       # prominence scaling constant C2
-  tprom     <<- 0       # target prominence 
-  dprom     <<- 0       # distractor prominence
-  # cl 				<<- 0 			# cue confusion level (0-100)
-  cuesim    <<- -1    # cue-feature similarity [-1..0]
-  #
-  VERBOSE <<- TRUE
-}
-reset_params()
-
-idnames <- c("Set","Iteration","Condition","Target","Distractor")
-actrnames <- c("weights","bl1","bl2","times1","times2","noise1","noise2","blact1","blact2","act1","act2","activation","latency","retrieved","acc","miss","fail")
[…]

### 2024-08-26 code_data/Model-comparison-using-Bayes-factors/Simulate_data_one_cue_model.R

@@ -1,113 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-
-################
-
-#### One CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-
-# Load model 
-source("Models/interACT-onecue.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1), c(1), c(1), c(1)),
-  distractor_match = list(c(0), c(0), c(0), c(0)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(1))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Generate prior predictions #####
-
-# Generate 200,000 simulations from model:
-nsims<-200000
-
-lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
-
-## check:
-hist(lfs,freq=FALSE)
-
-
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
-  lf<<- lfs[i]  
-  sims <- create_param_matrix(model_4cond, iterations=2000)
-  results <- run(sims)
-  condMeans <- compute_cond_means(results)
-  ## compute main effect of Syntactic manipulation
-  ## HiSyn-LoSyn:
-  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
-    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
-  ## save the estimate:
-  maineffectSyn[i]<-meSyn
-  ## compute main effect of Semantic manipulation
-  ## HiSem-LoSem:
[…]

### 2024-08-26 code_data/Model-comparison-using-Bayes-factors/Simulate_data_three_cues_model.R

@@ -1,105 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-
-################
-
-
-# Load model 
-source("Models/interACT-threecues.R")
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1,1), c(1,1,1), c(1,1,1), c(1,1,1)),
-  distractor_match = list(c(0,0,0), c(0,0,1), c(0,1,0), c(0,1,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(0.33,0.33,0.33))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-# Generate prior predictions #####
-# Generate 100,000 simulations from model:
-nsims<-200000
-
-
-lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
-
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
-  lf<<- lfs[i]  
-  sims <- create_param_matrix(model_4cond, iterations=2000)
-  results <- run(sims)
-  condMeans <- compute_cond_means(results)
-  ## compute main effect of Syntactic manipulation
-  ## HiSyn-LoSyn:
-  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
-    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
-  ## save the estimate:
-  maineffectSyn[i]<-meSyn
-  ## compute main effect of Semantic manipulation
-  ## HiSem-LoSem:
-  meSem<-mean(c(condMeans$Latency[2],condMeans$Latency[4]))-
-    mean(c(condMeans$Latency[1],condMeans$Latency[3]))
[…]

### 2024-08-26 code_data/Model-comparison-using-Bayes-factors/Simulate_data_two_cues_model.R

@@ -1,105 +1,0 @@
-# load packages
-library(truncnorm)
-library(ggplot2)
-library(cowplot)
-library(stringr)
-
-
-#### TWO CUE VERSION OF THE ORIGINAL INTERACT MODEL WITH COMPLEX SYNTACTIC CUE {+subject within same clause} #####
-
-# Load model 
-source("Models/interACT.R")
-
-reset_params()
-
-# Parameters to be passed to run() function for the simulations, including cue weights
-model_4cond <- list(
-  target_match = list(c(1,1), c(1,1), c(1,1), c(1,1)),
-  distractor_match = list(c(0,0), c(0,1), c(0,0), c(0,1)),
-  Target =     c("LoSyn","LoSyn","HiSyn","HiSyn"),
-  Distractor = c("LoSem","HiSem","LoSem","HiSem"),
-  weights = list(c(strWeight(),semWeight()))
-)
-
-mas <<- c(1.5) # Maximum associative strength
-mp <<- 0.5 # Mismatch penalty
-psc <<- 0 # Prominence scaling
-qcf <<- 1 # Match quality correction factor
-cuesim <<- -1 # Cue-feature similarity [-1..0]
-cueweighting <<- 1 # Cue weighting
-VERBOSE <<- FALSE
-
-#### Simulate data from the proposal density #####
-
-nsims <- 200000
-
-lfs <- rtruncnorm(nsims,a=0,b=0.05,mean=0,sd=0.025)
-
-maineffectSyn<-rep(NA,nsims)
-maineffectSem<-rep(NA,nsims)
-interactionSynSem<-rep(NA,nsims)
-
-for(i in 1:nsims){
-  # Run model
-  print(paste("run: ",i,sep=""))  
-  lf <<- lfs[i]  
-  sims <- create_param_matrix(model_4cond, iterations=2000)
-  results <- run(sims)
-  condMeans <- compute_cond_means(results)
-  ## compute main effect of Syntactic manipulation
-  ## HiSyn-LoSyn:
-  meSyn<-mean(c(condMeans$Latency[3],condMeans$Latency[4]))-
-    mean(c(condMeans$Latency[1],condMeans$Latency[2]))
-  ## save the estimate:
-  maineffectSyn[i]<-meSyn
-  ## compute main effect of Semantic manipulation
-  ## HiSem-LoSem:
-  meSem<-mean(c(condMeans$Latency[2],condMeans$Latency[4]))-
[…]

### 2024-08-26 code_data/previous_studies/models_G.R
Updated via the Open Science Framework

@@ -3,7 +3,7 @@
 
 # script to run models (to later compute Bayes Factors) for the previous studies
 
-# Mertzen et al., (2024) English and German eye-tracking data
+# Mertzen et al., (2023) English and German eye-tracking data
 # for comparability with self-paced reading times, we use the total fixation times (TFT) eye-tracking measure
 
 

### 2024-08-26 code_data/SPR/r_scripts/compute_BFs.R
Updated via the Open Science Framework

@@ -43,9 +43,9 @@ priors_l <- c(
   prior(normal(0, 0.1), class = sd)
 ) 
 
-load("E://model_fits/Fit_s_Full_precrit_slopes_symmetrical.Rda")
-load("E://model_fits/Fit_m_Full_precrit_slopes_symmetrical.Rda")
-load("E://model_fits/Fit_l_Full_precrit_slopes_symmetrical.Rda")
+load("model_fits/Fit_s_Full_precrit_slopes_symmetrical.Rda")
+load("model_fits/Fit_m_Full_precrit_slopes_symmetrical.Rda")
+load("model_fits/Fit_l_Full_precrit_slopes_symmetrical.Rda")
 
 
 # bfs
@@ -98,9 +98,9 @@ df.bf_schoknecht[nrow(df.bf_schoknecht)+1,] <- c("S-D", "German", "pre-critical"
 df.bf_schoknecht[nrow(df.bf_schoknecht)+1,] <- c("S-D", "German", "pre-critical", "interaction","Normal(0, 0.1)","no", round(SD_bf10.m_l_int,2))
 save(df.bf_schoknecht, file = paste("BFs_schoknecht_precrit.Rda"))
 
-load("E://model_fits/Fit_s_Full_crit_slopes_symmetrical.Rda")
-load("E://model_fits/Fit_m_Full_crit_slopes_symmetrical.Rda")
-load("E://model_fits/Fit_l_Full_crit_slopes_symmetrical.Rda")
+load("model_fits/Fit_s_Full_crit_slopes_symmetrical.Rda")
+load("model_fits/Fit_m_Full_crit_slopes_symmetrical.Rda")
+load("model_fits/Fit_l_Full_crit_slopes_symmetrical.Rda")
 
 # bfs
 # model_s
@@ -150,9 +150,9 @@ df.bf_schoknecht[nrow(df.bf_schoknecht)+1,] <- c("S-D", "German", "critical", "i
 save(df.bf_schoknecht, file = paste("BFs_schoknecht_crit.Rda"))
 
 
-load("E://model_fits/Fit_s_Full_spill_slopes_symmetrical.Rda")
-load("E://model_fits/Fit_m_Full_spill_slopes_symmetrical.Rda")
-load("E://model_fits/Fit_l_Full_spill_slopes_symmetrical.Rda")
+load("model_fits/Fit_s_Full_spill_slopes_symmetrical.Rda")
[…]

### 2024-08-26 code_data/previous_studies/compute_BFs.R
Updated via the Open Science Framework

@@ -1,13 +1,13 @@
 # author: Pia Schoknecht
 # date: 09.07.2024
 
-# script to compute Bayes Factors with the Savage-Dickey method
+# script to compute Bayes Factors with the Savage-Dickey method for Mertzen et al. (2023)
 library(tidyverse)
 library(brms)
 library(ggplot2)
 library(bayesplot)
 
-
+# load English model fits
 load("model_fits_E/Fit_s_Full_precrit.Rda")
 load("model_fits_E/Fit_m_Full_precrit.Rda")
 load("model_fits_E/Fit_l_Full_precrit.Rda")
@@ -19,6 +19,7 @@ load("model_fits_E/Fit_l_Full_crit.Rda")
 load("model_fits_E/Fit_s_Full_postcrit.Rda")
 load("model_fits_E/Fit_m_Full_postcrit.Rda")
 load("model_fits_E/Fit_l_Full_postcrit.Rda")
+
 
 #bfs 
 
@@ -36,44 +37,94 @@ compute_bfs <- function(model){
 }
 
 # save BFs in df
-df.bf_m23.E <- data.frame(matrix(ncol=8,nrow=0))
-colnames(df.bf_m23.E) <- c("Experiment", "BF_method", "Language", "Region", "Effect","Prior","truncated", "BF10")
+df.bf_m23 <- data.frame(matrix(ncol=8,nrow=0))
+colnames(df.bf_m23) <- c("Experiment", "BF_method", "Language", "Region", "Effect","Prior","truncated", "BF10")
 
-df.bf_m23.E[nrow(df.bf_m23.E)+1,] <- c("M23", "S-D", "English", "pre-critical", "syntactic","Normal(0, 0.01)","no", round(compute_bfs(m_s_full_precrit)[1],2))
-df.bf_m23.E[nrow(df.bf_m23.E)+1,] <- c("M23", "S-D", "English", "pre-critical", "semantic","Normal(0, 0.01)","no", round(compute_bfs(m_s_full_precrit)[2],2))
-df.bf_m23.E[nrow(df.bf_m23.E)+1,] <- c("M23", "S-D", "English", "pre-critical", "interaction","Normal(0, 0.01)","no", round(compute_bfs(m_s_full_precrit)[3],2))
[…]

### 2024-08-26 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 There was evidence against syntactic interference under all priors in the pre-critical and critical region.
-Only in the spill-over region was there anecdotal evidence (BF$_{10}$ = 2.8) for a small syntactic interference effect ($\pm$ 8 ms).
-Similarly, there was evidence against the interaction under all priors in the pre-critical and spill-over regions.
-Only in the critical region was there anecdotal evidence (BF$_{10}$ = 1.4) for a small interaction effect ($\pm$ 8 ms).
+Only in the spill-over region, there was anecdotal evidence (BF$_{10}$ = 2.8) for a small syntactic interference effect ($\pm$ 8 ms).
+There was evidence against the interaction under all priors in the pre-critical and spill-over regions.
+In the critical region, Bayes factors provided no evidence either way regarding the interaction (BF$_{10}$ = 1.4).
 
@@
 
-Figure \ref{fig:eeg_bfs} shows the Bayes factor analysis of the ERP data.
-There was anecdotal evidence for a very small syntactic interference effect ($\pm$ 0.2\,$\mu V$) on the N400 and evidence against larger effects ($>$ 1\,$\mu V$).
-There was evidence against an interaction under all priors in the N400 spatio-temporal window.
-In a new analysis of the P600 spatio-temporal window, there was anecdotal evidence for a very small interaction ($\pm$ 0.2\,$\mu V$) and evidence against larger interaction effects ($>$ 1\,$\mu V$).
-There was unequivocal evidence against an effect of syntactic interference in the P600.
+Figure \ref{fig:eeg_bfs} shows the Bayes factor analyses of the ERP data in the spatio-temporal windows of the N400 and P600 (the latter is a new analysis which was added during the revision).
+For a very small syntactic interference effect on the N400 and a very small interaction on the P600 ($\pm$ 0.2\,$\mu V$), Bayes factors were inconclusive (BF$_{10}$ close to 1).
+For all other investigated effect sizes ($>$ 1\,$\mu V$) on both ERP components, Bayes factors provided evidence against syntactic interference and the interaction.
 
@@
 \subsubsection*{Response to R1.8}
-We agree that syntactic interference and/or the interaction could be real, but our data does not provide conclusive evidence for them.
+The reviewer is right that syntactic interference and/or the interaction could represent real effects.
+However, our data does not provide conclusive evidence for them.
 Because it can be difficult to discern small differences in the Bayes factor plots (Figures 5 and 11 in Response to E.2), the Table below shows all Bayes factors for syntactic interference and the interaction.
 All Bayes factors above 1 are highlighted in bold.
-Following \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we regard Bayes factors around 1 as providing no evidence.
-The only Bayes factor remarkably above 1 is BF$_{10}$ = 2.8 for the syntactic interference effect in the reading times of the spill-over region.
-Because of the pre-critical effects in the reading times, this effect in the spill-over region is confounded.
-Therefore, we think that although the Bayes factor theoretically provided anecdotal evidence for syntactic interference in the spill-over region, there is no \textit{decisive} evidence for syntactic interference in our data.
+Following \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we regard Bayes factors around 1 as providing no evidence either way (inconclusive).
+The only Bayes factor above 1 is BF10 = 2.8 for the syntactic interference effect in the reading times of the spill-over region.
+Because of the pre-critical effects in reading times, this effect in the spill-over region is potentially confounded: we cannot be sure whether the observed effect is due to syntactic retrieval interference or was somehow caused by semantic encoding interference.
+Therefore, we think that although the Bayes factor does provide anecdotal evidence for the syntactic interference manipulation in the spill-over region, we cannot conclude that the syntactic interference effect in our data is due to syntactic retrieval interference
 
@@
 
+\newpage
 We have revised the description of the SPR results to include more nuance regarding the BFs under the different priors:

### 2024-08-26 paper/main.tex
Updated via the Open Science Framework

@@
 There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
-Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).
-Similarly,} \Copy{insig_interaction}{\textcolor{blue}{there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
-Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}
+Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).} \Copy{insig_interaction}{\textcolor{blue}{There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
+Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}
 
 \textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
-In contrast, the Bayes factor analysis provided at best weak evidence for syntactic interference and the interaction or even evidence against them.}}
+In contrast, the Bayes factor analysis across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}}
 
@@
 This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue.
-This result holds for prior assumptions that allows higher probability mass for small values of $F$., i.e., for $F<0.01$.}
+This result holds for prior assumptions that allows higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.}
 

### 2024-08-26 paper/main.tex
Updated via the Open Science Framework

@@
 In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms.
-Reading times at the critical verb showed very weak support for a small interaction of syntactic and semantic interference.
-There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading times differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
-This means that although it is interesting that small effects of syntactic interference and the interaction emerged only in these regions  any interpretation of these effects would be purely speculative.
-Therefore, we refrain from discussing the effects occurring in the critical and spill-over regions.}}\label{only_precritical} \textcolor{blue}{We will discuss the absence of decisive evidence for syntactic interference in our data in the general discussion.} The effects in the pre-critical region are discussed next.
+Reading times at the critical verb showed inconclusive results (neither support for nor against) regarding a very small interaction of syntactic and semantic interference.
+There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
+This means that although it is interesting that a small effect of syntactic interference emerged in the spill-over region any interpretation of this effect would be purely speculative.
+Therefore, we do not think that this effect can be clearly attributed to retrieval interference due to syntactic similarity and refrain from further discussing this effect occurring in the spill-over region.}}\label{only_precritical} \textcolor{blue}{We will discuss the absence of decisive evidence for syntactic interference in our data in the general discussion.} The effects in the pre-critical region are discussed next.
 

### 2024-08-26 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \subsection*{Response to E.2}
-We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide evidence for syntactic interference or the interaction.
+We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide convincing evidence for syntactic interference or the interaction.
 As an overview of the revised Bayes factor analyses, we show Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs} below.
@@
 Because of the pre-critical effects in reading times, this effect in the spill-over region is potentially confounded: we cannot be sure whether the observed effect is due to syntactic retrieval interference or was somehow caused by semantic encoding interference.
-Therefore, we think that although the Bayes factor does provide anecdotal evidence for the syntactic interference manipulation in the spill-over region, we cannot conclude that the syntactic interference effect in our data is due to syntactic retrieval interference
+Therefore, we think that although the Bayes factor provides anecdotal evidence for the syntactic interference manipulation in the spill-over region, we cannot conclude that this effect in our data is due to syntactic retrieval interference
 
@@
 \subsubsection*{Response to R2.2}
-We agree with the reviewer. We now refrain from interpreting the results at the critical region. To emphasize this point, we added the following to the manuscript on page \pageref{only_precritical}:\\
+We agree with the reviewer.
+We now refrain from interpreting the results at the critical and spill-over region.
+To emphasize this point, we added the following to the manuscript on page \pageref{only_precritical}:\\
 
@@
 
-We are not discussing the small interaction in the critical region because we refrained from interpreting effects after the pre-critical region due to pre-critical effects (see Response to R2.2).
+%We are not discussing the small interaction in the critical region because we refrained from interpreting effects after the pre-critical region due to pre-critical effects (see Response to R2.2).
 
@@
 
-\noindent We added the following text describing the ERPs for high vs.\ low syntactic interference, which show no difference, to the Results on page \pageref{ERP_results2}:
+\noindent We added the following text describing the ERPs for high vs.\ low syntactic interference, which show only slight differences, to the Results on page \pageref{ERP_results2}:
 
@@
 
-\noindent We added the following description of the Bayes factor analysis of the ERP data to the Results on page \pageref{ERP_results3}:
+\noindent We added the following description of the Bayes factor analysis of the ERP data (evidence for semantic interference, no evidence for syntactic interference or the interaction) to the Results on page \pageref{ERP_results3}:
 
@@
 We have extended the discussion of the potential reasons for the differences in results in the manuscript (cross-linguistic differences, different methods, differences in the items, statistical power, inter-individual differences, see below).
-it is unclear why the results differ between the studies.
+It is unclear why the results differ between the studies.
 
@@
 We agree with the reviewer regarding their overall conclusion, that there may be multiple ways to explain the absence of syntactic interference, and multiple reasons why our results differ from previous studies.
-Our opinion is that the present work still has value because it attempts to replicate  well-known results in sentence processing, uses a method different than self-paced reading and eye-tracking (EEG) to investigate these effects, and proposes one computational implementation that could explain the EEG results.
+Our opinion is that the present work still has value because it attempts to replicate well-known results in sentence processing, uses a method different than eye-tracking during reading to investigate these effects, and proposes one computational implementation that could explain the EEG results.
 

### 2024-08-27 paper/main.tex
Updated via the Open Science Framework

@@
 We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-\textcolor{blue}{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
+\Copy{syn_interf}{\textcolor{blue}{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
 This means that semantic interference occurred even before the retrieval event could begin.
 Reading times of the pre-critical word and critical verb showed evidence against syntactic interference.
-In the spill-over region, there was weak evidence for a small syntactic interference effect in the size of up to $\pm$8 ms.
-Reading times at the critical verb showed inconclusive results (neither support for nor against) regarding a very small interaction of syntactic and semantic interference.
+In the spill-over region, the syntactic manipulation slowed down the reading times, but to a lesser extent than the semantic manipulation (CrI of syntactic interference vs.
+CrI of semantic interference: [1,15] ms vs.\ [7,21] ms).
+The Bayes factors also provided less support for the syntactic effect than for the semantic effect (BFs under prior Normal(0, 0.01) in the spill-over region for the syntactic effect vs.
+the semantic effect: 2.8 vs.\ 246).
+The relatively small extent of the syntactic effect which only emerged in one region and which is also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference.
+However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.}}\label{syn_interf} \textcolor{blue}{We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
+For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) a very small effect ($\pm$8 ms).
 There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
-This means that although it is interesting that a small effect of syntactic interference emerged in the spill-over region any interpretation of this effect would be purely speculative.
-Therefore, we do not think that this effect can be clearly attributed to retrieval interference due to syntactic similarity and refrain from further discussing this effect occurring in the spill-over region.}}\label{only_precritical} \textcolor{blue}{We will discuss the absence of decisive evidence for syntactic interference in our data in the general discussion.} The effects in the pre-critical region are discussed next.
+Therefore, we refrain from further discussing effects occurring in the later regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.
 

### 2024-08-27 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 \subsection*{Response to E.2}
-We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses still do not provide convincing evidence for syntactic interference or the interaction.
+We have re-done our analyses in response to the concerns raised by the reviewers and the Bayes factors from the revised analyses do not provide evidence for an interaction of syntactic and semantic interference.
+The new analyses provide some evidence for syntactic interference in the spill-over region.
 As an overview of the revised Bayes factor analyses, we show Figures \ref{fig:spr_bfs} and \ref{fig:eeg_bfs} below.
@@
 
+Since the syntactic interference effect in the spill-over region was small and only weakly supported by the Bayes factor analysis, we believe that our SPR data predominantly showed semantic interference.
+We have added the following text summarizing the SPR results to the manuscript on page \pageref{syn_interf}:
+
+\begin{quote}
+    \Paste{syn_interf}
+\end{quote}
+
 Figure \ref{fig:eeg_bfs} shows the Bayes factor analyses of the ERP data in the spatio-temporal windows of the N400 and P600 (the latter is a new analysis which was added during the revision).
@@
 
-Given these Bayes factor results taken together (at best anecdotal evidence for small effects with magnitudes like an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), at least in our data, we don't feel that it is defensible to argue that, there is some evidence for syntactic interference (or for an interaction).
+Given these Bayes factor results taken together (at best anecdotal evidence for small effects with magnitudes like an 8\,ms reading time difference or an 0.2\,$\mu V$ difference in brain responses), at least in our data, we don't feel that it is defensible to argue that, there is convincing evidence for syntactic interference (or for an interaction).
 
@@
 The only Bayes factor above 1 is BF10 = 2.8 for the syntactic interference effect in the reading times of the spill-over region.
-Because of the pre-critical effects in reading times, this effect in the spill-over region is potentially confounded: we cannot be sure whether the observed effect is due to syntactic retrieval interference or was somehow caused by semantic encoding interference.
-Therefore, we think that although the Bayes factor provides anecdotal evidence for the syntactic interference manipulation in the spill-over region, we cannot conclude that this effect in our data is due to syntactic retrieval interference
+In contrast, the Bayes factor for semantic interference under the same prior and in the same region was 248.
+Therefore, we think that our SPR data predominantly showed semantic interference (see also our Response to E.2,).
+This view is further supported by the ERP results which unequivocally showed only semantic interference.
 
@@
 The meta-analysis of reading time studies by \textcite{jaeger_etal_2017} showed that the pooled estimate of inhibitory interference across studies with similar designs as ours was indeed small (13.1 ms, page 327), but some studies showed effects around 50 ms (e.g., two of the sub-experiments from \cite{vandyke07}) and one experiment by Franck et al., (2015) even showed an estimate above 100 ms (see Figure 2 on page 329 of \cite{jaeger_etal_2017}).
-Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}).
+Because such large interference effects have been found previously (see also our new Figure \ref{fig:previous_vs_present} showing all estimates of \cite{vandyke07} and \cite{mertzen}), it seems reasonable to not restrict the analyses to small effect sizes a priori.
 
-\newpage
 We have revised the description of the SPR results to include more nuance regarding the BFs under the different priors:
@@
 We agree with the reviewer regarding their overall conclusion, that there may be multiple ways to explain the absence of syntactic interference, and multiple reasons why our results differ from previous studies.
-Our opinion is that the present work still has value because it attempts to replicate well-known results in sentence processing, uses a method different than eye-tracking during reading to investigate these effects, and proposes one computational implementation that could explain the EEG results.
+Our opinion is that the present work still has value because it attempts to replicate well-known results in sentence processing, uses a method different than eye-tracking during reading (which was the method used in previous work) to investigate these effects, and proposes one computational implementation that could explain the EEG results.
 

### 2024-08-27 paper/ResponseToReviewers.tex
minor wording edits

@@
 
-\noindent \textbf{Prof. Jarrold}\\
+\noindent \textbf{Prof.\ Jarrold}\\
 \noindent \textbf{Associate Editor}\\
@@
 \vskip 1em
-Together with my co-authors Shravan Vasishth and Himanshu Yadav, who was added during the revision, I would like to submit our revision of the article titled ``Do syntactic and semantic similarity lead to interference effects?
+Together with my co-authors Shravan Vasishth and Himanshu Yadav, who was added as a co-author during the revision, I would like to submit our revision of the article titled ``Do syntactic and semantic similarity lead to interference effects?
 Evidence from self-paced reading and event-related potentials using German''.
@@
 
-You noted in your decision letter that the manuscript could potentially be a valuable contribution to the special issue, a view largely shared by the reviewers.
+You noted in your decision letter that the manuscript could potentially be a valuable contribution to the special issue, a view that seems to be shared by the reviewers.
 
@@
 In response to the concerns about our statistical analyses, we have completely re-done the analyses of the reading times and ERP data and added analyses of the previous data.
-We brought Himanshu Yadav into the project to refine our modeling, including the addition of a third model, and carry out model comparisons.
-We believe that these new analyses strengthen our initial interpretation that our study shows semantic interference, but no syntactic interference and no interaction.
+We brought Himanshu Yadav into the project to refine our modeling, including the addition of a third (baseline) model, and to carry out model comparisons.
+We believe that these new analyses strengthen our initial interpretation that our study shows clear evidence for semantic interference, but little to no evidence for syntactic interference, and no interaction between these two factors.
 
@@
 It is possible that the reviewer is right to not expect the design to lead to decisive syntactic interference effects because the syntactic manipulation might not be appropriate.
-However, we are building on the previously published claim that this design leads to syntactic interference.
+However, we are building on the previously published claim that this design does in fact lead to syntactic interference.
 We believe that our results are a valuable contribution to the literature on syntactic interference, because our work serves as an important counterpoint to the previously published claims \citep{vandyke07, mertzen}.
@@
 We built on this assumption and focus on the question which specific retrieval cues are being used during subject-verb dependency resolution.
-Our findings showed that the retrieval cues generated at the verb are highly specific (including hierarchical syntactic information) and can successfully identify the subject as the retrieval target.
+Our data suggest that the retrieval cues generated at the verb are highly specific (including hierarchical syntactic information) and can successfully identify the subject as the retrieval target.
 Therefore, the present findings further support the assumption that sentence processing relies on skilled memory retrieval.

### 2024-08-27 paper/main.tex
Minor wording edits

@@
 \noindent \textcolor{blue}{where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
-The term $\mathcal{L}{F_i|y}$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.
+The term $\mathcal{L}(F_i|y)$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.
 We compute marginal likelihoods under different prior assumptions $p(F)$.}
@@
 We choose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
-The $\delta$ parameter determines how good the approximation is: the lower the value of the $delta$, the better the approximation of the likelihood.
+The $\delta$ parameter determines how good the approximation is: the lower the value of $\delta$, the better the approximation of the likelihood.
 However, if $\delta$ is too low, the computational demand increases because too many proposals would end up with zero likelihood.
-We choose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising the computational efficiency.}
+We choose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising on computational efficiency.}
 
 \textcolor{blue}{From the marginal likelihoods, we computed Bayes factors for the two-cues model compared to the three-cues model and the one-cue model.
-That is, the marginal likelihood of the two-cues model $M_1$ was divided by the marginal likelihood of the three-cues model $M_2$ and one-cue model $M_3$ to obtain the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
+That is, the marginal likelihood of the two-cues model $M_1$ was divided by (i) the marginal likelihood of the three-cues model $M_2$,  and (ii) the marginal likelihood of the one-cue model $M_3$.
+This calculation resulted in the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
 Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above $10$ is interpreted as strong evidence in favor of the two-cues model, and a value between $3$ and $10$ is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).}

### 2024-08-28 paper/main.tex
minor wording change

@@
 To compute Bayes factors for \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) in the published paper.
-Then, using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method.
+Then, using these posteriors and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method.
 Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.

### 2024-08-28 paper/ResponseToReviewers.tex
minor wording change

@@
 We are simply reporting what we found.
-Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
+In psycholinguistics (as in other disciplines), contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.

### 2024-08-28 paper/main.tex
minor wording change

@@
 \Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature.
-Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{taboretal04, lcpaape2023}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
+In psycholinguistics alone, contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{taboretal04, lcpaape2023}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}

### 2024-08-28 paper/main.tex
Updated via the Open Science Framework

@@
 
-\Copy{intro}{It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval \textcolor{blue}{\parencite[see e.g., ][]{lewis06}}.
-As an example, consider
-(\ref{ex:1_vandyke07}).
+\Copy{intro}{
+
+It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval \textcolor{blue}{\parencite[see e.g., ][]{lewis06}.
+In fact, a key assumption in psycholinguistics is that the same constraints on memory retrieval that have been identified within memory research in cognitive psychology \parencite[e.g., ][]{anderson_etal_2004_ACT_R} may be applicable in sentence processing as well.
+An example of a key construct from cognitive psychology that is relevant for sentence processing is that spreading activation leads to difficulty in identifying the correct item to retrieve from memory.
+The present work investigates whether and how such independently posited, general constraints on memory retrieval impact sentence processing.
+As a concrete example of general memory constraints impacting sentence comprehension,} consider (\ref{ex:1_vandyke07}).
 
@@
 
-There is clear evidence in the literature \parencite[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
+There is clear evidence in the psycholinguistic literature \parencite[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
 A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues \textcolor{blue}{\citep{Lewis2005, mcelree}}.}\label{intro}
@@
 \textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method.
-A Bayesian hypothesis test for which the posterior density is divided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970}.} The Bayes factor is often written as $BF_{10}$.
+A Bayesian hypothesis test for which the posterior density is divided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970, dickey1971weighted, verdinelli1995computing}.} The Bayes factor is often written as $BF_{10}$.
 \textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account.
@@
 \textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as $\pm$1\,$\mu V$.
-}\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600.
-The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals included zero for both the N400 and the P600.
+}\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference for both the N400 and the P600.
+The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, for both the N400 and the P600, respectively.
 The 95\% credible interval of the interaction was centered around zero for the N400.
-For the P600, it included mostly negative values but crossed zero.}}\label{ERP_results1}
+For the P600, it included mostly negative values.}}\label{ERP_results1}
 
@@
 For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
-To compute Bayes factors for \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE) in the published paper.
-Then, using these posteriors and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method.
+To compute Bayes factors for \textcite{vandyke07}, we used the Normal-Normal conjugate case \parencite{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper.
+Then, using these posteriors and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970, dickey1971weighted, verdinelli1995computing}.
 Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.

### 2024-08-28 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

--- 2024-08-27 paper/ResponseToReviewers.tex
@@
 \subsection*{Response to E.3}
-Our work is situated in the cue-based retrieval framework which has been built over the last decades \citep{Lewis2005, mcelree}.
-Its main assumption is that sentence processing relies on skilled memory processes, with a focus on memory retrieval.
+Our work is situated in the cue-based retrieval framework which has been built over the last decades \citep{Lewis2005, mcelree} and is informed by memory research \citep[see e.g., ][]{anderson_etal_2004_ACT_R}.
+The key assumption is that sentence processing relies on skilled memory processes, with a focus on memory retrieval.
 We built on this assumption and focus on the question which specific retrieval cues are being used during subject-verb dependency resolution.
@@
 Therefore, the present findings further support the assumption that sentence processing relies on skilled memory retrieval.
-We have emphasized the theoretical grounding of our work by adding more references to the introduction on page \pageref{intro}:
+We have revised the introduction to emphasized the theoretical grounding of our work (see page \pageref{intro}):
 
@@
 \subsubsection*{Response to R1.5}
-About why our results diverge from previous studies' and whether we think our results are correct and the previous work is not, we have no position on this.
+In the revised manuscript, we only present computational modeling of retrieval interference in ERP data (see Response to 1.1).
+To our knowledge, our study is the first to investigate the present design with ERPs, therefore only the present data can be captured.
+
+About why our reading times results diverge from previous reading times studies' and whether we think our results are correct and the previous work is not, we have no position on this.
 We are simply reporting what we found.
-Contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{lcpaape2023, taboretal04}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
-The only way to resolve contradictions in the literature is by carrying out adequately powered studies in the future that can distinguish between the theoretical claims.
-Our work does not stand out as an anomaly; if anything, it's quite the opposite.
+We have added a section in which we directly compare our results to the previous results (see Response to R1.4).
+This comparison showed that the most remarkable difference between the studies concerns the syntactic interference effect in the pre-critical region, which was found in the previous studies but not in the present study.
+Contradictions like this can only be resolved over time by carrying out adequately powered studies in the future.
 We have added a discussion of divergent results in the literature to page \pageref{not_an_anomaly} in the General Discussion:

### 2024-08-29 paper/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 However, I am far from being an expert in this area, and the (therefore much more knowledgeable) reviewers have raised a number of issues with the paper, some of which I judge to be substantive.
-For this reason I have come to a 'revise and resubmit' decision.
+For this reason I have come to a `revise and resubmit' decision.
 If you do revise and resubmit your paper, as I hope you will, I will approach the same set of reviewers to see if they are able to re-review (and obviously there is no guarantee of acceptance at that stage).
@@
 For a very small syntactic interference effect on the N400 and a very small interaction on the P600 ($\pm$ 0.2\,$\mu V$), Bayes factors were inconclusive (BF$_{10}$ close to 1).
-For all other investigated effect sizes ($>$ 1\,$\mu V$) on both ERP components, Bayes factors provided evidence against syntactic interference and the interaction.
+For all other investigated effect sizes ($\geq$ $\pm$1\,$\mu V$) on both ERP components, Bayes factors provided evidence against syntactic interference and the interaction.
 
@@
 The key assumption is that sentence processing relies on skilled memory processes, with a focus on memory retrieval.
-We built on this assumption and focus on the question which specific retrieval cues are being used during subject-verb dependency resolution.
+We built on this assumption and focused on the question which specific retrieval cues are being used during subject-verb dependency resolution.
 Our data suggest that the retrieval cues generated at the verb are highly specific (including hierarchical syntactic information) and can successfully identify the subject as the retrieval target.
@@
 \textbf{Please ensure that you have included a ``Data Availability" statement immediately before your empirical work with a link to materials, data, and analysis code} (see our `guide for authors' for instructions).
-Please make sure that you have updated the data and analysis code following any revisions to your manuscript.
-Please confirm in your next covering letter that all results presented in your manuscript can be reproduced using the materials that you have made available.''
+\textbf{Please make sure that you have updated the data and analysis code following any revisions to your manuscript.
+Please confirm in your next covering letter that all results presented in your manuscript can be reproduced using the materials that you have made available}.''
 \end{quote}
@@
 \subsection*{Response to E.4}
-We have moved our data availability statement to page \pageref{statement}, before Experiments 1a and 1b:
+We have updated the files in our OSF repository during the revisions, so that all analyses can be reproduced.
+The data availability statement has been moved to page \pageref{statement}, before Experiments 1a and 1b:
 
@@
 \begin{quote}
-``In the manuscript "Do syntactic and semantic similarity lead to interference effects?
+``In the manuscript ``Do syntactic and semantic similarity lead to interference effects?
 Evidence from self-paced reading and event-related potentials using German" the authors investigate the effect of syntactic and semantic interference in a large-sample self-paced reading and EEG studies that follow in design Van Dyke (2007) and Mertzen et al.
@@
 Could the effect of increased reading times simply be due to the fact that the +animate distractor is not as good as the -animate distractor in this context?
-That is, the effect has nothing to do with encoding interference, it's just a consequence of the fact that, e.g., in (2a) the statement that "the burglar was awful" is harder to make sense of than that "the loss was awful".
+That is, the effect has nothing to do with encoding interference, it's just a consequence of the fact that, e.g., in (2a) the statement that ``the burglar was awful" is harder to make sense of than that ``the loss was awful".
 \textbf{I think one easy way to check it would be to run an acceptability/plausibility study which checks whether the +animate version and the -animate version are comparable in their acceptability.} I know that the manuscript mentions a possible confound due to plausibility (p.
@@
 
-%BFs do not to need to be strongly in favor of an effect across all investigated priors to conclude that an effect is supported.
-However, to draw decisive conclusions in favor of an effect, the evidence for that effect should be stronger than anecdotal for at least a subset of reasonable priors.
-
 Regarding the choice of priors, we believe that it is reasonable to explore a wide range of a priori effect sizes, specifically from $\pm$8 to $\pm$80 ms for the reading times.
@@
 
-We have revised the description of the SPR results to include more nuance regarding the BFs under the different priors:
+We have revised the description of the SPR results on page \pageref{SPR_results} to include more nuance regarding the BFs under the different priors:
 
@@
 
-For the analysis of the ERP data, we changed the largest prior from N(0,5) to N(0,2), so from effects ranging up to 11 $\mu V$ to effects ranging up to 4 $\mu V$, to accommodate the concern that we might focus on too large effect sizes.
+For the analysis of the ERP data, we changed the largest prior from N(0,5) to N(0,2), so from effects ranging up to $\pm$11 $\mu V$ to effects ranging up to $\pm$4 $\mu V$, to accommodate the concern that we might focus on too large effect sizes.
 We have revised the description of the BF analysis for the ERP data as follows (see page \pageref{ERP_results3}):
@@
 Regarding the second issue, they address it by testing both syntactic and semantic cues.
-As a syntactic cue, the authors tested "being a subject".
+As a syntactic cue, the authors tested ``being a subject".
 In the high syntactic interference condition, the distractor functions as the subject of an embedded clause that serves as the complement of a non-restrictive relative clause (who had told her that the loss/ burglar was awful).
@@
 1.
-From a theoretical point of view, "being a subject" is not a primitive.
+From a theoretical point of view, ``being a subject" is not a primitive.
 Subjects are identified by a cluster of features.
@@
 
-Unsurprisingly, the authors reach this very same conclusion: "We speculate that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory." Yes, exactly, but this isn't new.
+Unsurprisingly, the authors reach this very same conclusion: ``We speculate that one possibility might be that the parser tracks hierarchical structure and that this hierarchical information is used when searching for a noun in memory." Yes, exactly, but this isn't new.
 
-I also agree with this statement (p.41): "[…] the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
+I also agree with this statement (p.41): ``[…] the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
 In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb." Once again, yes, that's accurate.
@@
 Therefore, we do not discuss this visual impression.
-See the following text in the manuscript on page \pageref{insig_interaction}:
+See the following text in the manuscript on page 25 for the Bayes factors results of the interaction:
 
@@
 \end{quote}
-
-%We are not discussing the small interaction in the critical region because we refrained from interpreting effects after the pre-critical region due to pre-critical effects (see Response to R2.2).
 
@@
 Any type of retrieval interference, may it be due to syntactic or semantic retrieval cues, might affect memory retrieval and therefore the N400.
-We added the following text to the paragraph about the regions of interest on page \pageref{why_n400}:
+We added the following text to the paragraph about the regions of interest on page 8:
 
@@
 We think that this can easily be represented in encodings and then be used during retrieval.
-We have added a more explicit description of this underlying mechanism to page \pageref{clause_tracker}:
+We have added a more explicit description of this underlying mechanism to page 44:
 
@@
 Maybe the verb ``drive" wants a subject with a [+driver] feature.
-But is this plausibly something that is part of the memory encoding of "teacher" but not ``little boy"?
+But is this plausibly something that is part of the memory encoding of ``teacher" but not ``little boy"?
 Does this require something more than a toy theory of semantic encoding?
@@
 
-\noindent Regarding the line break, we have added the following text to page \pageref{linebreak}:
+\noindent Regarding the line break, we have added the following text to page 53:
 
@@
 We manipulated syntactic interference in the same way as they did.
-Nevertheless, their data showed syntactic interference and our data do not.
+Nevertheless, their data showed syntactic interference and our data did not.
 We have extended the discussion of the potential reasons for the differences in results in the manuscript (cross-linguistic differences, different methods, differences in the items, statistical power, inter-individual differences, see below).
@@
 
-\setcounter{exx}{4}
+\setcounter{exx}{5}
 \begin{quote}

### 2024-08-29 paper/main.tex
Updated via the Open Science Framework

@@
     The\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} believed that the widower who told had that the burglar / loss awful was in.the.evening regularly drank in.order to forget \\
-    \trans `The neighbor believed that the widower, who had told her that the loss was awful, drank regularly in the evenings to forget.' \\
+    \trans `The neighbor believed that the widower, who had told her that the burglar / loss was awful, drank regularly in the evenings to forget.' \\
 
@@
     The\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} believed that the widower who about the awful burglar / loss told had in.the.evening regularly drank in.order to forget\\
-    \trans `The neighbor believed that the widower, who had told her about the awful loss, drank regularly in the evenings to forget.'  \\
+    \trans `The neighbor believed that the widower, who had told her about the awful burglar / loss, drank regularly in the evenings to forget.'  \\
     \end{xlist}
@@
 The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval.
-For experimental work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \textcite{vandyke_mcelree06, cunnings_sturt_2018retrieval}.
+For experimental work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \textcite{vandyke_mcelree06} and \textcite{cunnings_sturt_2018retrieval}.
 For a principled approach to feature selection using word embeddings, see \textcite{smith2020principled}.
@@
 This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.
-Thus, our focus in the \textcolor{blue}{SPR experiment} is on the processing difficulty observed at the critical verb and the surrounding regions.
-
-\Copy{why_n400}{\textcolor{blue}{In the EEG experiment, the region of interest is the critical verb.
+Thus, our focus in the \textcolor{blue}{SPR experiment was} on the processing difficulty observed at the critical verb and the surrounding regions.
+
+\Copy{why_n400}{\textcolor{blue}{In the EEG experiment, the region of interest was the critical verb.
 Due to the direct measurement of brain activity with excellent temporal resolution event-related potential (ERP) effects do not tend to spill-over into later regions.
-We focus specifically on the N400 elicited by the verb.
+We focused specifically on the N400 elicited by the verb.
 The N400 is a negative deflection in the ERP signal that peaks approximately 400 ms after any word was encountered.
@@
 The few ERP studies that have investigated retrieval interference so far have indeed found negativities \citep{lee_garnsey, martinetal2014, schoknecht2022, vasishth_drenhaus_2011}, although the topography and latency of these negativities was mostly not typical for the N400.
-It is important to highlight here that the previous ERP studies on interference either increased the number of distractors in a sentences or used grammatical gender or number as the critical retrieval cue.
+It is important to highlight here that the previous ERP studies on interference either increased the number of distractors in a sentence or used grammatical gender or number as the critical retrieval cue.
 So, interference due to non-semantic features elicited negativities which were more or less similiar to the N400.
-Therefore, we expect the interference manipulation regardless of the type of retrieval cue (syntactic or semantic) to elicit N400 or similar effects.
+Therefore, we expected the interference manipulation regardless of the type of retrieval cue (syntactic or semantic) to elicit N400 or similar effects.
 For a discussion of the relation between the N400 and other negativities, see \textcite{ina_n400}.}
@@
 
-\Copy{osf}{\textcolor{blue}{\section{Data availability statement}
-All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}}\label{statement}
+\textcolor{blue}{\section{Data availability statement}
+\Copy{osf}{\textcolor{blue}{All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}}}\label{statement}
 
@@
 This was done to bring all fixed effects to a comparable scale which decreases the run time of the models.
-The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items).}
+The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.}
 
@@
 
-Since the accuracies are not of primary interest, we do not carry out Bayes factors analyses for these.
+Since the accuracies were not of primary interest, we did not carry out Bayes factors analyses for these.
 
@@
 Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but will be harder to interpret because of possible spill-over from the syntactic manipulation.
-The pre-pre-critical and pre-critical region show longer reading times for high vs.\ low semantic interference conditions.
+The pre-pre-critical and pre-critical region showed longer reading times for high vs.\ low semantic interference conditions.
 The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low.
@@
 Additionally, the second session in Experiment 1b showed shorter reading times compared to the first session.
-As mentioned earlier, this attenuation in reading times and differences between conditions are likely due to  adaptation to the SPR task in the second session.
+As mentioned earlier, this attenuation in reading times and differences between conditions were likely due to adaptation to the SPR task in the second session.
 
@@
 Reading times were a bit faster when both syntactic and semantic interference were low.
-The other two conditions (high syntactic/low semantic and low syntactic/high semantic) lead to the fastest verb reading times in Experiment 1b with items 61-120.
+The other two conditions (high syntactic/low semantic and low syntactic/high semantic) led to the fastest verb reading times in Experiment 1b with items 61-120.
 
@@
 In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms).
-The posteriors for the interaction were mostly positive, but the 95\% credible intervals included zero in all regions.
+The posteriors for the interaction were mostly positive in all regions.
 The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}
@@
 This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment.
-While this adaption is interesting in itself, it is not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}
+While this adaption is interesting in itself, it was not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}
 
@@
 The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}.
-The analyses use all the data from Experiments 1a and 1b.
+The analyses used all the data from Experiments 1a and 1b.
 \Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
@@
 Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).} \Copy{insig_interaction}{\textcolor{blue}{There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
-Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4).}}\label{insig_interaction}
-
-\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
-In contrast, the Bayes factor analysis across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}}
+Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).}}\label{insig_interaction}
+
+\textcolor{blue}{All in all, the Bayes factors provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
+In contrast, the Bayes factors  across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}}
 
 \Copy{fig_bf_spr}{\begin{figure}[htpb]
-    \caption{Bayes Factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
+    \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
     \label{fig:spr_bfs}
@@
 the semantic effect: 2.8 vs.\ 246).
-The relatively small extent of the syntactic effect which only emerged in one region and which is also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference.
+The relatively small extent of the syntactic effect which only emerged in one region and which was also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference.
 However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.}}\label{syn_interf} \textcolor{blue}{We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
-For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) a very small effect ($\pm$8 ms).
+For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) for a very small effect ($\pm$8 ms).
 There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
@@
 
-At the distractor (which occurred well before the pre-critical region), high semantic interference (animate distractors) lead to longer reading times than low semantic interference (inanimate distractors, see Figure \ref{fig:whole_sentence} A and B).
+At the distractor (which occurred well before the pre-critical region), high semantic interference (animate distractors) led to longer reading times than low semantic interference (inanimate distractors, see Figure \ref{fig:whole_sentence} A and B).
 One might argue that this pattern could be due to differences in word length.
@@
     The\textsubscript{\sc{fem}} neighbor\textsubscript{\sc{fem}} believed that the widower who told had that the burglar / loss awful was in.the.evening regularly drank in.order to forget \\
-    \trans `The neighbor believed that the widower, who had told her that the loss was awful, drank regularly in the evenings to forget.' \\
+    \trans `The neighbor believed that the widower, who had told her that the burglar / loss was awful, drank regularly in the evenings to forget.' \\
     \end{xlist}
@@
 Figure \ref{fig:posteriors_plausibility} presents the posteriors from the model with the Normal(0, 0.05) prior on the slopes.
-Although plausibility influenced reading times (higher plausibility lead to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms).
+Although plausibility influenced reading times (higher plausibility led to faster reading times, CrI [-30, -6] ms), there was an independent effect of semantic interference (CrI [2, 24] ms).
 The Bayes factors provided evidence for semantic interference in this analysis under all priors except the most diffuse one (Normal(0, 0.01): BF$_{10}$ = 4.2,
@@
 The interaction of semantic interference and plausibility was centered around zero.
-In conclusion, the slowdown in the reading time results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
-However, the plausibility difference was not the only driver of the reading time slowdown, i.e., semantic interference leads to slower reading times independent of plausibility.}
+In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
+However, the plausibility difference was not the only driver of the reading times slowdown, i.e., semantic interference led to slower reading times independent of plausibility.}
 
@@
 \textcite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region.
-They discuss several alternative explanations for their observed effects.
-In addition to encoding interference, \textcite{mertzen} propose three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
+They discussed several alternative explanations for their observed effects.
+In addition to encoding interference, \textcite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
 Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
 
-Regarding the possibility that sentence structure confounds as a possible explanation for the effects in the pre-critical region, this also seems implausible because the observed difference between the high and low semantic interference conditions is independent of the syntactic manipulation; it is the syntactic manipulation that has the confound.
+Regarding the possibility that sentence structure confounds caused the effects in the pre-critical region, this also seems implausible because the observed difference between the high and low semantic interference conditions is independent of the syntactic manipulation; it is the syntactic manipulation that has the confound.
 Therefore, the sentence structure confound is also not a good explanation for the semantic interference effect observed in our study.
@@
 Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data.
-Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in the critical region as well.
+Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in \textcolor{blue}{later regions} as well.
 The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence.
-It is not possible to pinpoint whether the effect in the critical region is due to a retrieval initiated at this region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \parencite{Yadavetal2022}.
-Because of the unclear status of the effects at the critical verb, we refrain from interpreting the interaction which was only observed in this region.
+It is not possible to pinpoint whether \textcolor{blue}{the effects in later regions were} due to a retrieval initiated at \textcolor{blue}{the respective} region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \parencite{Yadavetal2022}.
+Because of the unclear status of the effects at the critical verb \textcolor{blue}{and spill-over region}, we refrain from interpreting the \textcolor{blue}{semantic interference effects which were observed in these regions.}
 
@@
 Additionally, the results suggest an interaction, i.e., the difference between high and low semantic interference conditions was larger when syntactic interference was high vs.\ when it was low.
-Because the accuracies are not of primary interest, we did not carry out Bayes factors analyses for these.
+Because the accuracies were not of primary interest, we did not carry out Bayes factors analyses for these.
 
@@
 Figure \ref{fig:erp_sem_syn} (c) shows that this effect had a broad distribution.} \Copy{ERP_results2}{\textcolor{blue}{Figure \ref{fig:erp_sem_syn} (d) shows that syntactic interference affected the brain response to the critical verb in the N400 and P600 spatio-temporal windows to a smaller extend than semantic interference did.
-Figures \ref{fig:erp_sem_syn} (d) and (e) show that high vs.\ low syntactic interference lead to a slightly more negative ERP response with a broad distribution in the N400 time window.
+Figures \ref{fig:erp_sem_syn} (d) and (e) show that high vs.\ low syntactic interference led to a slightly more negative ERP response with a broad distribution in the N400 time window.
 Figure  \ref{fig:erp_sem_syn} (f) shows a slightly reduced P600 amplitude for high vs.\ low syntactic interference with a centro-parietal distribution.
@@
 Normal(0, 2): BF$_{10}$ = 5).
-In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$1\,$\mu V$ and no evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$2\,$\mu V$.
+In the P600 spatio-temporal window, the Bayes factors provided anecdotal to moderate evidence (BF$_{10}$ between 2 and 4.3) for the semantic interference effect under priors assuming effects smaller than $\pm$2\,$\mu V$ and no evidence for it (BF$_{10}$ = 1.1) under priors assuming effects in the range of $\pm$4\,$\mu V$.
 By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatio-temporal windows (BF$_{10}$ $<$ 1.2).
@@
 \textcolor{blue}{This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}.
-The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference lead to a reduced P600 amplitude compared to low semantic interference.
+The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference led to a reduced P600 amplitude compared to low semantic interference.
 Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}.
@@
 
-\textcolor{blue}{Similar to the self-paced reading results,} there is clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}.
+\textcolor{blue}{Similar to the self-paced reading results,} there was clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}.
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
@@
 We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
-Because  Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions.
+Because the Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions.
 However, the ERP data can be compared to the model's predictions; all that has to be changed is that the measurement is now on the microvolt scale instead of the millisecond scale.
@@
 Given that in our ERP data, there was no evidence for syntactic interference or depending on the a priori assumed effect size even evidence against it, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
-We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
+We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \label{clause_tracker}\Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
 \textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, as a simplification, we assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}.
@@
 This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
-In the present case, it is possible that syntactically relevant noun is only the one in the same clause as the verb.
-\textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}}\label{clause_tracker} \textcolor{blue}{The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
+In the present case, it is possible that the syntactically relevant noun is only the one in the same clause as the verb.
+\textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}} \textcolor{blue}{The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.
@@
 \textcolor{blue}{Figure \ref{fig:model_predictions_ERP} shows the  prior predictive distributions of the effects of interest for the three competing models discussed above in comparison to our ERP data.
-The null model model using one cue predicts no interference effects.
+The null model using one cue predicts no interference effects.
 The new two-cues model, including a syntactic cue that utilizes syntactic hierarchy, predicts no effect of syntactic interference and no interaction, but a semantic interference effect.
@@
 
-\textcolor{blue}{The procedure for computing Bayes factors is as follows.
+\textcolor{blue}{The procedure for computing Bayes factors was as follows.
 The Bayes factor computation required us to estimate the marginal likelihood of the competing models.
 We used the Monte Carlo integration method to estimate the marginal likelihood of each model.
-For Monte Carlo integration, we use the importance sampling estimator where a large number of number proposals are obtained from an importance density $q(F)$ and the average $likelihood \times prior$ is calculated across all these proposals to obtain the approximate marginal likelihood.
+For Monte Carlo integration, we used the importance sampling estimator where a large number of proposals were obtained from an importance density $q(F)$ and the average $likelihood \times prior$ was calculated across all these proposals to obtain the approximate marginal likelihood.
 The importance density should be chosen such that it covers the entire parameter space where $likelihood \times prior$ is non-zero and does not draw too many samples from the space where $likelihood \times prior$ is zero.
@@
 
-\textcolor{blue}{First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we use a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
-This importance density allows us to obtain more proposals for smaller values of $F$ (for F<0.03) while maintaining fatter tails for the larger values of $F$ ($F>0.03$); effectively, the whole parameter space of $F$ between $0$ and $0.05$ is explored to draw $0.2$ million proposals.
-Then, for each of these proposed parameter values, we compute the likelihood and prior density.
+\textcolor{blue}{First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we used a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
+This importance density allowed us to obtain more proposals for smaller values of $F$ (for $F<0.03$) while maintaining fatter tails for the larger values of $F$ ($F>0.03$); effectively, the whole parameter space of $F$ between $0$ and $0.05$ was explored to draw $0.2$ million proposals.
+Then, for each of these proposed parameter values, we computed the likelihood and prior density.
 The estimated marginal likelihood of a model will be given by}
@@
 The term $\mathcal{L}(F_i|y)$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.
-We compute marginal likelihoods under different prior assumptions $p(F)$.}
+We computed marginal likelihoods under different prior assumptions $p(F)$.}
 
@@
 In this case, model comparison is carried out using Approximate Bayesian Computation (ABC)~\citep{sisson2018handbook,palestro2018likelihood}.
-Under this approach, the model-generated data conditional on $F=F_i$ is compared with the actually observed data $y$.
+Using this approach, the model-generated data conditional on $F=F_i$ was compared with the actually observed data $y$.
 If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
 The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weights a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.
-We choose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
+We chose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
 The $\delta$ parameter determines how good the approximation is: the lower the value of $\delta$, the better the approximation of the likelihood.
 However, if $\delta$ is too low, the computational demand increases because too many proposals would end up with zero likelihood.
-We choose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising on computational efficiency.}
+We chose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising on computational efficiency.}
 
@@
 Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior.
-This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the previously assumed \{$\pm$grammatical subject\} cue.
-This result holds for prior assumptions that allows higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.}
+This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the two separate, previously assumed cues \{$\pm$grammatical subject\} and \{$\pm$same clause\}.
+This result holds for prior assumptions that allow higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.}
 
@@
 
-\Copy{only_this_design}{Taken together, \textcolor{blue}{present} the SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and \textcolor{blue}{little or no effects of syntactic interference or an interaction}.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
+\Copy{only_this_design}{Taken together, the \textcolor{blue}{present} SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and \textcolor{blue}{little or no effects of syntactic interference or an interaction}.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
@@
 \textcite{mertzen} made their data available.
-In comparison to the previous estimates, the effects of the present study are relatively small and associated with less uncertainty in general.
-The present and previous estimates of semantic interference seem relatively consistent: They indicate a positive effect which is largest in the pre-critical region and smallest in the critical region.
-It seems like the present estimates of syntactic interference are smaller, i.e., closer to zero, than the previous estimates.
-For the interaction, the present estimates are larger, i.e., further from zero in the positive direction, than the previous ones.
+In comparison to the previous estimates, the effects of the present study were relatively small and associated with less uncertainty in general.
+The present and previous estimates of semantic interference seem relatively consistent: They indicated a positive effect which was largest in the pre-critical region and smallest in the critical region.
+It seems like the present estimates of syntactic interference were smaller, i.e., closer to zero, than most of the previous estimates.
+For the interaction, most estimates were close to zero, but some estimates indicated a positive and some indicated a negative interaction.
 To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies to compare them to the present Bayes factors.
@@
 
-\Copy{linebreak}{\textcolor{blue}{The difference in methodology lead to another difference between our study and the previous studies.
+\Copy{linebreak}{\textcolor{blue}{The difference in methodology led to another difference between our study and the previous studies.
 In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line.
@@
 Cue-based retrieval as a theory on the memory processes during sentence comprehension implicitly states that these memory processes do not differ cross-linguistically \citep{lewis06, Lewis2005}.
-This is in line with no apparent clustering of the English vs.\, German results (see Figure \ref{fig:previous_vs_present} A).
+This is in line with no apparent clustering of the English vs.\ German results (see Figure \ref{fig:previous_vs_present} A).
 So, the use of materials in different languages cannot account for the differences in results between the studies.}
@@
 This raises the question whether the ERP effect was also caused by encoding interference, i.e., whether the ERP effect originated prior to the critical region.
-To answer this question, it is necessary to look at the ERPs elicited by words earlier in the sentence, This is an unusual step in ERP research but makes sense as an exploratory step in the present context.
+To answer this question, it is necessary to look at the ERPs elicited by words earlier in the sentence.
+This is an unusual step in ERP research but makes sense as an exploratory step in the present context.
 
@@
 
-Thus, the words directly preceding the critical verb do not show a pattern consistent with semantic interference.
+Thus, the words directly preceding the critical verb did not show a pattern consistent with semantic interference.
 A further question that arises is: which brain response was elicited by the distractor itself (see Figure \ref{fig:erp_distractor})?
@@
 
-We believe that the differences in the ERPs elicited by the distractor in the different conditions are caused by two factors.
-First, the high and low syntactic interference conditions have different word orders, i.e., the distractors are embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase; see Example \ref{ex:materials}).
+We believe that the differences in the ERPs elicited by the distractor in the different conditions were caused by two factors.
+First, the high and low syntactic interference conditions had different word orders, i.e., the distractors were embedded in different phrase structures (simple noun phrase vs.\ noun phrase including an adjective within a prepositional phrase; see Example \ref{ex:materials}).
 This is also apparent from the large difference between high and low syntactic interference conditions before distractor onset.
@@
 The comparison of the results of our SPR and ERP experiments suggests that the SPR results were strongly affected by encoding interference, while the ERPs elicited by the critical verb were not.
-This difference between methods might be caused by the difference in presentation rate.
+This difference between methods might have been caused by the difference in presentation rate.
 In the self-paced reading experiment, as implied by the name, the participants had control over the presentation rate of the stimuli.

### 2024-11-01 paper/revision2/elsarticle-template-harv.tex
Add files via upload

--- 2024-08-29 paper/main.tex
@@
-\documentclass[a4paper, man, floatsintext]{apa7}
-
-% Language setting
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%% 
+%% Template article for Elsevier's document class `elsarticle'
+%% with harvard style bibliographic references
+
+\documentclass[review,preprint,12pt,authoryear,floatsintext]{elsarticle}
+
+% my packages
 \usepackage[english]{babel}
 
-% Useful packages
-\usepackage{amsmath}
-\usepackage{amssymb}
+\usepackage{booktabs}
 \usepackage{graphicx}
-%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-\usepackage[natbib=true,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
-\usepackage{csquotes}
-\DeclareLanguageMapping{american}{american-apa}
-\bibliography{interferencebib}
-
+\usepackage{rotating}
+\usepackage{subfig}
+\usepackage{color}
 \usepackage{gb4e}
@@
 \noautomath
-\usepackage{authblk}
+\usepackage{amsmath} 
 \usepackage{rotating}
@@
 \usepackage{bigdelim}
-
-\usepackage{clipboard}
-\newclipboard{output-reviews}
-\openclipboard{output-reviews}
-
-\title{\hspace{1cm} Do syntactic and semantic similarity lead to interference effects?
-\newline
+\usepackage{hyperref}
+
+%my packages end
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times,authoryear]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,authoryear]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,authoryear]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+%% \usepackage{lineno}
+
+\begin{document}
+
+\begin{frontmatter}
+\title{Do syntactic and semantic similarity lead to interference effects?\\
 Evidence from self-paced reading and event-related potentials using German}
-
-\authorsnames[1,{2},1]{Pia Schoknecht, Himanshu Yadav, Shravan Vasishth}
-\authorsaffiliations{{Department of Linguistics, University of Potsdam},
-{Department of Cognitive Science, Indian Institute of Technology Kanpur}}
-
-\shorttitle{Syntactic and semantic interference effects}
-
-\authornote{This research was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), Project ID 317633480, SFB 1287.
-We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition. We are grateful to Dario Paape for helpful discussions.
-Correspondence concerning this article should be addressed to Pia Schoknecht, e-mail: pia.schoknecht@uni-potsdam.de
-}
-
-\abstract{Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed to interpret the sentence.
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \affiliation for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \affiliation{organization={},
+%%            addressline={}, 
+%%            city={},
+%%            postcode={}, 
+%%            state={},
+%%            country={}}
+%% \fntext[label3]{}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \affiliation[label1]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+%%
+%% \affiliation[label2]{organization={},
+%%             addressline={},
+%%             city={},
+%%             postcode={},
+%%             state={},
+%%             country={}}
+
+\begin{abstract}
+%% Text of abstract
+Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed to interpret the sentence.
 If these retrieval cues match with not only the features on the noun that is the retrieval target but also with those on other nouns in the sentence, then processing difficulty arises at the verb.
@@
 In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, in our data, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
-
-\clearpage
-
-}
-
-\begin{document}
-
-\maketitle
+\end{abstract}
+
+\end{frontmatter}
+%% \linenumbers
+
+%% main text
+\newpage
 
@@
 
-\Copy{intro}{
-
-It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval \textcolor{blue}{\parencite[see e.g., ][]{lewis06}.
-In fact, a key assumption in psycholinguistics is that the same constraints on memory retrieval that have been identified within memory research in cognitive psychology \parencite[e.g., ][]{anderson_etal_2004_ACT_R} may be applicable in sentence processing as well.
+{
+
+It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval \citep[see e.g., ][]{lewis06}.
+In fact, a key assumption in psycholinguistics is that the same constraints on memory retrieval that have been identified within memory research in cognitive psychology \citep[e.g., ][]{anderson_etal_2004_ACT_R} may be applicable in sentence processing as well.
 An example of a key construct from cognitive psychology that is relevant for sentence processing is that spreading activation leads to difficulty in identifying the correct item to retrieve from memory.
 The present work investigates whether and how such independently posited, general constraints on memory retrieval impact sentence processing.
-As a concrete example of general memory constraints impacting sentence comprehension,} consider (\ref{ex:1_vandyke07}).
+As a concrete example of general memory constraints impacting sentence comprehension, consider (\ref{ex:1_vandyke07}).
 
@@
 \ex \label{ex:1_vandyke07} 
-The worker was surprised that the resident who was living near the dangerous warehouse was complaining about the investigation. \parencite{vandyke07}
+The worker was surprised that the resident who was living near the dangerous warehouse was complaining about the investigation. \citep{vandyke07}
 \end{exe}
 
-There is clear evidence in the psycholinguistic literature \parencite[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
-A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues \textcolor{blue}{\citep{Lewis2005, mcelree}}.}\label{intro}
+There is clear evidence in the psycholinguistic literature \citep[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
+A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues \citep{Lewis2005, mcelree}.}\label{intro}
 
 Here, we consider two kinds of cues used in retrieval: semantic and  syntactic cues.
-In a sentence like (\ref{ex:1_vandyke07}), the retrieval cues generated at the verb are standardly assumed \parencite[e.g.,][]{vandyke07} to include at least the semantic cue \{+animate\} and the syntactic cue \{+grammatical subject\}.
+In a sentence like (\ref{ex:1_vandyke07}), the retrieval cues generated at the verb are standardly assumed \citep[e.g.,][]{vandyke07} to include at least the semantic cue \{+animate\} and the syntactic cue \{+grammatical subject\}.
 These cues pick out the target noun \textit{the resident} as the subject.
@@
 
-Previous work has shown clear evidence for this kind of retrieval interference \parencite{vandyke07, vandyke_mcelree06, vandyke_mcelree2011, nicenboim,  vandyke_lewis03,jaeger_etal_2017}.
+Previous work has shown clear evidence for this kind of retrieval interference \citep{vandyke07, vandyke_mcelree06, vandyke_mcelree2011, nicenboim,  vandyke_lewis03,jaeger_etal_2017}.
 However, two important questions remain unanswered.
-First, as discussed in \textcite{jaeger_etal_2017}, almost all the published studies on interference effects are severely underpowered.
-Low power has the consequence that published estimates will tend to be overestimates or misestimates of the effect size \parencite{vasishth2018_signficancefilter}.
+First, as discussed in \cite{jaeger_etal_2017}, almost all the published studies on interference effects are severely underpowered.
+Low power has the consequence that published estimates will tend to be overestimates or misestimates of the effect size \citep{vasishth2018_signficancefilter}.
 Over-/misestimates have important implications for theory evaluation.
-One example is discussed in \textcite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \parencite{levy&keller_2013} could not be validated when a larger-sample study was conducted.
-By contrast, the original small-sample study \parencite{levy&keller_2013} showed suggestive evidence consistent with the theory.
-Another example is the recent heated debate \parencite{nieuwland_etal_2018} on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \parencite{nicenboim_etal_2020}.
-Second, although it has been established \parencite[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
-For example, \textcite{dillon2013} argued, following \textcite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built.
+One example is discussed in \cite{vasishth2018_signficancefilter}, where a key prediction of surprisal theory \citep{levy&keller_2013} could not be validated when a larger-sample study was conducted.
+By contrast, the original small-sample study \citep{levy&keller_2013} showed suggestive evidence consistent with the theory.
+Another example is the recent heated debate \citep{nieuwland_etal_2018} on prediction in sentence comprehension, which was largely focused on the results of low-powered studies, leading to potentially misleading conclusions \citep{nicenboim_etal_2020}.
+Second, although it has been established \citep[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
+For example, \cite{dillon2013} argued, following \cite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built.
 Instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
-Using computational modeling, \textcite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference, and that the reason for this is that principle A of the binding theory is the primary driver of the cue-based retrieval process \parencite[cf.
+Using computational modeling, \cite{dillon2013} suggested that antecedent-reflexive dependencies were immune to interference, and that the reason for this is that principle A of the binding theory is the primary driver of the cue-based retrieval process \citep[cf.
 ][]{jaeger_etal_2020,yadav2021individual}.
 This was a surprising discovery because the default assumption in cue-based retrieval accounts was that all cues were equally weighted, regardless of the dependency type; this had the consequence that retrieval interference was expected whenever a cue-overload configuration was present in a sentence.
-More broadly, an important insight in the work of \textcite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
+More broadly, an important insight in the work of \cite{dillon2013}, among others, is that  the precise details of which cues are used during retrieval can affect processing difficulty in possibly surprising ways.
 We build on this idea in the present paper.
@@
 
-Reading studies have been the dominant method used to investigate retrieval interference \parencite[see the review in][]{jaeger_etal_2017}.
+Reading studies have been the dominant method used to investigate retrieval interference \citep[see the review in][]{jaeger_etal_2017}.
 We aimed to investigate the role of syntactic and semantic cues during sentence comprehension with higher-powered experiments using  self-paced reading (SPR) as well as electroencephalography (EEG), which to our knowledge has never been used to investigate interference during subject-verb dependency resolution.
 
-\Copy{vd_design}{All experiments used the 2 $\times$ 2 design developed by \textcite{vandyke07}; the design crosses syntactic and semantic interference.} \label{vd_design}
+{All experiments used the 2 $\times$ 2 design developed by \cite{vandyke07}; the design crosses syntactic and semantic interference.} \label{vd_design}
 
@@
 Syntactic interference is manipulated via the syntactic status of the distractor, which is either the subject (\ref{ex:hisyn}) of an embedded clause, or part of a prepositional phrase (\ref{ex:losyn}) and hence a non-subject.
-\Copy{franck_refs}{Following \textcite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase \textcolor{blue}{ -- even if the distractor occurs in an embedded clause {\parencite[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary, franck2020hierarchical}.}}}\label{franck_refs}
+{Following \cite{vandyke07}, the assumption here is that  during the retrieval of the subject, a distractor in \emph{any} subject position causes more syntactic interference than a distractor inside a prepositional phrase -- even if the distractor occurs in an embedded clause {\citep[this assumption stands in contrast to findings on so-called agreement attraction, where syntactic hierarchy plays a crucial role, see e.g.,][]{franck2002subject, franck2006agreement,franck2011_commentary, franck2020hierarchical}.}}\label{franck_refs}
 
@@
 During the retrieval of the subject of the verb \textit{trank} (`drank'), the animate distractor  \textit{Einbrecher} (`burglar') is assumed to cause high semantic interference, while the inanimate distractor \textit{Verlust} (`loss') is assumed to cause low semantic interference.
-\Copy{semantic_features_encoding}{\textcolor{blue}{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained.
+{The animacy of the distractor is used as a stand-in feature for semantic features relevant for the specific retrieval event which might be more fine-grained.
 In \ref{ex:materials}, the semantic retrieval cue might be \{+animate\} or \{+human\} but it might as well be \{+can drink\}.
@@
 The present study is not designed to provide insight into the specific semantic features used during memory encoding and retrieval.
-For experimental work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \textcite{vandyke_mcelree06} and \textcite{cunnings_sturt_2018retrieval}.
-For a principled approach to feature selection using word embeddings, see \textcite{smith2020principled}.
+For experimental work on highly specific retrieval cues like \{+sailable\} and \{+shatterable\}, see \cite{vandyke_mcelree06} and \cite{cunnings_sturt_2018retrieval}.
+For a principled approach to feature selection using word embeddings, see \cite{smith2020principled}.
 A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
-}}\label{semantic_features_encoding}
+}\label{semantic_features_encoding}
 
 \paragraph{Regions of interest in the design}
-\textcolor{blue}{For both our SPR and EEG experiments, o}ur principal focus was on the processing of the verb (\textit{trank}, `drank').
-\textcolor{blue}{In the SPR experiment, t}he reading times of the immediately following spill-over region and of the pre-critical region were also of interest.
-Previous work using the present design \parencite{vandyke07,mertzen} has consistently found effects in the pre-critical region (we return to this point in the discussion sections\textcolor{blue}{, see Figure \ref{fig:previous_vs_present} A}).
-A further issue with German embedded clauses \parencite[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another (\textit{erzählt hatte, trank}, literally `told, drank').
+For both our SPR and EEG experiments, our principal focus was on the processing of the verb (\textit{trank}, `drank').
+In the SPR experiment, the reading times of the immediately following spill-over region and of the pre-critical region were also of interest.
+Previous work using the present design \citep{vandyke07,mertzen} has consistently found effects in the pre-critical region (we return to this point in the discussion sections, see Figure \ref{fig:previous_vs_present} A).
+A further issue with German embedded clauses \citep[also see][]{mertzen} was that, without an appropriate intervening phrase, the clause embedding would lead to two verbs appearing one after another (\textit{erzählt hatte, trank}, literally `told, drank').
 This is because in German, the verb is clause-final in embedded clauses.
-Such verb sequences are inherently difficult to process \parencite{VSLK11,bach1986crossed}.
+Such verb sequences are inherently difficult to process \citep{VSLK11,bach1986crossed}.
 
 For these reasons, we added two adverbs before the critical verb. Including these intervening regions had a further advantage: 
-It is well-known \parencite[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region.
-The spill-over of longer reading times can be problematic because the standard deviation increases as well \parencite{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power.
+It is well-known \citep[e.g., ][]{rayner2000effect} that the last word in a clause-final position causes a slowdown in reading; such a slowdown has the potential to spill over to the subsequent regions, possibly affecting the estimated effects at the critical region.
+The spill-over of longer reading times can be problematic because the standard deviation increases as well \citep{wagenmakers2007linear}; such an increase in standard deviation would drastically lower statistical power.
 Inserting two adverbs before the critical verb thus served to attenuate spill-over from the clause-final word in the embedded clause preceding the critical region.
@@
 This design also has the advantage that any potential effects seen at the pre-critical region (the adverb immediately preceding the verb) are less likely to be due to spill-over from the preceding region.
-Thus, our focus in the \textcolor{blue}{SPR experiment was} on the processing difficulty observed at the critical verb and the surrounding regions.
-
-\Copy{why_n400}{\textcolor{blue}{In the EEG experiment, the region of interest was the critical verb.
+Thus, our focus in the SPR experiment was on the processing difficulty observed at the critical verb and the surrounding regions.
+
+{In the EEG experiment, the region of interest was the critical verb.
 Due to the direct measurement of brain activity with excellent temporal resolution event-related potential (ERP) effects do not tend to spill-over into later regions.
@@
 Therefore, we expected the interference manipulation regardless of the type of retrieval cue (syntactic or semantic) to elicit N400 or similar effects.
-For a discussion of the relation between the N400 and other negativities, see \textcite{ina_n400}.}
-
-\textcolor{blue}{Alternately, retrieval interference -- especially syntactic retrieval interference -- might modulate the P600.
+For a discussion of the relation between the N400 and other negativities, see \cite{ina_n400}.
+
+Alternately, retrieval interference -- especially syntactic retrieval interference -- might modulate the P600.
 The P600 is a positive deflection in the ERP wave form that is typically observed at least 600 ms after stimulus onset.
@@
 High interference could increase difficulty during syntactic integration of subject and verb in our materials.
-Potential misretrievals of the distractor might lead to reanalysis which could manifest as an increased P600 amplitude in high compared to low interference conditions.}} \label{why_n400}
+Potential misretrievals of the distractor might lead to reanalysis which could manifest as an increased P600 amplitude in high compared to low interference conditions.} \label{why_n400}
 
@@
 The participants received \pounds 3.40 as compensation for the norming experiment.
-A demonstration of the plausibility norming experiment can be accessed via \hyperlink{https://farm.pcibex.net/r/WIeKYM/}{https://farm.pcibex.net/r/WIeKYM/}.
+A demonstration of the plausibility norming experiment can be accessed via \\\hyperlink{https://farm.pcibex.net/r/WIeKYM/}{https://farm.pcibex.net/r/WIeKYM/}.
 
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
 \caption{Plausibility ratings for the animate (high semantic interference conditions) and inanimate distractor (low semantic interference conditions), the introduction noun (\textit{X believed that \dots}) and the subject.
@@
 \centering
-\includegraphics[width=0.6\textwidth]{images/pandora_plausibility_ratings.jpg}
+\includegraphics[width=0.6\textwidth]{pandora_plausibility_ratings.jpg}
 \end{figure}
@@
 
-As discussed in detail below, the second session showed an adaptation effect \parencite{prasad2021rapid}: Reading times decreased overall.
-The adaptation effect is theoretically interesting per se \parencite{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
+As discussed in detail below, the second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
+The adaptation effect is theoretically interesting per se \citep{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
 For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both).
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
     \caption{The structure of the present study.
@@
     \centering
-    \includegraphics[width=\textwidth]{images/pandora_project_structure_figure.pdf}
+    \includegraphics[width=\textwidth]{pandora_project_structure_figure.pdf}
 \end{figure}
 
-\textcolor{blue}{\section{Data availability statement}
-\Copy{osf}{\textcolor{blue}{All materials, data and analysis scripts can be accessed via https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}}}\label{statement}
+\section{Data availability statement}
+{All materials, data and analysis scripts can be accessed via \\ https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}
 
@@
 In total, 908 participants took part in the self-paced reading experiments E1a and E1b.
-117 participants were excluded due to low comprehension question accuracy (< 70\,\%).
+117 participants were excluded due to low comprehension question accuracy ($<$ 70\,\%).
 Additionally, 16 participants were excluded because the demographic information in their Prolific profile and the information which they provided during the experiment did not match (regarding, e.g., their native language, language impairment, or age).
@@
 
-\Copy{fillers1}{In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
+{In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
 The experimental sessions were separated by 1 - 20 days.
 The procedure of the sessions was identical.
-\textcolor{blue}{In session 1, the experimental items 1-60 were presented interspersed with 40 filler sentences.
+In session 1, the experimental items 1-60 were presented interspersed with 40 filler sentences.
 In session 2, the experimental items 61-120 were presented with another set of 40 filler sentences.
-The fillers were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers1} Each session lasted approximately 25 minutes.
+The fillers were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}\label{fillers1} Each session lasted approximately 25 minutes.
 
@@
 The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
-\textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task.
+In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task.
 The trial id of the 100 trials per session was rescaled to span from 0 to 1.
 This was done to bring all fixed effects to a comparable scale which decreases the run time of the models.
-The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.}
+The models were run with full random effects, i.e., varying intercepts and slopes of all fixed effects and their interactions by participants and items.
 
@@
 These intervals represent the range over which we can be 95\% certain that the values of the parameter lies, given the statistical model and the data.
-However, formal hypothesis testing cannot be carried out without a likelihood ratio test that compares two alternative models \parencite{schad_etal_2022_BF,Royall}.
+However, formal hypothesis testing cannot be carried out without a likelihood ratio test that compares two alternative models \citep{schad_etal_2022_BF,Royall}.
 For this reason, formal hypothesis tests for the presence or absence of the effects of interest were carried out using Bayes factors.
-Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF},
+Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \citep{schad_etal_2022_BF},
 we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}).
-The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, \textcolor{blue}{ranging from -8 to 8 ms, -40 to 40 ms, or -81 to 81 ms.
-These priors are based on the observed sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}, and on meta-analyses relating to interference effects \parencite{jaeger_etal_2017}.}
-
-\begin{table}[h]
+The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from -8 to 8 ms, -40 to 40 ms, or -81 to 81 ms.
+These priors are based on the observed sizes of effects and uncertainties in previous reading studies that use the present design \citep{vandyke07,mertzen}, and on meta-analyses relating to interference effects \citep{jaeger_etal_2017}.
+
+\begin{table}[ht]
     \caption{Priors used for the analysis of the self-paced reading time data.
@@
   \cmidrule{2-3}
-  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \textcolor{blue}{[-8, 8]}\\
-  &  Normal(0, 0.05)&  \textcolor{blue}{[-40, 40]}\\
-  & Normal(0, 0.1) & \textcolor{blue}{[-81, 81]}\\
+  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & [-8, 8]\\
+  &  Normal(0, 0.05)&  [-40, 40]\\
+  & Normal(0, 0.1) & [-81, 81]\\
   \cmidrule{2-3}
@@
 
-\textcolor{blue}{Bayes factors were computed using the Savage-Dickey density ratio method.
-A Bayesian hypothesis test for which the posterior density is divided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970, dickey1971weighted, verdinelli1995computing}.} The Bayes factor is often written as $BF_{10}$.
-\textcolor{blue}{To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.} One major advantage of the Bayes factor over \textcolor{blue}{frequentist ANOVA or likelihood ratio tests} is that it takes the uncertainty of the parameters into account.
-This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \parencite[see][for detailed discussion]{schad_etal_2022_BF}.
+Bayes factors were computed using the Savage-Dickey density ratio method.
+A Bayesian hypothesis test for which the posterior density is divided by the prior density at a specific parameter value of interest, e.g., zero \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970, dickey1971weighted, verdinelli1995computing}.
+The Bayes factor is often written as $BF_{10}$.
+To obtain $BF_{10}$ from the Savage-Dickey density ratio, we take its inverse.
+One major advantage of the Bayes factor over frequentist ANOVA or likelihood ratio tests is that it takes the uncertainty of the parameters into account.
+This leads to more conservative inferences compared to frequentist ANOVA, which only takes the maximum likelihood estimate of the parameter into account \citep[see][for detailed discussion]{schad_etal_2022_BF}.
 Another important advantage of the Bayes factor -- one that is very relevant for the present work -- is that it is possible to find evidence for or against an effect.
@@
 
-Bayes factors can be interpreted as follows \parencite[e.g., ][]{lee2014bayesian}: if BF$_{10}$ > 1, it provides evidence in favor of the effect of interest.
-If BF$_{10}$ < 1, it provides evidence \textcolor{blue}{against the effect of interest}.
-The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence \textcolor{blue}{against the effect of interest (see Table \ref{tab:bf_interpretation})}.
+Bayes factors can be interpreted as follows \citep[e.g., ][]{lee2014bayesian}: if BF$_{10}$ $>$ 1, it provides evidence in favor of the effect of interest.
+If BF$_{10}$ $<$ 1, it provides evidence against the effect of interest.
+The larger the value of BF$_{10}$, the stronger the evidence for the effect and the smaller the value of BF$_{10}$ is, the stronger the evidence against the effect of interest (see Table \ref{tab:bf_interpretation}).
 In general, a large number of iterations is needed in the brms package in order to obtain stable estimates of the Bayes factor  \citep{schad_etal_2022_BF}.
@@
 
-\begin{table}[h]
-    \centering
-    \caption{Interpretation of Bayes factors \parencite{lee2014bayesian}.}
+\begin{table}[ht]
+    \centering
+    \caption{Interpretation of Bayes factors \citep{lee2014bayesian}.}
     \label{tab:bf_interpretation}
@@
 
-\begin{table}[]
+\begin{table}[ht]
     \caption{By-condition accuracy in critical trials in the SPR experiment (E1a and E1b combined).}
@@
 
-\begin{table}[H]
+\begin{table}[ht]
     \caption{Results in log-odds from the Bayesian generalized model analyzing the comprehension accuracy in the SPR Experiments 1a and 1b.}
@@
 
-\Copy{figwhole_sentence}{
+{
 \begin{sidewaysfigure}[h]
@@
     \centering
-    \includegraphics[width=0.85\textwidth]{images/Pandora_all_wholesentence_pooled_zoom_exp.jpg}
+    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom_exp.jpg}
 \end{sidewaysfigure}
@@
 Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
-\Copy{SPR_results}{\textcolor{blue}{The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions.
+{The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions.
 In the spill-over region, it included only positive values ([1, 15] ms).
@@
 The posteriors for the interaction were mostly positive in all regions.
-The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}
-
-\begin{figure}[H]
+The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.
+
+\begin{figure}[ht]
     \caption{Posteriors for the syntactic and semantic interference effects and their interaction at the critical verb and surrounding regions in the pooled self-paced reading data (Experiments 1a and 1b combined).
@@
     \centering
-    \includegraphics[width=\textwidth]{images/posteriors_spr_pooled_774.png}
+    \includegraphics[width=\textwidth]{posteriors_spr_pooled_774.png}
 \end{figure}
@@
 
-\Copy{adaptation}{\textcolor{blue}{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last trial (100th), reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
+{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last trial (100th), reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
 This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment.
-While this adaption is interesting in itself, it was not the focus of the present study, therefore we will not discuss it further.}}\label{adaptation}
+While this adaption is interesting in itself, it was not the focus of the present study, therefore we will not discuss it further.}\label{adaptation}
 
@@
 
-Was there any evidence for the effects \textcolor{blue}{of interest }estimated above?
+Was there any evidence for the effects of interest estimated above?
 The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}.
 The analyses used all the data from Experiments 1a and 1b.
-\Copy{SPR_bfs}{\textcolor{blue}{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
+{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
 In the critical region, the evidence for semantic interference was weak (Normal(0, 0.01): BF$_{10}$ = 3.5,
@@
 There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
-Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).} \Copy{insig_interaction}{\textcolor{blue}{There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
-Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).}}\label{insig_interaction}
-
-\textcolor{blue}{All in all, the Bayes factors provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
-In contrast, the Bayes factors  across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}}
-
-\Copy{fig_bf_spr}{\begin{figure}[htpb]
+Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).
+{There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
+Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).}\label{insig_interaction}
+
+All in all, the Bayes factors provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
+In contrast, the Bayes factors  across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}
+
+\begin{figure}[ht]
     \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the reading times (combined Experiments 1a and 1b) at the critical verb and surrounding regions, under a range of priors on the target parameters.}
@@
     \centering
-    \includegraphics[width=\textwidth]{images/BF_plot_spr_774_allregions.png}
+    \includegraphics[width=\textwidth]{BF_plot_spr_774_allregions.png}
 \end{figure}
-}
 
@@
 We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-\Copy{syn_interf}{\textcolor{blue}{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
+{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
 This means that semantic interference occurred even before the retrieval event could begin.
@@
 The relatively small extent of the syntactic effect which only emerged in one region and which was also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference.
-However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.}}\label{syn_interf} \textcolor{blue}{We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
+However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.}\label{syn_interf} We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
 For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) for a very small effect ($\pm$8 ms).
-There was evidence against the interaction under wider priors and in the surrounding regions.} \Copy{only_precritical}{\textcolor{blue}{Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
-Therefore, we refrain from further discussing effects occurring in the later regions.}}\label{only_precritical} The effects in the pre-critical region are discussed next.
+There was evidence against the interaction under wider priors and in the surrounding regions.
+{Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
+Therefore, we refrain from further discussing effects occurring in the later regions.}\label{only_precritical} The effects in the pre-critical region are discussed next.
 
@@
 
-\label{plausib_anim_inan}\Copy{plausib_anim_inan}{\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
-To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}.
+\label{plausib_anim_inan}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
+To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}.
 Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}, repeated here as \ref{ex:hisyn_repeat} for convenience).
-}
 
@@
 
-\textcolor{blue}{Forty-four native speakers of German (mean age: 26 years old, age range: 18 - 35 years, 21 female, 23 male), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible).
+Forty-four native speakers of German (mean age: 26 years old, age range: 18 - 35 years, 21 female, 23 male), who were recruited over Prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible).
 The mean ratings per item ranged from 3.6 to 6.6.
@@
 The results of an ordinal brms model analyzing the ratings of the critical items are shown in Figure \ref{fig:plausibility_anim_inanim}.
-Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible.}
+Sentences with an animate distractor had a slightly lower probability to gain a high rating on the plausibility scale ($\beta$ = -0.48, CrI [-0.67, -0.3]), i.e., the high semantic interference conditions were rated to be less plausible.
 
@@
     \label{fig:plausibility_anim_inanim}
-    \includegraphics[width=0.8\linewidth]{images/plausibility_anim_inan.jpg}
+    \includegraphics[width=0.8\linewidth]{plausibility_anim_inan.jpg}
 \end{figure}
 
-\textcolor{blue}{To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
+To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
 This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings.
@@
 In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
-However, the plausibility difference was not the only driver of the reading times slowdown, i.e., semantic interference led to slower reading times independent of plausibility.}
-
-\begin{figure}
+However, the plausibility difference was not the only driver of the reading times slowdown, i.e., semantic interference led to slower reading times independent of plausibility.
+
+\begin{figure}[ht]
     \centering
@@
 The blue vertical lines represent the median and the blue shaded areas are 80\,\% credible intervals.}\label{fig:posteriors_plausibility}
-    \includegraphics[width=0.8\linewidth]{images/posteriors_spr_pooled_774_plausibility.png}
+    \includegraphics[width=0.8\linewidth]{posteriors_spr_pooled_774_plausibility.png}
 \end{figure}
@@
 At the pre-critical region, there was evidence for semantic interference.
-As this cannot be due to retrieval interference, the most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}\textcolor{blue}{, but see the potential confounds discussed above).} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
-
-Indeed, previous work using the same design as in the present paper has also found semantic interference effects at the pre-critical region: both \textcite{vandyke07} and \textcite{mertzen} found such effects.
+As this cannot be due to retrieval interference, the most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above).
+The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
+
+Indeed, previous work using the same design as in the present paper has also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
 \citeauthor{mertzen}'s (\citeyear{mertzen}) Figures 5 and 6 indicated reading time differences in earlier regions, especially for their German data (see their Figure 6).
-\textcite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \textcite{mertzen} also points out, encoding interference could be an explanation even in that study.
+\cite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \cite{mertzen} also points out, encoding interference could be an explanation even in that study.
 Given these earlier findings, our results are consistent with the encoding interference explanation.
 
-\textcite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region.
+\cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region.
 They discussed several alternative explanations for their observed effects.
-In addition to encoding interference, \textcite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
+In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
 Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
@@
 Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data.
-Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in \textcolor{blue}{later regions} as well.
+Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in later regions as well.
 The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence.
-It is not possible to pinpoint whether \textcolor{blue}{the effects in later regions were} due to a retrieval initiated at \textcolor{blue}{the respective} region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \parencite{Yadavetal2022}.
-Because of the unclear status of the effects at the critical verb \textcolor{blue}{and spill-over region}, we refrain from interpreting the \textcolor{blue}{semantic interference effects which were observed in these regions.}
+It is not possible to pinpoint whether the effects in later regions were due to a retrieval initiated at the respective region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.
+Because of the unclear status of the effects at the critical verb and spill-over region, we refrain from interpreting the semantic interference effects which were observed in these regions.
 
@@
 Sessions were separated by one to eight weeks for each participant (the median gap was two weeks, the first and third quartiles being one and three weeks).
-\Copy{fillers2}{The procedure of both sessions was identical \textcolor{blue}{and the same lists were used as in Experiment 1, with 60 critical sentences and 40 fillers per session}.
+{The procedure of both sessions was identical and the same lists were used as in Experiment 1, with 60 critical sentences and 40 fillers per session.
 
@@
 Participants were familiarized with the procedure with two practice sentences.
-After that, the experiment was conducted in four blocks of \textcolor{blue}{25} sentences each, presenting the items in pseudorandomized order, with breaks between the blocks.}\label{fillers2} Each trial started with the presentation of a fixation cross in the center of the screen for 500 ms.
+After that, the experiment was conducted in four blocks of 25 sentences each, presenting the items in pseudorandomized order, with breaks between the blocks.}\label{fillers2} Each trial started with the presentation of a fixation cross in the center of the screen for 500 ms.
 Next, each word of the sentence was presented in the center of the screen.
@@
 We used Bayesian linear mixed models to analyze the single trial EEG data in response to the critical verb.
-We averaged the activity of 12 centro-parietal electrodes (Cz, C3/4, CPz, CP1/2, CP5/6, Pz, P3/4, POz) in the standard \textcolor{blue}{time windows of the  N400 (300 to 500 ms post critical word onset) and P600 (600 to 900 ms post critical word onset)} for all analyses.
+We averaged the activity of 12 centro-parietal electrodes (Cz, C3/4, CPz, CP1/2, CP5/6, Pz, P3/4, POz) in the standard time windows of the  N400 (300 to 500 ms post critical word onset) and P600 (600 to 900 ms post critical word onset) for all analyses.
 The models included sum-contrast coded fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.
@@
 This functioned as a regression-based instead of traditional baseline correction \citep{alday2019}.
-Varying intercepts for participants and items as well as by-participant \textcolor{blue}{and by-item} random slopes for \textcolor{blue}{syntactic interference,} semantic interference \textcolor{blue}{and their interaction} were included.
+Varying intercepts for participants and items as well as by-participant and by-item random slopes for syntactic interference, semantic interference and their interaction were included.
 We used relatively informative priors for all parameters of the models (see Table \ref{tab:eeg_priors}).
@@
     \midrule
-  Intercept & \textcolor{blue}{Normal(0, 5)}& \textcolor{blue}{[-10, 10]}\\
+  Intercept & Normal(0, 5)& [-10, 10]\\
   \cmidrule{2-3}
-  \multirow{4}{1cm}{slope} & \textcolor{blue}{Normal(0, 0.1)} & \textcolor{blue}{[-0.2, 0.2}]\\
-  & \textcolor{blue}{Normal(0, 0.5)}& \textcolor{blue}{[-1, 1]}\\
-  & \textcolor{blue}{Normal(0, 1)} & \textcolor{blue}{[-2, 2]}\\
-  & \textcolor{blue}{Normal(0, 2)} & \textcolor{blue}{[-4, 4]}\\
+  \multirow{4}{1cm}{slope} & Normal(0, 0.1) & [-0.2, 0.2]\\
+  & Normal(0, 0.5)& [-1, 1]\\
+  & Normal(0, 1) & [-2, 2]\\
+  & Normal(0, 2) & [-4, 4]\\
   \cmidrule{2-3}
@@
 For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats,schad_etal_2022_BF}.
-We chose slope priors which assume a wide range of effect sizes from \textcolor{blue}{[$-0.2$, 0.2] to [$-4$, 4]\,$\mu V$}.
+We chose slope priors which assume a wide range of effect sizes from [$-0.2$, 0.2] to [$-4$, 4]\,$\mu V$.
 
@@
 
-\begin{table}[]
+\begin{table}[ht]
     \caption{By-condition accuracy in critical trials in the EEG experiment.}
@@
 
-\begin{table}[]
+\begin{table}[ht]
     \caption{Results in log-odds from a Bayesian generalized model analyzing the comprehension accuracy in critical trials of the EEG experiment.}
@@
 grand average ERPs elicited by the critical verb for all conditions.
-\textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as $\pm$1\,$\mu V$.
-}\Copy{ERP_results1}{\textcolor{blue}{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference for both the N400 and the P600.
+Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as $\pm$1\,$\mu V$.
+{The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference for both the N400 and the P600.
 The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, for both the N400 and the P600, respectively.
 The 95\% credible interval of the interaction was centered around zero for the N400.
-For the P600, it included mostly negative values.}}\label{ERP_results1}
-
-\begin{figure}[H]
+For the P600, it included mostly negative values.}\label{ERP_results1}
+
+\begin{figure}[ht]
     \caption{ERPs elicited by the critical verb (word onset at 0\,ms) at electrode Cz.}
@@
     \centering
-    \includegraphics[width=\textwidth]{images/N_103_Cz_crit.png}
+    \includegraphics[width=\textwidth]{N_103_Cz_crit.png}
 \end{figure}
 
-\begin{figure}[H]
-    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal \textcolor{blue}{windows of the N400 (A) and P600 (B)}.
+\begin{figure}[ht]
+    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal windows of the N400 (A) and P600 (B).
 Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
@@
     \centering
-    \includegraphics[width=\textwidth]{images/posteriors_eeg.jpg}
+    \includegraphics[width=\textwidth]{posteriors_eeg.jpg}
 \end{figure}
@@
 The timing and topography of this effect suggested that it was an N400 effect; thus, the N400 was modulated by semantic interference.
-\textcolor{blue}{Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced amplitude for high vs.\ low semantic interference in the P600 time window.
-Figure \ref{fig:erp_sem_syn} (c) shows that this effect had a broad distribution.} \Copy{ERP_results2}{\textcolor{blue}{Figure \ref{fig:erp_sem_syn} (d) shows that syntactic interference affected the brain response to the critical verb in the N400 and P600 spatio-temporal windows to a smaller extend than semantic interference did.
+Furthermore, Figure \ref{fig:erp_sem_syn} (a) shows a slightly reduced amplitude for high vs.\ low semantic interference in the P600 time window.
+Figure \ref{fig:erp_sem_syn} (c) shows that this effect had a broad distribution.
+{Figure \ref{fig:erp_sem_syn} (d) shows that syntactic interference affected the brain response to the critical verb in the N400 and P600 spatio-temporal windows to a smaller extend than semantic interference did.
 Figures \ref{fig:erp_sem_syn} (d) and (e) show that high vs.\ low syntactic interference led to a slightly more negative ERP response with a broad distribution in the N400 time window.
 Figure  \ref{fig:erp_sem_syn} (f) shows a slightly reduced P600 amplitude for high vs.\ low syntactic interference with a centro-parietal distribution.
-}}\label{ERP_results2}
-
-\begin{figure}[H]
+}\label{ERP_results2}
+
+\begin{figure}[htpb]
 \caption{Brain responses elicited by the critical verb.
@@
 \subfloat[]{\label{fig:erp_sem_ERP}
-\includegraphics[scale=.62]{images/ERP_animacy_Cz.png}}
+\includegraphics[scale=.56]{ERP_animacy_Cz.png}}
 \end{minipage}%
@@
 \centering
-\subfloat[]{\label{fig:erp_sem_topo}\includegraphics[scale=.65]{images/topo_animate-inanimate_300_500.png}}
+\subfloat[]{\label{fig:erp_sem_topo}\includegraphics[scale=.56]{topo_animate-inanimate_300_500.png}}
 \par\medskip
 \centering
-\subfloat[]{\label{fig:erp_sem_topo_p600}\includegraphics[scale=.65]{images/topo_animate-inanimate_600_900.png}}
+\subfloat[]{\label{fig:erp_sem_topo_p600}\includegraphics[scale=.56]{topo_animate-inanimate_600_900.png}}
 \end{minipage}
@@
 \subfloat[]{\label{fig:erp_syn_ERP}
-\includegraphics[scale=.62]{images/ERP_subj_Cz.png}}
+\includegraphics[scale=.56]{ERP_subj_Cz.png}}
 \end{minipage}%
@@
 \centering
-\subfloat[]{\label{fig:erp_syn_topo}\includegraphics[scale=.65]{images/topo_subj-nsubj_300_500.png}}
+\subfloat[]{\label{fig:erp_syn_topo}\includegraphics[scale=.56]{topo_subj-nsubj_300_500.png}}
 \par\medskip
 \centering
-\subfloat[]{\label{fig:erp_syn_topo_p600}\includegraphics[scale=.65]{images/topo_subj-nsubj_600_900.png}}
+\subfloat[]{\label{fig:erp_syn_topo_p600}\includegraphics[scale=.56]{topo_subj-nsubj_600_900.png}}
 \end{minipage}
@@
 
-Bayes factors are shown in Figure \ref{fig:eeg_bfs}.
-\Copy{ERP_results3}{\textcolor{blue}{In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect under all priors
+Bayes factors are shown in Figure \ref{fig:eeg_bfs}. {In the N400 spatio-temporal window, the Bayes factors provided moderate to strong evidence for the semantic interference effect under all priors 
 (Normal(0, 0.1): BF$_{10}$ = 6,
@@
 Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF$_{10}$ $<$ 1.2).
-In sum, in our ERP data, there was decisive evidence for the semantic interference effect and mainly evidence against syntactic interference and the interaction.}}\label{ERP_results3}
-
-\Copy{fig_bf_eeg}{
-\begin{figure}[H]
+In sum, in our ERP data, there was decisive evidence for the semantic interference effect and mainly evidence against syntactic interference and the interaction.}\label{ERP_results3}
+
+{
+\begin{figure}[h]
     \caption{Bayes factors for the effects of semantic interference, syntactic interference and their interaction in the N400 (300 - 500 ms post critical verb onset) and P600 (600 - 900 ms post critical verb onset) time windows at centro-parietal electrode sites.}\label{fig:eeg_bfs}
     \centering
-    \includegraphics[width=1\textwidth]{images/BF_plot_eeg.jpg}
+    \includegraphics[width=1\textwidth]{BF_plot_eeg.jpg}
 \end{figure} }
@@
 We presented a large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency completion.
-To our knowledge, this is the first ERP experiment on this classic interference design, which has only been investigated using reading studies so far \parencite{mertzen,vandyke07}.
+To our knowledge, this is the first ERP experiment on this classic interference design, which has only been investigated using reading studies so far \citep{mertzen,vandyke07}.
 
@@
 
-\Copy{ERP_discussion}{The Bayes factors analyses found \textcolor{blue}{moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$).
+{The Bayes factors analyses found moderate to strong evidence for semantic interference in the N400 amplitude for the whole range of investigated effects sizes ($\pm$0.2 to $\pm4$\,$\mu V$).
 In the P600 spatio-temporal window, there was anecdotal to moderate evidence for semantic interference.
@@
 In sum, the ERP results showed clear evidence for semantic interference and mostly evidence against syntactic interference and the interaction of syntactic and semantic interference.
-Consequently, we conclude that the ERP data showed no syntactic interference and no interaction.}}\label{ERP_discussion}
+Consequently, we conclude that the ERP data showed no syntactic interference and no interaction.}\label{ERP_discussion}
 
@@
 The more complex memory retrieval under high semantic interference resulted in a more negative N400.
-\textcolor{blue}{This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}.
+This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}.
 The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference led to a reduced P600 amplitude compared to low semantic interference.
@@
 In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to function as the subject.
-A similar effect was found by \textcite{Tanner_etal_2017}.
+A similar effect was found by \cite{Tanner_etal_2017}.
 In their study, the P600 elicited by ungrammatical verbs which were incongruent in number was reduced when the sentence included a matching distractor.
@@
 The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration.
-Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.}
-
-\textcolor{blue}{Similar to the self-paced reading results,} there was clear evidence for semantic interference, and \textcolor{blue}{no evidence for syntactic interference in our ERP data}.
+Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.
+
+Similar to the self-paced reading results, there was clear evidence for semantic interference, and no evidence for syntactic interference in our ERP data.
 How do these observed patterns compare to the theoretical predictions of cue-based retrieval theory?
-We turn to this point next by discussing the quantitative predictions of the \textcite{Lewis2005} model.
-
-\section{Quantitative predictions of the Lewis and Vasishth (2005) model for \textcolor{blue}{ERPs elicited in} the present design}
-
-\label{modeling}\Copy{modeling}{\textcolor{blue}{The \citet{Lewis2005} model is generally used to predict reading times.
-Recently, \textcite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
+We turn to this point next by discussing the quantitative predictions of the \cite{Lewis2005} model.
+
+\section{Quantitative predictions of the Lewis and Vasishth (2005) model for ERPs elicited in the present design}
+
+\label{modeling}{The \citet{Lewis2005} model is generally used to predict reading times.
+Recently, \cite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
 We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
@@
 In order to model the ERP data, we therefore adjust the scaling factor $F$, which rescales the activation of items in memory to the relevant dependent measure.
-The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.}
-
-We sampled the free scaling parameter $F$ from a \textcolor{blue}{truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05.
+The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data.
+
+We sampled the free scaling parameter $F$ from a truncated normal prior distribution ($Normal_{lb=0,ub=0.05}(0.01, 0.01)$) with a lower bound of 0 and an upper bound of 0.05.
 This prior reflects the a priori belief that effects on the microvolt scale will range from 0 to 5 $\mu V$ with higher likelihood for effect sizes below 3 $\mu V$ (see Figure \ref{fig:model_predictions_ERP}).
-This relatively uninformative prior was chosen because, although there exists no previously published ERP data on the Van Dyke (2007) design, it reflects the reasonable range of effect sizes in psycholinguistic ERP research.} After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
-\textcolor{blue}{We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study.
-Since the scaling parameter $F$ cannot be negative, an additional step was needed to reflect that larger predicted effect sizes correspond to more negative N400 amplitudes: The predicted effects from the model} were multiplied with $-1$.
-\textcolor{blue}{In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.}
-
-\textcolor{blue}{For their recent simulations, \textcite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \textcite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}.
+This relatively uninformative prior was chosen because, although there exists no previously published ERP data on the Van Dyke (2007) design, it reflects the reasonable range of effect sizes in psycholinguistic ERP research.
+After sampling the scaling parameter, simulated data was generated from the model while holding all other parameters in the model at fixed values.
+We decided to model N400 amplitude differences because Bayes factors provided stronger evidence for the semantic interference effect on the N400 than on the P600 in the present ERP study.
+Since the scaling parameter $F$ cannot be negative, an additional step was needed to reflect that larger predicted effect sizes correspond to more negative N400 amplitudes: The predicted effects from the model were multiplied with $-1$.
+In future work, the posterior distribution of the semantic interference effect in our ERP data, with range [-0.5, -0.1]\,$\mu V$, could be used as a basis for redefining a tighter prior on the scaling parameter $F$.
+
+For their recent simulations, \cite{mertzen} created a version of the Lewis and Vasishth (2005) model that assumes that in the \cite{vandyke07} design, three cues are used at the verb: \{$\pm$ animacy\}, \{$\pm$ grammatical subject\} and \{$\pm$ same clause\}.
 The \{$\pm$ same clause\} cue matches only with nouns that occurred in the same clause as the verb.
@@
 Given that in our ERP data, there was no evidence for syntactic interference or depending on the a priori assumed effect size even evidence against it, it is plausible that, at least in this experimental design and language, syntactic interference plays little to no role.
-We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \textcite{mertzen}.} \label{clause_tracker}\Copy{clause_tracker}{\textcolor{blue}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.} Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \textcite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
-\textcolor{blue}{As already noted by \textcite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, as a simplification, we assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}.
-This idea goes back to an assumption in the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and implies a simple clause tracking mechanism.} The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
-This proposal is \textcolor{blue}{furthermore} in the spirit of the \textcite{dillon2013} and \textcite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
+We speculate that this could be due to the parser tracking hierarchical structure and that this hierarchical information is used when searching for a noun in memory -- even more so than suggested by \cite{mertzen}.
+\label{clause_tracker}{So, we implemented a model which puts even more emphasis on hierarchical syntactic structure and thereby basically eliminates syntactic interference.
+Specifically, the relevant syntactic cues that might be used at the verb are not \{$\pm$ grammatical subject\} and \{$\pm$ same clause\} -- as implemented by \cite{mertzen} -- but rather a composite cue \{$\pm$ subject-in-same-clause\}.
+As already noted by \cite{mertzen}, structural information like [+ subject of clause$_i$] is not an inherent feature like for example [+ animate], but within in the cue-based retrieval framework, as a simplification, we assume that such a feature is encoded into memory and can consequently be used to match the retrieval cue \{$\pm$ subject of clause$_i$\}.
+This idea goes back to an assumption in the Lewis and Vasishth (2005) model by making the model ``sensitive to whether or not it is parsing an embedded clause'' \citep[][p.388]{Lewis2005} and implies a simple clause tracking mechanism.
+The resulting model assumes two retrieval cues: \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}.
+This proposal is furthermore in the spirit of the \cite{dillon2013} and \cite{Sturt2003} claim, that the parser intelligently targets only the syntactically relevant noun during antecedent retrieval.
 In the present case, it is possible that the syntactically relevant noun is only the one in the same clause as the verb.
-\textcolor{blue}{For more discussion of the role of structural retrieval cues, see \textcite{franck2020hierarchical} and \textcite{arnett2017subject}.}} \textcolor{blue}{The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
+For more discussion of the role of structural retrieval cues, see \cite{franck2020hierarchical} and \cite{arnett2017subject}.} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.
@@
 This model naturally does not predict any interference effects.
-It can be regarded the null model or no-interference model.}
-
-\textcolor{blue}{Figure \ref{fig:model_predictions_ERP} shows the  prior predictive distributions of the effects of interest for the three competing models discussed above in comparison to our ERP data.
+It can be regarded the null model or no-interference model.
+
+Figure \ref{fig:model_predictions_ERP} shows the  prior predictive distributions of the effects of interest for the three competing models discussed above in comparison to our ERP data.
 The null model using one cue predicts no interference effects.
@@
 The predictions of all models overlap with the effects of interest in the data.
-}
-
-\Copy{fig:model_predictions_ERP}{
-\begin{figure}[htpb]
-    \caption{The prior predictions on the microvolt scale for syntactic and semantic interference and their interaction \textcolor{blue}{from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.}
+
+\begin{figure}[ht]
+    \caption{The prior predictions on the microvolt scale for syntactic and semantic interference and their interaction from three versions of the Lewis and Vasishth (2005) cue-based retrieval model.
     %using the two cues \{$\pm$ subject-in-same-clause\} and \{$\pm$ animate\}. 
@@
         \centering
-         \includegraphics[width=\textwidth]{images/compmod_predictions.png}
+         \includegraphics[width=\textwidth]{compmod_predictions.png}
 \end{figure}
-}
-
-\textcolor{blue}{To quantitatively evaluate the relative predictive performances of the models on the semantic interference data, we used a Bayes factor analysis.
+
+To quantitatively evaluate the relative predictive performances of the models on the semantic interference data, we used a Bayes factor analysis.
 We computed Bayes factors for the two-cues model compared to the three-cues model and to the one-cue model, respectively.
@@
 $Normal$(0.02,0.02), and 
-$Normal$(0.02,0.005).}
-
-\textcolor{blue}{The procedure for computing Bayes factors was as follows.
+$Normal$(0.02,0.005).
+
+The procedure for computing Bayes factors was as follows.
 The Bayes factor computation required us to estimate the marginal likelihood of the competing models.
@@
 We estimated the marginal likelihoods using Monte Carlo importance sampling as follows.
-}
-
-\textcolor{blue}{First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we used a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
+
+First, we chose an importance density $q(F)$ for the parameter $F$ from which we draw $0.2$ million samples of $F$; we used a truncated normal distribution $Normal_{lb=0,ub=0.05}(0,0.025)$ for the importance density.
 This importance density allowed us to obtain more proposals for smaller values of $F$ (for $F<0.03$) while maintaining fatter tails for the larger values of $F$ ($F>0.03$); effectively, the whole parameter space of $F$ between $0$ and $0.05$ was explored to draw $0.2$ million proposals.
 Then, for each of these proposed parameter values, we computed the likelihood and prior density.
-The estimated marginal likelihood of a model will be given by}
+The estimated marginal likelihood of a model will be given by
 
@@
 
-\noindent \textcolor{blue}{where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+\noindent where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
 The term $\mathcal{L}(F_i|y)$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.
-We computed marginal likelihoods under different prior assumptions $p(F)$.}
-
-\textcolor{blue}{The Lewis and Vasishth (2005) model is a complex, non-deterministic model which can be simplified such that a likelihood function represents the model's generative process  \parencite[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}.
+We computed marginal likelihoods under different prior assumptions $p(F)$.
+
+The Lewis and Vasishth (2005) model is a complex, non-deterministic model which can be simplified such that a likelihood function represents the model's generative process  \citep[e.g., in][]{NicenboimRetrieval2018,lisson2020computational}.
 However, it is also possible to carry out model comparison without recasting the model as a likelihood function.
@@
 However, if $\delta$ is too low, the computational demand increases because too many proposals would end up with zero likelihood.
-We chose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising on computational efficiency.}
-
-\textcolor{blue}{From the marginal likelihoods, we computed Bayes factors for the two-cues model compared to the three-cues model and the one-cue model.
+We chose a reasonably small value of $\delta$ to obtain a good approximation of the likelihood without compromising on computational efficiency.
+
+From the marginal likelihoods, we computed Bayes factors for the two-cues model compared to the three-cues model and the one-cue model.
 That is, the marginal likelihood of the two-cues model $M_1$ was divided by (i) the marginal likelihood of the three-cues model $M_2$,  and (ii) the marginal likelihood of the one-cue model $M_3$.
 This calculation resulted in the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
-Following the convention from \textcite{jeffreys1998theory}, a Bayes factor above $10$ is interpreted as strong evidence in favor of the two-cues model, and a value between $3$ and $10$ is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).}
-
-\textcolor{blue}{The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
+Following the convention from \cite{jeffreys1998theory}, a Bayes factor above $10$ is interpreted as strong evidence in favor of the two-cues model, and a value between $3$ and $10$ is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).
+
+The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
 Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model; this comparison shows that there is strong evidence for interference effects.
-Compared to the three-cues model proposed by \textcite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior.
+Compared to the three-cues model proposed by \cite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior.
 This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the two separate, previously assumed cues \{$\pm$grammatical subject\} and \{$\pm$same clause\}.
-This result holds for prior assumptions that allow higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.}
-
-\begin{figure}
+This result holds for prior assumptions that allow higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.
+
+\begin{figure}[ht]
     \centering
@@
     \label{fig:BF_compmods}
-    \includegraphics[width=\linewidth]{images/BF_plot_compmodels.png}
+    \includegraphics[width=\linewidth]{BF_plot_compmodels.png}
 \end{figure}
@@
 
-\Copy{only_this_design}{Taken together, the \textcolor{blue}{present} SPR and ERP experiments \textcolor{blue}{using the 2 $\times$ 2 interference design developed by \textcite{vandyke07}} consistently showed effects of semantic interference and \textcolor{blue}{little or no effects of syntactic interference or an interaction}.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
+{Taken together, the present SPR and ERP experiments using the 2 $\times$ 2 interference design developed by \cite{vandyke07} consistently showed effects of semantic interference and little or no effects of syntactic interference or an interaction.}\label{only_this_design} This pattern was not only found in the dependent measures of the experiments that were of primary interest, but also in the comprehension accuracy.
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
@@
 
-\textcolor{blue}{In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design.
-Then we discuss two important issues that relate to the broader interference literature}.
+In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design.
+Then we discuss two important issues that relate to the broader interference literature.
 The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.
 
-\subsection{\textcolor{blue}{Comparison with previous findings}}\label{comparison}
-\Copy{comparison}{\textcolor{blue}{The present study used the same design as two previous studies \parencite{vandyke07, mertzen}.
+\subsection{Comparison with previous findings}\label{comparison}
+{The present study used the same design as two previous studies \citep{vandyke07, mertzen}.
 Figure \ref{fig:previous_vs_present} A shows the estimated reading times differences from the previous studies and the present self-paced reading experiment.
-Because the \textcite{vandyke07} data was not available to us, we derived estimates for their study from the summary statistics in the published paper.
-\textcite{mertzen} made their data available.
+Because the \cite{vandyke07} data was not available to us, we derived estimates for their study from the summary statistics in the published paper.
+\cite{mertzen} made their data available.
 In comparison to the previous estimates, the effects of the present study were relatively small and associated with less uncertainty in general.
@@
 For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
-To compute Bayes factors for \textcite{vandyke07}, we used the Normal-Normal conjugate case \parencite{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper.
+To compute Bayes factors for \cite{vandyke07}, we used the Normal-Normal conjugate case \citep{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper.
 Then, using these posteriors and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970, dickey1971weighted, verdinelli1995computing}.
 Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
-The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}
-
-\begin{sidewaysfigure}[h]
-    \caption{A) Reading time differences in the regions of interest from the present and previous studies using the design by \textcite{vandyke07}.
+The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.
+
+\begin{sidewaysfigure}[htpb]
+    \caption{A) Reading time differences in the regions of interest from the present and previous studies using the design by \cite{vandyke07}.
 VD07.E1, VD07.E2 and VD07.E3 stand for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiments 1-3.
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{images/present_previous_allplots.jpg}
+    \includegraphics[width=\linewidth]{present_previous_allplots.jpg}
 \end{sidewaysfigure}
@@
 
-\textcolor{blue}{In the pre-critical region, the Bayes factors from the present and previous studies showed no clear pattern regarding syntactic interference.
+In the pre-critical region, the Bayes factors from the present and previous studies showed no clear pattern regarding syntactic interference.
 The present study was the only one to provide evidence against the effect, while the previous studies supported it.
-The Bayes factors aligned qualitatively across studies for the semantic interference effect and the interaction -- consistently providing evidence for semantic interference and against the interaction.}
-
-\textcolor{blue}{In the critical region, the Bayes factors across studies showed a mixed picture for syntactic interference: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extremely strong evidence for it, while \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3 provided weak evidence against it.
+The Bayes factors aligned qualitatively across studies for the semantic interference effect and the interaction -- consistently providing evidence for semantic interference and against the interaction.
+
+In the critical region, the Bayes factors across studies showed a mixed picture for syntactic interference: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extremely strong evidence for it, while \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3 provided weak evidence against it.
 \citeauthor{mertzen}'s (\citeyear{mertzen}) experiments showed weak evidence for the effect.
@@
 Similarly, there is no clear pattern for semantic interference: Most experiments provided evidence against it, but the two self-paced reading experiments provided evidence for semantic interference.
-The Bayes factors across studies provided evidence against the interaction.}
-
-\textcolor{blue}{In the spill-over region, the Bayes factors for syntactic interference rather consistently provided evidence against the effect.
+The Bayes factors across studies provided evidence against the interaction.
+
+In the spill-over region, the Bayes factors for syntactic interference rather consistently provided evidence against the effect.
 Only \citeauthor{mertzen}'s (\citeyear{mertzen}) English experiment provided moderate evidence for it.
@@
 The Bayes factor results for the interaction showed evidence against it from all experiments but one.
-The outlier was \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment, which showed extreme evidence for the interaction.}
-
-\textcolor{blue}{In sum, while the present and previous studies mostly showed comparable region-wise results for semantic interference and the interaction, they differed remarkably regarding syntactic interference in the pre-critical and critical region.
+The outlier was \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment, which showed extreme evidence for the interaction.
+
+In sum, while the present and previous studies mostly showed comparable region-wise results for semantic interference and the interaction, they differed remarkably regarding syntactic interference in the pre-critical and critical region.
 In the pre-critical region, the present study was the obvious outlier, being the only study to provide evidence against the effect.
@@
 The lack of syntactic interference in the pre-critical region in the present study compared to the previous studies could be explained by differences in i) methodology, ii) items and/or iii) statistical power.
-We discuss each of these differences in turn.}
+We discuss each of these differences in turn.
 }
@@
 An obvious difference between the studies is the experimental methodology.
-The majority of the previous experiments employed eye-tracking while reading, while we used self-paced reading \textcolor{blue}{and EEG / ERPs.
-However, word-by-word} self-paced reading (SPR) \textcolor{blue}{and ERPs elicited by single words presented in the typical rapid serial presentation (RSVP) format, if anything} might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR} \textcolor{blue}{and RSVP.
-Word-by-word presentation} is likely to be more demanding on the comprehender's memory, which \textcolor{blue}{should} make it easier to detect interference effects.
-\textcolor{blue}{So, the differences in methodology do not offer a straightforward explanation for the lack of syntactic interference in our data compared to the previous studies.}
-
-\Copy{linebreak}{\textcolor{blue}{The difference in methodology led to another difference between our study and the previous studies.
-In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line.
+The majority of the previous experiments employed eye-tracking while reading, while we used self-paced reading and EEG / ERPs.
+However, word-by-word self-paced reading (SPR) and ERPs elicited by single words presented in the typical rapid serial presentation (RSVP) format, if anything might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR} and RSVP.
+Word-by-word presentation is likely to be more demanding on the comprehender's memory, which should make it easier to detect interference effects.
+So, the differences in methodology do not offer a straightforward explanation for the lack of syntactic interference in our data compared to the previous studies.
+
+{The difference in methodology led to another difference between our study and the previous studies.
+In their eye-tracking experiments, \cite{vandyke07} and \cite{mertzen} presented the stimuli on a single line.
 Whether \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment used line breaks or not is not reported.
@@
 The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor.
-It has been proposed by \textcite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability.
+It has been proposed by \cite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability.
 Therefore, increased prominence of the subject could have decreased interference.
@@
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.}
+Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.
 }\label{linebreak}
 
-\textcolor{blue}{We now turn to the differences regarding the experimental items.
+We now turn to the differences regarding the experimental items.
 The three studies investigating syntactic and semantic interference with the same design used either English or German items.
@@
 This is in line with no apparent clustering of the English vs.\ German results (see Figure \ref{fig:previous_vs_present} A).
-So, the use of materials in different languages cannot account for the differences in results between the studies.}
-
-\textcolor{blue}{The previous and present items differed in the pre-critical material.
+So, the use of materials in different languages cannot account for the differences in results between the studies.
+
+The previous and present items differed in the pre-critical material.
 \citet{vandyke07} used one region consisting of two words (e.g., `yesterday afternoon') which were analyzed together.
-\textcite{mertzen} used a one-word pre-critical region (e.g., `tatsächlich', indeed).
+\cite{mertzen} used a one-word pre-critical region (e.g., `tatsächlich', indeed).
 In the present study, we added a pre-pre-critical region to help us differentiate between the potential sources of the pre-critical effects in the literature as discussed in the Discussion of the SPR experiments.
@@
 While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region.
-However, Figure \ref{fig:whole_sentence} C shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.}
-
-\Copy{comma}{\textcolor{blue}{Another item-related difference is that all our experiments (SPR and EEG) displayed the subject with a comma.
+However, Figure \ref{fig:whole_sentence} C shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.
+
+{Another item-related difference is that all our experiments (SPR and EEG) displayed the subject with a comma.
 Commas are mandatory in German to separate subordinate clauses from the main clause.
@@
 But \citet{mertzen} found syntactic interference and we did not.
-So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.}}\label{comma}
+So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.}\label{comma}
 
 The contrast between our results and those from  \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment is surprising given that both used German items and partially even the same linguistic material.
-\Copy{items_mertzen}{\textcolor{blue}{However, \citet{mertzen} changed the structure of the relative clause modifying the subject by adding an additional animate distractor to all conditions (see an example of one of their English items in \ref{ex:mertzen}; their items had this structure cross-linguistically).
+{However, \citet{mertzen} changed the structure of the relative clause modifying the subject by adding an additional animate distractor to all conditions (see an example of one of their English items in \ref{ex:mertzen}; their items had this structure cross-linguistically).
 This was done to ``increase the strength of the manipulation.'' \citep[][p.
 9]{mertzen}.
-}
 
@@
 \ex \label{ex:mertzen} 
-\textcolor{blue}{It turned out that the attorney whose secretary had forgotten that the visitor was important frequently complained about the salary at the firm. \parencite{mertzen}}
+It turned out that the attorney whose secretary had forgotten that the visitor was important frequently complained about the salary at the firm. \citep{mertzen}
 \end{exe}
@@
 
-\textcolor{blue}{The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
+The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
 But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
 This additional distractor might have increased syntactic interference compared to our study.
-But it would have increased syntactic interference in comparison to \citeauthor{vandyke07}'s study as well and this was not the case (see Figure \ref{fig:previous_vs_present} A).}\label{items_mertzen} \textcolor{blue}{All in all, differences in the items provide no clear explanation why the previous studies found syntactic interference and the present study did not.}
-
-\textcolor{blue}{A plausible explanation for the divergence between studies is a difference in statistical power:} The present study has higher statistical power than the previous studies.
+But it would have increased syntactic interference in comparison to \citeauthor{vandyke07}'s study as well and this was not the case (see Figure \ref{fig:previous_vs_present} A).\label{items_mertzen} All in all, differences in the items provide no clear explanation why the previous studies found syntactic interference and the present study did not.
+
+A plausible explanation for the divergence between studies is a difference in statistical power: The present study has higher statistical power than the previous studies.
 \citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively.
-\textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
-In contrast, \textcolor{blue}{in our SPR experiments,} we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
+\cite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each.
+In contrast, in our SPR experiments, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items).
 Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design.
 Consequently, the estimates from the present study are the most precise ones so far for this design.
-\textcolor{blue}{This is reflected by the tighter estimates in Figure \ref{fig:previous_vs_present} A}.
+This is reflected by the tighter estimates in Figure \ref{fig:previous_vs_present} A.
 The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.
-\textcolor{blue}{In contrast, the large estimates of the previous studies are likely Type M errors, i.e., exaggerations, which are common in low-powered studies \parencite{gelman_carlin}.
-That at least some of the previous studies had low power is further emphasized by the large uncertainty associated with some of the previous estimates (see Figure \ref{fig:previous_vs_present} A).}
-
-\textcolor{blue}{In sum, while there are differences in the employed methodology and items between the present and previous studies, the most likely cause of the different results is the difference in statistical power.
-However, an alternative explanation could be that syntactic interference might differ inter-individually \parencite{yadav2021individual} and that the previous studies by chance sampled disproportionally many participants who showed syntactic interference, while we by chance sampled participants who did not.}
-
-\Copy{not_an_anomaly}{\textcolor{blue}{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature.
+In contrast, the large estimates of the previous studies are likely Type M errors, i.e., exaggerations, which are common in low-powered studies \citep{gelman_carlin}.
+That at least some of the previous studies had low power is further emphasized by the large uncertainty associated with some of the previous estimates (see Figure \ref{fig:previous_vs_present} A).
+
+In sum, while there are differences in the employed methodology and items between the present and previous studies, the most likely cause of the different results is the difference in statistical power.
+However, an alternative explanation could be that syntactic interference might differ inter-individually \citep{yadav2021individual} and that the previous studies by chance sampled disproportionally many participants who showed syntactic interference, while we by chance sampled participants who did not.
+
+{While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature.
 In psycholinguistics alone, contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{taboretal04, lcpaape2023}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
-The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}}\label{not_an_anomaly}
+The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}\label{not_an_anomaly}
 
@@
 In the SPR data, we found that the semantic interference effect originated at the distractor and persisted throughout the following regions.
-Due to this time course, encoding interference \parencite{Yadavetal2022,hammerly2019grammaticality} is the best explanation for the interference effect\textcolor{blue}{s in our SPR data}.
+Due to this time course, encoding interference \citep{Yadavetal2022,hammerly2019grammaticality} is the best explanation for the interference effects in our SPR data.
 This raises the question whether the ERP effect was also caused by encoding interference, i.e., whether the ERP effect originated prior to the critical region.
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
     \caption{ERPs elicited by the pre-pre-critical adverb (word onset at -1800 ms), pre-critical adverb (word onset at -900 ms) and critical verb (word onset at 0\,ms) at electrode Cz.}
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{images/N_103_Cz_precrit.png}
+    \includegraphics[width=0.95\textwidth]{N_103_Cz_precrit.png}
 \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}[ht]
     \centering
         \caption{ERPs elicited by the distractor (word onset at 0 ms) at eight selected electrodes.}
-    \label{fig:erp_distractor}\includegraphics[width=\textwidth]{images/Distractor_N_100_some_elec.png}
+    \label{fig:erp_distractor}\includegraphics[width=\textwidth]{Distractor_N_100_some_elec.png}
 \end{figure}
@@
 Having said that, we cannot rule out the possibility that -- due to the auto-paced nature of the ERP experiment -- any encoding interference that started at or after the distractor is mixed in with the cost of retrieval interference.
-Indeed, recent modeling work on agreement attraction in reading \parencite{Yadavetal2022} has shown that some kind of feature distortion as well as cue-based retrieval are needed to explain the data-sets that were publicly available at the time that the modeling was done.
-If the  \textcite{Yadavetal2022} account extends to interference effects in general,
+Indeed, recent modeling work on agreement attraction in reading \citep{Yadavetal2022} has shown that some kind of feature distortion as well as cue-based retrieval are needed to explain the data-sets that were publicly available at the time that the modeling was done.
+If the  \cite{Yadavetal2022} account extends to interference effects in general,
 it is reasonable to assume that the semantic interference effect observed at the verb in the present ERP study is driven by both encoding and retrieval interference.
@@
 
-The \textcolor{blue}{lack of decisive evidence for} a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \textcite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
-\Copy{future_work}{\textcolor{blue}{A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
-This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.}}\label{future_work}
+The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
+{A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
+This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.}\label{future_work}
 
@@
 The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands.
-It is reasonable to assume that the participants might have -- at least occasionally -- adopted a good-enough processing mode when faced with high task demands \parencite{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \parencite{MalsburgVasishth2013}, or both.
+It is reasonable to assume that the participants might have -- at least occasionally -- adopted a good-enough processing mode when faced with high task demands \citep{swets2008underspecification,LogacevMultiple,LogacevVasishthQJEP2016} or due to working memory capacity limitations \citep{MalsburgVasishth2013}, or both.
 So, given a good-enough processing mode which leaves some syntactic relations within the sentence underspecified, it makes sense for the comprehender to use a simple heuristic for subject-verb dependency resolution.
@@
 representation of event structures.
-This view has been put forward by \textcite{rabovsky_etal_2018}, when they used the neural-network sentence gestalt model \parencite{mcclelland1989_sentence_gestalt}, to model the N400 amplitude.
+This view has been put forward by \cite{rabovsky_etal_2018}, when they used the neural-network sentence gestalt model \citep{mcclelland1989_sentence_gestalt}, to model the N400 amplitude.
 This model assumes that language comprehension relies on associative form-to-meaning mapping instead of syntactic rules.
@@
 First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}.
-Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \parencite[e.g.,][]{Avetisyan_etal_2020_agree,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \parencite[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and
-the importance of word order \parencite[e.g.,][]{meng2000mode}.
+Second, there is plenty of independent evidence for the central role of syntactic information in comprehension; some examples are the role of case marking \citep[e.g.,][]{Avetisyan_etal_2020_agree,HusainEtAl2014,bhatia2022processing,bader2000case,bader2006case,miyamoto2002case}, syntactic constraints like Principle A of the binding theory \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual},  and
+the importance of word order \citep[e.g.,][]{meng2000mode}.
 
 There are of course important examples of the parser ignoring syntactic information.
-Three dramatic examples are the grammaticality illusion in multiple embeddings in English \parencite{gibsonthomas99}, syntactic local coherence effects \parencite{taboretal04}, and the agreement attraction phenomenon across different languages \parencite[e.g.,][]{wagersetal,lago_etal_2021,tucker2015representing}.
+Three dramatic examples are the grammaticality illusion in multiple embeddings in English \citep{gibsonthomas99}, syntactic local coherence effects \citep{taboretal04}, and the agreement attraction phenomenon across different languages \citep[e.g.,][]{wagersetal,lago_etal_2021,tucker2015representing}.
 However, there exists  evidence inconsistent with these findings.
-Regarding grammaticality illusions in double-center embeddings, the grammaticality illusion does not occur in German and Dutch \parencite{VSLK11,FrankEtAl2015}.
-Regarding syntactic local coherence effects, a large-sample study \parencite{lcpaape2023} presents evidence against syntactic local coherence.
-Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of number agreement attraction \parencite{chromy2023number}.
+Regarding grammaticality illusions in double-center embeddings, the grammaticality illusion does not occur in German and Dutch \citep{VSLK11,FrankEtAl2015}.
+Regarding syntactic local coherence effects, a large-sample study \citep{lcpaape2023} presents evidence against syntactic local coherence.
+Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of number agreement attraction \citep{chromy2023number}.
 Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.
@@
 The observed differences in the time course of the effect are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-Overall, we found \textcolor{blue}{no decisive evidence for} the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+Overall, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
 A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
@@
 \newpage
-\printbibliography
-
+
+%% If you have bibdatabase file and want bibtex to generate the
+%% bibitems, please use
+%%
+\bibliographystyle{elsarticle-harv} 
+\bibliography{interferencebib}
+
+%% else use the following coding to input the bibitems directly in the
+%% TeX file.
+
+% \begin{thebibliography}{00}
+
+% %% \bibitem[Author(year)]{label}
+% %% Text of bibliographic item
+
+% \bibitem[ ()]{}
+
+% \end{thebibliography}
 \end{document}
+
+\endinput
+%%
+%% End of file `elsarticle-template-harv.tex'.

### 2024-11-01 paper/revision2/ResponseToReviewers.tex
Add files via upload

\documentclass[12pt]{article}
\usepackage[a4paper,top=2cm,bottom=2cm,left=2.5cm,right=2.5cm,marginparwidth=1.75cm]{geometry}
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{float}
\usepackage{subfigure}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{color}
\usepackage{gb4e}
\noautomath

%\usepackage[backend=biber,useprefix=true,style=apa,url=false,sorting=nyt,eprint=false]{biblatex}

%\addbibresource{interferencebib.bib}

\usepackage[natbib=true,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
\usepackage{lineno,xcolor,clipboard,graphicx}

\usepackage{csquotes}
\DeclareLanguageMapping{american}{american-apa}
\bibliography{interferencebib}

\usepackage{marginnote}
\setcounter{section}{-1}

\openclipboard{output-reviews}

\renewcommand{\qedsymbol}{\rule{0.7em}{0.7em}}

\usepackage{epstopdf}
\newcommand{\revised}[1]{{\color{black}{#1}}}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\title{Response to editor and reviewers (JML-24-17)}
\author{Pia Schoknecht and Shravan Vasishth}
%\date{}                                           % Activate to display a given date or no date

\usepackage{xr}

\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
\typeout{(#1)}% latexmk will find this if $recorder=0
% however, in that case, it will ignore #1 if it is a .aux or 
% .pdf file etc and it exists! If it doesn't exist, it will appear 
% in the list of dependents regardless)
%
% Write the following if you want it to appear in \listfiles 
% --- although not really necessary and latexmk doesn't use this
%
\@addtofilelist{#1}
%
% latexmk will find this message if #1 doesn't exist (yet)
\IfFileExists{#1}{}{\typeout{No file #1.}}
}\makeatother

\newcommand*{\myexternaldocument}[1]{%
\externaldocument{#1}%
\addFileDependency{#1.tex}%
\addFileDependency{#1.aux}%
}

\myexternaldocument{elsarticle-template-harv}

\begin{document}
\hfill \today
%\section{}
%\subsection{}

\noindent \textbf{Prof.\ Jarrold}\\
\noindent \textbf{Associate Editor}\\
\noindent \textbf{Journal of Memory and Language}
\vskip 1em

\noindent \textbf{Response letter for Submission JML-24-17, 2nd round of revisions}
\vskip 2em
\noindent Dear Prof. Jarrold, 
\vskip 1em
Together with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit our second revision of the article titled ``Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German''.
We are grateful for thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.

%add description of revisions here

Below, we separately address each comment from the reviewers and yourself. All changes have been highlighted in the revised manuscript. The actionable parts of the comments are highlighted in bold.

\vskip 1em
\noindent Sincerely,
\vskip 1em
\noindent Pia Schoknecht\\
Postdoctoral researcher\\
Department of Linguistics\\
University of Potsdam, Germany
\thispagestyle{empty}
\newpage

\section*{Editor's comments} 
\subsection*{Comment E.1}
\begin{quote}
``Dear Dr. Schoknecht,

Thank you for submitting your revised manuscript to the Journal of Memory and Language. This has now been reviewed by the same three experts who reviewed the previous submission, and you can find their comments below.

You will see from these that all three reviewers appreciate the thoroughness with which you responded to the issues raised on the initial submission. Reviewers 1 and 3 raise some remaining points but recommend acceptance, while Reviewer 2 is less persuaded that your revisions have adequately addressed their concerns. Given this set of reviews, I have come to a `revise' decision because I would like you to have one more go at addressing the remaining points made by all three reviewers.

I do not anticipate sending out the next version for review again, but rather anticipate being in a position to accept it at that point. However, it's important to note that that is not guaranteed, and one reason I have come to a `revise' rather than `accept subject to minor revisions' is that, like Reviewer 1, I was unable to access your data via your OSF link. \textbf{It is very important to us at JML that data are readily accessible, and that accessibility needs to be in place before any potential acceptance decision}. 
''
\end{quote}

\subsubsection*{Response to E.1}
\textcolor{red}{We have made the OSF repository public and updated the link in the manuscript.}

\subsection*{Comment E.2}
\begin{quote}
``The other key points I would like to ask you to pay particular attention to in any response to the current set of reviews are:

\textbf{Reviewer 2 and 3's clear view that you still overstate the implications of your findings in places}.''
\end{quote}

\subsubsection*{Response to E.2}
\textcolor{red}{add *in this design* whenever we say that there is no syn interference}

\subsection*{Comment E.3}
\begin{quote}
``\textbf{Reviewer 2's additional point 1 that asks whether you have changed that particular analysis}.''
\end{quote}

\subsubsection*{Response to E.3}
Yes, we changed all analyses in the 1st revision. As we had explained in the response to reviewers on August 29, 2024 (specifically in the response letter and our responses to the comments E.2 and R3.5), the priors were changed from directional priors in the original version to symmetrical priors in the revised version.

\subsection*{Comment E.4}
\begin{quote}
``and \textbf{their final point about the length of the current version of the manuscript}.''
\end{quote}

\subsubsection*{Response to E.4}
\textcolor{red}{remove repetitions}

\section*{Reviewer \#1}

\subsection*{Comment R1.1}
\begin{quote}
``I was the Reviewer \#1 of the previous version of the manuscript.

All my concerns are addressed in this revision, so I am happy to say I don't see reasons to ask for resubmission or for another round of major, significant revisions. I found it a solid paper in the first round already, but with the revisions and especially given the modified and extended analysis, I find the whole story more convincing and also, fair.

I do have some suggestions. In my view these are only minor. They mainly concern just the way things are presented or discussed. I list my comments by page.

p. 16: osf data etc.

I wanted to check those, but \textbf{when I clicked the link, it said I don't have permissions and that I need to request access}.''
\end{quote}

\subsubsection*{Response to R1.1}
\textcolor{red}{We have made the OSF repository public and updated the link in the manuscript.}

\subsection*{Comment R1.2}

\begin{quote}
``p. 17, SPR experiment

\textbf{Did the analysis include any cut-off point for reading times to exclude data? I could not find this in the manuscript. If it did, could this information be included?} If not, I would strongly suggest to use this. In my own experience with Prolific, I noticed the following very common pattern: people basically do not read around 10-20\% of experimental items - that is, for quite a significant subset of participants I noticed that they have super fast reading times on some subset of items, most likely because they just rushed through those items by just pressing the space bar and keeping it pressed. This is sometimes followed by a long waiting time at some other point, so those participants would not be outliers when we check the total time in the experiment. These people and their data would not be excluded by checking their answers to comprehension questions, especially if only a small subset of items has comprehension questions, like one third, as in this experiment (basically, they could still easily pass with around 90\% success rate). So if this was not checked, it would be good to check and \textbf{consider removing RTs based on small/large cut-off points and to see whether that affects the analysis. If this was done already, it would be good to report more details, in particular, the cut-off points and also the amount of data that was removed this way}.''
\end{quote}

\subsubsection*{Response to R1.2}
\textcolor{red}{We trimmed the reading times (between X and Y ms) before analysis and have now added this information to the manuscript.}

\textcolor{red}{Paste text}

\subsection*{Comment R1.3}
\begin{quote}
``p. 36: ``Regarding the possibility that sentence structure confounds caused the
effects in the pre-critical region(.)"

\textbf{I wonder whether it would make sense to briefly summarize what the sentence structure confound should be.} Otherwise the whole paragraph is hard to follow for readers who did not read or do not remember Mertzen et al. (2023).''
\end{quote}
\subsubsection*{Response to R1.3}
\textcolor{red}{We have added the following text to the manuscript on page X:}

\subsection*{Comment R1.4}
\begin{quote}
``p. 51: ``For more discussion of the role of structural retrieval cues, see Franck and Wagers (2020) and Arnett and Wagers (2017)."

\textbf{It would be good to cite Kush et al. (2015) here} (after all, a significant portion of Franck and Wagers, 2020, builds on Kush et al. 2015 to explore ways to deal with structural cues).

Kush et al. (2015) Relation-sensitive retrieval: Evidence from bound variable pronouns. Journal of memory and language.''
\end{quote}
\subsubsection*{Response to R1.4}
We have added the suggested reference (now on page \pageref{cite_kush}):

\begin{quote}
    \Paste{cite_kush}
\end{quote}

\subsection*{Comment R1.5}
\begin{quote}
``p. 55: ``$\Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$"

I was a bit confused here, since the formula  ``$\Psi(d(x_i, y), 0 , \delta)$ " did not have any  ``$\Psi(.|\delta)$" part.''
\end{quote}

\subsubsection*{Response to R1.5}
\textcolor{red}{Opened issue and assigned to Himanshu}

\subsection*{Comment R1.6}

\begin{quote}
``p. 61, \textbf{Figure 14: one value is outside the scale of the bottom left graph. If possible, it would be good to make it visible}.''
\end{quote}

\subsubsection*{Response to R1.6}
\textcolor{red}{Check how the figure changes}

\subsection*{Comment R1.7}
\begin{quote}
``p. 64: ``Word-by-word presentation is likely to be more demanding on the comprehender's memory, which should make it easier to detect interference effects. So, the differences in methodology do not offer a straightforward
explanation for the lack of syntactic interference in our data compared to the
previous studies."

Actually, I thought here the authors basically provided an explanation of the differences between their findings and the previous findings because of differences in the methodology, even though their last sentence said otherwise. Let me spell out it: since the current methodology (SPR, EEG) is likely more demanding on memory, it is possible that people give up on detailed processing; rather, they do something akin to good-enough processing and do not fully process; consequently, they could still get interference from semantics (since the semantic interference is really simple, it is just lexical semantics and readers don't need to fully parse for that) but they don't correctly parse embedded subjects, hence the syntactic interference is gone. \textbf{The authors in fact come to good-enough processing in Section 7.3, but then it is not linked back to differences between their findings and the other findings (which I think should be, I do think this is a possible explanation).}''
\end{quote}

\subsubsection*{Response to R1.7}
\textcolor{red}{Follow the suggestion.}

\subsection*{Comment R1.8}
\begin{quote}
``p. 76: \textbf{Chromy should appear like this: ``Chromý" (note the correct diacritics above the "y")}

Thank you for your manuscript!''
\end{quote}
\subsubsection*{Response to R1.8}
We have corrected the spelling of Jan Chromý's last name.

\section*{Reviewer \#2} 
\subsection*{Comment R2.1}
\begin{quote}
``I appreciate the authors' responses to my previous comments. However, despite the revisions, significant weaknesses remain, rendering in my opinion this study primarily a methodological contribution that adds little value to our understanding of interference effects. Additionally, the authors' strong claims often come through as overstatements that could result in serious misconceptions.
My overall assessment is as follows.

1. Interpretability of the SPR results: The results from the self-paced reading (SPR) task remain unclear. The only observable effect—the semantic interference—emerges well before the critical region of interest (specifically, at the distractor region) and persists without change beyond the critical region. As a result, any effect at the critical region is uninterpretable.

The authors acknowledge the interpretability issue with the SPR results in their response to my comment, and they now include the following statement: ``Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over regions, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them. Therefore, we refrain from further discussing effects occurring in the later regions." \textbf{However, this comes after more than 10 full pages (pp. 25-37) of discussion on the reaction times (RTs) in the critical and pre-critical regions, which would lead any reader to believe these findings are meaningful. If, as the authors now realize, the effect emerges much earlier, this entire discussion is misleading: the SPR results after the distractor region are simply uninterpretable.}''
\end{quote}

\subsubsection*{Response to R2.1}
We agree with the review that the discussion of the SPR results was misleading and have revised it accordingly. We now start the discussion by discussing the effects at the distractor (see Response to R2.2) and then briefly discuss the later effects.

\textcolor{red}{Revise text and paste here}

\subsection*{Comment R2.2}
\begin{quote}
``However, there are also problems with the interpretation of the effect at the distractor region. If the authors want to attribute this effect to encoding interference, then they must argue for the existence of long-lasting encoding effects, as this very same effect persists throughout the entire sentence. Yet, the mechanism behind such enduring encoding effects remains unclear, and the authors do not address this challenge. Feature overlap, a commonly proposed mechanism for encoding interference, seems unlikely. It would suggest that the two elements compete for the same feature throughout the entire sentence, which is not plausible. Activation leveling, another potential mechanism, is also improbable, as it equalizes the activation of competing elements and should result in similar reaction times at some point, which we do not see here. \textbf{What is evident from these findings is that the inanimate distractors are read faster than the animate ones (but the opposite pattern is found in ERPs, see point 3), and this preference creates a reaction time difference that persists until and beyond the critical verb. If the authors wish to interpret this as evidence of encoding interference, they must provide compelling reasons to support the idea that encoding interference can have such a prolonged effect, along a clear mechanism for it}.''
\end{quote}
\subsubsection*{Response to R2.2}
The focus of the present study was to investigate retrieval interference due to syntactic and semantic similarity. The finding that the reading times were likely affected by encoding interference and not by retrieval interference was unexpected (but see \citeauthor{mertzen}'s (\citeyear{mertzen}) discussion). Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that an in-depth discussion of encoding interference is beyond its scope. To address the reviewer's concern that encoding interference would not cause such long-lasting effects as found in our SPR data, we have added the following text to the manuscript about a recent joint account of encoding and retrieval interference:

\textcolor{red}{Write someting about Himanshu's work in relation to this}

\subsection*{Comment R2.3}

\begin{quote}
``2. Lack of syntactic interference evidence: No evidence for syntactic interference was found in either task, contradicting two previous studies that tested similar structures (Van Dyke 2007; Mertzen et al. 2023). \textbf{In my previous review, I noted that the syntactic manipulation used was not appropriate and thus unlikely to induce interference effects. Although the authors concede that this might be true, they did not address the logical implications of this acknowledgment, as evidenced by their repeated strong assertions}, such as the one in the abstract: ``Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to syntactic cues".
If the syntactic manipulation is irrelevant, then repeatedly concluding that syntactic interference is absent is not only an overstatement but also misleading. All this study allows us to conclude is that when cues that are irrelevant to syntax are manipulated (like [+being a subject of whichever clause at whichever level of embedding], underpowered studies (Van Dyke 2007; Mertzen et al. 2023) may falsely report interference effects. While this is a valid conclusion and a sound methodological critique, it does not significantly advance our theoretical understanding of syntactic interference. Specifically, it fails to demonstrate that when relevant syntactic cues are manipulated, syntactic interference is not attested. Yet, this is the conclusion the authors consistently imply throughout their manuscript, leading to significant misconceptions. As I mentioned in my previous review, beyond the two studies the authors are building on, numerous studies in the literature show that hierarchically intervening elements do generate strong syntactic interference.
It is a fine enterprise to correct inaccurate claims from previously underpowered studies that used irrelevant manipulations; however, this should not not suggest that the findings presented here have a broader significance than they do.''
\end{quote}
\subsubsection*{Response to R2.3}
We agree with the reviewer that our present work does not rule out syntactic interference in general, but that the present data just suggests that the specific design used in this work did not induce considerable syntactic interference. We have made the following edits in the manuscript to emphasize this point:

\textcolor{red}{revise text and paste here}

\textcolor{red}{SV email: i think we didn't write clearly enough about the fact that syntactic interference *in this design* is expected given the previous work. we did say it but needs to be clearer. also, when we say we find evidence against syn int, we should always add, in this design. and then we should say that this does not mean that there is no syntactic manipulation that would cause no interference, there could be. i think that point is valid. what i found strange in her comment was that she says both that van dyke and mertzen found syn int, but also that syn int is never going to show up in this design because it is not the right manipulation.  maybe her point is that we only show that the van dyke and mertzen result was Type M error (i.e., there is no real effect in this design).}

\subsection*{Comment R2.4}
\begin{quote}
``3. Misalignment of SPR and ERP results: The results from the self-paced reading (SPR) and event-related potential (ERP) methods do not align, with the only observed effect—semantic interference—pointing in opposite directions: inhibitory in SPR but facilitatory in ERP. Specifically, the authors report a reversed effect in the P600 component, showing a reduced P600 in conditions of higher semantic interference. This contradicts typical predictions, where the P600 is usually greater under high interference. The authors interpret this as a facilitatory interference effect (i.e., easier processing with high semantic interference). However, if this were the case, we would also expect faster reaction times (RTs) in high semantic interference conditions, while the opposite is observed in the SPR data.

4. Conflicting directions in ERP components: Within the ERP results, the different components also point in opposite directions. The P600 component shows a reversed effect, with a reduced P600 in conditions of high semantic interference. However, the N400 component follows the expected pattern, with stronger activation under high semantic interference. It remains unclear how the directionality of these two can be reconciled within the same theoretical framework.''
\end{quote}
\subsubsection*{Response to R2.4}
\textcolor{red}{Not sure whether this comment requires action.}

\subsection*{Comment R2.5}
\begin{quote}
``5. Unclear conflicting mechanisms: The authors attribute the SPR results to encoding interference and the ERP results to retrieval interference, leading to a contradictory conclusion. I won't delve further into this point, as the issues raised earlier already highlight more fundamental problems.''
\end{quote}
\subsubsection*{Response to R2.5}
\textcolor{red}{Not sure whether this comment requires action. Maybe 1-2 sentences about Himanshu's work.}

\subsection*{Comment R2.6}
\begin{quote}
``Additional points:

1. Original version: ``By contrast, the syntactic interference effect had only anecdotal evidence under the two narrow priors (Normal-(0, 0.1): BF10 = 2.5, Normal-(0, 0.5): BF10 = 1.2). There was evidence against syntactic interference under wider priors (Normal-(0, 1): BF10 = 0.66, Normal-(0, 5): BF10 = 0.14). What this means is that, in our data, only if we assume a priori that the effect size is relatively small, there is very weak evidence in favor of syntactic interference."

Revised version: ``By contrast, Bayes factors provided either no evidence for or even evidence against syntactic interference in both spatiotemporal windows (BF10 $<$ 1.2). Similarly, Bayes factors provided either no evidence for or even evidence against the interaction in both spatio-temporal windows (BF10 $<$ 1.2)."

\textbf{Did the authors change the analyses? Why what was reported as anecdotal evidence under narrow priors is now reported as no evidence or evidence against?}''
\end{quote}
\subsubsection*{Response to R2.6}
Yes, the priors were changed in the 1st revision (directional priors in the original version and symmetrical priors in the revised version, see also our Response to E.3). This resulted in small numerical changes of the results, e.g., BF10 = 2.5 in the original analysis decreased to BF10 = 1.2 in the revised analysis. Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2. Since the first revision and in line with \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence.

\subsection*{Comment R2.7}
\begin{quote}
``2. As a side note, on p. 58, the authors conclude that ``the parser searches for a subject that is within the same clause but uses the animacy cue without reference to the clause in which a noun appears." Does this mechanism truly seem plausible to the authors? What would this imply in practice? \textbf{Does it suggest that semantic and syntactic information are entirely encapsulated from each other?} If we were to set aside the issues I raised earlier, this would be one of the paper's central conclusions, yet the authors fail to provide a credible mechanism to support it.''
\end{quote}
\subsubsection*{Response to R2.7}
\textcolor{red}{SV email: the syn / sem comment is also interesting because there are some theories that only assume semantic cue association--there isn't even any syntax. some discussion of the theoretically possible connections between syn and sem processing would help. e.g., one extreme is that syn and sem operate in tandem (e.g., categorial grammars), others believe that sem processing strictly follows syntactic structure building (classical montagovian stuff), then we have the sem only models (e.g., sentence gestalt model). there is no consensus on what the relationship is between syn and sem, so these could operate independently. local coherence is also an example--if we assume syntactic processing and no independent semantic processing, local coherence would not arise (depending on how one assumes that syntactic processing works).}

\subsection*{Comment R2.8}
\begin{quote}
``3. How do the fillers differ from the experimental items? In the example the authors reported there is retrieval at place again:

Experimental: The neighbor believed that the widower, who had told her that the loss was awful, regularly drank in the evenings to forget.\\
Filler: The carpet maker, who came to the workshop early, repaired the especially beautiful old carpet while listening to the news.

The authors state, ``The fillers were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety." \textbf{In what way did they provide more variety? In which respects? Both fillers and test items involve retrieval; what is the role of the fillers in this context?}''
\end{quote}
\subsubsection*{Response to R2.8}
\textcolor{red}{Describe why fillers should be comparable in complexity to the experimental items and how they distracted the participants from the purpose of the experiment}

\subsection*{Comment R2.9}
\begin{quote}
``4. There are numerous repetitions throughout the paper. A prime example is the paragraph on ``hyp testing using BF" (p. 45), which is repeated almost verbatim in the ``Discussion" section (p. 47). This redundancy occurs several times, making the paper excessively and unnecessarily long. The authors should streamline the content to enhance clarity and conciseness.''
\end{quote}

\subsubsection*{Response to R2.9}
\textcolor{red}{remove repetitions}

\section*{Reviewer \#3}

\subsection*{Comment R3.1}
\begin{quote}
``This is a revision of a paper that I previously reviewed. (I was Reviewer \#3 in the previous round.)

In the previous round of reviews, I was generally happy with the authors' findings. My queries were mostly related to the interpretation and generalizability of those findings. This is a routine issue in psycholinguistic research. A study looks at effects in one very specific (and often complicated) sentence type, and then draws conclusions about very broad notions such as `syntax' and `semantics'.

The authors have offered a thorough (daunting?) response to the reviews. I am more satisfied with some responses than others. But I do not think that this should stand in the way of publication in the JML special issue.

The authors acknowledge the concern about using relational notions such as `subject of the same clause' as a memory retrieval cue. They provide some text that clarifies that they need to implement a clause tracking mechanism. \textbf{I would prefer it if the text was clearer about the unfeasibility of a `subject of the same clause' feature.} But this does not impact their findings.''
\end{quote}

\subsubsection*{Response to R3.1}
\textcolor{red}{I'm not sure whether this comment requires action.}

\subsection*{Comment R3.2}
\begin{quote}
``Another concern that I raised involves the scalability of the notion that semantic interference effects reflect the use of semantic retrieval cues such as [+animate]. At issue is how this extends to finer-grained properties that are routinely responsible for plausibility violations. In my previous review I used the example of ``the teacher drove" vs. ``the little boy drove". It is, of course, clear that teachers are more plausible drivers than young children. This is part of our world knowledge. But it is less plausible that every time the noun `teacher' is encountered the property [+can drive] is activated. \textbf{The authors' response seems to be that animacy was sufficient for the materials in the current study, and that maybe other finer-grained features ``become activated only when relevant". To my mind, this point removes much of the value of the cue-based retrieval theory.} Nevertheless, this concern is about the generalizability of the authors' claims, and not about the soundness of their results.''
\end{quote}
\subsubsection*{Response to R3.2}
\textcolor{red}{I'm not sure whether this comment requires action.}

\subsection*{Comment R3.3}
\begin{quote}
``Finally, and in a similar vein, I raised questions about whether there might be other reasons for the selective interference effects found in this study. Maybe a different mechanism could capture the special status of the subject-of-the-same-clause, or maybe some specific property of the current experimental materials could be responsible for the lack of interference from other subjects. The authors seem skeptical of the first of these, and they seem more sympathetic to the second. \textbf{This is all reasonable enough, and their text conveys some caution. The abstract, on the other hand, is rather more confident.} I am sympathetic to the authors' preferred conclusion. But I am less confident than they are about how well justified it is based on the current findings. This has nothing to do with the quantitative wizardry that the authors display in the paper. It's all about the issue of generalizing from a single sentence configuration (that participants saw in 60\% of trials in the study).''
\end{quote}

\subsubsection*{Response to R3.3}
\textcolor{red}{Revise abstract to be more cautious.}

%\begin{quote}
%    \Paste{}
%\end{quote}

\end{document}

### 2024-11-01 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 \usepackage{hyperref}
+
+%% packages added by PS for the response to reviewers
+\usepackage{clipboard}
+\newclipboard{output-reviews}
+\openclipboard{output-reviews}
+%%
 
@@
 {
-
 It is well-established in sentence processing that completing linguistic dependencies requires memory retrieval \citep[see e.g., ][]{lewis06}.
@@
 \end{exe}
-
-There is clear evidence in the psycholinguistic literature \citep[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
-A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues \citep{Lewis2005, mcelree}.}\label{intro}
+}
+\Copy{intro}{There is clear evidence in the psycholinguistic literature \citep[e.g.,][]{jaeger_etal_2020, nicenboim, vandyke_mcelree06, vandyke_lewis03} that, at the verb \textit{was complaining}, it is necessary to retrieve the subject \textit{the resident} from memory in order to interpret who did what to whom.
+A widely-held assumption is that such memory retrievals during sentence comprehension are guided by retrieval cues \citep{Lewis2005, mcelree}.}
 
@@
 In the present case, it is possible that the syntactically relevant noun is only the one in the same clause as the verb.
-For more discussion of the role of structural retrieval cues, see \cite{franck2020hierarchical} and \cite{arnett2017subject}.} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
+{\Copy{cite_kush}{For more discussion of the role of structural retrieval cues, see \textcolor{blue}{\citet{kush2015relation}}, \citet{franck2020hierarchical} and \citet{arnett2017subject}.}}\label{cite_kush} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.

### 2024-11-04 code_data/SPR/r_scripts/Plot_ReadingTimes.R
Updated via the Open Science Framework

@@ -33,6 +33,9 @@ df_nofill <- droplevels(subset(df, label != "filler"))
 # trim all rts below 150 and above 3000 ms
 df_trim<- subset(df_nofill, df_nofill$rt > 150 & df_nofill$rt < 3000)
 
+# how much data was affected by the trimming?
+round(100 - (nrow(df_trim) / nrow(df_nofill)) *100,1)
+
 # subsets for hisyn and losyn (because they have different lengths)
 df_hisyn <- df_trim %>% filter(condition == "c" | condition == "d")
 df_losyn <- df_trim %>% filter(condition == "a" | condition == "b")

### 2024-11-04 code_data/previous_studies/PreviousFindings.Rmd
Updated via the Open Science Framework

@@
                   guides(shape = guide_legend(order = 2),col = guide_legend(order = 1))+
-                  coord_cartesian(ylim=c(-75, 150)) +
+                  coord_cartesian(ylim=c(-75, 250)) +
                   ggtitle("Reading time differences") 
@@
 
-load("BFs_schoknecht.Rda")
+load("../SPR/r_scripts/BFs_schoknecht.Rda")
 df.bf_schoknecht$Experiment <- "present"

### 2024-11-04 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 
-%\usepackage[backend=biber,useprefix=true,style=apa,url=false,sorting=nyt,eprint=false]{biblatex}
-
-%\addbibresource{interferencebib.bib}
-
-\usepackage[natbib=true,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
+%\usepackage[natbib=true,maxcitenames=2,bibstyle=authoryear, style=apa]{biblatex}
+\usepackage{natbib}
+\bibliography{interferencebib}
 \usepackage{lineno,xcolor,clipboard,graphicx}
@@
 \usepackage{csquotes}
-\DeclareLanguageMapping{american}{american-apa}
-\bibliography{interferencebib}
+%\DeclareLanguageMapping{american}{american-apa}
 
@@
 \subsubsection*{Response to E.1}
-\textcolor{red}{We have made the OSF repository public and updated the link in the manuscript.}
+We have made the OSF repository public and updated the link in the manuscript. This should grant general accessibility.
+
+\textcolor{red}{wait for ok from shravan}
 
@@
 \subsubsection*{Response to E.3}
-Yes, we changed all analyses in the 1st revision.
+Yes, we changed all analyses in the first revision.
 As we had explained in the response to reviewers on August 29, 2024 (specifically in the response letter and our responses to the comments E.2 and R3.5), the priors were changed from directional priors in the original version to symmetrical priors in the revised version.
@@
 \subsubsection*{Response to R1.1}
-\textcolor{red}{We have made the OSF repository public and updated the link in the manuscript.}
+We have made the OSF repository public and updated the link in the manuscript. This should grant general accessibility.
 
@@
 \subsubsection*{Response to R1.2}
-\textcolor{red}{We trimmed the reading times (between X and Y ms) before analysis and have now added this information to the manuscript.}
+We excluded reading times below 150 ms and above 3000 ms. This affected 4.9\,\% of the data. We have added this information to the manuscript on page \pageref{trim}:
 
-\textcolor{red}{Paste text}
+\begin{quote}
+    \Paste{trim}
+\end{quote}
 
@@
 \begin{quote}
-``p. 36: ``Regarding the possibility that sentence structure confounds caused the
-effects in the pre-critical region(.)"
+``p. 36: ``Regarding the possibility that sentence structure confounds caused the effects in the pre-critical region(.)"
 
@@
 \subsubsection*{Response to R1.3}
-\textcolor{red}{We have added the following text to the manuscript on page X:}
+We have added the following text to the manuscript on page \pageref{confound} (above the sentence which was quoted by the reviewer):
+
+\begin{quote}
+    \Paste{confound}
+\end{quote}
 
@@
 \subsubsection*{Response to R1.6}
-\textcolor{red}{Check how the figure changes}
+We have changed the y-axis limits to include the mean estimate of the respective data point (VD07.E2, regression path duration). For convenience, we paste the revised Figure here.
+
+\setcounter{figure}{13}
+\begin{sidewaysfigure}[htpb]
+    \caption{A) Reading time differences in the regions of interest from the present and previous studies using the design by \cite{vandyke07}.
+VD07.E1, VD07.E2 and VD07.E3 stand for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiments 1-3.
+The intervals for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) estimates are 95\% confidence intervals.
+M23.E and M23.G stand for \citeauthor{mertzen}'s (\citeyear{mertzen}) English and German experiment, respectively.
+The intervals for their study and the present study are Bayesian 95\% credible intervals.
+There are less estimates for the pre-critical region because \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1 and 2 did not have a pre-critical region.
+B) Bayes factors for the effects of interest in the present and previous studies under prior Normal(0,0.5).}
+    \label{fig:previous_vs_present}
+    \centering
+    \includegraphics[width=\linewidth]{present_previous_allplots.jpg}
+\end{sidewaysfigure}
+\clearpage
 
@@
 Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2.
-Since the first revision and in line with \textcite{jeffreys1998theory} and \textcite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence.
+Since the first revision and in line with \cite{jeffreys1998theory} and \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence.
 
@@
 \subsubsection*{Response to R3.3}
-\textcolor{red}{Revise abstract to be more cautious.}
+We have revised abstract to emphasize that our findings only apply to the present design. We paste the revised abstract here:
 
-%\begin{quote}
-%    \Paste{}
-%\end{quote}
+\begin{quote}
+    \Paste{abstract}
+\end{quote}
 

### 2024-11-04 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 \usepackage{hyperref}
+\usepackage{soul}
 
@@
 %% Text of abstract
+\Copy{abstract}{
 Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed to interpret the sentence.
@@
 We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07,mertzen}, to investigate interference due to syntactic and semantic cues in German.
+\textcolor{blue}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
 A Bayes factor analysis showed that both the self-paced reading times as well as event-related potentials provided clear evidence for a semantic interference effect.
-In the self-paced reading experiment, interference due to the semantic cues caused a reading time slow-down starting at the distractor.
+In the self-paced reading experiment, \textcolor{blue}{semantic interference due to the semantic cue \{+animate\}} caused a reading time slow-down starting at the distractor.
 The event-related potentials showed a more negative N400 amplitude at the critical verb, i.e., the retrieval site, for high vs.\ low semantic interference.
-Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to syntactic cues.
+Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to \textcolor{blue}{the syntactic cue \{+subject\}}.
 This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play an equal role in retrieval.
-We show through computational modeling that cue-based retrieval also shows no syntactic interference if the parser is assumed to keep track of which clause the subject occurs in.
+We show through computational modeling that cue-based retrieval also shows no syntactic interference \textcolor{blue}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in.
 Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model would be consistent with the observed patterns in our data.
 
-In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, in our data, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
+In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, in our data, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects \textcolor{blue}{in sentence configurations like the ones investigated in the present study}.
+}
 \end{abstract}
@@
 
-Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structures.
+\Copy{trim}{{\textcolor{blue}{Reading times below 150 ms and above 3000 ms were excluded before analysis (and visualization) because these overly short or long reaction times were likely caused by inattentive reading.
+The procedure excluded 4.9\,\% of the data.}}}\label{trim} Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structures.
 
@@
 They discussed several alternative explanations for their observed effects.
-In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions, and predictive processing effects.
-Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
+{\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{blue}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
 
@@
 In the present case, it is possible that the syntactically relevant noun is only the one in the same clause as the verb.
-{\Copy{cite_kush}{For more discussion of the role of structural retrieval cues, see \textcolor{blue}{\citet{kush2015relation}}, \citet{franck2020hierarchical} and \citet{arnett2017subject}.}}\label{cite_kush} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
+{\Copy{cite_kush}{For more discussion of the role of structural retrieval cues, see \textcolor{blue}{\cite{kush2015relation}}, \cite{franck2020hierarchical} and \cite{arnett2017subject}.}}\label{cite_kush} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.

### 2024-11-04 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 The focus of the present study was to investigate retrieval interference due to syntactic and semantic similarity.
-The finding that the reading times were likely affected by encoding interference and not by retrieval interference was unexpected (but see \citeauthor{mertzen}'s (\citeyear{mertzen}) discussion).
-Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that an in-depth discussion of encoding interference is beyond its scope.
+The finding that the reading times were primarily affected by encoding interference was unexpected (but see \citeauthor{mertzen}'s (\citeyear{mertzen}) discussion).
+Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that it is outside the scope of this manuscript to provide an in-depth discussion of a mechanism of encoding interference.
 To address the reviewer's concern that encoding interference would not cause such long-lasting effects as found in our SPR data, we have added the following text to the manuscript about a recent joint account of encoding and retrieval interference:

### 2024-11-04 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
-This pattern persisted in the following regions.
-
-Because the distractor and the immediately following regions differ between conditions, we did not compute Bayes factors for the observed reading time differences.
-Instead we focused our Bayes factors analyses on regions which were identical across conditions.
-These are the pre-critical, critical and spill-over regions.
-
-Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but will be harder to interpret because of possible spill-over from the syntactic manipulation.
-The pre-pre-critical and pre-critical region showed longer reading times for high vs.\ low semantic interference conditions.
+This pattern persisted in the following regions \textcolor{blue}{-- almost until the end of the sentence.} The distractor and the immediately following regions differ between conditions, \textcolor{blue}{therefore we did not carry out any statistical tests for these observed reading time differences.
+To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference.
+This sustained difference in reading times is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region, because any effects that might be present there can not be attributed clearly to the processing of that specific region.
+However, for comparability with previous work, we briefly report the reading times of the later regions which were identical across conditions, i.e., pre-critical, critical and spill-over region, below.}
+
+Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but \textcolor{blue}{it was only used in the present study.} The pre-pre-critical and pre-critical region showed longer reading times for high vs.\ low semantic interference conditions.
 The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low.
@@
 
-Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b.
+\textcolor{blue}{Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b, respectively.
+Across the sub-experiments, there is some variation in the reading time patterns, but all of them show a pattern consistent with semantic interference in at least some of the regions.
+Experiment 1a showed a remarkable difference between the experimental sessions, with overall shorter reading times in the second session.
+This attenuation in reading times was likely due to adaptation to the SPR task.}
+
+\textcolor{blue}{\st{Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b.
 Panel D shows the results from Experiment 1a.
@@
 As mentioned earlier, this attenuation in reading times and differences between conditions were likely due to adaptation to the SPR task in the second session.
-
-Panel E shows that Experiment 1b with items 1-60 (N = 350, left sub-figure) showed a pattern consistent with semantic interference in all regions.
+}}
+
+\textcolor{blue}{\st{Panel E shows that Experiment 1b with items 1-60 (N = 350, left sub-figure) showed a pattern consistent with semantic interference in all regions.
 Experiment 1b with items 61-120 (N = 220, right sub-figure) showed a pattern consistent with semantic interference in the regions surrounding the critical verb, but not at the verb itself.
@@
 The other two conditions (high syntactic/low semantic and low syntactic/high semantic) led to the fastest verb reading times in Experiment 1b with items 61-120.
-
-Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
-{The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions.
+}}
+
+\textcolor{blue}{\st{Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure} \ref{fig:spr_posteriors} \st{shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
+The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions.
 In the spill-over region, it included only positive values ([1, 15] ms).
@@
 The posteriors for the interaction were mostly positive in all regions.
-The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.
+The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}}
+
+\textcolor{blue}{\st{In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms).
+The posteriors for the interaction were mostly positive in all regions.
+The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}}
+
+Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions.
+The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions.
+The estimates of syntactic interference were less consistent across regions.
+They were almost centered around zero in the critical region, showed mostly positive values in the pre-critical region and fully positive values in the spill-over region.}
 
@@
 \end{figure}
-}\label{SPR_results}
-
-{The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last trial (100th), reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
+
+The fixed effect of trial id which was included in the models as a co-variate revealed a huge adaptation effect: From the first to the last trial (100th), reading times in the regions of interest decreased by on average around 250\,ms (pre-critical region: -219\,ms [-233, -204], critical region: -296\,ms [-317, -276], spill-over region: -246\,ms [-260, -233]).
 This adaptation effect was probably caused by increasing familiarity with the task and employed sentence structures over the time course of the experiment.
-While this adaption is interesting in itself, it was not the focus of the present study, therefore we will not discuss it further.}\label{adaptation}
+While this adaption is interesting in itself, it was not the focus of the present study, therefore we will not discuss it further.
 
 \paragraph{Hypothesis testing using Bayes factors}
-
-Was there any evidence for the effects of interest estimated above?
-The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}.
-The analyses used all the data from Experiments 1a and 1b.
-{We found very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
-In the critical region, the evidence for semantic interference was weak (Normal(0, 0.01): BF$_{10}$ = 3.5,
+\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the sustained reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
+There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
+In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5,
 Normal(0, 0.05): BF$_{10}$ = 1.3,
@@
 {There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
-Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).}\label{insig_interaction}
+Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).
 
 All in all, the Bayes factors provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but at best weak evidence in the critical region.
-In contrast, the Bayes factors  across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.}
+In contrast, the Bayes factors  across all regions provided at best weak evidence for syntactic interference or even evidence against it and evidence against the interaction.
 
@@
 We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
+\textcolor{blue}{ While Bayes factors provided mostly evidence for semantic interference and evidence against syntactic interference and the interaction at the critical and surrounding regions, reading times started to differ earlier in the sentence.
+Starting at the distractor, high semantic interference conditions (animate distractor) were read slower than low semantic interference conditions (inanimate distractor).
+Because the distractors differed between conditions, we did not analyze the reading time difference in that region.
+}
+
+\textcolor{blue}{\st{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
 This means that semantic interference occurred even before the retrieval event could begin.
@@
 In the spill-over region, the syntactic manipulation slowed down the reading times, but to a lesser extent than the semantic manipulation (CrI of syntactic interference vs.
-CrI of semantic interference: [1,15] ms vs.\ [7,21] ms).
+CrI of semantic interference: [1,15] ms vs.
+[7,21] ms).
 The Bayes factors also provided less support for the syntactic effect than for the semantic effect (BFs under prior Normal(0, 0.01) in the spill-over region for the syntactic effect vs.
-the semantic effect: 2.8 vs.\ 246).
+the semantic effect: 2.8 vs.
+246).
 The relatively small extent of the syntactic effect which only emerged in one region and which was also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference.
-However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.}\label{syn_interf} We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
+However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.
+We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
 For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) for a very small effect ($\pm$8 ms).
 There was evidence against the interaction under wider priors and in the surrounding regions.
-{Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
-Therefore, we refrain from further discussing effects occurring in the later regions.}\label{only_precritical} The effects in the pre-critical region are discussed next.
+Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
+Therefore, we refrain from further discussing effects occurring in the later regions.
+The effects in the pre-critical region are discussed next.}}
 

### 2024-11-08 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 %add description of revisions here
+%To address reviewer \# 2 concern that the detailed description of the SPR results in the critical and surrounding regions might be misleading the reader, we have removed the panels of Figure 3 that highlighted the reading times of these regions.
+Instead, we now show the long-lasting reading time difference starting at the distractor in each of the self-paced reading sub-experiments.
 
@@
 \subsubsection*{Response to R1.7}
-\textcolor{red}{Follow the suggestion.}
+We have added the reviewer's suggested explanation that the employed methods could have especially led to good-enough processing and that this might have contributed to the different findings between studies (see page \pageref{SPR_goodenough}):
+
+\begin{quote}
+    \Paste{SPR_goodenough}
+\end{quote}
 
@@
 \subsubsection*{Response to R2.1}
-We agree with the review that the discussion of the SPR results was misleading and have revised it accordingly.
-We now start the discussion by discussing the effects at the distractor (see Response to R2.2) and then briefly discuss the later effects.
+We agree with the reviewer that the presentation and discussion of the SPR results was misleading and have revised it accordingly.
+We have revised Figure 3 to remove the focus on the critical and surrounding regions (see below).
+At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference and that the long-lasting effect caused by that confounds the results at later effects.
+For comparability with previous findings, we then briefly report the later effects.
+We have revised the Results section in the following way, staring on page \pageref{revised_SPR}:
 
-\textcolor{red}{Revise text and paste here}
+\begin{quote}
+    \Paste{revised_SPR}
+\end{quote}
+
+\setcounter{figure}{2}
+
+\begin{sidewaysfigure}[h]
+    \caption{Self-paced reading times with 95\% confidence intervals.
+Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure.
+\textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiments 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
+    \label{fig:whole_sentence}
+    \centering
+    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
+\end{sidewaysfigure}
+\clearpage
+
+We do not share the reviewer's interpretation that the effects in the later regions are meaningless.
+We use the pre-critical reading times as a proxy to investigate the effect of encoding interference caused by the distractors.
+We emphasize the link between the distractor and the pre-critical region by discussing them together in a new section with the title ``Reading times difference prior to retrieval" (starting on page \pageref{prior_to_retrieval}).
+This new section replaces the former section ``Pre-critical effects".
+
+\begin{quote}
+    \Paste{SPR_discussion_revised}
+\end{quote}
+
+\textcolor{red}{Figure out how much of the revised text to paste here}
 
@@
 
-\textcolor{red}{Write someting about Himanshu's work in relation to this}
+\textcolor{red}{Write something about Himanshu's work in relation to this}
 
@@
 Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2.
-Since the first revision and in line with \cite{jeffreys1998theory} and \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence.
+Since the first revision and in line with \cite{jeffreys1998theory} and \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence either way.
 

### 2024-11-08 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 
-As discussed in detail below, the second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
+The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
 The adaptation effect is theoretically interesting per se \citep{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
 For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both).
-In the Results section, we present both sets of results as well as the pooled data (N=774) from E1a and E1b.
+In the Results section, we present the pooled data (N=774) from E1a and E1b.
 As mentioned above, because the ERP method requires a large number of items, all the ERP data come from participants who completed both sessions.
@@
 \section{Data availability statement}
-{All materials, data and analysis scripts can be accessed via \\ https://osf.io/4fpru/?view$\_$only=2d2a004d3e4147bc975b5e8c455c27df.}
+All materials, data and analysis scripts can be accessed via \\ \href{https://osf.io/4fpru/?view\_only=2d2a004d3e4147bc975b5e8c455c27df}{https://osf.io/4fpru/?view\_only=2d2a004d3e4147bc975b5e8c455c27df}
 
@@
 
-\Copy{trim}{{\textcolor{blue}{Reading times below 150 ms and above 3000 ms were excluded before analysis (and visualization) because these overly short or long reaction times were likely caused by inattentive reading.
-The procedure excluded 4.9\,\% of the data.}}}\label{trim} Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence} A and B, separately for high and low syntactic interference because these conditions had partially different sentence structures.
-
-{
+\Copy{trim}{{\textcolor{blue}{Reading times below 150 ms and above 3000 ms were excluded before analysis (and visualization) because these overly short or long reaction times were likely caused by inattentively performing the task.
+The procedure excluded 4.9\,\% of the data.}}}\label{trim} Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panels A and C)} and low \textcolor{blue}{(panels B and D)} syntactic interference because these conditions had partially different sentence structures.
+
 \begin{sidewaysfigure}[h]
@@
 Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure.
-Panel C shows the pooled reading times of all conditions in the critical and surrounding regions.
-Panels D and E show the reading times of all conditions in the critical and surrounding regions in the different parts of the self-paced reading experiment.}
+\textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiments 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
     \label{fig:whole_sentence}
     \centering
-    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom_exp.jpg}
+    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
 \end{sidewaysfigure}
-\clearpage}
-
-It is apparent from Figure \ref{fig:whole_sentence} A and B that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
-This pattern persisted in the following regions \textcolor{blue}{-- almost until the end of the sentence.} The distractor and the immediately following regions differ between conditions, \textcolor{blue}{therefore we did not carry out any statistical tests for these observed reading time differences.
-To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference.
-This sustained difference in reading times is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region, because any effects that might be present there can not be attributed clearly to the processing of that specific region.
-However, for comparability with previous work, we briefly report the reading times of the later regions which were identical across conditions, i.e., pre-critical, critical and spill-over region, below.}
-
-Figure \ref{fig:whole_sentence} C shows the reading times of these regions; here, we also show the pre-pre-critical region, which was identical across conditions as well, but \textcolor{blue}{it was only used in the present study.} The pre-pre-critical and pre-critical region showed longer reading times for high vs.\ low semantic interference conditions.
-The critical verb was read slowest when syntactic and semantic interference were both high and fastest when syntactic interference was high and semantic interference low.
-The other two conditions showed intermediate reading times.
-The spill-over region again primarily showed a difference between high and low semantic interference conditions.
-
-\textcolor{blue}{Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b, respectively.
-Across the sub-experiments, there is some variation in the reading time patterns, but all of them show a pattern consistent with semantic interference in at least some of the regions.
-Experiment 1a showed a remarkable difference between the experimental sessions, with overall shorter reading times in the second session.
-This attenuation in reading times was likely due to adaptation to the SPR task.}
-
-\textcolor{blue}{\st{Panels D and E show the reading times of the critical and surrounding regions for each of the two Experiments 1a and 1b.
-Panel D shows the results from Experiment 1a.
-The first session of Experiment 1a (items 1-60, N = 204, left sub-figure) showed results similar to the pooled results presented in Panel C.
-In the second session of Experiment 1a (items 61-120, N = 160, right sub-figure) a semantic interference effect was seen at the critical verb and the patterns in the surrounding regions were less clear than in the first session.
-Additionally, the second session in Experiment 1b showed shorter reading times compared to the first session.
-As mentioned earlier, this attenuation in reading times and differences between conditions were likely due to adaptation to the SPR task in the second session.
-}}
-
-\textcolor{blue}{\st{Panel E shows that Experiment 1b with items 1-60 (N = 350, left sub-figure) showed a pattern consistent with semantic interference in all regions.
-Experiment 1b with items 61-120 (N = 220, right sub-figure) showed a pattern consistent with semantic interference in the regions surrounding the critical verb, but not at the verb itself.
-At the verb, reading times were longest when both syntactic and semantic interference were high.
-Reading times were a bit faster when both syntactic and semantic interference were low.
-The other two conditions (high syntactic/low semantic and low syntactic/high semantic) led to the fastest verb reading times in Experiment 1b with items 61-120.
-}}
-
-\textcolor{blue}{\st{Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure} \ref{fig:spr_posteriors} \st{shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters.
-The 95\% credible intervals for syntactic interference were almost centered around zero in the pre-critical and critical regions.
-In the spill-over region, it included only positive values ([1, 15] ms).
-In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms).
-The posteriors for the interaction were mostly positive in all regions.
-The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}}
-
-\textcolor{blue}{\st{In contrast, for semantic interference, the 95\% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms) and smaller estimates in the critical region ([1, 19] ms).
-The posteriors for the interaction were mostly positive in all regions.
-The 95\% credible interval of the interaction was the widest in the critical region ([-3, 31 ms]) with a mean estimate of 14 ms.}}
-
-Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions.
+\clearpage
+
+\label{revised_SPR}\Copy{revised_SPR}{
+It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
+This pattern \textcolor{blue}{was found in the pooled data (panels A and B) as well as in the data of each experimental session (panels C and D).
+To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}.
+Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session.
+As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.
+
+\textcolor{blue}{Crucially, the reading times difference starting at the distractor} persisted in the following regions.
+The distractor and the immediately following regions differ between conditions, \textcolor{blue}{therefore we did not carry out any statistical tests for these observed reading time differences.
+The reading time difference starting at the distractor and persisting almost until the end of the sentence is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region, because any effects that might be present there cannot be attributed clearly to the processing of that specific region.
+However, for comparability with previous work, we briefly report analyses of the reading times of the later regions, i.e., pre-critical, critical and spill-over region, below.}
+
+Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions.
 The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions.
@@
 They were almost centered around zero in the critical region, showed mostly positive values in the pre-critical region and fully positive values in the spill-over region.}
+}
 
@@
 \paragraph{Hypothesis testing using Bayes factors}
-\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the sustained reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
+\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
 There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
@@
 \subsection{Discussion}
-We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation.
-\textcolor{blue}{ While Bayes factors provided mostly evidence for semantic interference and evidence against syntactic interference and the interaction at the critical and surrounding regions, reading times started to differ earlier in the sentence.
-Starting at the distractor, high semantic interference conditions (animate distractor) were read slower than low semantic interference conditions (inanimate distractor).
-Because the distractors differed between conditions, we did not analyze the reading time difference in that region.
-}
-
-\textcolor{blue}{\st{The reading time results at the critical verb showed anecdotal evidence against or moderate evidence for semantic interference (BF$_{10}$ between 0.7 and 3.5 depending on a priori assumed effect size), but an effect of semantic interference was present at the regions surrounding the critical verb, i.e., the pre-critical and spill-over region.
-This means that semantic interference occurred even before the retrieval event could begin.
-Reading times of the pre-critical word and critical verb showed evidence against syntactic interference.
-In the spill-over region, the syntactic manipulation slowed down the reading times, but to a lesser extent than the semantic manipulation (CrI of syntactic interference vs.
-CrI of semantic interference: [1,15] ms vs.
-[7,21] ms).
-The Bayes factors also provided less support for the syntactic effect than for the semantic effect (BFs under prior Normal(0, 0.01) in the spill-over region for the syntactic effect vs.
-the semantic effect: 2.8 vs.
-246).
-The relatively small extent of the syntactic effect which only emerged in one region and which was also only weakly supported by the Bayes factors leads us to the conclusion that our SPR data predominantly showed semantic interference.
-However, there was some anecdotal evidence for a small syntactic interference effect ($\pm$8 ms) in the spill-over region.
-We will discuss the absence of more decisive evidence for syntactic interference in our data in the general discussion.
-For the interaction of syntactic and semantic interference, reading times at the critical verb showed inconclusive results (neither support for nor against) for a very small effect ($\pm$8 ms).
-There was evidence against the interaction under wider priors and in the surrounding regions.
-Given the pre-critical reading time differences, effects in the later regions, i.e., the critical and spill-over region, cannot be attributed clearly to the stimuli in these regions and sentence processing mechanisms associated with them.
-Therefore, we refrain from further discussing effects occurring in the later regions.
-The effects in the pre-critical region are discussed next.}}
-
-\subsubsection{The effects in the pre-critical region}
-
-At the distractor (which occurred well before the pre-critical region), high semantic interference (animate distractors) led to longer reading times than low semantic interference (inanimate distractors, see Figure \ref{fig:whole_sentence} A and B).
-One might argue that this pattern could be due to differences in word length.
+\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation.
+\textcolor{blue}{However, reading times started to differ before this critical dependency could be formed.
+Starting at the distractor which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors).
+Because the distractor differed between conditions, we did not analyze distractor reading times.
+Instead, we treat the pre-critical region which was identical across conditions and preceded the critical verb as a proxy to investigate the difference starting at the distractor.
+Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region.
+We refrain from interpreting effects which were observed in later regions, i.e., the critical verb and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes.
+The reading times differences at the distractor and the pre-critical region are discussed below.}}
+
+\subsubsection{\textcolor{blue}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}
+
+\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence} A and B).
+One might argue that this could be due to differences in word length.
 The animate distractors in our materials were on average 1.1 letters longer than the inanimate distractors (mean length of animate distractors: 9.3, sd: 2.6, mean length of inanimate distractors: 8.2, sd: 2.9).
@@
 
-\label{plausib_anim_inan}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions.
+Another possible explanation for the reading time differences starting at the distractor could be \textcolor{blue}{that the distractors induced} a difference in plausibility between the high and low semantic interference conditions.
 To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}.
@@
 
-To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
+To investigate whether this plausibility difference is driving the \textcolor{blue}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
 This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings.
@@
 The interaction of semantic interference and plausibility was centered around zero.
-In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference was partially caused by reduced plausibility of the high semantic interference conditions.
-However, the plausibility difference was not the only driver of the reading times slowdown, i.e., semantic interference led to slower reading times independent of plausibility.
+In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference \textcolor{blue}{starting at the distractor and lasting into later regions} was partially caused by reduced plausibility of the high semantic interference conditions.
+However, the plausibility difference was not the only driver of the reading times slowdown, i.e., \textcolor{blue}{high} semantic interference led to slower reading times independent of plausibility.
 
@@
 \end{figure}
-}
-
-At the pre-critical region, there was evidence for semantic interference.
-As this cannot be due to retrieval interference, the most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above).
+
+\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.
+T}he most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above).
 The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
@@
 \cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region.
-They discussed several alternative explanations for their observed effects.
 {\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{blue}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
@@
 Regarding the predictive-processing explanation, \citet{mertzen} argued that the pre-critical adverb must attach to the upcoming verbal phrase; this leads to an anticipatory creation of a verb phrase chunk in memory, which triggers a retrieval of the subject already at the pre-critical region.
-However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence} A and B).
+However, in our study, the effect started at the distractor and became smaller as the two pre-critical adverbs were read (see Figure \ref{fig:whole_sentence}).
 The predictive-processing explanation would incorrectly predict that the effect begins at the first adverb, which is the first word of the verbal phrase.
 Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data.
-Crucially, this line of reasoning should not be limited to the explanation of the pre-critical effects, but should be extended to the effects in later regions as well.
-The reading times for high vs.\ low semantic interference conditions started to differ at the distractor and differed throughout the rest of the sentence.
-It is not possible to pinpoint whether the effects in later regions were due to a retrieval initiated at the respective region, or still due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.
-Because of the unclear status of the effects at the critical verb and spill-over region, we refrain from interpreting the semantic interference effects which were observed in these regions.
+Crucially and as mentioned above, \textcolor{blue}{due to the long-lasting reading time difference starting at the distractor} it is not possible to pinpoint \textcolor{blue}{the driver of effects observed in later regions.
+The effects in the critical region (and spill-over region) might be} due to a retrieval initiated at the respective region, or still \textcolor{blue}{be} due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.
 
@@
 However, word-by-word self-paced reading (SPR) and ERPs elicited by single words presented in the typical rapid serial presentation (RSVP) format, if anything might have a greater tendency to be affected by interference compared to natural reading (as in eye-tracking studies), due to the restricted reading format: Comprehenders have no opportunity to revisit previous material during conventional self-paced reading \citep[but see][]{BSPR} and RSVP.
-Word-by-word presentation is likely to be more demanding on the comprehender's memory, which should make it easier to detect interference effects.
-So, the differences in methodology do not offer a straightforward explanation for the lack of syntactic interference in our data compared to the previous studies.
+Word-by-word presentation is likely more demanding on the comprehender's memory, which should make it easier to detect interference effects.
+So, the differences in methodology do not offer a straightforward
+explanation for the lack of syntactic interference in our data compared to the previous studies.
+\Copy{SPR_goodenough}{\textcolor{blue}{However, due to the demanding task comprehenders might trade off processing depth and adapt good-enough processing \citep{ferreira2007goodenough}.
+Consequently, they would not fully parse embedded structures, hence there would be no syntactic interference from embedded subjects.
+Under this account, semantic interference might still arise because animacy information is readily available from the lexical entries with no need for detailed processing.}}\label{SPR_goodenough}
 
@@
 {While it seems unfortunate that the present and previous results together do not create a clear and unambiguous picture of the investigated phenomena, this divergence of results is not an anomaly within the scientific literature.
-In psycholinguistics alone, contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{taboretal04, lcpaape2023}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
+In psycholinguistics alone, contradictory results have been reported for many phenomena: e.g., predictive parsing effects \citep{nieuwland_etal_2018, Delong_etal_2005}, reflexives processing \citep{jaeger_etal_2020, dillon2013}, main verb/reduced relative garden paths \citep{trueswell1994semantic, ferreira1986independence}, local coherence \citep{taboretal04, lcpaape2025}, locality effects in German \citep{levy&keller_2013, vasishth2018_signficancefilter}, interference from extra-sentential distractors \citep{vandyke_mcelree06, mertzen_repVDM2006}, attenuation of agreement attraction due to case marking \citep{hartsuiker2001_agree, Avetisyan_etal_2020_agree}, agreement attraction in grammatical sentences \citep{wagers2009agrttr, nicenboim} and relative clauses in Chinese \citep{hsiao03, gibsonwu, VasishthetalPLoSOne2013, jaeger2015subject}.
 The only way to resolve contradictions in the literature -- like the contradictory results between the present and previous studies regarding syntactic interference -- is by carrying out adequately powered studies in the future to reach consensus over time.}\label{not_an_anomaly}
@@
 Regarding grammaticality illusions in double-center embeddings, the grammaticality illusion does not occur in German and Dutch \citep{VSLK11,FrankEtAl2015}.
-Regarding syntactic local coherence effects, a large-sample study \citep{lcpaape2023} presents evidence against syntactic local coherence.
+Regarding syntactic local coherence effects, a large-sample study \citep{lcpaape2025} presents evidence against syntactic local coherence.
 Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of number agreement attraction \citep{chromy2023number}.

### 2024-11-11 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 \subsubsection*{Response to R1.1}
-We have made the OSF repository public and updated the link in the manuscript. This should grant general accessibility.
+A formatting error broke the link in the previous version of the manuscript. We have made the OSF repository public and updated the link in the manuscript. This should grant general accessibility.
 
@@
 For comparability with previous findings, we then briefly report the later effects.
-We have revised the Results section in the following way, staring on page \pageref{revised_SPR}:
+We have revised the Results section in the following way, starting on page \pageref{revised_SPR}:
 
@@
     \Paste{revised_SPR}
+\end{quote}
+
+We do not share the reviewer's interpretation that the effects in the later regions are not meaningful.
+We use the pre-critical reading times as a proxy to investigate the effect of encoding interference caused by the distractor.
+To emphasize this we have renamed the former section ``Pre-critical effects" to ``Reading times difference prior to retrieval".
+We have revised the beginning of the Discussion on page \pageref{prior_to_retrieval} accordingly:
+
+\begin{quote}
+    \Paste{SPR_discussion_revised}
 \end{quote}
@@
 
-We do not share the reviewer's interpretation that the effects in the later regions are meaningless.
-We use the pre-critical reading times as a proxy to investigate the effect of encoding interference caused by the distractors.
-We emphasize the link between the distractor and the pre-critical region by discussing them together in a new section with the title ``Reading times difference prior to retrieval" (starting on page \pageref{prior_to_retrieval}).
-This new section replaces the former section ``Pre-critical effects".
-
-\begin{quote}
-    \Paste{SPR_discussion_revised}
-\end{quote}
-
-\textcolor{red}{Figure out how much of the revised text to paste here}
-
 \subsection*{Comment R2.2}
@@
 \end{quote}
+
 \subsubsection*{Response to R2.2}
@@
 \subsubsection*{Response to R2.3}
-We agree with the reviewer that our present work does not rule out syntactic interference in general, but that the present data just suggests that the specific design used in this work did not induce considerable syntactic interference.
+We agree with the reviewer that our present work does not rule out syntactic interference in general, but the present study using the present design did not induce considerable syntactic interference.
 We have made the following edits in the manuscript to emphasize this point:
@@
 The authors interpret this as a facilitatory interference effect (i.e., easier processing with high semantic interference).
-However, if this were the case, we would also expect faster reaction times (RTs) in high semantic interference conditions, while the opposite is observed in the SPR data.
+However, if this were the case, \textbf{we would also expect faster reaction times (RTs) in high semantic interference conditions, while the opposite is observed in the SPR data}.''
+\end{quote}
 
-4.
+\subsubsection*{Response to R2.4}
+ \textcolor{red}{write response: no facilitation in SPR data because 1.
+distractor effect overshadows everything 2.
+mean effect in RTs cannot show inhibition and facilitation at once, inhibition was stronger}
+
+\subsection*{Comment R2.5}
+\begin{quote}
+``4.
 Conflicting directions in ERP components: Within the ERP results, the different components also point in opposite directions.
@@
 \end{quote}
-\subsubsection*{Response to R2.4}
-\textcolor{red}{Not sure whether this comment requires action.}
 
-\subsection*{Comment R2.5}
+\subsubsection*{Response to R2.5}
+
+  \textcolor{red}{write response: Fleshed out more how inhibition and facilitation could both occur and manifest in ERPs}
+
+\begin{quote}
+    \Paste{inhibition_facilitation}
+\end{quote}
+
+\subsection*{Comment R2.6}
 \begin{quote}
@@
 \end{quote}
-\subsubsection*{Response to R2.5}
+\subsubsection*{Response to R2.6}
 \textcolor{red}{Not sure whether this comment requires action. Maybe 1-2 sentences about Himanshu's work.}
 
-\subsection*{Comment R2.6}
+\subsection*{Comment R2.7}
 \begin{quote}
@@
 \end{quote}
-\subsubsection*{Response to R2.6}
+\subsubsection*{Response to R2.7}
 Yes, the priors were changed in the 1st revision (directional priors in the original version and symmetrical priors in the revised version, see also our Response to E.3).
@@
 
-\subsection*{Comment R2.7}
+\subsection*{Comment R2.8}
 \begin{quote}
@@
 \end{quote}
-\subsubsection*{Response to R2.7}
+\subsubsection*{Response to R2.8}
 \textcolor{red}{SV email: the syn / sem comment is also interesting because there are some theories that only assume semantic cue association--there isn't even any syntax.
@@
 
-\subsection*{Comment R2.8}
+\subsection*{Comment R2.9}
 \begin{quote}
@@
 \end{quote}
-\subsubsection*{Response to R2.8}
-\textcolor{red}{Describe why fillers should be comparable in complexity to the experimental items and how they distracted the participants from the purpose of the experiment}
+\subsubsection*{Response to R2.9}
+The fillers included different types of subordinate clauses and (more) modifiers at different positions within the sentence.
+By deviating from the rigid structure of the critical items, they provided more variety in regard to the employed sentence structures.
+To avoid creating an obvious juxtaposition of very complex sentences (critical items) and very simple sentences (fillers), the fillers included long-distance dependencies and at least one subordinate clause.
+We have revised the description of the fillers in the manuscript on page \pageref{fillers} in the following way:
 
-\subsection*{Comment R2.9}
+\begin{quote}
+    \Paste{fillers}
+\end{quote}
+
+\subsection*{Comment R2.10}
 \begin{quote}
@@
 
-\subsubsection*{Response to R2.9}
+\subsubsection*{Response to R2.10}
 \textcolor{red}{remove repetitions}
@@
 \subsubsection*{Response to R3.3}
-We have revised abstract to emphasize that our findings only apply to the present design. We paste the revised abstract here:
+We have revised the abstract to emphasize that our findings only apply to the present design. We paste the revised abstract here:
 

### 2024-11-11 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 The event-related potentials showed a more negative N400 amplitude at the critical verb, i.e., the retrieval site, for high vs.\ low semantic interference.
-Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to \textcolor{blue}{the syntactic cue \{+subject\}}.
+Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to \textcolor{blue}{the syntactic cue \{+ grammatical subject\} in this design}.
 This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play an equal role in retrieval.
@@
 Second, although it has been established \citep[e.g.,][]{vandyke07,mertzen} that syntactic and semantic cues are involved in sentence processing, it is not obvious which cues are relevant for a particular retrieval event.
-For example, \cite{dillon2013} argued, following \cite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be built.
+For example, \cite{dillon2013} argued, following \cite{Sturt2003}, that the gender cue was either downweighted or not used at all when an antecedent-reflexive dependency needs to be build.
 Instead, the antecedent of a reflexive could be identified using the syntactic principle A of the binding theory, which correctly identifies the target for retrieval.
@@
 The length of the sentences required line breaks.
-These were hard-coded so that in each sentence a line break appeared i) between the subject and the relative clause and ii) two words after the critical verb (see \ref{ex:linebreak}, critical word in bold, | indicate line breaks in the moving window display).
+These were hard-coded so that in each sentence a line break appeared i) between the subject and the relative clause and ii) two words after the critical verb (see \ref{ex:linebreak}, critical word in bold, $\mid$ indicate line breaks in the moving window display).
 In a third of the trials, the sentence was followed by a yes/no comprehension question.
@@
 \ex \label{ex:linebreak}
-    \gll Die Nachbarin glaubte, dass der Witwer, | der erzählt hatte, dass der Verlust schrecklich war, regelmäßig abends \textbf{trank}, um zu | vergessen. \\ 
-    The neighbor believed that the widower | who told had that the loss awful was regularly in.the.evening drank in.order to | forget\\
+    \gll Die Nachbarin glaubte, dass der Witwer, $\mid$ der erzählt hatte, dass der Verlust schrecklich war, regelmäßig abends \textbf{trank}, um zu $\mid$  vergessen. \\ 
+    The neighbor believed that the widower $\mid$  who told had that the loss awful was regularly in.the.evening drank in.order to $\mid$  forget\\
     \trans  `The neighbor believed that the widower, who had told her that the loss was awful, regularly drank in the evenings to forget.'\\  
@@
 
-{In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
+In Experiment 1a, participants were re-invited for session 2 after they completed session 1.
 The experimental sessions were separated by 1 - 20 days.
@@
 In session 2, the experimental items 61-120 were presented with another set of 40 filler sentences.
-The fillers were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}\label{fillers1} Each session lasted approximately 25 minutes.
+{\Copy{fillers}{\textcolor{blue}{The fillers were included to mask the critical manipulation.
+They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{blue}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.
 
@@
 
-Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}.
+\Copy{trim}{{\textcolor{blue}{Self-paced reading times below 150 ms and above 3000 ms were excluded before analysis because these overly short or long reaction times were likely caused by inattentively performing the task.
+The procedure excluded 4.9\,\% of the data.}}}\label{trim} Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}.
 The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site.
@@
 
-\Copy{trim}{{\textcolor{blue}{Reading times below 150 ms and above 3000 ms were excluded before analysis (and visualization) because these overly short or long reaction times were likely caused by inattentively performing the task.
-The procedure excluded 4.9\,\% of the data.}}}\label{trim} Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panels A and C)} and low \textcolor{blue}{(panels B and D)} syntactic interference because these conditions had partially different sentence structures.
+Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panels A and C)} and low \textcolor{blue}{(panels B and D)} syntactic interference because these conditions had partially different sentence structures.
 
@@
 Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure.
-\textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiments 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
+\textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiment 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
     \label{fig:whole_sentence}
@@
 \textcolor{blue}{Crucially, the reading times difference starting at the distractor} persisted in the following regions.
-The distractor and the immediately following regions differ between conditions, \textcolor{blue}{therefore we did not carry out any statistical tests for these observed reading time differences.
-The reading time difference starting at the distractor and persisting almost until the end of the sentence is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region, because any effects that might be present there cannot be attributed clearly to the processing of that specific region.
-However, for comparability with previous work, we briefly report analyses of the reading times of the later regions, i.e., pre-critical, critical and spill-over region, below.}
+The distractor and the immediately following regions differed between conditions, \textcolor{blue}{therefore we did not analyze their reading times.
+The reading times difference starting at the distractor and persisting almost until the end of the sentence is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region.
+Any effects that might be present there cannot be attributed clearly to the processing of the respective region.
+However, for comparability with previous work, we briefly report analyses of the reading times of the later regions, i.e., the pre-critical, critical and spill-over region.}
 
@@
 Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region.
+In contrast, there was evidence against syntactic interference and an interaction in that region.
 We refrain from interpreting effects which were observed in later regions, i.e., the critical verb and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes.
@@
 
-To investigate whether this plausibility difference is driving the \textcolor{blue}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
+To investigate whether this plausibility difference was driving the \textcolor{blue}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
 This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings.
@@
 
-Indeed, previous work using the same design as in the present paper has also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
+Indeed, previous work using the same design as the present paper had also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
 \citeauthor{mertzen}'s (\citeyear{mertzen}) Figures 5 and 6 indicated reading time differences in earlier regions, especially for their German data (see their Figure 6).
-\cite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \cite{mertzen} also points out, encoding interference could be an explanation even in that study.
+\cite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \cite{mertzen} also pointed out, encoding interference could be an explanation even in that study.
 Given these earlier findings, our results are consistent with the encoding interference explanation.
@@
 This finding is in line with previous ERP studies on interference \citep{lee_garnsey, vasishth_drenhaus_2011, martinetal2014, schoknecht2022}.
-The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference led to a reduced P600 amplitude compared to low semantic interference.
-Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}.
+\Copy{inhibition_facilitation}{{The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference led to a reduced P600 amplitude compared to low semantic interference.
+\textcolor{blue}{We provide a speculative interpretation of this unexpected result.} Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}.
 The reduced P600 amplitude under high semantic interference in the present study might be explained by facilitatory interference \citep[see e.g.,][]{jaeger_etal_2017}.
-In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to function as the subject.
-A similar effect was found by \cite{Tanner_etal_2017}.
+\textcolor{blue}{We are not aware of any other study that found inhibitory as well facilitatory interference in the same sentence configuration.
+Facilitatory interference is typically only found in ungrammatical sentences.
+Nevertheless, we think that the reduced P600 amplitude for high vs.\ low semantic interference in our data could be explained by facilitatory interference.} In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to function as the subject.
+\textcolor{blue}{Even if the wrong noun, i.e., the distractor, was retrieved, it could be easily integrated with the verb.
+In contrast, misretrievals of the distractor in the low semantic interference conditions which might happen due to noisy processing would require reanalysis because the distractor did not match the semantic requirements of the verb.} A similar effect was found by \cite{Tanner_etal_2017}.
 In their study, the P600 elicited by ungrammatical verbs which were incongruent in number was reduced when the sentence included a matching distractor.
 However, the present study investigated only grammatical subject-verb integration and the size of the P600 effect was rather small.
-The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was not ruled out during memory retrieval prior to integration.
-Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.
+The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was \textcolor{blue}{misretrieved.
+Thus semantic cue overload would render memory retrieval more difficult causing a more negative N400 and if the distractor is falsely retrieved, integration with the verb would be easy resulting in a reduced, i.e., more negative, P600.
+This speculative claim should be tested in future work.}}}\label{inhibition_facilitation} Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.
 

### 2024-11-12 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 As we had explained in the response to reviewers on August 29, 2024 (specifically in the response letter and our responses to the comments E.2 and R3.5), the priors were changed from directional priors in the original version to symmetrical priors in the revised version.
+Additionally, we changed the method with which the Bayes factors were computed.
+Initially, we had used bridge sampling, but in the revision, we switched to the Savage-Dickey method because only using that method we were able to compare models with random slopes (this was also explained in the previous response letter, Response to R1.7).
+These changes did not change the results qualitatively, but led to some numerical changes which probably piqued the reviewer's interest.
 
@@
 \subsubsection*{Response to E.4}
-\textcolor{red}{remove repetitions}
+We have removed repetitions in the description of the statistical analyses and have shortened the description of the self-paced reading results (and the comparison to the previous studies).
+This overall reduced the length of the manuscript; after the first round of revisions the main text ended on page 77, now it ends on page \textcolor{red}{XX.}
 
@@
 We agree with the reviewer that the presentation and discussion of the SPR results was misleading and have revised it accordingly.
-We have revised Figure 3 to remove the focus on the critical and surrounding regions (see below).
+\textcolor{red}{We have revised Figure 3 to remove the focus on the critical and surrounding regions (see below)}.
 At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference and that the long-lasting effect caused by that confounds the results at later effects.
@@
 
-\textcolor{red}{revise text and paste here}
-
-\textcolor{red}{SV email: i think we didn't write clearly enough about the fact that syntactic interference *in this design* is expected given the previous work.
-we did say it but needs to be clearer.
-also, when we say we find evidence against syn int, we should always add, in this design.
-and then we should say that this does not mean that there is no syntactic manipulation that would cause no interference, there could be.
-i think that point is valid.
-what i found strange in her comment was that she says both that van dyke and mertzen found syn int, but also that syn int is never going to show up in this design because it is not the right manipulation.
-maybe her point is that we only show that the van dyke and mertzen result was Type M error (i.e., there is no real effect in this design).}
+\textcolor{red}{paste here}
 
@@
 \subsubsection*{Response to R2.6}
-\textcolor{red}{Not sure whether this comment requires action. Maybe 1-2 sentences about Himanshu's work.}
+This comment does not call for further action.
+In section 7.2 ``Encoding and retrieval interference", we discuss that we believe that the difference in presentation rate and whether it was controlled by the participants or not led to different types of interference (encoding vs.\ retrieval).
 
@@
 \subsubsection*{Response to R2.7}
-Yes, the priors were changed in the 1st revision (directional priors in the original version and symmetrical priors in the revised version, see also our Response to E.3).
+Yes, all analyses were changed in the first round of revision.
+Firstly, the priors were changed (directional priors in the original version and symmetrical priors in the revised version).
+Secondly, the method to compute Bayes factors was changed (from bridge sampling to Savage-Dickey, see also our Response to E.3).
 This resulted in small numerical changes of the results, e.g., BF10 = 2.5 in the original analysis decreased to BF10 = 1.2 in the revised analysis.
 Furthermore in the original version of the manuscript, we were inconsistent in our description of small BFs, i.e., BF10 = 1.2.
-Since the first revision and in line with \cite{jeffreys1998theory} and \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence either way.
+Since the first revision and in line with \cite{lee2014bayesian}, we now consistently interpret such small BFs as providing no evidence either way.
 
@@
 This redundancy occurs several times, making the paper excessively and unnecessarily long.
-The authors should streamline the content to enhance clarity and conciseness.''
+\textbf{The authors should streamline the content to enhance clarity and conciseness.}''
 \end{quote}
@@
 \subsubsection*{Response to R2.10}
-\textcolor{red}{remove repetitions}
+We have removed repetitions and shortened the presentation of the SPR results to improve clarity and conciseness (see also Response to E.4).
 

### 2024-11-12 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 
-The SPR study, hereafter Experiment 1a, was initially designed so that participants read items in the first session (1-60) and then, after a gap of 1-20 days, they read the the second set of items (61-120).
-The variation in the gap is due to the fact that we could not control when the participants would do the second session; however, the mean and median gaps were  about 11 days, with the first and third quartiles being 9 and 12 days.
-
-The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
+The SPR study, hereafter Experiment 1a, was initially designed so that participants read \textcolor{blue}{half of the items (1-60) in the first session and the other half of the items (61-120) in the second session on another day.} The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
 The adaptation effect is theoretically interesting per se \citep{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
@@
 
-\clearpage
 \section{Experiment 2: EEG}
@@
 Recently, \cite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
-We found that the reading times at the critical verb were confounded by pre-critical effects, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
+We found that the reading times at the critical verb were confounded by \textcolor{blue}{reading times differences starting earlier in the sentence}, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
 Because the Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions.
@@
 This model naturally does not predict any interference effects.
-It can be regarded the null model or no-interference model.
+It can be regarded the null model.
 
@@
 To quantitatively evaluate the relative predictive performances of the models on the semantic interference data, we used a Bayes factor analysis.
-We computed Bayes factors for the two-cues model compared to the three-cues model and to the one-cue model, respectively.
 Since Bayes factors are sensitive to the choice of priors on the parameters of interest \citep{kass1995bayes}, we computed Bayes factors under six different priors for the scaling parameter $F$,
@@
 
-\noindent where $n$ is the total number of proposals, i.e., $0.2$ million here; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+\noindent where $n$ is the total number of proposals; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
 The term $\mathcal{L}(F_i|y)$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.
-We computed marginal likelihoods under different prior assumptions $p(F)$.
 
@@
 From the marginal likelihoods, we computed Bayes factors for the two-cues model compared to the three-cues model and the one-cue model.
-That is, the marginal likelihood of the two-cues model $M_1$ was divided by (i) the marginal likelihood of the three-cues model $M_2$,  and (ii) the marginal likelihood of the one-cue model $M_3$.
-This calculation resulted in the Bayes factors $BF_{12}$ and $BF_{13}$, respectively.
-Following the convention from \cite{jeffreys1998theory}, a Bayes factor above $10$ is interpreted as strong evidence in favor of the two-cues model, and a value between $3$ and $10$ is interpreted as moderate evidence in favor of the two-cues model compared to the competing model (see also Table \ref{tab:bf_interpretation}).
+That is, the marginal likelihood of the two-cues model was divided by (i) the marginal likelihood of the three-cues model,  and (ii) the marginal likelihood of the one-cue model.
 
 The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
-Compared to the null model, the one-cue model, the Bayes factors indicate strong evidence for the new two-cues model; this comparison shows that there is strong evidence for interference effects.
-Compared to the three-cues model proposed by \cite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior.
+The Bayes factors indicate strong evidence for the new two-cues model \textcolor{blue}{compared to the one-cue model (null model)}; this comparison shows that there is strong evidence for interference effects.
+Compared to the three-cues model proposed by \cite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior \textcolor{blue}{(see Table \ref{tab:bf_interpretation} for the interpretation of Bayes factors)}.
 This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the two separate, previously assumed cues \{$\pm$grammatical subject\} and \{$\pm$same clause\}.
@@
 In comparison to the previous estimates, the effects of the present study were relatively small and associated with less uncertainty in general.
-The present and previous estimates of semantic interference seem relatively consistent: They indicated a positive effect which was largest in the pre-critical region and smallest in the critical region.
-It seems like the present estimates of syntactic interference were smaller, i.e., closer to zero, than most of the previous estimates.
-For the interaction, most estimates were close to zero, but some estimates indicated a positive and some indicated a negative interaction.
-To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies to compare them to the present Bayes factors.
-For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
-To compute Bayes factors for \cite{vandyke07}, we used the Normal-Normal conjugate case \citep{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper.
-Then, using these posteriors and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes factors with the Savage-Dickey method \citep{wagenmakers_savagedickey, vuorre2017_savagedickey, dickey1970, dickey1971weighted, verdinelli1995computing}.
-Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
-The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.
+\textcolor{blue}{We used Bayes factors to compare whether there was evidence for the estimated effects in the previous studies vs.\ the present study.} For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
+\textcolor{blue}{Bayes factors for \cite{vandyke07} were computed using} the Normal-Normal conjugate case \citep{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms).
+Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
+
+\textcolor{blue}{The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.
+The region-wise Bayes factors across studies were consistent with only a few exceptions (e.g., the majority of studies showed no evidence for the interaction), but in some cases there was no clear clustering in favor or against an effect (e.g., regarding semantic interference in the spill-over region).
+Only once did the present study deviate from a clear result if it was consistent across the previous studies: All previous studies showed evidence for syntactic interference in the pre-critical region, while the present study provided evidence against the effect.
+This difference in the results between the present and} previous studies could be explained by differences in i) methodology, ii) items and/or iii) statistical power.
+We discuss each of these differences in turn.
 
@@
 \end{sidewaysfigure}
-\clearpage
-
-In the pre-critical region, the Bayes factors from the present and previous studies showed no clear pattern regarding syntactic interference.
-The present study was the only one to provide evidence against the effect, while the previous studies supported it.
-The Bayes factors aligned qualitatively across studies for the semantic interference effect and the interaction -- consistently providing evidence for semantic interference and against the interaction.
-
-In the critical region, the Bayes factors across studies showed a mixed picture for syntactic interference: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provided extremely strong evidence for it, while \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3 provided weak evidence against it.
-\citeauthor{mertzen}'s (\citeyear{mertzen}) experiments showed weak evidence for the effect.
-The two self-paced reading experiments (present and \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1) provided evidence against it.
-Similarly, there is no clear pattern for semantic interference: Most experiments provided evidence against it, but the two self-paced reading experiments provided evidence for semantic interference.
-The Bayes factors across studies provided evidence against the interaction.
-
-In the spill-over region, the Bayes factors for syntactic interference rather consistently provided evidence against the effect.
-Only \citeauthor{mertzen}'s (\citeyear{mertzen}) English experiment provided moderate evidence for it.
-The Bayes factor results for semantic interference were mixed: the two self-paced reading experiments provided extremely strong evidence for the effect, two of the eye-tracking experiments (\citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 3, \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment) provided weak evidence for the effect and the two other eye-tracking experiments (\citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2, \citeauthor{mertzen}'s (\citeyear{mertzen}) English experiment) provided evidence against the effect.
-Importantly, the clustering in the results did neither align with the investigated language nor with the authors of the studies.
-The Bayes factor results for the interaction showed evidence against it from all experiments but one.
-The outlier was \citeauthor{vandyke07}'s (\citeyear{vandyke07}) SPR experiment, which showed extreme evidence for the interaction.
-
-In sum, while the present and previous studies mostly showed comparable region-wise results for semantic interference and the interaction, they differed remarkably regarding syntactic interference in the pre-critical and critical region.
-In the pre-critical region, the present study was the obvious outlier, being the only study to provide evidence against the effect.
-In the critical region, the Bayes factors for the syntactic interference were very mixed with no clear clustering.
-In conclusion, the most remarkable difference between the present study and previous findings concerns syntactic interference in the pre-critical region.
-The lack of syntactic interference in the pre-critical region in the present study compared to the previous studies could be explained by differences in i) methodology, ii) items and/or iii) statistical power.
-We discuss each of these differences in turn.
-}
 
@@
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.
+Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference \textcolor{blue}{in the present design}.
 }\label{linebreak}
@@
 So, the present study was the only one which had a pre-critical region that did not directly follow a clause boundary and it was also the only one to not find syntactic interference in the pre-critical region.
-These two observations combined suggest that the pre-critical syntactic interference effect in the previous studies might be a spill-over effect from the syntactic manipulation.
+These two observations combined suggest that the pre-critical syntactic interference effect in the previous studies might have been a spill-over effect from the syntactic manipulation.
 The syntactic interference manipulation in all the studies using this design was confounded with syntactic complexity of the material intervening between the subject and the (pre-)critical region.
@@
 While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region.
-However, Figure \ref{fig:whole_sentence} C shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.
+However, Figure \ref{fig:whole_sentence} \textcolor{blue}{C} shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.
 
@@
 The additional distractor (`the secretary' in \ref{ex:mertzen}) was animate and the subject of the relative clause, therefore theoretically it should have affected both syntactic and semantic interference equally.
-But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of verb phrase consisting of an auxiliar and an adjective.
+But it is possible that it increased syntactic interference in \citeauthor{mertzen}'s (\citeyear{mertzen}) experiment to a larger extent because the additional distractor (`the secretary') was a ``stronger'' subject than the manipulated distractor (`the visitor') because it was the subject of a full finite verb while the latter was always the subject of a verb phrase consisting of an auxiliar and an adjective.
 This additional distractor might have increased syntactic interference compared to our study.
@@
 
-Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role in sentence processing.
+Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role  \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations}.
 However, this conclusion would be problematic.
@@
 This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials.
-Both our experiments consistently showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor.
-This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions at the critical verb.
-The observed differences in the time course of the effect are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-Overall, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+Both our experiments consistently showed semantic interference, i.e., more processing difficulty in the presence of a semantically matching distractor.
+This manifested in longer self-paced reading times starting at the distractor \textcolor{blue}{itself which persisted almost until the end of the sentence} and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions at the critical verb.
+The observed differences in the time course of the effect are likely due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
+Overall, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\} \textcolor{blue}{in the present design}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
 A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.

### 2024-11-15 code_data/SPR/r_scripts/Plot_ReadingTimes.R
Updated via the Open Science Framework

@@ -1,5 +1,5 @@
 #author: "Pia Schoknecht"
-#date: "05 11 2024"
+#date: "15 11 2024"
 
 # load packages
 library(plyr)
@@ -8,6 +8,7 @@ library(dplyr)
 library(tidyverse)
 library(cowplot)
 library(ggh4x)
+
 
 # load data
 df1 <- read.csv("../data/pandora_web_data_allsessions_short.csv", sep=";")
@@ -76,166 +77,94 @@ losyn_summ_by_participants$syntactic.interference <- ifelse(losyn_summ_by_partic
 losyn_summ_by_participants$semantic.interference <- ifelse(losyn_summ_by_participants$condition %in% c("a","c"), "low", "high")
 
 # add nice region labels
-losyn_summ_by_participants$region<- ifelse(losyn_summ_by_participants$wordno == 16, "critical",
-                                           ifelse(losyn_summ_by_participants$wordno == 15, "pre-critical", 
-                                                  ifelse(losyn_summ_by_participants$wordno == 14, "pre-pre-critical",        
-                                                         ifelse(losyn_summ_by_participants$wordno == 17, "spill-over",        
-                                                                "other"))))
+lo_wordno <- c(10:18) 
+hi_wordno <- c(11:19) 
+region <- c("distractor-1", "distractor", "distractor+1", "distractor+2",
+            "pre-pre-critical", "pre-critical",
+            "critical", "critical+1", "critical+2")
+fake_wordno <- c(1:9)
+lo_dict.df<- data.frame(lo_wordno, fake_wordno, region) 
+hi_dict.df<- data.frame(hi_wordno, fake_wordno, region) 
 
-hisyn_summ_by_participants$region<- ifelse(hisyn_summ_by_participants$wordno == 17, "critical",
-                                           ifelse(hisyn_summ_by_participants$wordno == 16, "pre-critical", 
-                                                  ifelse(hisyn_summ_by_participants$wordno == 15, "pre-pre-critical",        
-                                                         ifelse(hisyn_summ_by_participants$wordno == 18, "spill-over",        
-                                                                "other"))))
+lo <- left_join(losyn_summ_by_participants, lo_dict.df, by=c("wordno" = "lo_wordno"))
[…]

### 2024-11-15 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 \vskip 1em
-Together with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit our second revision of the article titled ``Do syntactic and semantic similarity lead to interference effects?
+with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit our second revision of our article titled ``Do syntactic and semantic similarity lead to interference effects?
 Evidence from self-paced reading and event-related potentials using German''.
+
 We are grateful for thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
-
-%add description of revisions here
-%To address reviewer \# 2 concern that the detailed description of the SPR results in the critical and surrounding regions might be misleading the reader, we have removed the panels of Figure 3 that highlighted the reading times of these regions.
-Instead, we now show the long-lasting reading time difference starting at the distractor in each of the self-paced reading sub-experiments.
+To address the concern that we overstate the implications of our findings, we have clarified at various places in the abstract and discussion of our results that our findings only apply to the present design.
+Although we have added clarifying text in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), the manuscript length could be shortened by four pages, without compromising on content.
+We have removed repetitions and have revised unnecessary verbose text passages (e.g., in the description of the SPR results).
+We have made our OSF repository public to ensure general accessibility and have updated the link in the manuscript accordingly.
 
@@
 
-\textcolor{red}{wait for ok from shravan}
+\textcolor{red}{do this actually}
 
@@
 \subsubsection*{Response to E.2}
-\textcolor{red}{add *in this design* whenever we say that there is no syn interference}
+We have revised the manuscript to be more cautious regarding the implications and generalizability of our findings. See the revised abstract and related text edits below.
+
+\begin{quote}
+    \Paste{abstract}
+\end{quote}
+
+\noindent Related text edits on pages \pageref{thisdesign1}, \pageref{synsem}, \pageref{thisdesign2} and \pageref{thisdesign3}:
+
+\begin{quote}
+    \Paste{thisdesign1}
+\end{quote}
+
+\begin{quote}
+    \Paste{thisdesign4}
+\end{quote}
+
+\begin{quote}
+    \Paste{thisdesign2}
+\end{quote}
+
+\begin{quote}
+    \Paste{thisdesign3}
+\end{quote}
 
@@
 We have removed repetitions in the description of the statistical analyses and have shortened the description of the self-paced reading results (and the comparison to the previous studies).
-This overall reduced the length of the manuscript; after the first round of revisions the main text ended on page 77, now it ends on page \textcolor{red}{XX.}
+This overall reduced the length of the manuscript; after the first round of revisions the main text ended on page 77, now it ends on page 73.
 
@@
 \begin{quote}
-``p. 76: \textbf{Chromy should appear like this: ``Chromý" (note the correct diacritics above the "y")}
+``p. 76: \textbf{Chromy should appear like this: ``Chromý" (note the correct diacritics above the ``y")}
 
@@
 We agree with the reviewer that the presentation and discussion of the SPR results was misleading and have revised it accordingly.
-\textcolor{red}{We have revised Figure 3 to remove the focus on the critical and surrounding regions (see below)}.
+We have revised Figure 3 to remove the focus on the critical region (see below).
 At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference and that the long-lasting effect caused by that confounds the results at later effects.
@@
 Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure.
-\textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiments 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
+\textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
     \label{fig:whole_sentence}
     \centering
-    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
+    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom.jpg}
 \end{sidewaysfigure}
@@
 Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that it is outside the scope of this manuscript to provide an in-depth discussion of a mechanism of encoding interference.
-To address the reviewer's concern that encoding interference would not cause such long-lasting effects as found in our SPR data, we have added the following text to the manuscript about a recent joint account of encoding and retrieval interference:
-
-\textcolor{red}{Write something about Himanshu's work in relation to this}
 
@@
 We agree with the reviewer that our present work does not rule out syntactic interference in general, but the present study using the present design did not induce considerable syntactic interference.
-We have made the following edits in the manuscript to emphasize this point:
-
-\textcolor{red}{paste here}
+We have revised the abstract and manuscript main body accordingly (see response to E.2).
 
@@
 \subsubsection*{Response to R2.4}
- \textcolor{red}{write response: no facilitation in SPR data because 1.
-distractor effect overshadows everything 2.
-mean effect in RTs cannot show inhibition and facilitation at once, inhibition was stronger}
+The SPR results and the ERP results in the N400 window both show inhibitory interference, i.e., more processing difficulty under high vs.\ low semantic interference, manifesting in longer reading times and a more negative N400.
+We explain below in our response to R2.5 that the reduced P600 for high interference might be explained by facilitatory interference because under high semantic interference whatever noun was retrieved can be easily integrated.
+
+SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region.
+Additionally, facilitated integration could only be observed in the critical (or spill-over) region and in line with the previous reviewer comments, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.
 
@@
 However, the N400 component follows the expected pattern, with stronger activation under high semantic interference.
-It remains unclear how the directionality of these two can be reconciled within the same theoretical framework.''
+\textbf{It remains unclear how the directionality of these two can be reconciled within the same theoretical framework.}''
 \end{quote}
@@
 
-  \textcolor{red}{write response: Fleshed out more how inhibition and facilitation could both occur and manifest in ERPs}
+We have revised our discussion of the ERP results on page \pageref{inhibition_facilitation} to clarify how the seemingly conflicting results can be reconciled (see below).
+Interference might enhance the N400 amplitude because it makes memory retrieval more difficult and it might reduce the P600 because no matter which noun was retrieved it can be easily integrated with the verb.
 
@@
 \subsubsection*{Response to R2.8}
-\textcolor{red}{SV email: the syn / sem comment is also interesting because there are some theories that only assume semantic cue association--there isn't even any syntax.
-some discussion of the theoretically possible connections between syn and sem processing would help.
-e.g., one extreme is that syn and sem operate in tandem (e.g., categorial grammars), others believe that sem processing strictly follows syntactic structure building (classical montagovian stuff), then we have the sem only models (e.g., sentence gestalt model).
-there is no consensus on what the relationship is between syn and sem, so these could operate independently.
-local coherence is also an example--if we assume syntactic processing and no independent semantic processing, local coherence would not arise (depending on how one assumes that syntactic processing works).}
+
+To clarify the assumed independence of syntactic and semantic processing, we have revised section 7.3 starting on page \pageref{synsem}.
+We have also renamed the section to `The use of syntactic and semantic information' (the former title was `Relation to other sentence processing accounts').
+
+\begin{quote}
+    \Paste{synsem}
+\end{quote}
 
@@
 
-\subsection*{Comment R3.1}
 \begin{quote}
@@
 They provide some text that clarifies that they need to implement a clause tracking mechanism.
-\textbf{I would prefer it if the text was clearer about the unfeasibility of a `subject of the same clause' feature.} But this does not impact their findings.''
-\end{quote}
+I would prefer it if the text was clearer about the unfeasibility of a `subject of the same clause' feature.
+But this does not impact their findings.
 
-\subsubsection*{Response to R3.1}
-\textcolor{red}{I'm not sure whether this comment requires action.}
-
-\subsection*{Comment R3.2}
-\begin{quote}
-``Another concern that I raised involves the scalability of the notion that semantic interference effects reflect the use of semantic retrieval cues such as [+animate].
+Another concern that I raised involves the scalability of the notion that semantic interference effects reflect the use of semantic retrieval cues such as [+animate].
 At issue is how this extends to finer-grained properties that are routinely responsible for plausibility violations.
@@
 But it is less plausible that every time the noun `teacher' is encountered the property [+can drive] is activated.
-\textbf{The authors' response seems to be that animacy was sufficient for the materials in the current study, and that maybe other finer-grained features ``become activated only when relevant".
-To my mind, this point removes much of the value of the cue-based retrieval theory.} Nevertheless, this concern is about the generalizability of the authors' claims, and not about the soundness of their results.''
-\end{quote}
-\subsubsection*{Response to R3.2}
-\textcolor{red}{I'm not sure whether this comment requires action.}
+The authors' response seems to be that animacy was sufficient for the materials in the current study, and that maybe other finer-grained features ``become activated only when relevant".
+To my mind, this point removes much of the value of the cue-based retrieval theory.
+\textbf{Nevertheless, this concern is about the generalizability of the authors' claims, and not about the soundness of their results.}
 
-\subsection*{Comment R3.3}
-\begin{quote}
-``Finally, and in a similar vein, I raised questions about whether there might be other reasons for the selective interference effects found in this study.
+Finally, and in a similar vein, I raised questions about whether there might be other reasons for the selective interference effects found in this study.
 Maybe a different mechanism could capture the special status of the subject-of-the-same-clause, or maybe some specific property of the current experimental materials could be responsible for the lack of interference from other subjects.
@@
 
-\subsubsection*{Response to R3.3}
-We have revised the abstract to emphasize that our findings only apply to the present design. We paste the revised abstract here:
-
-\begin{quote}
-    \Paste{abstract}
-\end{quote}
+\subsubsection*{Response to Reviewer \#3}
+We have revised the manuscript (including the abstract) to show more caution regarding the generalizability of our findings (see our response to E.2).
 

### 2024-11-15 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure.
-\textcolor{blue}{Panel C and D show the reading times of the sentence focusing on the regions between the distractor and critical verb for each session of Experiment 1a and 1b; again separately for high (C) and low (D) syntactic interference.}}
+\textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
     \label{fig:whole_sentence}
     \centering
-    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_exp.jpg}
+    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom.jpg}
 \end{sidewaysfigure}
@@
 It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
-This pattern \textcolor{blue}{was found in the pooled data (panels A and B) as well as in the data of each experimental session (panels C and D).
-To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}.
+\textcolor{blue}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}.
 Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session.
@@
 There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
-In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5,
-Normal(0, 0.05): BF$_{10}$ = 1.3,
-Normal(0, 0.1): BF$_{10}$ = 0.7).
+In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5, Normal(0, 0.05): BF$_{10}$ = 1.3, Normal(0, 0.1): BF$_{10}$ = 0.7).
 There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
 Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was anecdotal evidence for syntactic interference (BF$_{10}$ = 2.8).
-{There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
+There was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ $<$ 1).
 Under the narrowest prior in the critical region, the Bayes factor provided no evidence either way (neither for nor against the interaction (BF$_{10}$ = 1.4)).
@@
 
-\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence} A and B).
+\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence}).
 One might argue that this could be due to differences in word length.
@@
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
-A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
+\textcolor{blue}{We revisit the connection between syntax and semantics in section 7.3.} A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
 
 In the remainder of this section, we first provide an in-depth comparison of the present results to the previous studies' results which used the same design.
-Then we discuss two important issues that relate to the broader interference literature.
-The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.
+Then we discuss two important issues that relate to the broader literature.
+The first is the role of encoding vs.\ retrieval interference, and the second is \textcolor{blue}{the use of syntactic and semantic information during sentence processing}.
 
@@
 
-{The difference in methodology led to another difference between our study and the previous studies.
+The difference in methodology led to another difference between our study and the previous studies.
 In their eye-tracking experiments, \cite{vandyke07} and \cite{mertzen} presented the stimuli on a single line.
@@
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference \textcolor{blue}{in the present design}.
-}\label{linebreak}
+\Copy{thisdesign1}{Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference \textcolor{blue}{in the present design}.
+}\label{thisdesign1}
 
@@
 While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region.
-However, Figure \ref{fig:whole_sentence} \textcolor{blue}{C} shows that the pre-pre-critical region (n-2) in our items did not show a syntactic effect.
+However, Figure \ref{fig:whole_sentence} \textcolor{blue}{C} shows that the pre-pre-critical region in our items did not show a syntactic effect.
 
@@
 
-\subsection{Relation to other sentence processing accounts}
-
-The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
-{A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
-This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.}\label{future_work}
-
-However, the semantic interference effect we found can be easily reconciled with previous work in  sentence processing.
-In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model.
- Our results could also be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
+\subsection{\textcolor{blue}{The use of syntactic and semantic information}}
+
+\label{synsem}\Copy{synsem}{\Copy{thisdesign4}{\textcolor{blue}{Our results suggest a remarkable disconnect in the use of syntactic and semantic cues during subject-verb dependency resolution in sentence configurations like the ones investigated in the present study.}} The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
+A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
+This future work should utilize the findings from the literature on agreement attraction which has shown that the hierarchical position of the distractor can determine whether or not it affects processing \citep{franck2002subject, franck2006agreement, franck2020hierarchical, parker2018not}.
+
+\textcolor{blue}{Although the syntactic cue included hierarchical information to match only with the correct subject within the same clause, this did not block the semantic cue from matching with the embedded distractor.
+This finding is not consistent with syntax-first models like garden-path theory \citep{frazier1987} which assume that syntactic processing precedes all other levels of linguistic evaluation (e.g., plausibility).
+Our findings rather suggest parallel, fully independent matching of different retrieval cues with the candidates in memory.}  In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model \textcolor{blue}{assuming independent cues.}
+
+The semantic interference effect we found can be easily reconciled with \textcolor{blue}{other sentence processing theories that do not assign a special status to syntactic processing}.} Our results could be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
 The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands.
@@
 
-Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role  \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations}.
+\Copy{thisdesign2}{Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role  \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations}}\label{thisdesign2}.
 However, this conclusion would be problematic.
@@
 The observed differences in the time course of the effect are likely due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-Overall, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\} \textcolor{blue}{in the present design}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.
+\Copy{thisdesign3}{Overall, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\} \textcolor{blue}{in the present design}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval}\label{thisdesign3}.
 A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.

### 2024-11-20 paper/revision2/elsarticle-template-harv.tex
Fixed a typo

@@
 \begin{equation}
-ML = \frac{1}{n} \sum_{i=1}{n} \frac{\mathcal{L}(F_i|y) p(F_i)}{q(F_i)} \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0.0.025) 
+ML = \frac{1}{n} \sum_{i=1}{n} \frac{\mathcal{L}(F_i|y) p(F_i)}{q(F_i)} \text{ where } F_i \sim Normal_{lb=0,ub=0.05}(0, 0.025)
 \end{equation}
 
-\noindent where $n$ is the total number of proposals; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0.0.025) $.
+\noindent where $n$ is the total number of proposals; $y$ is the observed data, i.e., the observed semantic interference effect in our case; and $F_i$ is the $i^{th}$ proposal from the importance density $Normal_{lb=0,ub=0.05}(0, 0.025) $.
 The term $\mathcal{L}(F_i|y)$ gives the likelihood of proposal $F_i$ given the data $y$, $p(F_i)$ gives the prior density of $F_i$, and $q(F_i)$ gives the importance density of $F_i$.

### 2024-11-22 elsarticle-template-harv.tex
Add files via upload

--- 2024-11-20 paper/revision2/elsarticle-template-harv.tex
@@
 \textcolor{blue}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
-A Bayes factor analysis showed that both the self-paced reading times as well as event-related potentials provided clear evidence for a semantic interference effect.
+A Bayes factor analysis showed that both the self-paced reading times as well as event-related potentials provided evidence for a semantic interference effect.
 In the self-paced reading experiment, \textcolor{blue}{semantic interference due to the semantic cue \{+animate\}} caused a reading time slow-down starting at the distractor.
@@
 \section{Data availability statement}
-All materials, data and analysis scripts can be accessed via \\ \href{https://osf.io/4fpru/?view\_only=2d2a004d3e4147bc975b5e8c455c27df}{https://osf.io/4fpru/?view\_only=2d2a004d3e4147bc975b5e8c455c27df}
+\Copy{link}{All materials, data and analysis scripts can be accessed via \href{https://osf.io/4fpru/}{https://osf.io/4fpru/}}
 
@@
 If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
-The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, where $\Psi(.|\delta)$ represents a density kernel that weights a proposal $F_i$ based on the distance between model-generated data $y$ and actual data, $d(x_i,y)$.
-We chose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
+\Copy{modeling}{The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, \textcolor{blue}{where the function $\Psi(d, 0, \delta)$} represents a density kernel that weights a proposal $F_i$ based on the \textcolor{blue}{distance $d$ between model-generated data $y$ and actual data $x_i$}.}\label{modeling} We chose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
 The $\delta$ parameter determines how good the approximation is: the lower the value of $\delta$, the better the approximation of the likelihood.

### 2024-11-22 paper/revision2/ResponseToReviewers.tex
Add files via upload

@@
 \usepackage{float}
+\usepackage{hyperref}
 \usepackage{subfigure}
@@
 \vskip 1em
-with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit our second revision of our article titled ``Do syntactic and semantic similarity lead to interference effects?
+with my co-authors Shravan Vasishth and Himanshu Yadav, I would like to submit the second revision of our article titled ``Do syntactic and semantic similarity lead to interference effects?
 Evidence from self-paced reading and event-related potentials using German''.
 
-We are grateful for thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
+We are grateful for the thoughtful and constructive comments and suggestions from yourself and the reviewers, and feel that the quality of the manuscript has been greatly improved by the revisions.
 To address the concern that we overstate the implications of our findings, we have clarified at various places in the abstract and discussion of our results that our findings only apply to the present design.
-Although we have added clarifying text in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), the manuscript length could be shortened by four pages, without compromising on content.
-We have removed repetitions and have revised unnecessary verbose text passages (e.g., in the description of the SPR results).
+Although we have added clarifying text in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), the manuscript length could be shortened by four pages.
+We have removed repetitions and revised unnecessary verbose text passages (e.g., in the description of the SPR results).
 We have made our OSF repository public to ensure general accessibility and have updated the link in the manuscript accordingly.
@@
 However, it's important to note that that is not guaranteed, and one reason I have come to a `revise' rather than `accept subject to minor revisions' is that, like Reviewer 1, I was unable to access your data via your OSF link.
-\textbf{It is very important to us at JML that data are readily accessible, and that accessibility needs to be in place before any potential acceptance decision}.
-''
+\textbf{It is very important to us at JML that data are readily accessible, and that accessibility needs to be in place before any potential acceptance decision}.''
 \end{quote}
@@
 \subsubsection*{Response to E.1}
-We have made the OSF repository public and updated the link in the manuscript. This should grant general accessibility.
+We have made the OSF repository public to grant general accessibility. We paste our data availability statement below.
 
-\textcolor{red}{do this actually}
+\begin{quote}
+    \Paste{link}
+\end{quote}
 
@@
 \subsubsection*{Response to R1.1}
-A formatting error broke the link in the previous version of the manuscript. We have made the OSF repository public and updated the link in the manuscript. This should grant general accessibility.
+We have made the OSF repository public to grant general accessibility (see also our Response to E.1).
 
@@
 \subsubsection*{Response to R1.4}
-We have added the suggested reference (now on page \pageref{cite_kush}):
+We have added the suggested reference on page \pageref{cite_kush}:
 
@@
 \subsubsection*{Response to R1.5}
-\textcolor{red}{Opened issue and assigned to Himanshu}
+We have changed the expression to make it clearer. On page \pageref{modeling}, we now say:
+
+\begin{quote}
+    \Paste{modeling}
+\end{quote}
 
@@
 
-SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region.
-Additionally, facilitated integration could only be observed in the critical (or spill-over) region and in line with the previous reviewer comments, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.
+In general, SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region.
+Additionally, only at the critical (or spill-over) region facilitated integration might be observed and in line with the reviews, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.
 
@@
 
-To clarify the assumed independence of syntactic and semantic processing, we have revised section 7.3 starting on page \pageref{synsem}.
+To clarify the assumed independence of syntactic and semantic processing, we have revised section 7.3, starting on page \pageref{synsem}.
 We have also renamed the section to `The use of syntactic and semantic information' (the former title was `Relation to other sentence processing accounts').
@@
 \subsubsection*{Response to R2.10}
-We have removed repetitions and shortened the presentation of the SPR results to improve clarity and conciseness (see also Response to E.4).
+We have removed the repetitions to improve clarity and conciseness (see also Response to E.4).
 

### 2024-11-22 code_data/SPR/r_scripts/Plot_stats.R
Updated via the Open Science Framework

@@ -12,7 +12,7 @@ library(cowplot)
 # Pre-critical region
 
 # load model fits
-load("E://model_fits/Fit_m_Full_precrit_slopes_symmetrical.Rda")
+load("model_fits/Fit_m_Full_precrit_slopes_symmetrical.Rda")
 m_precrit <- m_m_full
 rm(list = c("m_m_full"))
 
@@ -51,7 +51,7 @@ posteriors_precrit <- mcmc_areas(
                           annotate("text", x = 35, y = 1.8, label = "-5 — 8 — 22", size=3) 
 
 # Critical region
-load("E://model_fits/Fit_m_Full_crit_slopes_symmetrical.Rda")
+load("model_fits/Fit_m_Full_crit_slopes_symmetrical.Rda")
 m_crit <- m_m_full
 rm(list = c("m_m_full"))
 
@@ -86,7 +86,7 @@ posteriors_crit <- mcmc_areas(
                       annotate("text", x = 35, y = 1.8, label = "-3 — 14 — 31", size=3) 
 
 # Spill-over region
-load("E://model_fits/Fit_m_Full_spill_slopes_symmetrical.Rda")
+load("model_fits/Fit_m_Full_spill_slopes_symmetrical.Rda")
 m_spill <- m_m_full
 rm(list = c("m_m_full"))
 

### 2024-11-22 code_data/SPR/r_scripts/Plot_ReadingTimes.R
Updated via the Open Science Framework

@@ -45,27 +45,27 @@ df_losyn <- df_trim %>% filter(condition == "a" | condition == "b")
 #### Plot whole sentence ####
 
 #calculate mean reading time per subject per word per condition
-hisyn_by_participants <- summarize(
-  group_by(
+hisyn_by_participants <- dplyr::summarize(
+  dplyr::group_by(
     filter(df_hisyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "21")),
     wordno,condition,subj, exp),
   logrt=mean(log(rt)))
 
-losyn_by_participants <- summarize(
-  group_by(
+losyn_by_participants <- dplyr::summarize(
+  dplyr::group_by(
     filter(df_losyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "20")),
     wordno,condition,subj, exp),
   logrt=mean(log(rt)))
 
 # calculate mean per word per condition (averaged over subjects)
-hisyn_summ_by_participants <- summarize(
-  group_by(hisyn_by_participants,wordno,condition),
+hisyn_summ_by_participants <- dplyr::summarize(
+  dplyr::group_by(hisyn_by_participants,wordno,condition),
   meanlogrt=mean(logrt), 
   SE = sd(logrt)/sqrt(nrow(hisyn_by_participants)), 
   low2SE= meanlogrt - 2*SE, high2SE= meanlogrt + 2*SE)
 
-losyn_summ_by_participants <- summarize(
-  group_by(losyn_by_participants,wordno,condition),
+losyn_summ_by_participants <- dplyr::summarize(
+  dplyr::group_by(losyn_by_participants,wordno,condition),
   meanlogrt=mean(logrt), 
   SE = sd(logrt)/sqrt(nrow(losyn_by_participants)), 
   low2SE= meanlogrt - 2*SE, high2SE= meanlogrt + 2*SE)
@@ -166,5 +166,5 @@ p_zoom <- ggplot(data=zoom, aes(x=fake_wordno, y=meanlogrt,
 
 # Put everything together
 plot_grid(p_hisyn, p_losyn, p_zoom, ncol=1, rel_heights = c(0.9,0.9,1), labels="AUTO")
[…]

### 2024-11-22 code_data/SPR/r_scripts/acc_analysis.R
Updated via the Open Science Framework

@@ -26,13 +26,12 @@ df <- rbind(df1, df2, df3)
 # subset with only one row per subject and item with question
 acc <- df %>% 
   drop_na(question_rt) %>% 
-  filter(wordno == 01)# %>%
-# filter(question_rt < 20000)
+  filter(wordno == 01)
 
 #calculate accuracy per participant
 acc_subj <- table(acc$correct, acc$fullname, acc$session) %>%
   as.data.frame() %>%
-  rename(answ = "Var1", subj = "Var2", session = "Var3")
+  dplyr::rename(answ = "Var1", subj = "Var2", session = "Var3")
 
 #answers in percent (number of questions per subject: 32)
 acc_subj$percent <- acc_subj$Freq/32*100
@@ -71,7 +70,7 @@ acc <- df2 %>%
 #calculate accuracy per participant
 acc_subj <- table(acc$correct, acc$fullname, acc$condition, acc$session) %>%
   as.data.frame() %>%
-  rename(answ = "Var1", subj = "Var2", cond = "Var3", session ="Var4")
+  dplyr::rename(answ = "Var1", subj = "Var2", cond = "Var3", session ="Var4")
 
 
 #answers in percent (
@@ -87,7 +86,7 @@ round(range(acc_cor$percent),1)
 # calculate accuracy per condition
 acc_per_cond <- acc_cor %>%
   group_by(cond) %>%
-  summarise(percent = round(mean(percent),1))
+  dplyr::summarise(percent = round(mean(percent),1))
 
 ggplot(data=acc_per_cond, aes(x=cond, y=percent)) +
   geom_bar(stat="identity")

### 2024-11-22 code_data/EEG/r_scripts/analyses.R
Update analyses.R

@@ -8,12 +8,13 @@ library(brms)
 library(bayesplot)
 library(ggplot2)
 library(cowplot)
+library(extraDistr)
 
 #### prepare ####
 #load data
-prestim <- read_csv("data/Pandora_mean_200_0.csv")
-n400 <- read_csv("data/Pandora_mean_300_500.csv")
-p600 <- read_csv("data/Pandora_mean_600_900.csv")
+prestim <- read_csv("../data/Pandora_mean_200_0.csv")
+n400 <- read_csv("../data/Pandora_mean_300_500.csv")
+p600 <- read_csv("../data/Pandora_mean_600_900.csv")
  
 # remove subjects who have only one good session
 xtabs(~ session + subject, n400)

### 2024-11-22 code_data/plausibility_norming/analysis_ordinal.R
Update analysis_ordinal.R

@@ -7,6 +7,7 @@ library(tidyverse)
 library(ordinal)
 library(brms)
 library(bayesplot)
+library(extraDistr)
 
 # load data
 df <- read.csv("norming_data_HighvsLowSem.csv")

### 2024-11-22 paper/revision2/elsarticle-template-harv.tex
Add files via upload

@@
 
-Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panels A and C)} and low \textcolor{blue}{(panels B and D)} syntactic interference because these conditions had partially different sentence structures.
+Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panel A)} and low \textcolor{blue}{(panel B)} syntactic interference because these conditions had partially different sentence structures.
 

### 2024-11-23 paper/revision2/ResponseToReviewers.tex
minor edits

@@
 \title{Response to editor and reviewers (JML-24-17)}
-\author{Pia Schoknecht and Shravan Vasishth}
+\author{Pia Schoknecht, Himanshu Yadav, and Shravan Vasishth}
 %\date{}                                           % Activate to display a given date or no date
@@
 
-\noindent \textbf{Prof.\ Jarrold}\\
+\noindent \textbf{Prof.\ Christopher Jarrold}\\
 \noindent \textbf{Associate Editor}\\
@@
 To address the concern that we overstate the implications of our findings, we have clarified at various places in the abstract and discussion of our results that our findings only apply to the present design.
-Although we have added clarifying text in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), the manuscript length could be shortened by four pages.
-We have removed repetitions and revised unnecessary verbose text passages (e.g., in the description of the SPR results).
-We have made our OSF repository public to ensure general accessibility and have updated the link in the manuscript accordingly.
+
+Clarifying text has been added in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), and we have shortened the manuscript length by four pages.
+We have removed redundant text passages (e.g., in the description of the SPR results).
+
+The OSF repository has bee made public to ensure general accessibility. We have updated the link in the manuscript accordingly.
 

### 2024-11-24 paper/revision2/elsarticle-template-harv.tex
edited sections

@@
 The event-related potentials showed a more negative N400 amplitude at the critical verb, i.e., the retrieval site, for high vs.\ low semantic interference.
-Surprisingly, in both experiments, Bayes factor analyses showed evidence against interference due to \textcolor{blue}{the syntactic cue \{+ grammatical subject\} in this design}.
+Surprisingly, \textcolor{blue}{in this particular design,} in both experiments Bayes factor analyses showed evidence  against interference due to \textcolor{blue}{the syntactic cue \{+ grammatical subject\}}.
 This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play an equal role in retrieval.
-We show through computational modeling that cue-based retrieval also shows no syntactic interference \textcolor{blue}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in.
+We show through computational modeling that cue-based retrieval \textcolor{blue}{will} also shows no syntactic interference \textcolor{blue}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in.
 Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model would be consistent with the observed patterns in our data.
-
-In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, in our data, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects \textcolor{blue}{in sentence configurations like the ones investigated in the present study}.
+In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, \textcolor{blue}{at least in sentence configurations like the ones investigated in the present study}, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
 }
@@
 Thus semantic cue overload would render memory retrieval more difficult causing a more negative N400 and if the distractor is falsely retrieved, integration with the verb would be easy resulting in a reduced, i.e., more negative, P600.
-This speculative claim should be tested in future work.}}}\label{inhibition_facilitation} Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.
+This speculative claim would of course need to be tested in future work.}}}\label{inhibition_facilitation} Since the Bayes factor analysis provided stronger support for the semantic interference effect in the N400 (largest BF$_{10}$ = 17.5) than in the P600 (largest BF$_{10}$ = 4.3), we focus on the N400 results in the remainder of this paper.
 
@@
 explanation for the lack of syntactic interference in our data compared to the previous studies.
-\Copy{SPR_goodenough}{\textcolor{blue}{However, due to the demanding task comprehenders might trade off processing depth and adapt good-enough processing \citep{ferreira2007goodenough}.
+\Copy{SPR_goodenough}{\textcolor{blue}{However, due to the demanding task comprehenders might trade off processing depth and adopt good-enough processing \citep{ferreira2007goodenough}.
 Consequently, they would not fully parse embedded structures, hence there would be no syntactic interference from embedded subjects.
@@
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-\Copy{thisdesign1}{Nevertheless, both the SPR and the EEG experiment showed predominantly evidence against syntactic interference \textcolor{blue}{in the present design}.
-}\label{thisdesign1}
+\Copy{thisdesign1}{Nevertheless, \textcolor{blue}{in the present design,} both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.}\label{thisdesign1}
 
@@
 
-\label{synsem}\Copy{synsem}{\Copy{thisdesign4}{\textcolor{blue}{Our results suggest a remarkable disconnect in the use of syntactic and semantic cues during subject-verb dependency resolution in sentence configurations like the ones investigated in the present study.}} The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
+\label{synsem}\Copy{synsem}{\Copy{thisdesign4}{\textcolor{blue}{In the sentence configurations that were investigated in the present study, we find a remarkable disconnect in the use of syntactic and semantic cues during subject-verb dependency resolution.}} The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
 A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
@@
 Regarding syntactic local coherence effects, a large-sample study \citep{lcpaape2025} presents evidence against syntactic local coherence.
-Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show any evidence of number agreement attraction \citep{chromy2023number}.
-Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.
-Of course, this does not rule out the possibility that, under certain conditions, other factors dominate.
+Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show convincing evidence of number agreement attraction \citep{chromy2023number}.
+\Copy{gendisccaveat}{Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.}\label{gendisccaveat} \Copy{gendisccaveat2}{\textcolor{blue}{Moreover, it is entirely possible that a stronger syntactic manipulation than the one we used ends up showing a syntactic interference effect of the type that \citet{vandyke07} originally reported.
+Indeed, our own findings would need to be replicated, ideally by an independent research group, if we want to be sure that in the present design there is no evidence for syntactic interference.
+If such a replication attempt is carried out, it would also be useful to conduct an lab-based self-paced reading study to investigate our speculation earlier that the absence of syntactic interference in our online SPR study may have been due to participants adopting a good-enough processing strategy, leading to only semantic interference being detectable.}}\label{gendisccaveat2}
 
@@
 The observed differences in the time course of the effect are likely due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-\Copy{thisdesign3}{Overall, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\} \textcolor{blue}{in the present design}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval}\label{thisdesign3}.
+\Copy{thisdesign3}{Overall, \textcolor{blue}{in the present design}, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as was previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval}\label{thisdesign3}.
 A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.

### 2024-11-24 paper/revision2/ResponseToReviewers.tex
edited responses

@@
 \subsubsection*{Response to E.2}
-We have revised the manuscript to be more cautious regarding the implications and generalizability of our findings. See the revised abstract and related text edits below.
+We have revised the manuscript to be more cautious regarding the implications and generalizability of our findings.
+The most that we can claim in this paper is that \textit{in the particular experiment design} that we use here, we have no decisive evidence for syntactic interference.
+But this does not mean that syntactic interference cannot occur in general.
+We make this point clear in our revision now.
+
+See the revised abstract and related text edits below.
+
+\bigskip
+\noindent Revised abstract:
 
@@
     \Paste{abstract}
+\end{quote}
+
+In the general discussion (page \pageref{gendisccaveat}), we already had an unambiguous statement 
+that, given the broader evidence in sentence processing, syntax generally does play a role in dependency completion:
+
+\begin{quote}
+    \Paste{gendisccaveat}
+\end{quote}
+
+We have now made this section even more explicit by stating that syntactic interference may well play a role in some other, stronger syntactic manipulation than the one that we used, and that our own results would need to be independently replicated in order to establish their robustness:
+
+\begin{quote}
+    \Paste{gendisccaveat2}
 \end{quote}
@@
 We have removed repetitions in the description of the statistical analyses and have shortened the description of the self-paced reading results (and the comparison to the previous studies).
-This overall reduced the length of the manuscript; after the first round of revisions the main text ended on page 77, now it ends on page 73.
+This reduced the length of the manuscript by four ages; after the first round of revisions the main text ended on page 77, now it ends on page 73.
 
@@
 \begin{quote}
-``p. 76: \textbf{Chromy should appear like this: ``Chromý" (note the correct diacritics above the ``y")}
-
-Thank you for your manuscript!''
+``p. 76: \textbf{Chromy should appear like this: ``Chromý" (note the correct diacritics above the ``y")}''
 \end{quote}
@@
 Given that the present manuscript contains two large-sample experiments and computational modeling (and its substantial length), we think that it is outside the scope of this manuscript to provide an in-depth discussion of a mechanism of encoding interference.
+Developing a complete treatment of encoding interference would require considerably more computational modeling, which would further increase the length and complexity of this paper.
+In our opinion, such a model should be developed, but that would need to be a stand-alone article, which would also require more experimental studies to validate and test competing models' predictions.
+For a recent attempt that we made to try to implement different versions of feature distortion, a phenomenon that is related to encoding interference, see:
+
+Himanshu Yadav, Garrett Smith, Sebastian Reich, and Shravan Vasishth. Number feature distortion modulates cue-based retrieval in reading. Journal of Memory and Language, 129, 2023.
+
+So, in summary, a proper treatment of encoding interference would require a modeling effort at the same scale as was carried out in the above-mentioned paper.
+Such a modeling effort, although very worthwhile, would take several years' of work and is therefore far beyond the scope of the present work.
 
@@
 \subsubsection*{Response to R2.3}
-We agree with the reviewer that our present work does not rule out syntactic interference in general, but the present study using the present design did not induce considerable syntactic interference.
+We agree with the reviewer that our present work does not rule out syntactic interference in general.
+We have tried to stress throughout the paper that the present study using the present design did not induce considerable syntactic interference.
 We have revised the abstract and manuscript main body accordingly (see response to E.2).
+In the general discussion, we have also added the following caveats:
+
+\begin{quote}
+    \Paste{gendisccaveat2}
+\end{quote}
+
+ We feel that the above quote and the other re-statements in the paper (including the abstract) appropriately limit the claims we make in the paper.
+We never state anywhere in the paper that syntactic interference is generally absent in sentence processing.
 
@@
 In general, SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region.
-Additionally, only at the critical (or spill-over) region facilitated integration might be observed and in line with the reviews, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.
+Additionally, only at the critical (or spill-over) region facilitated integration might be observed and in line with the reviewer's other comments, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.
 

### 2024-11-25 paper/revision2/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-To clarify the assumed independence of syntactic and semantic processing, we have revised section 7.3, starting on page \pageref{synsem}.
-We have also renamed the section to `The use of syntactic and semantic information' (the former title was `Relation to other sentence processing accounts').
+To clarify the assumed independence of syntactic and semantic processing, we have revised the former section `Relation to other sentence processing accounts', starting on page \pageref{synsem}.
+We have also renamed the section to `The use of syntactic and semantic information'.
 

### 2024-11-25 paper/revision2/elsarticle-template-harv.tex
Updated via the Open Science Framework

@@
 \Copy{abstract}{
-Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed to interpret the sentence.
-If these retrieval cues match with not only the features on the noun that is the retrieval target but also with those on other nouns in the sentence, then processing difficulty arises at the verb.
+Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed.
+If these retrieval cues match with not only the \textcolor{blue}{subject} but also with those on other nouns in the sentence, then processing difficulty arises at the verb.
 This difficulty in identifying the correct dependent is called similarity-based interference.
-We present large-sample self-paced reading (N = 774) and event-related potentials experiments (N = 103) using a well-established design \citep{vandyke07,mertzen}, to investigate interference due to syntactic and semantic cues in German.
+We present large-sample self-paced reading and event-related potentials experiments using a well-established design to investigate interference due to syntactic and semantic cues in German.
 \textcolor{blue}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
-A Bayes factor analysis showed that both the self-paced reading times as well as event-related potentials provided evidence for a semantic interference effect.
-In the self-paced reading experiment, \textcolor{blue}{semantic interference due to the semantic cue \{+animate\}} caused a reading time slow-down starting at the distractor.
-The event-related potentials showed a more negative N400 amplitude at the critical verb, i.e., the retrieval site, for high vs.\ low semantic interference.
-Surprisingly, \textcolor{blue}{in this particular design,} in both experiments Bayes factor analyses showed evidence  against interference due to \textcolor{blue}{the syntactic cue \{+ grammatical subject\}}.
+\textcolor{blue}{Bayes factors analyses showed} evidence for a semantic interference effect \textcolor{blue}{in both experiments}. 
+Surprisingly, Bayes factors provided evidence against interference due to \textcolor{blue}{the syntactic cue \{+grammatical subject\} in this particular design in both experiments}.
 This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play an equal role in retrieval.
@@
 Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model would be consistent with the observed patterns in our data.
-In sum, this large-sample study of German using self-paced reading and event-related potentials suggests that, \textcolor{blue}{at least in sentence configurations like the ones investigated in the present study}, although both syntactic and semantic cues drive retrieval at the verb, syntactic cues may include hierarchical syntactic information, which can render the parser immune to syntactic interference effects.
 }
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 
-\section{The present study}
+\section*{The present study}
 
@@
 
-\subsection{Materials}
+\subsection*{Materials}
 
@@
 
-\subsubsection{Plausibility norming} \label{plausib_normin_orig}
+\subsubsection*{Plausibility norming} \label{plausib_normin_orig}
 \citeauthor{vandyke07} conducted a pre-test to ensure that the distractor in high semantic interference conditions was a sufficiently probable subject of the critical verb, i.e., that it could potentially induce semantic interference.
@@
 
-\subsection{The structure and sample sizes of the SPR and EEG experiments}
+\subsection*{The structure and sample sizes of the SPR and EEG experiments}
 
@@
 
-\section{Data availability statement}
+\section*{Data Availability}
 \Copy{link}{All materials, data and analysis scripts can be accessed via \href{https://osf.io/4fpru/}{https://osf.io/4fpru/}}
 
-\section{Experiments 1a and 1b: SPR}
-
-\subsection{Methods}
-\subsubsection{Participants}
+\section*{Experiments 1a and 1b: SPR}
+
+\subsection*{Methods}
+\subsubsection*{Participants}
 
@@
 
-\subsubsection{Procedure}
+\subsubsection*{Procedure}
 Participants completed a moving window self-paced reading task \citep{just_etal_1982} on the PCIbex Farm \citep{pcibex}.
@@
 
-\subsubsection{Statistical analyses}
+\subsubsection*{Statistical analyses}
 
@@
 
-\subsection{Results}
-\subsubsection{Comprehension question accuracy (Experiments 1a and 1b combined)}
+\subsection*{Results}
+\subsubsection*{Comprehension question accuracy (Experiments 1a and 1b combined)}
 
@@
 
-\subsubsection{Self-paced reading times (Experiments 1a and 1b combined)}
+\subsubsection*{Self-paced reading times (Experiments 1a and 1b combined)}
 
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 \Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation.
@@
 
-\subsubsection{\textcolor{blue}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}
+\subsubsection*{\textcolor{blue}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}
 
@@
 
-\section{Experiment 2: EEG}
-\subsection{Methods}
-\subsubsection{Participants}
+\section*{Experiment 2: EEG}
+\subsection*{Methods}
+\subsubsection*{Participants}
 146 participants from the University of Potsdam participant pool took part in the experiment.
@@
 
-\subsubsection{Procedure}
+\subsubsection*{Procedure}
 The experiment was conducted in two experimental sessions for practical reasons (each of the sessions lasted approximately two hours).
@@
 
-\subsubsection{EEG recording and processing}
+\subsubsection*{EEG recording and processing}
 The EEG was recorded with 24 Ag/AgCl scalp electrodes, positioned according to the international 10-20 system.
@@
 
-\subsubsection{Statistical analyses}
+\subsubsection*{Statistical analyses}
 The statistical analyses of the comprehension accuracy in the EEG experiment were carried out in the same manner as the analyses of the comprehension accuracy in the SPR experiment.
@@
 
-\subsection{Results}
-\subsubsection{Comprehension question accuracy}
+\subsection*{Results}
+\subsubsection*{Comprehension question accuracy}
 After exclusion of participants with accuracy below 70\,\%, the overall accuracy (including fillers) was 82.8\,\% (range: [71.9, 100]\,\%).
@@
 
-\subsubsection{Event-related potentials}
+\subsubsection*{Event-related potentials}
 
@@
 
-\subsection{Discussion}
+\subsection*{Discussion}
 We presented a large-scale ERP experiment investigating syntactic and semantic interference during subject-verb dependency completion.
@@
 
-\section{Quantitative predictions of the Lewis and Vasishth (2005) model for ERPs elicited in the present design}
+\section*{Quantitative predictions of the Lewis and Vasishth (2005) model for ERPs elicited in the present design}
 
@@
 
-\section{General discussion}
+\section*{General discussion}
 
@@
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
-\textcolor{blue}{We revisit the connection between syntax and semantics in section 7.3.} A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
+\textcolor{blue}{We revisit the connection between syntax and semantics below.} A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
 
@@
 
-\subsection{Comparison with previous findings}\label{comparison}
+\subsection*{Comparison with previous findings}\label{comparison}
 {The present study used the same design as two previous studies \citep{vandyke07, mertzen}.
@@
 
-\subsection{Encoding and retrieval interference}
+\subsection*{Encoding and retrieval interference}
 In the SPR data, we found that the semantic interference effect originated at the distractor and persisted throughout the following regions.
@@
 
-\subsection{\textcolor{blue}{The use of syntactic and semantic information}}
+\subsection*{\textcolor{blue}{The use of syntactic and semantic information}}
 
@@
 
-The semantic interference effect we found can be easily reconciled with \textcolor{blue}{other sentence processing theories that do not assign a special status to syntactic processing}.} Our results could be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
+The semantic interference effect we found can be easily reconciled with \textcolor{blue}{other sentence processing theories that do not assign a special status to syntactic processing}.
+Our results could be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
 The high number of participants that needed to be excluded in our experiments due to accuracy below 70\,\% (117 out of 908 participants in the SPR experiment, 29 out of 146 participants in the EEG experiment), suggests that our materials led to high processing demands.
@@
 This model assumes that language comprehension relies on associative form-to-meaning mapping instead of syntactic rules.
-This assumption is consistent with our results.
+This assumption is consistent with our results.}
 
@@
 
-\section{Conclusion}
+\section*{Conclusion}
 This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials.
@@
 
+\section*{Acknowledgements}
+This research was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation), Project ID 317633480, SFB 1287 and the University of Potsdam, Germany.
+We thank Johanna Thieke, Romy Leue, Elise Oltrogge and Lisa Plagemann for assistance during data acquisition.
+We are grateful to Dario Paape for helpful discussions and three anonymous reviewers for thoughtful and constructive comments.
+
 %% If you have bibdatabase file and want bibtex to generate the

### 2024-11-26 paper/revision2/elsarticle-template-harv.tex
Updated via the Open Science Framework

@@
 The distractor and the immediately following regions differed between conditions, \textcolor{blue}{therefore we did not analyze their reading times.
-The reading times difference starting at the distractor and persisting almost until the end of the sentence is problematic for the analysis and interpretation of reading times in the later regions, i.e., the critical region.
+In our materials, there were two words between the distractor and the critical verb which were identical across conditions: the pre-pre-critical and pre-critical adverbs.
+The pre-pre-critical region was introduced in the present study to absorb potential effects caused by processing the clause boundary of the embedded clause, rendering it a problematic region for analyses.
+The pre-critical region was not adjacent to clause boundaries.
+The reading times of the pre-critical region were analyzed to investigate the reading times difference starting at the distractor.
+This difference starting at the distractor and persisting almost until the end of the sentence was problematic for the analysis and interpretation of reading times in the critical (and spill-over) region.
 Any effects that might be present there cannot be attributed clearly to the processing of the respective region.
-However, for comparability with previous work, we briefly report analyses of the reading times of the later regions, i.e., the pre-critical, critical and spill-over region.}
+However, for comparability with previous work, we briefly also report analyses of the reading times of the critical and spill-over region.}
 
@@
 \paragraph{Hypothesis testing using Bayes factors}
-\textcolor{blue}{To complete the analysis of the effects in the critical and surrounding regions -- although they are confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
+\textcolor{blue}{To complete the analysis of the effects in the pre-critical, critical and spill-over regions -- although they were confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
 There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
@@
 In contrast, there was evidence against syntactic interference and an interaction in that region.
-We refrain from interpreting effects which were observed in later regions, i.e., the critical verb and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes.
+We refrain from interpreting effects in the critical and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes, i.e., memory retrieval.
 The reading times differences at the distractor and the pre-critical region are discussed below.}}
@@
 
-Indeed, previous work using the same design as the present paper had also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
+\Copy{previous_encoding}{Indeed, previous work using the same design as the present paper had also found semantic interference effects at the pre-critical region: both \cite{vandyke07} and \cite{mertzen} found such effects.
 \citeauthor{mertzen}'s (\citeyear{mertzen}) Figures 5 and 6 indicated reading time differences in earlier regions, especially for their German data (see their Figure 6).
 \cite{vandyke07} did not report reading times for the whole sentences / distractors;  \citeauthor{vandyke07} attributed the effects observed at the pre-critical region to plausibility differences between conditions, but as \cite{mertzen} also pointed out, encoding interference could be an explanation even in that study.
-Given these earlier findings, our results are consistent with the encoding interference explanation.
+Given these earlier findings, our results are consistent with the encoding interference explanation.}\label{previous_encoding}
 
@@
 The observed differences in the time course of the effect are likely due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-\Copy{thisdesign3}{Overall, \textcolor{blue}{in the present design}, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as was previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval}\label{thisdesign3}.
-A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
+\Copy{thisdesign3}{Overall, \textcolor{blue}{in the present design}, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as was previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.}\label{thisdesign3}  A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
 

### 2024-11-26 paper/revision2/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-The OSF repository has bee made public to ensure general accessibility. We have updated the link in the manuscript accordingly.
+The OSF repository has been made public to ensure general accessibility. We have updated the link in the manuscript accordingly.
 
@@
 As we had explained in the response to reviewers on August 29, 2024 (specifically in the response letter and our responses to the comments E.2 and R3.5), the priors were changed from directional priors in the original version to symmetrical priors in the revised version.
+This was necessary because we added trial id as a predictor as was suggested by a reviewer.
+There was no a priori hypothesis of the sign of the effect of trial id (do the participants get faster or slower over the course of the experiment?), therefore the prior for trial id needed to be symmetrical (positive and negative values).
+It is not possible to include symmetrical and directional priors within a single brms model \citep{brms}, thus all priors were changed to symmetrical ones.
+
 Additionally, we changed the method with which the Bayes factors were computed.
@@
 We have revised Figure 3 to remove the focus on the critical region (see below).
-At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference and that the long-lasting effect caused by that confounds the results at later effects.
-For comparability with previous findings, we then briefly report the later effects.
-We have revised the Results section in the following way, starting on page \pageref{revised_SPR}:
-
-\begin{quote}
-    \Paste{revised_SPR}
-\end{quote}
-
-We do not share the reviewer's interpretation that the effects in the later regions are not meaningful.
-We use the pre-critical reading times as a proxy to investigate the effect of encoding interference caused by the distractor.
-To emphasize this we have renamed the former section ``Pre-critical effects" to ``Reading times difference prior to retrieval".
-We have revised the beginning of the Discussion on page \pageref{prior_to_retrieval} accordingly:
-
-\begin{quote}
-    \Paste{SPR_discussion_revised}
-\end{quote}
+At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference.
+We also state that the results at the critical (and spill-over) region are confounded because it is not possible to disentangle the long-lasting encoding effect from retrieval effects potentially arising in the critical region.
 
@@
 \clearpage
+
+The reviewer seems to believe that our results are not publication-worthy because encoding interference confounded the reading times of all regions after the distractor.
+However, as we had already discussed in the results of the SPR results, the results of the two published studies using the present design \citep{vandyke07, mertzen} show indications that they were affected by encoding interference as well (see page \pageref{previous_encoding}):
+
+\begin{quote}
+    \Paste{previous_encoding}
+\end{quote}
+
+Therefore, our study replicates the encoding interference effect of the previous studies but it is the only study to prominently address that the reading times of the critical region are confounded by encoding interference.
+To ensure comparability to the previous studies, we believe that it is mandatory that we report the reading times effects in the same regions which have been reported by them: the pre-critical, critical and spill-over region -- despite the confound discussed above.
+We have revised the Results section to be more concise, starting on page \pageref{revised_SPR}:
+
+\begin{quote}
+    \Paste{revised_SPR}
+\end{quote}
+
+Because the finding that the self-paced reading times were primarily affected by semantic encoding interference is a central finding of our study, we think that it is appropriate that we investigated it.
+We did not analyze the distractor itself and the immediately following words because these were not identical across conditions.
+Instead, we used the reading times of the pre-critical region which was the same word across conditions (\textit{abends}, `in the evening', in the example item used for the x-axis labels in Figure 3 above) as a proxy to investigate the effect of encoding interference caused by the distractor.
+This pre-critical word preceded the critical verb, thus the reading times at the pre-critical word are unlikely to show retrieval interference.
+To emphasize this we have renamed the former section ``Pre-critical effects" to ``Reading times difference prior to retrieval".
+We have revised the beginning of the Discussion on page \pageref{prior_to_retrieval} accordingly:
+
+\begin{quote}
+    \Paste{SPR_discussion_revised}
+\end{quote}
 
@@
 \subsubsection*{Response to R2.4}
-The SPR results and the ERP results in the N400 window both show inhibitory interference, i.e., more processing difficulty under high vs.\ low semantic interference, manifesting in longer reading times and a more negative N400.
+The SPR results and the ERP results in the N400 window both showed inhibitory interference, i.e., more processing difficulty under high vs.\ low semantic interference, manifesting in longer reading times and a more negative N400.
 We explain below in our response to R2.5 that the reduced P600 for high interference might be explained by facilitatory interference because under high semantic interference whatever noun was retrieved can be easily integrated.
@@
 In general, SPR results cannot show the same complex results pattern that the ERP results showed because they provide just one measure per region.
-Additionally, only at the critical (or spill-over) region facilitated integration might be observed and in line with the reviewer's other comments, we do not interpret the reading times of the later regions because these were overshadowed by the reading times differences starting at the distractor.
+Additionally, only at the critical (or spill-over) region facilitated integration might be observed and in line with the reviewer's other comments, we do not interpret the reading times of the these regions because they were overshadowed by the reading times differences starting at the distractor.
 
@@
 This comment does not call for further action.
-In section 7.2 ``Encoding and retrieval interference", we discuss that we believe that the difference in presentation rate and whether it was controlled by the participants or not led to different types of interference (encoding vs.\ retrieval).
+In the General Discussion (subsection ``Encoding and retrieval interference"), we discuss that we believe that the difference in presentation rate and whether it was controlled by the participants or not led to different types of interference (encoding vs.\ retrieval).
 

### 2024-11-27 paper/revision2/ResponseToReviewers.tex
edited R21 response

@@
 At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference.
-We also state that the results at the critical (and spill-over) region are confounded because it is not possible to disentangle the long-lasting encoding effect from retrieval effects potentially arising in the critical region.
+We also state that the results at the critical (and spill-over) region are potentially confounded because it is not possible to disentangle the long-lasting encoding effect from retrieval effects potentially arising in the critical region.
 
@@
 
-The reviewer seems to believe that our results are not publication-worthy because encoding interference confounded the reading times of all regions after the distractor.
-However, as we had already discussed in the results of the SPR results, the results of the two published studies using the present design \citep{vandyke07, mertzen} show indications that they were affected by encoding interference as well (see page \pageref{previous_encoding}):
+The reviewer suggests that encoding interference confounded the reading times of all regions after the distractor and argues that this makes the results uninterpretable.
+The implication seems to be that the patterns observed at the pre-critical, critical, and spillover regions should not be reported.
+ However, as we had already discussed in the results of the SPR results, the two published studies that used the present design \citep{vandyke07, mertzen} show clear indications that these regions of interest were affected by encoding interference as well.
+Thus, across the three studies (Van Dyke et al., Mertzen et al., and ours) one consistent pattern is that the observed effects could be driven at least partly by encoding interference.
+We feel that this consistent pattern across studies in the three regions of interest is worth reporting.
+
+We discuss this on page \pageref{previous_encoding}:
 
@@
 
-Therefore, our study replicates the encoding interference effect of the previous studies but it is the only study to prominently address that the reading times of the critical region are confounded by encoding interference.
-To ensure comparability to the previous studies, we believe that it is mandatory that we report the reading times effects in the same regions which have been reported by them: the pre-critical, critical and spill-over region -- despite the confound discussed above.
+To summarize, one valuable finding in our study is the replication of the pattern found in the previous studies, which is attributable to encoding interference.
+Ours is the only study among these three to prominently raise the possibility that the reading times of the critical region may be confounded by encoding interference--the reviewer would agree with this discussion.
+To ensure comparability to the previous studies, we consider it very important to report the reading times effects in the same regions which the previous studies reported, namely: the pre-critical, critical and spill-over region.
+This seems necessary to us despite the possible confound discussed above that the reviewer justifiably raises.
 We have revised the Results section to be more concise, starting on page \pageref{revised_SPR}:

### 2024-11-27 paper/revision2/ResponseToReviewers.tex
edited R21 response

@@
 
-Because the finding that the self-paced reading times were primarily affected by semantic encoding interference is a central finding of our study, we think that it is appropriate that we investigated it.
+The main finding in our SPR study is that reading times were primarily affected by semantic encoding interference.
+In order to fully understand the effect of semantic interference and to relate it to the previous literature, the reader needs to see the analyses at the pre-critical, critical, and spillover regions.
 We did not analyze the distractor itself and the immediately following words because these were not identical across conditions.
-Instead, we used the reading times of the pre-critical region which was the same word across conditions (\textit{abends}, `in the evening', in the example item used for the x-axis labels in Figure 3 above) as a proxy to investigate the effect of encoding interference caused by the distractor.
+Instead, we start the analysis at the pre-critical region, which was the same word across conditions (\textit{abends}, `in the evening', in the example item used for the x-axis labels in Figure 3 above).
+This pre-critical region serves as a proxy for investigating the effect of encoding interference caused by the distractor.
 This pre-critical word preceded the critical verb, thus the reading times at the pre-critical word are unlikely to show retrieval interference.

### 2024-11-27 paper/revision2/elsarticle-template-harv.tex
Updated via the Open Science Framework

@@
 
-We turn next to the event-related potentials experiment.
+\textcolor{blue}{In summary, the main finding of the Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.
 
@@
 \Copy{gendisccaveat}{Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.}\label{gendisccaveat} \Copy{gendisccaveat2}{\textcolor{blue}{Moreover, it is entirely possible that a stronger syntactic manipulation than the one we used ends up showing a syntactic interference effect of the type that \citet{vandyke07} originally reported.
-Indeed, our own findings would need to be replicated, ideally by an independent research group, if we want to be sure that in the present design there is no evidence for syntactic interference.
-If such a replication attempt is carried out, it would also be useful to conduct an lab-based self-paced reading study to investigate our speculation earlier that the absence of syntactic interference in our online SPR study may have been due to participants adopting a good-enough processing strategy, leading to only semantic interference being detectable.}}\label{gendisccaveat2}
+Indeed, our own findings would need to be replicated, ideally by an independent research group, if we want to be sure that in the present design there is no evidence for syntactic interference.}}\label{gendisccaveat2}
 

### 2024-11-27 paper/revision2/ResponseToReviewers.tex
Updated via the Open Science Framework

@@
 
-Clarifying text has been added in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), and we have shortened the manuscript length by four pages.
+Clarifying text has been added in this second round of revisions (e.g., regarding the implications, the prolonged reading times starting at the distractor and the use of syntactic and semantic cues), and we have shortened the manuscript length by three pages.
 We have removed redundant text passages (e.g., in the description of the SPR results).
@@
 We have removed repetitions in the description of the statistical analyses and have shortened the description of the self-paced reading results (and the comparison to the previous studies).
-This reduced the length of the manuscript by four ages; after the first round of revisions the main text ended on page 77, now it ends on page 73.
+This reduced the length of the manuscript by three pages; after the first round of revisions the main text ended on page 77, now it ends on page 74.
 
@@
 The implication seems to be that the patterns observed at the pre-critical, critical, and spillover regions should not be reported.
- However, as we had already discussed in the results of the SPR results, the two published studies that used the present design \citep{vandyke07, mertzen} show clear indications that these regions of interest were affected by encoding interference as well.
+ However, as we had already reported in the discussion of the SPR results, the two published studies that used the present design \citep{vandyke07, mertzen} show clear indications that these regions of interest were affected by encoding interference as well.
 Thus, across the three studies (Van Dyke et al., Mertzen et al., and ours) one consistent pattern is that the observed effects could be driven at least partly by encoding interference.

### 2024-11-30 paper/revision2/elsarticle-template-harv.tex
minor edits

@@
 This difficulty in identifying the correct dependent is called similarity-based interference.
-We present large-sample self-paced reading and event-related potentials experiments using a well-established design to investigate interference due to syntactic and semantic cues in German.
+We present relatively large-sample self-paced reading and event-related potentials experiments using a well-established design to investigate interference due to syntactic and semantic cues in German.
 \textcolor{blue}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
@@
 Surprisingly, Bayes factors provided evidence against interference due to \textcolor{blue}{the syntactic cue \{+grammatical subject\} in this particular design in both experiments}.
-This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assume that both syntactic and semantic cues play an equal role in retrieval.
+This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assumes that both syntactic and semantic cues play an equal role in retrieval.
 We show through computational modeling that cue-based retrieval \textcolor{blue}{will} also shows no syntactic interference \textcolor{blue}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in.
-Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the predictions of the cue-based retrieval model would be consistent with the observed patterns in our data.
+Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the cue-based retrieval model would exhibit patterns consistent with the observed patterns in our data.
 }
@@
 Because the distractor differed between conditions, we did not analyze distractor reading times.
-Instead, we treat the pre-critical region which was identical across conditions and preceded the critical verb as a proxy to investigate the difference starting at the distractor.
+Instead, to investigate the difference starting after the distractor was read, we investigate the pre-critical region, which was identical across conditions and preceded the critical verb.
+The pre-critical region serves as a proxy for any encoding interference effects that arose after the distractor was read.
 Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region.
@@
 
-\Copy{thisdesign2}{Without the modeling results reported above, it would be easy to conclude that our data show that syntactic cues play no role  \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations}}\label{thisdesign2}.
-However, this conclusion would be problematic.
+\Copy{thisdesign2}{Without the modeling results reported above, it would be easy to conclude that, \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations},  our data show that syntactic cues play no role.}\label{thisdesign2} However, this conclusion would be problematic.
 First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}.

### 2024-11-30 paper/revision2/ResponseToReviewers.tex
minor edits

@@
 The implication seems to be that the patterns observed at the pre-critical, critical, and spillover regions should not be reported.
- However, as we had already reported in the discussion of the SPR results, the two published studies that used the present design \citep{vandyke07, mertzen} show clear indications that these regions of interest were affected by encoding interference as well.
+ However, as we had already reported in the discussion of the SPR results, the two published studies that used the present design \citep{vandyke07, mertzen} show clear indications that these regions of interest were affected as well, possibly by encoding interference.
 Thus, across the three studies (Van Dyke et al., Mertzen et al., and ours) one consistent pattern is that the observed effects could be driven at least partly by encoding interference.
@@
 To summarize, one valuable finding in our study is the replication of the pattern found in the previous studies, which is attributable to encoding interference.
-Ours is the only study among these three to prominently raise the possibility that the reading times of the critical region may be confounded by encoding interference--the reviewer would agree with this discussion.
+Ours is the only study among these three to prominently raise the possibility that the reading times of the critical region may be confounded by encoding interference--we believe the reviewer would agree with this discussion.
 To ensure comparability to the previous studies, we consider it very important to report the reading times effects in the same regions which the previous studies reported, namely: the pre-critical, critical and spill-over region.
 This seems necessary to us despite the possible confound discussed above that the reviewer justifiably raises.
+
 We have revised the Results section to be more concise, starting on page \pageref{revised_SPR}:

### 2024-12-01 paper/revision2/elsarticle-template-harv.tex
final edits

@@
 \textcolor{blue}{However, reading times started to differ before this critical dependency could be formed.
-Starting at the distractor which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors).
+Starting at the distractor, which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors).
 Because the distractor differed between conditions, we did not analyze distractor reading times.
@@
 \textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.
-T}he most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above).
+T}he most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.
 The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.

### 2024-12-02 code_data/SPR/r_scripts/Plot_ReadingTimes.R
Updated via the Open Science Framework

@@ -45,27 +45,27 @@ df_losyn <- df_trim %>% filter(condition == "a" | condition == "b")
 #### Plot whole sentence ####
 
 #calculate mean reading time per subject per word per condition
-hisyn_by_participants <- dplyr::summarize(
-  dplyr::group_by(
+hisyn_by_participants <- summarize(
+  group_by(
     filter(df_hisyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "21")),
     wordno,condition,subj, exp),
   logrt=mean(log(rt)))
 
-losyn_by_participants <- dplyr::summarize(
-  dplyr::group_by(
+losyn_by_participants <- summarize(
+  group_by(
     filter(df_losyn,wordno %in% c("0", "1", "2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "20")),
     wordno,condition,subj, exp),
   logrt=mean(log(rt)))
 
 # calculate mean per word per condition (averaged over subjects)
-hisyn_summ_by_participants <- dplyr::summarize(
-  dplyr::group_by(hisyn_by_participants,wordno,condition),
+hisyn_summ_by_participants <- summarize(
+  group_by(hisyn_by_participants,wordno,condition),
   meanlogrt=mean(logrt), 
   SE = sd(logrt)/sqrt(nrow(hisyn_by_participants)), 
   low2SE= meanlogrt - 2*SE, high2SE= meanlogrt + 2*SE)
 
-losyn_summ_by_participants <- dplyr::summarize(
-  dplyr::group_by(losyn_by_participants,wordno,condition),
+losyn_summ_by_participants <- summarize(
+  group_by(losyn_by_participants,wordno,condition),
   meanlogrt=mean(logrt), 
   SE = sd(logrt)/sqrt(nrow(losyn_by_participants)), 
   low2SE= meanlogrt - 2*SE, high2SE= meanlogrt + 2*SE)
@@ -151,7 +151,7 @@ p_zoom <- ggplot(data=zoom, aes(x=fake_wordno, y=meanlogrt,
                                 shape=syntactic.interference,
                                 linetype=semantic.interference,
                                 ymin=low2SE,ymax=high2SE)) +
-                geom_point(aes(group = interaction(syntactic.interference, semantic.interference)), size=3, position= position_dodge(0.4)) + 
[…]

### 2024-12-02 paper/revision2/elsarticle-template-harv.tex
Updated via the Open Science Framework

@@
 For a principled approach to feature selection using word embeddings, see \cite{smith2020principled}.
-A plausibility norming experiment (see page \pageref{plausib_normin_orig}) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
+A plausibility norming experiment (see below) was conducted to ensure that the semantic manipulation did work, i.e., that the semantic requirements of the verb matched the features of the subject and the animate distractor, but not the features of the inanimate distractor.
 }\label{semantic_features_encoding}
@@
 
-Alternately, retrieval interference -- especially syntactic retrieval interference -- might modulate the P600.
+Alternatively, retrieval interference -- especially syntactic retrieval interference -- might modulate the P600.
 The P600 is a positive deflection in the ERP wave form that is typically observed at least 600 ms after stimulus onset.
@@
 The SPR study, hereafter Experiment 1a, was initially designed so that participants read \textcolor{blue}{half of the items (1-60) in the first session and the other half of the items (61-120) in the second session on another day.} The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
-The adaptation effect is theoretically interesting per se \citep{fine2013rapid} and we discuss it below; but it was problematic for our design because the average effect would then be confounded by adaptation.
+The adaptation effect is theoretically interesting per se \citep{fine2013rapid}; but it was problematic for our design because the average effect would then be confounded by adaptation.
 For this reason, we ran a second version of the SPR study, hereafter Experiment 1b, in which participants (N=570) were either shown items 1-60 or items 61-120 (but not both).
@@
 {\Copy{fillers}{\textcolor{blue}{The fillers were included to mask the critical manipulation.
-They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{blue}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.
+They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{blue}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich, während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom.jpg}
+    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled.jpg}
 \end{sidewaysfigure}
@@
 It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
-\textcolor{blue}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006}.
-Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session.
+\Copy{smith1}{\textcolor{blue}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006, smith2021encoding}}}\label{smith1}.
+\textcolor{blue}{Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session.
 As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.
@@
 
-\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.
-T}he most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.
-The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
+\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.} \Copy{smith2}{The most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.}\label{smith2} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
 
@@
 
-\textcolor{blue}{In summary, the main finding of the Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.
+\textcolor{blue}{In summary, the main finding of Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.
 

### 2024-12-02 paper/revision2/ResponseToReviewers.tex
minor edits
Updated via the Open Science Framework

@@
 \subsubsection*{Response to E.3}
-Yes, we changed all analyses in the first revision.
-As we had explained in the response to reviewers on August 29, 2024 (specifically in the response letter and our responses to the comments E.2 and R3.5), the priors were changed from directional priors in the original version to symmetrical priors in the revised version.
-This was necessary because we added trial id as a predictor as was suggested by a reviewer.
-There was no a priori hypothesis of the sign of the effect of trial id (do the participants get faster or slower over the course of the experiment?), therefore the prior for trial id needed to be symmetrical (positive and negative values).
-It is not possible to include symmetrical and directional priors within a single brms model \citep{brms}, thus all priors were changed to symmetrical ones.
+Yes, as we had explained in the response to reviewers on August 29, 2024, all analyses were changed in the first round of revisions.
+For a fair comparison of our results to the results of the previous studies, especially Van Dyke (2007), it was necessary to compute Bayes factors using the Savage-Dickey ratio (in the original version of the manuscript we had used bridge sampling).
+Furthermore, for the Van Dyke (2007) data, the first on this design, it only made sense to use symmetrical priors which are agnostic regarding the sign of the effects.
+For consistency, we then used symmetrical priors for all our analyses (compared to the directional priors which we had used originally).
 
-Additionally, we changed the method with which the Bayes factors were computed.
-Initially, we had used bridge sampling, but in the revision, we switched to the Savage-Dickey method because only using that method we were able to compare models with random slopes (this was also explained in the previous response letter, Response to R1.7).
 These changes did not change the results qualitatively, but led to some numerical changes which probably piqued the reviewer's interest.
@@
 
-So, in summary, a proper treatment of encoding interference would require a modeling effort at the same scale as was carried out in the above-mentioned paper.
+Another interesting model of encoding interference is
+
+Smith, Garrett, Julie Franck, and Whitney Tabor. ``Encoding interference effects support self-organized sentence processing." Cognitive Psychology 124 (2021): 101356.
+
+We have now cited this paper on pages \pageref{revised_SPR} and \pageref{smith2}:
+
+\Paste{smith1}
+\\
+
+\Paste{smith2}\\
+
+So, in summary, a proper treatment of encoding interference would require a modeling effort at the same scale as was carried out in \citet{Yadavetal2022}.
 Such a modeling effort, although very worthwhile, would take several years' of work and is therefore far beyond the scope of the present work.

### 2024-12-16 paper/revision2/elsarticle-template-harv.tex
Updated via the Open Science Framework

@@
-%% 
-%% Copyright 2007-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
-
-\documentclass[review,preprint,12pt,authoryear,floatsintext]{elsarticle}
+\documentclass[12pt,authoryear,floatsintext]{elsarticle}
 
@@
 \begin{frontmatter}
-\title{Do syntactic and semantic similarity lead to interference effects?\\
-Evidence from self-paced reading and event-related potentials using German}
+\title{Do syntactic and semantic similarity lead to interference effects? Evidence from self-paced reading and event-related potentials using German}
+
+\author[add1]{Pia Schoknecht\corref{cor1}}
+\ead{pia.schoknecht@uni-potsdam.de}
+\author[add2]{Himanshu Yadav}
+\author[add1]{Shravan Vasishth}
+
+\cortext[cor1]{Corresponding author}
+
+\address[add1]{University of Potsdam, Department of Linguistics, Karl-Liebknecht-Straße 24-25, 14476 Potsdam, Germany}
+\address[add2]{Indian Institute of Technology Kanpur, Department of Cognitive Science, Kanpur 208016, India}
+
 %% Title, authors and addresses
@@
 Cue-based retrieval accounts of sentence processing postulate that at a verb, retrieval cues are generated to complete a dependency with the verb's argument(s); for example, the dependency between the subject and the verb must be completed.
-If these retrieval cues match with not only the \textcolor{blue}{subject} but also with those on other nouns in the sentence, then processing difficulty arises at the verb.
+If these retrieval cues match with not only the \textcolor{black}{subject} but also with those on other nouns in the sentence, then processing difficulty arises at the verb.
 This difficulty in identifying the correct dependent is called similarity-based interference.
 We present relatively large-sample self-paced reading and event-related potentials experiments using a well-established design to investigate interference due to syntactic and semantic cues in German.
-\textcolor{blue}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
-\textcolor{blue}{Bayes factors analyses showed} evidence for a semantic interference effect \textcolor{blue}{in both experiments}. 
-Surprisingly, Bayes factors provided evidence against interference due to \textcolor{blue}{the syntactic cue \{+grammatical subject\} in this particular design in both experiments}.
+\textcolor{black}{In this design, the syntactic cue \{+subject\} and the semantic cue \{+animate\} are manipulated.}
+\textcolor{black}{Bayes factors analyses showed} evidence for a semantic interference effect \textcolor{black}{in both experiments}. 
+Surprisingly, Bayes factors provided evidence against interference due to \textcolor{black}{the syntactic cue \{+grammatical subject\} in this particular design in both experiments}.
 This finding contradicts the predictions of the standard implementations of cue-based retrieval theory, which (implicitly) assumes that both syntactic and semantic cues play an equal role in retrieval.
-We show through computational modeling that cue-based retrieval \textcolor{blue}{will} also shows no syntactic interference \textcolor{blue}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in.
+We show through computational modeling that cue-based retrieval \textcolor{black}{will} also show no syntactic interference \textcolor{black}{in the present design} if the parser is assumed to keep track of which clause the subject occurs in.
 Thus, if syntactic retrieval cues include hierarchical syntactic information (is the noun in the same clause as the verb?), the cue-based retrieval model would exhibit patterns consistent with the observed patterns in our data.
@@
 
-The SPR study, hereafter Experiment 1a, was initially designed so that participants read \textcolor{blue}{half of the items (1-60) in the first session and the other half of the items (61-120) in the second session on another day.} The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
+The SPR study, hereafter Experiment 1a, was initially designed so that participants read \textcolor{black}{half of the items (1-60) in the first session and the other half of the items (61-120) in the second session on another day.} The second session showed an adaptation effect \citep{prasad2021rapid}: Reading times decreased overall.
 The adaptation effect is theoretically interesting per se \citep{fine2013rapid}; but it was problematic for our design because the average effect would then be confounded by adaptation.
@@
 In session 2, the experimental items 61-120 were presented with another set of 40 filler sentences.
-{\Copy{fillers}{\textcolor{blue}{The fillers were included to mask the critical manipulation.
-They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{blue}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich, während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.
+{\Copy{fillers}{\textcolor{black}{The fillers were included to mask the critical manipulation.
+They} were less syntactically complex than the experimental items but had at least one embedded clause and generally provided more variety\textcolor{black}{, e.g., by including modifiers and different types of subordinate clauses} (e.g., ``Der Teppichmacher, der früh in die Werkstatt gekommen war, reparierte den besonders schönen, alten Teppich, während er die Nachrichten hörte'', `The carpet maker who came to the workshop early repaired the especially beautiful old carpet while listening to the news.'; ``Der Jugendliche war genervt, weil seine Freundin, die nicht studieren wollte, oft die Schule schwänzte, um Computer zu spielen.'', `The teenager was annoyed because his girlfriend who did not want to go to University  skipped school frequently to play video games.').}}\label{fillers} Each session lasted approximately 25 minutes.
 
@@
 
-\Copy{trim}{{\textcolor{blue}{Self-paced reading times below 150 ms and above 3000 ms were excluded before analysis because these overly short or long reaction times were likely caused by inattentively performing the task.
+\Copy{trim}{{\textcolor{black}{Self-paced reading times below 150 ms and above 3000 ms were excluded before analysis because these overly short or long reaction times were likely caused by inattentively performing the task.
 The procedure excluded 4.9\,\% of the data.}}}\label{trim} Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}.
@@
 
-Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{blue}{(panel A)} and low \textcolor{blue}{(panel B)} syntactic interference because these conditions had partially different sentence structures.
+Reading times across the whole sentence are shown in Figure \ref{fig:whole_sentence}, separately for high \textcolor{black}{(panel A)} and low \textcolor{black}{(panel B)} syntactic interference because these conditions had partially different sentence structures.
 
@@
 Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure.
-\textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
+\textcolor{black}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
     \label{fig:whole_sentence}
@@
 It is apparent from Figure \ref{fig:whole_sentence} that, regardless of the syntactic manipulation, at the distractor, reading times between high and low semantic interference started to differ: distractors in the high semantic interference conditions induced longer reading times than in the low semantic interference conditions.
-\Copy{smith1}{\textcolor{blue}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006, smith2021encoding}}}\label{smith1}.
-\textcolor{blue}{Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session.
+\Copy{smith1}{\textcolor{black}{To anticipate our discussion of the SPR results here, we believe that the difference in reading times starting at the distractor was caused by encoding interference \citep{Oberauer_Kliegl_2006, smith2021encoding}}}\label{smith1}.
+\textcolor{black}{Additionally, the second session in Experiment 1a} showed shorter reading times compared to the first session.
 As mentioned earlier, this attenuation in reading times was likely due to adaptation to the SPR task in the second session.
 
-\textcolor{blue}{Crucially, the reading times difference starting at the distractor} persisted in the following regions.
-The distractor and the immediately following regions differed between conditions, \textcolor{blue}{therefore we did not analyze their reading times.
+\textcolor{black}{Crucially, the reading times difference starting at the distractor} persisted in the following regions.
+The distractor and the immediately following regions differed between conditions, \textcolor{black}{therefore we did not analyze their reading times.
 In our materials, there were two words between the distractor and the critical verb which were identical across conditions: the pre-pre-critical and pre-critical adverbs.
@@
 
-Figure \ref{fig:spr_posteriors} shows that \textcolor{blue}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions.
+Figure \ref{fig:spr_posteriors} shows that \textcolor{black}{the posterior distributions for the parameters are very similar in the pre-critical, critical and spill-over regions.
 The estimates of semantic interference were positive in all regions and the estimates of the interaction were mostly positive in all regions.
@@
 \paragraph{Hypothesis testing using Bayes factors}
-\textcolor{blue}{To complete the analysis of the effects in the pre-critical, critical and spill-over regions -- although they were confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
+\textcolor{black}{To complete the analysis of the effects in the pre-critical, critical and spill-over regions -- although they were confounded by the long-lasting reading time differences starting at the distractor, we here present a Bayes factor analysis using all the data from Experiments 1a and 1b (see Figure \ref{fig:spr_bfs}).
 There was} very strong to extreme evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231).
-In the critical region, the evidence for semantic interference was weak \textcolor{blue}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5, Normal(0, 0.05): BF$_{10}$ = 1.3, Normal(0, 0.1): BF$_{10}$ = 0.7).
+In the critical region, the evidence for semantic interference was weak \textcolor{black}{at best} (Normal(0, 0.01): BF$_{10}$ = 3.5, Normal(0, 0.05): BF$_{10}$ = 1.3, Normal(0, 0.1): BF$_{10}$ = 0.7).
 There was evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ $<$ 0.5).
@@
 \subsection*{Discussion}
-\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation.
-\textcolor{blue}{However, reading times started to differ before this critical dependency could be formed.
+\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{black}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation.
+\textcolor{black}{However, reading times started to differ before this critical dependency could be formed.
 Starting at the distractor, which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors).
@@
 
-\subsubsection*{\textcolor{blue}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}
-
-\textcolor{blue}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence}).
+\subsubsection*{\textcolor{black}{Reading times differences prior to retrieval}}\label{prior_to_retrieval}
+
+\textcolor{black}{Animate distractors led to longer reading times than inanimate distractors} (see Figure \ref{fig:whole_sentence}).
 One might argue that this could be due to differences in word length.
@@
 
-Another possible explanation for the reading time differences starting at the distractor could be \textcolor{blue}{that the distractors induced} a difference in plausibility between the high and low semantic interference conditions.
+Another possible explanation for the reading time differences starting at the distractor could be \textcolor{black}{that the distractors induced} a difference in plausibility between the high and low semantic interference conditions.
 To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \citep{pcibex}.
@@
 
-To investigate whether this plausibility difference was driving the \textcolor{blue}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
+To investigate whether this plausibility difference was driving the \textcolor{black}{reading time difference between the semantic interference conditions}, we ran a Bayesian linear mixed model with log-normal likelihood on the reading times of the pre-critical region.
 This model included fixed effects for semantic interference (high +0.5, low -0.5), trial id (recoded to span from 0 to 1) and the centered plausibility ratings.
@@
 The interaction of semantic interference and plausibility was centered around zero.
-In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference \textcolor{blue}{starting at the distractor and lasting into later regions} was partially caused by reduced plausibility of the high semantic interference conditions.
-However, the plausibility difference was not the only driver of the reading times slowdown, i.e., \textcolor{blue}{high} semantic interference led to slower reading times independent of plausibility.
+In conclusion, the slowdown in the reading times results for high vs.\ low semantic interference \textcolor{black}{starting at the distractor and lasting into later regions} was partially caused by reduced plausibility of the high semantic interference conditions.
+However, the plausibility difference was not the only driver of the reading times slowdown, i.e., \textcolor{black}{high} semantic interference led to slower reading times independent of plausibility.
 
@@
 
-\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.} \Copy{smith2}{The most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.}\label{smith2} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
+\textcolor{black}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference.} \Copy{smith2}{The most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}.}\label{smith2} The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
 
@@
 \cite{mertzen} observed both syntactic and semantic interference effects in the pre-critical region.
-{\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{blue}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
+{\Copy{confound}{In addition to encoding interference, \cite{mertzen} proposed three other alternative explanations for their data: parafoveal-on-foveal effects, sentence structure confounds across conditions \textcolor{black}{(two vs.\ one embedded clause between subject and verb in the high vs.\ low syntactic interference conditions)}, and predictive processing effects.}}\label{confound} Our results cannot be explained by the parafoveal-on-foveal explanation, because there is no parafoveal preview in self-paced reading.
 
@@
 Consequently, the prediction explanation is also ruled out in our data, rendering encoding interference the most likely explanation of the effects found prior to the critical verb in our data.
-Crucially and as mentioned above, \textcolor{blue}{due to the long-lasting reading time difference starting at the distractor} it is not possible to pinpoint \textcolor{blue}{the driver of effects observed in later regions.
-The effects in the critical region (and spill-over region) might be} due to a retrieval initiated at the respective region, or still \textcolor{blue}{be} due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.
-
-\textcolor{blue}{In summary, the main finding of Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.
+Crucially and as mentioned above, \textcolor{black}{due to the long-lasting reading time difference starting at the distractor} it is not possible to pinpoint \textcolor{black}{the driver of effects observed in later regions.
+The effects in the critical region (and spill-over region) might be} due to a retrieval initiated at the respective region, or still \textcolor{black}{be} due to encoding and maintaining the distractor in memory \citep{ness2017}, or a combination of encoding and retrieval interference \citep{Yadavetal2022}.
+
+\textcolor{black}{In summary, the main finding of Experiment 1 is that the reading times were primarily affected by semantic encoding interference.} We turn next to the event-related potentials experiment.
 
@@
 \Copy{inhibition_facilitation}{{The semantic interference effect on the P600 is in the opposite direction than expected: High semantic interference led to a reduced P600 amplitude compared to low semantic interference.
-\textcolor{blue}{We provide a speculative interpretation of this unexpected result.} Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}.
+\textcolor{black}{We provide a speculative interpretation of this unexpected result.} Typically, the P600 amplitude is more positive when syntactic processing is complex, e.g., due to reanalysis \citep{osterhout&holcomb_1992}.
 The reduced P600 amplitude under high semantic interference in the present study might be explained by facilitatory interference \citep[see e.g.,][]{jaeger_etal_2017}.
-\textcolor{blue}{We are not aware of any other study that found inhibitory as well facilitatory interference in the same sentence configuration.
+\textcolor{black}{We are not aware of any other study that found inhibitory as well facilitatory interference in the same sentence configuration.
 Facilitatory interference is typically only found in ungrammatical sentences.
 Nevertheless, we think that the reduced P600 amplitude for high vs.\ low semantic interference in our data could be explained by facilitatory interference.} In the present design, subject-verb integration might be facilitated, i.e., easier, under high semantic interference because there are two semantically suitable candidates to function as the subject.
-\textcolor{blue}{Even if the wrong noun, i.e., the distractor, was retrieved, it could be easily integrated with the verb.
+\textcolor{black}{Even if the wrong noun, i.e., the distractor, was retrieved, it could be easily integrated with the verb.
 In contrast, misretrievals of the distractor in the low semantic interference conditions which might happen due to noisy processing would require reanalysis because the distractor did not match the semantic requirements of the verb.} A similar effect was found by \cite{Tanner_etal_2017}.
@@
 However, the present study investigated only grammatical subject-verb integration and the size of the P600 effect was rather small.
-The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was \textcolor{blue}{misretrieved.
+The small size of the effect on average could indicate that the facilitation occurred only occasionally when the distractor was \textcolor{black}{misretrieved.
 Thus semantic cue overload would render memory retrieval more difficult causing a more negative N400 and if the distractor is falsely retrieved, integration with the verb would be easy resulting in a reduced, i.e., more negative, P600.
@@
 Recently, \cite{mertzen} presented simulations for the present design that predicted approximately equally sized effects of syntactic and semantic interference in the reading times of the critical verb.
-We found that the reading times at the critical verb were confounded by \textcolor{blue}{reading times differences starting earlier in the sentence}, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
+We found that the reading times at the critical verb were confounded by \textcolor{black}{reading times differences starting earlier in the sentence}, which were best explained by \textit{encoding} interference \citep[see also][]{mertzen}.
 Because the Lewis and Vasishth (2005) model is a model of retrieval interference and not encoding interference, and because we cannot disentangle encoding and retrieval interference in the present design, it would not make sense to compare the results from the present SPR experiment to the model's predictions.
@@
 In the present case, it is possible that the syntactically relevant noun is only the one in the same clause as the verb.
-{\Copy{cite_kush}{For more discussion of the role of structural retrieval cues, see \textcolor{blue}{\cite{kush2015relation}}, \cite{franck2020hierarchical} and \cite{arnett2017subject}.}}\label{cite_kush} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
+{\Copy{cite_kush}{For more discussion of the role of structural retrieval cues, see \textcolor{black}{\cite{kush2015relation}}, \cite{franck2020hierarchical} and \cite{arnett2017subject}.}}\label{cite_kush} The proposed model here using two cues predicts no syntactic interference and no interaction, but does predict semantic interference effects.
 Lastly, we implemented a model with one cue which unambiguously matches the subject.
@@
 If the model-generated data $x_{i} \sim Model(F_i)$ for $F_i$ is close to the actual data $y$, then $F_i$ has higher likelihood and vice versa.
-\Copy{modeling}{The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, \textcolor{blue}{where the function $\Psi(d, 0, \delta)$} represents a density kernel that weights a proposal $F_i$ based on the \textcolor{blue}{distance $d$ between model-generated data $y$ and actual data $x_i$}.}\label{modeling} We chose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
+\Copy{modeling}{The likelihood term can be rewritten as $\mathcal{L}(F_i|y) = \Psi(d(x_i, y), 0 , \delta)$, \textcolor{black}{where the function $\Psi(d, 0, \delta)$} represents a density kernel that weights a proposal $F_i$ based on the \textcolor{black}{distance $d$ between model-generated data $y$ and actual data $x_i$}.}\label{modeling} We chose a Gaussian kernel, $Normal(0,\delta)$ where $\delta=0.01$, such that the distance between model-generated data and observed data $d(x_i,y)$ is assumed to come from the distribution $Normal(0,0.01)$.
 The $\delta$ parameter determines how good the approximation is: the lower the value of $\delta$, the better the approximation of the likelihood.
@@
 The Bayes factors under different priors on $F$ are shown in Figure \ref{fig:BF_compmods}.
-The Bayes factors indicate strong evidence for the new two-cues model \textcolor{blue}{compared to the one-cue model (null model)}; this comparison shows that there is strong evidence for interference effects.
-Compared to the three-cues model proposed by \cite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior \textcolor{blue}{(see Table \ref{tab:bf_interpretation} for the interpretation of Bayes factors)}.
+The Bayes factors indicate strong evidence for the new two-cues model \textcolor{black}{compared to the one-cue model (null model)}; this comparison shows that there is strong evidence for interference effects.
+Compared to the three-cues model proposed by \cite{mertzen}, the Bayes factors provide either no evidence for the two-cues model or anecdotal to moderate evidence for the new two-cues model depending on the prior \textcolor{black}{(see Table \ref{tab:bf_interpretation} for the interpretation of Bayes factors)}.
 This model comparison shows that there is some support for the idea that the relevant syntactic cue is the \{$\pm$subject-in-same-clause\} cue – rather than the two separate, previously assumed cues \{$\pm$grammatical subject\} and \{$\pm$same clause\}.
-This result holds for prior assumptions that allow higher probability mass for small values of the scaling parameter $F$., i.e., for $F<0.01$.
+This result holds for prior assumptions that allow higher probability mass for small values of the scaling parameter $F$, i.e., for $F<0.01$.
 
@@
 The modeling reported here suggests that one way to explain the observed pattern in the context of cue-based retrieval is to assume the following: the parser searches for a subject that is within the same clause, but uses the animacy cue without reference to the clause that a noun appears in.
-\textcolor{blue}{We revisit the connection between syntax and semantics below.} A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
+\textcolor{black}{We revisit the connection between syntax and semantics below.} A broader implication is that the human sentence comprehension system may be using the hierarchical syntactic structure to selectively target nouns for retrieval; this is an idea that has independent support in the literature \citep[e.g.,][]{Sturt2003,dillon2013,yadav2021individual}.
 
@@
 Then we discuss two important issues that relate to the broader literature.
-The first is the role of encoding vs.\ retrieval interference, and the second is \textcolor{blue}{the use of syntactic and semantic information during sentence processing}.
+The first is the role of encoding vs.\ retrieval interference, and the second is \textcolor{black}{the use of syntactic and semantic information during sentence processing}.
 
@@
 In comparison to the previous estimates, the effects of the present study were relatively small and associated with less uncertainty in general.
-\textcolor{blue}{We used Bayes factors to compare whether there was evidence for the estimated effects in the previous studies vs.\ the present study.} For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
-\textcolor{blue}{Bayes factors for \cite{vandyke07} were computed using} the Normal-Normal conjugate case \citep{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms).
+\textcolor{black}{We used Bayes factors to compare whether there was evidence for the estimated effects in the previous studies vs.\ the present study.} For maximal comparability, we used only the self-paced reading times and total fixation times from the previous studies for this comparison.
+\textcolor{black}{Bayes factors for \cite{vandyke07} were computed using} the Normal-Normal conjugate case \citep{lee2012bayesian} to derive posteriors from the summary statistics (mean and SE) in the published paper and a reasonable prior (Normal(0,0.5), which amounts to an a priori assumed effect size in the range of -40 ms to 40 ms).
 Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed from hierarchical mixed models which we ran on their data analogously to our own analyses described in Statistical Analyses for Experiments 1a and b.
 
-\textcolor{blue}{The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.
+\textcolor{black}{The Bayes factors for the effects of interest of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.
 The region-wise Bayes factors across studies were consistent with only a few exceptions (e.g., the majority of studies showed no evidence for the interaction), but in some cases there was no clear clustering in favor or against an effect (e.g., regarding semantic interference in the spill-over region).
@@
 explanation for the lack of syntactic interference in our data compared to the previous studies.
-\Copy{SPR_goodenough}{\textcolor{blue}{However, due to the demanding task comprehenders might trade off processing depth and adopt good-enough processing \citep{ferreira2007goodenough}.
+\Copy{SPR_goodenough}{\textcolor{black}{However, due to the demanding task comprehenders might trade off processing depth and adopt good-enough processing \citep{ferreira2007goodenough}.
 Consequently, they would not fully parse embedded structures, hence there would be no syntactic interference from embedded subjects.
@@
 So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment.
-\Copy{thisdesign1}{Nevertheless, \textcolor{blue}{in the present design,} both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.}\label{thisdesign1}
+\Copy{thisdesign1}{Nevertheless, \textcolor{black}{in the present design,} both the SPR and the EEG experiment showed predominantly evidence against syntactic interference.}\label{thisdesign1}
 
@@
 While this seems like a plausible explanation for the previous effects, it would also apply to our pre-pre-critical region.
-However, Figure \ref{fig:whole_sentence} \textcolor{blue}{C} shows that the pre-pre-critical region in our items did not show a syntactic effect.
+However, Figure \ref{fig:whole_sentence} \textcolor{black}{C} shows that the pre-pre-critical region in our items did not show a syntactic effect.
 
@@
 
-\subsection*{\textcolor{blue}{The use of syntactic and semantic information}}
-
-\label{synsem}\Copy{synsem}{\Copy{thisdesign4}{\textcolor{blue}{In the sentence configurations that were investigated in the present study, we find a remarkable disconnect in the use of syntactic and semantic cues during subject-verb dependency resolution.}} The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
+\subsection*{\textcolor{black}{The use of syntactic and semantic information}}
+
+\label{synsem}\Copy{synsem}{\Copy{thisdesign4}{\textcolor{black}{In the sentence configurations that were investigated in the present study, we find a remarkable disconnect in the use of syntactic and semantic cues during subject-verb dependency resolution.}} The lack of decisive evidence for a main effect of syntactic interference for subject-verb dependency resolution in our study is not consistent with the assumption in previous work, e.g., \cite{vandyke07,mertzen}, that the parser searches for a subject simply by setting the retrieval cue \{$\pm$ grammatical subject\}.
 A future avenue of research could be to find out what an appropriate syntactic manipulation would be to consistently trigger syntactic interference.
@@
 
-\textcolor{blue}{Although the syntactic cue included hierarchical information to match only with the correct subject within the same clause, this did not block the semantic cue from matching with the embedded distractor.
+\textcolor{black}{Although the syntactic cue included hierarchical information to match only with the correct subject within the same clause, this did not block the semantic cue from matching with the embedded distractor.
 This finding is not consistent with syntax-first models like garden-path theory \citep{frazier1987} which assume that syntactic processing precedes all other levels of linguistic evaluation (e.g., plausibility).
-Our findings rather suggest parallel, fully independent matching of different retrieval cues with the candidates in memory.}  In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model \textcolor{blue}{assuming independent cues.}
-
-The semantic interference effect we found can be easily reconciled with \textcolor{blue}{other sentence processing theories that do not assign a special status to syntactic processing}.
+Our findings rather suggest parallel, fully independent matching of different retrieval cues with the candidates in memory.}  In the modeling section above, we have already discussed how our results relate to predictions from the \citet{Lewis2005} cue-based retrieval model \textcolor{black}{assuming independent cues.}
+
+The semantic interference effect we found can be easily reconciled with \textcolor{black}{other sentence processing theories that do not assign a special status to syntactic processing}.
 Our results could be seen as consistent with the good-enough processing account \citep{ferreira2007goodenough}, which assumes that comprehenders do not always aim to build a fully fleshed-out analysis of a sentence, but instead might accept incomplete, underspecified, or even incorrect analyses.
@@
 
-\Copy{thisdesign2}{Without the modeling results reported above, it would be easy to conclude that, \textcolor{blue}{during subject-verb dependency resolution in the investigated sentence configurations},  our data show that syntactic cues play no role.}\label{thisdesign2} However, this conclusion would be problematic.
+\Copy{thisdesign2}{Without the modeling results reported above, it would be easy to conclude that, \textcolor{black}{during subject-verb dependency resolution in the investigated sentence configurations},  our data show that syntactic cues play no role.}\label{thisdesign2} However, this conclusion would be problematic.
 First, as we showed above, the cue-based retrieval model suggests that -- at least in our data -- the \{$\pm$ grammatical subject\} cue is used, but differently than previously assumed: the relevant cue may be \{$\pm$ subject-in-same-clause\}.
@@
 Regarding agreement attraction, there seems to be at least one language with rich case marking (Czech) that does not show convincing evidence of number agreement attraction \citep{chromy2023number}.
-\Copy{gendisccaveat}{Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.}\label{gendisccaveat} \Copy{gendisccaveat2}{\textcolor{blue}{Moreover, it is entirely possible that a stronger syntactic manipulation than the one we used ends up showing a syntactic interference effect of the type that \citet{vandyke07} originally reported.
+\Copy{gendisccaveat}{Thus, it seems that, cross-linguistically, syntax does generally play a central role in building incremental structure and in completing dependencies.}\label{gendisccaveat} \Copy{gendisccaveat2}{\textcolor{black}{Moreover, it is entirely possible that a stronger syntactic manipulation than the one we used ends up showing a syntactic interference effect of the type that \citet{vandyke07} originally reported.
 Indeed, our own findings would need to be replicated, ideally by an independent research group, if we want to be sure that in the present design there is no evidence for syntactic interference.}}\label{gendisccaveat2}
@@
 Both our experiments consistently showed semantic interference, i.e., more processing difficulty in the presence of a semantically matching distractor.
-This manifested in longer self-paced reading times starting at the distractor \textcolor{blue}{itself which persisted almost until the end of the sentence} and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions at the critical verb.
+This manifested in longer self-paced reading times starting at the distractor \textcolor{black}{itself which persisted almost until the end of the sentence} and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions at the critical verb.
 The observed differences in the time course of the effect are likely due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference.
-\Copy{thisdesign3}{Overall, \textcolor{blue}{in the present design}, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as was previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.}\label{thisdesign3}  A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
+\Copy{thisdesign3}{Overall, \textcolor{black}{in the present design}, we found no decisive evidence for the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as was previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.}\label{thisdesign3}  A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
 
